### 2023-01-02 draft/draft.Rmd
draft

---
title: "Pollution, agricultural productivity, and development: Evidence from coal in plants in India^[thanks to...]"
author:
  - Joshua D. Merfeld^[KDI School of Public Policy and Management and IZA; merfeld@kdis.ac.kr]
date: "`r Sys.Date()`"
abstract: |
  | \noindent abstract
  |
  | \textbf{\textit{Keywords}}: 
  | \textbf{\textit{JEL Codes}}:

# Output type and options (no TOC and yes fig captions)
output:
  bookdown::pdf_document2: 
    toc: false
    fig_caption: yes
    # This also keeps the .tex file that is used to make the output (we will eventually need this for publication)
    keep_tex: yes

# Set bibliography
bibliography: references.bib

# This includes latex arguments
header-includes:
  - \newcommand{\beginappendix}{   
      \setcounter{table}{0}  
      \renewcommand{\thetable}{A\arabic{table}} 
      \setcounter{figure}{0} 
      \renewcommand{\thefigure}{A\arabic{figure}}
    }  # Relabel appendix tables/figures as e.g. A1
  - \usepackage[capposition=top]{floatrow}
  - \usepackage{placeins}   # THis is for /FloatBarrier
  - \usepackage{setspace}
  - \usepackage{dcolumn}
  - \usepackage{booktabs}
  - \usepackage{siunitx}
  - \usepackage{amsmath}
  - \usepackage{enumerate}
  - \usepackage[shortlabels]{enumitem}
  - \usepackage[hang,flushmargin]{footmisc}		# no indents on footnotes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dev = "cairo_pdf")

library(sf)
library(sp)
library(tidyverse)
library(scales)
library(RColorBrewer)
library(kableExtra)
library(ggpubr)
library(rticles)
library(stats)
library(broom)
library(extrafont)
library(fixest)
library(modelsummary)
library(openxlsx)
library(raster)
loadfonts()  # Must be run once in each session (markdown is a new "session" each time you render/knit)

# Sets wd to folder this script is in so we can create relative paths instead of absolute paths
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
# We want to get the wd up to the main folder
# Need to go up two levels
setwd('..')
setwd('..')
# Double check
getwd()    # check

reportP <- function(pValue){
  if (pValue < 0.001){
    result <- "p < 0.001"
  } else {
    result <- sprintf("p = %.3f", pValue) # inserts a float into a string and simultaneously do rounding
  }
  return(result)
}

reportDec <- function(dec){

    result <- sprintf("%.3f", dec) # inserts a float into a string and simultaneously do rounding

  return(result)
}

reportPercent <- function(dec){

    result <- sprintf("%.1f", 100*dec) # inserts a float into a string and simultaneously do rounding

  return(result)
}

commaNum <- function(large){

  result <- formatC(large, big.mark = ",")

  return(result)
}

# Shapefile of india
india_shape <- read_sf("data/spatial/shapefiles/state.shp")

# Coal plants
### Load raw data ----------------------------------------------------------------------------------------------------------------------------------------------------------
plants <- read.xlsx("data/raw/coal_plants.xlsx", sheet = "Units")
# India only
plants <- plants %>% filter(Country=="India")
# Also want only things with a non-missing year
plants <- plants %>% filter(is.na(Year)==F)

# Just keep what we want 
plants <- plants %>% dplyr::select(plant_id = ParentID, 
                                   unit_id = Tracker.ID, 
                                   capacity = `Capacity.(MW)`, 
                                   year_built = Year, 
                                   year_retired = RETIRED, 
                                   lat = Latitude, 
                                   lon = Longitude)

gps_points_plants <- plants %>% dplyr::select(lon, lat)
plants <- st_as_sf(SpatialPointsDataFrame(gps_points_plants, plants %>% dplyr::select(-c(lon, lat)), proj4string = CRS("EPSG:4326")))
plants <- plants %>% mutate(capacity = capacity/1000)

plants1990 <- plants %>% 
                  filter(year_built<=1990 & (year_retired>1990 | is.na(year_retired)==T)) %>%
                  group_by(plant_id) %>%
                  mutate(capacity = sum(capacity)) %>%
                  filter(row_number()==1) %>%
                  ungroup()

plants2010 <- plants %>% 
                  filter(year_built<=2010 & (year_retired>2010 | is.na(year_retired)==T)) %>%
                  group_by(plant_id) %>%
                  mutate(capacity = sum(capacity)) %>%
                  filter(row_number()==1) %>%
                  ungroup()

# wind example
windexample <- readRDS("data/spatial/windexample.rds")

# Tables
pollutiontable <- readRDS("pollution_development/draft/tables/pollutiontable.rds")
yieldtable <- readRDS("pollution_development/draft/tables/yieldtable.rds")
yieldtabletwo <- readRDS("pollution_development/draft/tables/yieldtabletwo.rds")
pmeffects <- readRDS("pollution_development/draft/tables/yieldtablePM.rds")
labortable <- readRDS("pollution_development/draft/tables/labortable.rds")
#labortable_two <- readRDS("pollution_development/draft/tables/labortable_two.rds")
ntltable <- readRDS("pollution_development/draft/tables/ntltable.rds")

```

\newpage

# Introduction

# Data and methods

## Data
SHRUGS: @almn2021 
Ag productivity: @gangopadhyay2022new
PM estimates: @hammer2020global

## Methodology

# Results

# Conclusion

```{r plants, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Coal plants in India from 1990 to 2010"}

plant1 <- ggplot() +
                theme_minimal() +
                geom_sf(data = india_shape) +
                geom_sf(data = plants1990)

plant2 <- ggplot() +
                theme_minimal() +
                geom_sf(data = india_shape) +
                geom_sf(data = plants2010)

ggarrange(plant1, plant2,
          labels = c("1990", "2010"))

```

```{r windexample, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Wind direction (2010-01-01)"}

ggplot() + 
  theme_minimal() +
  geom_sf(data = india_shape, fill = NA) +
  geom_tile(data = windexample, aes(x = x, y = y, fill = wind_example), alpha = 0.3) +
  scale_fill_viridis_c("wind direction\n(degrees)", option = "viridis") +
  geom_sf(data = plants2010) +
  labs(x = "lon",
       y = "lat")

```

\newpage

```{r pollutiontable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(pollutiontable) <- c("(1)", "(2)", "(3)", "(4)")
kable(
      pollutiontable, caption = "Wind direction and particulate matter",
      align = "cccc", booktabs = TRUE, linesep = ""
      ) %>%
  add_footnote(
               c(
                 "Note: Standard errors are in parentheses and are clustered at the village level.",
                 "* p<0.10 ** p<0.05 *** p<0.01"
                 ), 
               notation = "none",
               threeparttable = TRUE
               ) %>%
  column_spec(1, width = "4cm") %>%
  column_spec(c(2:5),width = "2cm") %>%
  row_spec(c(3, 6), bold = TRUE) %>%
  row_spec(7, hline_after = TRUE) %>%
  add_header_above(c(" ", "1998-2015" = 2, "2002-2013" = 2)) %>%
  kable_classic_2()

```

```{r yieldtable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(yieldtable) <- c("(1)", "(2)", "(3)", "(4)")
kable(
      yieldtable, caption = "Wind direction and agricultural productivity",
      align = "cccc", booktabs = TRUE, linesep = ""
      ) %>%
  add_footnote(
               c(
                 "Note: Standard errors are in parentheses and are clustered at the village level.",
                 "* p<0.10 ** p<0.05 *** p<0.01"
                 ), 
               notation = "none",
               threeparttable = TRUE
               ) %>%
  column_spec(1, width = "4cm") %>%
  column_spec(c(2:5),width = "2cm") %>%
  row_spec(6, bold = TRUE) %>%
  row_spec(8, hline_after = TRUE) %>%
  add_header_above(c(" ", "all" = 2, "monsoon" = 1, "winter" = 1)) %>%
  kable_classic_2()

```

```{r yieldtabletwo, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(yieldtabletwo) <- c("(1)", "(2)", "(3)", "(4)")
kable(
      yieldtabletwo, caption = "Wind direction and agricultural productivity",
      align = "cccc", booktabs = TRUE, linesep = ""
      ) %>%
  add_footnote(
               c(
                 "Note: Standard errors are in parentheses and are clustered at the village level.",
                 "* p<0.10 ** p<0.05 *** p<0.01"
                 ), 
               notation = "none",
               threeparttable = TRUE
               ) %>%
  column_spec(1, width = "4cm") %>%
  column_spec(c(2:5),width = "2cm") %>%
  row_spec(c(9, 10), hline_after = TRUE) %>%
  row_spec(c(6, 11), bold = TRUE) %>%
  add_header_above(c(" ", "all" = 2, "monsoon" = 1, "winter" = 1)) %>%
  kable_classic_2()

```

```{r labortable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(labortable) <- c("(1)", "(2)", "(3)")
kable(
      labortable, caption = "Wind direction and labor allocation",
      align = "ccc", booktabs = TRUE, linesep = ""
      ) %>%
  add_footnote(
               c(
                 "Note: Standard errors are in parentheses and are clustered at the village level.",
                 "* p<0.10 ** p<0.05 *** p<0.01"
                 ), 
               notation = "none",
               threeparttable = TRUE
               ) %>%
  column_spec(1, width = "4cm") %>%
  column_spec(c(2:4),width = "2cm") %>%
  row_spec(c(8), hline_after = TRUE) %>%
  row_spec(c(4, 7), bold = TRUE) %>%
  kable_classic_2()

```

\FloatBarrier
\newpage 
# References {.unnumbered}

::: {#refs}
:::
\FloatBarrier

```{=tex}
\beginappendix
```
\newpage
# Appendix A {.unnumbered}

\FloatBarrier
\singlespacing

```{r pmeffects, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(pmeffects) <- c("(1)", "(2)", "(3)", "(4)")
kable(
      pmeffects, caption = "Particulate matter and agricultural productivity",
      align = "cccc", booktabs = TRUE, linesep = ""
      ) %>%
  add_footnote(
               c(
                 "Note: Standard errors are in parentheses and are clustered at the village level.",
                 "* p<0.10 ** p<0.05 *** p<0.01"
                 ), 
               notation = "none"
               ) %>%
  column_spec(1, width = "4cm") %>%
  column_spec(c(2:5),width = "2cm") %>%
  row_spec(c(9, 10), hline_after = TRUE) %>%
  row_spec(c(6, 11), bold = TRUE) %>%
  add_header_above(c(" ", "all" = 2, "monsoon" = 1, "winter" = 1)) %>%
  kable_classic_2()

```

### 2023-01-02 draft/draft.tex
draft

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\newcommand{\beginappendix}{ \setcounter{table}{0} \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}} }
\usepackage[capposition=top]{floatrow}
\usepackage{placeins}
\usepackage{setspace}
\usepackage{dcolumn}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{amsmath}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage[hang,flushmargin]{footmisc}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Pollution, agricultural productivity, and development: Evidence from coal in plants in India},
  pdfauthor={Joshua D. Merfeld},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Pollution, agricultural productivity, and development: Evidence from coal in plants in India\footnote{thanks to\ldots{}}}
\author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
\date{2023-01-02}

\begin{document}
\maketitle
\begin{abstract}
\noindent abstract\\
\strut \\
\textbf{\textit{Keywords}}:\\
\textbf{\textit{JEL Codes}}:
\end{abstract}

\newpage

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\hypertarget{data-and-methods}{%
\section{Data and methods}\label{data-and-methods}}

\hypertarget{data}{%
\subsection{Data}\label{data}}

SHRUGS: Asher et al. (2021)
Ag productivity: Gangopadhyay et al. (2022)
PM estimates: Hammer et al. (2020)

\hypertarget{methodology}{%
\subsection{Methodology}\label{methodology}}

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

\begin{figure}
\centering
\includegraphics{draft_files/figure-latex/plants-1.pdf}
\caption{\label{fig:plants}Coal plants in India from 1990 to 2010}
\end{figure}

\begin{figure}
\centering
\includegraphics{draft_files/figure-latex/windexample-1.pdf}
\caption{\label{fig:windexample}Wind direction (2010-01-01)}
\end{figure}

\newpage

\begin{table}

\begin{threeparttable}
\caption{\label{tab:pollutiontable}Wind direction and particulate matter}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{1998-2015} & \multicolumn{2}{c}{2002-2013} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
  & (1) & (2) & (3) & (4)\\
\midrule
wind & 0.045*** & 0.040*** & 0.063*** & 0.037***\\
 & (0.004) & (0.006) & (0.005) & (0.007)\\
\textbf{fixed effects:} & \textbf{---------} & \textbf{---------} & \textbf{---------} & \textbf{---------}\\
village & Yes & Yes & Yes & Yes\\
month & Yes & Yes & Yes & Yes\\
\textbf{varying slopes:} & \textbf{---------} & \textbf{---------} & \textbf{---------} & \textbf{---------}\\
month (village) & No & Yes & No & Yes\\
\midrule
observations & 22,345,092 & 22,345,092 & 14,896,728 & 14,896,728\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\small
\item [] Note: Standard errors are in parentheses and are clustered at the village level.
\item [] * p<0.10 ** p<0.05 *** p<0.01
\end{tablenotes}
\end{threeparttable}
\end{table}

\begin{table}

\begin{threeparttable}
\caption{\label{tab:yieldtable}Wind direction and agricultural productivity}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{all} & \multicolumn{1}{c}{monsoon} & \multicolumn{1}{c}{winter} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5}
  & (1) & (2) & (3) & (4)\\
\midrule
wind & -0.003*** & -0.003*** & -0.002*** & -0.003***\\
 & (0.0002) & (0.0002) & (0.0002) & (0.0003)\\
rain (z) &  & 0.029*** & 0.082*** & 0.016***\\
 &  & (0.0004) & (0.002) & (0.0004)\\
sub-sample & all & all & monsoon & winter\\
\textbf{fixed effects:} & \textbf{---------} & \textbf{---------} & \textbf{---------} & \textbf{---------}\\
village-season & Yes & Yes & No & No\\
year & Yes & Yes & Yes & Yes\\
\midrule
village & No & No & Yes & Yes\\
observations & 2,391,533 & 2,375,337 & 1,259,123 & 1,116,214\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\small
\item [] Note: Standard errors are in parentheses and are clustered at the village level.
\item [] * p<0.10 ** p<0.05 *** p<0.01
\end{tablenotes}
\end{threeparttable}
\end{table}

\begin{table}

\begin{threeparttable}
\caption{\label{tab:yieldtabletwo}Wind direction and agricultural productivity}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{all} & \multicolumn{1}{c}{monsoon} & \multicolumn{1}{c}{winter} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5}
  & (1) & (2) & (3) & (4)\\
\midrule
pm25 & -0.021*** & -0.020*** & -0.013*** & -0.024***\\
 & (0.001) & (0.002) & (0.001) & (0.003)\\
rain (z) &  & 0.004** & 0.086*** & -0.016***\\
 &  & (0.002) & (0.002) & (0.004)\\
sub-sample & all & all & monsoon & winter\\
\textbf{fixed effects:} & \textbf{---------} & \textbf{---------} & \textbf{---------} & \textbf{---------}\\
village-season & Yes & Yes & No & No\\
year & Yes & Yes & Yes & Yes\\
village & No & No & Yes & Yes\\
\midrule
observations & 2,391,533 & 2,375,337 & 1,259,123 & 1,116,214\\
\midrule
\textbf{first stage:} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
wind & 0.143*** & 0.126*** & 0.155*** & 0.105***\\
 & (0.003) & (0.003) & (0.003) & (0.004)\\
rain (z) &  & -1.23*** & 0.301*** & -1.36***\\
 &  & (0.010) & (0.015) & (0.009)\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\small
\item [] Note: Standard errors are in parentheses and are clustered at the village level.
\item [] * p<0.10 ** p<0.05 *** p<0.01
\end{tablenotes}
\end{threeparttable}
\end{table}

\begin{table}

\begin{threeparttable}
\caption{\label{tab:labortable}Wind direction and labor allocation}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
\toprule
  & (1) & (2) & (3)\\
\midrule
wind & -0.028* & -0.035** & -0.027*\\
 & (0.015) & (0.015) & (0.015)\\
controls & No & Yes & Yes\\
\textbf{fixed effects:} & \textbf{-------} & \textbf{--------} & \textbf{-------}\\
district & Yes & Yes & Yes\\
year & Yes & Yes & Yes\\
\textbf{varying slopes:} & \textbf{-------} & \textbf{--------} & \textbf{-------}\\
year (district) & No & No & Yes\\
\midrule
observations & 899,045 & 898,856 & 898,856\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\small
\item [] Note: Standard errors are in parentheses and are clustered at the village level.
\item [] * p<0.10 ** p<0.05 *** p<0.01
\end{tablenotes}
\end{threeparttable}
\end{table}

\FloatBarrier
\newpage

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-almn2021}{}}%
Asher, Sam, Tobias Lunt, Ryu Matsuura, and Paul Novosad. 2021. {``{Development Research at High Geographic Resolution: An Analysis of Night Lights, Firms, and Poverty in India using the SHRUG Open Data Platform}.''} \emph{{The World Bank Economic Review}}.

\leavevmode\vadjust pre{\hypertarget{ref-gangopadhyay2022new}{}}%
Gangopadhyay, Prasun K, Paresh B Shirsath, Vinay K Dadhwal, and Pramod K Aggarwal. 2022. {``{A new two-decade (2001--2019) high-resolution agricultural primary productivity dataset for India}.''} \emph{{Scientific Data}} 9 (1): 1--12.

\leavevmode\vadjust pre{\hypertarget{ref-hammer2020global}{}}%
Hammer, Melanie S, Aaron van Donkelaar, Chi Li, Alexei Lyapustin, Andrew M Sayer, N Christina Hsu, Robert C Levy, et al. 2020. {``Global Estimates and Long-Term Trends of Fine Particulate Matter Concentrations (1998--2018).''} \emph{{Environmental Science \& Technology}} 54 (13): 7879--90.

\end{CSLReferences}

\FloatBarrier

\beginappendix

\newpage

\hypertarget{appendix-a}{%
\section*{Appendix A}\label{appendix-a}}
\addcontentsline{toc}{section}{Appendix A}

\FloatBarrier
\singlespacing

\begin{table}

\caption{\label{tab:pmeffects}Particulate matter and agricultural productivity}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{all} & \multicolumn{1}{c}{monsoon} & \multicolumn{1}{c}{winter} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5}
  & (1) & (2) & (3) & (4)\\
\midrule
particulate matter & 0.045 & 0.501*** & 2.05*** & 0.472***\\
(PM 2.5, '000s) & (0.031) & (0.031) & (0.099) & (0.037)\\
rain (z) &  & 0.030*** & 0.083*** & 0.017***\\
 &  & (0.0004) & (0.002) & (0.0004)\\
sub-sample & all & all & monsoon & winter\\
\textbf{fixed effects:} & \textbf{-------} & \textbf{---------} & \textbf{---------} & \textbf{---------}\\
village-season & Yes & Yes & No & No\\
year & Yes & Yes & Yes & Yes\\
village & No & No & Yes & Yes\\
\midrule
observations & 2,391,533 & 2,375,337 & 1,259,123 & 1,116,214\\
\midrule
\bottomrule
\multicolumn{5}{l}{\textsuperscript{} Note: Standard errors are in parentheses and are clustered at the village}\\
\multicolumn{5}{l}{level.}\\
\multicolumn{5}{l}{\textsuperscript{} * p<0.10 ** p<0.05 *** p<0.01}\\
\end{tabular}
\end{table}

\end{document}

### 2023-01-02 analyses/regressions.R

@@ -1,4 +1,4 @@
-# Purpose: This script estimates effects
+# Purpose: This script estimates effects. I save the results to load into the markdown script
 # Author: Josh Merfeld
 # Date: December 12th, 2022
 
@@ -18,43 +18,173 @@ setwd("../../")
 getwd()    # check
 
 
-# Nightlights
-ntl <- read_csv("data/clean/village_ntl.csv")
-ntl <- ntl %>% mutate(state = substr(shrid, 1, 2), 
-                      sums_proportion = sums/366)
-ntl <- panel(ntl, ~ shrid + year, duplicate.method = "first")
+
+
+pollution <- read_csv(paste0("data/clean/pm25/all_combined.csv"))
+pollution <- pollution %>% mutate(
+                                  date = as_date(paste0(year, "-", month, "-01")),
+                                  state = substr(shrid, 1, 2),
+                                  village = shrid,
+                                  month = as.numeric(date),
+                                  wind = sums
+                                  )
+pollution <- pollution %>% mutate()
+pollution <- panel(pollution, ~ village + date)
+
+
+pol1 <- feols(mean ~ wind | village + month, data = pollution, cluster = c("village"))
+pol2 <- feols(mean ~ wind | village[month] + month, data = pollution, cluster = c("village"))
+pol3 <- feols(mean ~ wind | village + month, data = pollution %>% filter(year>=2002 & year<=2013), cluster = c("village"))
+pol4 <- feols(mean ~ wind | village[month] + month, data = pollution %>% filter(year>=2002 & year<=2013), cluster = c("village"))
+
+pollutiontable <- etable(
+                          pol1, pol2, pol3, pol4, 
+                          se.below = TRUE,
+                          depvar = FALSE,
+                          signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.1),
+                          digits = 3,
+                          fitstat = c("n"),
+                          coefstat = "se"
+                          )
+pollutiontable <- pollutiontable[-9,]
+pollutiontable <- pollutiontable[-8,]
+pollutiontable <- as.matrix(pollutiontable)
+rownames(pollutiontable) <- c("wind", "", "fixed effects:",
+                               "village", "month", "varying slopes:",
+                               "month (village)", "", "observations")
+saveRDS(pollutiontable, "pollution_development/draft/tables/pollutiontable.rds") 
 
 
 
 
 
[…]

### 2023-01-02 cleaning/cleaning_ag_productivity_village.R

@@ -1,4 +1,4 @@
-# Purpose: This script doespulls EVI
+# Purpose: This script cleans data for estimates of agricultural productivity
 # Author: Josh Merfeld
 # Date: December 12th, 2022
 
@@ -73,8 +73,9 @@ for (year in 2002:2013){
 for (year in 2002:2013){
   # Monsoon is this year only
   rainfall <- read_csv(paste0("data/clean/terra/precip", year, ".csv"))
-  rainfall$rain_total <- apply(rainfall[,c(7:11)], 1, sum)
-  rainfall <- rainfall %>% dplyr::select(shrid, rain_total)
+  rainfall <- rainfall[, c(1, 7:11)]
+  rainfall$rain_total <- apply(rainfall[,c(2:6)], 1, sum)
+  colnames(rainfall) <- c("shrid", "rain_m1", "rain_m2", "rain_m3", "rain_m4", "rain_m5", "rain_total")
   rainfall$season <- "monsoon"
   # Load the extracted values from above
   extracted_values <- read_csv(paste0("data/clean/ag_productivity/monsoon", year, ".csv"))
@@ -88,15 +89,14 @@ for (year in 2002:2013){
   
   # Winter is TWO years
   rainfall1 <- read_csv(paste0("data/clean/terra/precip", year, ".csv"))
-  rainfall1$rain_total1 <- apply(rainfall1[,c(12:13)], 1, sum)
-  rainfall1 <- rainfall1 %>% dplyr::select(shrid, rain_total1)
+  rainfall1 <- rainfall1[, c(1, 12:13)]
   # Second year
   rainfall2 <- read_csv(paste0("data/clean/terra/precip", year + 1, ".csv"))
-  rainfall2$rain_total2 <- apply(rainfall2[,c(2:4)], 1, sum)
-  rainfall2 <- rainfall2 %>% dplyr::select(shrid, rain_total2)
+  rainfall2 <- rainfall2[, c(1, 2:4)]
+  
   rainfall <- cbind(rainfall1, rainfall2[,-1])
-  rainfall <- rainfall %>% mutate(rain_total = rain_total1 + rain_total2)
-  rainfall <- rainfall %>% dplyr::select(shrid, rain_total)
+  rainfall$rain_total <- apply(rainfall[,c(2:6)], 1, sum)
+  colnames(rainfall) <- c("shrid", "rain_m1", "rain_m2", "rain_m3", "rain_m4", "rain_m5", "rain_total")
[…]

### 2023-01-02 cleaning/cleaning_overlap_merging_nss.R

@@ -71,7 +71,9 @@ df_labor <- df_labor %>% filter(is.na(merge)==F)
 
 
 # Going to go through the dates and process date by date
-df_labor_merged <-
+pollution <- read_csv(paste0("data/clean/pm25/all_combined.csv"))
+pollution <- pollution %>% rename(pm25 = mean)
+df_labor_merged <- c()
 for (day in 1:length(date_vec)){
   # First, load villages
   village_wind <- read_csv(paste0("data/clean/wind_ntl/days/date_", year(date_vec[day]), "-", month(date_vec[day]), "-", day(date_vec[day]), ".csv")) %>% as_tibble()
@@ -81,16 +83,24 @@ for (day in 1:length(date_vec)){
   village_wind <- villages %>% left_join(village_wind, by = "shrid")
   # Missings as zero
   village_wind$days_sum[is.na(village_wind$days_sum)==T] <- 0
+  # And also pm
+  pollution_merge <- pollution %>% filter(year==year(date_vec[day]) & month==month(date_vec[day]))
+  village_wind <- village_wind %>% left_join(pollution, by = "shrid")
+  
   
   # Weighted mean by AREA. Any parts of the district without villages are ZEROS
   village_wind <- village_wind %>% 
                   group_by(state, district) %>%
                   mutate(days_sum = weighted.mean(days_sum, area_weight),
+                         pm25_sum = weighted.mean(pm25, area_weight),
                          state_merge = as.numeric(state),
-                         district_merge = as.numeric(district)) %>%
+                         district_merge = as.numeric(district),
+                         tot_weight = sum(area_weight),
+                         days_sum = days_sum*tot_weight,
+                         pm25_sum = days_sum*tot_weight) %>% # Last one adjusts for the zeros for areas not here
                   filter(row_number()==1) %>%
                   ungroup() %>%
-                  select(state_merge, district_merge, days_sum)
+                  select(state_merge, district_merge, days_sum, pm25_sum)
   
   
   temp <- df_labor %>% filter(date==date_vec[day])
@@ -108,239 +118,6 @@ write.csv(df_labor_merged, "data/clean/nss/merged_week.csv")
 
 
 
-
-
-# NOW DO DAILY ---------------------------------------------------------
-
-df_labor <- c()
-for (nsswave in c(61, 62, 64, 66, 68)){
[…]

### 2023-01-02 cleaning/cleaning_pm25.R

@@ -1,0 +1,137 @@
+# Purpose: This script cleans 
+# Author: Josh Merfeld
+# Date: December 12th, 2022
+
+rm(list = ls())
+
+# Load packages
+library(tidyverse)
+library(sf)
+library(sp)
+library(raster)
+library(ncdf4)
+library(lubridate)
+library(exactextractr)  
+
+
+# Sets wd to folder this script is in so we can create relative paths instead of absolute paths
+setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
+# We want to get the wd up to the main folder
+# Need to go up two levels
+setwd("../../")
+# Double check
+getwd()    # check
+
+
+
+villages_centroids <- read_sf("data/spatial/centroids_villages_30km")
+villages_centroids <- as_tibble(villages_centroids) %>% dplyr::select(shrid)
+villages_centroids$within30 <- 1
+villages <- read_sf("data/spatial/shapefiles/village.shp")
+villages$shrid <- paste0(villages$pc11_s_id, "-", villages$pc11_tv_id)
+villages <- villages %>% left_join(villages_centroids, by = "shrid")
+villages <- villages %>% filter(within30==1) %>% dplyr::select(shrid)
+rm(villages_centroids)
+
+
+
+months <- c()
+for (month in 1:9){
+  months <- c(months, paste0("0", month))
+}
+months <- c(months, "11", "12")
+pollution <- c()
+for (year in 1998:2015){
+  for (month in months){
+    # MONSOON --------------------------------------------------------------------
+    # This changes by year
+    r <- raster(paste0("data/raw/pm25/V5GL03.HybridPM25.Asia.", year, month, "-", year, month, ".nc"))
+    
+    # crop first (makes extraction MUCH faster)
+    r <- crop(r, villages)
+    
+    # extract
+    extracted_values <- exact_extract(r, villages, fun = "mean", append_cols = "shrid")
+    extracted_values <- as_tibble(extracted_values) %>% 
+                        group_by(shrid) %>% 
+                        mutate(mean = mean(mean)) %>% 
+                        filter(row_number()==1) %>%
+                        ungroup()
+    
+    # And wind direction
+    village_wind <- read_csv(paste0("data/clean/wind_ntl/months/y", year, "m", as.numeric(month), ".csv")) %>% as_tibble()
+    
+    # Merger
+    extracted_values <- extracted_values %>% left_join(village_wind, by = "shrid")
[…]

### 2023-01-03 draft/draft.Rmd
draft and finalizing regressions

@@
 
+# district
+districts <- read_sf(paste0("data/spatial/districts/districts.shp"))
+districts <- districts %>% 
+                dplyr::select(state = ST_CEN_CD, district = DT_CEN_CD, state_name = ST_NM, district_name = DISTRICT)
+districts <- districts %>% filter(state=="02" & district=="02")
+# Villages
+villages <- read_sf(paste0("data/spatial/villages_overlap/villages_overlap.shp"))
+villages <- villages %>% mutate(shrid = paste0(pc11_s_id, "-", pc11_tv_id)) %>%
+                          dplyr::select(shrid, state = ST_CEN_CD, district = DT_CEN_CD, state_name = ST_NM, district_name = DISTRICT)
+villages <- villages %>% filter(state=="02" & district=="02")
+wind <- read_csv(paste0("data/clean/wind_ntl/days/date_2010-1-1.csv")) %>% as_tibble()
+wind$days_sum <- apply(wind[,2:8], 1, FUN = "sum")
+wind <- wind %>% dplyr::select(shrid, days_sum)
+villages <- villages %>% left_join(wind, by = "shrid")
+villages <- villages %>% st_set_crs(st_crs(districts))
+# and crs for plants
+plants1990 <- plants1990 %>% st_transform(st_crs(districts))
+plants2010 <- plants2010 %>% st_transform(st_crs(districts))
+india_shape2 <- india_shape %>% st_transform(st_crs(districts))
+
 # Tables
@@
 yieldtabletwo <- readRDS("pollution_development/draft/tables/yieldtabletwo.rds")
+yieldtablehet <- readRDS("pollution_development/draft/tables/yieldtablehet.rds")
 pmeffects <- readRDS("pollution_development/draft/tables/yieldtablePM.rds")
@@
 ntltable <- readRDS("pollution_development/draft/tables/ntltable.rds")
+yieldtableleads <- readRDS("pollution_development/draft/tables/yieldtableleads.rds")
+
+```
+
+```{r, echo = F}
+# This chunk will allow us to add footers below figures
+library(ggplot2)
+
+hook_plot_tex_footer <- function(x, options) {  
+  x_out <- knitr:::hook_plot_tex(x, options) 
+  if(!is.null(options$fig.footer)) {
+    inter <- sprintf("\\floatfoot*{%s}\n\\end{figure}", options$fig.footer[1])
+    x_out <- gsub(x=x_out, pattern="\n\\end{figure}", replacement=inter, fixed=TRUE)
+  }
+  x_out
+}
+
+knitr::knit_hooks$set(plot=hook_plot_tex_footer)
 
@@
 \newpage
+\doublespacing
 
 # Introduction
+
+pollution could be bad for agriculture [@heck1982assessment,@marshall1997hidden]
+
+lower agricultural productivity near gold mines in Ghana [@aragon2016polluting]; apparently lots of pollution
+  - "The main identification assumption is that the change in agricultural productivity over time in both areas would be similar in the absence of mining."
+  - within 20km
+  - also finds suggestive evidence it's not all labor productivity changes
+
+"Much of what we know about the marginal effect of pollution on infant mortality is derived from developed country data.
+However, given the lower levels of air pollution in developed countries, these estimates may not be externally valid to the developing country context if there is a non-linear dose relationship between pollution and mortality or if the costs of avoidance behaviour differ considerably between the two contexts.
+In this article, we estimate the relationship between pollution and infant mortality using data from Mexico.
+Our estimates for PM10 tend to be similar (or even smaller) than the US estimates, while our findings on CO tend to be larger than those derived from the US context." [@arceo2016does]
+
+We have known about the effects of pollution on human health for many years [@brunekreef2002air,@kampa2008human,@pope2006health]
+good review on the effects of pollution early in life in @currie2014we
+  - also in developing countries, where there is a consistent effect of exposure on infant mortality [@heft2018robust]
+
+"Air pollution has both acute and chronic effects on human health, affecting a number of different systems and organs.
+It ranges from minor upper respiratory irritation to chronic respiratory and heart disease, lung cancer, acute respiratory infections in children and chronic bronchitis in adults, aggravating pre-existing heart and lung disease, or asthmatic attacks.
+In addition, short- and long-term exposures have also been linked with premature mortality and reduced life expectancy.
+These effects of air pollutants on human health and their mechanism of action are briefly discussed." @kampa2008human
+
+worker productivity and cognitive function:
+  - call center workers in China, and at common levels of pollution [@chang2019effect]
+  - tests [@ebenstein2016long]
+  - farm workers in California [@graff2012impact]
+  - increase in extensive margin [@hanna2015effect]
+  - manufacturing in China [@he2019severe]; "We uncover statistically significant adverse output effects from more prolonged exposure, but effects are not large.
+A substantial +10 mug/m3 PM2.5 variation sustained over 25 days reduces daily output by 1 percent."
+
+induces migration [@chen2022effect]
+
+overall effects on agricultural productivity taking into accounts any changes in input allocation caused by the increase in pollution.
 
@@
 ## Data
-SHRUGS: @almn2021 
+
+@almn2021
 Ag productivity: @gangopadhyay2022new
@@
 
+```{r data, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+data <- matrix(NA, nrow = 3, ncol = 5)
+colnames(data) <- c("shapefile", "wind", "pollution", "agriculture", "nightlights")
+rownames(data) <- c("source", "geographic coverage", "temporal coverage")
+data[1,1] <- "Asher et al. (2021)"
+data[1,2] <- "NCAR"
+data[1,3] <- "Hammer et al. (2020)"
+data[1,4] <- "Angopadhyay et al. (2022)"
+data[1,5] <- "Asher et al. (2021)"
+data[2,1] <- "India"
+data[2,2] <- "global"
+data[2,3] <- "global"
+data[2,4] <- "global"
+data[2,5] <- "India"
+data[3,1] <- " "
+data[3,2] <- "daily"
+data[3,3] <- "monthly"
+data[3,4] <- "two seasons/year"
+data[3,5] <- "yearly"
+
+# Table
+kable(
+      data, caption = "Remote sensing data sources",
+      align = "ccccc", booktabs = TRUE, linesep = ""
+      ) %>%
+  footnote(
+           general = "NCAR: https://climatedataguide.ucar.edu/", 
+           general_title = " ",
+           threeparttable = TRUE
+           ) %>%
+  column_spec(1, width = "3cm") %>%
+  column_spec(c(2:6),width = "2cm") %>%
+  kable_classic_2()
+
+```
+
 ## Methodology
 
-# Results
-
-# Conclusion
-
-```{r plants, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Coal plants in India from 1990 to 2010"}
+```{r plants, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Coal plants in India from 1990 to 2010", fig.footer = list("Note: The top figure shows the location of coal plants in 1990.
+The bottom figure shows the location of coal plants in 2010.")}
 
@@
                 theme_minimal() +
-                geom_sf(data = india_shape) +
-                geom_sf(data = plants1990)
+                geom_sf(data = india_shape2) +
+                geom_sf(data = plants1990, size = 0.75) +
+                labs(x = "lon",
+                     y = "lat",
+                     subtitle = "Figure A: 1990")
 
@@
                 theme_minimal() +
-                geom_sf(data = india_shape) +
-                geom_sf(data = plants2010)
+                geom_sf(data = india_shape2) +
+                geom_sf(data = plants2010, size = 0.75) +
+                labs(x = "lon",
+                     y = "lat",
+                     subtitle = "Figure B: 2010")
 
-ggarrange(plant1, plant2,
-          labels = c("1990", "2010"))
+ggarrange(plant1, plant2)
 
@@
 
-```{r windexample, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Wind direction (2010-01-01)"}
+```{r windexample, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 8, fig.cap = "Wind direction and aggregation examples (2010-01-01)", fig.footer = list("Note: The top figure shows the average wind direction on January 1st, 2010.
+The points are the location of coal plants on that date.
+The bottom figure shows the distribution of pollution exposure in a specific district -- Kangra district in Himachal Pradesh -- on the same date.")}
 
-ggplot() + 
+p1 <- ggplot() + 
         theme_minimal() +
@@
         scale_fill_viridis_c("wind direction\n(degrees)", option = "viridis") +
-  geom_sf(data = plants2010) +
+        geom_sf(data = plants2010, size = 0.5) +
         labs(x = "lon",
-       y = "lat")
+             y = "lat",
+             subtitle = "Figure A: Wind direction")
+
+p2 <- ggplot() + 
+        theme_minimal() +
+        geom_sf(data = districts, fill = NA) +
+        geom_sf(data = villages %>% filter(is.na(days_sum)==F), aes(fill = days_sum), lwd = NA) +
+        scale_fill_viridis_c("exposure (days)", option = "viridis") +
+        labs(x = "lon",
+             y = "lat",
+             subtitle = "Figure B: Aggregation to district")
+
+ggarrange(p1, p2,
+          nrow = 2)
 
@@
       ) %>%
-  add_footnote(
-               c(
+  footnote(
+           general = c(
                        "Note: Standard errors are in parentheses and are clustered at the village level.",
@@
                        ), 
-               notation = "none",
+           general_title = "",
            threeparttable = TRUE
@@
   column_spec(c(2:5),width = "2cm") %>%
-  row_spec(c(3, 6), bold = TRUE) %>%
-  row_spec(7, hline_after = TRUE) %>%
+  row_spec(c(3), bold = TRUE) %>%
+  row_spec(6, hline_after = TRUE) %>%
   add_header_above(c(" ", "1998-2015" = 2, "2002-2013" = 2)) %>%
@@
 
+# Results
+
 ```{r yieldtable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
@@
 # Table
-colnames(yieldtable) <- c("(1)", "(2)", "(3)", "(4)")
+colnames(yieldtable) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
 kable(
       yieldtable, caption = "Wind direction and agricultural productivity",
-      align = "cccc", booktabs = TRUE, linesep = ""
+      align = "ccccc", booktabs = TRUE, linesep = ""
       ) %>%
-  add_footnote(
-               c(
+  footnote(
+           general = c(
                        "Note: Standard errors are in parentheses and are clustered at the village level.",
@@
                        ), 
-               notation = "none",
+           general_title = "",
            threeparttable = TRUE
            ) %>%
-  column_spec(1, width = "4cm") %>%
-  column_spec(c(2:5),width = "2cm") %>%
-  row_spec(6, bold = TRUE) %>%
-  row_spec(8, hline_after = TRUE) %>%
-  add_header_above(c(" ", "all" = 2, "monsoon" = 1, "winter" = 1)) %>%
+  column_spec(1, width = "3cm") %>%
+  column_spec(c(2:6),width = "2cm") %>%
+  row_spec(5, bold = TRUE) %>%
+  row_spec(9, hline_after = TRUE) %>%
+  add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
   kable_classic_2()
@@
 # Table
-colnames(yieldtabletwo) <- c("(1)", "(2)", "(3)", "(4)")
+colnames(yieldtabletwo) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
 kable(
-      yieldtabletwo, caption = "Wind direction and agricultural productivity",
+      yieldtabletwo, caption = "Pollution and agricultural productivity, IV estimates",
       align = "cccc", booktabs = TRUE, linesep = ""
       ) %>%
-  add_footnote(
-               c(
+  footnote(
+           general = c(
                        "Note: Standard errors are in parentheses and are clustered at the village level.",
@@
                        ), 
-               notation = "none",
+           general_title = "",
+           threeparttable = TRUE
+           ) %>%
+  column_spec(1, width = "3.5cm") %>%
+  column_spec(c(2:6),width = "2cm") %>%
+  row_spec(5, bold = TRUE) %>%
+  row_spec(c(9:10), hline_after = TRUE) %>%
+  add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
+  kable_classic_2()
+
+```
+
+```{r yieldtablehet, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+# Table
+colnames(yieldtablehet) <- c("(1)", "(2)", "(3)", "(4)", "(5")
+kable(
+      yieldtablehet, caption = "Heterogeneity in the effects of pollution on productivity",
+      align = "ccccc", booktabs = TRUE, linesep = ""
+      ) %>%
+  footnote(
+           general = c(
+                       "Note: Standard errors are in parentheses and are clustered at the village level.",
+                       "* p<0.10 ** p<0.05 *** p<0.01"
+                       ), 
+           general_title = "",
            threeparttable = TRUE
@@
   column_spec(1, width = "4cm") %>%
-  column_spec(c(2:5),width = "2cm") %>%
-  row_spec(c(9, 10), hline_after = TRUE) %>%
-  row_spec(c(6, 11), bold = TRUE) %>%
-  add_header_above(c(" ", "all" = 2, "monsoon" = 1, "winter" = 1)) %>%
+  column_spec(c(2:6),width = "2cm") %>%
+  row_spec(c(13), hline_after = TRUE) %>%
+  row_spec(c(11), bold = TRUE) %>%
+  add_header_above(c(" ", ">p(50)" = 1, "<=p(50)" = 1, "all" = 3)) %>%
   kable_classic_2()
@@
 # Table
-colnames(labortable) <- c("(1)", "(2)", "(3)")
+colnames(labortable) <- c("all", "all", "self", "wage", "farm", "non-farm")
 kable(
       labortable, caption = "Wind direction and labor allocation",
-      align = "ccc", booktabs = TRUE, linesep = ""
+      align = "cccccc", booktabs = TRUE, linesep = ""
       ) %>%
-  add_footnote(
-               c(
-                 "Note: Standard errors are in parentheses and are clustered at the village level.",
+  footnote(
+           general = c(
+                       "Note: Standard errors are in parentheses and are clustered at the district level. Control variables include female, age, age squared, and (years of) education.",
                        "* p<0.10 ** p<0.05 *** p<0.01"
                        ), 
-               notation = "none",
+           general_title = "",
            threeparttable = TRUE
            ) %>%
-  column_spec(1, width = "4cm") %>%
-  column_spec(c(2:4),width = "2cm") %>%
+  column_spec(1, width = "3cm") %>%
+  column_spec(c(2:7),width = "1.5cm") %>%
   row_spec(c(8), hline_after = TRUE) %>%
-  row_spec(c(4, 7), bold = TRUE) %>%
+  row_spec(c(4,7), bold = TRUE) %>%
   kable_classic_2()
@@
 
+```{r ntltable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+# Table
+colnames(ntltable) <- c("(1)", "(2)")
+kable(
+      ntltable, caption = "Wind direction and nightlight growth",
+      align = "cc", booktabs = TRUE, linesep = ""
+      ) %>%
+  footnote(
+           general = c(
+                       "Note: Standard errors are in parentheses and are clustered at the village level.",
+                       "* p<0.10 ** p<0.05 *** p<0.01"
+                       ), 
+           general_title = "",
+           threeparttable = TRUE
+           ) %>%
+  column_spec(1, width = "3cm") %>%
+  column_spec(c(2:3),width = "2cm") %>%
+  row_spec(c(9), hline_after = TRUE) %>%
+  row_spec(c(5,8), bold = TRUE) %>%
+  kable_classic_2()
+
+```
+
+# Conclusion
+
 \FloatBarrier
 \newpage
+
+\newpage
+\singlespacing
 # References {.unnumbered}
@@
 
-\FloatBarrier
-\singlespacing
-
-```{r pmeffects, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+```{r yieldtableleads, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
 
 # Table
-colnames(pmeffects) <- c("(1)", "(2)", "(3)", "(4)")
+colnames(yieldtableleads) <- c("(1)", "(2)")
 kable(
-      pmeffects, caption = "Particulate matter and agricultural productivity",
-      align = "cccc", booktabs = TRUE, linesep = ""
+      yieldtableleads, caption = "Agricultural productivity and pollution leads",
+      align = "cc", booktabs = TRUE, linesep = ""
       ) %>%
-  add_footnote(
-               c(
+  footnote(
+           general = c(
                        "Note: Standard errors are in parentheses and are clustered at the village level.",
@@
                        ), 
-               notation = "none"
+           general_title = "",
+           threeparttable = TRUE
            ) %>%
-  column_spec(1, width = "4cm") %>%
-  column_spec(c(2:5),width = "2cm") %>%
-  row_spec(c(9, 10), hline_after = TRUE) %>%
-  row_spec(c(6, 11), bold = TRUE) %>%
-  add_header_above(c(" ", "all" = 2, "monsoon" = 1, "winter" = 1)) %>%
+  column_spec(1, width = "3cm") %>%
+  column_spec(c(2:3),width = "2cm") %>%
+  row_spec(c(12), hline_after = TRUE) %>%
+  row_spec(c(9), bold = TRUE) %>%
   kable_classic_2()

### 2023-01-03 draft/draft.tex
draft and finalizing regressions

@@
 \author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
-\date{2023-01-02}
+\date{2023-01-03}
 
@@
 \newpage
+\doublespacing
 
@@
 \section{Introduction}\label{introduction}}
+
+pollution could be bad for agriculture Marshall et al. (1997)
+
+lower agricultural productivity near gold mines in Ghana (Aragón and Rud 2016); apparently lots of pollution
+- ``The main identification assumption is that the change in agricultural productivity over time in both areas would be similar in the absence of mining.''
+- within 20km
+- also finds suggestive evidence it's not all labor productivity changes
+
+``Much of what we know about the marginal effect of pollution on infant mortality is derived from developed country data.
+However, given the lower levels of air pollution in developed countries, these estimates may not be externally valid to the developing country context if there is a non-linear dose relationship between pollution and mortality or if the costs of avoidance behaviour differ considerably between the two contexts.
+In this article, we estimate the relationship between pollution and infant mortality using data from Mexico.
+Our estimates for PM10 tend to be similar (or even smaller) than the US estimates, while our findings on CO tend to be larger than those derived from the US context.'' (Arceo, Hanna, and Oliva 2016)
+
+We have known about the effects of pollution on human health for many years Pope III and Dockery (2006)
+good review on the effects of pollution early in life in Currie et al. (2014)
+- also in developing countries, where there is a consistent effect of exposure on infant mortality (Heft-Neal et al. 2018)
+
+``Air pollution has both acute and chronic effects on human health, affecting a number of different systems and organs.
+It ranges from minor upper respiratory irritation to chronic respiratory and heart disease, lung cancer, acute respiratory infections in children and chronic bronchitis in adults, aggravating pre-existing heart and lung disease, or asthmatic attacks.
+In addition, short- and long-term exposures have also been linked with premature mortality and reduced life expectancy.
+These effects of air pollutants on human health and their mechanism of action are briefly discussed.'' Kampa and Castanas (2008)
+
+worker productivity and cognitive function:
+- call center workers in China, and at common levels of pollution (Chang et al. 2019)
+- tests (Ebenstein, Lavy, and Roth 2016)
+- farm workers in California (Graff Zivin and Neidell 2012)
+- increase in extensive margin (Hanna and Oliva 2015)
+- manufacturing in China (He, Liu, and Salvo 2019); ``We uncover statistically significant adverse output effects from more prolonged exposure, but effects are not large.
+A substantial +10 mug/m3 PM2.5 variation sustained over 25 days reduces daily output by 1 percent.''
+
+induces migration (Chen, Oliva, and Zhang 2022)
+
+overall effects on agricultural productivity taking into accounts any changes in input allocation caused by the increase in pollution.
 
@@
 
-SHRUGS: Asher et al. (2021)
+Asher et al. (2021)
 Ag productivity: Gangopadhyay et al. (2022)
 PM estimates: Hammer et al. (2020)
+
+\begin{table}
+
+\caption{\label{tab:data}Remote sensing data sources}
+\centering
+\begin{threeparttable}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\toprule
+  & shapefile & wind & pollution & agriculture & nightlights\\
+\midrule
+source & Asher et al. (2021) & NCAR & Hammer et al. (2020) & Angopadhyay et al. (2022) & Asher et al. (2021)\\
+geographic coverage & India & global & global & global & India\\
+temporal coverage &  & daily & monthly & two seasons/year & yearly\\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}
+\item \textit{ } 
+\item NCAR: https://climatedataguide.ucar.edu/
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
 
@@
 
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
-
 \begin{figure}
-\centering
-\includegraphics{draft_files/figure-latex/plants-1.pdf}
-\caption{\label{fig:plants}Coal plants in India from 1990 to 2010}
+\includegraphics{draft_files/figure-latex/plants-1} \caption[Coal plants in India from 1990 to 2010]{Coal plants in India from 1990 to 2010}\label{fig:plants}\floatfoot*{Note: The top figure shows the location of coal plants in 1990.
+The bottom figure shows the location of coal plants in 2010.}
 \end{figure}
@@
 \begin{figure}
-\centering
-\includegraphics{draft_files/figure-latex/windexample-1.pdf}
-\caption{\label{fig:windexample}Wind direction (2010-01-01)}
+\includegraphics{draft_files/figure-latex/windexample-1} \caption[Wind direction and aggregation examples (2010-01-01)]{Wind direction and aggregation examples (2010-01-01)}\label{fig:windexample}\floatfoot*{Note: The top figure shows the average wind direction on January 1st, 2010.
+The points are the location of coal plants on that date.
+The bottom figure shows the distribution of pollution exposure in a specific district -- Kangra district in Himachal Pradesh -- on the same date.}
 \end{figure}
@@
 
-\begin{threeparttable}
 \caption{\label{tab:pollutiontable}Wind direction and particulate matter}
 \centering
+\begin{threeparttable}
 \begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
@@
 \midrule
-wind & 0.045*** & 0.040*** & 0.063*** & 0.037***\\
- & (0.004) & (0.006) & (0.005) & (0.007)\\
-\textbf{fixed effects:} & \textbf{---------} & \textbf{---------} & \textbf{---------} & \textbf{---------}\\
+wind & 0.045*** & 0.014*** & 0.063*** & 0.015***\\
+ & (0.004) & (0.001) & (0.005) & (0.002)\\
+\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
 village & Yes & Yes & Yes & Yes\\
-month & Yes & Yes & Yes & Yes\\
-\textbf{varying slopes:} & \textbf{---------} & \textbf{---------} & \textbf{---------} & \textbf{---------}\\
-month (village) & No & Yes & No & Yes\\
+month & Yes & No & Yes & No\\
+district-month & No & Yes & No & Yes\\
 \midrule
@@
 \begin{tablenotes}
-\small
-\item [] Note: Standard errors are in parentheses and are clustered at the village level.
-\item [] * p<0.10 ** p<0.05 *** p<0.01
+\item Note: Standard errors are in parentheses and are clustered at the village level.
+\item * p<0.10 ** p<0.05 *** p<0.01
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+\begin{table}
+
+\caption{\label{tab:yieldtable}Wind direction and agricultural productivity}
+\centering
+\begin{threeparttable}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\toprule
+\multicolumn{1}{c}{ } & \multicolumn{3}{c}{all} & \multicolumn{1}{c}{monsoon} & \multicolumn{1}{c}{winter} \\
+\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6}
+  & (1) & (2) & (3) & (4) & (5)\\
+\midrule
+wind & -0.003*** & -0.003*** & -0.0007*** & -0.002*** & -0.003***\\
+ & (0.0002) & (0.0002) & (9.21e-5) & (0.0002) & (0.0003)\\
+rain (z) &  & 0.029*** & 0.009*** & 0.082*** & 0.016***\\
+ &  & (0.0004) & (0.001) & (0.002) & (0.0004)\\
+\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+village-season & Yes & Yes & Yes & No & No\\
+year & Yes & Yes & No & Yes & Yes\\
+district-year-season & No & No & Yes & No & No\\
+village & No & No & No & Yes & Yes\\
+\midrule
+observations & 2,391,533 & 2,375,337 & 2,375,337 & 1,259,123 & 1,116,214\\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}
+\item Note: Standard errors are in parentheses and are clustered at the village level.
+\item * p<0.10 ** p<0.05 *** p<0.01
 \end{tablenotes}
@@
 
+\caption{\label{tab:yieldtabletwo}Pollution and agricultural productivity, IV estimates}
+\centering
 \begin{threeparttable}
-\caption{\label{tab:yieldtable}Wind direction and agricultural productivity}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3.5cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
 \toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{all} & \multicolumn{1}{c}{monsoon} & \multicolumn{1}{c}{winter} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5}
-  & (1) & (2) & (3) & (4)\\
+\multicolumn{1}{c}{ } & \multicolumn{3}{c}{all} & \multicolumn{1}{c}{monsoon} & \multicolumn{1}{c}{winter} \\
+\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6}
+  & (1) & (2) & (3) & (4) & (5)\\
 \midrule
-wind & -0.003*** & -0.003*** & -0.002*** & -0.003***\\
- & (0.0002) & (0.0002) & (0.0002) & (0.0003)\\
-rain (z) &  & 0.029*** & 0.082*** & 0.016***\\
- &  & (0.0004) & (0.002) & (0.0004)\\
-sub-sample & all & all & monsoon & winter\\
-\textbf{fixed effects:} & \textbf{---------} & \textbf{---------} & \textbf{---------} & \textbf{---------}\\
-village-season & Yes & Yes & No & No\\
-year & Yes & Yes & Yes & Yes\\
+particulate matter & -0.021*** & -0.020*** & -0.033*** & -0.013*** & -0.024***\\
+(PM 2.5, '000s) & (0.001) & (0.002) & (0.005) & (0.001) & (0.003)\\
+rain (z) &  & 0.004** & 0.002 & 0.086*** & -0.016***\\
+ &  & (0.002) & (0.002) & (0.002) & (0.004)\\
+\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+village-season & Yes & Yes & Yes & No & No\\
+year & Yes & Yes & No & Yes & Yes\\
+district-year-season & No & No & Yes & No & No\\
+village & No & No & No & Yes & Yes\\
 \midrule
-village & No & No & Yes & Yes\\
-observations & 2,391,533 & 2,375,337 & 1,259,123 & 1,116,214\\
+observations & 2,391,533 & 2,375,337 & 2,375,337 & 1,259,123 & 1,116,214\\
+\midrule
+first stage: &  &  &  &  & \\
+wind & 0.143*** & 0.126*** & 0.022*** & 0.155*** & 0.105***\\
+ & (0.003) & (0.003) & (0.002) & (0.003) & (0.004)\\
+rain (z) &  & -1.23*** & -0.235*** & 0.301*** & -1.36***\\
+ &  & (0.010) & (0.018) & (0.015) & (0.009)\\
 \bottomrule
@@
 \begin{tablenotes}
-\small
-\item [] Note: Standard errors are in parentheses and are clustered at the village level.
-\item [] * p<0.10 ** p<0.05 *** p<0.01
+\item Note: Standard errors are in parentheses and are clustered at the village level.
+\item * p<0.10 ** p<0.05 *** p<0.01
 \end{tablenotes}
@@
 
+\caption{\label{tab:yieldtablehet}Heterogeneity in the effects of pollution on productivity}
+\centering
 \begin{threeparttable}
-\caption{\label{tab:yieldtabletwo}Wind direction and agricultural productivity}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
 \toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{all} & \multicolumn{1}{c}{monsoon} & \multicolumn{1}{c}{winter} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5}
-  & (1) & (2) & (3) & (4)\\
+\multicolumn{1}{c}{ } & \multicolumn{1}{c}{>p(50)} & \multicolumn{1}{c}{<=p(50)} & \multicolumn{3}{c}{all} \\
+\cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-6}
+  & (1) & (2) & (3) & (4) & (5\\
 \midrule
-pm25 & -0.021*** & -0.020*** & -0.013*** & -0.024***\\
- & (0.001) & (0.002) & (0.001) & (0.003)\\
-rain (z) &  & 0.004** & 0.086*** & -0.016***\\
- &  & (0.002) & (0.002) & (0.004)\\
-sub-sample & all & all & monsoon & winter\\
-\textbf{fixed effects:} & \textbf{---------} & \textbf{---------} & \textbf{---------} & \textbf{---------}\\
-village-season & Yes & Yes & No & No\\
-year & Yes & Yes & Yes & Yes\\
-village & No & No & Yes & Yes\\
+wind & -0.003*** & -0.032*** & -0.004*** & 0.0004 & -0.0001\\
+ & (0.0002) & (0.002) & (0.0003) & (0.0005) & (0.001)\\
+rain (z) & 0.031*** & 0.030*** & 0.029*** & 0.029*** & 0.029***\\
+ & (0.0006) & (0.0005) & (0.0004) & (0.0004) & (0.0004)\\
+wind x rain &  &  & 0.0005*** &  & \\
+ &  &  & (6.67e-5) &  & \\
+wind squared &  &  &  & -0.0002*** & \\
+ &  &  &  & (3.54e-5) & \\
+wind x starting yield &  &  &  &  & -0.001**\\
+ &  &  &  &  & (0.0006)\\
+\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+village-season & Yes & Yes & Yes & Yes & Yes\\
+year & Yes & Yes & Yes & Yes & Yes\\
 \midrule
-observations & 2,391,533 & 2,375,337 & 1,259,123 & 1,116,214\\
-\midrule
-\textbf{first stage:} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-wind & 0.143*** & 0.126*** & 0.155*** & 0.105***\\
- & (0.003) & (0.003) & (0.003) & (0.004)\\
-rain (z) &  & -1.23*** & 0.301*** & -1.36***\\
- &  & (0.010) & (0.015) & (0.009)\\
+observations & 1,115,694 & 1,259,643 & 2,375,337 & 2,375,337 & 2,371,364\\
 \bottomrule
@@
 
-\begin{threeparttable}
 \caption{\label{tab:labortable}Wind direction and labor allocation}
 \centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\begin{threeparttable}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}}
 \toprule
-  & (1) & (2) & (3)\\
+  & all & all & self & wage & farm & non-farm\\
 \midrule
-wind & -0.028* & -0.035** & -0.027*\\
- & (0.015) & (0.015) & (0.015)\\
-controls & No & Yes & Yes\\
-\textbf{fixed effects:} & \textbf{-------} & \textbf{--------} & \textbf{-------}\\
-district & Yes & Yes & Yes\\
-year & Yes & Yes & Yes\\
-\textbf{varying slopes:} & \textbf{-------} & \textbf{--------} & \textbf{-------}\\
-year (district) & No & No & Yes\\
+wind & -0.021 & -0.027* & 0.008 & -0.034** & 0.035* & -0.061**\\
+ & (0.015) & (0.015) & (0.017) & (0.016) & (0.019) & (0.025)\\
+controls & No & Yes & Yes & Yes & Yes & Yes\\
+\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+district & Yes & Yes & Yes & Yes & Yes & Yes\\
+year & Yes & Yes & Yes & Yes & Yes & Yes\\
+\textbf{varying slopes:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+year (by district) & Yes & Yes & Yes & Yes & Yes & Yes\\
 \midrule
-observations & 899,045 & 898,856 & 898,856\\
+observations & 899,045 & 898,856 & 898,856 & 898,856 & 898,856 & 898,856\\
 \bottomrule
@@
 \begin{tablenotes}
-\small
-\item [] Note: Standard errors are in parentheses and are clustered at the village level.
-\item [] * p<0.10 ** p<0.05 *** p<0.01
+\item Note: Standard errors are in parentheses and are clustered at the district level. Control variables include female, age, age squared, and (years of) education.
+\item * p<0.10 ** p<0.05 *** p<0.01
 \end{tablenotes}
@@
 
+\begin{table}
+
+\caption{\label{tab:ntltable}Wind direction and nightlight growth}
+\centering
+\begin{threeparttable}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\toprule
+  & (1) & (2)\\
+\midrule
+wind (lagged) & -0.080*** & -0.083***\\
+ & (0.015) & (0.015)\\
+rain (z, lagged) &  & -0.035***\\
+ &  & (0.002)\\
+\textbf{fixed effects:} & \textbf{} & \textbf{}\\
+village & Yes & Yes\\
+year & Yes & Yes\\
+\textbf{varying slopes:} & \textbf{} & \textbf{}\\
+year (by village) & Yes & Yes\\
+\midrule
+observations & 2,146,715 & 2,146,715\\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}
+\item Note: Standard errors are in parentheses and are clustered at the village level.
+\item * p<0.10 ** p<0.05 *** p<0.01
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+
+\hypertarget{conclusion}{%
+\section{Conclusion}\label{conclusion}}
+
 \FloatBarrier
 \newpage
+
+\newpage
+\singlespacing
 
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-aragon2016polluting}{}}%
+Aragón, Fernando M, and Juan Pablo Rud. 2016. {``{Polluting industries and agricultural productivity: Evidence from mining in Ghana}.''} \emph{{The Economic Journal}} 126 (597): 1980--2011.
+
+\leavevmode\vadjust pre{\hypertarget{ref-arceo2016does}{}}%
+Arceo, Eva, Rema Hanna, and Paulina Oliva.
+2016.
+{``{Does the effect of pollution on infant mortality differ between developing and developed countries?
+Evidence from Mexico City}.''} \emph{{The Economic Journal}} 126 (591): 257--80.
+
 \leavevmode\vadjust pre{\hypertarget{ref-almn2021}{}}%
@@
 {``{Development Research at High Geographic Resolution: An Analysis of Night Lights, Firms, and Poverty in India using the SHRUG Open Data Platform}.''} \emph{{The World Bank Economic Review}}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-brunekreef2002air}{}}%
+Brunekreef, Bert, and Stephen T Holgate. 2002. {``{Air pollution and health}.''} \emph{{The Lancet}} 360 (9341): 1233--42.
+
+\leavevmode\vadjust pre{\hypertarget{ref-chang2019effect}{}}%
+Chang, Tom Y, Joshua Graff Zivin, Tal Gross, and Matthew Neidell.
+2019.
+{``{The effect of pollution on worker productivity: evidence from call center workers in China}.''} \emph{{American Economic Journal: Applied Economics}} 11 (1): 151--72.
+
+\leavevmode\vadjust pre{\hypertarget{ref-chen2022effect}{}}%
+Chen, Shuai, Paulina Oliva, and Peng Zhang. 2022. {``{The effect of air pollution on migration: evidence from China}.''} \emph{{Journal of Development Economics}} 156: 102833.
+
+\leavevmode\vadjust pre{\hypertarget{ref-currie2014we}{}}%
+Currie, Janet, Joshua Graff Zivin, Jamie Mullins, and Matthew Neidell.
+2014.
+{``{What do we know about short-and long-term effects of early-life exposure to pollution?}''} \emph{{Annual Review Resource Economics}} 6 (1): 217--47.
+
+\leavevmode\vadjust pre{\hypertarget{ref-ebenstein2016long}{}}%
+Ebenstein, Avraham, Victor Lavy, and Sefi Roth.
+2016.
+{``{The long-run economic consequences of high-stakes examinations: Evidence from transitory variation in pollution}.''} \emph{{American Economic Journal: Applied Economics}} 8 (4): 36--65.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-graff2012impact}{}}%
+Graff Zivin, Joshua, and Matthew Neidell. 2012. {``{The impact of pollution on worker productivity}.''} \emph{{American Economic Review}} 102 (7): 3652--73.
+
 \leavevmode\vadjust pre{\hypertarget{ref-hammer2020global}{}}%
@@
 {``Global Estimates and Long-Term Trends of Fine Particulate Matter Concentrations (1998--2018).''} \emph{{Environmental Science \& Technology}} 54 (13): 7879--90.
+
+\leavevmode\vadjust pre{\hypertarget{ref-hanna2015effect}{}}%
+Hanna, Rema, and Paulina Oliva. 2015. {``{The effect of pollution on labor supply: Evidence from a natural experiment in Mexico City}.''} \emph{{Journal of Public Economics}} 122: 68--79.
+
+\leavevmode\vadjust pre{\hypertarget{ref-he2019severe}{}}%
+He, Jiaxiu, Haoming Liu, and Alberto Salvo.
+2019.
+{``{Severe air pollution and labor productivity: Evidence from industrial towns in China}.''} \emph{{American Economic Journal: Applied Economics}} 11 (1): 173--201.
+
+\leavevmode\vadjust pre{\hypertarget{ref-heck1982assessment}{}}%
+Heck, Walter W, OC Taylor, Richard Adams, Gail Bingham, Joseph Miller, Eric Preston, and Leonard Weinstein.
+1982.
+{``{Assessment of crop loss from ozone}.''} \emph{{Journal of the Air Pollution Control Association}} 32 (4): 353--61.
+
+\leavevmode\vadjust pre{\hypertarget{ref-heft2018robust}{}}%
+Heft-Neal, Sam, Jennifer Burney, Eran Bendavid, and Marshall Burke. 2018. {``{Robust relationship between air quality and infant mortality in Africa}.''} \emph{Nature} 559 (7713): 254--58.
+
+\leavevmode\vadjust pre{\hypertarget{ref-kampa2008human}{}}%
+Kampa, Marilena, and Elias Castanas. 2008. {``{Human health effects of air pollution}.''} \emph{{Environmental Pollution}} 151 (2): 362--67.
+
+\leavevmode\vadjust pre{\hypertarget{ref-marshall1997hidden}{}}%
+Marshall, Fiona, Mike Ashmore, Fiona Hinchcliffe, et al.
+1997.
+\emph{{A hidden threat to food production: Air pollution and agriculture in the developing world}}.
+{International Institute for Environment and Development.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-pope2006health}{}}%
+Pope III, C Arden, and Douglas W Dockery.
+2006.
+{``{Health effects of fine particulate air pollution: lines that connect}.''} \emph{{Journal of the Air \& Waste Management Association}} 56 (6): 709--42.
 
@@
 
-\FloatBarrier
-\singlespacing
-
 \begin{table}
 
-\caption{\label{tab:pmeffects}Particulate matter and agricultural productivity}
+\caption{\label{tab:yieldtableleads}Agricultural productivity and pollution leads}
 \centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\begin{threeparttable}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
 \toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{all} & \multicolumn{1}{c}{monsoon} & \multicolumn{1}{c}{winter} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5}
-  & (1) & (2) & (3) & (4)\\
+  & (1) & (2)\\
 \midrule
-particulate matter & 0.045 & 0.501*** & 2.05*** & 0.472***\\
-(PM 2.5, '000s) & (0.031) & (0.031) & (0.099) & (0.037)\\
-rain (z) &  & 0.030*** & 0.083*** & 0.017***\\
- &  & (0.0004) & (0.002) & (0.0004)\\
-sub-sample & all & all & monsoon & winter\\
-\textbf{fixed effects:} & \textbf{-------} & \textbf{---------} & \textbf{---------} & \textbf{---------}\\
-village-season & Yes & Yes & No & No\\
-year & Yes & Yes & Yes & Yes\\
-village & No & No & Yes & Yes\\
+wind (lag) & 0.004*** & 0.0006*\\
+ & (0.0007) & (0.0003)\\
+wind & -0.002*** & -0.0007***\\
+ & (0.0004) & (0.0002)\\
+wind (lead) & -0.004*** & -0.0001\\
+ & (0.0001) & (9.8e-5)\\
+rain (z) & 0.021*** & 0.008***\\
+ & (0.0005) & (0.001)\\
+\textbf{fixed effects:} & \textbf{} & \textbf{}\\
+village-season & Yes & Yes\\
+year & Yes & No\\
+district-year-season & No & Yes\\
 \midrule
-observations & 2,391,533 & 2,375,337 & 1,259,123 & 1,116,214\\
-\midrule
+observations & 1,979,635 & 1,979,635\\
 \bottomrule
-\multicolumn{5}{l}{\textsuperscript{} Note: Standard errors are in parentheses and are clustered at the village}\\
-\multicolumn{5}{l}{level.}\\
-\multicolumn{5}{l}{\textsuperscript{} * p<0.10 ** p<0.05 *** p<0.01}\\
 \end{tabular}
+\begin{tablenotes}
+\item Note: Standard errors are in parentheses and are clustered at the village level.
+\item * p<0.10 ** p<0.05 *** p<0.01
+\end{tablenotes}
+\end{threeparttable}
 \end{table}

### 2023-01-03 analyses/regressions.R

@@ -7,6 +7,7 @@ rm(list = ls())
 library(tidyverse)
 library(lubridate)
 library(fixest)
+library(sf)
 
 
 # Sets wd to folder this script is in so we can create relative paths instead of absolute paths
@@ -31,11 +32,25 @@ pollution <- pollution %>% mutate(
 pollution <- pollution %>% mutate()
 pollution <- panel(pollution, ~ village + date)
 
+# And district identifiers
+villages_overlap <- read_sf("data/spatial/villages_overlap/villages_overlap.shp")
+villages_overlap <- villages_overlap %>% mutate(shrid = paste0(pc11_s_id, "-", pc11_tv_id),
+                                                distfe = paste0(pc11_s_id, "-", pc11_d_id)) %>%
+                                          as_tibble() %>%
+                                          dplyr::select(shrid, distfe) %>%
+                                          group_by(shrid) %>%
+                                          filter(row_number()==1) %>%
+                                          ungroup()
+pollution <- pollution %>% left_join(villages_overlap, by = "shrid")
+
+
+
+
 
 pol1 <- feols(mean ~ wind | village + month, data = pollution, cluster = c("village"))
-pol2 <- feols(mean ~ wind | village[month] + month, data = pollution, cluster = c("village"))
+pol2 <- feols(mean ~ wind | village + month^distfe, data = pollution, cluster = c("village"))
 pol3 <- feols(mean ~ wind | village + month, data = pollution %>% filter(year>=2002 & year<=2013), cluster = c("village"))
-pol4 <- feols(mean ~ wind | village[month] + month, data = pollution %>% filter(year>=2002 & year<=2013), cluster = c("village"))
+pol4 <- feols(mean ~ wind | village + month^distfe, data = pollution %>% filter(year>=2002 & year<=2013), cluster = c("village"))
 
 pollutiontable <- etable(
                           pol1, pol2, pol3, pol4, 
@@ -46,12 +61,11 @@ pollutiontable <- etable(
                           fitstat = c("n"),
                           coefstat = "se"
                           )
-pollutiontable <- pollutiontable[-9,]
-pollutiontable <- pollutiontable[-8,]
+pollutiontable <- pollutiontable[-c(7:8),]
 pollutiontable <- as.matrix(pollutiontable)
 rownames(pollutiontable) <- c("wind", "", "fixed effects:",
[…]

### 2023-01-03 cleaning/cleaning_overlap_merging_nss.R

@@ -71,8 +71,6 @@ df_labor <- df_labor %>% filter(is.na(merge)==F)
 
 
 # Going to go through the dates and process date by date
-pollution <- read_csv(paste0("data/clean/pm25/all_combined.csv"))
-pollution <- pollution %>% rename(pm25 = mean)
 df_labor_merged <- c()
 for (day in 1:length(date_vec)){
   # First, load villages
@@ -114,6 +112,41 @@ df_labor_merged <- df_labor_merged %>% mutate(distfe = paste0("s", state_merge, 
 write.csv(df_labor_merged, "data/clean/nss/merged_week.csv")
 
 
+
+
+
+
+
+pollution <- read_csv(paste0("data/clean/pm25/all_combined.csv"))
+pollution <- pollution %>% rename(pm25 = mean)
+df_labor_merged <- read_csv("data/clean/nss/merged_week.csv")
+
+# And also pm; all years covering the nss data
+pollution_merged <- c()
+for (y in year(min(df_labor_merged$date)):year(max(df_labor_merged$date))){
+  for (m in 1:12){
+    pollution_merge <- pollution %>% filter(year==y & month==m)
+    
+    village_wind <- villages %>% left_join(pollution_merge, by = "shrid")
+    # Weighted mean by AREA. Any parts of the district without villages are ZEROS
+    village_wind <- village_wind %>% 
+                    group_by(state, district) %>%
+                    mutate(pm25 = weighted.mean(pm25, area_weight, na.rm = T),
+                           state_merge = as.numeric(state),
+                           district_merge = as.numeric(district),
+                           tot_weight = sum(area_weight),
+                           pm25 = pm25*tot_weight,
+                           year = y, month_int = m) %>% # Last one adjusts for the zeros for areas not here
+                    filter(row_number()==1) %>%
+                    ungroup() %>%
+                    select(state_merge, district_merge, year, month_int, pm25)
+    
+    pollution_merged <- rbind(pollution_merged, village_wind)
+    
+    print(paste0(y, "-", m))
+  }
+}
+write_csv(pollution_merged, paste0("data/clean/pm25/nss_merge.csv"))
 
 
 

### 2023-01-09 draft/draft.Rmd
updated draft

@@
 
+reportDecOne <- function(dec){
+
+    result <- sprintf("%.1f", dec) # inserts a float into a string and simultaneously do rounding
+
+  return(result)
+}
+
 reportPercent <- function(dec){
@@
 labortable <- readRDS("pollution_development/draft/tables/labortable.rds")
-#labortable_two <- readRDS("pollution_development/draft/tables/labortable_two.rds")
+labortablemonsoon <- readRDS("pollution_development/draft/tables/labortablemonsoon.rds")
+labortablewinter <- readRDS("pollution_development/draft/tables/labortablewinter.rds")
 ntltable <- readRDS("pollution_development/draft/tables/ntltable.rds")
 yieldtableleads <- readRDS("pollution_development/draft/tables/yieldtableleads.rds")
+yieldtablepostplant <- readRDS("pollution_development/draft/tables/yieldtablepostplant.rds")
+laborplot <- readRDS("pollution_development/draft/tables/laborplot.rds")
+
+plants1990_capacity <- plants %>% filter(year_built<=1990 & (year_retired>1990 | is.na(year_retired)==TRUE))
+plants2010_capacity <- plants %>% filter(year_built<=2010 & (year_retired>2010 | is.na(year_retired)==TRUE))
 
@@
 
-pollution could be bad for agriculture [@heck1982assessment,@marshall1997hidden]
+pollution could be bad for agriculture [@heck1982assessment;@marshall1997hidden]
 
@@
 
-We have known about the effects of pollution on human health for many years [@brunekreef2002air,@kampa2008human,@pope2006health]
+We have known about the effects of pollution on human health for many years [@brunekreef2002air;@kampa2008human;@pope2006health]
 good review on the effects of pollution early in life in @currie2014we
-  - also in developing countries, where there is a consistent effect of exposure on infant mortality [@heft2018robust]
+  - also in developing countries, where there is a consistent effect of exposure on infant mortality [@heft2018robust;@heft2019air]
 
@@
   - call center workers in China, and at common levels of pollution [@chang2019effect]
-  - tests [@ebenstein2016long]
+  - tests [@ebenstein2016long;@wen2022lower]
   - farm workers in California [@graff2012impact]
@@
 
-@almn2021
-Ag productivity: @gangopadhyay2022new
-PM estimates: @hammer2020global
+The main goal of this paper is to examine whether exposure to exhaust from coal plants affects agricultural productivity, labor allocation, and economic growth.
+I use several sources of data, which I describe here and in Table \@ref(tab:data) in the appendix.
 
-```{r data, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-data <- matrix(NA, nrow = 3, ncol = 5)
-colnames(data) <- c("shapefile", "wind", "pollution", "agriculture", "nightlights")
-rownames(data) <- c("source", "geographic coverage", "temporal coverage")
-data[1,1] <- "Asher et al. (2021)"
-data[1,2] <- "NCAR"
-data[1,3] <- "Hammer et al. (2020)"
-data[1,4] <- "Angopadhyay et al. (2022)"
-data[1,5] <- "Asher et al. (2021)"
-data[2,1] <- "India"
-data[2,2] <- "global"
-data[2,3] <- "global"
-data[2,4] <- "global"
-data[2,5] <- "India"
-data[3,1] <- " "
-data[3,2] <- "daily"
-data[3,3] <- "monthly"
-data[3,4] <- "two seasons/year"
-data[3,5] <- "yearly"
-
-# Table
-kable(
-      data, caption = "Remote sensing data sources",
-      align = "ccccc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           general = "NCAR: https://climatedataguide.ucar.edu/", 
-           general_title = " ",
-           threeparttable = TRUE
-           ) %>%
-  column_spec(1, width = "3cm") %>%
-  column_spec(c(2:6),width = "2cm") %>%
-  kable_classic_2()
-
-```
-
-## Methodology
+The first set of data lists the location of coal plant across the globe.
+This data comes from Global Energy Monitor.^[https://globalenergymonitor.org/projects/global-coal-plant-tracker/] and lists all units generating at least 30 megawatts of electricity.
+The data on coal plants includes the year of opening (and, if applicable, the year of retirement), the GPS (latitude/longitude) location of each plant, and the amount of power produced by the units at each plant.
+For this paper, I do not use the information on the capacity of the plant.
 
@@
 
-```{r windexample, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 8, fig.cap = "Wind direction and aggregation examples (2010-01-01)", fig.footer = list("Note: The top figure shows the average wind direction on January 1st, 2010.
+Figure \@ref(fig:plants) presents the location of coal plants in India for two specific years: 1990 and 2010.
+There is a clear increase in the prevalence of coal plants across the country over the two decades.
+Additionally, the overall capacity from coal plants increased from `r paste0(reportDecOne(sum(plants1990_capacity$capacity)))` gigawatts to `r paste0(reportDecOne(sum(plants2010_capacity$capacity)))` gigawats, an increase of `r paste0(reportPercent((sum(plants2010_capacity$capacity) - sum(plants1990_capacity$capacity))/sum(plants1990_capacity$capacity)))` percent in just 20 years.
+
+The second dataset includes agricultural productivity for both the monsoon and winter seasons, from 2002 to 2018.
+To match other data, I only use the data up to 2013.
+This data comes from @gangopadhyay2022new and estimates land productivity (i.e.
+yield, in tons per hectare) for the major crops in India.
+They define the monsoon season as June to October and the winter season as November to March.
+I keep these definitions when matching across data below.
+This data is also publicly available from Nature's data-sharing website.^[https://springernature.figshare.com]
+
+The agricultural productivity data is available as raster files with a resolution of 500m.
+To aggregate this data up to a useful administrative unit, I use village-level shapefiles provided by @almn2021 and publicly available on the SHRUG platform.^[https://www.devdatalab.org/shrug_download/] I aggregate the agricultural productivity data to the village level by extracting mean productivity for each feature in the shapefile.
+I do this separately for each season -- monsoon and winter -- and each year, from 2002 to 2013.
+
+To measure exposure to pollution from coal plants, I first locate all village centroids located within 30km from a coal plant in a given year.
+I choose 30km due to previous research on the effects of (air) pollution [@aragon2016polluting;@bencko1977health;@li2014health].
+I then calculate the direction from coal plants to all village centroids within that 30km radius.
+To define exposure, I then pull daily wind direction data from the National Center for Atmospheric Research.^[https://climatedataguide.ucar.edu/] For each day, I document whether the wind is blowing towards each village centroid.^[I define "towards" as within five degrees to help take into account that I am using village centroids, using the x and y components of wind speed.] I then temporally aggregate this daily data depending on the temporal definition of the corresponding outcome.
+For example, agricultural productivity is defined across five months (e.g.
+the monsoon season is from June to October) so I count the total days a given village is exposed to wind during those five months.
+
+```{r windexample, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 8, fig.cap = "Wind direction and aggregation examples (2010-01-01)", fig.footer = list("Note: The left figure shows the average wind direction on January 1st, 2010.
 The points are the location of coal plants on that date.
-The bottom figure shows the distribution of pollution exposure in a specific district -- Kangra district in Himachal Pradesh -- on the same date.")}
+The right figure shows the distribution of pollution exposure in a specific district -- Kangra district in Himachal Pradesh -- on the same date.")}
 
@@
 
-\newpage
+The top panel in Figure \@ref(fig:windexample) is an example wind direction raster.
+The raster shows wind direction for the entirety of India on January 1st, 2010, as well as the location of coal power plants on that date.
+The prevailing winds on the date differ across the country and, though not shown in the figure, across days.
+This means that the overall exposure for a given area to coal plant emissions changes across time.
+I also pull data on particulate matter to help validate this proxy for exposure.
+This data comes from @hammer2020global and I also aggregate this to the village using the same method detailed above.
+
+To help understand some of the mechanisms driving the relationship between exposure to emissions and agricultural productivity, I also include household survey data.
+Unfortunately, I am not aware of any publicly available survey data that covers multiple years and has identifiers below the district level.
+For example, the ARIS-REDS data^[https://www.ncaer.org/project/additional-rural-incomes-survey-aris-rural-economic-demographic-survey-reds] includes GPS identifiers for one year, but that only allows GPS matching for a total of two years: 1999 and 2007/2008.
+The DHS, on the other hand, only has district-level identifiers (except for the newest two waves) and is also not a labor survey.
+Instead, I use the National Sample Survey (NSS) from the Ministry of Statistics and Programme Implementation.
+
+The NSS is a nationally representative survey that was conducted every couple of years until relatively recently.
+Many rounds of the NSS also included a module on employment that asked about labor allocation over the previous seven days.
+I use the 61st, 62nd, 64th, 66th, and 68th rounds of the NSS, all of which include the employment module.
+The publicly available data also includes the date of survey, which allows me to match pollution exposure to the exact day of the survey.
+
+The biggest issue with using the NSS is that it includes only district-level identifiers.
+The main exposure variables, on the other hand, are defined at a lower level of aggregation, at the village.
+I choose to aggregate this village-level exposure to the district level as shown in the second panel of Figure \@ref(fig:windexample).
+I use the area overlap of the villages with the district to define a weighted average of exposure for the district at a given time.
+I calculate the area of overlap for each village and the district as well as the overall area of the district and use these variables to define weights for individual villages.
+Any part of the district that does not overlap with one of the villages -- the white area in Figure \@ref(fig:windexample) -- does not have any exposure and, as such, receives a value of zero for exposure.
+
+Finally, I use nightlights as a measure of economic growth [@henderson2012measuring].
+I also take this data directly from SHRUG and @almn2021.
+This covers 1992 through 2013.
+Some of the older data may not be ideal [@gibson2021night], but this is the only way to document changes in economic growth at a more disaggregated level and at a regular basis.
+Although satellite imagery is an attractive alternative [@burke2021using], high-resolution satellite imagery does not go back to the early 1990s.
+
+## Identification
+
+To calculate the effects of exposure to pollution on different outcomes, I rely on fixed effects and the differential exposure driven by plausibly exogenous wind direction, conditional on total rainfall.
+The regession is of the form
+\begin{gather} y_{it} = \alpha_{i} + \gamma_{t} + \beta wind_{it} + \delta rain_{it} + \varepsilon_{it}, \end{gather}
+where $y_{it}$ is outcome $y$ for unit $i$ at time $t$, $\alpha$ is a geographic fixed effects, $\gamma_t$ is time fixed effects, $wind_{it}$ is exposure to pollution -- defined using wind direction -- and $rain_{it}$ is rainfall.
+Although I use $i$ and $t$ as subscripts for the fixed effects, note that these change depending on the unit of analysis and the outcome.
+I explicitly state the level of fixed effects in the tables and results below.
+
+Identification relies on changes in wind direction across time for the same geographic units.
+In other words, conditional on the fixed effects and rainfall, I assume changes in wind direction are as good as random and uncorrelated with the outcome -- e.g.
+agricultural productivity -- except through exposure to pollution.
+One major threat to identification is that the building of a coal plant could lead to a spurious correlation between the exposure variable and the outcome.
+Specifically, since exposure is defined as zero prior to the building of a coal plant, the exposure variable itself might be endogenous even though wind direction is not.
+To help allay these concerns, I estimate additional regressions restricting the sample to only time periods after which a given location has a coal plant.
+I also present results using leads to show that future wind direction does not seem to predict current outcomes.
+
+## An example: validating wind direction
+
+Before moving on to the main results, it is worth taking the time to better understand the fixed effects strategy with an example.
+Consider the data used for the regressions presented in Table \@ref(tab:pollutiontable).
+The outcome variable is particulate matter -- specifically, $\mathrm{PM_{2.5}}$, which is particulate matter no larger than 2.5 micrometers in diameter -- which comes from @hammer2020global.
+$\mathrm{PM_{2.5}}$ is one of the harmful byproducts from coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.^[https://www.epa.gov/airmarkets/power-plants-and-neighboring-communities] This data is available at the monthly level, so in Table \@ref(tab:pollutiontable), I aggregate wind exposure to the monthly level, as well.
 
@@
 
+Both the particulate matter data and the wind data is at the village level, which means the regressions in Table \@ref(tab:pollutiontable) are at the village level.
+As such, the geographic fixed effects are village fixed effects and the temporal fixed effects are month fixed effects.
+Since the exposure variable is at the village level and we follow villages over time, the standard errors are also clustered at the village level.
+
+In addition to serving as an example, Table \@ref(tab:pollutiontable) also helps validate the use of wind direction as a proxy for exposure to pollution.
+The first two columns include all available years of data while the last two columns restrict estimation only to the years for which the agricultural productivity regressions are estimated.
+Across all four columns, the story is the same: when the wind is blowing in the direction of a village from a coal plant within 30 kilometers, estimated $\mathrm{PM_{2.5}}$ is substantially higher.
+The regressions indicate that one additional day of wind in the direction of the village increases mean $\mathrm{PM_{2.5}}$ for the month by between 0.014 and 0.045 $\mathrm{\mu g/m^3}$.
+
+To put these numbers into perspective, the World Health Organization's updated guidelines are that average annual exposure should not exceed 15 $\mathrm{\mu g/m^3}$.
+For the month, an increase at the midpoint of the estimated range is equal to an increase of 0.2 percent of the maximum recommended mean concentration from the WHO.
+This is just a single day of wind in the direction of a village.
+Since this includes all villages located within 30km of a coal plant, this is evidence that wind direction can have serious repercussions on the health of hundreds of millions of people in India.
+
 # Results
+
+This section presents the main results of this paper.
+I present three sets of results: one set looking at agricultural productivity, another looking at labor allocation, and a final set looking at nightlights, which are a proxy for economic development.
+I go through each of these in turn.
+
+## Agricultural productivity
+
+I first present results for agricultural productivity.
+The outcome of all the regressions in this section is the log of agricultural land productivity, which is defined as tons per hectare.
+The first set of results are in Table \@ref(tab:yieldtable).
+The unit of analysis is the year season -- from 2002 to 2013 -- the geographic fixed effect is the village-season, and the temporal fixed effect is the year.
+The exposure variable is defined for the entire season.
+Since @gangopadhyay2022new define both the monsoon (kharif) season and the winter (rabi) season as five months long, I do as well.
+This implies that the exposure variable ranges from zero to slightly more than 150.
+
+The first column presents the most simple results, with only the wind exposure variable and the fixed effects.
+The second column adds a rainfall variable -- defined as the total rainfall in the season -- the addition of which is motivated by the fact that changes in wind direction could be correlated with changes in weather.
+However, the coefficient on the wind exposure variable is completely unchanged by the addition of rainfall, indicating that weather may not be as worrisome of a confounder as thought.
+
+Column three adds district-by-year-by-season fixed effects as an additional robustness checks.
+This addition decreases the magnitude of the exposure variable but it remains negative and significant.
+However, part of this may be due to the fact that the new fixed effects soak up much of the variation in the exposure variable, as shown in Figure \@ref(fig:windexample).
+The last two columns separate the effects into the monsoon season and the winter season, but the overall effect is roughly similar, though rainfall has a larger effect in the monsoon season than the winter season, which is unsurprising.
 
@@
 
+To put the size of the coefficient in context, it is worth digging a little deeper into the exposure variable.
+The mean within-village absolute deviation in wind exposure is approximately 8.06 days, meaning that the average change in agricultural productivity from year-to-year due to nothing but changes in wind patterns carrying particulate matter is around 2.4 percent relative to the mean.
+This implies that swings in agricultural productivity -- due to wind direction -- of more than five percent are probably quite common, since the absolute deviation includes deviations below and above the mean.
+
 ```{r yieldtabletwo, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
@@
   column_spec(c(2:6),width = "2cm") %>%
-  row_spec(5, bold = TRUE) %>%
+  row_spec(c(5,11), bold = TRUE) %>%
   row_spec(c(9:10), hline_after = TRUE) %>%
@@
 
+If you believe that wind direction is exogenous and, conditional on the fixed effects and rainfall, only affects agricultural productivity through pollution -- which seems reasonable since it is hard to imagine what other variables it might affect -- then it is a valid instrument for
+pollution.
+While I opt to present the reduced form results for the rest of this paper, it is nonetheless an interesting exercise to also use wind direction as an instrument for pollution, which I measure using estimated $\mathrm{PM_{2.5}}$ from @hammer2020global.
+
+The IV results appear in Table \@ref(tab:yieldtabletwo), with the second-stage results in the top panel and the first-stage results in the bottom panel.
+The first-stage results make clear how strong of an instrument wind is; also note that the coefficients are quite different from those in Table \@ref(tab:pollutiontable) because the unit of analysis in Table \@ref(tab:yieldtabletwo) is the season, while in Table \@ref(tab:pollutiontable) the unit of analysis is the month.
+
+The second-stage results indicate that $\mathrm{PM_{2.5}}$ is a very strong predictor of agricultural productivity.
+The smallest coefficient -- that for the monsoon season in column (4) -- and the mean absolute deviation for $\mathrm{PM_{2.5}}$ point to a change in agricultural productivity of around 23 percent.
+
 ```{r yieldtablehet, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
@@
 # Table
-colnames(yieldtablehet) <- c("(1)", "(2)", "(3)", "(4)", "(5")
+colnames(yieldtablehet) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
 kable(
@@
 ```
+
+Having established that exposure to pollution has a negative effect on agricultural productivity, I now move to heterogeneity in these effects.
+Table \@ref(tab:yieldtablehet) presents four different analyses: effects of exposure by maximum exposure, multiplicative effects of shocks, the possibility of non-linearities, and the differential effects of exposure based on initial agricultural productivity.
+
+The first two columns split the sample based on average exposure; villages with maximum exposure above the median are included in column (1) while those below the median are in column (2).
+Interestingly, the largest effects are seen on villages in the lower half of the distribution.
+Apparently, smaller maximum effects lead to larger overall effects of pollution exposure on agricultural productivity.
+In other words, it is not the places where wind is most likely to blow from a coal plant that experience the largest decrease in productivity in response to an additional day of wind-driven pollution.
+
+Column three looks add the possible multiplicative effects of rainfall shocks and pollution shocks on agricultural productivity.
+In previous tables, rainfall always entered the regression positively -- as we would expect -- indicating that higher rainfall in a given season leads to higher agricultural productivity.
+However, an interesting question is whether rainfall and pollution shocks might interact.
+The positive coefficient on the interaction term between days of wind and rainfall during the agricultural season indicates that there is a multiplicative effect of shocks; when wind days are higher (bad) and rainfall is lower (bad), the overall negative effects are even worse than when just one shock happens.
+Considering that climate change is leading to more variable rainfall and more negative rainfall shocks, these results are potentially worrisome, especially for smallholders.
+
+Column four adds a quadratic in wind to see if the negative effects of pollution on agricultural productivity are increase or decreasing in wind.
+In column four, the coefficient on the linear term is 0.0004 while the coefficient on the quadratic term is -0.0002, meaning that the effect of wind is decreasing for any days of wind over two.
+The negative coefficient also indicates that this negative effect is consistently decreasing, meaning the effect gets worse as average wind days gets higher.
+Note that this is somewhat different from the regressions in the first two columns, which split the sample by the *maximum* days of wind observed in the sample, not the mean.
+
+In the final column I add to the regression an interaction between wind and initial agricultural productivity (defined in 2002, the first year observed in the productivity data).
+The linear term drops out of the regression since it is invariant within villages, leaving just the interaction term.
+The negative coefficient on the interaction term of -0.001 indicates that the effect of wind is more negative in areas where agricultural productivity is higher.
+Note that agricultural productivity is defined in logs, so this can be interpreted as a percentage effect, not a level effect.
+Apparently, wind-driven pollution is particularly detrimental to agricultural productivity in the most productive places.
+
+## Robustness checks
+
+I present two separate robustness checks in the appendix.
+The first set of analyses attempts to separate the possible endogenous location of coal plants from outcomes.
+Specifically, prior to the opening of a new coal plant in an area, all villages receive a zero for wind.
+This means that, even if wind direction is exogenous, the opening of the coal plant could lead to correlations between the error term and the wind variable.
+
+Although the results with district-year-season help mitigate this concern, Table \@ref(tab:yieldtablepostplant) in the appendix checks this possibility in a different way.
+Specifically, I restrict the sample to only village-years *after* the introduction of a coal plant.
+This drops any pre-plant-opening observations -- all of which are mechanically zeros for wind -- to prevent endogeneity due to the opening.
+The coefficients in Table \@ref(tab:yieldtablepostplant) show that the overall effects are essentially unchanged, with just a small change in coefficient magnitude (driven mostly by rounding).
+
+The second robustness check includes leads for wind to see if they predict current year productivity.
+I present these results in \@ref(tab:yieldtableleads).
+The second column includes district-year-season fixed effects, while the first does not.
+The results make clear the importance of including trends or district-year fixed effects.
+In the first column, leads seem to predict productivity.
+However, this relationship disappears in the second column, indicating that the previous results with district-year fixed effects are likely the most reliable.
+
+## Labor allocation
+
+The previous set of results show a consistent negative effect of wind-driven pollution on agricultural productivity.
+A key question is the mechanism through which this happens; is this driven by changes in labor allocation/productivity or by changes in land productivity?
+While I cannot definitely prove the mechanism, I can shed some light on possibilities by looking at labor allocation decisions.
+Since these shocks generally happen after initial planting decisions, it seems unlikely that changes in land *allocation* could drive the results.
+
+Table \@ref(tab:labortable) presents results for labor allocation over the last seven days using the National Sample Survey and the methodology described above.
+Since the labor recall is for seven days, labor allocation ranges from zero to seven, as does the wind variable.
+Recall that the data is at the district level, not the village level.
+However, it is also more temporally disaggregated, with rolling interviews throughout the year-long survey.
+Figure \@ref(fig:laborplot) in the appendix shows this temporal distribution across the five waves of the survey.
+
+There are some interesting results in Table \@ref(tab:labortable).
+The first two columns show a marginally significant decrease in overall labor allocation (where "labor" is defined as non-domestic self and wage employment).
+The third and fourth columns show that this overall decrease is predominantly from wage employment, not self employment.
+An additional day of wind over the last seven decreases wage employment by 0.034 days in the district.
+Mean wage employment is 1.458 days, meaning one day of wind decreases wage employment by 2.3 percent of the mean.
+Since much of the district is not covered by "treated" villages, this is actually a rather substantive effect on the villages involved.
+
+The last two columns split the effects by sector, into farm and non-farm employment.
+Although agricultural (land) productivity decreases when wind blows in the direction of villages, overall labor allocation to farm labor actually goes *up*, by around 4.5 percent of the mean of 0.784 days.
+Non-farm labor, on the other hand, goes down, by around 2.4 percent of the mean of 2.527 days.
+
+There are two possible explanations for this pattern of results.
+First, it could be that land productivity goes down, leading farmers to increase their farm labor allocation in an attempt to compensate for the decreased productivity.
+A second possibility is that pollution does not affect land productivity at all, at least not directly.
+In this scenario, the worsening air quality could lead to decreased labor productivity, meaning that "effective labor" essentially decreases when pollution is higher.
+Unfortunately, I am not able to disentangle these two possiblities with the data available.
+
+Table \@ref(tab:labortablemonsoon) and Table \@ref(tab:labortablewinter) in the appendix disaggregate these effects into the monsoon and winter seasons.
+Apparently, the entirety of the effect is driven by changes in labor allocation during the monsoon season.
+The dynamics across seasons apparently lead to large differences in the effects of pollution on labor allocation.
 
@@
 
+## Nightlights
+
+Pollution leads to lower agricultural productivity as well as changes in labor allocation.
+This leads to an obvious question: might higher levels of pollution also directly affect economic growth?
+To explore this question, I look at changes in nightlights relative to changes in pollution.
+It seems reasonable to think that pollution might affect nightlights with a lag, so I include the lag of pollution rather than contempraneous pollution.
+
 ```{r ntltable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
@@
 
+I present these results in Table \@ref(tab:ntltable).
+To help simplify the coefficients, I recode wind to be the proportion of days in the year that a given village experiences wind in their direction, from a coal plant.
+Mean wind in this new specification is 0.066, indicating that a village in the sample experiences wind in its direction on around 6.6 percent of days, or 24 days of the year.
+
+We see a strong, negative relationship between pollution (lagged wind) and nightlights.
+Going from the 25th percentile to the 75th percentile of wind exposure is synonymous with a decrease in the growth of nightlights within a village by around 0.7 percent.
+While this is a relatively small magnitude, it is worth noting that there are more than 100,000 villages in this sample.
+While the effect for a single village might be small, the overall, aggregate effect on growth is quite substantial.
+
 # Conclusion
@@
 \FloatBarrier
-\newpage
-
 \newpage
@@
 :::
+
+\FloatBarrier
+\newpage
+
+# Appendix A {.unnumbered}
+\beginappendix
 \FloatBarrier
 
-```{=tex}
-\beginappendix
+```{r data, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+data <- matrix(NA, nrow = 6, ncol = 3)
+rownames(data) <- c("shapefile", "coal plants", "wind", "pollution", "agriculture", "nightlights")
+colnames(data) <- c("source", "geographic coverage", "temporal coverage")
+data[1,1] <- "Asher et al. (2021)"
+data[2,1] <- "Global Energy Monitor"
+data[3,1] <- "NCAR"
+data[4,1] <- "Hammer et al. (2020)"
+data[5,1] <- "Angopadhyay et al. (2022)"
+data[6,1] <- "Asher et al. (2021)"
+data[1,2] <- "India"
+data[2,2] <- "global"
+data[3,2] <- "global"
+data[4,2] <- "global"
+data[5,2] <- "global"
+data[6,2] <- "India"
+data[1,3] <- " "
+data[2,3] <- "yearly"
+data[3,3] <- "daily"
+data[4,3] <- "monthly"
+data[5,3] <- "two seasons/year"
+data[6,3] <- "yearly"
+
+# Table
+kable(
+      data, caption = "Remote sensing data sources",
+      align = "ccccc", booktabs = TRUE, linesep = ""
+      ) %>%
+  footnote(
+           general = c(
+                       "Note: NCAR stands for the National Center for Atmospheric Research.
+Additional information on Global Energy Monitor and NCAR can be found on their websites, at https://globalenergymonitor.org/projects/global-coal-plant-tracker and https://climatedataguide.ucar.edu/, respectively."
+                       ), 
+           general_title = "",
+           threeparttable = TRUE
+           ) %>%
+  column_spec(1, width = "2cm") %>%
+  column_spec(2,width = "4.5cm") %>%
+  column_spec(c(3:4),width = "3.5cm") %>%
+  kable_styling(latex_options = "HOLD_position") %>%
+  kable_classic_2()
+
 ```
+
 \newpage
-# Appendix A {.unnumbered}
+```{r yieldtablepostplant, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
 
+# Table
+colnames(yieldtablepostplant) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
+kable(
+      yieldtablepostplant, caption = "Wind direction and agricultural productivity, post-coal plant only",
+      align = "ccccc", booktabs = TRUE, linesep = ""
+      ) %>%
+  footnote(
+           general = c(
+                       "Note: Standard errors are in parentheses and are clustered at the village level. The sample is restricted to village-years after a coal plant was built near the village.",
+                       "* p<0.10 ** p<0.05 *** p<0.01"
+                       ), 
+           general_title = "",
+           threeparttable = TRUE
+           ) %>%
+  column_spec(1, width = "3cm") %>%
+  column_spec(c(2:6),width = "2cm") %>%
+  row_spec(5, bold = TRUE) %>%
+  row_spec(9, hline_after = TRUE) %>%
+  add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
+  kable_styling(latex_options = "HOLD_position") %>%
+  kable_classic_2()
+
+```
+
+\FloatBarrier
+\newpage
 ```{r yieldtableleads, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
@@
 ```
+
+\FloatBarrier
+\newpage
+```{r laborplot, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Timing of household surveys for the National Sample Survey", fig.footer = list("Note: The histogram shows the distribution of survey dates for the four waves of the National Sample Survey (NSS) for the five waves used in this paper.")}
+
+laborplot
+
+```
+
+\FloatBarrier
+\newpage
+```{r labortablemonsoon, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+# Table
+colnames(labortablemonsoon) <- c("all", "all", "self", "wage", "farm", "non-farm")
+kable(
+      labortablemonsoon, caption = "Wind direction and labor allocation, monsoon season only",
+      align = "cccccc", booktabs = TRUE, linesep = ""
+      ) %>%
+  footnote(
+           general = c(
+                       "Note: Standard errors are in parentheses and are clustered at the district level. Control variables include female, age, age squared, and (years of) education.",
+                       "* p<0.10 ** p<0.05 *** p<0.01"
+                       ), 
+           general_title = "",
+           threeparttable = TRUE
+           ) %>%
+  column_spec(1, width = "3cm") %>%
+  column_spec(c(2:7),width = "1.5cm") %>%
+  row_spec(c(8), hline_after = TRUE) %>%
+  row_spec(c(4,7), bold = TRUE) %>%
+  kable_classic_2()
+
+```
+
+\FloatBarrier
+\newpage
+```{r labortablewinter, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+# Table
+colnames(labortablewinter) <- c("all", "all", "self", "wage", "farm", "non-farm")
+kable(
+      labortablewinter, caption = "Wind direction and labor allocation, winter season only",
+      align = "cccccc", booktabs = TRUE, linesep = ""
+      ) %>%
+  footnote(
+           general = c(
+                       "Note: Standard errors are in parentheses and are clustered at the district level. Control variables include female, age, age squared, and (years of) education.",
+                       "* p<0.10 ** p<0.05 *** p<0.01"
+                       ), 
+           general_title = "",
+           threeparttable = TRUE
+           ) %>%
+  column_spec(1, width = "3cm") %>%
+  column_spec(c(2:7),width = "1.5cm") %>%
+  row_spec(c(8), hline_after = TRUE) %>%
+  row_spec(c(4,7), bold = TRUE) %>%
+  kable_classic_2()
+
+```

### 2023-01-09 draft/draft.tex
updated draft

@@
 \author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
-\date{2023-01-03}
+\date{2023-01-09}
 
@@
 
-pollution could be bad for agriculture Marshall et al. (1997)
+pollution could be bad for agriculture (Heck et al. 1982; Marshall et al. 1997)
 
@@
 
-We have known about the effects of pollution on human health for many years Pope III and Dockery (2006)
+We have known about the effects of pollution on human health for many years (Brunekreef and Holgate 2002; Kampa and Castanas 2008; Pope III and Dockery 2006)
 good review on the effects of pollution early in life in Currie et al. (2014)
-- also in developing countries, where there is a consistent effect of exposure on infant mortality (Heft-Neal et al. 2018)
+- also in developing countries, where there is a consistent effect of exposure on infant mortality (Heft-Neal et al. 2018, 2019)
 
@@
 - call center workers in China, and at common levels of pollution (Chang et al. 2019)
-- tests (Ebenstein, Lavy, and Roth 2016)
+- tests (Ebenstein, Lavy, and Roth 2016; Wen and Burke 2022)
 - farm workers in California (Graff Zivin and Neidell 2012)
@@
 
-Asher et al. (2021)
-Ag productivity: Gangopadhyay et al. (2022)
-PM estimates: Hammer et al. (2020)
+The main goal of this paper is to examine whether exposure to exhaust from coal plants affects agricultural productivity, labor allocation, and economic growth.
+I use several sources of data, which I describe here and in Table \ref{tab:data} in the appendix.
 
-\begin{table}
-
-\caption{\label{tab:data}Remote sensing data sources}
-\centering
-\begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
-\toprule
-  & shapefile & wind & pollution & agriculture & nightlights\\
-\midrule
-source & Asher et al. (2021) & NCAR & Hammer et al. (2020) & Angopadhyay et al. (2022) & Asher et al. (2021)\\
-geographic coverage & India & global & global & global & India\\
-temporal coverage &  & daily & monthly & two seasons/year & yearly\\
-\bottomrule
-\end{tabular}
-\begin{tablenotes}
-\item \textit{ } 
-\item NCAR: https://climatedataguide.ucar.edu/
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
-\hypertarget{methodology}{%
-\subsection{Methodology}\label{methodology}}
+The first set of data lists the location of coal plant across the globe.
+This data comes from Global Energy Monitor.\footnote{\url{https://globalenergymonitor.org/projects/global-coal-plant-tracker/}} and lists all units generating at least 30 megawatts of electricity.
+The data on coal plants includes the year of opening (and, if applicable, the year of retirement), the GPS (latitude/longitude) location of each plant, and the amount of power produced by the units at each plant.
+For this paper, I do not use the information on the capacity of the plant.
 
@@
 
+Figure \ref{fig:plants} presents the location of coal plants in India for two specific years: 1990 and 2010.
+There is a clear increase in the prevalence of coal plants across the country over the two decades.
+Additionally, the overall capacity from coal plants increased from 42.4 gigawatts to 100.4 gigawats, an increase of 136.6 percent in just 20 years.
+
+The second dataset includes agricultural productivity for both the monsoon and winter seasons, from 2002 to 2018.
+To match other data, I only use the data up to 2013.
+This data comes from Gangopadhyay et al.
+(2022) and estimates land productivity (i.e.~yield, in tons per hectare) for the major crops in India.
+They define the monsoon season as June to October and the winter season as November to March.
+I keep these definitions when matching across data below.
+This data is also publicly available from Nature's data-sharing website.\footnote{\url{https://springernature.figshare.com}}
+
+The agricultural productivity data is available as raster files with a resolution of 500m.
+To aggregate this data up to a useful administrative unit, I use village-level shapefiles provided by Asher et al.
+(2021) and publicly available on the SHRUG platform.\footnote{\url{https://www.devdatalab.org/shrug_download/}} I aggregate the agricultural productivity data to the village level by extracting mean productivity for each feature in the shapefile.
+I do this separately for each season -- monsoon and winter -- and each year, from 2002 to 2013.
+
+To measure exposure to pollution from coal plants, I first locate all village centroids located within 30km from a coal plant in a given year.
+I choose 30km due to previous research on the effects of (air) pollution (Aragón and Rud 2016; Bencko and Symon 1977; Li and Gibson 2014).
+I then calculate the direction from coal plants to all village centroids within that 30km radius.
+To define exposure, I then pull daily wind direction data from the National Center for Atmospheric Research.\footnote{\url{https://climatedataguide.ucar.edu/}} For each day, I document whether the wind is blowing towards each village centroid.\footnote{I define ``towards'' as within five degrees to help take into account that I am using village centroids, using the x and y components of wind speed.} I then temporally aggregate this daily data depending on the temporal definition of the corresponding outcome.
+For example, agricultural productivity is defined across five months (e.g.~the monsoon season is from June to October) so I count the total days a given village is exposed to wind during those five months.
+
 \begin{figure}
-\includegraphics{draft_files/figure-latex/windexample-1} \caption[Wind direction and aggregation examples (2010-01-01)]{Wind direction and aggregation examples (2010-01-01)}\label{fig:windexample}\floatfoot*{Note: The top figure shows the average wind direction on January 1st, 2010.
+\includegraphics{draft_files/figure-latex/windexample-1} \caption[Wind direction and aggregation examples (2010-01-01)]{Wind direction and aggregation examples (2010-01-01)}\label{fig:windexample}\floatfoot*{Note: The left figure shows the average wind direction on January 1st, 2010.
 The points are the location of coal plants on that date.
-The bottom figure shows the distribution of pollution exposure in a specific district -- Kangra district in Himachal Pradesh -- on the same date.}
+The right figure shows the distribution of pollution exposure in a specific district -- Kangra district in Himachal Pradesh -- on the same date.}
 \end{figure}
 
-\newpage
+The top panel in Figure \ref{fig:windexample} is an example wind direction raster.
+The raster shows wind direction for the entirety of India on January 1st, 2010, as well as the location of coal power plants on that date.
+The prevailing winds on the date differ across the country and, though not shown in the figure, across days.
+This means that the overall exposure for a given area to coal plant emissions changes across time.
+I also pull data on particulate matter to help validate this proxy for exposure.
+This data comes from Hammer et al.
+(2020) and I also aggregate this to the village using the same method detailed above.
+
+To help understand some of the mechanisms driving the relationship between exposure to emissions and agricultural productivity, I also include household survey data.
+Unfortunately, I am not aware of any publicly available survey data that covers multiple years and has identifiers below the district level.
+For example, the ARIS-REDS data\footnote{\url{https://www.ncaer.org/project/additional-rural-incomes-survey-aris-rural-economic-demographic-survey-reds}} includes GPS identifiers for one year, but that only allows GPS matching for a total of two years: 1999 and 2007/2008.
+The DHS, on the other hand, only has district-level identifiers (except for the newest two waves) and is also not a labor survey.
+Instead, I use the National Sample Survey (NSS) from the Ministry of Statistics and Programme Implementation.
+
+The NSS is a nationally representative survey that was conducted every couple of years until relatively recently.
+Many rounds of the NSS also included a module on employment that asked about labor allocation over the previous seven days.
+I use the 61st, 62nd, 64th, 66th, and 68th rounds of the NSS, all of which include the employment module.
+The publicly available data also includes the date of survey, which allows me to match pollution exposure to the exact day of the survey.
+
+The biggest issue with using the NSS is that it includes only district-level identifiers.
+The main exposure variables, on the other hand, are defined at a lower level of aggregation, at the village.
+I choose to aggregate this village-level exposure to the district level as shown in the second panel of Figure \ref{fig:windexample}.
+I use the area overlap of the villages with the district to define a weighted average of exposure for the district at a given time.
+I calculate the area of overlap for each village and the district as well as the overall area of the district and use these variables to define weights for individual villages.
+Any part of the district that does not overlap with one of the villages -- the white area in Figure \ref{fig:windexample} -- does not have any exposure and, as such, receives a value of zero for exposure.
+
+Finally, I use nightlights as a measure of economic growth (Henderson, Storeygard, and Weil 2012).
+I also take this data directly from SHRUG and Asher et al.
+(2021).
+This covers 1992 through 2013.
+Some of the older data may not be ideal (Gibson et al.
+2021), but this is the only way to document changes in economic growth at a more disaggregated level and at a regular basis.
+Although satellite imagery is an attractive alternative (Burke et al.
+2021), high-resolution satellite imagery does not go back to the early 1990s.
+
+\hypertarget{identification}{%
+\subsection{Identification}\label{identification}}
+
+To calculate the effects of exposure to pollution on different outcomes, I rely on fixed effects and the differential exposure driven by plausibly exogenous wind direction, conditional on total rainfall.
+The regession is of the form
+\begin{gather} y_{it} = \alpha_{i} + \gamma_{t} + \beta wind_{it} + \delta rain_{it} + \varepsilon_{it}, \end{gather}
+where \(y_{it}\) is outcome \(y\) for unit \(i\) at time \(t\), \(\alpha\) is a geographic fixed effects, \(\gamma_t\) is time fixed effects, \(wind_{it}\) is exposure to pollution -- defined using wind direction -- and \(rain_{it}\) is rainfall.
+Although I use \(i\) and \(t\) as subscripts for the fixed effects, note that these change depending on the unit of analysis and the outcome.
+I explicitly state the level of fixed effects in the tables and results below.
+
+Identification relies on changes in wind direction across time for the same geographic units.
+In other words, conditional on the fixed effects and rainfall, I assume changes in wind direction are as good as random and uncorrelated with the outcome -- e.g.~agricultural productivity -- except through exposure to pollution.
+One major threat to identification is that the building of a coal plant could lead to a spurious correlation between the exposure variable and the outcome.
+Specifically, since exposure is defined as zero prior to the building of a coal plant, the exposure variable itself might be endogenous even though wind direction is not.
+To help allay these concerns, I estimate additional regressions restricting the sample to only time periods after which a given location has a coal plant.
+I also present results using leads to show that future wind direction does not seem to predict current outcomes.
+
+\hypertarget{an-example-validating-wind-direction}{%
+\subsection{An example: validating wind direction}\label{an-example-validating-wind-direction}}
+
+Before moving on to the main results, it is worth taking the time to better understand the fixed effects strategy with an example.
+Consider the data used for the regressions presented in Table \ref{tab:pollutiontable}.
+The outcome variable is particulate matter -- specifically, \(\mathrm{PM_{2.5}}\), which is particulate matter no larger than 2.5 micrometers in diameter -- which comes from Hammer et al.
+(2020).
+\(\mathrm{PM_{2.5}}\) is one of the harmful byproducts from coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.\footnote{\url{https://www.epa.gov/airmarkets/power-plants-and-neighboring-communities}} This data is available at the monthly level, so in Table \ref{tab:pollutiontable}, I aggregate wind exposure to the monthly level, as well.
 
@@
 
+Both the particulate matter data and the wind data is at the village level, which means the regressions in Table \ref{tab:pollutiontable} are at the village level.
+As such, the geographic fixed effects are village fixed effects and the temporal fixed effects are month fixed effects.
+Since the exposure variable is at the village level and we follow villages over time, the standard errors are also clustered at the village level.
+
+In addition to serving as an example, Table \ref{tab:pollutiontable} also helps validate the use of wind direction as a proxy for exposure to pollution.
+The first two columns include all available years of data while the last two columns restrict estimation only to the years for which the agricultural productivity regressions are estimated.
+Across all four columns, the story is the same: when the wind is blowing in the direction of a village from a coal plant within 30 kilometers, estimated \(\mathrm{PM_{2.5}}\) is substantially higher.
+The regressions indicate that one additional day of wind in the direction of the village increases mean \(\mathrm{PM_{2.5}}\) for the month by between 0.014 and 0.045 \(\mathrm{\mu g/m^3}\).
+
+To put these numbers into perspective, the World Health Organization's updated guidelines are that average annual exposure should not exceed 15 \(\mathrm{\mu g/m^3}\).
+For the month, an increase at the midpoint of the estimated range is equal to an increase of 0.2 percent of the maximum recommended mean concentration from the WHO.
+This is just a single day of wind in the direction of a village.
+Since this includes all villages located within 30km of a coal plant, this is evidence that wind direction can have serious repercussions on the health of hundreds of millions of people in India.
+
 \hypertarget{results}{%
 \section{Results}\label{results}}
+
+This section presents the main results of this paper.
+I present three sets of results: one set looking at agricultural productivity, another looking at labor allocation, and a final set looking at nightlights, which are a proxy for economic development.
+I go through each of these in turn.
+
+\hypertarget{agricultural-productivity}{%
+\subsection{Agricultural productivity}\label{agricultural-productivity}}
+
+I first present results for agricultural productivity.
+The outcome of all the regressions in this section is the log of agricultural land productivity, which is defined as tons per hectare.
+The first set of results are in Table \ref{tab:yieldtable}.
+The unit of analysis is the year season -- from 2002 to 2013 -- the geographic fixed effect is the village-season, and the temporal fixed effect is the year.
+The exposure variable is defined for the entire season.
+Since Gangopadhyay et al.
+(2022) define both the monsoon (kharif) season and the winter (rabi) season as five months long, I do as well.
+This implies that the exposure variable ranges from zero to slightly more than 150.
+
+The first column presents the most simple results, with only the wind exposure variable and the fixed effects.
+The second column adds a rainfall variable -- defined as the total rainfall in the season -- the addition of which is motivated by the fact that changes in wind direction could be correlated with changes in weather.
+However, the coefficient on the wind exposure variable is completely unchanged by the addition of rainfall, indicating that weather may not be as worrisome of a confounder as thought.
+
+Column three adds district-by-year-by-season fixed effects as an additional robustness checks.
+This addition decreases the magnitude of the exposure variable but it remains negative and significant.
+However, part of this may be due to the fact that the new fixed effects soak up much of the variation in the exposure variable, as shown in Figure \ref{fig:windexample}.
+The last two columns separate the effects into the monsoon season and the winter season, but the overall effect is roughly similar, though rainfall has a larger effect in the monsoon season than the winter season, which is unsurprising.
 
@@
 
+To put the size of the coefficient in context, it is worth digging a little deeper into the exposure variable.
+The mean within-village absolute deviation in wind exposure is approximately 8.06 days, meaning that the average change in agricultural productivity from year-to-year due to nothing but changes in wind patterns carrying particulate matter is around 2.4 percent relative to the mean.
+This implies that swings in agricultural productivity -- due to wind direction -- of more than five percent are probably quite common, since the absolute deviation includes deviations below and above the mean.
+
 \begin{table}
@@
 particulate matter & -0.021*** & -0.020*** & -0.033*** & -0.013*** & -0.024***\\
-(PM 2.5, '000s) & (0.001) & (0.002) & (0.005) & (0.001) & (0.003)\\
+(PM 2.5) & (0.001) & (0.002) & (0.005) & (0.001) & (0.003)\\
 rain (z) &  & 0.004** & 0.002 & 0.086*** & -0.016***\\
@@
 \midrule
-first stage: &  &  &  &  & \\
+\textbf{first stage:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
 wind & 0.143*** & 0.126*** & 0.022*** & 0.155*** & 0.105***\\
@@
 
+If you believe that wind direction is exogenous and, conditional on the fixed effects and rainfall, only affects agricultural productivity through pollution -- which seems reasonable since it is hard to imagine what other variables it might affect -- then it is a valid instrument for
+pollution.
+While I opt to present the reduced form results for the rest of this paper, it is nonetheless an interesting exercise to also use wind direction as an instrument for pollution, which I measure using estimated \(\mathrm{PM_{2.5}}\) from Hammer et al.
+(2020).
+
+The IV results appear in Table \ref{tab:yieldtabletwo}, with the second-stage results in the top panel and the first-stage results in the bottom panel.
+The first-stage results make clear how strong of an instrument wind is; also note that the coefficients are quite different from those in Table \ref{tab:pollutiontable} because the unit of analysis in Table \ref{tab:yieldtabletwo} is the season, while in Table \ref{tab:pollutiontable} the unit of analysis is the month.
+
+The second-stage results indicate that \(\mathrm{PM_{2.5}}\) is a very strong predictor of agricultural productivity.
+The smallest coefficient -- that for the monsoon season in column (4) -- and the mean absolute deviation for \(\mathrm{PM_{2.5}}\) point to a change in agricultural productivity of around 23 percent.
+
 \begin{table}
@@
 \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-6}
-  & (1) & (2) & (3) & (4) & (5\\
+  & (1) & (2) & (3) & (4) & (5)\\
 \midrule
@@
 \end{table}
+
+Having established that exposure to pollution has a negative effect on agricultural productivity, I now move to heterogeneity in these effects.
+Table \ref{tab:yieldtablehet} presents four different analyses: effects of exposure by maximum exposure, multiplicative effects of shocks, the possibility of non-linearities, and the differential effects of exposure based on initial agricultural productivity.
+
+The first two columns split the sample based on average exposure; villages with maximum exposure above the median are included in column (1) while those below the median are in column (2).
+Interestingly, the largest effects are seen on villages in the lower half of the distribution.
+Apparently, smaller maximum effects lead to larger overall effects of pollution exposure on agricultural productivity.
+In other words, it is not the places where wind is most likely to blow from a coal plant that experience the largest decrease in productivity in response to an additional day of wind-driven pollution.
+
+Column three looks add the possible multiplicative effects of rainfall shocks and pollution shocks on agricultural productivity.
+In previous tables, rainfall always entered the regression positively -- as we would expect -- indicating that higher rainfall in a given season leads to higher agricultural productivity.
+However, an interesting question is whether rainfall and pollution shocks might interact.
+The positive coefficient on the interaction term between days of wind and rainfall during the agricultural season indicates that there is a multiplicative effect of shocks; when wind days are higher (bad) and rainfall is lower (bad), the overall negative effects are even worse than when just one shock happens.
+Considering that climate change is leading to more variable rainfall and more negative rainfall shocks, these results are potentially worrisome, especially for smallholders.
+
+Column four adds a quadratic in wind to see if the negative effects of pollution on agricultural productivity are increase or decreasing in wind.
+In column four, the coefficient on the linear term is 0.0004 while the coefficient on the quadratic term is -0.0002, meaning that the effect of wind is decreasing for any days of wind over two.
+The negative coefficient also indicates that this negative effect is consistently decreasing, meaning the effect gets worse as average wind days gets higher.
+Note that this is somewhat different from the regressions in the first two columns, which split the sample by the \emph{maximum} days of wind observed in the sample, not the mean.
+
+In the final column I add to the regression an interaction between wind and initial agricultural productivity (defined in 2002, the first year observed in the productivity data).
+The linear term drops out of the regression since it is invariant within villages, leaving just the interaction term.
+The negative coefficient on the interaction term of -0.001 indicates that the effect of wind is more negative in areas where agricultural productivity is higher.
+Note that agricultural productivity is defined in logs, so this can be interpreted as a percentage effect, not a level effect.
+Apparently, wind-driven pollution is particularly detrimental to agricultural productivity in the most productive places.
+
+\hypertarget{robustness-checks}{%
+\subsection{Robustness checks}\label{robustness-checks}}
+
+I present two separate robustness checks in the appendix.
+The first set of analyses attempts to separate the possible endogenous location of coal plants from outcomes.
+Specifically, prior to the opening of a new coal plant in an area, all villages receive a zero for wind.
+This means that, even if wind direction is exogenous, the opening of the coal plant could lead to correlations between the error term and the wind variable.
+
+Although the results with district-year-season help mitigate this concern, Table \ref{tab:yieldtablepostplant} in the appendix checks this possibility in a different way.
+Specifically, I restrict the sample to only village-years \emph{after} the introduction of a coal plant.
+This drops any pre-plant-opening observations -- all of which are mechanically zeros for wind -- to prevent endogeneity due to the opening.
+The coefficients in Table \ref{tab:yieldtablepostplant} show that the overall effects are essentially unchanged, with just a small change in coefficient magnitude (driven mostly by rounding).
+
+The second robustness check includes leads for wind to see if they predict current year productivity.
+I present these results in \ref{tab:yieldtableleads}.
+The second column includes district-year-season fixed effects, while the first does not.
+The results make clear the importance of including trends or district-year fixed effects.
+In the first column, leads seem to predict productivity.
+However, this relationship disappears in the second column, indicating that the previous results with district-year fixed effects are likely the most reliable.
+
+\hypertarget{labor-allocation}{%
+\subsection{Labor allocation}\label{labor-allocation}}
+
+The previous set of results show a consistent negative effect of wind-driven pollution on agricultural productivity.
+A key question is the mechanism through which this happens; is this driven by changes in labor allocation/productivity or by changes in land productivity?
+While I cannot definitely prove the mechanism, I can shed some light on possibilities by looking at labor allocation decisions.
+Since these shocks generally happen after initial planting decisions, it seems unlikely that changes in land \emph{allocation} could drive the results.
+
+Table \ref{tab:labortable} presents results for labor allocation over the last seven days using the National Sample Survey and the methodology described above.
+Since the labor recall is for seven days, labor allocation ranges from zero to seven, as does the wind variable.
+Recall that the data is at the district level, not the village level.
+However, it is also more temporally disaggregated, with rolling interviews throughout the year-long survey.
+Figure \ref{fig:laborplot} in the appendix shows this temporal distribution across the five waves of the survey.
+
+There are some interesting results in Table \ref{tab:labortable}.
+The first two columns show a marginally significant decrease in overall labor allocation (where ``labor'' is defined as non-domestic self and wage employment).
+The third and fourth columns show that this overall decrease is predominantly from wage employment, not self employment.
+An additional day of wind over the last seven decreases wage employment by 0.034 days in the district.
+Mean wage employment is 1.458 days, meaning one day of wind decreases wage employment by 2.3 percent of the mean.
+Since much of the district is not covered by ``treated'' villages, this is actually a rather substantive effect on the villages involved.
+
+The last two columns split the effects by sector, into farm and non-farm employment.
+Although agricultural (land) productivity decreases when wind blows in the direction of villages, overall labor allocation to farm labor actually goes \emph{up}, by around 4.5 percent of the mean of 0.784 days.
+Non-farm labor, on the other hand, goes down, by around 2.4 percent of the mean of 2.527 days.
+
+There are two possible explanations for this pattern of results.
+First, it could be that land productivity goes down, leading farmers to increase their farm labor allocation in an attempt to compensate for the decreased productivity.
+A second possibility is that pollution does not affect land productivity at all, at least not directly.
+In this scenario, the worsening air quality could lead to decreased labor productivity, meaning that ``effective labor'' essentially decreases when pollution is higher.
+Unfortunately, I am not able to disentangle these two possiblities with the data available.
+
+Table \ref{tab:labortablemonsoon} and Table \ref{tab:labortablewinter} in the appendix disaggregate these effects into the monsoon and winter seasons.
+Apparently, the entirety of the effect is driven by changes in labor allocation during the monsoon season.
+The dynamics across seasons apparently lead to large differences in the effects of pollution on labor allocation.
 
@@
 \end{table}
+
+\hypertarget{nightlights}{%
+\subsection{Nightlights}\label{nightlights}}
+
+Pollution leads to lower agricultural productivity as well as changes in labor allocation.
+This leads to an obvious question: might higher levels of pollution also directly affect economic growth?
+To explore this question, I look at changes in nightlights relative to changes in pollution.
+It seems reasonable to think that pollution might affect nightlights with a lag, so I include the lag of pollution rather than contempraneous pollution.
 
@@
 
+I present these results in Table \ref{tab:ntltable}.
+To help simplify the coefficients, I recode wind to be the proportion of days in the year that a given village experiences wind in their direction, from a coal plant.
+Mean wind in this new specification is 0.066, indicating that a village in the sample experiences wind in its direction on around 6.6 percent of days, or 24 days of the year.
+
+We see a strong, negative relationship between pollution (lagged wind) and nightlights.
+Going from the 25th percentile to the 75th percentile of wind exposure is synonymous with a decrease in the growth of nightlights within a village by around 0.7 percent.
+While this is a relatively small magnitude, it is worth noting that there are more than 100,000 villages in this sample.
+While the effect for a single village might be small, the overall, aggregate effect on growth is quite substantial.
+
 \hypertarget{conclusion}{%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-bencko1977health}{}}%
+Bencko, Vladimir, and Karel Symon. 1977. {``Health Aspects of Burning Coal with a High Arsenic Content.''} \emph{{Environmental Research}} 13 (3): 378--85.
+
 \leavevmode\vadjust pre{\hypertarget{ref-brunekreef2002air}{}}%
 Brunekreef, Bert, and Stephen T Holgate. 2002. {``{Air pollution and health}.''} \emph{{The Lancet}} 360 (9341): 1233--42.
+
+\leavevmode\vadjust pre{\hypertarget{ref-burke2021using}{}}%
+Burke, Marshall, Anne Driscoll, David B Lobell, and Stefano Ermon. 2021. {``{Using satellite imagery to understand and promote sustainable development}.''} \emph{Science} 371 (6535): eabe8628.
 
@@
 {``{A new two-decade (2001--2019) high-resolution agricultural primary productivity dataset for India}.''} \emph{{Scientific Data}} 9 (1): 1--12.
+
+\leavevmode\vadjust pre{\hypertarget{ref-gibson2021night}{}}%
+Gibson, John, Susan Olivia, Geua Boe-Gibson, and Chao Li. 2021. {``{Which night lights data should we use in economics, and where?}''} \emph{{Journal of Development Economics}} 149: 102602.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-heft2019air}{}}%
+Heft-Neal, Sam, Jennifer Burney, Eran Bendavid, Kara Voss, and Marshall Burke. 2019. {``{Air pollution and infant mortality: Evidence from Saharan dust}.''} {National Bureau of Economic Research}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-henderson2012measuring}{}}%
+Henderson, J Vernon, Adam Storeygard, and David N Weil. 2012. {``Measuring Economic Growth from Outer Space.''} \emph{{American Economic Review}} 102 (2): 994--1028.
+
 \leavevmode\vadjust pre{\hypertarget{ref-kampa2008human}{}}%
 Kampa, Marilena, and Elias Castanas. 2008. {``{Human health effects of air pollution}.''} \emph{{Environmental Pollution}} 151 (2): 362--67.
+
+\leavevmode\vadjust pre{\hypertarget{ref-li2014health}{}}%
+Li, Ya-Ru, and Jacqueline MacDonald Gibson.
+2014.
+{``{Health and air quality benefits of policies to reduce coal-fired power plant emissions: a case study in North Carolina}.''} \emph{{Environmental Science \& Technology}} 48 (17): 10019--27.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-wen2022lower}{}}%
+Wen, Jeff, and Marshall Burke. 2022. {``{Lower test scores from wildfire smoke exposure}.''} \emph{{Nature Sustainability}} 5 (11): 947--55.
+
 \end{CSLReferences}
@@
 \FloatBarrier
-
-\beginappendix
-
 \newpage
@@
 
+\setcounter{table}{0} \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}} 
+\FloatBarrier
+
+\begin{table}[H]
+
+\caption{\label{tab:data}Remote sensing data sources}
+\centering
+\begin{threeparttable}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2cm}>{\centering\arraybackslash}p{4.5cm}>{\centering\arraybackslash}p{3.5cm}>{\centering\arraybackslash}p{3.5cm}}
+\toprule
+  & source & geographic coverage & temporal coverage\\
+\midrule
+shapefile & Asher et al. (2021) & India & \\
+coal plants & Global Energy Monitor & global & yearly\\
+wind & NCAR & global & daily\\
+pollution & Hammer et al. (2020) & global & monthly\\
+agriculture & Angopadhyay et al. (2022) & global & two seasons/year\\
+nightlights & Asher et al. (2021) & India & yearly\\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}
+\item Note: NCAR stands for the National Center for Atmospheric Research.
+Additional information on Global Energy Monitor and NCAR can be found on their websites, at https://globalenergymonitor.org/projects/global-coal-plant-tracker and https://climatedataguide.ucar.edu/, respectively.
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+
+\newpage
+\begin{table}[H]
+
+\caption{\label{tab:yieldtablepostplant}Wind direction and agricultural productivity, post-coal plant only}
+\centering
+\begin{threeparttable}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\toprule
+\multicolumn{1}{c}{ } & \multicolumn{3}{c}{all} & \multicolumn{1}{c}{monsoon} & \multicolumn{1}{c}{winter} \\
+\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6}
+  & (1) & (2) & (3) & (4) & (5)\\
+\midrule
+wind & -0.003*** & -0.002*** & -0.0007*** & -0.002*** & -0.002***\\
+ & (0.0002) & (0.0002) & (9.7e-5) & (0.0002) & (0.0003)\\
+rain (z) &  & 0.031*** & 0.010*** & 0.103*** & 0.016***\\
+ &  & (0.0005) & (0.001) & (0.002) & (0.0005)\\
+\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+village-season & Yes & Yes & Yes & No & No\\
+year & Yes & Yes & No & Yes & Yes\\
+district-year-season & No & No & Yes & No & No\\
+village & No & No & No & Yes & Yes\\
+\midrule
+observations & 2,062,036 & 2,046,104 & 2,046,104 & 1,089,013 & 957,091\\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}
+\item Note: Standard errors are in parentheses and are clustered at the village level. The sample is restricted to village-years after a coal plant was built near the village.
+\item * p<0.10 ** p<0.05 *** p<0.01
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+
+\FloatBarrier
+\newpage
 \begin{table}
@@
 
+\FloatBarrier
+\newpage
+\begin{figure}
+\includegraphics{draft_files/figure-latex/laborplot-1} \caption[Timing of household surveys for the National Sample Survey]{Timing of household surveys for the National Sample Survey}\label{fig:laborplot}\floatfoot*{Note: The histogram shows the distribution of survey dates for the four waves of the National Sample Survey (NSS) for the five waves used in this paper.}
+\end{figure}
+
+\FloatBarrier
+\newpage
+\begin{table}
+
+\caption{\label{tab:labortablemonsoon}Wind direction and labor allocation, monsoon season only}
+\centering
+\begin{threeparttable}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}}
+\toprule
+  & all & all & self & wage & farm & non-farm\\
+\midrule
+wind & -0.021 & -0.033 & 0.038 & -0.071** & 0.072** & -0.105**\\
+ & (0.028) & (0.029) & (0.034) & (0.033) & (0.036) & (0.046)\\
+controls & No & Yes & Yes & Yes & Yes & Yes\\
+\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+district & Yes & Yes & Yes & Yes & Yes & Yes\\
+year & Yes & Yes & Yes & Yes & Yes & Yes\\
+\textbf{varying slopes:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+year (by district) & Yes & Yes & Yes & Yes & Yes & Yes\\
+\midrule
+observations & 359,652 & 359,576 & 359,576 & 359,576 & 359,576 & 359,576\\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}
+\item Note: Standard errors are in parentheses and are clustered at the district level. Control variables include female, age, age squared, and (years of) education.
+\item * p<0.10 ** p<0.05 *** p<0.01
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+
+\FloatBarrier
+\newpage
+\begin{table}
+
+\caption{\label{tab:labortablewinter}Wind direction and labor allocation, winter season only}
+\centering
+\begin{threeparttable}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}}
+\toprule
+  & all & all & self & wage & farm & non-farm\\
+\midrule
+wind & 0.003 & 0.002 & 0.011 & -0.009 & 0.024 & -0.022\\
+ & (0.029) & (0.026) & (0.028) & (0.025) & (0.034) & (0.039)\\
+controls & No & Yes & Yes & Yes & Yes & Yes\\
+\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+district & Yes & Yes & Yes & Yes & Yes & Yes\\
+year & Yes & Yes & Yes & Yes & Yes & Yes\\
+\textbf{varying slopes:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+year (by district) & Yes & Yes & Yes & Yes & Yes & Yes\\
+\midrule
+observations & 375,956 & 375,883 & 375,883 & 375,883 & 375,883 & 375,883\\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}
+\item Note: Standard errors are in parentheses and are clustered at the district level. Control variables include female, age, age squared, and (years of) education.
+\item * p<0.10 ** p<0.05 *** p<0.01
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+
 \end{document}

### 2023-01-09 analyses/regressions.R

@@ -43,9 +43,11 @@ villages_overlap <- villages_overlap %>% mutate(shrid = paste0(pc11_s_id, "-", p
                                           ungroup()
 pollution <- pollution %>% left_join(villages_overlap, by = "shrid")
 
-
-
-
+pollution <- pollution %>% group_by(shrid) %>%
+                              mutate(pm_abs_dev = abs(mean - mean(mean))) %>%
+                              ungroup()
+summary(pollution$pm_abs_dev)
+# mean 21.1
 
 pol1 <- feols(mean ~ wind | village + month, data = pollution, cluster = c("village"))
 pol2 <- feols(mean ~ wind | village + month^distfe, data = pollution, cluster = c("village"))
@@ -114,6 +116,11 @@ villages_overlap <- villages_overlap %>% mutate(shrid = paste0(pc11_s_id, "-", p
 df <- df %>% left_join(villages_overlap, by = "shrid")
 
 
+df <- df %>% group_by(shrid) %>%
+                mutate(wind_abs_dev = abs(wind*10 - mean(wind*10))) %>%
+                ungroup()
+summary(df$wind_abs_dev)
+# mean is 8.056, max is 131.58
 
 summary(df$wind*10)
 # Min. 1st Qu.  Median    Mean  3rd Qu.   Max. 
@@ -183,11 +190,10 @@ colnames(new_row) <- colnames(yieldtabletwo)
 
 
 yieldtabletwo_merge <- rbind(yieldtabletwo[-c(10:11),], new_row, yieldtabletwo1[1:4,])
-yieldtabletwo_merge <- yieldtabletwo_merge[-10,]
 # Turn to matrix so that duplicate row names are allowed
 yieldtabletwo_merge <- as.matrix(yieldtabletwo_merge)
 
-rownames(yieldtabletwo_merge) <- c("particulate matter", "(PM 2.5, '000s)", "rain (z)", "",
+rownames(yieldtabletwo_merge) <- c("particulate matter", "(PM 2.5)", "rain (z)", "",
                                    "fixed effects:", "village-season", "year", "district-year-season", "village", 
                                    "observations",
                                    "first stage:", "wind", " ", "rain (z)", " ")
@@ -296,6 +302,63 @@ saveRDS(yieldtableleads, "pollution_development/draft/tables/yieldtableleads.rds
 
 
 
+
+
+
+# Create a dataset that is only POST plant building
+df_new <- df %>% group_by(shrid) %>% 
[…]

### 2023-01-13 cleaning/cleaning_nss.R
nothing big

@@ -64,6 +64,7 @@ df_labor$Total_no_of_days_in_current_acti[is.na(df_labor$Total_no_of_days_in_cur
 df_labor$days_self <- 0
 df_labor$days_wage <- 0
 df_labor$days_domestic <- 0
+df_labor$days_sick <- 0
 # Now labor types
 df_labor$days_self[df_labor$Current_day_activity_Status %in% c("11", "12", "21")] <- df_labor$days_self[df_labor$Current_day_activity_Status %in% c("11", "12", "21")] + 
                                                                                       df_labor$Total_no_of_days_in_current_acti[df_labor$Current_day_activity_Status %in% c("11", "12", "21")]

### 2023-01-13 draft/draft.Rmd
first draft!

@@
 abstract: |
-  | \noindent abstract
+  | \noindent The rapid increase in population in India from the 1990s through the 2010s led to the rapid construction of new coal power plants throughout the country.
+Using this roll-out combined with randdom changes in wind direction, I document negative effects of pollution from coal power plants on agricultural productivity throughout the country.
+Results suggest this decrease may be driven by a decline in labor allocation of the oldest half of my sample, though I cannot rule out direct effects of pollution on land productivity.
+These impacts spill over into aggregate economic growth, as well; nightlight growth is slower following years of higher exposure to pollution.
   |
-  | \textbf{\textit{Keywords}}: 
-  | \textbf{\textit{JEL Codes}}:
+  | \textbf{\textit{Keywords}}: pollution, productivity, agriculture, labor, India
+  | \textbf{\textit{JEL Codes}}: H40, I15, J22, O13, Q52, Q53
 
@@
 labortablewinter <- readRDS("pollution_development/draft/tables/labortablewinter.rds")
+labortableold <- readRDS("pollution_development/draft/tables/labortableold.rds")
 ntltable <- readRDS("pollution_development/draft/tables/ntltable.rds")
@@
 
-pollution could be bad for agriculture [@heck1982assessment;@marshall1997hidden]
+From 1990 to 2010, India's population grew by more than 40 percent.^[data.worldbank.org] This population growth led to a large increase in demand for power.
+While natural gas has become a popular alternative in much of the world, the Indian government met much of the increased demand through the construction of coal power plants.
+During those same two decades, power generation from all coal units of at least 30 megawatts increased by almost 140 percent, from 42.4 gigawatts to more than 100 gigawatts.^[https://globalenergymonitor.org/] There are no doubt positive effects from this increase in power generation; for example, electricity may lead to economic and productivity growth [@dinkelman2011effects;@kline2014local;@rud2012electricity;@van2017long], though electrification alone may not be a sufficient condition [@lee2020does].
 
-lower agricultural productivity near gold mines in Ghana [@aragon2016polluting]; apparently lots of pollution
-  - "The main identification assumption is that the change in agricultural productivity over time in both areas would be similar in the absence of mining."
-  - within 20km
-  - also finds suggestive evidence it's not all labor productivity changes
+These improvements do not come without downsides, however, especially when it comes to power derived from coal.
+Pollutants from coal contribute to climate change and the phase out of coal energy is therefore a major climate goal [@iea2022].
+Emissions from the production and burning of coal are also directly harmful to health and the environment, including pollutants like sulfur dioxide, nitrogen oxides, and mercury, to name but a few.^[https://www.eia.gov/energyexplained/coal/coal-and-the-environment.php] While the use of coal had been declining, recent events -- including Russia's invasion of Ukraine -- have led to an increase in the demand for coal.^[https://www.npr.org/2022/08/15/1117560560/a-rising-demand-for-coal-amidst-war-in-ukraine]
 
-"Much of what we know about the marginal effect of pollution on infant mortality is derived from developed country data.
-However, given the lower levels of air pollution in developed countries, these estimates may not be externally valid to the developing country context if there is a non-linear dose relationship between pollution and mortality or if the costs of avoidance behaviour differ considerably between the two contexts.
-In this article, we estimate the relationship between pollution and infant mortality using data from Mexico.
-Our estimates for PM10 tend to be similar (or even smaller) than the US estimates, while our findings on CO tend to be larger than those derived from the US context." [@arceo2016does]
+In this paper, I study the effects of emissions from coal power plants on agricultural productivity, labor allocation, and economic growth in India.
+Previous work has suggested that there may be harmful direct effects of air pollution [@heck1982assessment;@marshall1997hidden] and water pollution [@reddy2006impact] on agricultural productivity, including work showing that gold mines -- which release large amounts of pollutants into the surrounding environment -- lead to lower agricultural productivity in Ghana [@aragon2016polluting].
+This paper, however, tackles the broader case of coal plants, which are a leading source of power throughout the world.
+I show that pollution from these plants can lead to decreased agricultural productivity in the case of India.
 
-We have known about the effects of pollution on human health for many years [@brunekreef2002air;@kampa2008human;@pope2006health]
-good review on the effects of pollution early in life in @currie2014we
-  - also in developing countries, where there is a consistent effect of exposure on infant mortality [@heft2018robust;@heft2019air]
+I combine the roll out of coal plants in India with plausibly exogenous changes in wind direction -- in a similar spirit to [@deryugina2019mortality] -- to measure the effects of pollution from coal plants on key development outcomes.
+I first locate the location of coal plants in India by year.
+Using village-level shapefiles, I identify all villages that are located within 30km of a coal plant at any timn during the sample period, which comprises the 1990s to the 2010s, with differences depending on data availability.
+I then download daily wind data and create an identifier for whether a village is located downwind from a coal plant on any given day, aggregating these daily values to different levels depending on the analysis.
 
-"Air pollution has both acute and chronic effects on human health, affecting a number of different systems and organs.
-It ranges from minor upper respiratory irritation to chronic respiratory and heart disease, lung cancer, acute respiratory infections in children and chronic bronchitis in adults, aggravating pre-existing heart and lung disease, or asthmatic attacks.
-In addition, short- and long-term exposures have also been linked with premature mortality and reduced life expectancy.
-These effects of air pollutants on human health and their mechanism of action are briefly discussed." @kampa2008human
+After validating that wind direction is a strong predictor of particulate matter, I first document that the total number of days of being downwind from a coal power plant has a negative effect on overall agricultural productivity, using newly available estimates of India-wide agricultural (land) productivity.
+The within-village absolute deviation of wind exposure is around eight days, meaning that median changes in wind direction are responsible for fluctuations in agricultural productivity of around 2.4 percent, with changes of more than five percent from season to season being common.
+Using wind direction as an instrument for particulate matter, an increase of PM 2.5 by one microgram leads to an even larger decrease in agricultural productivity, at around 20 percent.
+This IV strategy is similar to that in @deryugina2019mortality, though they use general "high-pollution" wind directions while I concentrate only on wind directions from the location of a coal power plant.
 
-worker productivity and cognitive function:
-  - call center workers in China, and at common levels of pollution [@chang2019effect]
-  - tests [@ebenstein2016long;@wen2022lower]
-  - farm workers in California [@graff2012impact]
-  - increase in extensive margin [@hanna2015effect]
-  - manufacturing in China [@he2019severe]; "We uncover statistically significant adverse output effects from more prolonged exposure, but effects are not large.
-A substantial +10 mug/m3 PM2.5 variation sustained over 25 days reduces daily output by 1 percent."
+These effects are apparently concentrated in villages that do not experience the most extreme exposure to pollution; villages with higher maximum exposure experience effects only around 10 percent as large as villages with lower maximum exposure.
+Additional heterogeneity analysis shows that, despite the results with respect to the maximum, the effects are increasing in *mean* exposure.
+I also show that the effects are largest in villages with the highest initial agricultural productivity and that pollution and rainfall shocks are compounding; higher exposure along with less rainfall increases the negative effect of either shock individually.
 
-induces migration [@chen2022effect]
+Robustness checks show that this effect is concentrated in the years after the construction of the power plant, which helps decrease concerns around endogeneity since, conditional on the village fixed effect and rainfall, windfall direction across years and seasons is exogenous.
+Additionally, leads for wind direction do not predict current agricultural productivity, again supporting the identifying assumptions.
 
-overall effects on agricultural productivity taking into accounts any changes in input allocation caused by the increase in pollution.
+While I am not able to explicitly test effects on land productivity, I look at how adults allocate labor over the last seven days using the National Sample Survey and creating a district-level exposure measure based on the timing of the survey for each household.
+An additional day of wind over the previous week decreases overall labor allocation by 0.027 days.
+However, all effects on labor allocation are driven by the older half of my sample; these adults see much large decreases in overall labor allocation, self employment, farm employment, and non-farm employment.
+This is suggestive evidence of productivity effects being driven by the health effects on individuals, though I am not able to rule out direct effects on land productivity with my data.
 
-# Data and methods
+The final set of results explores whether there are impacts of pollution on aggregate economic development.
+Using nightlights, I show that lagged wind -- i.e.
+pollution exposure in the prior year -- has a negative impact on economic development as proxied by nightlights.
+Going from the 25th percentile to the 75th percentile of wind exposure is synonymous with a decrease in the growth of nightlights within a village by around 0.7 percent.
+While this is a relatively small magnitude, it is worth noting that there are more than 100,000 villages in this sample.
+While the effect for a single village might be small, the overall, aggregate effect on growth is quite substantial.
 
-## Data
+This paper contributes to several strands of literature.
+First, we already have evidence of effects of pollution on different forms of labor activity.
+For example, general levels of pollution decrease productivity of call center works in China [@chang2019effect], lead people to perform worse on cognitive functions [@ebenstein2016long;@wen2022lower], and also lead to lower levels of farm labor productivity in California, driven by changes in ozone concentration [@graff2012impact].
+Also in China, @he2019severe find small but significant negative effects of pollution on manufacturing productivity, while @chen2022effect show pollution leads to labor out migration.
+Perhaps relevant to my labor allocation results, @hanna2015effect show that decreases in pollution lead to increases in labor allocation, while I show the same relationship among the oldest cohort.
+The paper most related to this one is @aragon2016polluting, who show that gold mines in Ghana have substantial negative impacts on agricultural productivity in the surrounding areas.
+This paper differs in two important ways.
+First, I explore effects of pollution from coal plants, which are much more ubiquitous than goald mines across the world.
+Second, I use a different identification strategy, relying on wind blowing pollution from coal plants.
+
+I also contribute to the more general literature on the effects of pollution on health.
+Pollution has substantial negative effects on health in both developing and developed countries [@arceo2016does] and we have known of these negative effects for many years [@brunekreef2002air;@kampa2008human;@pope2006health].
+@currie2014we review the literature on the health effects of pollution during adolescence; we have robust evidence of pollution exposure leading to increases in infant mortality, including in developing countries [@heft2018robust;@heft2019air].
+Since these effects are often driven by heart and lung disease and asthma [@kampa2008human], this my explain the labor allocation effects I identify in this paper.
+
+The rest of the paper is organized as follows.
+I first discuss data and methodology.
+I break this section up into \autoref{data2}, which covers the multiple sources of data used in this paper; \autoref{identification}, which discusses the empirical strategy; and \autoref{validation}, which validates the wind direction variable using data on particulate matter.
+I present the main results in \autoref{results} before concluding in \autoref{conclusion}.
+
+# Data and methods \label{data}
+
+## Data \label{data2}
 
@@
 
-## Identification
+## Identification \label{identification}
 
@@
 
-## An example: validating wind direction
+## An example: validating wind direction \label{validation}
 
@@
 
-# Results
+# Results \label{results}
 
@@
 
+Previous research has shown larger effects of pollution on the health of the oldest [@deryugina2019mortality].
+The labor sample includes all adults between 15 and 60 years of age.
+To look at heterogeneity by age, I present an additional set of results interacting the wind variable with an indicator for whether an individual is over 37 years of age.^[This splits the sample into two equal bins of age: those who are 15-37 and those who are 38-60.] These results are in Table \@ref(tab:labortableold), which otherwise presents the same specifications in the last five columns as \@ref(tab:labortable).
+
+```{r labortableold, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+# Table
+labortableold <- labortableold[,-1]
+colnames(labortableold) <- c("all", "self", "wage", "farm", "non-farm")
+kable(
+      labortableold, caption = "Wind direction and labor allocation, heterogeneity by age",
+      align = "ccccc", booktabs = TRUE, linesep = ""
+      ) %>%
+  footnote(
+           general = c(
+                       "Note: Standard errors are in parentheses and are clustered at the district level. Control variables include female, age, age squared, and (years of) education.",
+                       "* p<0.10 ** p<0.05 *** p<0.01"
+                       ), 
+           general_title = "",
+           threeparttable = TRUE
+           ) %>%
+  column_spec(1, width = "4cm") %>%
+  column_spec(c(2:6),width = "1.5cm") %>%
+  row_spec(c(10), hline_after = TRUE) %>%
+  row_spec(c(6,9), bold = TRUE) %>%
+  kable_classic_2()
+
+```
+
+Table \@ref(tab:labortableold) presents a very clear set of results: the effects are always larger (more negative) for older adults.
+Particularly interestingly, we now see that the mean-zero effect on self employment in the previous table is driven by two stark contrasts: a large increase for the younger group and a large decrease for the older group.
+For the youngest group, this increase appears in farm labor, indicating likely increases in farm self-employment.
+The oldest, on the other hand, have large decreases in both farm and non-farm labor, with larger decreases in the latter.
+
+While speculative, one possible explanation for this pattern is a decrease in labor productivity.
+While the youngest may be able to readjust and increase their overall self-employment labor allocation, perhaps the oldest cannot due to, for example, health effects of pollution, which are well documented [@bentayeb2012adverse;@pope2000epidemiology].^[Much of the literature also suggests the youngest are at significant risk of mortality and morbidities.
+However, given my focus on labor allocation, I only discuss older adults here.]
+
 ## Nightlights
@@
 To explore this question, I look at changes in nightlights relative to changes in pollution.
-It seems reasonable to think that pollution might affect nightlights with a lag, so I include the lag of pollution rather than contempraneous pollution.
+It seems reasonable to think that pollution might affect nightlights with a lag, so I include the lag of pollution rather than contemporaneous pollution.
 
@@
 
-# Conclusion
+# Conclusion \label{conclusion}
+
+This paper documents the negative effects of pollution on agricultural productivity in India.
+I find clear indications that pollution from coal plants leads to lower agricultural productivity.
+Importantly, I use exogenous changes in wind direction to measure exposure to pollution and show that these effects are robust to a range of alternative specifications.
+Results indicate that these effects may be driven by changes in labor allocation patterns, but I am unable to rule out direct effects on land productivity with the data available.
+Finally, these negative effects are also seen in aggregate economic development, with pollution leading to decreases in the growth of nightlights.
+
+These results suggest caution in building new power plants that emit substantial amounts of pollution.
+While the building of additional power generation facilities may be an economic necessity, the location of these facilities may be particular important for multiple aspects of economic activity, not just health and mortality.
 

### 2023-01-13 draft/draft.tex
first draft!

@@
 \author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
-\date{2023-01-09}
+\date{2023-01-13}
 
@@
 \begin{abstract}
-\noindent abstract\\
+\noindent The rapid increase in population in India from the 1990s through the 2010s led to the rapid construction of new coal power plants throughout the country.
+Using this roll-out combined with randdom changes in wind direction, I document negative effects of pollution from coal power plants on agricultural productivity throughout the country.
+Results suggest this decrease may be driven by a decline in labor allocation of the oldest half of my sample, though I cannot rule out direct effects of pollution on land productivity.
+These impacts spill over into aggregate economic growth, as well; nightlight growth is slower following years of higher exposure to pollution.\\
 \strut \\
-\textbf{\textit{Keywords}}:\\
-\textbf{\textit{JEL Codes}}:
+\textbf{\textit{Keywords}}: pollution, productivity, agriculture, labor, India\\
+\textbf{\textit{JEL Codes}}: H40, I15, J22, O13, Q52, Q53
 \end{abstract}
@@
 
-pollution could be bad for agriculture (Heck et al. 1982; Marshall et al. 1997)
+From 1990 to 2010, India's population grew by more than 40 percent.\footnote{data.worldbank.org} This population growth led to a large increase in demand for power.
+While natural gas has become a popular alternative in much of the world, the Indian government met much of the increased demand through the construction of coal power plants.
+During those same two decades, power generation from all coal units of at least 30 megawatts increased by almost 140 percent, from 42.4 gigawatts to more than 100 gigawatts.\footnote{\url{https://globalenergymonitor.org/}} There are no doubt positive effects from this increase in power generation; for example, electricity may lead to economic and productivity growth (Dinkelman 2011; Kline and Moretti 2014; Rud 2012; Van de Walle et al.
+2017), though electrification alone may not be a sufficient condition (Lee, Miguel, and Wolfram 2020).
 
-lower agricultural productivity near gold mines in Ghana (Aragón and Rud 2016); apparently lots of pollution
-- ``The main identification assumption is that the change in agricultural productivity over time in both areas would be similar in the absence of mining.''
-- within 20km
-- also finds suggestive evidence it's not all labor productivity changes
+These improvements do not come without downsides, however, especially when it comes to power derived from coal.
+Pollutants from coal contribute to climate change and the phase out of coal energy is therefore a major climate goal (IEA 2022).
+Emissions from the production and burning of coal are also directly harmful to health and the environment, including pollutants like sulfur dioxide, nitrogen oxides, and mercury, to name but a few.\footnote{\url{https://www.eia.gov/energyexplained/coal/coal-and-the-environment.php}} While the use of coal had been declining, recent events -- including Russia's invasion of Ukraine -- have led to an increase in the demand for coal.\footnote{\url{https://www.npr.org/2022/08/15/1117560560/a-rising-demand-for-coal-amidst-war-in-ukraine}}
 
-``Much of what we know about the marginal effect of pollution on infant mortality is derived from developed country data.
-However, given the lower levels of air pollution in developed countries, these estimates may not be externally valid to the developing country context if there is a non-linear dose relationship between pollution and mortality or if the costs of avoidance behaviour differ considerably between the two contexts.
-In this article, we estimate the relationship between pollution and infant mortality using data from Mexico.
-Our estimates for PM10 tend to be similar (or even smaller) than the US estimates, while our findings on CO tend to be larger than those derived from the US context.'' (Arceo, Hanna, and Oliva 2016)
+In this paper, I study the effects of emissions from coal power plants on agricultural productivity, labor allocation, and economic growth in India.
+Previous work has suggested that there may be harmful direct effects of air pollution (Heck et al.
+1982; Marshall et al.
+1997) and water pollution (Reddy and Behera 2006) on agricultural productivity, including work showing that gold mines -- which release large amounts of pollutants into the surrounding environment -- lead to lower agricultural productivity in Ghana (Aragón and Rud 2016).
+This paper, however, tackles the broader case of coal plants, which are a leading source of power throughout the world.
+I show that pollution from these plants can lead to decreased agricultural productivity in the case of India.
 
-We have known about the effects of pollution on human health for many years (Brunekreef and Holgate 2002; Kampa and Castanas 2008; Pope III and Dockery 2006)
-good review on the effects of pollution early in life in Currie et al. (2014)
-- also in developing countries, where there is a consistent effect of exposure on infant mortality (Heft-Neal et al. 2018, 2019)
+I combine the roll out of coal plants in India with plausibly exogenous changes in wind direction -- in a similar spirit to (Deryugina et al.
+2019) -- to measure the effects of pollution from coal plants on key development outcomes.
+I first locate the location of coal plants in India by year.
+Using village-level shapefiles, I identify all villages that are located within 30km of a coal plant at any timn during the sample period, which comprises the 1990s to the 2010s, with differences depending on data availability.
+I then download daily wind data and create an identifier for whether a village is located downwind from a coal plant on any given day, aggregating these daily values to different levels depending on the analysis.
 
-``Air pollution has both acute and chronic effects on human health, affecting a number of different systems and organs.
-It ranges from minor upper respiratory irritation to chronic respiratory and heart disease, lung cancer, acute respiratory infections in children and chronic bronchitis in adults, aggravating pre-existing heart and lung disease, or asthmatic attacks.
-In addition, short- and long-term exposures have also been linked with premature mortality and reduced life expectancy.
-These effects of air pollutants on human health and their mechanism of action are briefly discussed.'' Kampa and Castanas (2008)
+After validating that wind direction is a strong predictor of particulate matter, I first document that the total number of days of being downwind from a coal power plant has a negative effect on overall agricultural productivity, using newly available estimates of India-wide agricultural (land) productivity.
+The within-village absolute deviation of wind exposure is around eight days, meaning that median changes in wind direction are responsible for fluctuations in agricultural productivity of around 2.4 percent, with changes of more than five percent from season to season being common.
+Using wind direction as an instrument for particulate matter, an increase of PM 2.5 by one microgram leads to an even larger decrease in agricultural productivity, at around 20 percent.
+This IV strategy is similar to that in Deryugina et al.
+(2019), though they use general ``high-pollution'' wind directions while I concentrate only on wind directions from the location of a coal power plant.
 
-worker productivity and cognitive function:
-- call center workers in China, and at common levels of pollution (Chang et al. 2019)
-- tests (Ebenstein, Lavy, and Roth 2016; Wen and Burke 2022)
-- farm workers in California (Graff Zivin and Neidell 2012)
-- increase in extensive margin (Hanna and Oliva 2015)
-- manufacturing in China (He, Liu, and Salvo 2019); ``We uncover statistically significant adverse output effects from more prolonged exposure, but effects are not large.
-A substantial +10 mug/m3 PM2.5 variation sustained over 25 days reduces daily output by 1 percent.''
+These effects are apparently concentrated in villages that do not experience the most extreme exposure to pollution; villages with higher maximum exposure experience effects only around 10 percent as large as villages with lower maximum exposure.
+Additional heterogeneity analysis shows that, despite the results with respect to the maximum, the effects are increasing in \emph{mean} exposure.
+I also show that the effects are largest in villages with the highest initial agricultural productivity and that pollution and rainfall shocks are compounding; higher exposure along with less rainfall increases the negative effect of either shock individually.
 
-induces migration (Chen, Oliva, and Zhang 2022)
+Robustness checks show that this effect is concentrated in the years after the construction of the power plant, which helps decrease concerns around endogeneity since, conditional on the village fixed effect and rainfall, windfall direction across years and seasons is exogenous.
+Additionally, leads for wind direction do not predict current agricultural productivity, again supporting the identifying assumptions.
 
-overall effects on agricultural productivity taking into accounts any changes in input allocation caused by the increase in pollution.
+While I am not able to explicitly test effects on land productivity, I look at how adults allocate labor over the last seven days using the National Sample Survey and creating a district-level exposure measure based on the timing of the survey for each household.
+An additional day of wind over the previous week decreases overall labor allocation by 0.027 days.
+However, all effects on labor allocation are driven by the older half of my sample; these adults see much large decreases in overall labor allocation, self employment, farm employment, and non-farm employment.
+This is suggestive evidence of productivity effects being driven by the health effects on individuals, though I am not able to rule out direct effects on land productivity with my data.
+
+The final set of results explores whether there are impacts of pollution on aggregate economic development.
+Using nightlights, I show that lagged wind -- i.e.~pollution exposure in the prior year -- has a negative impact on economic development as proxied by nightlights.
+Going from the 25th percentile to the 75th percentile of wind exposure is synonymous with a decrease in the growth of nightlights within a village by around 0.7 percent.
+While this is a relatively small magnitude, it is worth noting that there are more than 100,000 villages in this sample.
+While the effect for a single village might be small, the overall, aggregate effect on growth is quite substantial.
+
+This paper contributes to several strands of literature.
+First, we already have evidence of effects of pollution on different forms of labor activity.
+For example, general levels of pollution decrease productivity of call center works in China (Chang et al.
+2019), lead people to perform worse on cognitive functions (Ebenstein, Lavy, and Roth 2016; Wen and Burke 2022), and also lead to lower levels of farm labor productivity in California, driven by changes in ozone concentration (Graff Zivin and Neidell 2012).
+Also in China, He, Liu, and Salvo (2019) find small but significant negative effects of pollution on manufacturing productivity, while Chen, Oliva, and Zhang (2022) show pollution leads to labor out migration.
+Perhaps relevant to my labor allocation results, Hanna and Oliva (2015) show that decreases in pollution lead to increases in labor allocation, while I show the same relationship among the oldest cohort.
+The paper most related to this one is Aragón and Rud (2016), who show that gold mines in Ghana have substantial negative impacts on agricultural productivity in the surrounding areas.
+This paper differs in two important ways.
+First, I explore effects of pollution from coal plants, which are much more ubiquitous than goald mines across the world.
+Second, I use a different identification strategy, relying on wind blowing pollution from coal plants.
+
+I also contribute to the more general literature on the effects of pollution on health.
+Pollution has substantial negative effects on health in both developing and developed countries (Arceo, Hanna, and Oliva 2016) and we have known of these negative effects for many years (Brunekreef and Holgate 2002; Kampa and Castanas 2008; Pope III and Dockery 2006).
+Currie et al.
+(2014) review the literature on the health effects of pollution during adolescence; we have robust evidence of pollution exposure leading to increases in infant mortality, including in developing countries (Heft-Neal et al.
+2018, 2019).
+Since these effects are often driven by heart and lung disease and asthma (Kampa and Castanas 2008), this my explain the labor allocation effects I identify in this paper.
+
+The rest of the paper is organized as follows.
+I first discuss data and methodology.
+I break this section up into \autoref{data2}, which covers the multiple sources of data used in this paper; \autoref{identification}, which discusses the empirical strategy; and \autoref{validation}, which validates the wind direction variable using data on particulate matter.
+I present the main results in \autoref{results} before concluding in \autoref{conclusion}.
 
 \hypertarget{data-and-methods}{%
-\section{Data and methods}\label{data-and-methods}}
+\section{\texorpdfstring{Data and methods \label{data}}{Data and methods }}\label{data-and-methods}}
 
 \hypertarget{data}{%
-\subsection{Data}\label{data}}
+\subsection{\texorpdfstring{Data \label{data2}}{Data }}\label{data}}
 
@@
 \hypertarget{identification}{%
-\subsection{Identification}\label{identification}}
+\subsection{\texorpdfstring{Identification \label{identification}}{Identification }}\label{identification}}
 
@@
 \hypertarget{an-example-validating-wind-direction}{%
-\subsection{An example: validating wind direction}\label{an-example-validating-wind-direction}}
+\subsection{\texorpdfstring{An example: validating wind direction \label{validation}}{An example: validating wind direction }}\label{an-example-validating-wind-direction}}
 
@@
 \hypertarget{results}{%
-\section{Results}\label{results}}
+\section{\texorpdfstring{Results \label{results}}{Results }}\label{results}}
 
@@
 
+Previous research has shown larger effects of pollution on the health of the oldest (Deryugina et al.
+2019).
+The labor sample includes all adults between 15 and 60 years of age.
+To look at heterogeneity by age, I present an additional set of results interacting the wind variable with an indicator for whether an individual is over 37 years of age.\footnote{This splits the sample into two equal bins of age: those who are 15-37 and those who are 38-60.} These results are in Table \ref{tab:labortableold}, which otherwise presents the same specifications in the last five columns as \ref{tab:labortable}.
+
+\begin{table}
+
+\caption{\label{tab:labortableold}Wind direction and labor allocation, heterogeneity by age}
+\centering
+\begin{threeparttable}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}}
+\toprule
+  & all & self & wage & farm & non-farm\\
+\midrule
+wind & 0.061*** & 0.096*** & -0.034 & 0.068*** & -0.006\\
+ & (0.023) & (0.022) & (0.024) & (0.020) & (0.030)\\
+wind times age>=38 & -0.207*** & -0.207*** & -0.0001 & -0.078*** & -0.129***\\
+ & (0.037) & (0.033) & (0.040) & (0.020) & (0.030)\\
+controls & Yes & Yes & Yes & Yes & Yes\\
+\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+district & Yes & Yes & Yes & Yes & Yes\\
+year & Yes & Yes & Yes & Yes & Yes\\
+\textbf{varying slopes:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+year (by district) & Yes & Yes & Yes & Yes & Yes\\
+\midrule
+observations & 898,856 & 898,856 & 898,856 & 898,856 & 898,856\\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}
+\item Note: Standard errors are in parentheses and are clustered at the district level. Control variables include female, age, age squared, and (years of) education.
+\item * p<0.10 ** p<0.05 *** p<0.01
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+
+Table \ref{tab:labortableold} presents a very clear set of results: the effects are always larger (more negative) for older adults.
+Particularly interestingly, we now see that the mean-zero effect on self employment in the previous table is driven by two stark contrasts: a large increase for the younger group and a large decrease for the older group.
+For the youngest group, this increase appears in farm labor, indicating likely increases in farm self-employment.
+The oldest, on the other hand, have large decreases in both farm and non-farm labor, with larger decreases in the latter.
+
+While speculative, one possible explanation for this pattern is a decrease in labor productivity.
+While the youngest may be able to readjust and increase their overall self-employment labor allocation, perhaps the oldest cannot due to, for example, health effects of pollution, which are well documented (Bentayeb et al.
+2012; Pope 3rd 2000).\footnote{Much of the literature also suggests the youngest are at significant risk of mortality and morbidities.
+However, given my focus on labor allocation, I only discuss older adults here.}
+
 \hypertarget{nightlights}{%
@@
 \hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
+\section{\texorpdfstring{Conclusion \label{conclusion}}{Conclusion }}\label{conclusion}}
+
+This paper documents the negative effects of pollution on agricultural productivity in India.
+I find clear indications that pollution from coal plants leads to lower agricultural productivity.
+Importantly, I use exogenous changes in wind direction to measure exposure to pollution and show that these effects are robust to a range of alternative specifications.
+Results indicate that these effects may be driven by changes in labor allocation patterns, but I am unable to rule out direct effects on land productivity with the data available.
+Finally, these negative effects are also seen in aggregate economic development, with pollution leading to decreases in the growth of nightlights.
+
+These results suggest caution in building new power plants that emit substantial amounts of pollution.
+While the building of additional power generation facilities may be an economic necessity, the location of these facilities may be particular important for multiple aspects of economic activity, not just health and mortality.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-bentayeb2012adverse}{}}%
+Bentayeb, Malek, Marzia Simoni, Nour Baiz, Dan Norback, Sandra Baldacci, Sara Maio, Giovanni Viegi, Isabella Annesi-Maesano, Geriatric Study in Europe on Health Effects of Air Quality in Nursing Homes (GERIE) Group, et al.
+2012.
+{``Adverse Respiratory Effects of Outdoor Air Pollution in the Elderly.''} \emph{{The International Journal of Tuberculosis and Lung Disease}} 16 (9): 1149--61.
+
 \leavevmode\vadjust pre{\hypertarget{ref-brunekreef2002air}{}}%
@@
 {``{What do we know about short-and long-term effects of early-life exposure to pollution?}''} \emph{{Annual Review Resource Economics}} 6 (1): 217--47.
+
+\leavevmode\vadjust pre{\hypertarget{ref-deryugina2019mortality}{}}%
+Deryugina, Tatyana, Garth Heutel, Nolan H Miller, David Molitor, and Julian Reif.
+2019.
+{``{The mortality and medical costs of air pollution: Evidence from changes in wind direction}.''} \emph{{American Economic Review}} 109 (12): 4178--4219.
+
+\leavevmode\vadjust pre{\hypertarget{ref-dinkelman2011effects}{}}%
+Dinkelman, Taryn. 2011. {``{The effects of rural electrification on employment: New evidence from South Africa}.''} \emph{{American Economic Review}} 101 (7): 3078--108.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-iea2022}{}}%
+IEA. 2022. {``{Coal-fired electricity}.''} {International Energy Association}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-kampa2008human}{}}%
 Kampa, Marilena, and Elias Castanas. 2008. {``{Human health effects of air pollution}.''} \emph{{Environmental Pollution}} 151 (2): 362--67.
+
+\leavevmode\vadjust pre{\hypertarget{ref-kline2014local}{}}%
+Kline, Patrick, and Enrico Moretti.
+2014.
+{``{Local economic development, agglomeration economies, and the big push: 100 years of evidence from the Tennessee Valley Authority}.''} \emph{The Quarterly Journal of Economics} 129 (1): 275--331.
+
+\leavevmode\vadjust pre{\hypertarget{ref-lee2020does}{}}%
+Lee, Kenneth, Edward Miguel, and Catherine Wolfram. 2020. {``Does Household Electrification Supercharge Economic Development?''} \emph{{Journal of Economic Perspectives}} 34 (1): 122--44.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-pope2000epidemiology}{}}%
+Pope 3rd, CA. 2000. {``Epidemiology of Fine Particulate Air Pollution and Human Health: Biologic Mechanisms and Who's at Risk?''} \emph{{Environmental Health Perspectives}} 108 (suppl 4): 713--23.
+
 \leavevmode\vadjust pre{\hypertarget{ref-pope2006health}{}}%
@@
 {``{Health effects of fine particulate air pollution: lines that connect}.''} \emph{{Journal of the Air \& Waste Management Association}} 56 (6): 709--42.
+
+\leavevmode\vadjust pre{\hypertarget{ref-reddy2006impact}{}}%
+Reddy, V Ratna, and Bhagirath Behera. 2006. {``{Impact of water pollution on rural communities: An economic analysis}.''} \emph{{Ecological Economics}} 58 (3): 520--37.
+
+\leavevmode\vadjust pre{\hypertarget{ref-rud2012electricity}{}}%
+Rud, Juan Pablo. 2012. {``{Electricity provision and industrial development: Evidence from India}.''} \emph{{Journal of Development Economics}} 97 (2): 352--67.
+
+\leavevmode\vadjust pre{\hypertarget{ref-van2017long}{}}%
+Van de Walle, Dominique, Martin Ravallion, Vibhuti Mendiratta, and Gayatri Koolwal.
+2017.
+{``{Long-term gains from electrification in rural India}.''} \emph{{The World Bank Economic Review}} 31 (2): 385--411.
 

### 2023-01-13 analyses/regressions.R

@@ -526,6 +526,61 @@ saveRDS(labortable, "pollution_development/draft/tables/labortablewinter.rds")
 
 
 
+# control variables
+setFixest_fml(..ctrl = ~ poly(female, 1) + poly(age, 2) + educ)
+df <- df %>% mutate(old = as.numeric(age>37.5),
+                    wind_old = wind*old)
+
+labor1 <- feols((days_self + days_wage) ~ wind+ wind_old | district[year] + year,
+                  data = df,
+                  cluster = "district")
+labor2 <- feols((days_self + days_wage) ~ wind + wind_old + ..ctrl | district[year] + year,
+                  data = df,
+                  cluster = "district")
+labor3 <- feols(days_self ~ wind + wind_old + ..ctrl | district[year] + year,
+                  data = df,
+                  cluster = "district")
+labor4 <- feols(days_wage ~ wind + wind_old + ..ctrl | district[year] + year,
+                  data = df,
+                  cluster = "district")
+labor5 <- feols(days_f ~ wind + wind_old + ..ctrl | district[year] + year,
+                  data = df,
+                  cluster = "district")
+labor6 <- feols(days_nf ~ wind + wind_old + ..ctrl | district[year] + year,
+                  data = df,
+                  cluster = "district")
+
+
+labortable <- etable(
+                      labor1, labor2, labor3, labor4, labor5, labor6,
+                      se.below = TRUE,
+                      depvar = FALSE,
+                      signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.1),
+                      digits = 3,
+                      fitstat = c("n"),
+                      coefstat = "se",
+                      group = list(controls = "poly"), drop = "educ"
+                      )
+labortable <- labortable[-c(11:12),]
+labortable <- as.matrix(labortable)
+rownames(labortable) <- c("wind", "", "wind times age>=38", "", "controls", "fixed effects:",
+                          "district", "year", "varying slopes:", "year (by district)", 
+                          "observations")
+labortable[c(6,9),] <- " "
+saveRDS(labortable, "pollution_development/draft/tables/labortableold.rds")
+
+
+
+
+
+
+
+
+
+
+
+
+
 
 
 # Nightlights

### 2023-01-13 draft/draft.Rmd
First draft!!

@@
 ---
-title: "Pollution, agricultural productivity, and development: Evidence from coal in plants in India^[thanks to...]"
+title: "Pollution, agricultural productivity, and development: Evidence from coal in plants in India^[All scripts used in this paper are publicly available on my GitHub page: github.com/JoshMerfeld/pollution_development]"
 author:

### 2023-01-13 draft/draft.tex
First draft!!

@@
 
-\title{Pollution, agricultural productivity, and development: Evidence from coal in plants in India\footnote{thanks to\ldots{}}}
+\title{Pollution, agricultural productivity, and development: Evidence from coal in plants in India\footnote{All scripts used in this paper are publicly available on my GitHub page: github.com/JoshMerfeld/pollution\_development}}
 \author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}

### 2023-01-14 draft/draft.Rmd
updated draft

@@
   - \usepackage[hang,flushmargin]{footmisc}		# no indents on footnotes
+  - \usepackage{makecell}
 ---
@@
 # Tables
+plantresultstable <- readRDS("pollution_development/draft/tables/plantresultstable.rds")
 pollutiontable <- readRDS("pollution_development/draft/tables/pollutiontable.rds")
@@
 
+## Where do coal plants open?
+
+A large question looms with respect to identification: what are the characteristics of areas near the construction of new coal plants?
+While I examine the identifying assumptions in a panel context below, I begin with this static question here.
+Table \@ref(tab:plantresultstable) presents four separate regressions.
+The first two columns use 1991 census data -- downloaded from SHRUG -- and the second two columns use 2001 census data, all at the village level.
+The first column presents results from regressing a dummy for whether there is a coal plant within 30km in 1991 on several census variables.
+Villages with higher populations and higher literacy in 1991 are more likely to be near a coal plant at the time.
+Interestingly, however, they are less likely to have a paved road.
+
+```{r plantresultstable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+# Table
+colnames(plantresultstable) <- c("1991", "2001", "2001", "2011")
+kable(
+      plantresultstable, caption = "Local characteristics and the construction of coal plants",
+      align = "cccc", booktabs = TRUE, linesep = "", escape = FALSE
+      ) %>%
+  footnote(
+           "Note: Robust standard errors are in parentheses.
+The outcome in the first column is whether the village is within 30km of a coal plant in 1991.
+The second column is whether a village in 1991 will have a coal plant in 2001, conditional on not having one in 1991.
+The last two columns are similarly defined, except using 2001 and 2011 as the years.
+* p<0.10 ** p<0.05 *** p<0.01",
+           general_title = "",
+           threeparttable = TRUE,
+           footnote_as_chunk = TRUE,
+           escape = FALSE
+           ) %>%
+  column_spec(1, width = "3cm") %>%
+  column_spec(c(2:5),width = "2cm") %>%
+  row_spec(9, hline_after = TRUE) %>%
+  add_header_above(c(" ", "1991 census" = 2, "2001 census" = 2)) %>%
+  kable_classic_2()
+
+```
+
+The second column keeps the 1991 census data but changes the outcome variable to a dummy for whether there *will be* a coal plant constructed before 2001.
+This column restricts the sample to only villages that do not already have a coal plant in 1991.
+Population is no longer a significant predictor while the coefficients on literacy and the presence of a paved road have flipped.
+
+The last two columns are particularly relevant to the current results.
+They present similar results to those in the first two columns, but use the 2001 census data and a dummy for the presence of a coal plant in 2001 and 2011 in columns three and four, respectively.
+Importantly, I include agricultural productivity in the monsoon season of 2002 -- the earliest year available -- as an additional predictor.
+In 2001, agricultural productivity is a strong predictor for the presence of a coal plant.
+Moreover, it is also a significant predictor of the construction of a coal plant between 2001 and 2011 (column four).
+Given that the coefficient is positive, this indicates that coal plants are more likely to be built in areas with higher agricultural productivity.
+In other words, any selection in the overall location of coal plants would bias against finding negative effects of pollution on agricultural productivity.
+
+However, note that this is not the same as looking at wind direction. I discuss this more in the results section below.
+
 ## An example: validating wind direction \label{validation}
@@
   footnote(
-           general = c(
-                       "Note: Standard errors are in parentheses and are clustered at the village level.",
-                       "* p<0.10 ** p<0.05 *** p<0.01"
-                       ), 
+           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
            general_title = "",
-           threeparttable = TRUE
+           threeparttable = TRUE,
+           footnote_as_chunk = TRUE
            ) %>%
@@
   footnote(
-           general = c(
-                       "Note: Standard errors are in parentheses and are clustered at the district level. Control variables include female, age, age squared, and (years of) education.",
-                       "* p<0.10 ** p<0.05 *** p<0.01"
-                       ), 
+           "Note: Standard errors are in parentheses and are clustered at the district level.
+Control variables include female, age, age squared, and (years of) education.
+* p<0.10 ** p<0.05 *** p<0.01",
            general_title = "",
-           threeparttable = TRUE
+           threeparttable = TRUE,
+           footnote_as_chunk = TRUE
            ) %>%
@@
   footnote(
-           general = c(
            "Note: NCAR stands for the National Center for Atmospheric Research.
-Additional information on Global Energy Monitor and NCAR can be found on their websites, at https://globalenergymonitor.org/projects/global-coal-plant-tracker and https://climatedataguide.ucar.edu/, respectively."
-                       ), 
+Additional information on Global Energy Monitor and NCAR can be found on their websites, at https://globalenergymonitor.org/projects/global-coal-plant-tracker and https://climatedataguide.ucar.edu/, respectively.",
            general_title = "",
-           threeparttable = TRUE
+           threeparttable = TRUE,
+           footnote_as_chunk = TRUE
            ) %>%
@@
   footnote(
-           general = c(
-                       "Note: Standard errors are in parentheses and are clustered at the village level. The sample is restricted to village-years after a coal plant was built near the village.",
-                       "* p<0.10 ** p<0.05 *** p<0.01"
-                       ), 
+           "Note: Standard errors are in parentheses and are clustered at the village level.
+The sample is restricted to village-years after a coal plant was built near the village.
+* p<0.10 ** p<0.05 *** p<0.01",
            general_title = "",
-           threeparttable = TRUE
+           threeparttable = TRUE,
+           footnote_as_chunk = TRUE
            ) %>%

### 2023-01-14 draft/draft.tex
updated draft

@@
 \usepackage[hang,flushmargin]{footmisc}
+\usepackage{makecell}
 \usepackage{booktabs}
@@
 \author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
-\date{2023-01-13}
+\date{2023-01-15}
 
@@
 
+\hypertarget{where-do-coal-plants-open}{%
+\subsection{Where do coal plants open?}\label{where-do-coal-plants-open}}
+
+A large question looms with respect to identification: what are the characteristics of areas near the construction of new coal plants?
+While I examine the identifying assumptions in a panel context below, I begin with this static question here.
+Table \ref{tab:plantresultstable} presents four separate regressions.
+The first two columns use 1991 census data -- downloaded from SHRUG -- and the second two columns use 2001 census data, all at the village level.
+The first column presents results from regressing a dummy for whether there is a coal plant within 30km in 1991 on several census variables.
+Villages with higher populations and higher literacy in 1991 are more likely to be near a coal plant at the time.
+Interestingly, however, they are less likely to have a paved road.
+
+\begin{table}
+
+\caption{\label{tab:plantresultstable}Local characteristics and the construction of coal plants}
+\centering
+\begin{threeparttable}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\toprule
+\multicolumn{1}{c}{ } & \multicolumn{2}{c}{1991 census} & \multicolumn{2}{c}{2001 census} \\
+\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
+  & 1991 & 2001 & 2001 & 2011\\
+\midrule
+pop (log) & 0.068*** & 0.0003 & 0.009*** & 0.002***\\
+ & (0.0002) & (0.0002) & (0.0003) & (0.0002)\\
+literacy (prop) & 0.127*** & -0.005*** & 0.043*** & 0.030***\\
+ & (0.002) & (0.001) & (0.002) & (0.002)\\
+has paved road & -0.036*** & 0.003*** & -0.008*** & -2.68e-7\\
+ & (0.0008) & (0.0004) & (0.0008) & (0.0006)\\
+ag productivity &  &  & 0.033*** & 0.005***\\
+ &  &  & (0.0005) & (0.0004)\\
+sub-sample & all & no plant & all & no plant\\
+\midrule
+observations & 490,882 & 447,288 & 516,798 & 476,517\\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}[para]
+\item Note: Robust standard errors are in parentheses.
+The outcome in the first column is whether the village is within 30km of a coal plant in 1991.
+The second column is whether a village in 1991 will have a coal plant in 2001, conditional on not having one in 1991.
+The last two columns are similarly defined, except using 2001 and 2011 as the years.
+* p<0.10 ** p<0.05 *** p<0.01
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+
+The second column keeps the 1991 census data but changes the outcome variable to a dummy for whether there \emph{will be} a coal plant constructed before 2001.
+This column restricts the sample to only villages that do not already have a coal plant in 1991.
+Population is no longer a significant predictor while the coefficients on literacy and the presence of a paved road have flipped.
+
+The last two columns are particularly relevant to the current results.
+They present similar results to those in the first two columns, but use the 2001 census data and a dummy for the presence of a coal plant in 2001 and 2011 in columns three and four, respectively.
+Importantly, I include agricultural productivity in the monsoon season of 2002 -- the earliest year available -- as an additional predictor.
+In 2001, agricultural productivity is a strong predictor for the presence of a coal plant.
+Moreover, it is also a significant predictor of the construction of a coal plant between 2001 and 2011 (column four).
+Given that the coefficient is positive, this indicates that coal plants are more likely to be built in areas with higher agricultural productivity.
+In other words, any selection in the overall location of coal plants would bias against finding negative effects of pollution on agricultural productivity.
+
+However, note that this is not the same as looking at wind direction. I discuss this more in the results section below.
+
 \hypertarget{an-example-validating-wind-direction}{%
@@
 \end{tabular}
-\begin{tablenotes}
-\item Note: Standard errors are in parentheses and are clustered at the village level.
-\item * p<0.10 ** p<0.05 *** p<0.01
+\begin{tablenotes}[para]
+\item Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01
 \end{tablenotes}
@@
 \end{tabular}
-\begin{tablenotes}
-\item Note: Standard errors are in parentheses and are clustered at the district level. Control variables include female, age, age squared, and (years of) education.
-\item * p<0.10 ** p<0.05 *** p<0.01
+\begin{tablenotes}[para]
+\item Note: Standard errors are in parentheses and are clustered at the district level. Control variables include female, age, age squared, and (years of) education. * p<0.10 ** p<0.05 *** p<0.01
 \end{tablenotes}
@@
 \end{tabular}
-\begin{tablenotes}
+\begin{tablenotes}[para]
 \item Note: NCAR stands for the National Center for Atmospheric Research.
@@
 \end{tabular}
-\begin{tablenotes}
-\item Note: Standard errors are in parentheses and are clustered at the village level. The sample is restricted to village-years after a coal plant was built near the village.
-\item * p<0.10 ** p<0.05 *** p<0.01
+\begin{tablenotes}[para]
+\item Note: Standard errors are in parentheses and are clustered at the village level.
+The sample is restricted to village-years after a coal plant was built near the village.
+* p<0.10 ** p<0.05 *** p<0.01
 \end{tablenotes}

### 2023-01-14 analyses/regressions.R

@@ -21,6 +21,61 @@ getwd()    # check
 
 
 
+# Where do plants open?
+villages90 <- read_csv("data/clean/census_plants/villages90.csv")
+villages90 <- villages90 %>% mutate(
+                                    literacy = pc91_pca_p_lit/pc91_pca_tot_p,
+                                    pop = pc91_pca_tot_p,
+                                    scst_prop = (pc91_pca_p_sc + pc91_pca_p_st)/pop,
+                                    tar_road = pc91_vd_tar_road
+                                    )
+villages00 <- read_csv("data/clean/census_plants/villages00.csv")
+villages00 <- villages00 %>% group_by(shrid) %>%
+                             mutate(
+                                    literacy = sum(pc01_pca_p_lit)/sum(pc01_pca_tot_p),
+                                    pop = sum(pc01_pca_tot_p),
+                                    scst_prop = (sum(pc01_pca_p_sc) + sum(pc01_pca_p_st))/pop,
+                                    tar_road = max(pc01_vd_tar_road)
+                                    ) %>%
+                              filter(row_number()==1) %>%
+                              ungroup()
+monsoon <- read_csv("data/clean/census_plants/monsoon2002.csv")
+monsoon <- monsoon %>% group_by(shrid) %>% 
+                        mutate(ag_prod = log(mean(mean))) %>%
+                        filter(row_number()==1) %>%
+                        ungroup()
+villages00 <- villages00 %>% left_join(monsoon, by = "shrid")
+
+
+reg1 <- feols(plants90 ~ log(pop) + literacy + tar_road, data = villages90, vcov = "hetero")
+reg2 <- feols(plants00 ~ log(pop) + literacy + tar_road, data = villages90 %>% filter(plants90==0), vcov = "hetero")
+reg3 <- feols(plants00 ~ ag_prod + log(pop) + literacy + tar_road, data = villages00, vcov = "hetero")
+reg4 <- feols(plants10 ~ ag_prod + log(pop) + literacy + tar_road, data = villages00 %>% filter(plants00==0), vcov = "hetero")
+
+plantresultstable <- etable(
+                          reg1, reg2, reg3, reg4,
+                          se.below = TRUE,
+                          depvar = FALSE,
+                          signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.1),
+                          digits = 3,
+                          fitstat = c("n"),
+                          coefstat = "se",
[…]

### 2023-01-14 cleaning/cleaning_coalplantopenings.R

@@ -1,0 +1,147 @@
+# Purpose: This script cleans the data for looking at where coal plants open
+# Author: Josh Merfeld
+# Date: December 12th, 2022
+
+rm(list = ls())
+
+library(tidyverse)
+library(sf)
+library(rgeos)
+library(sp)
+library(rgdal)
+library(raster)
+library(openxlsx)
+library(exactextractr)  
+
+
+# Sets wd to folder this script is in so we can create relative paths instead of absolute paths
+setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
+# We want to get the wd up to the main folder
+# Need to go up two levels
+setwd("../../")
+# Double check
+getwd()    # check
+
+
+# Shapefile
+villages <- read_sf("data/spatial/shapefiles/village.shp")
+villages <- villages %>% mutate(shrid = paste0(pc11_s_id, "-", pc11_tv_id)) %>% dplyr::select(shrid)
+
+# Extract productivity just for FIRST season
+r <- raster(paste0("data/raw/ag_productivity/NGPPCY-IN-M2002.tif"))
+# extract
+monsoon <- exact_extract(r, villages, fun = "mean", append_cols = "shrid")
+write_csv(monsoon, "data/clean/census_plants/monsoon2002.csv")
+rm(r, monsoon)
+
+
+
+# centroids
+centroids <- gCentroid( as_Spatial(villages), byid = TRUE )
+# Project to meters
+centroids <- st_transform(st_as_sf(centroids), "EPSG:24378")
+
+# coal plants
+plants <- read.xlsx("data/raw/coal_plants.xlsx", sheet = "Units")
+# India only
+plants <- plants %>% filter(Country=="India")
+# Also want only things with a non-missing year
+plants <- plants %>% filter(is.na(Year)==F)
+
+# Just keep what we want 
+plants <- plants %>% dplyr::select(plant_id = ParentID, 
+                                   unit_id = Tracker.ID, 
+                                   capacity = `Capacity.(MW)`, 
+                                   year_built = Year, 
+                                   year_retired = RETIRED, 
+                                   lat = Latitude, 
+                                   lon = Longitude)
+# Plants exist IN 1991
+plants90 <- plants %>% filter(year_built<1991 & (year_retired>1991 | is.na(year_retired)==T))
+gps <- plants90 %>% dplyr::select(lon, lat)
[…]

### 2023-01-16 draft/draft.Rmd
updating draft

@@
   | \noindent The rapid increase in population in India from the 1990s through the 2010s led to the rapid construction of new coal power plants throughout the country.
-Using this roll-out combined with randdom changes in wind direction, I document negative effects of pollution from coal power plants on agricultural productivity throughout the country.
+Using this roll-out combined with random changes in wind direction, I document negative effects of pollution from coal power plants on agricultural productivity throughout the country.
 Results suggest this decrease may be driven by a decline in labor allocation of the oldest half of my sample, though I cannot rule out direct effects of pollution on land productivity.
@@
 yieldtable <- readRDS("pollution_development/draft/tables/yieldtable.rds")
+yieldtablepm <- readRDS("pollution_development/draft/tables/yieldtablepm.rds")
 yieldtabletwo <- readRDS("pollution_development/draft/tables/yieldtabletwo.rds")
@@
 This paper, however, tackles the broader case of coal plants, which are a leading source of power throughout the world.
-I show that pollution from these plants can lead to decreased agricultural productivity in the case of India.
+I show that air pollution from these plants can lead to decreased agricultural productivity.
 
-I combine the roll out of coal plants in India with plausibly exogenous changes in wind direction -- in a similar spirit to [@deryugina2019mortality] -- to measure the effects of pollution from coal plants on key development outcomes.
+I combine the roll out of coal plants in India with plausibly exogenous changes in wind direction -- in a similar spirit to @deryugina2019mortality -- to measure the effects of pollution from coal plants on key development outcomes.
 I first locate the location of coal plants in India by year.
-Using village-level shapefiles, I identify all villages that are located within 30km of a coal plant at any timn during the sample period, which comprises the 1990s to the 2010s, with differences depending on data availability.
+Using a village-level shapefile, I identify all villages that are located within 30km of a coal plant at any time during the sample period, which comprises the 1990s to the 2010s, with differences depending on data availability.
 I then download daily wind data and create an identifier for whether a village is located downwind from a coal plant on any given day, aggregating these daily values to different levels depending on the analysis.
@@
 The first set of data lists the location of coal plant across the globe.
-This data comes from Global Energy Monitor.^[https://globalenergymonitor.org/projects/global-coal-plant-tracker/] and lists all units generating at least 30 megawatts of electricity.
+This data comes from Global Energy Monitor^[https://globalenergymonitor.org/projects/global-coal-plant-tracker/] and lists all units generating at least 30 megawatts of electricity.
 The data on coal plants includes the year of opening (and, if applicable, the year of retirement), the GPS (latitude/longitude) location of each plant, and the amount of power produced by the units at each plant.
@@
 Additionally, the overall capacity from coal plants increased from `r paste0(reportDecOne(sum(plants1990_capacity$capacity)))` gigawatts to `r paste0(reportDecOne(sum(plants2010_capacity$capacity)))` gigawats, an increase of `r paste0(reportPercent((sum(plants2010_capacity$capacity) - sum(plants1990_capacity$capacity))/sum(plants1990_capacity$capacity)))` percent in just 20 years.
+This increase is driven both by the rollout of new plants as well as the construction of new units at existing plants, the latter of which is not shown in Figure \@ref(fig:plants).
 
@@
 I keep these definitions when matching across data below.
-This data is also publicly available from Nature's data-sharing website.^[https://springernature.figshare.com]
+This data is also publicly available from Nature's data-sharing website.^[https://springernature.figshare.com] Since the main regressions below use village fixed effects, any differences in cropping patterns across areas should not bias estimates on changes, as long as these patterns do not change markedly in response to pollution.
 
@@
 For example, agricultural productivity is defined across five months (e.g.
-the monsoon season is from June to October) so I count the total days a given village is exposed to wind during those five months.
+the monsoon season is from June to October) so I count the total days a given village is exposed to wind blowing from _any_ coal plant within 30km during those five months.
 
@@
 For example, the ARIS-REDS data^[https://www.ncaer.org/project/additional-rural-incomes-survey-aris-rural-economic-demographic-survey-reds] includes GPS identifiers for one year, but that only allows GPS matching for a total of two years: 1999 and 2007/2008.
-The DHS, on the other hand, only has district-level identifiers (except for the newest two waves) and is also not a labor survey.
+The DHS, on the other hand, only has state-level identifiers (except for the newest two waves) and is also not a labor survey.
 Instead, I use the National Sample Survey (NSS) from the Ministry of Statistics and Programme Implementation.
@@
 
-The second column keeps the 1991 census data but changes the outcome variable to a dummy for whether there *will be* a coal plant constructed before 2001.
-This column restricts the sample to only villages that do not already have a coal plant in 1991.
+The second column keeps the 1991 census data but changes the outcome variable to a dummy for whether there *will be* a coal plant constructed within 30km before 2001.
+This column restricts the sample to only villages that do not already have a coal plant within 30km in 1991.
 Population is no longer a significant predictor while the coefficients on literacy and the presence of a paved road have flipped.
@@
 The first set of results are in Table \@ref(tab:yieldtable).
-The unit of analysis is the year season -- from 2002 to 2013 -- the geographic fixed effect is the village-season, and the temporal fixed effect is the year.
+The unit of analysis is the vilalge-year-season -- from 2002 to 2013 -- the geographic fixed effect is the village-season, and the temporal fixed effect is the year.
 The exposure variable is defined for the entire season.
-Since @gangopadhyay2022new define both the monsoon (kharif) season and the winter (rabi) season as five months long, I do as well.
+Since @gangopadhyay2022new define both the monsoon (kharif) season and the winter (rabi) season as five months long, I do as well, using the same five months as the authors.
 This implies that the exposure variable ranges from zero to slightly more than 150.
@@
 
+```{r yieldtablepm, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+# Table
+colnames(yieldtablepm) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
+kable(
+      yieldtablepm, caption = "Particulate matter (PM 2.5) and agricultural productivity",
+      align = "ccccc", booktabs = TRUE, linesep = ""
+      ) %>%
+  footnote(
+           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
+           general_title = "",
+           threeparttable = TRUE,
+           footnote_as_chunk = TRUE
+           ) %>%
+  column_spec(1, width = "3cm") %>%
+  column_spec(c(2:6),width = "2cm") %>%
+  row_spec(5, bold = TRUE) %>%
+  row_spec(9, hline_after = TRUE) %>%
+  add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
+  kable_classic_2()
+
+```
+
+Of course, it is not wind, per se, that we are interested in.
+Instead, we are actually interested in better understanding the effects of exposure to pollution on agricultural productivity.
+Table \@ref(tab:yieldtablepm) presents the same regressions as Table \@ref(fig:yieldtable), with a key difference: instead of wind exposure, it uses average concentration of $\mathrm{PM_{2.5}}$ -- from @hammer2020global -- during the season as the key explanatory variable.
+I rescale particulate matter by dividing by one thousand to help with presentation of the coefficients.
+Several interesting patterns emerge.
+In the first column we see no relationship between particulate matter and agricultural productivity, while there is actually a positive relationship in columns two, four, and five.
+There is a small, negative relationship between pollution and productivity in column three.
+
+Recall, however, the results from above: coal plants were placed in places with higher average agricultural productivity.
+This means that the simple regressions in \@ref(tab:yieldtablepm) might actually be biased upwards: although pollution might negatively affect productivity, pollution is higher in more productive areas due to the endogenous placement of coal plants in the study areas.
+To recover the actual effect of pollution, we need an instrument.
+If you believe that wind direction is exogenous and, conditional on the fixed effects and rainfall, only affects agricultural productivity through pollution -- which seems reasonable since it is hard to imagine what other variables it might affect -- then it is a valid instrument for pollution.
+
+I present the IV results in Table \@ref(tab:yieldtabletwo), with the second-stage results in the top panel and the first-stage results in the bottom panel.
+The first-stage results make clear how strong of an instrument wind is; the t-statistic in column one, for example, is almost 48 while it is still around 11 in column three, meeting even the higher bar for the first-stage F-test argued for by @lee2022valid.
+Also note that the coefficients are quite different from those in Table \@ref(tab:pollutiontable) because the unit of analysis in Table \@ref(tab:yieldtabletwo) is the season, while in Table \@ref(tab:pollutiontable) the unit of analysis is the month.
+
+The second-stage results first indicate that $\mathrm{PM_{2.5}}$ is a very strong predictor of agricultural productivity.
+The smallest coefficient -- that for the monsoon season in column (4) -- and the mean absolute deviation for $\mathrm{PM_{2.5}}$ point to a change in agricultural productivity of around 23 percent, which is much higher than the wind results documented above.
+This makes sense, of course, since the IV is essentially scaling up the effects of wind.
+The comparison of the results to those in the previous table also confirm the concerns around endogenous placement of coal plants.
+The coefficients in Table \@ref(tab:yieldtabletwo) are always more negative than those in Table \@ref(tab:yieldtablepm), and sometimes substantially so.
+In other words, the simple regression of agricultural productivity on pollution results in coefficients on pollution that are upwards biased, since coal plants are placed in more productive areas.
+The IV strategy corrects this bias and leads to a robust, negative relationship between pollution and agricultural productivity.
+
 ```{r yieldtabletwo, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
@@
 
-If you believe that wind direction is exogenous and, conditional on the fixed effects and rainfall, only affects agricultural productivity through pollution -- which seems reasonable since it is hard to imagine what other variables it might affect -- then it is a valid instrument for
-pollution.
-While I opt to present the reduced form results for the rest of this paper, it is nonetheless an interesting exercise to also use wind direction as an instrument for pollution, which I measure using estimated $\mathrm{PM_{2.5}}$ from @hammer2020global.
+Having established that exposure to pollution has a negative effect on agricultural productivity, I now move to heterogeneity in these effects.
+Table \@ref(tab:yieldtablehet) presents four different analyses: effects of exposure by maximum exposure, multiplicative effects of shocks, the possibility of non-linearities, and the differential effects of exposure based on initial agricultural productivity.
 
-The IV results appear in Table \@ref(tab:yieldtabletwo), with the second-stage results in the top panel and the first-stage results in the bottom panel.
-The first-stage results make clear how strong of an instrument wind is; also note that the coefficients are quite different from those in Table \@ref(tab:pollutiontable) because the unit of analysis in Table \@ref(tab:yieldtabletwo) is the season, while in Table \@ref(tab:pollutiontable) the unit of analysis is the month.
-
-The second-stage results indicate that $\mathrm{PM_{2.5}}$ is a very strong predictor of agricultural productivity.
-The smallest coefficient -- that for the monsoon season in column (4) -- and the mean absolute deviation for $\mathrm{PM_{2.5}}$ point to a change in agricultural productivity of around 23 percent.
+The first two columns split the sample based on maximum exposure; villages with maximum exposure above the median are included in column (1) while those below the median are in column (2).
+Interestingly, the largest effects are seen on villages in the lower half of the distribution.
+Apparently, smaller maximum effects lead to larger overall effects of pollution exposure on agricultural productivity.
+In other words, it is not the places where wind is most likely to blow from a coal plant that experience the largest decrease in productivity in response to an additional day of wind-driven pollution.
 
@@
 
-Having established that exposure to pollution has a negative effect on agricultural productivity, I now move to heterogeneity in these effects.
-Table \@ref(tab:yieldtablehet) presents four different analyses: effects of exposure by maximum exposure, multiplicative effects of shocks, the possibility of non-linearities, and the differential effects of exposure based on initial agricultural productivity.
-
-The first two columns split the sample based on average exposure; villages with maximum exposure above the median are included in column (1) while those below the median are in column (2).
-Interestingly, the largest effects are seen on villages in the lower half of the distribution.
-Apparently, smaller maximum effects lead to larger overall effects of pollution exposure on agricultural productivity.
-In other words, it is not the places where wind is most likely to blow from a coal plant that experience the largest decrease in productivity in response to an additional day of wind-driven pollution.
-
 Column three looks add the possible multiplicative effects of rainfall shocks and pollution shocks on agricultural productivity.
@@
 The linear term drops out of the regression since it is invariant within villages, leaving just the interaction term.
-The negative coefficient on the interaction term of -0.001 indicates that the effect of wind is more negative in areas where agricultural productivity is higher.
+The negative coefficient on the interaction term of -0.001 indicates that the effect of pollution is more negative in areas where agricultural productivity is higher.
 Note that agricultural productivity is defined in logs, so this can be interpreted as a percentage effect, not a level effect.

### 2023-01-16 draft/draft.tex
updating draft

@@
 \author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
-\date{2023-01-15}
+\date{2023-01-16}
 
@@
 \noindent The rapid increase in population in India from the 1990s through the 2010s led to the rapid construction of new coal power plants throughout the country.
-Using this roll-out combined with randdom changes in wind direction, I document negative effects of pollution from coal power plants on agricultural productivity throughout the country.
+Using this roll-out combined with random changes in wind direction, I document negative effects of pollution from coal power plants on agricultural productivity throughout the country.
 Results suggest this decrease may be driven by a decline in labor allocation of the oldest half of my sample, though I cannot rule out direct effects of pollution on land productivity.
@@
 During those same two decades, power generation from all coal units of at least 30 megawatts increased by almost 140 percent, from 42.4 gigawatts to more than 100 gigawatts.\footnote{\url{https://globalenergymonitor.org/}} There are no doubt positive effects from this increase in power generation; for example, electricity may lead to economic and productivity growth (Dinkelman 2011; Kline and Moretti 2014; Rud 2012; Van de Walle et al.
-2017), though electrification alone may not be a sufficient condition (Lee, Miguel, and Wolfram 2020).
+2017), though electrification alone may not be a sufficient condition (K.
+Lee, Miguel, and Wolfram 2020).
 
@@
 This paper, however, tackles the broader case of coal plants, which are a leading source of power throughout the world.
-I show that pollution from these plants can lead to decreased agricultural productivity in the case of India.
-
-I combine the roll out of coal plants in India with plausibly exogenous changes in wind direction -- in a similar spirit to (Deryugina et al.
-2019) -- to measure the effects of pollution from coal plants on key development outcomes.
+I show that air pollution from these plants can lead to decreased agricultural productivity.
+
+I combine the roll out of coal plants in India with plausibly exogenous changes in wind direction -- in a similar spirit to Deryugina et al.
+(2019) -- to measure the effects of pollution from coal plants on key development outcomes.
 I first locate the location of coal plants in India by year.
-Using village-level shapefiles, I identify all villages that are located within 30km of a coal plant at any timn during the sample period, which comprises the 1990s to the 2010s, with differences depending on data availability.
+Using a village-level shapefile, I identify all villages that are located within 30km of a coal plant at any time during the sample period, which comprises the 1990s to the 2010s, with differences depending on data availability.
 I then download daily wind data and create an identifier for whether a village is located downwind from a coal plant on any given day, aggregating these daily values to different levels depending on the analysis.
@@
 The first set of data lists the location of coal plant across the globe.
-This data comes from Global Energy Monitor.\footnote{\url{https://globalenergymonitor.org/projects/global-coal-plant-tracker/}} and lists all units generating at least 30 megawatts of electricity.
+This data comes from Global Energy Monitor\footnote{\url{https://globalenergymonitor.org/projects/global-coal-plant-tracker/}} and lists all units generating at least 30 megawatts of electricity.
 The data on coal plants includes the year of opening (and, if applicable, the year of retirement), the GPS (latitude/longitude) location of each plant, and the amount of power produced by the units at each plant.
@@
 Additionally, the overall capacity from coal plants increased from 42.4 gigawatts to 100.4 gigawats, an increase of 136.6 percent in just 20 years.
+This increase is driven both by the rollout of new plants as well as the construction of new units at existing plants, the latter of which is not shown in Figure \ref{fig:plants}.
 
@@
 I keep these definitions when matching across data below.
-This data is also publicly available from Nature's data-sharing website.\footnote{\url{https://springernature.figshare.com}}
+This data is also publicly available from Nature's data-sharing website.\footnote{\url{https://springernature.figshare.com}} Since the main regressions below use village fixed effects, any differences in cropping patterns across areas should not bias estimates on changes, as long as these patterns do not change markedly in response to pollution.
 
@@
 To define exposure, I then pull daily wind direction data from the National Center for Atmospheric Research.\footnote{\url{https://climatedataguide.ucar.edu/}} For each day, I document whether the wind is blowing towards each village centroid.\footnote{I define ``towards'' as within five degrees to help take into account that I am using village centroids, using the x and y components of wind speed.} I then temporally aggregate this daily data depending on the temporal definition of the corresponding outcome.
-For example, agricultural productivity is defined across five months (e.g.~the monsoon season is from June to October) so I count the total days a given village is exposed to wind during those five months.
+For example, agricultural productivity is defined across five months (e.g.~the monsoon season is from June to October) so I count the total days a given village is exposed to wind blowing from \emph{any} coal plant within 30km during those five months.
 
@@
 For example, the ARIS-REDS data\footnote{\url{https://www.ncaer.org/project/additional-rural-incomes-survey-aris-rural-economic-demographic-survey-reds}} includes GPS identifiers for one year, but that only allows GPS matching for a total of two years: 1999 and 2007/2008.
-The DHS, on the other hand, only has district-level identifiers (except for the newest two waves) and is also not a labor survey.
+The DHS, on the other hand, only has state-level identifiers (except for the newest two waves) and is also not a labor survey.
 Instead, I use the National Sample Survey (NSS) from the Ministry of Statistics and Programme Implementation.
@@
 
-The second column keeps the 1991 census data but changes the outcome variable to a dummy for whether there \emph{will be} a coal plant constructed before 2001.
-This column restricts the sample to only villages that do not already have a coal plant in 1991.
+The second column keeps the 1991 census data but changes the outcome variable to a dummy for whether there \emph{will be} a coal plant constructed within 30km before 2001.
+This column restricts the sample to only villages that do not already have a coal plant within 30km in 1991.
 Population is no longer a significant predictor while the coefficients on literacy and the presence of a paved road have flipped.
@@
 The first set of results are in Table \ref{tab:yieldtable}.
-The unit of analysis is the year season -- from 2002 to 2013 -- the geographic fixed effect is the village-season, and the temporal fixed effect is the year.
+The unit of analysis is the vilalge-year-season -- from 2002 to 2013 -- the geographic fixed effect is the village-season, and the temporal fixed effect is the year.
 The exposure variable is defined for the entire season.
 Since Gangopadhyay et al.
-(2022) define both the monsoon (kharif) season and the winter (rabi) season as five months long, I do as well.
+(2022) define both the monsoon (kharif) season and the winter (rabi) season as five months long, I do as well, using the same five months as the authors.
 This implies that the exposure variable ranges from zero to slightly more than 150.
@@
 
+\caption{\label{tab:yieldtablepm}Particulate matter (PM 2.5) and agricultural productivity}
+\centering
+\begin{threeparttable}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\toprule
+\multicolumn{1}{c}{ } & \multicolumn{3}{c}{all} & \multicolumn{1}{c}{monsoon} & \multicolumn{1}{c}{winter} \\
+\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6}
+  & (1) & (2) & (3) & (4) & (5)\\
+\midrule
+particulate matter & 0.0000 & 0.0005*** & -0.0004*** & 0.002*** & 0.0005***\\
+(PM 2.5, '000s) & (0.0000) & (3.11e-5) & (5.71e-5) & (9.88e-5) & (3.69e-5)\\
+rain (z) &  & 0.030*** & 0.009*** & 0.083*** & 0.017***\\
+ &  & (0.0004) & (0.001) & (0.002) & (0.0004)\\
+\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+village-season & Yes & Yes & Yes & No & No\\
+year & Yes & Yes & No & Yes & Yes\\
+district-year-season & No & No & Yes & No & No\\
+village & No & No & No & Yes & Yes\\
+\midrule
+observations & 2,391,533 & 2,375,337 & 2,375,337 & 1,259,123 & 1,116,214\\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}[para]
+\item Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+
+Of course, it is not wind, per se, that we are interested in.
+Instead, we are actually interested in better understanding the effects of exposure to pollution on agricultural productivity.
+Table \ref{tab:yieldtablepm} presents the same regressions as Table \ref{fig:yieldtable}, with a key difference: instead of wind exposure, it uses average concentration of \(\mathrm{PM_{2.5}}\) -- from Hammer et al.
+(2020) -- during the season as the key explanatory variable.
+I rescale particulate matter by dividing by one thousand to help with presentation of the coefficients.
+Several interesting patterns emerge.
+In the first column we see no relationship between particulate matter and agricultural productivity, while there is actually a positive relationship in columns two, four, and five.
+There is a small, negative relationship between pollution and productivity in column three.
+
+Recall, however, the results from above: coal plants were placed in places with higher average agricultural productivity.
+This means that the simple regressions in \ref{tab:yieldtablepm} might actually be biased upwards: although pollution might negatively affect productivity, pollution is higher in more productive areas due to the endogenous placement of coal plants in the study areas.
+To recover the actual effect of pollution, we need an instrument.
+If you believe that wind direction is exogenous and, conditional on the fixed effects and rainfall, only affects agricultural productivity through pollution -- which seems reasonable since it is hard to imagine what other variables it might affect -- then it is a valid instrument for pollution.
+
+I present the IV results in Table \ref{tab:yieldtabletwo}, with the second-stage results in the top panel and the first-stage results in the bottom panel.
+The first-stage results make clear how strong of an instrument wind is; the t-statistic in column one, for example, is almost 48 while it is still around 11 in column three, meeting even the higher bar for the first-stage F-test argued for by D.
+S.
+Lee et al.
+(2022).
+Also note that the coefficients are quite different from those in Table \ref{tab:pollutiontable} because the unit of analysis in Table \ref{tab:yieldtabletwo} is the season, while in Table \ref{tab:pollutiontable} the unit of analysis is the month.
+
+The second-stage results first indicate that \(\mathrm{PM_{2.5}}\) is a very strong predictor of agricultural productivity.
+The smallest coefficient -- that for the monsoon season in column (4) -- and the mean absolute deviation for \(\mathrm{PM_{2.5}}\) point to a change in agricultural productivity of around 23 percent, which is much higher than the wind results documented above.
+This makes sense, of course, since the IV is essentially scaling up the effects of wind.
+The comparison of the results to those in the previous table also confirm the concerns around endogenous placement of coal plants.
+The coefficients in Table \ref{tab:yieldtabletwo} are always more negative than those in Table \ref{tab:yieldtablepm}, and sometimes substantially so.
+In other words, the simple regression of agricultural productivity on pollution results in coefficients on pollution that are upwards biased, since coal plants are placed in more productive areas.
+The IV strategy corrects this bias and leads to a robust, negative relationship between pollution and agricultural productivity.
+
+\begin{table}
+
 \caption{\label{tab:yieldtabletwo}Pollution and agricultural productivity, IV estimates}
@@
 
-If you believe that wind direction is exogenous and, conditional on the fixed effects and rainfall, only affects agricultural productivity through pollution -- which seems reasonable since it is hard to imagine what other variables it might affect -- then it is a valid instrument for
-pollution.
-While I opt to present the reduced form results for the rest of this paper, it is nonetheless an interesting exercise to also use wind direction as an instrument for pollution, which I measure using estimated \(\mathrm{PM_{2.5}}\) from Hammer et al.
-(2020).
-
-The IV results appear in Table \ref{tab:yieldtabletwo}, with the second-stage results in the top panel and the first-stage results in the bottom panel.
-The first-stage results make clear how strong of an instrument wind is; also note that the coefficients are quite different from those in Table \ref{tab:pollutiontable} because the unit of analysis in Table \ref{tab:yieldtabletwo} is the season, while in Table \ref{tab:pollutiontable} the unit of analysis is the month.
-
-The second-stage results indicate that \(\mathrm{PM_{2.5}}\) is a very strong predictor of agricultural productivity.
-The smallest coefficient -- that for the monsoon season in column (4) -- and the mean absolute deviation for \(\mathrm{PM_{2.5}}\) point to a change in agricultural productivity of around 23 percent.
+Having established that exposure to pollution has a negative effect on agricultural productivity, I now move to heterogeneity in these effects.
+Table \ref{tab:yieldtablehet} presents four different analyses: effects of exposure by maximum exposure, multiplicative effects of shocks, the possibility of non-linearities, and the differential effects of exposure based on initial agricultural productivity.
+
+The first two columns split the sample based on maximum exposure; villages with maximum exposure above the median are included in column (1) while those below the median are in column (2).
+Interestingly, the largest effects are seen on villages in the lower half of the distribution.
+Apparently, smaller maximum effects lead to larger overall effects of pollution exposure on agricultural productivity.
+In other words, it is not the places where wind is most likely to blow from a coal plant that experience the largest decrease in productivity in response to an additional day of wind-driven pollution.
 
@@
 
-Having established that exposure to pollution has a negative effect on agricultural productivity, I now move to heterogeneity in these effects.
-Table \ref{tab:yieldtablehet} presents four different analyses: effects of exposure by maximum exposure, multiplicative effects of shocks, the possibility of non-linearities, and the differential effects of exposure based on initial agricultural productivity.
-
-The first two columns split the sample based on average exposure; villages with maximum exposure above the median are included in column (1) while those below the median are in column (2).
-Interestingly, the largest effects are seen on villages in the lower half of the distribution.
-Apparently, smaller maximum effects lead to larger overall effects of pollution exposure on agricultural productivity.
-In other words, it is not the places where wind is most likely to blow from a coal plant that experience the largest decrease in productivity in response to an additional day of wind-driven pollution.
-
 Column three looks add the possible multiplicative effects of rainfall shocks and pollution shocks on agricultural productivity.
@@
 {``{Local economic development, agglomeration economies, and the big push: 100 years of evidence from the Tennessee Valley Authority}.''} \emph{The Quarterly Journal of Economics} 129 (1): 275--331.
+
+\leavevmode\vadjust pre{\hypertarget{ref-lee2022valid}{}}%
+Lee, David S, Justin McCrary, Marcelo J Moreira, and Jack Porter. 2022. {``{Valid t-ratio Inference for IV}.''} \emph{{American Economic Review}} 112 (10): 3260--90.
 

### 2023-01-16 analyses/regressions.R

@@ -260,6 +260,36 @@ saveRDS(yieldtabletwo_merge, "pollution_development/draft/tables/yieldtabletwo.r
 
 
 
+df <- df %>% mutate(pm = (pm25/1000))
+yield1 <- feols(log(yield) ~ pm25 | village^season + year, data = df, cluster = c("village"))
+yield2 <- feols(log(yield) ~ pm25 + rain_z | village^season + year, data = df, cluster = c("village"))
+yield3 <- feols(log(yield) ~ pm25 + rain_z | village^season + year^season^distfe, data = df, cluster = c("village"))
+yield4 <- feols(log(yield) ~ pm25 + rain_z | village + year, data = df %>% filter(season=="monsoon"), cluster = c("village"))
+yield5 <- feols(log(yield) ~ pm25 + rain_z | village + year, data = df %>% filter(season=="winter"), cluster = c("village"))
+
+yieldtable <- etable(
+                      yield1, yield2, yield3, yield4, yield5,
+                      se.below = TRUE,
+                      depvar = FALSE,
+                      signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.1),
+                      digits = 3,
+                      fitstat = c("n"),
+                      coefstat = "se"
+                      #extralines = list("Sub-sample" = c("all", "all", "all", "monsoon", "winter"))
+                      )
+yieldtable <- yieldtable[-c(10,11),]
+# Turn to matrix so that duplicate row names are allowed
+yieldtable <- as.matrix(yieldtable)
+rownames(yieldtable) <- c("particulate matter", "(PM 2.5, '000s)", "rain (z)", "", 
+                          "fixed effects:", "village-season", "year", "district-year-season", "village", "observations")
+yieldtable[5,] <- ""
+yieldtable[1,1] <- "0.0000"
+yieldtable[2,1] <- "(0.0000)"
+saveRDS(yieldtable, "pollution_development/draft/tables/yieldtablepm.rds")
+
+
+
+
 
 
 df <- df %>% mutate(pm = (pm25/1000))

### 2023-01-17 draft/draft.Rmd
small changes to draft

@@
                 dplyr::select(state = ST_CEN_CD, district = DT_CEN_CD, state_name = ST_NM, district_name = DISTRICT)
-districts <- districts %>% filter(state=="02" & district=="02")
+districts <- districts %>% filter(state=="23" & district=="07")
 # Villages
@@
                           dplyr::select(shrid, state = ST_CEN_CD, district = DT_CEN_CD, state_name = ST_NM, district_name = DISTRICT)
-villages <- villages %>% filter(state=="02" & district=="02")
+villages <- villages %>% filter(state=="23" & district=="07")
 wind <- read_csv(paste0("data/clean/wind_ntl/days/date_2010-1-1.csv")) %>% as_tibble()
@@
 india_shape2 <- india_shape %>% st_transform(st_crs(districts))
+
+villages <- st_snap(villages, districts, tolerance = 1250)
 
@@
 
-Robustness checks show that this effect is concentrated in the years after the construction of the power plant, which helps decrease concerns around endogeneity since, conditional on the village fixed effect and rainfall, windfall direction across years and seasons is exogenous.
+Robustness checks show that this effect is concentrated in the years after the construction of the power plant, which helps decrease concerns around endogeneity since, conditional on the village fixed effect and weather (rainfall and temperature), wind direction across years and seasons is exogenous.
 Additionally, leads for wind direction do not predict current agricultural productivity, again supporting the identifying assumptions.
@@
 While I am not able to explicitly test effects on land productivity, I look at how adults allocate labor over the last seven days using the National Sample Survey and creating a district-level exposure measure based on the timing of the survey for each household.
-An additional day of wind over the previous week decreases overall labor allocation by 0.027 days.
+An additional day of wind over the previous week decreases overall labor allocation by `r paste0(substr(labortable[1,2],2,6))` days.
 However, all effects on labor allocation are driven by the older half of my sample; these adults see much large decreases in overall labor allocation, self employment, farm employment, and non-farm employment.
@@
 The points are the location of coal plants on that date.
-The right figure shows the distribution of pollution exposure in a specific district -- Kangra district in Himachal Pradesh -- on the same date.")}
+The right figure shows the distribution of pollution exposure in a specific district -- Guna district in Madhya Pradesh -- on the same date.")}
 
@@
 The main exposure variables, on the other hand, are defined at a lower level of aggregation, at the village.
-I choose to aggregate this village-level exposure to the district level as shown in the second panel of Figure \@ref(fig:windexample).
+I choose to aggregate this village-level exposure to the district level as shown in the second panel of Figure \@ref(fig:windexample) -- the district is a randomly selected district in the data, Guna district in Madhya Pradesh.
 I use the area overlap of the villages with the district to define a weighted average of exposure for the district at a given time.
@@
 
-To calculate the effects of exposure to pollution on different outcomes, I rely on fixed effects and the differential exposure driven by plausibly exogenous wind direction, conditional on total rainfall.
-The regession is of the form
-\begin{gather} y_{it} = \alpha_{i} + \gamma_{t} + \beta wind_{it} + \delta rain_{it} + \varepsilon_{it}, \end{gather}
-where $y_{it}$ is outcome $y$ for unit $i$ at time $t$, $\alpha$ is a geographic fixed effects, $\gamma_t$ is time fixed effects, $wind_{it}$ is exposure to pollution -- defined using wind direction -- and $rain_{it}$ is rainfall.
+To calculate the effects of exposure to pollution on different outcomes, I rely on fixed effects and the differential exposure driven by plausibly exogenous wind direction, conditional on total rainfall and average temperature.
+The regression is of the form
+\begin{gather} y_{it} = \alpha_{i} + \gamma_{t} + \beta wind_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it}, \end{gather}
+where $y_{it}$ is outcome $y$ for unit $i$ at time $t$, $\alpha$ is a geographic fixed effects, $\gamma_t$ is time fixed effects, $wind_{it}$ is exposure to pollution -- defined using wind direction -- $rain_{it}$ is total rainfall during the season, and $temp_{it}$ is average temperature during the season.
 Although I use $i$ and $t$ as subscripts for the fixed effects, note that these change depending on the unit of analysis and the outcome.
@@
 Identification relies on changes in wind direction across time for the same geographic units.
-In other words, conditional on the fixed effects and rainfall, I assume changes in wind direction are as good as random and uncorrelated with the outcome -- e.g.
+In other words, conditional on the fixed effects and weather, I assume changes in wind direction are as good as random and uncorrelated with the outcome -- e.g.
 agricultural productivity -- except through exposure to pollution.
@@
 Across all four columns, the story is the same: when the wind is blowing in the direction of a village from a coal plant within 30 kilometers, estimated $\mathrm{PM_{2.5}}$ is substantially higher.
-The regressions indicate that one additional day of wind in the direction of the village increases mean $\mathrm{PM_{2.5}}$ for the month by between 0.014 and 0.045 $\mathrm{\mu g/m^3}$.
+The regressions indicate that one additional day of wind in the direction of the village increases mean $\mathrm{PM_{2.5}}$ for the month by between `r paste0(substr(pollutiontable[1,2],1,5))` and `r paste0(substr(pollutiontable[1,3],1,5))` $\mathrm{\mu g/m^3}$.
 
@@
 The first column presents the most simple results, with only the wind exposure variable and the fixed effects.
-The second column adds a rainfall variable -- defined as the total rainfall in the season -- the addition of which is motivated by the fact that changes in wind direction could be correlated with changes in weather.
-However, the coefficient on the wind exposure variable is completely unchanged by the addition of rainfall, indicating that weather may not be as worrisome of a confounder as thought.
+The second column adds a rainfall -- defined as the total rainfall in the season -- and average temperature variable, the addition of which is motivated by the fact that changes in wind direction could be correlated with changes in weather.
+However, the coefficient on the wind exposure variable is completely unchanged by the addition of the two weather variables, indicating that weather may not be as worrisome of a confounder as thought.
+The direction of the weather coefficients is also consistent with expectations; higher rainfall is good for agricultural productivity while higher temperature is not.
 
@@
 However, part of this may be due to the fact that the new fixed effects soak up much of the variation in the exposure variable, as shown in Figure \@ref(fig:windexample).
-The last two columns separate the effects into the monsoon season and the winter season, but the overall effect is roughly similar, though rainfall has a larger effect in the monsoon season than the winter season, which is unsurprising.
+The last two columns separate the effects into the monsoon season and the winter season, but the overall effect is roughly similar, though rainfall has a larger effect in the monsoon season than the winter season, which is unsurprising given the winter season is driven by irrigation in many areas.
 
@@
   column_spec(c(2:6),width = "2cm") %>%
-  row_spec(5, bold = TRUE) %>%
-  row_spec(9, hline_after = TRUE) %>%
+  row_spec(7, bold = TRUE) %>%
+  row_spec(11, hline_after = TRUE) %>%
   add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
@@
 To recover the actual effect of pollution, we need an instrument.
-If you believe that wind direction is exogenous and, conditional on the fixed effects and rainfall, only affects agricultural productivity through pollution -- which seems reasonable since it is hard to imagine what other variables it might affect -- then it is a valid instrument for pollution.
+If you believe that wind direction is exogenous and, conditional on the fixed effects and weather variables, only affects agricultural productivity through pollution -- which seems reasonable since it is hard to imagine what other variables it might affect -- then it is a valid instrument for pollution.
 
@@
            ) %>%
-  column_spec(1, width = "3.5cm") %>%
+  column_spec(1, width = "3cm") %>%
   column_spec(c(2:6),width = "2cm") %>%
-  row_spec(c(5,11), bold = TRUE) %>%
-  row_spec(c(9:10), hline_after = TRUE) %>%
+  row_spec(c(7,13), bold = TRUE) %>%
+  row_spec(c(11:12), hline_after = TRUE) %>%
   add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
@@
 # Table
+yieldtablehet[1,5] <- "0.000"
 colnames(yieldtablehet) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
@@
            ) %>%
-  column_spec(1, width = "4cm") %>%
+  column_spec(1, width = "3cm") %>%
   column_spec(c(2:6),width = "2cm") %>%
-  row_spec(c(13), hline_after = TRUE) %>%
-  row_spec(c(11), bold = TRUE) %>%
+  row_spec(c(15), hline_after = TRUE) %>%
+  row_spec(c(13), bold = TRUE) %>%
   add_header_above(c(" ", ">p(50)" = 1, "<=p(50)" = 1, "all" = 3)) %>%
@@
 Column four adds a quadratic in wind to see if the negative effects of pollution on agricultural productivity are increase or decreasing in wind.
-In column four, the coefficient on the linear term is 0.0004 while the coefficient on the quadratic term is -0.0002, meaning that the effect of wind is decreasing for any days of wind over two.
-The negative coefficient also indicates that this negative effect is consistently decreasing, meaning the effect gets worse as average wind days gets higher.
+In column four, the coefficient on the linear term is 0.0006 while the coefficient on the quadratic term is -0.0002, meaning that the effect of wind is decreasing for any days of wind over three The negative coefficient also indicates that this negative effect is consistently decreasing, meaning the effect gets worse as average wind days gets higher.
 Note that this is somewhat different from the regressions in the first two columns, which split the sample by the *maximum* days of wind observed in the sample, not the mean.
@@
 
-data <- matrix(NA, nrow = 6, ncol = 3)
-rownames(data) <- c("shapefile", "coal plants", "wind", "pollution", "agriculture", "nightlights")
+data <- matrix(NA, nrow = 7, ncol = 3)
+rownames(data) <- c("shapefile", "coal plants", "wind", "pollution", "agriculture", "nightlights", "weather")
 colnames(data) <- c("source", "geographic coverage", "temporal coverage")
@@
 data[6,1] <- "Asher et al. (2021)"
+data[7,1] <- "TerraClimate"
 data[1,2] <- "India"
@@
 data[6,2] <- "India"
+data[7,2] <- "global"
 data[1,3] <- " "
@@
 data[6,3] <- "yearly"
-
-# Table
-kable(
-      data, caption = "Remote sensing data sources",
-      align = "ccccc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: NCAR stands for the National Center for Atmospheric Research.
-Additional information on Global Energy Monitor and NCAR can be found on their websites, at https://globalenergymonitor.org/projects/global-coal-plant-tracker and https://climatedataguide.ucar.edu/, respectively.",
+data[7,3] <- "monthly"
+
+# Table
+kable(
+      data, caption = "Data sources",
+      align = "ccccc", booktabs = TRUE, linesep = "", escape = TRUE
+      ) %>%
+  footnote(
+           general = "Global Energy Monitor: globalenergymonitor.org/projects/global-coal-plant-tracker. NCAR: climatedataguide.ucar.edu/. TerraClimate: www.climatologylab.org/terraclimate.html",
            general_title = "",
@@
   column_spec(c(2:3),width = "2cm") %>%
-  row_spec(c(12), hline_after = TRUE) %>%
-  row_spec(c(9), bold = TRUE) %>%
+  row_spec(c(14), hline_after = TRUE) %>%
+  row_spec(c(11), bold = TRUE) %>%
   kable_classic_2()

### 2023-01-17 draft/draft.tex
small changes to draft

@@
 \author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
-\date{2023-01-16}
+\date{2023-01-17}
 
@@
 The points are the location of coal plants on that date.
-The right figure shows the distribution of pollution exposure in a specific district -- Kangra district in Himachal Pradesh -- on the same date.}
+The right figure shows the distribution of pollution exposure in a specific district -- Guna district in Madhya Pradesh -- on the same date.}
 \end{figure}
@@
 The main exposure variables, on the other hand, are defined at a lower level of aggregation, at the village.
-I choose to aggregate this village-level exposure to the district level as shown in the second panel of Figure \ref{fig:windexample}.
+I choose to aggregate this village-level exposure to the district level as shown in the second panel of Figure \ref{fig:windexample} -- the district is a randomly selected district in the data, Guna district in Madhya Pradesh.
 I use the area overlap of the villages with the district to define a weighted average of exposure for the district at a given time.
@@
 
-To calculate the effects of exposure to pollution on different outcomes, I rely on fixed effects and the differential exposure driven by plausibly exogenous wind direction, conditional on total rainfall.
-The regession is of the form
-\begin{gather} y_{it} = \alpha_{i} + \gamma_{t} + \beta wind_{it} + \delta rain_{it} + \varepsilon_{it}, \end{gather}
-where \(y_{it}\) is outcome \(y\) for unit \(i\) at time \(t\), \(\alpha\) is a geographic fixed effects, \(\gamma_t\) is time fixed effects, \(wind_{it}\) is exposure to pollution -- defined using wind direction -- and \(rain_{it}\) is rainfall.
+To calculate the effects of exposure to pollution on different outcomes, I rely on fixed effects and the differential exposure driven by plausibly exogenous wind direction, conditional on total rainfall and average temperature.
+The regression is of the form
+\begin{gather} y_{it} = \alpha_{i} + \gamma_{t} + \beta wind_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it}, \end{gather}
+where \(y_{it}\) is outcome \(y\) for unit \(i\) at time \(t\), \(\alpha\) is a geographic fixed effects, \(\gamma_t\) is time fixed effects, \(wind_{it}\) is exposure to pollution -- defined using wind direction -- \(rain_{it}\) is total rainfall during the season, and \(temp_{it}\) is average temperature during the season.
 Although I use \(i\) and \(t\) as subscripts for the fixed effects, note that these change depending on the unit of analysis and the outcome.
@@
 Identification relies on changes in wind direction across time for the same geographic units.
-In other words, conditional on the fixed effects and rainfall, I assume changes in wind direction are as good as random and uncorrelated with the outcome -- e.g.~agricultural productivity -- except through exposure to pollution.
+In other words, conditional on the fixed effects and weather, I assume changes in wind direction are as good as random and uncorrelated with the outcome -- e.g.~agricultural productivity -- except through exposure to pollution.
 One major threat to identification is that the building of a coal plant could lead to a spurious correlation between the exposure variable and the outcome.
@@
 Across all four columns, the story is the same: when the wind is blowing in the direction of a village from a coal plant within 30 kilometers, estimated \(\mathrm{PM_{2.5}}\) is substantially higher.
-The regressions indicate that one additional day of wind in the direction of the village increases mean \(\mathrm{PM_{2.5}}\) for the month by between 0.014 and 0.045 \(\mathrm{\mu g/m^3}\).
+The regressions indicate that one additional day of wind in the direction of the village increases mean \(\mathrm{PM_{2.5}}\) for the month by between 0.01 and 0.06 \(\mathrm{\mu g/m^3}\).
 
@@
 However, part of this may be due to the fact that the new fixed effects soak up much of the variation in the exposure variable, as shown in Figure \ref{fig:windexample}.
-The last two columns separate the effects into the monsoon season and the winter season, but the overall effect is roughly similar, though rainfall has a larger effect in the monsoon season than the winter season, which is unsurprising.
+The last two columns separate the effects into the monsoon season and the winter season, but the overall effect is roughly similar, though rainfall has a larger effect in the monsoon season than the winter season, which is unsurprising given the winter season is driven by irrigation in many areas.
 
@@
 \midrule
-wind & -0.003*** & -0.003*** & -0.0007*** & -0.002*** & -0.003***\\
- & (0.0002) & (0.0002) & (9.21e-5) & (0.0002) & (0.0003)\\
-rain (z) &  & 0.029*** & 0.009*** & 0.082*** & 0.016***\\
- &  & (0.0004) & (0.001) & (0.002) & (0.0004)\\
+wind & -0.003*** & -0.003*** & -0.0007*** & -0.002*** & -0.004***\\
+ & (0.0002) & (0.0002) & (9.24e-5) & (0.0002) & (0.0003)\\
+rain (z) &  & 0.029*** & 0.008*** & 0.075*** & 0.006***\\
+ &  & (0.0004) & (0.001) & (0.002) & (0.0005)\\
+mean temp (10s) &  & -0.0004*** & -0.007*** & -0.002*** & -0.008***\\
+ &  & (5.53e-5) & (0.0004) & (0.0001) & (9.83e-5)\\
 \textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
@@
 \midrule
-observations & 2,391,533 & 2,375,337 & 2,375,337 & 1,259,123 & 1,116,214\\
+observations & 2,387,417 & 2,371,221 & 2,371,221 & 1,256,951 & 1,114,270\\
 \bottomrule
@@
 \midrule
-particulate matter & 0.0000 & 0.0005*** & -0.0004*** & 0.002*** & 0.0005***\\
-(PM 2.5, '000s) & (0.0000) & (3.11e-5) & (5.71e-5) & (9.88e-5) & (3.69e-5)\\
-rain (z) &  & 0.030*** & 0.009*** & 0.083*** & 0.017***\\
- &  & (0.0004) & (0.001) & (0.002) & (0.0004)\\
+particulate matter & 0.0000 & 0.0005*** & -0.0004*** & 0.002*** & 0.0004***\\
+(PM 2.5, '000s) & (0.0000) & (3.17e-5) & (5.72e-5) & (9.87e-5) & (3.7e-5)\\
+rain (z) &  & 0.030*** & 0.008*** & 0.075*** & 0.007***\\
+ &  & (0.0004) & (0.001) & (0.002) & (0.0005)\\
+mean temp (10s) &  & -0.0003*** & -0.007*** & -0.002*** & -0.008***\\
+ &  & (5.63e-5) & (0.0004) & (0.0001) & (9.79e-5)\\
 \textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
@@
 \begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3.5cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
 \toprule
@@
 \midrule
-particulate matter & -0.021*** & -0.020*** & -0.033*** & -0.013*** & -0.024***\\
+particulate matter & -0.021*** & -0.024*** & -0.032*** & -0.012*** & -0.035***\\
 (PM 2.5) & (0.001) & (0.002) & (0.005) & (0.001) & (0.003)\\
-rain (z) &  & 0.004** & 0.002 & 0.086*** & -0.016***\\
- &  & (0.002) & (0.002) & (0.002) & (0.004)\\
+rain (z) &  & -0.011*** & -4.24e-5 & 0.078*** & -0.042***\\
+ &  & (0.003) & (0.002) & (0.002) & (0.004)\\
+mean temp (10s) &  & -0.005*** & -0.007*** & -0.002*** & -0.008***\\
+ &  & (0.0003) & (0.0004) & (0.0001) & (0.0001)\\
 \textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
@@
 \midrule
-observations & 2,391,533 & 2,375,337 & 2,375,337 & 1,259,123 & 1,116,214\\
+observations & 2,387,417 & 2,371,221 & 2,371,221 & 1,256,951 & 1,114,270\\
 \midrule
 \textbf{first stage:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-wind & 0.143*** & 0.126*** & 0.022*** & 0.155*** & 0.105***\\
- & (0.003) & (0.003) & (0.002) & (0.003) & (0.004)\\
-rain (z) &  & -1.23*** & -0.235*** & 0.301*** & -1.36***\\
- &  & (0.010) & (0.018) & (0.015) & (0.009)\\
+wind & 0.145*** & 0.111*** & 0.022*** & 0.157*** & 0.105***\\
+ & (0.002) & (0.002) & (0.002) & (0.002) & (0.004)\\
+rain (z) &  & -1.68*** & -0.239*** & 0.225*** & -1.37***\\
+ &  & (0.011) & (0.017) & (0.015) & (0.010)\\
+mean temp (10s) &  & -0.192*** & 0.002 & -0.021*** & -0.007***\\
+ &  & (0.0009) & (0.004) & (0.001) & (0.002)\\
 \bottomrule
@@
 \begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
 \toprule
@@
 \midrule
-wind & -0.003*** & -0.032*** & -0.004*** & 0.0004 & -0.0001\\
+wind & -0.003*** & -0.033*** & -0.004*** & 0.0006 & 0.000\\
  & (0.0002) & (0.002) & (0.0003) & (0.0005) & (0.001)\\
-rain (z) & 0.031*** & 0.030*** & 0.029*** & 0.029*** & 0.029***\\
+rain (z) & 0.028*** & 0.032*** & 0.028*** & 0.028*** & 0.028***\\
  & (0.0006) & (0.0005) & (0.0004) & (0.0004) & (0.0004)\\
+mean temp (10s) & -0.001*** & 0.0006*** & -0.0004*** & -0.0004*** & -0.0004***\\
+ & (7.66e-5) & (7.9e-5) & (5.55e-5) & (5.56e-5) & (5.49e-5)\\
 wind x rain &  &  & 0.0005*** &  & \\
- &  &  & (6.67e-5) &  & \\
+ &  &  & (6.74e-5) &  & \\
 wind squared &  &  &  & -0.0002*** & \\
- &  &  &  & (3.54e-5) & \\
+ &  &  &  & (3.58e-5) & \\
 wind x starting yield &  &  &  &  & -0.001**\\
@@
 \midrule
-observations & 1,115,694 & 1,259,643 & 2,375,337 & 2,375,337 & 2,371,364\\
+observations & 1,112,082 & 1,259,139 & 2,371,221 & 2,371,221 & 2,367,248\\
 \bottomrule
@@
 Column four adds a quadratic in wind to see if the negative effects of pollution on agricultural productivity are increase or decreasing in wind.
-In column four, the coefficient on the linear term is 0.0004 while the coefficient on the quadratic term is -0.0002, meaning that the effect of wind is decreasing for any days of wind over two.
-The negative coefficient also indicates that this negative effect is consistently decreasing, meaning the effect gets worse as average wind days gets higher.
+In column four, the coefficient on the linear term is 0.0006 while the coefficient on the quadratic term is -0.0002, meaning that the effect of wind is decreasing for any days of wind over three The negative coefficient also indicates that this negative effect is consistently decreasing, meaning the effect gets worse as average wind days gets higher.
 Note that this is somewhat different from the regressions in the first two columns, which split the sample by the \emph{maximum} days of wind observed in the sample, not the mean.
@@
 
-\caption{\label{tab:data}Remote sensing data sources}
+\caption{\label{tab:data}Data sources}
 \centering
@@
 nightlights & Asher et al. (2021) & India & yearly\\
-\bottomrule
-\end{tabular}
-\begin{tablenotes}[para]
-\item Note: NCAR stands for the National Center for Atmospheric Research.
-Additional information on Global Energy Monitor and NCAR can be found on their websites, at https://globalenergymonitor.org/projects/global-coal-plant-tracker and https://climatedataguide.ucar.edu/, respectively.
+weather & TerraClimate & global & monthly\\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}[para]
+\item Global Energy Monitor: globalenergymonitor.org/projects/global-coal-plant-tracker. NCAR: climatedataguide.ucar.edu/. TerraClimate: www.climatologylab.org/terraclimate.html
 \end{tablenotes}
@@
 \midrule
-wind & -0.003*** & -0.002*** & -0.0007*** & -0.002*** & -0.002***\\
- & (0.0002) & (0.0002) & (9.7e-5) & (0.0002) & (0.0003)\\
-rain (z) &  & 0.031*** & 0.010*** & 0.103*** & 0.016***\\
+wind & -0.003*** & -0.002*** & -0.0006*** & -0.002*** & -0.003***\\
+ & (0.0002) & (0.0002) & (9.56e-5) & (0.0002) & (0.0003)\\
+rain (z) &  & 0.018*** & 0.008*** & 0.061*** & 0.003***\\
  &  & (0.0005) & (0.001) & (0.002) & (0.0005)\\
+mean temp (10s) &  & 0.0008*** & -0.005*** & 0.005*** & -0.008***\\
+ &  & (5.28e-5) & (0.0004) & (0.0001) & (0.0001)\\
 \textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
@@
 \midrule
-observations & 2,062,036 & 2,046,104 & 2,046,104 & 1,089,013 & 957,091\\
+observations & 1,886,956 & 1,872,177 & 1,872,177 & 996,509 & 875,668\\
 \bottomrule
@@
 \midrule
-wind (lag) & 0.004*** & 0.0006*\\
+wind (lag) & 0.004*** & 0.0005\\
  & (0.0007) & (0.0003)\\
-wind & -0.002*** & -0.0007***\\
+wind & -0.002*** & -0.0006***\\
  & (0.0004) & (0.0002)\\
-wind (lead) & -0.004*** & -0.0001\\
- & (0.0001) & (9.8e-5)\\
-rain (z) & 0.021*** & 0.008***\\
- & (0.0005) & (0.001)\\
+wind (lead) & -0.004*** & 0.0000\\
+ & (0.0001) & (0.0000)\\
+rain (z) & 0.022*** & 0.006***\\
+ & (0.0004) & (0.001)\\
+mean temp (10s) & 0.0005*** & -0.007***\\
+ & (6.04e-5) & (0.0004)\\
 \textbf{fixed effects:} & \textbf{} & \textbf{}\\
@@
 \midrule
-observations & 1,979,635 & 1,979,635\\
+observations & 1,976,205 & 1,976,205\\
 \bottomrule

### 2023-01-17 analyses/regressions.R

@@ -182,12 +182,15 @@ summary(df$wind*10)
 # Min. 1st Qu.  Median    Mean  3rd Qu.   Max. 
 # 0.00    0.00    2.00   14.33   14.00  153.00 
 
+# Mean, divided by ten (for presentation)
+df <- df %>% mutate(temp_mean = (tmax_mean + tmin_mean)/20)
+
 
 yield1 <- feols(log(yield) ~ wind | village^season + year, data = df, cluster = c("village"))
-yield2 <- feols(log(yield) ~ wind + rain_z | village^season + year, data = df, cluster = c("village"))
-yield3 <- feols(log(yield) ~ wind + rain_z | village^season + year^season^distfe, data = df, cluster = c("village"))
-yield4 <- feols(log(yield) ~ wind + rain_z | village + year, data = df %>% filter(season=="monsoon"), cluster = c("village"))
-yield5 <- feols(log(yield) ~ wind + rain_z | village + year, data = df %>% filter(season=="winter"), cluster = c("village"))
+yield2 <- feols(log(yield) ~ wind + rain_z + temp_mean | village^season + year, data = df, cluster = c("village"))
+yield3 <- feols(log(yield) ~ wind + rain_z + temp_mean | village^season + year^season^distfe, data = df, cluster = c("village"))
+yield4 <- feols(log(yield) ~ wind + rain_z + temp_mean | village + year, data = df %>% filter(season=="monsoon"), cluster = c("village"))
+yield5 <- feols(log(yield) ~ wind + rain_z + temp_mean | village + year, data = df %>% filter(season=="winter"), cluster = c("village"))
 
 yieldtable <- etable(
                       yield1, yield2, yield3, yield4, yield5,
@@ -199,22 +202,22 @@ yieldtable <- etable(
                       coefstat = "se"
                       #extralines = list("Sub-sample" = c("all", "all", "all", "monsoon", "winter"))
                       )
-yieldtable <- yieldtable[-c(10,11),]
+yieldtable <- yieldtable[-c(12,13),]
 # Turn to matrix so that duplicate row names are allowed
 yieldtable <- as.matrix(yieldtable)
-rownames(yieldtable) <- c("wind", "", "rain (z)", "", 
+rownames(yieldtable) <- c("wind", "", "rain (z)", "", "mean temp (10s)", "",
[…]

### 2023-01-17 cleaning/cleaning_ag_productivity_village.R

@@ -77,6 +77,17 @@ for (year in 2002:2013){
   rainfall$rain_total <- apply(rainfall[,c(2:6)], 1, sum)
   colnames(rainfall) <- c("shrid", "rain_m1", "rain_m2", "rain_m3", "rain_m4", "rain_m5", "rain_total")
   rainfall$season <- "monsoon"
+  # temp
+  tempMax <- read_csv(paste0("data/clean/terra/tmax", year, ".csv"))
+  tempMax <- tempMax[, c(1, 7:11)]
+  tempMax$tmax_mean <- apply(tempMax[,c(2:6)], 1, mean)
+  colnames(tempMax) <- c("shrid", "tmax_m1", "tmax_m2", "tmax_m3", "tmax_m4", "tmax_m5", "tmax_mean")
+  tempMax$season <- "monsoon"
+  tempMin <- read_csv(paste0("data/clean/terra/tmax", year, ".csv"))
+  tempMin <- tempMin[, c(1, 7:11)]
+  tempMin$tmin_mean <- apply(tempMin[,c(2:6)], 1, mean)
+  colnames(tempMin) <- c("shrid", "tmin_m1", "tmin_m2", "tmin_m3", "tmin_m4", "tmin_m5", "tmin_mean")
+  tempMin$season <- "monsoon"
   # Load the extracted values from above
   extracted_values <- read_csv(paste0("data/clean/ag_productivity/monsoon", year, ".csv"))
   # Fix some mistakes in the extraction
@@ -85,7 +96,45 @@ for (year in 2002:2013){
                         mutate(mean = mean(mean)) %>% 
                         filter(row_number()==1) %>% 
                         ungroup()
+  rainfall <- rainfall %>% 
+              group_by(shrid) %>% 
+              mutate(
+                     rain_m1 = mean(rain_m1),
+                     rain_m2 = mean(rain_m2),
+                     rain_m3 = mean(rain_m3),
+                     rain_m4 = mean(rain_m4),
+                     rain_m5 = mean(rain_m5),
+                     rain_total = mean(rain_total)
+                     ) %>% 
+              filter(row_number()==1) %>% 
+              ungroup()
+  tempMax <- tempMax %>% 
+              group_by(shrid) %>% 
+              mutate(
+                     tmax_m1 = mean(tmax_m1),
+                     tmax_m2 = mean(tmax_m2),
+                     tmax_m3 = mean(tmax_m3),
+                     tmax_m4 = mean(tmax_m4),
[…]

### 2023-01-17 cleaning/cleaning_precip_temp.R

@@ -75,6 +75,6 @@ for (year in 1990:2020){
   print(year)
 }
   
-  
+
   
   

### 2023-01-18 presentation/slides/slides.Rmd
slides

--- 2023-01-17 draft/draft.Rmd
@@
 ---
-title: "Pollution, agricultural productivity, and development: Evidence from coal in plants in India^[All scripts used in this paper are publicly available on my GitHub page: github.com/JoshMerfeld/pollution_development]"
+title: "Pollution, agricultural productivity, and development: Evidence from coal in plants in India"
+subtitle: 
 author: 
-  - Joshua D. Merfeld^[KDI School of Public Policy and Management and IZA; merfeld@kdis.ac.kr]
-date: "`r Sys.Date()`"
-abstract: |
-  | \noindent The rapid increase in population in India from the 1990s through the 2010s led to the rapid construction of new coal power plants throughout the country.
-Using this roll-out combined with random changes in wind direction, I document negative effects of pollution from coal power plants on agricultural productivity throughout the country.
-Results suggest this decrease may be driven by a decline in labor allocation of the oldest half of my sample, though I cannot rule out direct effects of pollution on land productivity.
-These impacts spill over into aggregate economic growth, as well; nightlight growth is slower following years of higher exposure to pollution.
-  |
-  | \textbf{\textit{Keywords}}: pollution, productivity, agriculture, labor, India
-  | \textbf{\textit{JEL Codes}}: H40, I15, J22, O13, Q52, Q53
-
-# Output type and options (no TOC and yes fig captions)
+  - "Joshua D. Merfeld <br> KDI School and IZA"
+date: '`r Sys.Date()`'
 output:
-  bookdown::pdf_document2: 
-    toc: false
-    fig_caption: yes
-    # This also keeps the .tex file that is used to make the output (we will eventually need this for publication)
-    keep_tex: yes
-
-# Set bibliography
-bibliography: references.bib
-
-# This includes latex arguments
-header-includes:
-  - \newcommand{\beginappendix}{   
-      \setcounter{table}{0}  
-      \renewcommand{\thetable}{A\arabic{table}} 
-      \setcounter{figure}{0} 
-      \renewcommand{\thefigure}{A\arabic{figure}}
-    }  # Relabel appendix tables/figures as e.g. A1
-  - \usepackage[capposition=top]{floatrow}
-  - \usepackage{placeins}   # THis is for /FloatBarrier
-  - \usepackage{setspace}
-  - \usepackage{dcolumn}
-  - \usepackage{booktabs}
-  - \usepackage{siunitx}
-  - \usepackage{amsmath}
-  - \usepackage{enumerate}
-  - \usepackage[shortlabels]{enumitem}
-  - \usepackage[hang,flushmargin]{footmisc}		# no indents on footnotes
-  - \usepackage{makecell}
+  xaringan::moon_reader:
+    css: xaringan-themer.css
+    nature:
+      slideNumberFormat: "%current%"
+      highlightStyle: github
+      highlightLines: true
+      ratio: 16:9
+      countIncrementalSlides: true
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, dev = "cairo_pdf")
+options(htmltools.dir.version = FALSE)
+knitr::opts_chunk$set(
+  fig.width=9, fig.height=3.5, fig.retina=3,
+  out.width = "100%",
+  cache = FALSE,
+  echo = TRUE,
+  message = FALSE, 
+  warning = FALSE,
+  hiline = TRUE
+)
 
@@
 library(tidyverse)
+library(gganimate)
+library(transformr)
+library(gifski)
+library(ggthemes)
 library(scales)
@@
 library(raster)
+library(lubridate)
 loadfonts()  # Must be run once in each session (markdown is a new "session" each time you render/knit)
@@
 # We want to get the wd up to the main folder
-# Need to go up two levels
+# Need to go up three levels
+setwd('..')
 setwd('..')
@@
 # Coal plants
-### Load raw data ----------------------------------------------------------------------------------------------------------------------------------------------------------
+### Load raw data ----------------------------------------------------------------------------------
+
 plants <- read.xlsx("data/raw/coal_plants.xlsx", sheet = "Units")
@@
 
-gps_points_plants <- plants %>% dplyr::select(lon, lat)
-plants <- st_as_sf(SpatialPointsDataFrame(gps_points_plants, plants %>% dplyr::select(-c(lon, lat)), proj4string = CRS("EPSG:4326")))
 plants <- plants %>% mutate(capacity = capacity/1000)
 
-plants1990 <- plants %>% 
-                  filter(year_built<=1990 & (year_retired>1990 | is.na(year_retired)==T)) %>%
+# long
+plantsyearly <- c()
+for (year in 1990:2015){
+  plantstemp <- plants %>% filter(year_built<=year & (year_retired>year | is.na(year_retired)==TRUE))
+  plantstemp <- plantstemp %>%
                   group_by(plant_id) %>%
@@
                   filter(row_number()==1) %>%
-                  ungroup()
+                  ungroup() %>%
+                  dplyr::select(plant_id, capacity, lat, lon)
+  plantstemp$year <- year
+  plantsyearly <- rbind(plantsyearly, plantstemp)
+}
 
-plants2010 <- plants %>% 
-                  filter(year_built<=2010 & (year_retired>2010 | is.na(year_retired)==T)) %>%
-                  group_by(plant_id) %>%
-                  mutate(capacity = sum(capacity)) %>%
-                  filter(row_number()==1) %>%
-                  ungroup()
-
-# wind example
-windexample <- readRDS("data/spatial/windexample.rds")
-
+# Wind example
 # district
@@
 villages <- villages %>% filter(state=="23" & district=="07")
-wind <- read_csv(paste0("data/clean/wind_ntl/days/date_2010-1-1.csv")) %>% as_tibble()
-wind$days_sum <- apply(wind[,2:8], 1, FUN = "sum")
-wind <- wind %>% dplyr::select(shrid, days_sum)
-villages <- villages %>% left_join(wind, by = "shrid")
-villages <- villages %>% st_set_crs(st_crs(districts))
-# and crs for plants
-plants1990 <- plants1990 %>% st_transform(st_crs(districts))
-plants2010 <- plants2010 %>% st_transform(st_crs(districts))
-india_shape2 <- india_shape %>% st_transform(st_crs(districts))
+# Go through two months worth of days
+date <- as_date("2010-01-01")
+villageswind <- c()
+for (add_day in 1:100){
+  wind <- read_csv(paste0("data/clean/wind_ntl/days/date_2010-", month(date), "-", day(date), ".csv")) %>% as_tibble()
+  wind <- wind[,1:2]
+  colnames(wind) <- c("shrid", "indicator")
+  villagestemp <- villages %>% left_join(wind, by = "shrid")
+  villagestemp$date <- date
 
-villages <- st_snap(villages, districts, tolerance = 1250)
+  villageswind <- rbind(villageswind, villagestemp)
+
+  date <- date + 1
+}
+
+villageswind <- villageswind %>% st_set_crs(st_crs(districts))
 
@@
 
-plants1990_capacity <- plants %>% filter(year_built<=1990 & (year_retired>1990 | is.na(year_retired)==TRUE))
-plants2010_capacity <- plants %>% filter(year_built<=2010 & (year_retired>2010 | is.na(year_retired)==TRUE))
+```
+
+```{r xaringan-themer, include=FALSE, warning=FALSE}
+library(xaringanthemer)
+# style_duo_accent(
+#   primary_color = "#006334",
+#   secondary_color = "#A7A9AC",
+#   inverse_header_color = "#FFFFFF"
+# )
+# style_mono_light(
+#                  base_color = "#006334"
+#                  )
+style_duo(
+          primary_color = "#A7A9AC", 
+          secondary_color = "#006334"
+          )
+```
+
+## Coal plants over time
+
+```{r plants, echo = FALSE, message = FALSE, warning = FALSE,}
+
+ggplot() +
+    geom_sf(data = india_shape, color = "gray85", fill = "gray80") +
+    geom_point(data = plantsyearly, 
+               aes(x = lon, y = lat, size = capacity),
+               colour = "#006334", alpha = .5) +
+    # Here comes the gganimate code
+    transition_manual(
+                      frames = year,
+                      cumulative = FALSE
+                      ) +
+    labs(title = "year: {current_frame}") +
+    theme(legend.position = "none")
 
@@
 
-```{r, echo = F}
-# This chunk will allow us to add footers below figures
-library(ggplot2)
+---
 
-hook_plot_tex_footer <- function(x, options) {  
-  x_out <- knitr:::hook_plot_tex(x, options) 
-  if(!is.null(options$fig.footer)) {
-    inter <- sprintf("\\floatfoot*{%s}\n\\end{figure}", options$fig.footer[1])
-    x_out <- gsub(x=x_out, pattern="\n\\end{figure}", replacement=inter, fixed=TRUE)
-  }
-  x_out
-}
+## Wind direction - first 100 days of 2010 in Guna district (MP)
 
-knitr::knit_hooks$set(plot=hook_plot_tex_footer)
+```{r wind, echo = FALSE, message = FALSE, warning = FALSE,}
+
+ggplot() +
+    geom_sf(data = districts, color = "gray85", fill = "gray80") +
+    geom_sf(data = villageswind, 
+            aes(fill = indicator),
+            alpha = .5) +
+    # Here comes the gganimate code
+    transition_manual(
+                      frames = date,
+                      cumulative = FALSE
+                      ) +
+    labs(title = "date: {current_frame}") +
+    theme(legend.position = "none")
 
@@
 
-\newpage
-\doublespacing
-
-# Introduction
-
-From 1990 to 2010, India's population grew by more than 40 percent.^[data.worldbank.org] This population growth led to a large increase in demand for power.
-While natural gas has become a popular alternative in much of the world, the Indian government met much of the increased demand through the construction of coal power plants.
-During those same two decades, power generation from all coal units of at least 30 megawatts increased by almost 140 percent, from 42.4 gigawatts to more than 100 gigawatts.^[https://globalenergymonitor.org/] There are no doubt positive effects from this increase in power generation; for example, electricity may lead to economic and productivity growth [@dinkelman2011effects;@kline2014local;@rud2012electricity;@van2017long], though electrification alone may not be a sufficient condition [@lee2020does].
-
-These improvements do not come without downsides, however, especially when it comes to power derived from coal.
-Pollutants from coal contribute to climate change and the phase out of coal energy is therefore a major climate goal [@iea2022].
-Emissions from the production and burning of coal are also directly harmful to health and the environment, including pollutants like sulfur dioxide, nitrogen oxides, and mercury, to name but a few.^[https://www.eia.gov/energyexplained/coal/coal-and-the-environment.php] While the use of coal had been declining, recent events -- including Russia's invasion of Ukraine -- have led to an increase in the demand for coal.^[https://www.npr.org/2022/08/15/1117560560/a-rising-demand-for-coal-amidst-war-in-ukraine]
-
-In this paper, I study the effects of emissions from coal power plants on agricultural productivity, labor allocation, and economic growth in India.
-Previous work has suggested that there may be harmful direct effects of air pollution [@heck1982assessment;@marshall1997hidden] and water pollution [@reddy2006impact] on agricultural productivity, including work showing that gold mines -- which release large amounts of pollutants into the surrounding environment -- lead to lower agricultural productivity in Ghana [@aragon2016polluting].
-This paper, however, tackles the broader case of coal plants, which are a leading source of power throughout the world.
-I show that air pollution from these plants can lead to decreased agricultural productivity.
-
-I combine the roll out of coal plants in India with plausibly exogenous changes in wind direction -- in a similar spirit to @deryugina2019mortality -- to measure the effects of pollution from coal plants on key development outcomes.
-I first locate the location of coal plants in India by year.
-Using a village-level shapefile, I identify all villages that are located within 30km of a coal plant at any time during the sample period, which comprises the 1990s to the 2010s, with differences depending on data availability.
-I then download daily wind data and create an identifier for whether a village is located downwind from a coal plant on any given day, aggregating these daily values to different levels depending on the analysis.
-
-After validating that wind direction is a strong predictor of particulate matter, I first document that the total number of days of being downwind from a coal power plant has a negative effect on overall agricultural productivity, using newly available estimates of India-wide agricultural (land) productivity.
-The within-village absolute deviation of wind exposure is around eight days, meaning that median changes in wind direction are responsible for fluctuations in agricultural productivity of around 2.4 percent, with changes of more than five percent from season to season being common.
-Using wind direction as an instrument for particulate matter, an increase of PM 2.5 by one microgram leads to an even larger decrease in agricultural productivity, at around 20 percent.
-This IV strategy is similar to that in @deryugina2019mortality, though they use general "high-pollution" wind directions while I concentrate only on wind directions from the location of a coal power plant.
-
-These effects are apparently concentrated in villages that do not experience the most extreme exposure to pollution; villages with higher maximum exposure experience effects only around 10 percent as large as villages with lower maximum exposure.
-Additional heterogeneity analysis shows that, despite the results with respect to the maximum, the effects are increasing in *mean* exposure.
-I also show that the effects are largest in villages with the highest initial agricultural productivity and that pollution and rainfall shocks are compounding; higher exposure along with less rainfall increases the negative effect of either shock individually.
-
-Robustness checks show that this effect is concentrated in the years after the construction of the power plant, which helps decrease concerns around endogeneity since, conditional on the village fixed effect and weather (rainfall and temperature), wind direction across years and seasons is exogenous.
-Additionally, leads for wind direction do not predict current agricultural productivity, again supporting the identifying assumptions.
-
-While I am not able to explicitly test effects on land productivity, I look at how adults allocate labor over the last seven days using the National Sample Survey and creating a district-level exposure measure based on the timing of the survey for each household.
-An additional day of wind over the previous week decreases overall labor allocation by `r paste0(substr(labortable[1,2],2,6))` days.
-However, all effects on labor allocation are driven by the older half of my sample; these adults see much large decreases in overall labor allocation, self employment, farm employment, and non-farm employment.
-This is suggestive evidence of productivity effects being driven by the health effects on individuals, though I am not able to rule out direct effects on land productivity with my data.
-
-The final set of results explores whether there are impacts of pollution on aggregate economic development.
-Using nightlights, I show that lagged wind -- i.e.
-pollution exposure in the prior year -- has a negative impact on economic development as proxied by nightlights.
-Going from the 25th percentile to the 75th percentile of wind exposure is synonymous with a decrease in the growth of nightlights within a village by around 0.7 percent.
-While this is a relatively small magnitude, it is worth noting that there are more than 100,000 villages in this sample.
-While the effect for a single village might be small, the overall, aggregate effect on growth is quite substantial.
-
-This paper contributes to several strands of literature.
-First, we already have evidence of effects of pollution on different forms of labor activity.
-For example, general levels of pollution decrease productivity of call center works in China [@chang2019effect], lead people to perform worse on cognitive functions [@ebenstein2016long;@wen2022lower], and also lead to lower levels of farm labor productivity in California, driven by changes in ozone concentration [@graff2012impact].
-Also in China, @he2019severe find small but significant negative effects of pollution on manufacturing productivity, while @chen2022effect show pollution leads to labor out migration.
-Perhaps relevant to my labor allocation results, @hanna2015effect show that decreases in pollution lead to increases in labor allocation, while I show the same relationship among the oldest cohort.
-The paper most related to this one is @aragon2016polluting, who show that gold mines in Ghana have substantial negative impacts on agricultural productivity in the surrounding areas.
-This paper differs in two important ways.
-First, I explore effects of pollution from coal plants, which are much more ubiquitous than goald mines across the world.
-Second, I use a different identification strategy, relying on wind blowing pollution from coal plants.
-
-I also contribute to the more general literature on the effects of pollution on health.
-Pollution has substantial negative effects on health in both developing and developed countries [@arceo2016does] and we have known of these negative effects for many years [@brunekreef2002air;@kampa2008human;@pope2006health].
-@currie2014we review the literature on the health effects of pollution during adolescence; we have robust evidence of pollution exposure leading to increases in infant mortality, including in developing countries [@heft2018robust;@heft2019air].
-Since these effects are often driven by heart and lung disease and asthma [@kampa2008human], this my explain the labor allocation effects I identify in this paper.
-
-The rest of the paper is organized as follows.
-I first discuss data and methodology.
-I break this section up into \autoref{data2}, which covers the multiple sources of data used in this paper; \autoref{identification}, which discusses the empirical strategy; and \autoref{validation}, which validates the wind direction variable using data on particulate matter.
-I present the main results in \autoref{results} before concluding in \autoref{conclusion}.
-
-# Data and methods \label{data}
-
-## Data \label{data2}
-
-The main goal of this paper is to examine whether exposure to exhaust from coal plants affects agricultural productivity, labor allocation, and economic growth.
-I use several sources of data, which I describe here and in Table \@ref(tab:data) in the appendix.
-
-The first set of data lists the location of coal plant across the globe.
-This data comes from Global Energy Monitor^[https://globalenergymonitor.org/projects/global-coal-plant-tracker/] and lists all units generating at least 30 megawatts of electricity.
-The data on coal plants includes the year of opening (and, if applicable, the year of retirement), the GPS (latitude/longitude) location of each plant, and the amount of power produced by the units at each plant.
-For this paper, I do not use the information on the capacity of the plant.
-
-```{r plants, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Coal plants in India from 1990 to 2010", fig.footer = list("Note: The top figure shows the location of coal plants in 1990.
-The bottom figure shows the location of coal plants in 2010.")}
-
-plant1 <- ggplot() +
-                theme_minimal() +
-                geom_sf(data = india_shape2) +
-                geom_sf(data = plants1990, size = 0.75) +
-                labs(x = "lon",
-                     y = "lat",
-                     subtitle = "Figure A: 1990")
-
-plant2 <- ggplot() +
-                theme_minimal() +
-                geom_sf(data = india_shape2) +
-                geom_sf(data = plants2010, size = 0.75) +
-                labs(x = "lon",
-                     y = "lat",
-                     subtitle = "Figure B: 2010")
-
-ggarrange(plant1, plant2)
-
-```
-
-Figure \@ref(fig:plants) presents the location of coal plants in India for two specific years: 1990 and 2010.
-There is a clear increase in the prevalence of coal plants across the country over the two decades.
-Additionally, the overall capacity from coal plants increased from `r paste0(reportDecOne(sum(plants1990_capacity$capacity)))` gigawatts to `r paste0(reportDecOne(sum(plants2010_capacity$capacity)))` gigawats, an increase of `r paste0(reportPercent((sum(plants2010_capacity$capacity) - sum(plants1990_capacity$capacity))/sum(plants1990_capacity$capacity)))` percent in just 20 years.
-This increase is driven both by the rollout of new plants as well as the construction of new units at existing plants, the latter of which is not shown in Figure \@ref(fig:plants).
-
-The second dataset includes agricultural productivity for both the monsoon and winter seasons, from 2002 to 2018.
-To match other data, I only use the data up to 2013.
-This data comes from @gangopadhyay2022new and estimates land productivity (i.e.
-yield, in tons per hectare) for the major crops in India.
-They define the monsoon season as June to October and the winter season as November to March.
-I keep these definitions when matching across data below.
-This data is also publicly available from Nature's data-sharing website.^[https://springernature.figshare.com] Since the main regressions below use village fixed effects, any differences in cropping patterns across areas should not bias estimates on changes, as long as these patterns do not change markedly in response to pollution.
-
-The agricultural productivity data is available as raster files with a resolution of 500m.
-To aggregate this data up to a useful administrative unit, I use village-level shapefiles provided by @almn2021 and publicly available on the SHRUG platform.^[https://www.devdatalab.org/shrug_download/] I aggregate the agricultural productivity data to the village level by extracting mean productivity for each feature in the shapefile.
-I do this separately for each season -- monsoon and winter -- and each year, from 2002 to 2013.
-
-To measure exposure to pollution from coal plants, I first locate all village centroids located within 30km from a coal plant in a given year.
-I choose 30km due to previous research on the effects of (air) pollution [@aragon2016polluting;@bencko1977health;@li2014health].
-I then calculate the direction from coal plants to all village centroids within that 30km radius.
-To define exposure, I then pull daily wind direction data from the National Center for Atmospheric Research.^[https://climatedataguide.ucar.edu/] For each day, I document whether the wind is blowing towards each village centroid.^[I define "towards" as within five degrees to help take into account that I am using village centroids, using the x and y components of wind speed.] I then temporally aggregate this daily data depending on the temporal definition of the corresponding outcome.
-For example, agricultural productivity is defined across five months (e.g.
-the monsoon season is from June to October) so I count the total days a given village is exposed to wind blowing from _any_ coal plant within 30km during those five months.
-
-```{r windexample, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 8, fig.cap = "Wind direction and aggregation examples (2010-01-01)", fig.footer = list("Note: The left figure shows the average wind direction on January 1st, 2010.
-The points are the location of coal plants on that date.
-The right figure shows the distribution of pollution exposure in a specific district -- Guna district in Madhya Pradesh -- on the same date.")}
-
-p1 <- ggplot() + 
-        theme_minimal() +
-        geom_sf(data = india_shape, fill = NA) +
-        geom_tile(data = windexample, aes(x = x, y = y, fill = wind_example), alpha = 0.3) +
-        scale_fill_viridis_c("wind direction\n(degrees)", option = "viridis") +
-        geom_sf(data = plants2010, size = 0.5) +
-        labs(x = "lon",
-             y = "lat",
-             subtitle = "Figure A: Wind direction")
-
-p2 <- ggplot() + 
-        theme_minimal() +
-        geom_sf(data = districts, fill = NA) +
-        geom_sf(data = villages %>% filter(is.na(days_sum)==F), aes(fill = days_sum), lwd = NA) +
-        scale_fill_viridis_c("exposure (days)", option = "viridis") +
-        labs(x = "lon",
-             y = "lat",
-             subtitle = "Figure B: Aggregation to district")
-
-ggarrange(p1, p2,
-          nrow = 2)
-
-```
-
-The top panel in Figure \@ref(fig:windexample) is an example wind direction raster.
-The raster shows wind direction for the entirety of India on January 1st, 2010, as well as the location of coal power plants on that date.
-The prevailing winds on the date differ across the country and, though not shown in the figure, across days.
-This means that the overall exposure for a given area to coal plant emissions changes across time.
-I also pull data on particulate matter to help validate this proxy for exposure.
-This data comes from @hammer2020global and I also aggregate this to the village using the same method detailed above.
-
-To help understand some of the mechanisms driving the relationship between exposure to emissions and agricultural productivity, I also include household survey data.
-Unfortunately, I am not aware of any publicly available survey data that covers multiple years and has identifiers below the district level.
-For example, the ARIS-REDS data^[https://www.ncaer.org/project/additional-rural-incomes-survey-aris-rural-economic-demographic-survey-reds] includes GPS identifiers for one year, but that only allows GPS matching for a total of two years: 1999 and 2007/2008.
-The DHS, on the other hand, only has state-level identifiers (except for the newest two waves) and is also not a labor survey.
-Instead, I use the National Sample Survey (NSS) from the Ministry of Statistics and Programme Implementation.
-
-The NSS is a nationally representative survey that was conducted every couple of years until relatively recently.
-Many rounds of the NSS also included a module on employment that asked about labor allocation over the previous seven days.
-I use the 61st, 62nd, 64th, 66th, and 68th rounds of the NSS, all of which include the employment module.
-The publicly available data also includes the date of survey, which allows me to match pollution exposure to the exact day of the survey.
-
-The biggest issue with using the NSS is that it includes only district-level identifiers.
-The main exposure variables, on the other hand, are defined at a lower level of aggregation, at the village.
-I choose to aggregate this village-level exposure to the district level as shown in the second panel of Figure \@ref(fig:windexample) -- the district is a randomly selected district in the data, Guna district in Madhya Pradesh.
-I use the area overlap of the villages with the district to define a weighted average of exposure for the district at a given time.
-I calculate the area of overlap for each village and the district as well as the overall area of the district and use these variables to define weights for individual villages.
-Any part of the district that does not overlap with one of the villages -- the white area in Figure \@ref(fig:windexample) -- does not have any exposure and, as such, receives a value of zero for exposure.
-
-Finally, I use nightlights as a measure of economic growth [@henderson2012measuring].
-I also take this data directly from SHRUG and @almn2021.
-This covers 1992 through 2013.
-Some of the older data may not be ideal [@gibson2021night], but this is the only way to document changes in economic growth at a more disaggregated level and at a regular basis.
-Although satellite imagery is an attractive alternative [@burke2021using], high-resolution satellite imagery does not go back to the early 1990s.
-
-## Identification \label{identification}
-
-To calculate the effects of exposure to pollution on different outcomes, I rely on fixed effects and the differential exposure driven by plausibly exogenous wind direction, conditional on total rainfall and average temperature.
-The regression is of the form
-\begin{gather} y_{it} = \alpha_{i} + \gamma_{t} + \beta wind_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it}, \end{gather}
-where $y_{it}$ is outcome $y$ for unit $i$ at time $t$, $\alpha$ is a geographic fixed effects, $\gamma_t$ is time fixed effects, $wind_{it}$ is exposure to pollution -- defined using wind direction -- $rain_{it}$ is total rainfall during the season, and $temp_{it}$ is average temperature during the season.
-Although I use $i$ and $t$ as subscripts for the fixed effects, note that these change depending on the unit of analysis and the outcome.
-I explicitly state the level of fixed effects in the tables and results below.
-
-Identification relies on changes in wind direction across time for the same geographic units.
-In other words, conditional on the fixed effects and weather, I assume changes in wind direction are as good as random and uncorrelated with the outcome -- e.g.
-agricultural productivity -- except through exposure to pollution.
-One major threat to identification is that the building of a coal plant could lead to a spurious correlation between the exposure variable and the outcome.
-Specifically, since exposure is defined as zero prior to the building of a coal plant, the exposure variable itself might be endogenous even though wind direction is not.
-To help allay these concerns, I estimate additional regressions restricting the sample to only time periods after which a given location has a coal plant.
-I also present results using leads to show that future wind direction does not seem to predict current outcomes.
+---
 
 ## Where do coal plants open?
-
-A large question looms with respect to identification: what are the characteristics of areas near the construction of new coal plants?
-While I examine the identifying assumptions in a panel context below, I begin with this static question here.
-Table \@ref(tab:plantresultstable) presents four separate regressions.
-The first two columns use 1991 census data -- downloaded from SHRUG -- and the second two columns use 2001 census data, all at the village level.
-The first column presents results from regressing a dummy for whether there is a coal plant within 30km in 1991 on several census variables.
-Villages with higher populations and higher literacy in 1991 are more likely to be near a coal plant at the time.
-Interestingly, however, they are less likely to have a paved road.
 
@@
 
-The second column keeps the 1991 census data but changes the outcome variable to a dummy for whether there *will be* a coal plant constructed within 30km before 2001.
-This column restricts the sample to only villages that do not already have a coal plant within 30km in 1991.
-Population is no longer a significant predictor while the coefficients on literacy and the presence of a paved road have flipped.
+---
 
-The last two columns are particularly relevant to the current results.
-They present similar results to those in the first two columns, but use the 2001 census data and a dummy for the presence of a coal plant in 2001 and 2011 in columns three and four, respectively.
-Importantly, I include agricultural productivity in the monsoon season of 2002 -- the earliest year available -- as an additional predictor.
-In 2001, agricultural productivity is a strong predictor for the presence of a coal plant.
-Moreover, it is also a significant predictor of the construction of a coal plant between 2001 and 2011 (column four).
-Given that the coefficient is positive, this indicates that coal plants are more likely to be built in areas with higher agricultural productivity.
-In other words, any selection in the overall location of coal plants would bias against finding negative effects of pollution on agricultural productivity.
-
-However, note that this is not the same as looking at wind direction. I discuss this more in the results section below.
-
-## An example: validating wind direction \label{validation}
-
-Before moving on to the main results, it is worth taking the time to better understand the fixed effects strategy with an example.
-Consider the data used for the regressions presented in Table \@ref(tab:pollutiontable).
-The outcome variable is particulate matter -- specifically, $\mathrm{PM_{2.5}}$, which is particulate matter no larger than 2.5 micrometers in diameter -- which comes from @hammer2020global.
-$\mathrm{PM_{2.5}}$ is one of the harmful byproducts from coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.^[https://www.epa.gov/airmarkets/power-plants-and-neighboring-communities] This data is available at the monthly level, so in Table \@ref(tab:pollutiontable), I aggregate wind exposure to the monthly level, as well.
+## Validating wind direction
 
@@
 
-Both the particulate matter data and the wind data is at the village level, which means the regressions in Table \@ref(tab:pollutiontable) are at the village level.
-As such, the geographic fixed effects are village fixed effects and the temporal fixed effects are month fixed effects.
-Since the exposure variable is at the village level and we follow villages over time, the standard errors are also clustered at the village level.
+---
 
-In addition to serving as an example, Table \@ref(tab:pollutiontable) also helps validate the use of wind direction as a proxy for exposure to pollution.
-The first two columns include all available years of data while the last two columns restrict estimation only to the years for which the agricultural productivity regressions are estimated.
-Across all four columns, the story is the same: when the wind is blowing in the direction of a village from a coal plant within 30 kilometers, estimated $\mathrm{PM_{2.5}}$ is substantially higher.
-The regressions indicate that one additional day of wind in the direction of the village increases mean $\mathrm{PM_{2.5}}$ for the month by between `r paste0(substr(pollutiontable[1,2],1,5))` and `r paste0(substr(pollutiontable[1,3],1,5))` $\mathrm{\mu g/m^3}$.
-
-To put these numbers into perspective, the World Health Organization's updated guidelines are that average annual exposure should not exceed 15 $\mathrm{\mu g/m^3}$.
-For the month, an increase at the midpoint of the estimated range is equal to an increase of 0.2 percent of the maximum recommended mean concentration from the WHO.
-This is just a single day of wind in the direction of a village.
-Since this includes all villages located within 30km of a coal plant, this is evidence that wind direction can have serious repercussions on the health of hundreds of millions of people in India.
-
-# Results \label{results}
-
-This section presents the main results of this paper.
-I present three sets of results: one set looking at agricultural productivity, another looking at labor allocation, and a final set looking at nightlights, which are a proxy for economic development.
-I go through each of these in turn.
-
-## Agricultural productivity
-
-I first present results for agricultural productivity.
-The outcome of all the regressions in this section is the log of agricultural land productivity, which is defined as tons per hectare.
-The first set of results are in Table \@ref(tab:yieldtable).
-The unit of analysis is the vilalge-year-season -- from 2002 to 2013 -- the geographic fixed effect is the village-season, and the temporal fixed effect is the year.
-The exposure variable is defined for the entire season.
-Since @gangopadhyay2022new define both the monsoon (kharif) season and the winter (rabi) season as five months long, I do as well, using the same five months as the authors.
-This implies that the exposure variable ranges from zero to slightly more than 150.
-
-The first column presents the most simple results, with only the wind exposure variable and the fixed effects.
-The second column adds a rainfall -- defined as the total rainfall in the season -- and average temperature variable, the addition of which is motivated by the fact that changes in wind direction could be correlated with changes in weather.
-However, the coefficient on the wind exposure variable is completely unchanged by the addition of the two weather variables, indicating that weather may not be as worrisome of a confounder as thought.
-The direction of the weather coefficients is also consistent with expectations; higher rainfall is good for agricultural productivity while higher temperature is not.
-
-Column three adds district-by-year-by-season fixed effects as an additional robustness checks.
-This addition decreases the magnitude of the exposure variable but it remains negative and significant.
-However, part of this may be due to the fact that the new fixed effects soak up much of the variation in the exposure variable, as shown in Figure \@ref(fig:windexample).
-The last two columns separate the effects into the monsoon season and the winter season, but the overall effect is roughly similar, though rainfall has a larger effect in the monsoon season than the winter season, which is unsurprising given the winter season is driven by irrigation in many areas.
+## Effects of wind direction on agricultural productivity
 
@@
 
-To put the size of the coefficient in context, it is worth digging a little deeper into the exposure variable.
-The mean within-village absolute deviation in wind exposure is approximately 8.06 days, meaning that the average change in agricultural productivity from year-to-year due to nothing but changes in wind patterns carrying particulate matter is around 2.4 percent relative to the mean.
-This implies that swings in agricultural productivity -- due to wind direction -- of more than five percent are probably quite common, since the absolute deviation includes deviations below and above the mean.
+---
+
+## Effects of PM on agricultural productivity
 
@@
 
-Of course, it is not wind, per se, that we are interested in.
-Instead, we are actually interested in better understanding the effects of exposure to pollution on agricultural productivity.
-Table \@ref(tab:yieldtablepm) presents the same regressions as Table \@ref(fig:yieldtable), with a key difference: instead of wind exposure, it uses average concentration of $\mathrm{PM_{2.5}}$ -- from @hammer2020global -- during the season as the key explanatory variable.
-I rescale particulate matter by dividing by one thousand to help with presentation of the coefficients.
-Several interesting patterns emerge.
-In the first column we see no relationship between particulate matter and agricultural productivity, while there is actually a positive relationship in columns two, four, and five.
-There is a small, negative relationship between pollution and productivity in column three.
+---
 
-Recall, however, the results from above: coal plants were placed in places with higher average agricultural productivity.
-This means that the simple regressions in \@ref(tab:yieldtablepm) might actually be biased upwards: although pollution might negatively affect productivity, pollution is higher in more productive areas due to the endogenous placement of coal plants in the study areas.
-To recover the actual effect of pollution, we need an instrument.
-If you believe that wind direction is exogenous and, conditional on the fixed effects and weather variables, only affects agricultural productivity through pollution -- which seems reasonable since it is hard to imagine what other variables it might affect -- then it is a valid instrument for pollution.
-
-I present the IV results in Table \@ref(tab:yieldtabletwo), with the second-stage results in the top panel and the first-stage results in the bottom panel.
-The first-stage results make clear how strong of an instrument wind is; the t-statistic in column one, for example, is almost 48 while it is still around 11 in column three, meeting even the higher bar for the first-stage F-test argued for by @lee2022valid.
-Also note that the coefficients are quite different from those in Table \@ref(tab:pollutiontable) because the unit of analysis in Table \@ref(tab:yieldtabletwo) is the season, while in Table \@ref(tab:pollutiontable) the unit of analysis is the month.
-
-The second-stage results first indicate that $\mathrm{PM_{2.5}}$ is a very strong predictor of agricultural productivity.
-The smallest coefficient -- that for the monsoon season in column (4) -- and the mean absolute deviation for $\mathrm{PM_{2.5}}$ point to a change in agricultural productivity of around 23 percent, which is much higher than the wind results documented above.
-This makes sense, of course, since the IV is essentially scaling up the effects of wind.
-The comparison of the results to those in the previous table also confirm the concerns around endogenous placement of coal plants.
-The coefficients in Table \@ref(tab:yieldtabletwo) are always more negative than those in Table \@ref(tab:yieldtablepm), and sometimes substantially so.
-In other words, the simple regression of agricultural productivity on pollution results in coefficients on pollution that are upwards biased, since coal plants are placed in more productive areas.
-The IV strategy corrects this bias and leads to a robust, negative relationship between pollution and agricultural productivity.
+## Effects of PM on agricultural productivity, IV estimates
 
@@
 
-Having established that exposure to pollution has a negative effect on agricultural productivity, I now move to heterogeneity in these effects.
-Table \@ref(tab:yieldtablehet) presents four different analyses: effects of exposure by maximum exposure, multiplicative effects of shocks, the possibility of non-linearities, and the differential effects of exposure based on initial agricultural productivity.
+---
 
-The first two columns split the sample based on maximum exposure; villages with maximum exposure above the median are included in column (1) while those below the median are in column (2).
-Interestingly, the largest effects are seen on villages in the lower half of the distribution.
-Apparently, smaller maximum effects lead to larger overall effects of pollution exposure on agricultural productivity.
-In other words, it is not the places where wind is most likely to blow from a coal plant that experience the largest decrease in productivity in response to an additional day of wind-driven pollution.
-
-```{r yieldtablehet, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-# Table
-yieldtablehet[1,5] <- "0.000"
-colnames(yieldtablehet) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
-kable(
-      yieldtablehet, caption = "Heterogeneity in the effects of pollution on productivity",
-      align = "ccccc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
-           ) %>%
-  column_spec(1, width = "3cm") %>%
-  column_spec(c(2:6),width = "2cm") %>%
-  row_spec(c(15), hline_after = TRUE) %>%
-  row_spec(c(13), bold = TRUE) %>%
-  add_header_above(c(" ", ">p(50)" = 1, "<=p(50)" = 1, "all" = 3)) %>%
-  kable_classic_2()
-
-```
-
-Column three looks add the possible multiplicative effects of rainfall shocks and pollution shocks on agricultural productivity.
-In previous tables, rainfall always entered the regression positively -- as we would expect -- indicating that higher rainfall in a given season leads to higher agricultural productivity.
-However, an interesting question is whether rainfall and pollution shocks might interact.
-The positive coefficient on the interaction term between days of wind and rainfall during the agricultural season indicates that there is a multiplicative effect of shocks; when wind days are higher (bad) and rainfall is lower (bad), the overall negative effects are even worse than when just one shock happens.
-Considering that climate change is leading to more variable rainfall and more negative rainfall shocks, these results are potentially worrisome, especially for smallholders.
-
-Column four adds a quadratic in wind to see if the negative effects of pollution on agricultural productivity are increase or decreasing in wind.
-In column four, the coefficient on the linear term is 0.0006 while the coefficient on the quadratic term is -0.0002, meaning that the effect of wind is decreasing for any days of wind over three The negative coefficient also indicates that this negative effect is consistently decreasing, meaning the effect gets worse as average wind days gets higher.
-Note that this is somewhat different from the regressions in the first two columns, which split the sample by the *maximum* days of wind observed in the sample, not the mean.
-
-In the final column I add to the regression an interaction between wind and initial agricultural productivity (defined in 2002, the first year observed in the productivity data).
-The linear term drops out of the regression since it is invariant within villages, leaving just the interaction term.
-The negative coefficient on the interaction term of -0.001 indicates that the effect of pollution is more negative in areas where agricultural productivity is higher.
-Note that agricultural productivity is defined in logs, so this can be interpreted as a percentage effect, not a level effect.
-Apparently, wind-driven pollution is particularly detrimental to agricultural productivity in the most productive places.
-
-## Robustness checks
-
-I present two separate robustness checks in the appendix.
-The first set of analyses attempts to separate the possible endogenous location of coal plants from outcomes.
-Specifically, prior to the opening of a new coal plant in an area, all villages receive a zero for wind.
-This means that, even if wind direction is exogenous, the opening of the coal plant could lead to correlations between the error term and the wind variable.
-
-Although the results with district-year-season help mitigate this concern, Table \@ref(tab:yieldtablepostplant) in the appendix checks this possibility in a different way.
-Specifically, I restrict the sample to only village-years *after* the introduction of a coal plant.
-This drops any pre-plant-opening observations -- all of which are mechanically zeros for wind -- to prevent endogeneity due to the opening.
-The coefficients in Table \@ref(tab:yieldtablepostplant) show that the overall effects are essentially unchanged, with just a small change in coefficient magnitude (driven mostly by rounding).
-
-The second robustness check includes leads for wind to see if they predict current year productivity.
-I present these results in \@ref(tab:yieldtableleads).
-The second column includes district-year-season fixed effects, while the first does not.
-The results make clear the importance of including trends or district-year fixed effects.
-In the first column, leads seem to predict productivity.
-However, this relationship disappears in the second column, indicating that the previous results with district-year fixed effects are likely the most reliable.
-
-## Labor allocation
-
-The previous set of results show a consistent negative effect of wind-driven pollution on agricultural productivity.
-A key question is the mechanism through which this happens; is this driven by changes in labor allocation/productivity or by changes in land productivity?
-While I cannot definitely prove the mechanism, I can shed some light on possibilities by looking at labor allocation decisions.
-Since these shocks generally happen after initial planting decisions, it seems unlikely that changes in land *allocation* could drive the results.
-
-Table \@ref(tab:labortable) presents results for labor allocation over the last seven days using the National Sample Survey and the methodology described above.
-Since the labor recall is for seven days, labor allocation ranges from zero to seven, as does the wind variable.
-Recall that the data is at the district level, not the village level.
-However, it is also more temporally disaggregated, with rolling interviews throughout the year-long survey.
-Figure \@ref(fig:laborplot) in the appendix shows this temporal distribution across the five waves of the survey.
-
-There are some interesting results in Table \@ref(tab:labortable).
-The first two columns show a marginally significant decrease in overall labor allocation (where "labor" is defined as non-domestic self and wage employment).
-The third and fourth columns show that this overall decrease is predominantly from wage employment, not self employment.
-An additional day of wind over the last seven decreases wage employment by 0.034 days in the district.
-Mean wage employment is 1.458 days, meaning one day of wind decreases wage employment by 2.3 percent of the mean.
-Since much of the district is not covered by "treated" villages, this is actually a rather substantive effect on the villages involved.
-
-The last two columns split the effects by sector, into farm and non-farm employment.
-Although agricultural (land) productivity decreases when wind blows in the direction of villages, overall labor allocation to farm labor actually goes *up*, by around 4.5 percent of the mean of 0.784 days.
-Non-farm labor, on the other hand, goes down, by around 2.4 percent of the mean of 2.527 days.
-
-There are two possible explanations for this pattern of results.
-First, it could be that land productivity goes down, leading farmers to increase their farm labor allocation in an attempt to compensate for the decreased productivity.
-A second possibility is that pollution does not affect land productivity at all, at least not directly.
-In this scenario, the worsening air quality could lead to decreased labor productivity, meaning that "effective labor" essentially decreases when pollution is higher.
-Unfortunately, I am not able to disentangle these two possiblities with the data available.
-
-Table \@ref(tab:labortablemonsoon) and Table \@ref(tab:labortablewinter) in the appendix disaggregate these effects into the monsoon and winter seasons.
-Apparently, the entirety of the effect is driven by changes in labor allocation during the monsoon season.
-The dynamics across seasons apparently lead to large differences in the effects of pollution on labor allocation.
-
-```{r labortable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-# Table
-colnames(labortable) <- c("all", "all", "self", "wage", "farm", "non-farm")
-kable(
-      labortable, caption = "Wind direction and labor allocation",
-      align = "cccccc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the district level.
-Control variables include female, age, age squared, and (years of) education.
-* p<0.10 ** p<0.05 *** p<0.01",
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
-           ) %>%
-  column_spec(1, width = "3cm") %>%
-  column_spec(c(2:7),width = "1.5cm") %>%
-  row_spec(c(8), hline_after = TRUE) %>%
-  row_spec(c(4,7), bold = TRUE) %>%
-  kable_classic_2()
-
-```
-
-Previous research has shown larger effects of pollution on the health of the oldest [@deryugina2019mortality].
-The labor sample includes all adults between 15 and 60 years of age.
-To look at heterogeneity by age, I present an additional set of results interacting the wind variable with an indicator for whether an individual is over 37 years of age.^[This splits the sample into two equal bins of age: those who are 15-37 and those who are 38-60.] These results are in Table \@ref(tab:labortableold), which otherwise presents the same specifications in the last five columns as \@ref(tab:labortable).
+## Wind direction and labor allocation, by age
 
@@
 
-Table \@ref(tab:labortableold) presents a very clear set of results: the effects are always larger (more negative) for older adults.
-Particularly interestingly, we now see that the mean-zero effect on self employment in the previous table is driven by two stark contrasts: a large increase for the younger group and a large decrease for the older group.
-For the youngest group, this increase appears in farm labor, indicating likely increases in farm self-employment.
-The oldest, on the other hand, have large decreases in both farm and non-farm labor, with larger decreases in the latter.
+---
 
-While speculative, one possible explanation for this pattern is a decrease in labor productivity.
-While the youngest may be able to readjust and increase their overall self-employment labor allocation, perhaps the oldest cannot due to, for example, health effects of pollution, which are well documented [@bentayeb2012adverse;@pope2000epidemiology].^[Much of the literature also suggests the youngest are at significant risk of mortality and morbidities.
-However, given my focus on labor allocation, I only discuss older adults here.]
-
-## Nightlights
-
-Pollution leads to lower agricultural productivity as well as changes in labor allocation.
-This leads to an obvious question: might higher levels of pollution also directly affect economic growth?
-To explore this question, I look at changes in nightlights relative to changes in pollution.
-It seems reasonable to think that pollution might affect nightlights with a lag, so I include the lag of pollution rather than contemporaneous pollution.
+## Wind direction and growth of nightlights
 
@@
 
-I present these results in Table \@ref(tab:ntltable).
-To help simplify the coefficients, I recode wind to be the proportion of days in the year that a given village experiences wind in their direction, from a coal plant.
-Mean wind in this new specification is 0.066, indicating that a village in the sample experiences wind in its direction on around 6.6 percent of days, or 24 days of the year.
+---
 
-We see a strong, negative relationship between pollution (lagged wind) and nightlights.
-Going from the 25th percentile to the 75th percentile of wind exposure is synonymous with a decrease in the growth of nightlights within a village by around 0.7 percent.
-While this is a relatively small magnitude, it is worth noting that there are more than 100,000 villages in this sample.
-While the effect for a single village might be small, the overall, aggregate effect on growth is quite substantial.
-
-# Conclusion \label{conclusion}
-
-This paper documents the negative effects of pollution on agricultural productivity in India.
-I find clear indications that pollution from coal plants leads to lower agricultural productivity.
-Importantly, I use exogenous changes in wind direction to measure exposure to pollution and show that these effects are robust to a range of alternative specifications.
-Results indicate that these effects may be driven by changes in labor allocation patterns, but I am unable to rule out direct effects on land productivity with the data available.
-Finally, these negative effects are also seen in aggregate economic development, with pollution leading to decreases in the growth of nightlights.
-
-These results suggest caution in building new power plants that emit substantial amounts of pollution.
-While the building of additional power generation facilities may be an economic necessity, the location of these facilities may be particular important for multiple aspects of economic activity, not just health and mortality.
-
-\FloatBarrier
-\newpage
-\singlespacing
-# References {.unnumbered}
-
-::: {#refs} 
-:::
-
-\FloatBarrier
-\newpage
-
-# Appendix A {.unnumbered}
-\beginappendix
-\FloatBarrier
-
-```{r data, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-data <- matrix(NA, nrow = 7, ncol = 3)
-rownames(data) <- c("shapefile", "coal plants", "wind", "pollution", "agriculture", "nightlights", "weather")
-colnames(data) <- c("source", "geographic coverage", "temporal coverage")
-data[1,1] <- "Asher et al. (2021)"
-data[2,1] <- "Global Energy Monitor"
-data[3,1] <- "NCAR"
-data[4,1] <- "Hammer et al. (2020)"
-data[5,1] <- "Angopadhyay et al. (2022)"
-data[6,1] <- "Asher et al. (2021)"
-data[7,1] <- "TerraClimate"
-data[1,2] <- "India"
-data[2,2] <- "global"
-data[3,2] <- "global"
-data[4,2] <- "global"
-data[5,2] <- "global"
-data[6,2] <- "India"
-data[7,2] <- "global"
-data[1,3] <- " "
-data[2,3] <- "yearly"
-data[3,3] <- "daily"
-data[4,3] <- "monthly"
-data[5,3] <- "two seasons/year"
-data[6,3] <- "yearly"
-data[7,3] <- "monthly"
-
-# Table
-kable(
-      data, caption = "Data sources",
-      align = "ccccc", booktabs = TRUE, linesep = "", escape = TRUE
-      ) %>%
-  footnote(
-           general = "Global Energy Monitor: globalenergymonitor.org/projects/global-coal-plant-tracker. NCAR: climatedataguide.ucar.edu/. TerraClimate: www.climatologylab.org/terraclimate.html",
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
-           ) %>%
-  column_spec(1, width = "2cm") %>%
-  column_spec(2,width = "4.5cm") %>%
-  column_spec(c(3:4),width = "3.5cm") %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  kable_classic_2()
-
-```
-
-\newpage
-```{r yieldtablepostplant, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-# Table
-colnames(yieldtablepostplant) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
-kable(
-      yieldtablepostplant, caption = "Wind direction and agricultural productivity, post-coal plant only",
-      align = "ccccc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the village level.
-The sample is restricted to village-years after a coal plant was built near the village.
-* p<0.10 ** p<0.05 *** p<0.01",
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
-           ) %>%
-  column_spec(1, width = "3cm") %>%
-  column_spec(c(2:6),width = "2cm") %>%
-  row_spec(7, bold = TRUE) %>%
-  row_spec(11, hline_after = TRUE) %>%
-  add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  kable_classic_2()
-
-```
-
-\FloatBarrier
-\newpage
-```{r yieldtableleads, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-# Table
-colnames(yieldtableleads) <- c("(1)", "(2)")
-kable(
-      yieldtableleads, caption = "Agricultural productivity and pollution leads",
-      align = "cc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
-           ) %>%
-  column_spec(1, width = "3cm") %>%
-  column_spec(c(2:3),width = "2cm") %>%
-  row_spec(c(14), hline_after = TRUE) %>%
-  row_spec(c(11), bold = TRUE) %>%
-  kable_classic_2()
-
-```
-
-\FloatBarrier
-\newpage
-```{r laborplot, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Timing of household surveys for the National Sample Survey", fig.footer = list("Note: The histogram shows the distribution of survey dates for the four waves of the National Sample Survey (NSS) for the five waves used in this paper.")}
-
-laborplot
-
-```
-
-\FloatBarrier
-\newpage
-```{r labortablemonsoon, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-# Table
-colnames(labortablemonsoon) <- c("all", "all", "self", "wage", "farm", "non-farm")
-kable(
-      labortablemonsoon, caption = "Wind direction and labor allocation, monsoon season only",
-      align = "cccccc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the district level.
-Control variables include female, age, age squared, and (years of) education.
-* p<0.10 ** p<0.05 *** p<0.01",
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
-           ) %>%
-  column_spec(1, width = "3cm") %>%
-  column_spec(c(2:7),width = "1.5cm") %>%
-  row_spec(c(8), hline_after = TRUE) %>%
-  row_spec(c(4,7), bold = TRUE) %>%
-  kable_classic_2()
-
-```
-
-\FloatBarrier
-\newpage
-```{r labortablewinter, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-# Table
-colnames(labortablewinter) <- c("all", "all", "self", "wage", "farm", "non-farm")
-kable(
-      labortablewinter, caption = "Wind direction and labor allocation, winter season only",
-      align = "cccccc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the district level.
-Control variables include female, age, age squared, and (years of) education.
-* p<0.10 ** p<0.05 *** p<0.01",
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
-           ) %>%
-  column_spec(1, width = "3cm") %>%
-  column_spec(c(2:7),width = "1.5cm") %>%
-  row_spec(c(8), hline_after = TRUE) %>%
-  row_spec(c(4,7), bold = TRUE) %>%
-  kable_classic_2()
-
-```
+## Conclusion

### 2023-01-18 draft/draft.Rmd
abstract

@@
 
+Coal is one of the most common energy sources throughout the world.
+In 2021, energy generation from coal reached an all-time high,^[https://www.iea.org/reports/coal-fired-electricity] driven in part by rapid population growth throughout the world.
+Along with record energy production, coal plants also emit large amounts of pollution, leading to record coal-derived carbon dioxide emissions in 2021.^[*ibid.*]
+
+India is no exception to these trends.
 From 1990 to 2010, India's population grew by more than 40 percent.^[data.worldbank.org] This population growth led to a large increase in demand for power.

### 2023-01-18 draft/draft.tex
abstract

@@
 \author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
-\date{2023-01-17}
+\date{2023-01-18}
 
@@
 
+Coal is one of the most common energy sources throughout the world.
+In 2021, energy generation from coal reached an all-time high,\footnote{\url{https://www.iea.org/reports/coal-fired-electricity}} driven in part by rapid population growth throughout the world.
+Along with record energy production, coal plants also emit large amounts of pollution, leading to record coal-derived carbon dioxide emissions in 2021.\footnote{\emph{ibid.}}
+
+India is no exception to these trends.
 From 1990 to 2010, India's population grew by more than 40 percent.\footnote{data.worldbank.org} This population growth led to a large increase in demand for power.

### 2023-01-19 cleaning/1.master.R
changing names of cleaning scripts and adding 1.master to help with reproducibility

@@ -1,0 +1,63 @@
+# Purpose: This script runs all of the other scripts in the proper order
+# Author: Josh Merfeld
+# Date: January 19th, 2023
+
+rm(list = ls())
+
+# Sets wd to folder this script is in so we can create relative paths instead of absolute paths
+setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
+# just make sure WD is the same as current:
+getwd()    # check
+
+
+
+
+
+##################
+# This cleans the nightlight data, taken from SHRUG
+source("2.cleaning_ntl.R")
+
+
+##################
+# This cleans the NSS data
+source("3.cleaning_nss.R")
+
+
+##################
+# This calculates the directions from plants to villages and downloads/cleans wind data.
+# This takes a LONG TIME TO RUN. Like a week.
+source("4.cleaning_plants.R")
+
+
+##################
+# This cleans the precip/temp data (using rasters)
+source("5.cleaning_precip_temp.R")
+
+
+##################
+# This cleans the particulate matter data, and also adds wind to it.
+source("6.cleaning_pm25.R")
+
+
+##################
+# This cleans the agricultural productivity data and adds rain/temp to it so it is ready for analysis at season level
+source("7.cleaning_ag_productivity_village.R")
+
+
+##################
+# This adds the wind data to the NSS data, which requires some aggregation
+source("8.cleaning_overlap_merging_nss.R")
+
+
+##################
+# This gets data ready for analyzing characteristics of villages correlated with plant openings. 1991 and 2001 census data from SHRUG.
+source("9.cleaning_coalplantopenings.R")
+
+
+
+
+
+
+
+
+

### 2023-01-19 cleaning/old/4.cleaning_plants.R

@@ -152,6 +152,7 @@ for (vil in 1:nrow(villages)){
 # Add india
 india_wind <- read_sf("data/spatial/shapefiles/state.shp")
 
+# NTOE: have since learned I could do this much easier just using lubridate. It is what it is. This still works.
 years <- c("2011", "2012", "2013", "2014", "2015")
 months <- c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10", "11", "12")
 month_days <- c("01", "02", "03", "04", "05", "06", "07", "08", "09")

### 2023-01-19 draft/draft.Rmd
draft

--- 2023-01-17 draft/draft.Rmd
@@
 abstract: |
-  | \noindent The rapid increase in population in India from the 1990s through the 2010s led to the rapid construction of new coal power plants throughout the country.
+  | \noindent I show that emissions from a common energy source lead to decreased productivity and growth.
+The rapid increase in population in India from the 1990s through the 2010s led to the rapid construction of new coal power plants throughout the country.
 Using this roll-out combined with random changes in wind direction, I document negative effects of pollution from coal power plants on agricultural productivity throughout the country.
@@
 library(ggpubr)
+library(ggthemes)
 library(rticles)
@@
 
+Coal is one of the most common energy sources throughout the world.
+In 2021, energy generation from coal reached an all-time high,^[www.iea.org/reports/coal-fired-electricity] driven in part by rapid population growth throughout the world.
+India is no exception to these trends.
 From 1990 to 2010, India's population grew by more than 40 percent.^[data.worldbank.org] This population growth led to a large increase in demand for power.
 While natural gas has become a popular alternative in much of the world, the Indian government met much of the increased demand through the construction of coal power plants.
-During those same two decades, power generation from all coal units of at least 30 megawatts increased by almost 140 percent, from 42.4 gigawatts to more than 100 gigawatts.^[https://globalenergymonitor.org/] There are no doubt positive effects from this increase in power generation; for example, electricity may lead to economic and productivity growth [@dinkelman2011effects;@kline2014local;@rud2012electricity;@van2017long], though electrification alone may not be a sufficient condition [@lee2020does].
+During those same two decades, power generation from all coal units of at least 30 megawatts increased by almost 140 percent, from 42.4 gigawatts to more than 100 gigawatts.^[globalenergymonitor.org/] There are no doubt positive effects from this increase in power generation; for example, electricity may lead to economic and productivity growth [@dinkelman2011effects;@kline2014local;@rud2012electricity;@van2017long], though electrification alone may not be a sufficient condition [@lee2020does].
 
@@
 Pollutants from coal contribute to climate change and the phase out of coal energy is therefore a major climate goal [@iea2022].
-Emissions from the production and burning of coal are also directly harmful to health and the environment, including pollutants like sulfur dioxide, nitrogen oxides, and mercury, to name but a few.^[https://www.eia.gov/energyexplained/coal/coal-and-the-environment.php] While the use of coal had been declining, recent events -- including Russia's invasion of Ukraine -- have led to an increase in the demand for coal.^[https://www.npr.org/2022/08/15/1117560560/a-rising-demand-for-coal-amidst-war-in-ukraine]
+Emissions from the production and burning of coal are also directly harmful to health and the environment, including pollutants like sulfur dioxide, nitrogen oxides, and mercury, to name but a few.^[www.eia.gov/energyexplained/coal/coal-and-the-environment.php] While the use of coal had been declining, recent events -- including Russia's invasion of Ukraine -- have led to an increase in the demand for coal.^[www.npr.org/2022/08/15/1117560560/a-rising-demand-for-coal-amidst-war-in-ukraine]
 
@@
 
-I combine the roll out of coal plants in India with plausibly exogenous changes in wind direction -- in a similar spirit to @deryugina2019mortality -- to measure the effects of pollution from coal plants on key development outcomes.
+Identifying the causal effects of pollution is difficult.
+For starters, the construction of coal plants is likely to be endogenous, though the relationship between construction and economic conditions is not a priori clear.
+On the one hand, governments may locate new plants in fast-growing areas, while on the other hand, they may decide to exact location of plants based on the political power of local citizens [@kopas2020environmental].
+In the case of the data used here, for example, coal plants are constructed in areas with higher population and higher literacy (Table \@ref(tab:plantresultstable)).
+In terms of agriculture, higher productivity predicts more coal plants in a given year as well as the construction of a coal plant over the next decade, indicating possible bias when estimating OLS regressions of agricultural productivity on pollution levels.
+
+To identify the effects of air pollution on agricultural productivity, I instead combine the roll out of coal plants in India with plausibly exogenous changes in wind direction -- in a similar spirit to @deryugina2019mortality -- to measure the effects of pollution from coal plants on key development outcomes.
 I first locate the location of coal plants in India by year.
 Using a village-level shapefile, I identify all villages that are located within 30km of a coal plant at any time during the sample period, which comprises the 1990s to the 2010s, with differences depending on data availability.
-I then download daily wind data and create an identifier for whether a village is located downwind from a coal plant on any given day, aggregating these daily values to different levels depending on the analysis.
-
-After validating that wind direction is a strong predictor of particulate matter, I first document that the total number of days of being downwind from a coal power plant has a negative effect on overall agricultural productivity, using newly available estimates of India-wide agricultural (land) productivity.
+To measure exposure to emissions from coal plants, I calculate the direction from each coal plant to all villages within 30km.
+I then download daily wind data and create an identifier separately for each of the one-hundred-thousand-plus villages for whether a given village is located downwind from a coal plant on any given day.
+I then aggregate this exposure variable based on the desired analysis; for agricultural productivity, for example, I create an exposure variable measuring the total number of exposure days in the five-month agricultural season.
+Consistent with expectations, wind direction is a strong predictor of particulate matter at the monthly level.
+More days being located downwind from a coal power plant leads to higher average concentrations of particulate matter 2.5.
+
+I then document that the total number of days of being downwind from a coal power plant has a negative effect on overall agricultural productivity, using newly available estimates of India-wide agricultural (land) productivity.
+Just a single additional day of being downwind from a coal power plant leads to a decrease in predicted yields of 0.3 percent.
 The within-village absolute deviation of wind exposure is around eight days, meaning that median changes in wind direction are responsible for fluctuations in agricultural productivity of around 2.4 percent, with changes of more than five percent from season to season being common.
-Using wind direction as an instrument for particulate matter, an increase of PM 2.5 by one microgram leads to an even larger decrease in agricultural productivity, at around 20 percent.
+Using wind direction as an instrument for particulate matter, an increase of PM 2.5 by one microgram per cubic meter leads to an even larger decrease in agricultural productivity, at around 20 percent.
 This IV strategy is similar to that in @deryugina2019mortality, though they use general "high-pollution" wind directions while I concentrate only on wind directions from the location of a coal power plant.
+The IV estimates indicate that the effect of pollution is more negative than na\"{i}ve OLS estimates indicate, which is consistent with the evidence on the construction of coal plants.
 
@@
 
-# Data and methods \label{data}
-
-## Data \label{data2}
+# Data and methods 
+\label{data}
+
+## Data 
+\label{data2}
 
@@
 The first set of data lists the location of coal plant across the globe.
-This data comes from Global Energy Monitor^[https://globalenergymonitor.org/projects/global-coal-plant-tracker/] and lists all units generating at least 30 megawatts of electricity.
+This data comes from Global Energy Monitor^[globalenergymonitor.org/projects/global-coal-plant-tracker/] and lists all units generating at least 30 megawatts of electricity.
 The data on coal plants includes the year of opening (and, if applicable, the year of retirement), the GPS (latitude/longitude) location of each plant, and the amount of power produced by the units at each plant.
@@
 plant1 <- ggplot() +
-                theme_minimal() +
-                geom_sf(data = india_shape2) +
-                geom_sf(data = plants1990, size = 0.75) +
+                theme_map() +
+                geom_sf(data = india_shape2, fill = NA, lwd = 0.01) +
+                geom_sf(data = plants1990, size = 0.75, color = "#440154FF") +
                 labs(x = "lon",
@@
 plant2 <- ggplot() +
-                theme_minimal() +
-                geom_sf(data = india_shape2) +
-                geom_sf(data = plants2010, size = 0.75) +
+                theme_map() +
+                geom_sf(data = india_shape2, fill = NA, lwd = 0.01) +
+                geom_sf(data = plants2010, size = 0.75, color = "#440154FF") +
                 labs(x = "lon",
@@
 I keep these definitions when matching across data below.
-This data is also publicly available from Nature's data-sharing website.^[https://springernature.figshare.com] Since the main regressions below use village fixed effects, any differences in cropping patterns across areas should not bias estimates on changes, as long as these patterns do not change markedly in response to pollution.
+This data is also publicly available from Nature's data-sharing website.^[springernature.figshare.com] Since the main regressions below use village fixed effects, any differences in cropping patterns across areas should not bias estimates on changes, as long as these patterns do not change markedly in response to pollution.
 
 The agricultural productivity data is available as raster files with a resolution of 500m.
-To aggregate this data up to a useful administrative unit, I use village-level shapefiles provided by @almn2021 and publicly available on the SHRUG platform.^[https://www.devdatalab.org/shrug_download/] I aggregate the agricultural productivity data to the village level by extracting mean productivity for each feature in the shapefile.
+To aggregate this data up to a useful administrative unit, I use village-level shapefiles provided by @almn2021 and publicly available on the SHRUG platform.^[www.devdatalab.org/shrug_download/] I aggregate the agricultural productivity data to the village level by extracting mean productivity for each feature in the shapefile.
 I do this separately for each season -- monsoon and winter -- and each year, from 2002 to 2013.
@@
 I then calculate the direction from coal plants to all village centroids within that 30km radius.
-To define exposure, I then pull daily wind direction data from the National Center for Atmospheric Research.^[https://climatedataguide.ucar.edu/] For each day, I document whether the wind is blowing towards each village centroid.^[I define "towards" as within five degrees to help take into account that I am using village centroids, using the x and y components of wind speed.] I then temporally aggregate this daily data depending on the temporal definition of the corresponding outcome.
+To define exposure, I then pull daily wind direction data from the National Center for Atmospheric Research.^[climatedataguide.ucar.edu/] For each day, I document whether the wind is blowing towards each village centroid.^[I define "towards" as within five degrees to help take into account that I am using village centroids, using the x and y components of wind speed.] I then temporally aggregate this daily data depending on the temporal definition of the corresponding outcome.
 For example, agricultural productivity is defined across five months (e.g.
@@
 
-```{r windexample, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 8, fig.cap = "Wind direction and aggregation examples (2010-01-01)", fig.footer = list("Note: The left figure shows the average wind direction on January 1st, 2010.
+```{r windexample, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 8, fig.cap = "Wind direction and aggregation examples (2010-01-01)", fig.footer = list("Note: The top figure shows the average wind direction on January 1st, 2010, with zero degrees indicating north.
 The points are the location of coal plants on that date.
-The right figure shows the distribution of pollution exposure in a specific district -- Guna district in Madhya Pradesh -- on the same date.")}
+The bottom figure shows the distribution of pollution exposure in a specific district -- Guna district in Madhya Pradesh -- on the same date.")}
 
 p1 <- ggplot() + 
-        theme_minimal() +
+        theme_map() +
         geom_sf(data = india_shape, fill = NA) +
@@
              y = "lat",
-             subtitle = "Figure A: Wind direction")
+             subtitle = "Figure A: Wind direction") +
+        theme(legend.position = c(0.6, 0.05),
+              legend.direction = "vertical",
+              legend.key.height = unit(0.4, 'cm'),
+              legend.text = element_text(size = 8),
+              legend.background=element_blank() )
 
 p2 <- ggplot() + 
-        theme_minimal() +
+        theme_map() +
         geom_sf(data = districts, fill = NA) +
@@
              y = "lat",
-             subtitle = "Figure B: Aggregation to district")
+             subtitle = "Figure B: Aggregation to district") +
+        theme(legend.position = c(0.5, 0.05),
+              legend.direction = "horizontal",
+              legend.key.width = unit(0.3, 'cm'),
+              legend.text = element_text(size = 8),
+              legend.background=element_blank() )
 
@@
 Unfortunately, I am not aware of any publicly available survey data that covers multiple years and has identifiers below the district level.
-For example, the ARIS-REDS data^[https://www.ncaer.org/project/additional-rural-incomes-survey-aris-rural-economic-demographic-survey-reds] includes GPS identifiers for one year, but that only allows GPS matching for a total of two years: 1999 and 2007/2008.
+For example, the ARIS-REDS data^[www.ncaer.org/project/additional-rural-incomes-survey-aris-rural-economic-demographic-survey-reds] includes GPS identifiers for one year, but that only allows GPS matching for a total of two years: 1999 and 2007/2008.
 The DHS, on the other hand, only has state-level identifiers (except for the newest two waves) and is also not a labor survey.
@@
 
-## Identification \label{identification}
+## Identification 
+\label{identification}
 
@@
 
-## Where do coal plants open?
-
 A large question looms with respect to identification: what are the characteristics of areas near the construction of new coal plants?
@@
 In other words, any selection in the overall location of coal plants would bias against finding negative effects of pollution on agricultural productivity.
-
-However, note that this is not the same as looking at wind direction. I discuss this more in the results section below.
-
-## An example: validating wind direction \label{validation}
-
-Before moving on to the main results, it is worth taking the time to better understand the fixed effects strategy with an example.
+However, note that this is not the same as looking at wind direction, though comparing the OLS and IV results below leads to the same conclusion regarding selection.
+
+It is worth taking the time to validate the use of wind direction as a measure of exposure to pollution.
 Consider the data used for the regressions presented in Table \@ref(tab:pollutiontable).
 The outcome variable is particulate matter -- specifically, $\mathrm{PM_{2.5}}$, which is particulate matter no larger than 2.5 micrometers in diameter -- which comes from @hammer2020global.
-$\mathrm{PM_{2.5}}$ is one of the harmful byproducts from coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.^[https://www.epa.gov/airmarkets/power-plants-and-neighboring-communities] This data is available at the monthly level, so in Table \@ref(tab:pollutiontable), I aggregate wind exposure to the monthly level, as well.
+$\mathrm{PM_{2.5}}$ is one of the harmful byproducts from coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.^[www.epa.gov/airmarkets/power-plants-and-neighboring-communities] This data is available at the monthly level, so in Table \@ref(tab:pollutiontable), I aggregate wind exposure to the monthly level, as well.
 
@@
 Since the exposure variable is at the village level and we follow villages over time, the standard errors are also clustered at the village level.
+In columns two and four, I switch the month fixed effects for district-month fixed effects, which considers only villages in the same district in the same month.
 
@@
 
-# Results \label{results}
+# Results 
+\label{results}
 
@@
 Figure \@ref(fig:laborplot) in the appendix shows this temporal distribution across the five waves of the survey.
+Since I can no longer include district-by-month fixed effects, I instead opt for district-level trends to help take into account aggregate temporal changes in outcomes.
+
+```{r labortable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+# Table
+colnames(labortable) <- c("all", "all", "self", "wage", "farm", "non-farm")
+kable(
+      labortable, caption = "Wind direction and labor allocation",
+      align = "cccccc", booktabs = TRUE, linesep = ""
+      ) %>%
+  footnote(
+           "Note: Standard errors are in parentheses and are clustered at the district level.
+Control variables include female, age, age squared, and (years of) education.
+* p<0.10 ** p<0.05 *** p<0.01",
+           general_title = "",
+           threeparttable = TRUE,
+           footnote_as_chunk = TRUE
+           ) %>%
+  column_spec(1, width = "3cm") %>%
+  column_spec(c(2:7),width = "1.5cm") %>%
+  row_spec(c(8), hline_after = TRUE) %>%
+  row_spec(c(4,7), bold = TRUE) %>%
+  kable_classic_2()
+
+```
 
@@
 
-```{r labortable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-# Table
-colnames(labortable) <- c("all", "all", "self", "wage", "farm", "non-farm")
-kable(
-      labortable, caption = "Wind direction and labor allocation",
-      align = "cccccc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the district level.
-Control variables include female, age, age squared, and (years of) education.
-* p<0.10 ** p<0.05 *** p<0.01",
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
-           ) %>%
-  column_spec(1, width = "3cm") %>%
-  column_spec(c(2:7),width = "1.5cm") %>%
-  row_spec(c(8), hline_after = TRUE) %>%
-  row_spec(c(4,7), bold = TRUE) %>%
-  kable_classic_2()
-
-```
-
 Previous research has shown larger effects of pollution on the health of the oldest [@deryugina2019mortality].
@@
 
-# Conclusion \label{conclusion}
-
-This paper documents the negative effects of pollution on agricultural productivity in India.
+# Conclusion 
+\label{conclusion}
+
+The rapid growth of coal power plants has also led to increases in emissions from the burning of coal.
+In this paper, I show some of the negative effects associated with this burning: decreased agricultural productivity.
 I find clear indications that pollution from coal plants leads to lower agricultural productivity.
 Importantly, I use exogenous changes in wind direction to measure exposure to pollution and show that these effects are robust to a range of alternative specifications.
-Results indicate that these effects may be driven by changes in labor allocation patterns, but I am unable to rule out direct effects on land productivity with the data available.
+
+Results indicate that these effects may be driven by changes in labor allocation patterns.
+In addition, the negative labor effects are concentrated in older adults, which is consistent with previous research on the effects of pollution on health in adults.
+Despite this, I am unable to rule out direct effects on land productivity with the data available.
+In other words,
+
 Finally, these negative effects are also seen in aggregate economic development, with pollution leading to decreases in the growth of nightlights.
+These effects, though small for an individual village, are quite large in overall terms, indicating that pollution from coal plants can have effects on overall economic development.
 

### 2023-01-19 draft/draft.tex
draft

--- 2023-01-17 draft/draft.tex
@@
 \author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
-\date{2023-01-17}
+\date{2023-01-19}
 
@@
 \begin{abstract}
-\noindent The rapid increase in population in India from the 1990s through the 2010s led to the rapid construction of new coal power plants throughout the country.
+\noindent I show that emissions from a common energy source lead to decreased productivity and growth.
+The rapid increase in population in India from the 1990s through the 2010s led to the rapid construction of new coal power plants throughout the country.
 Using this roll-out combined with random changes in wind direction, I document negative effects of pollution from coal power plants on agricultural productivity throughout the country.
@@
 
+Coal is one of the most common energy sources throughout the world.
+In 2021, energy generation from coal reached an all-time high,\footnote{www.iea.org/reports/coal-fired-electricity} driven in part by rapid population growth throughout the world.
+India is no exception to these trends.
 From 1990 to 2010, India's population grew by more than 40 percent.\footnote{data.worldbank.org} This population growth led to a large increase in demand for power.
 While natural gas has become a popular alternative in much of the world, the Indian government met much of the increased demand through the construction of coal power plants.
-During those same two decades, power generation from all coal units of at least 30 megawatts increased by almost 140 percent, from 42.4 gigawatts to more than 100 gigawatts.\footnote{\url{https://globalenergymonitor.org/}} There are no doubt positive effects from this increase in power generation; for example, electricity may lead to economic and productivity growth (Dinkelman 2011; Kline and Moretti 2014; Rud 2012; Van de Walle et al.
+During those same two decades, power generation from all coal units of at least 30 megawatts increased by almost 140 percent, from 42.4 gigawatts to more than 100 gigawatts.\footnote{globalenergymonitor.org/} There are no doubt positive effects from this increase in power generation; for example, electricity may lead to economic and productivity growth (Dinkelman 2011; Kline and Moretti 2014; Rud 2012; Van de Walle et al.
 2017), though electrification alone may not be a sufficient condition (K.
@@
 Pollutants from coal contribute to climate change and the phase out of coal energy is therefore a major climate goal (IEA 2022).
-Emissions from the production and burning of coal are also directly harmful to health and the environment, including pollutants like sulfur dioxide, nitrogen oxides, and mercury, to name but a few.\footnote{\url{https://www.eia.gov/energyexplained/coal/coal-and-the-environment.php}} While the use of coal had been declining, recent events -- including Russia's invasion of Ukraine -- have led to an increase in the demand for coal.\footnote{\url{https://www.npr.org/2022/08/15/1117560560/a-rising-demand-for-coal-amidst-war-in-ukraine}}
+Emissions from the production and burning of coal are also directly harmful to health and the environment, including pollutants like sulfur dioxide, nitrogen oxides, and mercury, to name but a few.\footnote{www.eia.gov/energyexplained/coal/coal-and-the-environment.php} While the use of coal had been declining, recent events -- including Russia's invasion of Ukraine -- have led to an increase in the demand for coal.\footnote{www.npr.org/2022/08/15/1117560560/a-rising-demand-for-coal-amidst-war-in-ukraine}
 
@@
 
-I combine the roll out of coal plants in India with plausibly exogenous changes in wind direction -- in a similar spirit to Deryugina et al.
+Identifying the causal effects of pollution is difficult.
+For starters, the construction of coal plants is likely to be endogenous, though the relationship between construction and economic conditions is not a priori clear.
+On the one hand, governments may locate new plants in fast-growing areas, while on the other hand, they may decide to exact location of plants based on the political power of local citizens (Kopas et al.
+2020).
+In the case of the data used here, for example, coal plants are constructed in areas with higher population and higher literacy (Table \ref{tab:plantresultstable}).
+In terms of agriculture, higher productivity predicts more coal plants in a given year as well as the construction of a coal plant over the next decade, indicating possible bias when estimating OLS regressions of agricultural productivity on pollution levels.
+
+To identify the effects of air pollution on agricultural productivity, I instead combine the roll out of coal plants in India with plausibly exogenous changes in wind direction -- in a similar spirit to Deryugina et al.
 (2019) -- to measure the effects of pollution from coal plants on key development outcomes.
@@
 Using a village-level shapefile, I identify all villages that are located within 30km of a coal plant at any time during the sample period, which comprises the 1990s to the 2010s, with differences depending on data availability.
-I then download daily wind data and create an identifier for whether a village is located downwind from a coal plant on any given day, aggregating these daily values to different levels depending on the analysis.
-
-After validating that wind direction is a strong predictor of particulate matter, I first document that the total number of days of being downwind from a coal power plant has a negative effect on overall agricultural productivity, using newly available estimates of India-wide agricultural (land) productivity.
+To measure exposure to emissions from coal plants, I calculate the direction from each coal plant to all villages within 30km.
+I then download daily wind data and create an identifier separately for each of the one-hundred-thousand-plus villages for whether a given village is located downwind from a coal plant on any given day.
+I then aggregate this exposure variable based on the desired analysis; for agricultural productivity, for example, I create an exposure variable measuring the total number of exposure days in the five-month agricultural season.
+Consistent with expectations, wind direction is a strong predictor of particulate matter at the monthly level.
+More days being located downwind from a coal power plant leads to higher average concentrations of particulate matter 2.5.
+
+I then document that the total number of days of being downwind from a coal power plant has a negative effect on overall agricultural productivity, using newly available estimates of India-wide agricultural (land) productivity.
+Just a single additional day of being downwind from a coal power plant leads to a decrease in predicted yields of 0.3 percent.
 The within-village absolute deviation of wind exposure is around eight days, meaning that median changes in wind direction are responsible for fluctuations in agricultural productivity of around 2.4 percent, with changes of more than five percent from season to season being common.
-Using wind direction as an instrument for particulate matter, an increase of PM 2.5 by one microgram leads to an even larger decrease in agricultural productivity, at around 20 percent.
+Using wind direction as an instrument for particulate matter, an increase of PM 2.5 by one microgram per cubic meter leads to an even larger decrease in agricultural productivity, at around 20 percent.
 This IV strategy is similar to that in Deryugina et al.
 (2019), though they use general ``high-pollution'' wind directions while I concentrate only on wind directions from the location of a coal power plant.
+The IV estimates indicate that the effect of pollution is more negative than na"\{i\}ve OLS estimates indicate, which is consistent with the evidence on the construction of coal plants.
 
@@
 \hypertarget{data-and-methods}{%
-\section{\texorpdfstring{Data and methods \label{data}}{Data and methods }}\label{data-and-methods}}
+\section{Data and methods}\label{data-and-methods}}
+
+\label{data}
 
 \hypertarget{data}{%
-\subsection{\texorpdfstring{Data \label{data2}}{Data }}\label{data}}
+\subsection{Data}\label{data}}
+
+\label{data2}
 
@@
 The first set of data lists the location of coal plant across the globe.
-This data comes from Global Energy Monitor\footnote{\url{https://globalenergymonitor.org/projects/global-coal-plant-tracker/}} and lists all units generating at least 30 megawatts of electricity.
+This data comes from Global Energy Monitor\footnote{globalenergymonitor.org/projects/global-coal-plant-tracker/} and lists all units generating at least 30 megawatts of electricity.
 The data on coal plants includes the year of opening (and, if applicable, the year of retirement), the GPS (latitude/longitude) location of each plant, and the amount of power produced by the units at each plant.
@@
 I keep these definitions when matching across data below.
-This data is also publicly available from Nature's data-sharing website.\footnote{\url{https://springernature.figshare.com}} Since the main regressions below use village fixed effects, any differences in cropping patterns across areas should not bias estimates on changes, as long as these patterns do not change markedly in response to pollution.
+This data is also publicly available from Nature's data-sharing website.\footnote{springernature.figshare.com} Since the main regressions below use village fixed effects, any differences in cropping patterns across areas should not bias estimates on changes, as long as these patterns do not change markedly in response to pollution.
 
@@
 To aggregate this data up to a useful administrative unit, I use village-level shapefiles provided by Asher et al.
-(2021) and publicly available on the SHRUG platform.\footnote{\url{https://www.devdatalab.org/shrug_download/}} I aggregate the agricultural productivity data to the village level by extracting mean productivity for each feature in the shapefile.
+(2021) and publicly available on the SHRUG platform.\footnote{www.devdatalab.org/shrug\_download/} I aggregate the agricultural productivity data to the village level by extracting mean productivity for each feature in the shapefile.
 I do this separately for each season -- monsoon and winter -- and each year, from 2002 to 2013.
@@
 I then calculate the direction from coal plants to all village centroids within that 30km radius.
-To define exposure, I then pull daily wind direction data from the National Center for Atmospheric Research.\footnote{\url{https://climatedataguide.ucar.edu/}} For each day, I document whether the wind is blowing towards each village centroid.\footnote{I define ``towards'' as within five degrees to help take into account that I am using village centroids, using the x and y components of wind speed.} I then temporally aggregate this daily data depending on the temporal definition of the corresponding outcome.
+To define exposure, I then pull daily wind direction data from the National Center for Atmospheric Research.\footnote{climatedataguide.ucar.edu/} For each day, I document whether the wind is blowing towards each village centroid.\footnote{I define ``towards'' as within five degrees to help take into account that I am using village centroids, using the x and y components of wind speed.} I then temporally aggregate this daily data depending on the temporal definition of the corresponding outcome.
 For example, agricultural productivity is defined across five months (e.g.~the monsoon season is from June to October) so I count the total days a given village is exposed to wind blowing from \emph{any} coal plant within 30km during those five months.
@@
 \begin{figure}
-\includegraphics{draft_files/figure-latex/windexample-1} \caption[Wind direction and aggregation examples (2010-01-01)]{Wind direction and aggregation examples (2010-01-01)}\label{fig:windexample}\floatfoot*{Note: The left figure shows the average wind direction on January 1st, 2010.
+\includegraphics{draft_files/figure-latex/windexample-1} \caption[Wind direction and aggregation examples (2010-01-01)]{Wind direction and aggregation examples (2010-01-01)}\label{fig:windexample}\floatfoot*{Note: The top figure shows the average wind direction on January 1st, 2010, with zero degrees indicating north.
 The points are the location of coal plants on that date.
-The right figure shows the distribution of pollution exposure in a specific district -- Guna district in Madhya Pradesh -- on the same date.}
+The bottom figure shows the distribution of pollution exposure in a specific district -- Guna district in Madhya Pradesh -- on the same date.}
 \end{figure}
@@
 Unfortunately, I am not aware of any publicly available survey data that covers multiple years and has identifiers below the district level.
-For example, the ARIS-REDS data\footnote{\url{https://www.ncaer.org/project/additional-rural-incomes-survey-aris-rural-economic-demographic-survey-reds}} includes GPS identifiers for one year, but that only allows GPS matching for a total of two years: 1999 and 2007/2008.
+For example, the ARIS-REDS data\footnote{www.ncaer.org/project/additional-rural-incomes-survey-aris-rural-economic-demographic-survey-reds} includes GPS identifiers for one year, but that only allows GPS matching for a total of two years: 1999 and 2007/2008.
 The DHS, on the other hand, only has state-level identifiers (except for the newest two waves) and is also not a labor survey.
@@
 \hypertarget{identification}{%
-\subsection{\texorpdfstring{Identification \label{identification}}{Identification }}\label{identification}}
+\subsection{Identification}\label{identification}}
+
+\label{identification}
 
@@
 I also present results using leads to show that future wind direction does not seem to predict current outcomes.
-
-\hypertarget{where-do-coal-plants-open}{%
-\subsection{Where do coal plants open?}\label{where-do-coal-plants-open}}
 
@@
 In other words, any selection in the overall location of coal plants would bias against finding negative effects of pollution on agricultural productivity.
-
-However, note that this is not the same as looking at wind direction. I discuss this more in the results section below.
-
-\hypertarget{an-example-validating-wind-direction}{%
-\subsection{\texorpdfstring{An example: validating wind direction \label{validation}}{An example: validating wind direction }}\label{an-example-validating-wind-direction}}
-
-Before moving on to the main results, it is worth taking the time to better understand the fixed effects strategy with an example.
+However, note that this is not the same as looking at wind direction, though comparing the OLS and IV results below leads to the same conclusion regarding selection.
+
+It is worth taking the time to validate the use of wind direction as a measure of exposure to pollution.
 Consider the data used for the regressions presented in Table \ref{tab:pollutiontable}.
@@
 (2020).
-\(\mathrm{PM_{2.5}}\) is one of the harmful byproducts from coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.\footnote{\url{https://www.epa.gov/airmarkets/power-plants-and-neighboring-communities}} This data is available at the monthly level, so in Table \ref{tab:pollutiontable}, I aggregate wind exposure to the monthly level, as well.
+\(\mathrm{PM_{2.5}}\) is one of the harmful byproducts from coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.\footnote{www.epa.gov/airmarkets/power-plants-and-neighboring-communities} This data is available at the monthly level, so in Table \ref{tab:pollutiontable}, I aggregate wind exposure to the monthly level, as well.
 
@@
 \hypertarget{results}{%
-\section{\texorpdfstring{Results \label{results}}{Results }}\label{results}}
+\section{Results}\label{results}}
+
+\label{results}
 
@@
 Figure \ref{fig:laborplot} in the appendix shows this temporal distribution across the five waves of the survey.
-
-There are some interesting results in Table \ref{tab:labortable}.
-The first two columns show a marginally significant decrease in overall labor allocation (where ``labor'' is defined as non-domestic self and wage employment).
-The third and fourth columns show that this overall decrease is predominantly from wage employment, not self employment.
-An additional day of wind over the last seven decreases wage employment by 0.034 days in the district.
-Mean wage employment is 1.458 days, meaning one day of wind decreases wage employment by 2.3 percent of the mean.
-Since much of the district is not covered by ``treated'' villages, this is actually a rather substantive effect on the villages involved.
-
-The last two columns split the effects by sector, into farm and non-farm employment.
-Although agricultural (land) productivity decreases when wind blows in the direction of villages, overall labor allocation to farm labor actually goes \emph{up}, by around 4.5 percent of the mean of 0.784 days.
-Non-farm labor, on the other hand, goes down, by around 2.4 percent of the mean of 2.527 days.
-
-There are two possible explanations for this pattern of results.
-First, it could be that land productivity goes down, leading farmers to increase their farm labor allocation in an attempt to compensate for the decreased productivity.
-A second possibility is that pollution does not affect land productivity at all, at least not directly.
-In this scenario, the worsening air quality could lead to decreased labor productivity, meaning that ``effective labor'' essentially decreases when pollution is higher.
-Unfortunately, I am not able to disentangle these two possiblities with the data available.
-
-Table \ref{tab:labortablemonsoon} and Table \ref{tab:labortablewinter} in the appendix disaggregate these effects into the monsoon and winter seasons.
-Apparently, the entirety of the effect is driven by changes in labor allocation during the monsoon season.
-The dynamics across seasons apparently lead to large differences in the effects of pollution on labor allocation.
+Since I can no longer include district-by-month fixed effects, I instead opt for district-level trends to help take into account aggregate temporal changes in outcomes.
 
@@
 \end{table}
+
+There are some interesting results in Table \ref{tab:labortable}.
+The first two columns show a marginally significant decrease in overall labor allocation (where ``labor'' is defined as non-domestic self and wage employment).
+The third and fourth columns show that this overall decrease is predominantly from wage employment, not self employment.
+An additional day of wind over the last seven decreases wage employment by 0.034 days in the district.
+Mean wage employment is 1.458 days, meaning one day of wind decreases wage employment by 2.3 percent of the mean.
+Since much of the district is not covered by ``treated'' villages, this is actually a rather substantive effect on the villages involved.
+
+The last two columns split the effects by sector, into farm and non-farm employment.
+Although agricultural (land) productivity decreases when wind blows in the direction of villages, overall labor allocation to farm labor actually goes \emph{up}, by around 4.5 percent of the mean of 0.784 days.
+Non-farm labor, on the other hand, goes down, by around 2.4 percent of the mean of 2.527 days.
+
+There are two possible explanations for this pattern of results.
+First, it could be that land productivity goes down, leading farmers to increase their farm labor allocation in an attempt to compensate for the decreased productivity.
+A second possibility is that pollution does not affect land productivity at all, at least not directly.
+In this scenario, the worsening air quality could lead to decreased labor productivity, meaning that ``effective labor'' essentially decreases when pollution is higher.
+Unfortunately, I am not able to disentangle these two possiblities with the data available.
+
+Table \ref{tab:labortablemonsoon} and Table \ref{tab:labortablewinter} in the appendix disaggregate these effects into the monsoon and winter seasons.
+Apparently, the entirety of the effect is driven by changes in labor allocation during the monsoon season.
+The dynamics across seasons apparently lead to large differences in the effects of pollution on labor allocation.
 
@@
 While the youngest may be able to readjust and increase their overall self-employment labor allocation, perhaps the oldest cannot due to, for example, health effects of pollution, which are well documented (Bentayeb et al.
-2012; Pope 3rd 2000).\footnote{Much of the literature also suggests the youngest are at significant risk of mortality and morbidities.
+2012; Pope III 2000).\footnote{Much of the literature also suggests the youngest are at significant risk of mortality and morbidities.
 However, given my focus on labor allocation, I only discuss older adults here.}
@@
 \hypertarget{conclusion}{%
-\section{\texorpdfstring{Conclusion \label{conclusion}}{Conclusion }}\label{conclusion}}
-
-This paper documents the negative effects of pollution on agricultural productivity in India.
+\section{Conclusion}\label{conclusion}}
+
+\label{conclusion}
+
+The rapid growth of coal power plants has also led to increases in emissions from the burning of coal.
+In this paper, I show some of the negative effects associated with this burning: decreased agricultural productivity.
 I find clear indications that pollution from coal plants leads to lower agricultural productivity.
 Importantly, I use exogenous changes in wind direction to measure exposure to pollution and show that these effects are robust to a range of alternative specifications.
-Results indicate that these effects may be driven by changes in labor allocation patterns, but I am unable to rule out direct effects on land productivity with the data available.
+
+Results indicate that these effects may be driven by changes in labor allocation patterns.
+In addition, the negative labor effects are concentrated in older adults, which is consistent with previous research on the effects of pollution on health in adults.
+Despite this, I am unable to rule out direct effects on land productivity with the data available.
+In other words,
+
 Finally, these negative effects are also seen in aggregate economic development, with pollution leading to decreases in the growth of nightlights.
+These effects, though small for an individual village, are quite large in overall terms, indicating that pollution from coal plants can have effects on overall economic development.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-kopas2020environmental}{}}%
+Kopas, Jacob, Erin York, Xiaomeng Jin, SP Harish, Ryan Kennedy, Shiran Victoria Shen, and Johannes Urpelainen.
+2020.
+{``{Environmental justice in India: incidence of air pollution from coal-fired power plants}.''} \emph{{Ecological Economics}} 176: 106711.
+
 \leavevmode\vadjust pre{\hypertarget{ref-lee2022valid}{}}%
@@
 \leavevmode\vadjust pre{\hypertarget{ref-pope2000epidemiology}{}}%
-Pope 3rd, CA. 2000. {``Epidemiology of Fine Particulate Air Pollution and Human Health: Biologic Mechanisms and Who's at Risk?''} \emph{{Environmental Health Perspectives}} 108 (suppl 4): 713--23.
+Pope III, C Arden.
+2000.
+{``Epidemiology of Fine Particulate Air Pollution and Human Health: Biologic Mechanisms and Who's at Risk?''} \emph{{Environmental Health Perspectives}} 108 (suppl 4): 713--23.
 

### 2023-01-19 analyses/regressions.R

@@ -6,6 +6,7 @@ rm(list = ls())
 
 library(tidyverse)
 library(lubridate)
+library(viridis)
 library(fixest)
 library(sf)
 
@@ -99,10 +100,35 @@ villages_overlap <- villages_overlap %>% mutate(shrid = paste0(pc11_s_id, "-", p
 pollution <- pollution %>% left_join(villages_overlap, by = "shrid")
 
 pollution <- pollution %>% group_by(shrid) %>%
-                              mutate(pm_abs_dev = abs(mean - mean(mean))) %>%
+                              mutate(pm_abs_dev = abs(mean - mean(mean)),
+                                     pm_dev = (mean - mean(mean)),
+                                     wind_dev = (wind - mean(wind))) %>%
+                              ungroup()
+pollution$month_int <- month(pollution$date)
+pollution <- pollution %>% group_by(shrid, month_int) %>%
+                              mutate(m_pm_dev = (mean - mean(mean)),
+                                     m_wind_dev = (wind - mean(wind))) %>%
                               ungroup()
 summary(pollution$pm_abs_dev)
-# mean 21.1
+summary(pollution$wind_dev)
+summary(pollution$m_wind_dev)
+summary(pollution$wind)
+
+
+
+# figures
+# Random 10 percent sample
+set.seed(12350)
+ggplot(data = pollution[sample(1:nrow(pollution), nrow(pollution)/10, replace = F),]) +
+      geom_smooth(aes(x = m_wind_dev, y = m_pm_dev), color = "#440154FF", se = FALSE, alpha = 0.75) +
+      labs(x = "wind",
+           y = "PM") +
+      theme_minimal()
+# won't save as RDS because it's simply too large
+ggsave("pollution_development/draft/tables/devplot.png")
+
+
+
 
 pol1 <- feols(mean ~ wind | village + month, data = pollution, cluster = c("village"))
 pol2 <- feols(mean ~ wind | village + month^distfe, data = pollution, cluster = c("village"))
@@ -439,7 +465,7 @@ df <- df %>% left_join(pollution, by = c("state_merge", "district_merge", "year"
 laborplot <- ggplot(data = df) +
                 geom_histogram(aes(x = date), binwidth = 7, color = "gray") + 
                 labs(x = "date",
-                     y = "households interviewed")
+                     y = "households interviewed") +
                 theme_minimal()
 saveRDS(laborplot, "pollution_development/draft/tables/laborplot.rds")
 

### 2023-01-20 draft/draft.tex
more changes to draft

@@
 
-\title{Pollution, agricultural productivity, and development: Evidence from coal in plants in India\footnote{All scripts used in this paper are publicly available on my GitHub page: github.com/JoshMerfeld/pollution\_development}}
+\title{Pollution, agricultural productivity, and development: Evidence from coal in plants in India\footnote{I thank Marc Bellemare and David Sungho Park for comments and suggestions that greatly improved the quality of the paper.
+The usual disclaimer applies.
+All scripts used in this paper are publicly available on my GitHub page: github.com/JoshMerfeld/pollution\_development}}
 \author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
-\date{2023-01-19}
+\date{2023-01-20}
 
@@
 Using this roll-out combined with random changes in wind direction, I document negative effects of pollution from coal power plants on agricultural productivity throughout the country.
-Results suggest this decrease may be driven by a decline in labor allocation of the oldest half of my sample, though I cannot rule out direct effects of pollution on land productivity.
+Results suggest this decrease may be driven by a decline in labor allocation of the oldest half of my sample, who greatly decrease labor allocated to both farm and non-farm employment when exposed to pollution.
 These impacts spill over into aggregate economic growth, as well; nightlight growth is slower following years of higher exposure to pollution.\\
@@
 Coal is one of the most common energy sources throughout the world.
-In 2021, energy generation from coal reached an all-time high,\footnote{www.iea.org/reports/coal-fired-electricity} driven in part by rapid population growth throughout the world.
+In 2021, energy generation from coal reached an all-time high,\footnote{www.iea.org/reports/coal-fired-electricity} driven in part by rapid population growth and rising energy prices.
 India is no exception to these trends.
@@
 While natural gas has become a popular alternative in much of the world, the Indian government met much of the increased demand through the construction of coal power plants.
-During those same two decades, power generation from all coal units of at least 30 megawatts increased by almost 140 percent, from 42.4 gigawatts to more than 100 gigawatts.\footnote{globalenergymonitor.org/} There are no doubt positive effects from this increase in power generation; for example, electricity may lead to economic and productivity growth (Dinkelman 2011; Kline and Moretti 2014; Rud 2012; Van de Walle et al.
+During those same two decades, power generation from all coal units of at least 30 megawatts in the country increased by almost 140 percent, from 42.4 gigawatts to more than 100 gigawatts.\footnote{globalenergymonitor.org/} There are no doubt positive effects from this increase in power generation; for example, electricity may lead to economic and productivity growth (Dinkelman 2011; Kline and Moretti 2014; Rud 2012; Van de Walle et al.
 2017), though electrification alone may not be a sufficient condition (K.
@@
 
-In this paper, I study the effects of emissions from coal power plants on agricultural productivity, labor allocation, and economic growth in India.
+I study the effects of emissions from coal power plants on agricultural productivity, labor allocation, and economic growth in India.
 Previous work has suggested that there may be harmful direct effects of air pollution (Heck et al.
@@
 1997) and water pollution (Reddy and Behera 2006) on agricultural productivity, including work showing that gold mines -- which release large amounts of pollutants into the surrounding environment -- lead to lower agricultural productivity in Ghana (Aragón and Rud 2016).
-This paper, however, tackles the broader case of coal plants, which are a leading source of power throughout the world.
-I show that air pollution from these plants can lead to decreased agricultural productivity.
+These effects can work through the well known effects on human health and productivity, but also directly through land productivity.
+Black carbon in soot, for example, can directly affect solar radiation reaching plants on the surface (Ramanathan and Carmichael 2008; Burney and Ramanathan 2014).
+Relative to previous work in economics, this paper tackles the broader case of coal plants, which are a leading source of power throughout the world.
+I show that air pollution from these plants can lead to lower agricultural productivity and even decreased economic growth.
 
@@
 For starters, the construction of coal plants is likely to be endogenous, though the relationship between construction and economic conditions is not a priori clear.
-On the one hand, governments may locate new plants in fast-growing areas, while on the other hand, they may decide to exact location of plants based on the political power of local citizens (Kopas et al.
+On the one hand, governments may locate new plants in fast-growing areas, while on the other hand, they may decide the exact location of plants based on the political power of local citizens (Kopas et al.
 2020).
-In the case of the data used here, for example, coal plants are constructed in areas with higher population and higher literacy (Table \ref{tab:plantresultstable}).
+In the case of the data used here, coal plants are constructed in areas with higher population and higher literacy (Table \ref{tab:plantresultstable}).
 In terms of agriculture, higher productivity predicts more coal plants in a given year as well as the construction of a coal plant over the next decade, indicating possible bias when estimating OLS regressions of agricultural productivity on pollution levels.
@@
 (2019) -- to measure the effects of pollution from coal plants on key development outcomes.
-I first locate the location of coal plants in India by year.
+I first plot the location of coal plants in India by year.
 Using a village-level shapefile, I identify all villages that are located within 30km of a coal plant at any time during the sample period, which comprises the 1990s to the 2010s, with differences depending on data availability.
 To measure exposure to emissions from coal plants, I calculate the direction from each coal plant to all villages within 30km.
-I then download daily wind data and create an identifier separately for each of the one-hundred-thousand-plus villages for whether a given village is located downwind from a coal plant on any given day.
-I then aggregate this exposure variable based on the desired analysis; for agricultural productivity, for example, I create an exposure variable measuring the total number of exposure days in the five-month agricultural season.
+Using daily wind data and these spatial variables, I create an identifier separately for each of the one-hundred-thousand-plus villages for whether a given village is located downwind from a coal plant on any given day.
+I then aggregate this exposure variable based on the desired analysis; for agricultural productivity, for example, I create an exposure variable measuring the total number of exposure days in the five-month agricultural season, separately for the main monsoon season and the smaller winter season.
 Consistent with expectations, wind direction is a strong predictor of particulate matter at the monthly level.
@@
 
-I then document that the total number of days of being downwind from a coal power plant has a negative effect on overall agricultural productivity, using newly available estimates of India-wide agricultural (land) productivity.
+Using newly available estimates of India-wide agricultural (land) productivity, results show that the total number of days of being downwind from a coal power plant has a negative effect on overall agricultural productivity.
 Just a single additional day of being downwind from a coal power plant leads to a decrease in predicted yields of 0.3 percent.
@@
 (2019), though they use general ``high-pollution'' wind directions while I concentrate only on wind directions from the location of a coal power plant.
-The IV estimates indicate that the effect of pollution is more negative than na"\{i\}ve OLS estimates indicate, which is consistent with the evidence on the construction of coal plants.
+The IV estimates indicate that the effect of pollution is more negative than naïve OLS estimates indicate, which is consistent with the evidence on the construction of coal plants; higher levels of pollution are correlated with higher levels of agricultural productivity and economic growth, a relationship that the IV strategy adjusts downwards.
 
@@
 I also show that the effects are largest in villages with the highest initial agricultural productivity and that pollution and rainfall shocks are compounding; higher exposure along with less rainfall increases the negative effect of either shock individually.
+As weather shocks become more common across the globe, the rise of pollution in India may be particularly problematic for farmers.
 
@@
 An additional day of wind over the previous week decreases overall labor allocation by 0.027 days.
-However, all effects on labor allocation are driven by the older half of my sample; these adults see much large decreases in overall labor allocation, self employment, farm employment, and non-farm employment.
-This is suggestive evidence of productivity effects being driven by the health effects on individuals, though I am not able to rule out direct effects on land productivity with my data.
+However, the overall results show stark heterogeneity based on age; the older half of the sample greatly decreases labor allocation in response to pollution exposure while the younger half increases overall labor allocation.
+Older adults decrease labor allocation to both farm and non-farm labor, while younger adults increase labor allocated to farm labor.
+This is suggestive evidence of productivity effects being driven by the health and productivity effects on individuals, since older adults are more susceptible to pollution and they also greatly decrease non-farm employment, which presumably will no suffer any possible direct effects of pollution on productivity.
+The preponderance of results seems to indicate that effects on labor are the most likely mechanism.
 
@@
 While the effect for a single village might be small, the overall, aggregate effect on growth is quite substantial.
+To put this number in perspective, Castelló-Climent, Chaudhary, and Mukhopadhyay (2018) argue that a standard-deviation increase in the share of adults with a university education leads to 12 percent higher nightlights in India, while Storeygard (2016) shows that a four-fold increase in oil prices over six years leads to seven percent lower nightlights at a distance of around 500km from the nearest large port in sub-Saharan Africa.
+In other words, the 0.7 percent change in nightlights is quite large given this is annual variation.
 
@@
 Perhaps relevant to my labor allocation results, Hanna and Oliva (2015) show that decreases in pollution lead to increases in labor allocation, while I show the same relationship among the oldest cohort.
-The paper most related to this one is Aragón and Rud (2016), who show that gold mines in Ghana have substantial negative impacts on agricultural productivity in the surrounding areas.
+The papers most related to this one are Aragón and Rud (2016), who show that gold mines in Ghana have substantial negative impacts on agricultural productivity in the surrounding areas, and Burney and Ramanathan (2014), who model the effects of climate change and pollutants on agricultural productivity in India.
 This paper differs in two important ways.
-First, I explore effects of pollution from coal plants, which are much more ubiquitous than goald mines across the world.
-Second, I use a different identification strategy, relying on wind blowing pollution from coal plants.
+First, I explore effects of pollution from coal plants, which are much more ubiquitous than gold mines across the world.
+Second, I use a different identification strategy from either paper, relying on wind blowing pollution from coal plants.
 
@@
 I first discuss data and methodology.
-I break this section up into \autoref{data2}, which covers the multiple sources of data used in this paper; \autoref{identification}, which discusses the empirical strategy; and \autoref{validation}, which validates the wind direction variable using data on particulate matter.
+I break this section up into \autoref{data2}, which covers the multiple sources of data used in this paper; \autoref{identification}, which discusses the empirical strategy and validates the wind direction variable using data on particulate matter.
 I present the main results in \autoref{results} before concluding in \autoref{conclusion}.
@@
 
+In this section, I describe the different data sources and the identification strategy.
+I also spend time validating the use of wind direction as a proxy for particulate matter as well as looking at the correlates of construction of coal power plants.
+
 \hypertarget{data}{%
@@
 
-The main goal of this paper is to examine whether exposure to exhaust from coal plants affects agricultural productivity, labor allocation, and economic growth.
+The main goal of this paper is to examine whether exposure to emissions from coal plants affects agricultural productivity, labor allocation, and economic growth.
 I use several sources of data, which I describe here and in Table \ref{tab:data} in the appendix.
@@
 
-Figure \ref{fig:plants} presents the location of coal plants in India for two specific years: 1990 and 2010.
+Figure \ref{fig:plants} displays the location of coal plants in India for two specific years: 1990 and 2010.
 There is a clear increase in the prevalence of coal plants across the country over the two decades.
-Additionally, the overall capacity from coal plants increased from 42.4 gigawatts to 100.4 gigawats, an increase of 136.6 percent in just 20 years.
-This increase is driven both by the rollout of new plants as well as the construction of new units at existing plants, the latter of which is not shown in Figure \ref{fig:plants}.
+Additionally, the overall capacity from coal plants increased from 42.4 gigawatts to 100.4 gigawatts, an increase of 136.6 percent in just 20 years.
+This increase is driven both by the roll out of new plants as well as the construction of new units at existing plants, the latter of which is not shown in Figure \ref{fig:plants}.
 
@@
 This data comes from Gangopadhyay et al.
-(2022) and estimates land productivity (i.e.~yield, in tons per hectare) for the major crops in India.
-They define the monsoon season as June to October and the winter season as November to March.
+(2022) and estimates land productivity (i.e.~yield, in tons per hectare) for the major crops in India, using satellite-derived data.
+While there is a wealth of literature on measurement error in land (Desiere and Jolliffe 2018; Abay et al.
+2022), the authors use a static cropland mask for the productivity estimates.
+This means that any measurement error will be constant across time and subsumed by the village fixed effects.
+While the measure of output is also noisy, the use of satellite-derived data and a static cropland mask reduces concerns about measurement error being responsible for the results below.
+The authors define the monsoon season as June to October and the winter season as November to March.
 I keep these definitions when matching across data below.
@@
 I also present results using leads to show that future wind direction does not seem to predict current outcomes.
+
+The inclusion of rainfall and temperature is meant to reduce concerns that wind direction may not affect agricultural productivity but, instead, be correlated with it through changes in weather.
+For example, prevailing winds from one direction may be associated with dry weather, while winds from another direction may be more likely to carry moisture and rain.
+It is not a priori clear how this would bias the estimates with respect to pollution and coal plants, however.
+The differences in weather patterns is very context specific and can differ even for villages located near one another.
+Nonetheless, adding rainfall and temperature should alleviate some of these concerns.
 
@@
 To recover the actual effect of pollution, we need an instrument.
-If you believe that wind direction is exogenous and, conditional on the fixed effects and weather variables, only affects agricultural productivity through pollution -- which seems reasonable since it is hard to imagine what other variables it might affect -- then it is a valid instrument for pollution.
+If, conditional on the fixed effects and weather variables, wind direction is exogenous and only affects agricultural productivity through pollution -- which seems reasonable since it is hard to imagine what other variables it might affect -- then it is a valid instrument for pollution.
 
 I present the IV results in Table \ref{tab:yieldtabletwo}, with the second-stage results in the top panel and the first-stage results in the bottom panel.
-The first-stage results make clear how strong of an instrument wind is; the t-statistic in column one, for example, is almost 48 while it is still around 11 in column three, meeting even the higher bar for the first-stage F-test argued for by D.
+The first-stage results make clear how strong of an instrument wind is; the t-statistic in column one, for example, is almost 48 while it is still well above 10 in column three, which translates to a first-stage F-statistic of 195, meeting even the higher bar for the first-stage F-test argued for by D.
 S.
@@
 In this scenario, the worsening air quality could lead to decreased labor productivity, meaning that ``effective labor'' essentially decreases when pollution is higher.
-Unfortunately, I am not able to disentangle these two possiblities with the data available.
 
@@
 
+One way to provide suggestive evidence of mechanisms is to explore heterogeneity in the effects.
 Previous research has shown larger effects of pollution on the health of the oldest (Deryugina et al.
@@
 
+To put some of these changes in context, an additional day of exposure to pollution from wind increases self-employment for the younger half of the sample by around six percent of the mean and increases farm employment for the younger half by almost 10 percent of the mean.
+For the older half, on the other hand, an additional day of exposure decreases self-employment and non-farm employment by around five percent of their means.
+
 While speculative, one possible explanation for this pattern is a decrease in labor productivity.
@@
 2012; Pope III 2000).\footnote{Much of the literature also suggests the youngest are at significant risk of mortality and morbidities.
-However, given my focus on labor allocation, I only discuss older adults here.}
+However, given my focus on labor allocation, I only discuss older adults here.} We see decreases in overall agricultural productivity, despite an increase in overall farm labor.
+We also see decreases in non-farm employment, indicating that individuals are reallocating labor from the non-farm sector to the farm sector.
+
+It is worth noting that the larger negative effects for older adults is more consistent with a labor story than a land story; it is hard to imagine these patterns emerging if the only effects were through land productivity changes.
+The overall decrease in non-farm employment is more compatible with a labor health/productivity story, though with general equilibrium effects in agriculture spillover over into the non-farm sector, this need not necessarily be the case.
+
+In addition, Table \ref{tab:labortableoldstate} in the appendix presents one additional set of results to support a labor story.
+Changes in land productivity should be relatively fixed in a short period of time.
+For example, if we focus on just a single month in a single district, the overall effects driven by changes in land productivity should be constant even though pollution may have differential effects throughout the month on labor productivity, which responds more immediately to changes in ar quality.
+Table \ref{tab:labortableoldstate} changes the specification to include district-by-year-by-month fixed effects.
+In other words, the results compare people in the same district in the same month, but who may have differential exposure to pollution simply because they were interviewed at a slightly different time.
+
+The overall pattern of results stays the same when we shift the focus.
+Overall labor allocation increases for the youngest, but decreases for the oldest, and this is driven completely by self employment.
+There are decreases in both farm and non-farm employment for the oldest, while the youngest show increases in non-farm employment, but not farm employment -- which is a change from the previous results.
+The overall story remains the same, however, which is further evidence of effects being driven primarily through changes in labor allocation and productivity.
+
+Finally, Table \ref{tab:wagestable} in the appendix shows that there are no significant changes in daily wages, though the overall direction of wage changes would be ambiguous, anyway.
+While reduced productivity might lead to lower wages, decreased labor supply could also put upward pressure on wages.
 
@@
 Despite this, I am unable to rule out direct effects on land productivity with the data available.
-In other words,
 
@@
 
-These results suggest caution in building new power plants that emit substantial amounts of pollution.
-While the building of additional power generation facilities may be an economic necessity, the location of these facilities may be particular important for multiple aspects of economic activity, not just health and mortality.
+These results, of course, do not address possible positive effects of the building of additional power plants, which can lead to increased electricity access.
+As such, the main conclusion is that the location of new power generation facilities may be particularly important for multiple aspects of economic activity, not just health and mortality.
 
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-abay2022nonclassical}{}}%
+Abay, Kibrom A, Christopher B Barrett, Talip Kilic, Heather Moylan, John Ilukor, and Wilbert Drazi Vundru.
+2022.
+{``{Nonclassical Measurement Error and Farmers?
+Response to Information Reveal Behavioral Anomalies}.''}
+
 \leavevmode\vadjust pre{\hypertarget{ref-aragon2016polluting}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-burney2014recent}{}}%
+Burney, Jennifer, and V Ramanathan. 2014. {``{Recent climate and air pollution impacts on Indian agriculture}.''} \emph{{Proceedings of the National Academy of Sciences}} 111 (46): 16319--24.
+
+\leavevmode\vadjust pre{\hypertarget{ref-castello2018higher}{}}%
+Castelló-Climent, Amparo, Latika Chaudhary, and Abhiroop Mukhopadhyay.
+2018.
+{``{Higher education and prosperity: From Catholic missionaries to luminosity in India}.''} \emph{{The Economic Journal}} 128 (616): 3039--75.
+
 \leavevmode\vadjust pre{\hypertarget{ref-chang2019effect}{}}%
@@
 {``{The mortality and medical costs of air pollution: Evidence from changes in wind direction}.''} \emph{{American Economic Review}} 109 (12): 4178--4219.
+
+\leavevmode\vadjust pre{\hypertarget{ref-desiere2018land}{}}%
+Desiere, Sam, and Dean Jolliffe. 2018. {``{Land productivity and plot size: Is measurement error driving the inverse relationship?}''} \emph{{Journal of Development Economics}} 130: 84--98.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-ramanathan2008global}{}}%
+Ramanathan, Veerabhadran, and Gregory Carmichael. 2008. {``{Global and regional climate changes due to black carbon}.''} \emph{{Nature Geoscience}} 1 (4): 221--27.
+
 \leavevmode\vadjust pre{\hypertarget{ref-reddy2006impact}{}}%
@@
 Rud, Juan Pablo. 2012. {``{Electricity provision and industrial development: Evidence from India}.''} \emph{{Journal of Development Economics}} 97 (2): 352--67.
+
+\leavevmode\vadjust pre{\hypertarget{ref-storeygard2016farther}{}}%
+Storeygard, Adam. 2016. {``{Farther on down the road: transport costs, trade and urban growth in sub-Saharan Africa}.''} \emph{{The Review of Economic Studies}} 83 (3): 1263--95.
 
@@
 
+\FloatBarrier
 \newpage
@@
 
+\FloatBarrier
+\newpage
+\begin{table}
+
+\caption{\label{tab:labortableoldstate}Wind direction and labor allocation, heterogeneity by age}
+\centering
+\begin{threeparttable}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}}
+\toprule
+  & all & self & wage & farm & non-farm\\
+\midrule
+wind & 0.100*** & 0.117*** & -0.018 & 0.026 & 0.074***\\
+ & (0.027) & (0.029) & (0.026) & (0.024) & (0.028)\\
+wind times age>=38 & -0.203*** & -0.201*** & -0.002 & -0.070*** & -0.133***\\
+ & (0.037) & (0.032) & (0.039) & (0.018) & (0.031)\\
+controls & Yes & Yes & Yes & Yes & Yes\\
+\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+district-year-month & Yes & Yes & Yes & Yes & Yes\\
+\midrule
+observations & 898,856 & 898,856 & 898,856 & 898,856 & 898,856\\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}[para]
+\item Note: Standard errors are in parentheses and are clustered at the district level. Control variables include female, age, age squared, and (years of) education. * p<0.10 ** p<0.05 *** p<0.01
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+
+\FloatBarrier
+\newpage
+\begin{table}
+
+\caption{\label{tab:wagestable}Wind direction and wages}
+\centering
+\begin{threeparttable}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}}
+\toprule
+  & farm & non-farm & all\\
+\midrule
+wind & 0.037 & 0.010 & 0.009\\
+ & (0.031) & (0.010) & (0.011)\\
+controls & Yes & Yes & Yes\\
+\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{}\\
+district & Yes & Yes & Yes\\
+year & Yes & Yes & Yes\\
+\textbf{varying slopes:} & \textbf{} & \textbf{} & \textbf{}\\
+year (by district) & Yes & Yes & Yes\\
+\midrule
+observations & 16,534 & 149,973 & 133,526\\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}[para]
+\item Note: Standard errors are in parentheses and are clustered at the district level. Control variables include female, age, age squared, and (years of) education. * p<0.10 ** p<0.05 *** p<0.01
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+
 \end{document}

### 2023-01-20 draft/draft.Rmd
changes

@@
 ---
-title: "Pollution, agricultural productivity, and development: Evidence from coal in plants in India^[All scripts used in this paper are publicly available on my GitHub page: github.com/JoshMerfeld/pollution_development]"
+title: "Pollution, agricultural productivity, and development: Evidence from coal in plants in India^[I thank Marc Bellemare and David Sungho Park for comments and suggestions that greatly improved the quality of the paper.
+The usual disclaimer applies.
+All scripts used in this paper are publicly available on my GitHub page: github.com/JoshMerfeld/pollution_development]"
 author:
@@
 Using this roll-out combined with random changes in wind direction, I document negative effects of pollution from coal power plants on agricultural productivity throughout the country.
-Results suggest this decrease may be driven by a decline in labor allocation of the oldest half of my sample, though I cannot rule out direct effects of pollution on land productivity.
+Results suggest this decrease may be driven by a decline in labor allocation of the oldest half of my sample, who greatly decrease labor allocated to both farm and non-farm employment when exposed to pollution.
 These impacts spill over into aggregate economic growth, as well; nightlight growth is slower following years of higher exposure to pollution.
@@
 labortable <- readRDS("pollution_development/draft/tables/labortable.rds")
+wagestable <- readRDS("pollution_development/draft/tables/wagestable.rds")
 labortablemonsoon <- readRDS("pollution_development/draft/tables/labortablemonsoon.rds")
@@
 labortableold <- readRDS("pollution_development/draft/tables/labortableold.rds")
+labortableoldstate <- readRDS("pollution_development/draft/tables/labortableoldstate.rds")
 ntltable <- readRDS("pollution_development/draft/tables/ntltable.rds")
@@
 Coal is one of the most common energy sources throughout the world.
-In 2021, energy generation from coal reached an all-time high,^[www.iea.org/reports/coal-fired-electricity] driven in part by rapid population growth throughout the world.
+In 2021, energy generation from coal reached an all-time high,^[www.iea.org/reports/coal-fired-electricity] driven in part by rapid population growth and rising energy prices.
 India is no exception to these trends.
@@
 While natural gas has become a popular alternative in much of the world, the Indian government met much of the increased demand through the construction of coal power plants.
-During those same two decades, power generation from all coal units of at least 30 megawatts increased by almost 140 percent, from 42.4 gigawatts to more than 100 gigawatts.^[globalenergymonitor.org/] There are no doubt positive effects from this increase in power generation; for example, electricity may lead to economic and productivity growth [@dinkelman2011effects;@kline2014local;@rud2012electricity;@van2017long], though electrification alone may not be a sufficient condition [@lee2020does].
+During those same two decades, power generation from all coal units of at least 30 megawatts in the country increased by almost 140 percent, from 42.4 gigawatts to more than 100 gigawatts.^[globalenergymonitor.org/] There are no doubt positive effects from this increase in power generation; for example, electricity may lead to economic and productivity growth [@dinkelman2011effects;@kline2014local;@rud2012electricity;@van2017long], though electrification alone may not be a sufficient condition [@lee2020does].
 
@@
 
-In this paper, I study the effects of emissions from coal power plants on agricultural productivity, labor allocation, and economic growth in India.
+I study the effects of emissions from coal power plants on agricultural productivity, labor allocation, and economic growth in India.
 Previous work has suggested that there may be harmful direct effects of air pollution [@heck1982assessment;@marshall1997hidden] and water pollution [@reddy2006impact] on agricultural productivity, including work showing that gold mines -- which release large amounts of pollutants into the surrounding environment -- lead to lower agricultural productivity in Ghana [@aragon2016polluting].
-This paper, however, tackles the broader case of coal plants, which are a leading source of power throughout the world.
-I show that air pollution from these plants can lead to decreased agricultural productivity.
+These effects can work through the well known effects on human health and productivity, but also directly through land productivity.
+Black carbon in soot, for example, can directly affect solar radiation reaching plants on the surface [@ramanathan2008global;@burney2014recent].
+Relative to previous work in economics, this paper tackles the broader case of coal plants, which are a leading source of power throughout the world.
+I show that air pollution from these plants can lead to lower agricultural productivity and even decreased economic growth.
 
@@
 For starters, the construction of coal plants is likely to be endogenous, though the relationship between construction and economic conditions is not a priori clear.
-On the one hand, governments may locate new plants in fast-growing areas, while on the other hand, they may decide to exact location of plants based on the political power of local citizens [@kopas2020environmental].
-In the case of the data used here, for example, coal plants are constructed in areas with higher population and higher literacy (Table \@ref(tab:plantresultstable)).
+On the one hand, governments may locate new plants in fast-growing areas, while on the other hand, they may decide the exact location of plants based on the political power of local citizens [@kopas2020environmental].
+In the case of the data used here, coal plants are constructed in areas with higher population and higher literacy (Table \@ref(tab:plantresultstable)).
 In terms of agriculture, higher productivity predicts more coal plants in a given year as well as the construction of a coal plant over the next decade, indicating possible bias when estimating OLS regressions of agricultural productivity on pollution levels.
@@
 To identify the effects of air pollution on agricultural productivity, I instead combine the roll out of coal plants in India with plausibly exogenous changes in wind direction -- in a similar spirit to @deryugina2019mortality -- to measure the effects of pollution from coal plants on key development outcomes.
-I first locate the location of coal plants in India by year.
+I first plot the location of coal plants in India by year.
 Using a village-level shapefile, I identify all villages that are located within 30km of a coal plant at any time during the sample period, which comprises the 1990s to the 2010s, with differences depending on data availability.
 To measure exposure to emissions from coal plants, I calculate the direction from each coal plant to all villages within 30km.
-I then download daily wind data and create an identifier separately for each of the one-hundred-thousand-plus villages for whether a given village is located downwind from a coal plant on any given day.
-I then aggregate this exposure variable based on the desired analysis; for agricultural productivity, for example, I create an exposure variable measuring the total number of exposure days in the five-month agricultural season.
+Using daily wind data and these spatial variables, I create an identifier separately for each of the one-hundred-thousand-plus villages for whether a given village is located downwind from a coal plant on any given day.
+I then aggregate this exposure variable based on the desired analysis; for agricultural productivity, for example, I create an exposure variable measuring the total number of exposure days in the five-month agricultural season, separately for the main monsoon season and the smaller winter season.
 Consistent with expectations, wind direction is a strong predictor of particulate matter at the monthly level.
@@
 This IV strategy is similar to that in @deryugina2019mortality, though they use general "high-pollution" wind directions while I concentrate only on wind directions from the location of a coal power plant.
-The IV estimates indicate that the effect of pollution is more negative than na\"{i}ve OLS estimates indicate, which is consistent with the evidence on the construction of coal plants.
+The IV estimates indicate that the effect of pollution is more negative than naïve OLS estimates indicate, which is consistent with the evidence on the construction of coal plants; higher levels of pollution are correlated with higher levels of agricultural productivity and economic growth, a relationship that the IV strategy adjusts downwards.
 
@@
 An additional day of wind over the previous week decreases overall labor allocation by `r paste0(substr(labortable[1,2],2,6))` days.
-However, all effects on labor allocation are driven by the older half of my sample; these adults see much large decreases in overall labor allocation, self employment, farm employment, and non-farm employment.
-This is suggestive evidence of productivity effects being driven by the health effects on individuals, though I am not able to rule out direct effects on land productivity with my data.
+However, the overall results show stark heterogeneity based on age; the older half of the sample greatly decreases labor allocation in response to pollution exposure while the younger half increases overall labor allocation.
+Older adults decrease labor allocation to both farm and non-farm labor, while younger adults increase labor allocated to farm labor.
+This is suggestive evidence of productivity effects being driven by the health and productivity effects on individuals, since older adults are more susceptible to pollution and they also greatly decrease non-farm employment, which presumably will no suffer any possible direct effects of pollution on productivity.
+The preponderance of results seems to indicate that effects on labor are the most likely mechanism.
 
@@
 While the effect for a single village might be small, the overall, aggregate effect on growth is quite substantial.
+To put this number in perspective, @castello2018higher argue that a standard-deviation increase in the share of adults with a university education leads to 12 percent higher nightlights in India, while @storeygard2016farther shows that a four-fold increase in oil prices over six years leads to seven percent lower nightlights at a distance of around 500km from the nearest large port in sub-Saharan Africa.
+In other words, the 0.7 percent change in nightlights is quite large given this is annual variation.
 
@@
 Perhaps relevant to my labor allocation results, @hanna2015effect show that decreases in pollution lead to increases in labor allocation, while I show the same relationship among the oldest cohort.
-The paper most related to this one is @aragon2016polluting, who show that gold mines in Ghana have substantial negative impacts on agricultural productivity in the surrounding areas.
+The papers most related to this one are @aragon2016polluting, who show that gold mines in Ghana have substantial negative impacts on agricultural productivity in the surrounding areas, and @burney2014recent, who model the effects of climate change and pollutants on agricultural productivity in India.
 This paper differs in two important ways.
-First, I explore effects of pollution from coal plants, which are much more ubiquitous than goald mines across the world.
-Second, I use a different identification strategy, relying on wind blowing pollution from coal plants.
+First, I explore effects of pollution from coal plants, which are much more ubiquitous than gold mines across the world.
+Second, I use a different identification strategy from either paper, relying on wind blowing pollution from coal plants.
 
@@
 
+In this section, I describe the different data sources and the identification strategy.
+I also spend time validating the use of wind direction as a proxy for particulate matter as well as looking at the correlates of construction of coal power plants.
+
 ## Data 
@@
 
-The main goal of this paper is to examine whether exposure to exhaust from coal plants affects agricultural productivity, labor allocation, and economic growth.
+The main goal of this paper is to examine whether exposure to emissions from coal plants affects agricultural productivity, labor allocation, and economic growth.
 I use several sources of data, which I describe here and in Table \@ref(tab:data) in the appendix.
@@
 
-Figure \@ref(fig:plants) presents the location of coal plants in India for two specific years: 1990 and 2010.
+Figure \@ref(fig:plants) displays the location of coal plants in India for two specific years: 1990 and 2010.
 There is a clear increase in the prevalence of coal plants across the country over the two decades.
-Additionally, the overall capacity from coal plants increased from `r paste0(reportDecOne(sum(plants1990_capacity$capacity)))` gigawatts to `r paste0(reportDecOne(sum(plants2010_capacity$capacity)))` gigawats, an increase of `r paste0(reportPercent((sum(plants2010_capacity$capacity) - sum(plants1990_capacity$capacity))/sum(plants1990_capacity$capacity)))` percent in just 20 years.
-This increase is driven both by the rollout of new plants as well as the construction of new units at existing plants, the latter of which is not shown in Figure \@ref(fig:plants).
+Additionally, the overall capacity from coal plants increased from `r paste0(reportDecOne(sum(plants1990_capacity$capacity)))` gigawatts to `r paste0(reportDecOne(sum(plants2010_capacity$capacity)))` gigawatts, an increase of `r paste0(reportPercent((sum(plants2010_capacity$capacity) - sum(plants1990_capacity$capacity))/sum(plants1990_capacity$capacity)))` percent in just 20 years.
+This increase is driven both by the roll out of new plants as well as the construction of new units at existing plants, the latter of which is not shown in Figure \@ref(fig:plants).
 
@@
 This data comes from @gangopadhyay2022new and estimates land productivity (i.e.
-yield, in tons per hectare) for the major crops in India.
-They define the monsoon season as June to October and the winter season as November to March.
+yield, in tons per hectare) for the major crops in India, using satellite-derived data.
+While there is a wealth of literature on measurement error in land [@desiere2018land;@abay2022nonclassical], the authors use a static cropland mask for the productivity estimates.
+This means that any measurement error will be constant across time and subsumed by the village fixed effects.
+While the measure of output is also noisy, the use of satellite-derived data and a static cropland mask reduces concerns about measurement error being responsible for the results below.
+The authors define the monsoon season as June to October and the winter season as November to March.
 I keep these definitions when matching across data below.
@@
 
+The inclusion of rainfall and temperature is meant to reduce concerns that wind direction may not affect agricultural productivity but, instead, be correlated with it through changes in weather.
+For example, prevailing winds from one direction may be associated with dry weather, while winds from another direction may be more likely to carry moisture and rain.
+It is not a priori clear how this would bias the estimates with respect to pollution and coal plants, however.
+The differences in weather patterns is very context specific and can differ even for villages located near one another.
+Nonetheless, adding rainfall and temperature should alleviate some of these concerns.
+
 A large question looms with respect to identification: what are the characteristics of areas near the construction of new coal plants?
@@
 To recover the actual effect of pollution, we need an instrument.
-If you believe that wind direction is exogenous and, conditional on the fixed effects and weather variables, only affects agricultural productivity through pollution -- which seems reasonable since it is hard to imagine what other variables it might affect -- then it is a valid instrument for pollution.
+If, conditional on the fixed effects and weather variables, wind direction is exogenous and only affects agricultural productivity through pollution -- which seems reasonable since it is hard to imagine what other variables it might affect -- then it is a valid instrument for pollution.
 
 I present the IV results in Table \@ref(tab:yieldtabletwo), with the second-stage results in the top panel and the first-stage results in the bottom panel.
-The first-stage results make clear how strong of an instrument wind is; the t-statistic in column one, for example, is almost 48 while it is still around 11 in column three, meeting even the higher bar for the first-stage F-test argued for by @lee2022valid.
+The first-stage results make clear how strong of an instrument wind is; the t-statistic in column one, for example, is almost 48 while it is still well above 10 in column three, which translates to a first-stage F-statistic of 195, meeting even the higher bar for the first-stage F-test argued for by @lee2022valid.
 Also note that the coefficients are quite different from those in Table \@ref(tab:pollutiontable) because the unit of analysis in Table \@ref(tab:yieldtabletwo) is the season, while in Table \@ref(tab:pollutiontable) the unit of analysis is the month.
@@
 In this scenario, the worsening air quality could lead to decreased labor productivity, meaning that "effective labor" essentially decreases when pollution is higher.
-Unfortunately, I am not able to disentangle these two possiblities with the data available.
 
@@
 
+One way to provide suggestive evidence of mechanisms is to explore heterogeneity in the effects.
 Previous research has shown larger effects of pollution on the health of the oldest [@deryugina2019mortality].
@@
 
+To put some of these changes in context, an additional day of exposure to pollution from wind increases self-employment for the younger half of the sample by around six percent of the mean and increases farm employment for the younger half by almost 10 percent of the mean.
+For the older half, on the other hand, an additional day of exposure decreases self-employment and non-farm employment by around five percent of their means.
+
 While speculative, one possible explanation for this pattern is a decrease in labor productivity.
 While the youngest may be able to readjust and increase their overall self-employment labor allocation, perhaps the oldest cannot due to, for example, health effects of pollution, which are well documented [@bentayeb2012adverse;@pope2000epidemiology].^[Much of the literature also suggests the youngest are at significant risk of mortality and morbidities.
-However, given my focus on labor allocation, I only discuss older adults here.]
+However, given my focus on labor allocation, I only discuss older adults here.] We see decreases in overall agricultural productivity, despite an increase in overall farm labor.
+We also see decreases in non-farm employment, indicating that individuals are reallocating labor from the non-farm sector to the farm sector.
+
+It is worth noting that the larger negative effects for older adults is more consistent with a labor story than a land story; it is hard to imagine these patterns emerging if the only effects were through land productivity changes.
+The overall decrease in non-farm employment is more compatible with a labor health/productivity story, though with general equilibrium effects in agriculture spillover over into the non-farm sector, this need not necessarily be the case.
+
+In addition, Table \@ref(tab:labortableoldstate) in the appendix presents one additional set of results to support a labor story.
+Changes in land productivity should be relatively fixed in a short period of time.
+For example, if we focus on just a single month in a single district, the overall effects driven by changes in land productivity should be constant even though pollution may have differential effects throughout the month on labor productivity, which responds more immediately to changes in ar quality.
+Table \@ref(tab:labortableoldstate) changes the specification to include district-by-year-by-month fixed effects.
+In other words, the results compare people in the same district in the same month, but who may have differential exposure to pollution simply because they were interviewed at a slightly different time.
+
+The overall pattern of results stays the same when we shift the focus.
+Overall labor allocation increases for the youngest, but decreases for the oldest, and this is driven completely by self employment.
+There are decreases in both farm and non-farm employment for the oldest, while the youngest show increases in non-farm employment, but not farm employment -- which is a change from the previous results.
+The overall story remains the same, however, which is further evidence of effects being driven primarily through changes in labor allocation and productivity.
+
+Finally, Table \@ref(tab:wagestable) in the appendix shows that there are no significant changes in daily wages, though the overall direction of wage changes would be ambiguous, anyway.
+While reduced productivity might lead to lower wages, decreased labor supply could also put upward pressure on wages.
 
@@
 ```
+
+\FloatBarrier
+\newpage
+```{r labortableoldstate, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+# Table
+colnames(labortableoldstate) <- c("all", "self", "wage", "farm", "non-farm")
+kable(
+      labortableoldstate, caption = "Wind direction and labor allocation, heterogeneity by age",
+      align = "ccccc", booktabs = TRUE, linesep = ""
+      ) %>%
+  footnote(
+           "Note: Standard errors are in parentheses and are clustered at the district level.
+Control variables include female, age, age squared, and (years of) education.
+* p<0.10 ** p<0.05 *** p<0.01",
+           general_title = "",
+           threeparttable = TRUE,
+           footnote_as_chunk = TRUE
+           ) %>%
+  column_spec(1, width = "4cm") %>%
+  column_spec(c(2:6),width = "1.5cm") %>%
+  row_spec(c(7), hline_after = TRUE) %>%
+  row_spec(c(6), bold = TRUE) %>%
+  kable_classic_2()
+
+```
+
+\FloatBarrier
+\newpage
+```{r wagestable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+# Table
+colnames(wagestable) <- c("farm", "non-farm", "all")
+kable(
+      wagestable, caption = "Wind direction and wages",
+      align = "ccc", booktabs = TRUE, linesep = ""
+      ) %>%
+  footnote(
+           "Note: Standard errors are in parentheses and are clustered at the district level.
+Control variables include female, age, age squared, and (years of) education.
+* p<0.10 ** p<0.05 *** p<0.01",
+           general_title = "",
+           threeparttable = TRUE,
+           footnote_as_chunk = TRUE
+           ) %>%
+  column_spec(1, width = "3cm") %>%
+  column_spec(c(2:4),width = "1.5cm") %>%
+  row_spec(c(8), hline_after = TRUE) %>%
+  row_spec(c(4,7), bold = TRUE) %>%
+  kable_classic_2()
+
+```

### 2023-01-20 analyses/regressions.R

@@ -239,6 +239,9 @@ saveRDS(yieldtable, "pollution_development/draft/tables/yieldtable.rds")
 
 
 
+
+
+
 yield1 <- feols(log(yield) ~ 1 | village^season + year | pm25 ~ wind, data = df, cluster = c("village"))
 yield2 <- feols(log(yield) ~ rain_z + temp_mean | village^season + year | pm25 ~ wind, data = df, cluster = c("village"))
 yield3 <- feols(log(yield) ~ rain_z + temp_mean | village^season + year^season^distfe | pm25 ~ wind, data = df, cluster = c("village"))
@@ -264,7 +267,7 @@ yieldtabletwo1 <- etable(
                         depvar = FALSE,
                         signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.1),
                         digits = 3,
-                        fitstat = c("n"),
+                        fitstat = c("ivwald"),
                         coefstat = "se"
                         #extralines = list("Sub-sample" = c("all", "all", "monsoon", "winter"))
                         )
@@ -519,6 +522,43 @@ saveRDS(labortable, "pollution_development/draft/tables/labortable.rds")
 
 
 
+# WAGES
+df$f_wages_all <- df$f_wages
+df$f_wages_all[is.na(df$f_wages_all)==T] <- 0
+df$nf_wages_all <- df$nf_wages
+df$nf_wages_all[is.na(df$nf_wages_all)==T] <- 0
+
+labor2 <- feols(log(f_wages/days_f) ~ wind + ..ctrl | district[year] + year,
+                  data = df %>% filter(f_wages_all>0),
+                  cluster = "district")
+labor3 <- feols(log(nf_wages/days_nf) ~ wind + ..ctrl | district[year] + year,
+                  data = df %>% filter(nf_wages_all>0),
+                  cluster = "district")
+labor4 <- feols(log((f_wages_all + nf_wages_all)/(days_f + days_nf)) ~ wind + ..ctrl | district[year] + year,
+                  data = df %>% filter(f_wages_all>0 | nf_wages_all>0),
+                  cluster = "district")
+
+
+labortable <- etable(
+                      labor2, labor3, labor4,
+                      se.below = TRUE,
+                      depvar = FALSE,
+                      signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.1),
+                      digits = 3,
+                      fitstat = c("n"),
+                      coefstat = "se",
[…]

### 2023-01-20 cleaning/3.cleaning_nss.R

@@ -29,8 +29,9 @@ df_hh <- df_hh %>% mutate(hid = paste(Sector, State_region, District, Stratum, S
                           district61 = DISTRICT_CODE,
                           weight_combined = WEIGHT_COMBINED,
                           weight_subround = WEIGHT_SUB_ROUND,
-                          weight_subsample = WEIGHT_SUB_SAMPLE) %>%
-                    select(hid, state61, district61, date, weight_combined, weight_subround, weight_subsample)
+                          weight_subsample = WEIGHT_SUB_SAMPLE,
+                          rural = as.numeric(Sector==1)) %>%
+                    select(hid, state61, district61, date, weight_combined, weight_subround, weight_subsample, rural)
 df_hh <- df_hh[complete.cases(df_hh),]
 
 # To join, with concordance files, need to do some changes
@@ -326,8 +327,9 @@ df_hh <- df_hh %>% mutate(hid = hhid,
                           date = as.Date(paste0("20", substr(S10B2_v02i, 5, 6), "-", substr(S10B2_v02i, 3, 4), "-", substr(S10B2_v02i, 1, 2)), "%Y-%m-%d"),
                           state62 = state,
                           district62 = district,
-                          weight_combined = weight) %>%
-                    select(hid, state62, district62, date, weight_combined)
+                          weight_combined = weight,
+                          rural = as.numeric(sector==1)) %>%
+                    select(hid, state62, district62, date, weight_combined, rural)
 df_hh <- df_hh[complete.cases(df_hh),]
 
 # To join, with concordance files, need to do some changes
@@ -620,8 +622,9 @@ df_hh <- df_hh %>% mutate(hid = key_Hhold,
                           district64 = District,
                           weight_combined = wgt_combined,
                           weight = wgt,
-                          weight_subsample = wgt_ss) %>%
-                    select(hid, state64, district64, date, weight_combined, weight, weight_subsample)
+                          weight_subsample = wgt_ss,
+                          rural = as.numeric(Sector==1)) %>%
+                    select(hid, state64, district64, date, weight_combined, weight, weight_subsample, rural)
 df_hh <- df_hh[complete.cases(df_hh),]
 
 # To join, with concordance files, need to do some changes
@@ -911,8 +914,9 @@ df_hh <- df_hh %>% mutate(hid = paste(fsu, sector, state, district, stratum, sub
                           state66 = state,
                           district66 = district,
                           weight = mlt,
-                          weight_subsample = mlt_sr) %>%
[…]

### 2023-01-20 cleaning/8.cleaning_overlap_merging_nss.R

@@ -70,6 +70,11 @@ df_labor <- df_labor %>% mutate(state_merge = as.numeric(state_merge),
 df_labor <- df_labor %>% filter(is.na(merge)==F)
 
 
+
+pollution <- read_csv(paste0("data/clean/pm25/all_combined.csv"))
+pollution <- pollution %>% rename(pm25 = mean)
+
+
 # Going to go through the dates and process date by date
 df_labor_merged <- c()
 for (day in 1:length(date_vec)){
@@ -116,12 +121,8 @@ write.csv(df_labor_merged, "data/clean/nss/merged_week.csv")
 
 
 
-
-pollution <- read_csv(paste0("data/clean/pm25/all_combined.csv"))
-pollution <- pollution %>% rename(pm25 = mean)
-df_labor_merged <- read_csv("data/clean/nss/merged_week.csv")
-
 # And also pm; all years covering the nss data
+df_labor_merged <- read_csv("data/clean/nss/merged_week.csv")
 pollution_merged <- c()
 for (y in year(min(df_labor_merged$date)):year(max(df_labor_merged$date))){
   for (m in 1:12){

### 2023-02-11 draft/draft.Rmd
updating tables

@@
 ---
-title: "Pollution, agricultural productivity, and development: Evidence from coal in plants in India^[I thank Marc Bellemare and David Sungho Park for comments and suggestions that greatly improved the quality of the paper.
+title: "Pollution, agricultural productivity, and development: Evidence from coal plants in India^[I thank Marc Bellemare and David Sungho Park for comments and suggestions that greatly improved the quality of the paper.
 The usual disclaimer applies.
@@
 labortable <- readRDS("pollution_development/draft/tables/labortable.rds")
+labortablerural <- readRDS("pollution_development/draft/tables/labortablerural.rds")
+labortableurban <- readRDS("pollution_development/draft/tables/labortableurban.rds")
 wagestable <- readRDS("pollution_development/draft/tables/wagestable.rds")
@@
 
+```{=tex}
 \newpage
 \doublespacing
-
+```
 # Introduction
@@
 Coal is one of the most common energy sources throughout the world.
-In 2021, energy generation from coal reached an all-time high,^[www.iea.org/reports/coal-fired-electricity] driven in part by rapid population growth and rising energy prices.
+In 2021, energy generation from coal reached an all-time high,[^1] driven in part by rapid population growth and rising energy prices.
 India is no exception to these trends.
-From 1990 to 2010, India's population grew by more than 40 percent.^[data.worldbank.org] This population growth led to a large increase in demand for power.
+From 1990 to 2010, India's population grew by more than 40 percent.[^2] This population growth led to a large increase in demand for power.
 While natural gas has become a popular alternative in much of the world, the Indian government met much of the increased demand through the construction of coal power plants.
-During those same two decades, power generation from all coal units of at least 30 megawatts in the country increased by almost 140 percent, from 42.4 gigawatts to more than 100 gigawatts.^[globalenergymonitor.org/] There are no doubt positive effects from this increase in power generation; for example, electricity may lead to economic and productivity growth [@dinkelman2011effects;@kline2014local;@rud2012electricity;@van2017long], though electrification alone may not be a sufficient condition [@lee2020does].
+During those same two decades, power generation from all coal units of at least 30 megawatts in the country increased by almost 140 percent, from 42.4 gigawatts to more than 100 gigawatts.[^3] There are no doubt positive effects from this increase in power generation; for example, electricity may lead to economic and productivity growth [@dinkelman2011effects; @kline2014local; @rud2012electricity; @van2017long], though electrification alone may not be a sufficient condition [@lee2020does].
+
+[^1]: www.iea.org/reports/coal-fired-electricity
+
+[^2]: data.worldbank.org
+
+[^3]: globalenergymonitor.org/
 
@@
 Pollutants from coal contribute to climate change and the phase out of coal energy is therefore a major climate goal [@iea2022].
-Emissions from the production and burning of coal are also directly harmful to health and the environment, including pollutants like sulfur dioxide, nitrogen oxides, and mercury, to name but a few.^[www.eia.gov/energyexplained/coal/coal-and-the-environment.php] While the use of coal had been declining, recent events -- including Russia's invasion of Ukraine -- have led to an increase in the demand for coal.^[www.npr.org/2022/08/15/1117560560/a-rising-demand-for-coal-amidst-war-in-ukraine]
+Emissions from the production and burning of coal are also directly harmful to health and the environment, including pollutants like sulfur dioxide, nitrogen oxides, and mercury, to name but a few.[^4] While the use of coal had been declining, recent events -- including Russia's invasion of Ukraine -- have led to an increase in the demand for coal.[^5]
+
+[^4]: www.eia.gov/energyexplained/coal/coal-and-the-environment.php
+
+[^5]: www.npr.org/2022/08/15/1117560560/a-rising-demand-for-coal-amidst-war-in-ukraine
 
 I study the effects of emissions from coal power plants on agricultural productivity, labor allocation, and economic growth in India.
-Previous work has suggested that there may be harmful direct effects of air pollution [@heck1982assessment;@marshall1997hidden] and water pollution [@reddy2006impact] on agricultural productivity, including work showing that gold mines -- which release large amounts of pollutants into the surrounding environment -- lead to lower agricultural productivity in Ghana [@aragon2016polluting].
+Previous work has suggested that there may be harmful direct effects of air pollution [@heck1982assessment; @marshall1997hidden] and water pollution [@reddy2006impact] on agricultural productivity, including work showing that gold mines -- which release large amounts of pollutants into the surrounding environment -- lead to lower agricultural productivity in Ghana [@aragon2016polluting].
 These effects can work through the well known effects on human health and productivity, but also directly through land productivity.
-Black carbon in soot, for example, can directly affect solar radiation reaching plants on the surface [@ramanathan2008global;@burney2014recent].
+Black carbon in soot, for example, can directly affect solar radiation reaching plants on the surface [@ramanathan2008global; @burney2014recent].
 Relative to previous work in economics, this paper tackles the broader case of coal plants, which are a leading source of power throughout the world.
@@
 First, we already have evidence of effects of pollution on different forms of labor activity.
-For example, general levels of pollution decrease productivity of call center works in China [@chang2019effect], lead people to perform worse on cognitive functions [@ebenstein2016long;@wen2022lower], and also lead to lower levels of farm labor productivity in California, driven by changes in ozone concentration [@graff2012impact].
+For example, general levels of pollution decrease productivity of call center works in China [@chang2019effect], lead people to perform worse on cognitive functions [@ebenstein2016long; @wen2022lower], and also lead to lower levels of farm labor productivity in California, driven by changes in ozone concentration [@graff2012impact].
 Also in China, @he2019severe find small but significant negative effects of pollution on manufacturing productivity, while @chen2022effect show pollution leads to labor out migration.
@@
 I also contribute to the more general literature on the effects of pollution on health.
-Pollution has substantial negative effects on health in both developing and developed countries [@arceo2016does] and we have known of these negative effects for many years [@brunekreef2002air;@kampa2008human;@pope2006health].
-@currie2014we review the literature on the health effects of pollution during adolescence; we have robust evidence of pollution exposure leading to increases in infant mortality, including in developing countries [@heft2018robust;@heft2019air].
+Pollution has substantial negative effects on health in both developing and developed countries [@arceo2016does] and we have known of these negative effects for many years [@brunekreef2002air; @kampa2008human; @pope2006health].
+@currie2014we review the literature on the health effects of pollution during adolescence; we have robust evidence of pollution exposure leading to increases in infant mortality, including in developing countries [@heft2018robust; @heft2019air].
 Since these effects are often driven by heart and lung disease and asthma [@kampa2008human], this my explain the labor allocation effects I identify in this paper.
@@
 # Data and methods
+
 \label{data}
@@
 ## Data
+
 \label{data2}
@@
 The first set of data lists the location of coal plant across the globe.
-This data comes from Global Energy Monitor^[globalenergymonitor.org/projects/global-coal-plant-tracker/] and lists all units generating at least 30 megawatts of electricity.
+This data comes from Global Energy Monitor[^6] and lists all units generating at least 30 megawatts of electricity.
 The data on coal plants includes the year of opening (and, if applicable, the year of retirement), the GPS (latitude/longitude) location of each plant, and the amount of power produced by the units at each plant.
 For this paper, I do not use the information on the capacity of the plant.
+
+[^6]: globalenergymonitor.org/projects/global-coal-plant-tracker/
 
@@
                 geom_sf(data = india_shape2, fill = NA, lwd = 0.01) +
-                geom_sf(data = plants1990, size = 0.75, color = "#440154FF") +
+                geom_sf(data = plants1990, size = 0.75, color = "black") +
                 labs(x = "lon",
@@
                 geom_sf(data = india_shape2, fill = NA, lwd = 0.01) +
-                geom_sf(data = plants2010, size = 0.75, color = "#440154FF") +
+                geom_sf(data = plants2010, size = 0.75, color = "black") +
                 labs(x = "lon",
@@
 yield, in tons per hectare) for the major crops in India, using satellite-derived data.
-While there is a wealth of literature on measurement error in land [@desiere2018land;@abay2022nonclassical], the authors use a static cropland mask for the productivity estimates.
+While there is a wealth of literature on measurement error in land [@desiere2018land; @abay2022nonclassical], the authors use a static cropland mask for the productivity estimates.
 This means that any measurement error will be constant across time and subsumed by the village fixed effects.
@@
 I keep these definitions when matching across data below.
-This data is also publicly available from Nature's data-sharing website.^[springernature.figshare.com] Since the main regressions below use village fixed effects, any differences in cropping patterns across areas should not bias estimates on changes, as long as these patterns do not change markedly in response to pollution.
+This data is also publicly available from Nature's data-sharing website.[^7] Since the main regressions below use village fixed effects, any differences in cropping patterns across areas should not bias estimates on changes, as long as these patterns do not change markedly in response to pollution.
+
+[^7]: springernature.figshare.com
 
 The agricultural productivity data is available as raster files with a resolution of 500m.
-To aggregate this data up to a useful administrative unit, I use village-level shapefiles provided by @almn2021 and publicly available on the SHRUG platform.^[www.devdatalab.org/shrug_download/] I aggregate the agricultural productivity data to the village level by extracting mean productivity for each feature in the shapefile.
+To aggregate this data up to a useful administrative unit, I use village-level shapefiles provided by @almn2021 and publicly available on the SHRUG platform.[^8] I aggregate the agricultural productivity data to the village level by extracting mean productivity for each feature in the shapefile.
 I do this separately for each season -- monsoon and winter -- and each year, from 2002 to 2013.
 
+[^8]: www.devdatalab.org/shrug_download/
+
 To measure exposure to pollution from coal plants, I first locate all village centroids located within 30km from a coal plant in a given year.
-I choose 30km due to previous research on the effects of (air) pollution [@aragon2016polluting;@bencko1977health;@li2014health].
+I choose 30km due to previous research on the effects of (air) pollution [@aragon2016polluting; @bencko1977health; @li2014health].
 I then calculate the direction from coal plants to all village centroids within that 30km radius.
-To define exposure, I then pull daily wind direction data from the National Center for Atmospheric Research.^[climatedataguide.ucar.edu/] For each day, I document whether the wind is blowing towards each village centroid.^[I define "towards" as within five degrees to help take into account that I am using village centroids, using the x and y components of wind speed.] I then temporally aggregate this daily data depending on the temporal definition of the corresponding outcome.
+To define exposure, I then pull daily wind direction data from the National Center for Atmospheric Research.[^9] For each day, I document whether the wind is blowing towards each village centroid.[^10] I then temporally aggregate this daily data depending on the temporal definition of the corresponding outcome.
 For example, agricultural productivity is defined across five months (e.g.
-the monsoon season is from June to October) so I count the total days a given village is exposed to wind blowing from _any_ coal plant within 30km during those five months.
+the monsoon season is from June to October) so I count the total days a given village is exposed to wind blowing from *any* coal plant within 30km during those five months.
+
+[^9]: climatedataguide.ucar.edu/
+
+[^10]: I define "towards" as within five degrees to help take into account that I am using village centroids, using the x and y components of wind speed.
 
@@
 Unfortunately, I am not aware of any publicly available survey data that covers multiple years and has identifiers below the district level.
-For example, the ARIS-REDS data^[www.ncaer.org/project/additional-rural-incomes-survey-aris-rural-economic-demographic-survey-reds] includes GPS identifiers for one year, but that only allows GPS matching for a total of two years: 1999 and 2007/2008.
+For example, the ARIS-REDS data[^11] includes GPS identifiers for one year, but that only allows GPS matching for a total of two years: 1999 and 2007/2008.
 The DHS, on the other hand, only has state-level identifiers (except for the newest two waves) and is also not a labor survey.
 Instead, I use the National Sample Survey (NSS) from the Ministry of Statistics and Programme Implementation.
+
+[^11]: www.ncaer.org/project/additional-rural-incomes-survey-aris-rural-economic-demographic-survey-reds
 
@@
 ## Identification
+
 \label{identification}
@@
 To calculate the effects of exposure to pollution on different outcomes, I rely on fixed effects and the differential exposure driven by plausibly exogenous wind direction, conditional on total rainfall and average temperature.
-The regression is of the form
-\begin{gather} y_{it} = \alpha_{i} + \gamma_{t} + \beta wind_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it}, \end{gather}
-where $y_{it}$ is outcome $y$ for unit $i$ at time $t$, $\alpha$ is a geographic fixed effects, $\gamma_t$ is time fixed effects, $wind_{it}$ is exposure to pollution -- defined using wind direction -- $rain_{it}$ is total rainfall during the season, and $temp_{it}$ is average temperature during the season.
+The regression is of the form \begin{gather} y_{it} = \alpha_{i} + \gamma_{t} + \beta wind_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it}, \end{gather} where $y_{it}$ is outcome $y$ for unit $i$ at time $t$, $\alpha$ is a geographic fixed effects, $\gamma_t$ is time fixed effects, $wind_{it}$ is exposure to pollution -- defined using wind direction -- $rain_{it}$ is total rainfall during the season, and $temp_{it}$ is average temperature during the season.
 Although I use $i$ and $t$ as subscripts for the fixed effects, note that these change depending on the unit of analysis and the outcome.
@@
 The outcome variable is particulate matter -- specifically, $\mathrm{PM_{2.5}}$, which is particulate matter no larger than 2.5 micrometers in diameter -- which comes from @hammer2020global.
-$\mathrm{PM_{2.5}}$ is one of the harmful byproducts from coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.^[www.epa.gov/airmarkets/power-plants-and-neighboring-communities] This data is available at the monthly level, so in Table \@ref(tab:pollutiontable), I aggregate wind exposure to the monthly level, as well.
+$\mathrm{PM_{2.5}}$ is one of the harmful byproducts from coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.[^12] This data is available at the monthly level, so in Table \@ref(tab:pollutiontable), I aggregate wind exposure to the monthly level, as well.
+
+[^12]: www.epa.gov/airmarkets/power-plants-and-neighboring-communities
 
@@
 # Results
+
 \label{results}
@@
   footnote(
-           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
+           "Note: Standard errors are in parentheses and are clustered at the village level.
+The outcome in all regressions is estimated agricultural yield, defined as tons per hectare.
+* p<0.10 ** p<0.05 *** p<0.01",
            general_title = "",
@@
   footnote(
-           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
+           "Note: Standard errors are in parentheses and are clustered at the village level.
+The outcome in all regressions is estimated agricultural yield, defined as tons per hectare.
+The first two columns split the sample based on the *maximum* observed exposure.
+* p<0.10 ** p<0.05 *** p<0.01",
            general_title = "",
@@
 
-```{r labortable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-# Table
-colnames(labortable) <- c("all", "all", "self", "wage", "farm", "non-farm")
-kable(
-      labortable, caption = "Wind direction and labor allocation",
+```{r labortablerural, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+# Table
+colnames(labortablerural) <- c("all", "all", "self", "wage", "farm", "non-farm")
+colnames(labortableurban) <- c("all", "all", "self", "wage", "farm", "non-farm")
+
+mat1 <- matrix(c("", "", "", "", "", ""), nrow = 1)
+colnames(mat1) <- c("all", "all", "self", "wage", "farm", "non-farm")
+rownames(mat1) <- c("rural")
+mat2 <- matrix(c("", "", "", "", "", ""), nrow = 1)
+colnames(mat2) <- c("all", "all", "self", "wage", "farm", "non-farm")
+rownames(mat2) <- c("urban")
+
+labortablerural <- rbind(mat1, labortablerural[c(1:2, 7:8),], mat2, labortableurban[-c(3:6),], labortableurban[c(3:6),])
+
+kable(
+      labortablerural, caption = "Wind direction and labor allocation, rural only",
       align = "cccccc", booktabs = TRUE, linesep = ""
@@
 Control variables include female, age, age squared, and (years of) education.
+The outcome in each regression is days worked in the previous seven days.
 * p<0.10 ** p<0.05 *** p<0.01",
@@
   column_spec(c(2:7),width = "1.5cm") %>%
-  row_spec(c(8), hline_after = TRUE) %>%
-  row_spec(c(4,7), bold = TRUE) %>%
+  row_spec(c(5,10), hline_after = TRUE) %>%
+  row_spec(c(1,6,12), bold = TRUE) %>%
   kable_classic_2()
@@
 The labor sample includes all adults between 15 and 60 years of age.
-To look at heterogeneity by age, I present an additional set of results interacting the wind variable with an indicator for whether an individual is over 37 years of age.^[This splits the sample into two equal bins of age: those who are 15-37 and those who are 38-60.] These results are in Table \@ref(tab:labortableold), which otherwise presents the same specifications in the last five columns as \@ref(tab:labortable).
-
-```{r labortableold, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-# Table
-labortableold <- labortableold[,-1]
-colnames(labortableold) <- c("all", "self", "wage", "farm", "non-farm")
-kable(
-      labortableold, caption = "Wind direction and labor allocation, heterogeneity by age",
-      align = "ccccc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the district level.
-Control variables include female, age, age squared, and (years of) education.
-* p<0.10 ** p<0.05 *** p<0.01",
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
-           ) %>%
-  column_spec(1, width = "4cm") %>%
-  column_spec(c(2:6),width = "1.5cm") %>%
-  row_spec(c(10), hline_after = TRUE) %>%
-  row_spec(c(6,9), bold = TRUE) %>%
-  kable_classic_2()
-
-```
+To look at heterogeneity by age, I present an additional set of results interacting the wind variable with an indicator for whether an individual is over 37 years of age.[^13] These results are in Table \@ref(tab:labortableold), which otherwise presents the same specifications in the last five columns as \@ref(tab:labortable).
+
+[^13]: This splits the sample into two equal bins of age: those who are 15-37 and those who are 38-60.
 
@@
 While speculative, one possible explanation for this pattern is a decrease in labor productivity.
-While the youngest may be able to readjust and increase their overall self-employment labor allocation, perhaps the oldest cannot due to, for example, health effects of pollution, which are well documented [@bentayeb2012adverse;@pope2000epidemiology].^[Much of the literature also suggests the youngest are at significant risk of mortality and morbidities.
-However, given my focus on labor allocation, I only discuss older adults here.] We see decreases in overall agricultural productivity, despite an increase in overall farm labor.
+While the youngest may be able to readjust and increase their overall self-employment labor allocation, perhaps the oldest cannot due to, for example, health effects of pollution, which are well documented [@bentayeb2012adverse; @pope2000epidemiology].[^14] We see decreases in overall agricultural productivity, despite an increase in overall farm labor.
 We also see decreases in non-farm employment, indicating that individuals are reallocating labor from the non-farm sector to the farm sector.
+
+[^14]: Much of the literature also suggests the youngest are at significant risk of mortality and morbidities. However, given my focus on labor allocation, I only discuss older adults here.
 
@@
 # Table
-colnames(ntltable) <- c("(1)", "(2)")
+colnames(ntltable) <- c("(1)", "(2)", "(3)", "(4)")
 kable(
       ntltable, caption = "Wind direction and nightlight growth",
-      align = "cc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
+      align = "cccc", booktabs = TRUE, linesep = ""
+      ) %>%
+  footnote(
+           "Note: Standard errors are in parentheses and are clustered at the village level.
+The outcome in all regressions is logged nightlights.
+Rainfall is included in columns (2) and (4).
+* p<0.10 ** p<0.05 *** p<0.01",
            general_title = "",
@@
   column_spec(1, width = "3cm") %>%
-  column_spec(c(2:3),width = "2cm") %>%
+  column_spec(c(2:5),width = "2cm") %>%
   row_spec(c(9), hline_after = TRUE) %>%
@@
 # Conclusion
+
 \label{conclusion}
@@
 
+```{=tex}
 \FloatBarrier
@@
 \singlespacing
+```
 # References {.unnumbered}
@@
 
+```{=tex}
 \FloatBarrier
 \newpage
-
+```
 # Appendix A {.unnumbered}
+
+```{=tex}
 \beginappendix
 \FloatBarrier
-
+```
 ```{r data, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
@@
 
+```{=tex}
 \FloatBarrier
 \newpage
+```
 ```{r yieldtablepostplant, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
@@
 The sample is restricted to village-years after a coal plant was built near the village.
+The outcome in all regressions estimated agricultural yield (tons per hectare).
 * p<0.10 ** p<0.05 *** p<0.01",
@@
 
+```{=tex}
 \FloatBarrier
 \newpage
+```
 ```{r yieldtableleads, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
@@
   footnote(
-           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
+           "Note: Standard errors are in parentheses and are clustered at the village level.
+The outcome in all regressions estimated agricultural yield (tons per hectare).
+* p<0.10 ** p<0.05 *** p<0.01",
            general_title = "",
@@
 
+```{=tex}
 \FloatBarrier
 \newpage
+```
 ```{r laborplot, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Timing of household surveys for the National Sample Survey", fig.footer = list("Note: The histogram shows the distribution of survey dates for the four waves of the National Sample Survey (NSS) for the five waves used in this paper.")}
@@
 
+```{=tex}
 \FloatBarrier
 \newpage
+```
 ```{r labortablemonsoon, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
@@
 
+```{=tex}
 \FloatBarrier
 \newpage
+```
 ```{r labortablewinter, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
@@
 
+```{=tex}
 \FloatBarrier
 \newpage
+```
 ```{r labortableoldstate, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
@@
 
+```{=tex}
 \FloatBarrier
 \newpage
+```
 ```{r wagestable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
@@
 Control variables include female, age, age squared, and (years of) education.
+The outcome in each regression is the daily wage (log Rs per day).
 * p<0.10 ** p<0.05 *** p<0.01",

### 2023-02-11 draft/draft.tex
updating tables

@@
 \hypersetup{
-  pdftitle={Pollution, agricultural productivity, and development: Evidence from coal in plants in India},
+  pdftitle={Pollution, agricultural productivity, and development: Evidence from coal plants in India},
   pdfauthor={Joshua D. Merfeld},
@@
 
-\title{Pollution, agricultural productivity, and development: Evidence from coal in plants in India\footnote{I thank Marc Bellemare and David Sungho Park for comments and suggestions that greatly improved the quality of the paper.
+\title{Pollution, agricultural productivity, and development: Evidence from coal plants in India\footnote{I thank Marc Bellemare and David Sungho Park for comments and suggestions that greatly improved the quality of the paper.
 The usual disclaimer applies.
@@
 \author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
-\date{2023-01-20}
+\date{2023-02-10}
 
@@
 While I am not able to explicitly test effects on land productivity, I look at how adults allocate labor over the last seven days using the National Sample Survey and creating a district-level exposure measure based on the timing of the survey for each household.
-An additional day of wind over the previous week decreases overall labor allocation by 0.027 days.
+An additional day of wind over the previous week decreases overall labor allocation by 0.055 days.
 However, the overall results show stark heterogeneity based on age; the older half of the sample greatly decreases labor allocation in response to pollution exposure while the younger half increases overall labor allocation.
@@
 To calculate the effects of exposure to pollution on different outcomes, I rely on fixed effects and the differential exposure driven by plausibly exogenous wind direction, conditional on total rainfall and average temperature.
-The regression is of the form
-\begin{gather} y_{it} = \alpha_{i} + \gamma_{t} + \beta wind_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it}, \end{gather}
-where \(y_{it}\) is outcome \(y\) for unit \(i\) at time \(t\), \(\alpha\) is a geographic fixed effects, \(\gamma_t\) is time fixed effects, \(wind_{it}\) is exposure to pollution -- defined using wind direction -- \(rain_{it}\) is total rainfall during the season, and \(temp_{it}\) is average temperature during the season.
+The regression is of the form \begin{gather} y_{it} = \alpha_{i} + \gamma_{t} + \beta wind_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it}, \end{gather} where \(y_{it}\) is outcome \(y\) for unit \(i\) at time \(t\), \(\alpha\) is a geographic fixed effects, \(\gamma_t\) is time fixed effects, \(wind_{it}\) is exposure to pollution -- defined using wind direction -- \(rain_{it}\) is total rainfall during the season, and \(temp_{it}\) is average temperature during the season.
 Although I use \(i\) and \(t\) as subscripts for the fixed effects, note that these change depending on the unit of analysis and the outcome.
@@
 \begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01
+\item Note: Standard errors are in parentheses and are clustered at the village level.
+The outcome in all regressions is estimated agricultural yield, defined as tons per hectare.
+* p<0.10 ** p<0.05 *** p<0.01
 \end{tablenotes}
@@
 \begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01
+\item Note: Standard errors are in parentheses and are clustered at the village level.
+The outcome in all regressions is estimated agricultural yield, defined as tons per hectare.
+The first two columns split the sample based on the *maximum* observed exposure.
+* p<0.10 ** p<0.05 *** p<0.01
 \end{tablenotes}
@@
 
-\caption{\label{tab:labortable}Wind direction and labor allocation}
+\caption{\label{tab:labortablerural}Wind direction and labor allocation, rural only}
 \centering
@@
 \midrule
-wind & -0.021 & -0.027* & 0.008 & -0.034** & 0.035* & -0.061**\\
- & (0.015) & (0.015) & (0.017) & (0.016) & (0.019) & (0.025)\\
+\textbf{rural} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+wind & -0.112*** & -0.118*** & -0.050* & -0.068** & -0.002 & -0.116***\\
+ & (0.036) & (0.029) & (0.030) & (0.031) & (0.028) & (0.034)\\
+DV mean & 3.499 & 3.499 & 2.172 & 1.327 & 1.260 & 2.240\\
+observations & 529,562 & 529,466 & 529,466 & 529,466 & 529,466 & 529,466\\
+\midrule
+\textbf{urban} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+wind & -0.020 & -0.030 & -0.011 & -0.019 & 0.023* & -0.053\\
+ & (0.033) & (0.041) & (0.039) & (0.031) & (0.014) & (0.048)\\
+DV mean & 3.040 & 3.040 & 1.393 & 1.647 & 0.101 & 2.939\\
+observations & 369,483 & 369,390 & 369,390 & 369,390 & 369,390 & 369,390\\
+\midrule
 controls & No & Yes & Yes & Yes & Yes & Yes\\
@@
 district & Yes & Yes & Yes & Yes & Yes & Yes\\
-year & Yes & Yes & Yes & Yes & Yes & Yes\\
-\textbf{varying slopes:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-year (by district) & Yes & Yes & Yes & Yes & Yes & Yes\\
-\midrule
-observations & 899,045 & 898,856 & 898,856 & 898,856 & 898,856 & 898,856\\
-\bottomrule
-\end{tabular}
-\begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the district level. Control variables include female, age, age squared, and (years of) education. * p<0.10 ** p<0.05 *** p<0.01
+state-month & Yes & Yes & Yes & Yes & Yes & Yes\\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}[para]
+\item Note: Standard errors are in parentheses and are clustered at the district level.
+Control variables include female, age, age squared, and (years of) education.
+The outcome in each regression is days worked in the previous seven days.
+* p<0.10 ** p<0.05 *** p<0.01
 \end{tablenotes}
@@
 
-\begin{table}
-
-\caption{\label{tab:labortableold}Wind direction and labor allocation, heterogeneity by age}
-\centering
-\begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}}
-\toprule
-  & all & self & wage & farm & non-farm\\
-\midrule
-wind & 0.061*** & 0.096*** & -0.034 & 0.068*** & -0.006\\
- & (0.023) & (0.022) & (0.024) & (0.020) & (0.030)\\
-wind times age>=38 & -0.207*** & -0.207*** & -0.0001 & -0.078*** & -0.129***\\
- & (0.037) & (0.033) & (0.040) & (0.020) & (0.030)\\
-controls & Yes & Yes & Yes & Yes & Yes\\
-\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-district & Yes & Yes & Yes & Yes & Yes\\
-year & Yes & Yes & Yes & Yes & Yes\\
-\textbf{varying slopes:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-year (by district) & Yes & Yes & Yes & Yes & Yes\\
-\midrule
-observations & 898,856 & 898,856 & 898,856 & 898,856 & 898,856\\
-\bottomrule
-\end{tabular}
-\begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the district level. Control variables include female, age, age squared, and (years of) education. * p<0.10 ** p<0.05 *** p<0.01
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
 Table \ref{tab:labortableold} presents a very clear set of results: the effects are always larger (more negative) for older adults.
@@
 \begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
-\toprule
-  & (1) & (2)\\
-\midrule
-wind (lagged) & -0.080*** & -0.083***\\
- & (0.015) & (0.015)\\
-rain (z, lagged) &  & -0.035***\\
- &  & (0.002)\\
-\textbf{fixed effects:} & \textbf{} & \textbf{}\\
-village & Yes & Yes\\
-year & Yes & Yes\\
-\textbf{varying slopes:} & \textbf{} & \textbf{}\\
-year (by village) & Yes & Yes\\
-\midrule
-observations & 2,146,715 & 2,146,715\\
-\bottomrule
-\end{tabular}
-\begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\toprule
+  & (1) & (2) & (3) & (4)\\
+\midrule
+wind (lag) & -0.024*** & -0.024*** &  & \\
+ & (0.003) & (0.003) &  & \\
+wind (lead) &  &  & -0.001 & -0.001\\
+ &  &  & (0.002) & (0.002)\\
+\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+village & Yes & Yes & Yes & Yes\\
+year & Yes & Yes & Yes & Yes\\
+\textbf{varying slopes:} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+year (by village) & Yes & Yes & Yes & Yes\\
+\midrule
+observations & 1,694,535 & 1,694,535 & 1,807,504 & 1,807,504\\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}[para]
+\item Note: Standard errors are in parentheses and are clustered at the village level.
+The outcome in all regressions is logged nightlights.
+Rainfall is included in columns (2) and (4).
+* p<0.10 ** p<0.05 *** p<0.01
 \end{tablenotes}
@@
 
-\setcounter{table}{0} \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}} 
+\beginappendix
 \FloatBarrier
@@
 \newpage
+
 \begin{table}[H]
@@
 The sample is restricted to village-years after a coal plant was built near the village.
+The outcome in all regressions estimated agricultural yield (tons per hectare).
 * p<0.10 ** p<0.05 *** p<0.01
@@
 \newpage
+
 \begin{table}
@@
 \begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01
+\item Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions estimated agricultural yield (tons per hectare). * p<0.10 ** p<0.05 *** p<0.01
 \end{tablenotes}
@@
 \newpage
+
 \begin{figure}
@@
 \midrule
-wind & -0.021 & -0.033 & 0.038 & -0.071** & 0.072** & -0.105**\\
- & (0.028) & (0.029) & (0.034) & (0.033) & (0.036) & (0.046)\\
+wind & -0.120*** & -0.119** & -0.067 & -0.052 & 0.025 & -0.144**\\
+ & (0.046) & (0.046) & (0.042) & (0.045) & (0.036) & (0.062)\\
 controls & No & Yes & Yes & Yes & Yes & Yes\\
@@
 district & Yes & Yes & Yes & Yes & Yes & Yes\\
-year & Yes & Yes & Yes & Yes & Yes & Yes\\
-\textbf{varying slopes:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-year (by district) & Yes & Yes & Yes & Yes & Yes & Yes\\
-\midrule
+state-month & Yes & Yes & Yes & Yes & Yes & Yes\\
+\textbf{DV mean} & \textbf{3.324} & \textbf{3.324} & \textbf{1.883} & \textbf{1.441} & \textbf{0.810} & \textbf{2.513}\\
 observations & 359,652 & 359,576 & 359,576 & 359,576 & 359,576 & 359,576\\
+\midrule
 \bottomrule
@@
 \midrule
-wind & 0.003 & 0.002 & 0.011 & -0.009 & 0.024 & -0.022\\
- & (0.029) & (0.026) & (0.028) & (0.025) & (0.034) & (0.039)\\
+wind & -0.012 & -0.032 & -0.026 & -0.006 & -0.020 & -0.012\\
+ & (0.040) & (0.037) & (0.040) & (0.028) & (0.030) & (0.036)\\
 controls & No & Yes & Yes & Yes & Yes & Yes\\
@@
 district & Yes & Yes & Yes & Yes & Yes & Yes\\
-year & Yes & Yes & Yes & Yes & Yes & Yes\\
-\textbf{varying slopes:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-year (by district) & Yes & Yes & Yes & Yes & Yes & Yes\\
-\midrule
+state-month & Yes & Yes & Yes & Yes & Yes & Yes\\
+\textbf{DV mean} & \textbf{3.339} & \textbf{3.339} & \textbf{1.865} & \textbf{1.474} & \textbf{0.790} & \textbf{2.550}\\
 observations & 375,956 & 375,883 & 375,883 & 375,883 & 375,883 & 375,883\\
-\bottomrule
-\end{tabular}
-\begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the district level. Control variables include female, age, age squared, and (years of) education. * p<0.10 ** p<0.05 *** p<0.01
+\midrule
+\bottomrule
+\end{tabular}
+\begin{tablenotes}[para]
+\item Note: Standard errors are in parentheses and are clustered at the district level.
+Control variables include female, age, age squared, and (years of) education.
+The outcome in each regression is days worked in the previous seven days.
+* p<0.10 ** p<0.05 *** p<0.01
 \end{tablenotes}
@@
 \begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the district level. Control variables include female, age, age squared, and (years of) education. * p<0.10 ** p<0.05 *** p<0.01
+\item Note: Standard errors are in parentheses and are clustered at the district level.
+Control variables include female, age, age squared, and (years of) education.
+The outcome in each regression is days worked in the previous seven days.
+* p<0.10 ** p<0.05 *** p<0.01
 \end{tablenotes}
@@
 \begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the district level. Control variables include female, age, age squared, and (years of) education. * p<0.10 ** p<0.05 *** p<0.01
+\item Note: Standard errors are in parentheses and are clustered at the district level.
+Control variables include female, age, age squared, and (years of) education.
+The outcome in each regression is the daily wage (log Rs per day).
+* p<0.10 ** p<0.05 *** p<0.01
 \end{tablenotes}

### 2023-02-11 presentation/slides/slides.Rmd
updating tables

@@
 library(lubridate)
+library(icons)
+library(png)
 loadfonts()  # Must be run once in each session (markdown is a new "session" each time you render/knit)
@@
 
+```{css, echo=F}
+    /* Table width = 100% max-width */
+
+    <!-- .remark-slide table{ -->
+    <!--     width: 100%; -->
+    <!-- } -->
+
+    /* Change the background color to white for shaded rows (even rows) */
+
+    .remark-slide thead, .remark-slide tr:nth-child(n) {
+        background-color: white;
+    }
+
+    tfoot {
+      font-size: 80%;
+    }
+```
+
+## Coal and energy worldwide
+
+- Coal one of most common energy sources worldwide
+  - Reached an all-time high in 2021 <small><small>(www.iea.org/reports/coal-fired-electricity)</small></small>
+
+![]("../../../pollution_development/draft/tables/energy.png")
+
+---
+
+## Coal and energy in India
+
+![]("../../../pollution_development/draft/tables/energy.png")
+
+---
+
+## Coal and energy in India
+
+- From 1990 to 2010: coal generation from 42.4 gigawatts to more than 100
+- Obvious benefits to more power <small><small>(Dinkelman 2011; Kline and Moretti 2014; Rud 2012; Van de Walle et al. 2017)</small></small>
+
+---
+
+## Downsides, too
+
+- But power generation can have downsides, too
+  - Especially true near plants: pollution a major concern
+- Climate change and pollutants like SO2, mercury, etc. <small><small>(www.eia.gov/energyexplained/coal/coal-and-the-environment.php)</small></small>
+- Long literature on negative effects of pollution 
+  - Health <small><small>(Arceo, Hanna, and Oliva 2016; Brunekreef and Holgate 2002; Deryugina et al. 2019; Kampa and Castanas 2008; Pope III and Dockery 2006)</small></small>
+  - Productivity/labor allocation <small><small>(Chang et al. 2019; Ebenstein, Lavy, and Roth 2016; Hanna and Oliva 2015; He, Liu, and Salvo 2019; Wen and Burke 2022)</small></small>
+
+---
+
+## This paper
+
+- Effects of pollution from coal plants on agricultural productivity in India
+  - Previous evidence from gold mines (Aragon and Rud 2016) and workers in California (Graff Zivin and Neidell 2012)
+- Higher exposure to pollution leads to lower agricultural productivity
+  - Exposure defined as wind direction from coal plants towards villages
+  - One additional day of exposure per season: decrease of 0.3 percent (yields)
+
+---
+
+## This paper
+
+- Adults decrease labor allocation
+  - Driven by large decrease in older half of sample; younger half increases allocation to farm labor
+- More pollution in previous year also leads to lower nightlight growth
+  - 25th to 75th percentile of exposure leads to decrease in nightlights of 0.7 percent
+
+---
+
+## Data sources
+
+```{r data, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+data <- matrix(NA, nrow = 8, ncol = 3)
+rownames(data) <- c("shapefile", "coal plants", "wind", "pollution", "agriculture", "nightlights", "weather", "labor")
+colnames(data) <- c("source", "geographic coverage", "temporal coverage")
+data[1,1] <- "Asher et al. (2021)"
+data[2,1] <- "Global Energy Monitor"
+data[3,1] <- "NCAR"
+data[4,1] <- "Hammer et al. (2020)"
+data[5,1] <- "Angopadhyay et al. (2022)"
+data[6,1] <- "Asher et al. (2021)"
+data[7,1] <- "TerraClimate"
+data[8,1] <- "National Sample Survey (NSS)"
+data[1,2] <- "India"
+data[2,2] <- "global"
+data[3,2] <- "global"
+data[4,2] <- "global"
+data[5,2] <- "India"
+data[6,2] <- "India"
+data[7,2] <- "global"
+data[8,2] <- "India"
+data[1,3] <- " "
+data[2,3] <- "yearly"
+data[3,3] <- "daily"
+data[4,3] <- "monthly"
+data[5,3] <- "two seasons/year"
+data[6,3] <- "yearly"
+data[7,3] <- "monthly"
+data[8,3] <- "daily interviews"
+
+# Table
+kable(
+      data,
+      align = "ccccc", booktabs = TRUE, linesep = "", escape = TRUE
+      ) %>%
+  footnote(
+           general = "Global Energy Monitor: globalenergymonitor.org/projects/global-coal-plant-tracker. NCAR: climatedataguide.ucar.edu/. TerraClimate: www.climatologylab.org/terraclimate.html",
+           general_title = "",
+           threeparttable = TRUE,
+           footnote_as_chunk = TRUE
+           ) %>%
+  column_spec(1, width = "2cm") %>%
+  column_spec(2,width = "4.5cm") %>%
+  column_spec(c(3:4),width = "3.5cm") %>%
+  kable_styling(latex_options = "HOLD_position") %>%
+  kable_classic_2()
+
+```
+
+---
+
+## General idea
+- Coal plants open in a given year
+- Plot wind direction each day
+  - Exposure means wind blowing from coal plant to village (within 30km)
+  - Aggregate up to desired level (e.g. season, month, week)
+
+---
+
 ## Coal plants over time
@@
 
+```{r windbase, echo = FALSE, message = FALSE, warning = FALSE,}
+
+ggplot() +
+    geom_sf(data = districts, color = "gray85", fill = "gray80") +
+    geom_sf(data = villageswind %>% filter(date=="2010-01-01"),
+            alpha = .5) +
+    labs(title = "date: 2010-01-01") +
+    theme(legend.position = "none")
+
+```
+
+---
+
+## Wind direction - first 100 days of 2010 in Guna district (MP)
+
 ```{r wind, echo = FALSE, message = FALSE, warning = FALSE,}
@@
 ggplot() +
-    geom_sf(data = districts, color = "gray85", fill = "gray80") +
     geom_sf(data = villageswind, 
@@
 kable(
-      plantresultstable, caption = "Local characteristics and the construction of coal plants",
+      plantresultstable,
       align = "cccc", booktabs = TRUE, linesep = "", escape = FALSE
@@
 
-## Validating wind direction
+## Validating wind direction - particulate matter and exposure
 
@@
 kable(
-      pollutiontable, caption = "Wind direction and particulate matter",
+      pollutiontable,
       align = "cccc", booktabs = TRUE, linesep = ""
@@
 kable(
-      yieldtable, caption = "Wind direction and agricultural productivity",
+      yieldtable,
       align = "ccccc", booktabs = TRUE, linesep = ""
@@
 
+- Likely endogenous
+  - Upward biased?
+
 ```{r yieldtablepm, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
@@
 kable(
-      yieldtablepm, caption = "Particulate matter (PM 2.5) and agricultural productivity",
+      yieldtablepm[-c(3:6),],
       align = "ccccc", booktabs = TRUE, linesep = ""
       ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
-           ) %>%
+  # footnote(
+  #          "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
+  #          general_title = "",
+  #          threeparttable = TRUE,
+  #          footnote_as_chunk = TRUE
+  #          ) %>%
   column_spec(1, width = "3cm") %>%
   column_spec(c(2:6),width = "2cm") %>%
-  row_spec(7, bold = TRUE) %>%
-  row_spec(11, hline_after = TRUE) %>%
+  row_spec(3, bold = TRUE) %>%
+  row_spec(7, hline_after = TRUE) %>%
   add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
@@
 ```
+
+---
+
+## Effects of PM on agricultural productivity
+
+- We are really interested in the effect of _pollution_, not wind
+- Use wind as instrument? Assumes...
+  - Conditional on fixed effects and weather...
+  - Wind only affects agriculture through pollution
+- Reasonable?
 
@@
 kable(
-      yieldtabletwo, caption = "Pollution and agricultural productivity, IV estimates",
+      yieldtabletwo[-c(3:6, 12, 16:19),],
       align = "cccc", booktabs = TRUE, linesep = ""
       ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
-           ) %>%
+  # footnote(
+  #          "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
+  #          general_title = "",
+  #          threeparttable = TRUE,
+  #          footnote_as_chunk = TRUE
+  #          ) %>%
   column_spec(1, width = "3cm") %>%
   column_spec(c(2:6),width = "2cm") %>%
-  row_spec(c(7,13), bold = TRUE) %>%
-  row_spec(c(11:12), hline_after = TRUE) %>%
+  row_spec(c(3,8), bold = TRUE) %>%
+  row_spec(c(2), hline_after = TRUE) %>%
   add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
   kable_classic_2()
+
+```
+
+---
+
+## Robustness checks
+
+- Leads of wind direction
+- Only post coal plant opening
+
+---
+
+## Taking stock
+
+- We see clear negative effects of pollution on productivity
+  - Absolute deviation of exposure is approx. 8 days, meaning 2-3 percent change in productivity
+  - Using AD of PM, it's above 20 percent change
+- Key question: what is the driver?
+  - Land?
+  - Labor?
+- Hard to look explicitly at land, so let's look at labor
+
+---
+
+## National Sample Survey (NSS)
+
+```{r nss, echo = FALSE, message = FALSE, warning = FALSE,}
+
+laborplot
+
+```
+
+---
+
+## Problem: only have district identifiers, so aggregate up
+
+```{r aggregation, echo = FALSE, message = FALSE, warning = FALSE,}
+
+# Sets wd to folder this script is in so we can create relative paths instead of absolute paths
+setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
+# We want to get the wd up to the main folder
+# Need to go up three levels
+setwd('..')
+setwd('..')
+setwd('..')
+
+# district
+districts <- read_sf(paste0("data/spatial/districts/districts.shp"))
+districts <- districts %>% 
+                dplyr::select(state = ST_CEN_CD, district = DT_CEN_CD, state_name = ST_NM, district_name = DISTRICT)
+districts <- districts %>% filter(state=="23" & district=="07")
+# Villages
+villages <- read_sf(paste0("data/spatial/villages_overlap/villages_overlap.shp"))
+villages <- villages %>% mutate(shrid = paste0(pc11_s_id, "-", pc11_tv_id)) %>%
+                          dplyr::select(shrid, state = ST_CEN_CD, district = DT_CEN_CD, state_name = ST_NM, district_name = DISTRICT)
+villages <- villages %>% filter(state=="23" & district=="07")
+wind <- read_csv(paste0("data/clean/wind_ntl/days/date_2010-1-1.csv")) %>% as_tibble()
+wind$days_sum <- apply(wind[,2:8], 1, FUN = "sum")
+wind <- wind %>% dplyr::select(shrid, days_sum)
+villages <- villages %>% left_join(wind, by = "shrid")
+villages <- villages %>% st_set_crs(st_crs(districts))
+
+ggplot() + 
+  theme_map() +
+  geom_sf(data = districts, fill = NA) +
+  geom_sf(data = villages %>% filter(is.na(days_sum)==F), aes(fill = days_sum), lwd = NA) +
+  scale_fill_viridis_c("exposure (days)", option = "viridis") +
+  labs(x = "lon",
+       y = "lat") +
+  theme(legend.position = c(0.5, 0.05),
+        legend.direction = "horizontal",
+        legend.key.width = unit(0.3, 'cm'),
+        legend.text = element_text(size = 8),
+        legend.background=element_blank() )
 
@@
 kable(
-      labortableold, caption = "Wind direction and labor allocation, heterogeneity by age",
+      labortableold,
       align = "ccccc", booktabs = TRUE, linesep = ""
@@
 
+## Large negative effects on older half
+
+- Older half of sample decreases overall labor allocation
+  - Driven by self-employment and nonfarm employment
+- Younger half actually increases allocation!
+- Consistent with health effects only on older adults, though not conclusive
+
+---
+
+## Effects on aggregate growth?
+
+- We know pollution affects productivity and labor allocation
+- Might it also affect growth?
+- Proxy for growth with nightlights
+  - Look at previous year's exposure and current year's nightlights
+  - "Growth" of nightlights
+  - Village FE
+
+---
+
 ## Wind direction and growth of nightlights
@@
 kable(
-      ntltable, caption = "Wind direction and nightlight growth",
+      ntltable,
       align = "cc", booktabs = TRUE, linesep = ""
@@
 ## Conclusion
+
+- Large growth in coal power in India
+  - Other countries, as well
+- Increase in electricity, but downsides as well
+  - Pollution!
+- We already have plenty of evidence on effects of pollution on health
+
+---
+
+## Conclusion
+
+- Paper presents evidence that villages downwind from coal plants suffer in terms of agricultural productivity
+  - Identification relies on within-village changes across seasons
+- Wind exposure fluctuations lead to changes in agricultural productivity of around 2.4 percent based on mean absolute deviation
+  - Instrumenting for particulate matter, effects are even larger (more than 20 percent)
+- Clear changes in labor allocation
+  - Oldest half of sample
+- Slower growth of nightlights
+
+---
+
+## Conclusion
+
+- Overall cost/benefit way beyond scope of paper
+- Instead, results point to importance of the location of pollution sources
+  - Downwind villages suffer
+- 30km radius indicates relatively long-range effects

### 2023-02-11 analyses/regressions.R

@@ -58,7 +58,8 @@ plantresultstable <- etable(
                           se.below = TRUE,
                           depvar = FALSE,
                           signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.1),
-                          digits = 3,
+                          digits = "r3",
+                          digits.stats = "r3",
                           fitstat = c("n"),
                           coefstat = "se",
                           extralines = list("Sub-sample" = c("all", "no plant", "all", "no plant"))
@@ -140,7 +141,8 @@ pollutiontable <- etable(
                           se.below = TRUE,
                           depvar = FALSE,
                           signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.1),
-                          digits = 3,
+                          digits = "r3",
+                          digits.stats = "r3",
                           fitstat = c("n"),
                           coefstat = "se"
                           )
@@ -223,7 +225,8 @@ yieldtable <- etable(
                       se.below = TRUE,
                       depvar = FALSE,
                       signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.1),
-                      digits = 3,
+                      digits = "r3",
+                      digits.stats = "r3",
                       fitstat = c("n"),
                       coefstat = "se"
                       #extralines = list("Sub-sample" = c("all", "all", "all", "monsoon", "winter"))
@@ -255,7 +258,8 @@ yieldtabletwo <- etable(
                         se.below = TRUE,
                         depvar = FALSE,
                         signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.1),
-                        digits = 3,
+                        digits = "r3",
+                        digits.stats = "r3",
                         fitstat = c("n"),
                         coefstat = "se"
                         #extralines = list("Sub-sample" = c("all", "all", "monsoon", "winter"))
@@ -266,7 +270,8 @@ yieldtabletwo1 <- etable(
                         se.below = TRUE,
                         depvar = FALSE,
                         signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.1),
-                        digits = 3,
+                        digits = "r3",
+                        digits.stats = "r3",
                         fitstat = c("ivwald"),
                         coefstat = "se"
[…]

### 2023-02-11 cleaning/8.cleaning_overlap_merging_nss.R

@@ -86,24 +86,40 @@ for (day in 1:length(date_vec)){
   village_wind <- villages %>% left_join(village_wind, by = "shrid")
   # Missings as zero
   village_wind$days_sum[is.na(village_wind$days_sum)==T] <- 0
+  # there are a couple mistakes in terms of duplicate ids. fix those
+  village_wind <- village_wind %>% group_by(shrid) %>% filter(row_number()==1) %>% ungroup()
   # And also pm
   pollution_merge <- pollution %>% filter(year==year(date_vec[day]) & month==month(date_vec[day]))
-  village_wind <- village_wind %>% left_join(pollution, by = "shrid")
+  # there are a couple mistakes in terms of duplicate ids. fix those
+  pollution_merge <- pollution_merge %>% group_by(shrid) %>% filter(row_number()==1) %>% ungroup()
+  pollution_merge <- pollution_merge %>% select(shrid, pm25)
+  village_wind <- village_wind %>% left_join(pollution_merge, by = "shrid")
   
+  # monthly wind
+  village_month_wind <- read_csv(paste0("data/clean/wind_ntl/months/y", year(date_vec[day]), "m", month(date_vec[day]), ".csv")) %>% as_tibble()
+  # there are a couple mistakes in terms of duplicate ids. fix those
+  village_month_wind <- village_month_wind %>% group_by(shrid) %>% filter(row_number()==1) %>% ungroup()
+  # rename
+  colnames(village_month_wind) <- c("shrid", "month_sum")
+  village_month_wind <- village_month_wind %>% select(shrid, month_sum)
+  village_wind <- village_wind %>% left_join(village_month_wind, by = "shrid")
   
   # Weighted mean by AREA. Any parts of the district without villages are ZEROS
   village_wind <- village_wind %>% 
                   group_by(state, district) %>%
-                  mutate(days_sum = weighted.mean(days_sum, area_weight),
-                         pm25_sum = weighted.mean(pm25, area_weight),
+                  mutate(days_sum = weighted.mean(days_sum, area_weight, na.rm = T),
+                         month_sum = weighted.mean(month_sum, area_weight, na.rm = T),
+                         pm25_sum = weighted.mean(pm25, area_weight, na.rm = T),
                          state_merge = as.numeric(state),
                          district_merge = as.numeric(district),
-                         tot_weight = sum(area_weight),
+                         tot_weight = sum(area_weight, na.rm = T),
[…]

### 2023-02-21 draft/draft.Rmd
lots of updates: draft, regressions, and new cleaning for wind

@@
 # Table
-yieldtablehet[1,5] <- "0.000"
-colnames(yieldtablehet) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
+colnames(yieldtablehet) <- c(">p(50)", "<=p(50)", ">p(50)", "<=p(50)", "")
+yieldtablehet[7:8,] <- "Yes"
 kable(
@@
   column_spec(c(2:6),width = "2cm") %>%
-  row_spec(c(15), hline_after = TRUE) %>%
-  row_spec(c(13), bold = TRUE) %>%
-  add_header_above(c(" ", ">p(50)" = 1, "<=p(50)" = 1, "all" = 3)) %>%
+  row_spec(c(8), hline_after = TRUE) %>%
+  row_spec(c(6), bold = TRUE) %>%
+  add_header_above(c(" ", "exposure" = 2, "yield" = 2, " " = 1)) %>%
   kable_classic_2()
@@
 kable(
-      labortablerural, caption = "Wind direction and labor allocation, rural only",
+      labortablerural, caption = "Wind direction and labor allocation, by location",
       align = "cccccc", booktabs = TRUE, linesep = ""
@@
 colnames(ntltable) <- c("(1)", "(2)", "(3)", "(4)")
+newrow <- matrix(c("no", "yes", "no", "yes"), nrow = 1)
+colnames(newrow) <- c("(1)", "(2)", "(3)", "(4)")
+rownames(newrow) <- "weather"
+ntltable <- rbind(ntltable[1:4,], newrow, ntltable[5:nrow(ntltable),])
 kable(
@@
   column_spec(c(2:5),width = "2cm") %>%
-  row_spec(c(9), hline_after = TRUE) %>%
-  row_spec(c(5,8), bold = TRUE) %>%
+  row_spec(c(10), hline_after = TRUE) %>%
+  row_spec(c(6,9), bold = TRUE) %>%
   kable_classic_2()

### 2023-02-21 draft/draft.tex
lots of updates: draft, regressions, and new cleaning for wind

@@
 \author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
-\date{2023-02-10}
+\date{2023-02-11}
 
@@
 \toprule
-\multicolumn{1}{c}{ } & \multicolumn{1}{c}{>p(50)} & \multicolumn{1}{c}{<=p(50)} & \multicolumn{3}{c}{all} \\
-\cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-6}
-  & (1) & (2) & (3) & (4) & (5)\\
-\midrule
-wind & -0.003*** & -0.033*** & -0.004*** & 0.0006 & 0.000\\
- & (0.0002) & (0.002) & (0.0003) & (0.0005) & (0.001)\\
-rain (z) & 0.028*** & 0.032*** & 0.028*** & 0.028*** & 0.028***\\
- & (0.0006) & (0.0005) & (0.0004) & (0.0004) & (0.0004)\\
-mean temp (10s) & -0.001*** & 0.0006*** & -0.0004*** & -0.0004*** & -0.0004***\\
- & (7.66e-5) & (7.9e-5) & (5.55e-5) & (5.56e-5) & (5.49e-5)\\
-wind x rain &  &  & 0.0005*** &  & \\
- &  &  & (6.74e-5) &  & \\
-wind squared &  &  &  & -0.0002*** & \\
- &  &  &  & (3.58e-5) & \\
-wind x starting yield &  &  &  &  & -0.001**\\
- &  &  &  &  & (0.0006)\\
+\multicolumn{1}{c}{ } & \multicolumn{2}{c}{exposure} & \multicolumn{2}{c}{yield} & \multicolumn{1}{c}{ } \\
+\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
+  & >p(50) & <=p(50) & >p(50) & <=p(50) & \\
+\midrule
+pm & -0.074*** & -0.012*** & -0.006*** & -0.066*** & -0.021***\\
+ & (0.004) & (0.002) & (0.001) & (0.006) & (0.001)\\
+pm (10s) times rain &  &  &  &  & 0.005***\\
+ &  &  &  &  & (0.000)\\
+weather & yes & yes & yes & yes & yes\\
 \textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
@@
 \midrule
-observations & 1,112,082 & 1,259,139 & 2,371,221 & 2,371,221 & 2,367,248\\
+observations & 650,770 & 1,720,451 & 1,192,644 & 1,178,577 & 2,371,221\\
 \bottomrule
@@
 
-\caption{\label{tab:labortablerural}Wind direction and labor allocation, rural only}
+\caption{\label{tab:labortablerural}Wind direction and labor allocation, by location}
 \centering
@@
  &  &  & (0.002) & (0.002)\\
+weather & no & yes & no & yes\\
 \textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\

### 2023-02-21 analyses/regressions.R

@@ -333,17 +333,23 @@ saveRDS(yieldtable, "pollution_development/draft/tables/yieldtablepm.rds")
 
 
 df <- df %>% group_by(shrid) %>% 
-                mutate(wind_max = max(wind)) %>% 
+                mutate(wind_med = mean(wind, na.rm = T),
+                       yield_med = mean(yield, na.rm = T)) %>% 
                 ungroup() %>% 
-                mutate(wind_max_median = median(wind_max))
+                mutate(wind_med_median = median(wind_max, na.rm = T),
+                       yield_med_median = median(yield_med, na.rm = T))
 
 
+df$pmrain <- (df$pm25*df$rain_z)*0.1 # divided by ten for the coefficient
+df$windrain <- df$wind*df$rain_z
+df$pmyield <- df$pm25*df$original_yield
+df$windyield <- df$wind*df$original_yield
 
-yield1 <- feols(log(yield) ~ wind + rain_z + temp_mean | village^season + year, data = df %>% filter(wind_max>wind_max_median), cluster = c("village"))
-yield2 <- feols(log(yield) ~ wind + rain_z + temp_mean | village^season + year, data = df %>% filter(wind_max<=wind_max_median), cluster = c("village"))
-yield3 <- feols(log(yield) ~ wind*rain_z + temp_mean | village^season + year, data = df, cluster = c("village"))
-yield4 <- feols(log(yield) ~ wind + wind^2 + rain_z + temp_mean | village^season + year, data = df, cluster = c("village"))
-yield5 <- feols(log(yield) ~ wind*original_yield + rain_z + temp_mean | village^season + year, data = df, cluster = c("village"))
+yield1 <- feols(log(yield) ~ rain_z + temp_mean | village^season + year | pm25 ~ wind, data = df %>% filter(wind_med>wind_med_median), cluster = c("village"))
+yield2 <- feols(log(yield) ~ rain_z + temp_mean | village^season + year | pm25 ~ wind, data = df %>% filter(wind_med<=wind_med_median), cluster = c("village"))
+yield3 <- feols(log(yield) ~ rain_z + temp_mean | village^season + year | pm25 ~ wind, data = df %>% filter(yield_med>yield_med_median), cluster = c("village"))
+yield4 <- feols(log(yield) ~ rain_z + temp_mean | village^season + year | pm25 ~ wind, data = df %>% filter(yield_med<=yield_med_median), cluster = c("village"))
[…]

### 2023-02-21 draft/draft.Rmd
lots of updates!

--- 2023-02-11 draft/draft.Rmd
@@
 yieldtablehet <- readRDS("pollution_development/draft/tables/yieldtablehet.rds")
+yieldtablehet1st <- readRDS("pollution_development/draft/tables/yieldtablehet1st.rds")
+yieldtablemonthly <- readRDS("pollution_development/draft/tables/yieldtablemonthly.rds")
 pmeffects <- readRDS("pollution_development/draft/tables/yieldtablePM.rds")
@@
 labortableurban <- readRDS("pollution_development/draft/tables/labortableurban.rds")
+labortableruralyoung <- readRDS("pollution_development/draft/tables/labortableruralyoung.rds")
+labortableruralold <- readRDS("pollution_development/draft/tables/labortableruralold.rds")
 wagestable <- readRDS("pollution_development/draft/tables/wagestable.rds")
@@
 laborplot <- readRDS("pollution_development/draft/tables/laborplot.rds")
+monthplot <- readRDS("pollution_development/draft/tables/monthplot.rds")
 
@@
 Table \@ref(tab:yieldtablepm) presents the same regressions as Table \@ref(fig:yieldtable), with a key difference: instead of wind exposure, it uses average concentration of $\mathrm{PM_{2.5}}$ -- from @hammer2020global -- during the season as the key explanatory variable.
-I rescale particulate matter by dividing by one thousand to help with presentation of the coefficients.
+
 Several interesting patterns emerge.
-In the first column we see no relationship between particulate matter and agricultural productivity, while there is actually a positive relationship in columns two, four, and five.
-There is a small, negative relationship between pollution and productivity in column three.
+We only observe a negative relationship between particulate matter and agricultural productivity in column three; the coefficient in column two is actually positive.
+Breaking out the effects into the monsoon and winter seasons, we also see a positive relationship between pollution and agricultural productivity.
 
@@
 This makes sense, of course, since the IV is essentially scaling up the effects of wind.
+
 The comparison of the results to those in the previous table also confirm the concerns around endogenous placement of coal plants.
 The coefficients in Table \@ref(tab:yieldtabletwo) are always more negative than those in Table \@ref(tab:yieldtablepm), and sometimes substantially so.
-In other words, the simple regression of agricultural productivity on pollution results in coefficients on pollution that are upwards biased, since coal plants are placed in more productive areas.
+In other words, the simple regression of agricultural productivity on pollution results in coefficients on pollution that are upwards biased, since higher levels of pollution occur in more productive areas, on average.
 The IV strategy corrects this bias and leads to a robust, negative relationship between pollution and agricultural productivity.
@@
 
+## Heterogeneity
+
 Having established that exposure to pollution has a negative effect on agricultural productivity, I now move to heterogeneity in these effects.
-Table \@ref(tab:yieldtablehet) presents four different analyses: effects of exposure by maximum exposure, multiplicative effects of shocks, the possibility of non-linearities, and the differential effects of exposure based on initial agricultural productivity.
-
-The first two columns split the sample based on maximum exposure; villages with maximum exposure above the median are included in column (1) while those below the median are in column (2).
-Interestingly, the largest effects are seen on villages in the lower half of the distribution.
-Apparently, smaller maximum effects lead to larger overall effects of pollution exposure on agricultural productivity.
-In other words, it is not the places where wind is most likely to blow from a coal plant that experience the largest decrease in productivity in response to an additional day of wind-driven pollution.
+Table \@ref(tab:yieldtablehet) presents three different analyses: heterogeneity by median exposure, heterogeneity by initial yield, and possible multiplicative effects of shocks.
+All regressions use the IV strategy detailed above and I include the F statistic from the first stage regressions (of which there are two in the last column) at the bottom of the table.
+Table \@ref(tab:yieldtablehet1st) in the appendix presents the first-stage results for all columns.
+
+The first two columns split the sample based on median exposure.
+In other words, for each village, I calculate the median exposure (based on wind) and then use this single value for each village to split the sample into villages above and below median exposure.
+The coefficient in the first column is 2.5 times larger than the coefficient in the second column, indicating much larger effects of pollution in places where it is more common.
+One possible explanation is that exposure is cumulative, leading to larger and larger effects.
+I return to this in column five.
+
+Columns two and three split the sample based on median yield at the very beginning of the sample (the first year available, 2002-2003).
+Column three presents results for villages above the median while column four presents results for villages below the median.
+The largest effects are apparently concentrated in the lower half of the distribution.
+Perhaps more productive areas are more likely to have ways to combat the effects of shocks -- e.g.
+by adjusting the use of other inputs -- or maybe less productive areas are inherently more susceptible to the effects of pollution, perhaps due to direct effects on land productivity.
+
+Given ongoing changes in the climate, a key question is how welfare will change as shocks become more and more common.
+In the last column looks at possible multiplicative effects of shocks by interacting pollution with rainfall and seeing whether the effects compound one another.
+Since there are now two endogenous regressors, I instrument for both using wind exposure and an interaction between wind exposure and rainfall.
 
@@
 # Table
-yieldtablehet[1,5] <- "0.000"
-colnames(yieldtablehet) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
+colnames(yieldtablehet) <- c(">p(50)", "<=p(50)", ">p(50)", "<=p(50)", "")
+rownames(yieldtablehet)[c(1, 3)] <- c("PM 2.5", "PM time rain")
 kable(
@@
 The outcome in all regressions is estimated agricultural yield, defined as tons per hectare.
-The first two columns split the sample based on the *maximum* observed exposure.
+The first two columns split the sample based on the median exposure (by village).
+The second two columns split the sample based on initial -- the first year available -- agricultural productivity.
 * p<0.10 ** p<0.05 *** p<0.01",
@@
            ) %>%
-  column_spec(1, width = "3cm") %>%
-  column_spec(c(2:6),width = "2cm") %>%
-  row_spec(c(15), hline_after = TRUE) %>%
-  row_spec(c(13), bold = TRUE) %>%
-  add_header_above(c(" ", ">p(50)" = 1, "<=p(50)" = 1, "all" = 3)) %>%
-  kable_classic_2()
-
-```
-
-Column three looks add the possible multiplicative effects of rainfall shocks and pollution shocks on agricultural productivity.
-In previous tables, rainfall always entered the regression positively -- as we would expect -- indicating that higher rainfall in a given season leads to higher agricultural productivity.
-However, an interesting question is whether rainfall and pollution shocks might interact.
-The positive coefficient on the interaction term between days of wind and rainfall during the agricultural season indicates that there is a multiplicative effect of shocks; when wind days are higher (bad) and rainfall is lower (bad), the overall negative effects are even worse than when just one shock happens.
-Considering that climate change is leading to more variable rainfall and more negative rainfall shocks, these results are potentially worrisome, especially for smallholders.
-
-Column four adds a quadratic in wind to see if the negative effects of pollution on agricultural productivity are increase or decreasing in wind.
-In column four, the coefficient on the linear term is 0.0006 while the coefficient on the quadratic term is -0.0002, meaning that the effect of wind is decreasing for any days of wind over three The negative coefficient also indicates that this negative effect is consistently decreasing, meaning the effect gets worse as average wind days gets higher.
-Note that this is somewhat different from the regressions in the first two columns, which split the sample by the *maximum* days of wind observed in the sample, not the mean.
-
-In the final column I add to the regression an interaction between wind and initial agricultural productivity (defined in 2002, the first year observed in the productivity data).
-The linear term drops out of the regression since it is invariant within villages, leaving just the interaction term.
-The negative coefficient on the interaction term of -0.001 indicates that the effect of pollution is more negative in areas where agricultural productivity is higher.
-Note that agricultural productivity is defined in logs, so this can be interpreted as a percentage effect, not a level effect.
-Apparently, wind-driven pollution is particularly detrimental to agricultural productivity in the most productive places.
+  column_spec(1, width = "4.5cm") %>%
+  column_spec(c(2:6),width = "1.8cm") %>%
+  row_spec(c(8), hline_after = TRUE) %>%
+  row_spec(c(6), bold = TRUE) %>%
+  add_header_above(c(" ", "exposure" = 2, "yield" = 2, " " = 1)) %>%
+  kable_classic_2()
+
+```
+
+The coefficient on the level term for pollution is negative.
+Since rainfall is defined as a z score, this represents the effect of pollution only when rainfall is at its mean value for a given season.
+The interaction term, on the other hand, is positive.
+Since higher values of rainfall are generally good, this positive coefficient represents a compounding effect of pollution and rainfall; when rainfall is low (bad), the effects of particulate matter on agricultural productivity is larger (more negative).
+Another way to think about this, however, is that higher values of rainfall dampen the effects of pollution shocks.
+This could have very profound effects on welfare, since this essentially increases the variance of agricultural productivity across years (which is already increasing due to climate change).
 
@@
 This drops any pre-plant-opening observations -- all of which are mechanically zeros for wind -- to prevent endogeneity due to the opening.
-The coefficients in Table \@ref(tab:yieldtablepostplant) show that the overall effects are essentially unchanged, with just a small change in coefficient magnitude (driven mostly by rounding).
-
-The second robustness check includes leads for wind to see if they predict current year productivity.
+The coefficients in Table \@ref(tab:yieldtablepostplant) show that the overall effects are essentially unchanged across any specifications.
+Comparing changes in coefficients can also help reduce any effects driven by other polluting industries located near coal plants; since the pre-opening exposure variable is zero, the coefficients in Table \@ref(tab:yieldtablepostplant) would likely increase in magnitude if the effects were driven by non-coal pollution.
+
+The second robustness check includes leads and lags for wind to see if they predict current year productivity.
 I present these results in \@ref(tab:yieldtableleads).
-The second column includes district-year-season fixed effects, while the first does not.
-The results make clear the importance of including trends or district-year fixed effects.
-In the first column, leads seem to predict productivity.
-However, this relationship disappears in the second column, indicating that the previous results with district-year fixed effects are likely the most reliable.
+The first column is the reduced form effect and shows that only current wind predicts productivity; leads and lags have no impact.
+We see similar results in column two -- the IV results -- with a negative coefficient for current pollution, only.
+However, the standard errors are much larger than the main results -- around five times as large -- making it hard to come to any conclusions.
 
@@
 kable(
-      labortablerural, caption = "Wind direction and labor allocation, rural only",
+      labortablerural, caption = "Wind direction and labor allocation by rural status",
       align = "cccccc", booktabs = TRUE, linesep = ""
@@
 
+```{r labortableruralage, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+# Table
+colnames(labortableruralyoung) <- c("all", "all", "self", "wage", "farm", "non-farm")
+colnames(labortableruralold) <- c("all", "all", "self", "wage", "farm", "non-farm")
+
+mat1 <- matrix(c("", "", "", "", "", ""), nrow = 1)
+colnames(mat1) <- c("all", "all", "self", "wage", "farm", "non-farm")
+rownames(mat1) <- c("younger than 35")
+mat2 <- matrix(c("", "", "", "", "", ""), nrow = 1)
+colnames(mat2) <- c("all", "all", "self", "wage", "farm", "non-farm")
+rownames(mat2) <- c("45 or older")
+
+labortableruralyoung <- rbind(mat1, labortableruralyoung[c(1:2, 7:8),], mat2, labortableruralold[-c(3:6),], labortableruralold[c(3:6),])
+
+kable(
+      labortableruralyoung, caption = "Wind direction and labor allocation by age (rural only)",
+      align = "cccccc", booktabs = TRUE, linesep = ""
+      ) %>%
+  footnote(
+           "Note: Standard errors are in parentheses and are clustered at the district level.
+Control variables include female, age, age squared, and (years of) education.
+The outcome in each regression is days worked in the previous seven days.
+* p<0.10 ** p<0.05 *** p<0.01",
+           general_title = "",
+           threeparttable = TRUE,
+           footnote_as_chunk = TRUE
+           ) %>%
+  column_spec(1, width = "3cm") %>%
+  column_spec(c(2:7),width = "1.5cm") %>%
+  row_spec(c(5,10), hline_after = TRUE) %>%
+  row_spec(c(1,6,12), bold = TRUE) %>%
+  kable_classic_2()
+
+```
+
 To put some of these changes in context, an additional day of exposure to pollution from wind increases self-employment for the younger half of the sample by around six percent of the mean and increases farm employment for the younger half by almost 10 percent of the mean.
@@
 
-## Nightlights
+## Month-specific exposure
 
@@
 
-```{r ntltable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-# Table
-colnames(ntltable) <- c("(1)", "(2)", "(3)", "(4)")
-kable(
-      ntltable, caption = "Wind direction and nightlight growth",
-      align = "cccc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the village level.
-The outcome in all regressions is logged nightlights.
-Rainfall is included in columns (2) and (4).
-* p<0.10 ** p<0.05 *** p<0.01",
+```{r yieldtablemonthly, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+# Table
+
+kable(
+      yieldtablemonthly, caption = "Month-specific exposure",
+      align = "cc", booktabs = TRUE, linesep = ""
+      ) %>%
+  footnote(
+           "Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions is logged nightlights. * p<0.10 ** p<0.05 *** p<0.01", 
            general_title = "",
@@
   column_spec(1, width = "3cm") %>%
-  column_spec(c(2:5),width = "2cm") %>%
-  row_spec(c(9), hline_after = TRUE) %>%
-  row_spec(c(5,8), bold = TRUE) %>%
+  column_spec(c(2:3),width = "2cm") %>%
+  row_spec(c(13), hline_after = TRUE) %>%
+  row_spec(c(11), bold = TRUE) %>%
   kable_classic_2()
@@
 
-data <- matrix(NA, nrow = 7, ncol = 3)
-rownames(data) <- c("shapefile", "coal plants", "wind", "pollution", "agriculture", "nightlights", "weather")
+data <- matrix(NA, nrow = 8, ncol = 3)
+rownames(data) <- c("shapefile", "coal plants", "wind", "pollution", "agriculture", "nightlights", "weather", "labor")
 colnames(data) <- c("source", "geographic coverage", "temporal coverage")
@@
 data[7,1] <- "TerraClimate"
+data[8,1] <- "National Sample Survey (NSS)"
 data[1,2] <- "India"
@@
 data[4,2] <- "global"
-data[5,2] <- "global"
+data[5,2] <- "India"
 data[6,2] <- "India"
 data[7,2] <- "global"
+data[8,2] <- "India"
 data[1,3] <- " "
@@
 data[7,3] <- "monthly"
+data[8,3] <- "daily interviews"
 
@@
   column_spec(1, width = "2cm") %>%
-  column_spec(2,width = "4.5cm") %>%
-  column_spec(c(3:4),width = "3.5cm") %>%
+  column_spec(2,width = "5cm") %>%
+  column_spec(c(3:4),width = "3cm") %>%
   kable_styling(latex_options = "HOLD_position") %>%
+  kable_classic_2()
+
+```
+
+```{=tex}
+\FloatBarrier
+\newpage
+```
+```{r yieldtablehet1st, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+# Table
+colnames(yieldtablehet1st) <- c("PM", "PM", "PM", "PM", "PM", "PM*rain")
+kable(
+      yieldtablehet1st, caption = "Heterogeneity in the effects of pollution on productivity, first-stage results",
+      align = "cccccc", booktabs = TRUE, linesep = ""
+      ) %>%
+  footnote(
+           "Note: Standard errors are in parentheses and are clustered at the village level.
+The outcome in all regressions is estimated agricultural yield, defined as tons per hectare.
+The first two columns split the sample based on the median exposure (by village).
+The second two columns split the sample based on initial -- the first year available -- agricultural productivity.
+* p<0.10 ** p<0.05 *** p<0.01",
+           general_title = "",
+           threeparttable = TRUE,
+           footnote_as_chunk = TRUE
+           ) %>%
+  column_spec(1, width = "2.8cm") %>%
+  column_spec(c(2:7),width = "1.7cm") %>%
+  row_spec(c(8), hline_after = TRUE) %>%
+  row_spec(c(6), bold = TRUE) %>%
+  add_header_above(c(" ", "(1)" = 1, "(2)" = 1, "(3)" = 1, "(4)" = 1, "(5)" = 2)) %>%
   kable_classic_2()
@@
   column_spec(c(2:6),width = "2cm") %>%
-  row_spec(7, bold = TRUE) %>%
-  row_spec(11, hline_after = TRUE) %>%
+  row_spec(c(7, 13), bold = TRUE) %>%
+  row_spec(c(11:12), hline_after = TRUE) %>%
   add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
@@
 ```
+
 ```{r yieldtableleads, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
@@
   column_spec(c(2:3),width = "2cm") %>%
-  row_spec(c(14), hline_after = TRUE) %>%
-  row_spec(c(11), bold = TRUE) %>%
+  row_spec(c(16), hline_after = TRUE) %>%
+  row_spec(c(14), bold = TRUE) %>%
   kable_classic_2()
@@
 ```
+
+```{=tex}
+\FloatBarrier
+\newpage
+```
+```{r monthplot, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Average farm days by season, rural only", fig.footer = list("Note: The line graph shows the mean number of farm days worked per week based on the month of the relevant agricultural season.
+The monsoon season is defined as June (month 1) to October (month 5) while the rainy season is defined as November (month 1) to March (month 5) of the following year.
+The data is from the NSS.")}
+
+monthplot
+
+```

### 2023-02-21 draft/draft.tex
lots of updates!

@@
 \author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
-\date{2023-02-11}
+\date{2023-02-20}
 
@@
 \midrule
-pop (log) & 0.068*** & 0.0003 & 0.009*** & 0.002***\\
- & (0.0002) & (0.0002) & (0.0003) & (0.0002)\\
-literacy (prop) & 0.127*** & -0.005*** & 0.043*** & 0.030***\\
+pop (log) & 0.068*** & 0.000 & 0.009*** & 0.002***\\
+ & (0.000) & (0.000) & (0.000) & (0.000)\\
+literacy (prop) & 0.127*** & -0.005*** & 0.044*** & 0.030***\\
  & (0.002) & (0.001) & (0.002) & (0.002)\\
-has paved road & -0.036*** & 0.003*** & -0.008*** & -2.68e-7\\
- & (0.0008) & (0.0004) & (0.0008) & (0.0006)\\
+has paved road & -0.037*** & 0.003*** & -0.008*** & -0.000\\
+ & (0.001) & (0.000) & (0.001) & (0.001)\\
 ag productivity &  &  & 0.033*** & 0.005***\\
- &  &  & (0.0005) & (0.0004)\\
+ &  &  & (0.001) & (0.000)\\
 sub-sample & all & no plant & all & no plant\\
@@
 \midrule
-wind & 0.045*** & 0.014*** & 0.063*** & 0.015***\\
- & (0.004) & (0.001) & (0.005) & (0.002)\\
+wind (10s) & 0.002*** & 0.002*** & 0.006*** & 0.001***\\
+ & (0.000) & (0.000) & (0.001) & (0.000)\\
 \textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
@@
 \midrule
-observations & 22,345,092 & 22,345,092 & 14,896,728 & 14,896,728\\
+observations & 24,376,458 & 24,376,458 & 16,250,973 & 16,250,973\\
 \bottomrule
@@
 Across all four columns, the story is the same: when the wind is blowing in the direction of a village from a coal plant within 30 kilometers, estimated \(\mathrm{PM_{2.5}}\) is substantially higher.
-The regressions indicate that one additional day of wind in the direction of the village increases mean \(\mathrm{PM_{2.5}}\) for the month by between 0.01 and 0.06 \(\mathrm{\mu g/m^3}\).
+The regressions indicate that one additional day of wind in the direction of the village increases mean \(\mathrm{PM_{2.5}}\) for the month by between 0.00 and 0.00 \(\mathrm{\mu g/m^3}\).
 
@@
 \midrule
-wind & -0.003*** & -0.003*** & -0.0007*** & -0.002*** & -0.004***\\
- & (0.0002) & (0.0002) & (9.24e-5) & (0.0002) & (0.0003)\\
+wind & -0.003*** & -0.003*** & -0.001*** & -0.002*** & -0.004***\\
+ & (0.000) & (0.000) & (0.000) & (0.000) & (0.000)\\
 rain (z) &  & 0.029*** & 0.008*** & 0.075*** & 0.006***\\
- &  & (0.0004) & (0.001) & (0.002) & (0.0005)\\
-mean temp (10s) &  & -0.0004*** & -0.007*** & -0.002*** & -0.008***\\
- &  & (5.53e-5) & (0.0004) & (0.0001) & (9.83e-5)\\
+ &  & (0.000) & (0.001) & (0.002) & (0.000)\\
+mean temp (10s) &  & -0.000*** & -0.007*** & -0.002*** & -0.008***\\
+ &  & (0.000) & (0.000) & (0.000) & (0.000)\\
 \textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
@@
 \midrule
-particulate matter & 0.0000 & 0.0005*** & -0.0004*** & 0.002*** & 0.0004***\\
-(PM 2.5, '000s) & (0.0000) & (3.17e-5) & (5.72e-5) & (9.87e-5) & (3.7e-5)\\
-rain (z) &  & 0.030*** & 0.008*** & 0.075*** & 0.007***\\
- &  & (0.0004) & (0.001) & (0.002) & (0.0005)\\
-mean temp (10s) &  & -0.0003*** & -0.007*** & -0.002*** & -0.008***\\
- &  & (5.63e-5) & (0.0004) & (0.0001) & (9.79e-5)\\
+particulate matter & 0.0000 & 0.0329*** & -0.0787*** & -0.0509*** & 0.0216***\\
+(log PM 2.5) & (0.0000) & (0.0028) & (0.0040) & (0.0033) & (0.0052)\\
+rain (z) &  & 0.0290*** & 0.0079*** & 0.0760*** & 0.0064***\\
+ &  & (0.0004) & (0.0013) & (0.0017) & (0.0005)\\
+mean temp (10s) &  & -0.0003*** & -0.0071*** & -0.0020*** & -0.0080***\\
+ &  & (0.0001) & (0.0004) & (0.0001) & (0.0001)\\
 \textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
@@
 (2020) -- during the season as the key explanatory variable.
-I rescale particulate matter by dividing by one thousand to help with presentation of the coefficients.
+
 Several interesting patterns emerge.
-In the first column we see no relationship between particulate matter and agricultural productivity, while there is actually a positive relationship in columns two, four, and five.
-There is a small, negative relationship between pollution and productivity in column three.
+We only observe a negative relationship between particulate matter and agricultural productivity in column three; the coefficient in column two is actually positive.
+Breaking out the effects into the monsoon and winter seasons, we also see a positive relationship between pollution and agricultural productivity.
 
@@
 This makes sense, of course, since the IV is essentially scaling up the effects of wind.
+
 The comparison of the results to those in the previous table also confirm the concerns around endogenous placement of coal plants.
 The coefficients in Table \ref{tab:yieldtabletwo} are always more negative than those in Table \ref{tab:yieldtablepm}, and sometimes substantially so.
-In other words, the simple regression of agricultural productivity on pollution results in coefficients on pollution that are upwards biased, since coal plants are placed in more productive areas.
+In other words, the simple regression of agricultural productivity on pollution results in coefficients on pollution that are upwards biased, since higher levels of pollution occur in more productive areas, on average.
 The IV strategy corrects this bias and leads to a robust, negative relationship between pollution and agricultural productivity.
@@
 \midrule
-particulate matter & -0.021*** & -0.024*** & -0.032*** & -0.012*** & -0.035***\\
-(PM 2.5) & (0.001) & (0.002) & (0.005) & (0.001) & (0.003)\\
-rain (z) &  & -0.011*** & -4.24e-5 & 0.078*** & -0.042***\\
- &  & (0.003) & (0.002) & (0.002) & (0.004)\\
-mean temp (10s) &  & -0.005*** & -0.007*** & -0.002*** & -0.008***\\
- &  & (0.0003) & (0.0004) & (0.0001) & (0.0001)\\
+particulate matter & -1.190*** & -1.027*** & -1.671*** & -0.525*** & -1.741***\\
+(log PM 2.5) & (0.070) & (0.071) & (0.235) & (0.043) & (0.139)\\
+rain (z) &  & 0.028*** & 0.012*** & 0.080*** & 0.001**\\
+ &  & (0.000) & (0.001) & (0.002) & (0.001)\\
+mean temp (10s) &  & -0.001*** & -0.009*** & -0.002*** & -0.007***\\
+ &  & (0.000) & (0.000) & (0.000) & \vphantom{2} (0.000)\\
 \textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
@@
 \textbf{first stage:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-wind & 0.145*** & 0.111*** & 0.022*** & 0.157*** & 0.105***\\
- & (0.002) & (0.002) & (0.002) & (0.002) & (0.004)\\
-rain (z) &  & -1.68*** & -0.239*** & 0.225*** & -1.37***\\
- &  & (0.011) & (0.017) & (0.015) & (0.010)\\
-mean temp (10s) &  & -0.192*** & 0.002 & -0.021*** & -0.007***\\
- &  & (0.0009) & (0.004) & (0.001) & (0.002)\\
+wind & 0.003*** & 0.003*** & 0.000*** & 0.004*** & 0.002***\\
+ & (0.000) & (0.000) & (0.000) & (0.000) & (0.000)\\
+rain (z) &  & -0.001*** & 0.002*** & 0.009*** & -0.003***\\
+ &  & (0.000) & (0.000) & (0.000) & \vphantom{1} (0.000)\\
+mean temp (10s) &  & -0.000*** & -0.001*** & 0.001*** & 0.001***\\
+ &  & (0.000) & (0.000) & (0.000) & (0.000)\\
 \bottomrule
@@
 
+\hypertarget{heterogeneity}{%
+\subsection{Heterogeneity}\label{heterogeneity}}
+
 Having established that exposure to pollution has a negative effect on agricultural productivity, I now move to heterogeneity in these effects.
-Table \ref{tab:yieldtablehet} presents four different analyses: effects of exposure by maximum exposure, multiplicative effects of shocks, the possibility of non-linearities, and the differential effects of exposure based on initial agricultural productivity.
-
-The first two columns split the sample based on maximum exposure; villages with maximum exposure above the median are included in column (1) while those below the median are in column (2).
-Interestingly, the largest effects are seen on villages in the lower half of the distribution.
-Apparently, smaller maximum effects lead to larger overall effects of pollution exposure on agricultural productivity.
-In other words, it is not the places where wind is most likely to blow from a coal plant that experience the largest decrease in productivity in response to an additional day of wind-driven pollution.
+Table \ref{tab:yieldtablehet} presents three different analyses: heterogeneity by median exposure, heterogeneity by initial yield, and possible multiplicative effects of shocks.
+All regressions use the IV strategy detailed above and I include the F statistic from the first stage regressions (of which there are two in the last column) at the bottom of the table.
+Table \ref{tab:yieldtablehet1st} in the appendix presents the first-stage results for all columns.
+
+The first two columns split the sample based on median exposure.
+In other words, for each village, I calculate the median exposure (based on wind) and then use this single value for each village to split the sample into villages above and below median exposure.
+The coefficient in the first column is 2.5 times larger than the coefficient in the second column, indicating much larger effects of pollution in places where it is more common.
+One possible explanation is that exposure is cumulative, leading to larger and larger effects.
+I return to this in column five.
+
+Columns two and three split the sample based on median yield at the very beginning of the sample (the first year available, 2002-2003).
+Column three presents results for villages above the median while column four presents results for villages below the median.
+The largest effects are apparently concentrated in the lower half of the distribution.
+Perhaps more productive areas are more likely to have ways to combat the effects of shocks -- e.g.~by adjusting the use of other inputs -- or maybe less productive areas are inherently more susceptible to the effects of pollution, perhaps due to direct effects on land productivity.
+
+Given ongoing changes in the climate, a key question is how welfare will change as shocks become more and more common.
+In the last column looks at possible multiplicative effects of shocks by interacting pollution with rainfall and seeing whether the effects compound one another.
+Since there are now two endogenous regressors, I instrument for both using wind exposure and an interaction between wind exposure and rainfall.
 
@@
 \begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4.5cm}>{\centering\arraybackslash}p{1.8cm}>{\centering\arraybackslash}p{1.8cm}>{\centering\arraybackslash}p{1.8cm}>{\centering\arraybackslash}p{1.8cm}>{\centering\arraybackslash}p{1.8cm}}
 \toprule
@@
 \midrule
-pm & -0.074*** & -0.012*** & -0.006*** & -0.066*** & -0.021***\\
- & (0.004) & (0.002) & (0.001) & (0.006) & (0.001)\\
-pm (10s) times rain &  &  &  &  & 0.005***\\
- &  &  &  &  & (0.000)\\
-weather & yes & yes & yes & yes & yes\\
+PM 2.5 & -1.082*** & -2.523*** & -0.439*** & -1.711*** & -0.809***\\
+ & (0.073) & (0.624) & (0.046) & (0.146) & (0.060)\\
+PM time rain &  &  &  &  & 0.180***\\
+ &  &  &  &  & (0.008)\\
+weather & Yes & Yes & Yes & Yes & Yes\\
 \textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
@@
 \midrule
-observations & 650,770 & 1,720,451 & 1,192,644 & 1,178,577 & 2,371,221\\
+F (1st stage, PM) & 6,488 & 60 & 2,867 & 2,150 & 2,958\\
+F (1st stage, PM times rain) & -- & -- & -- & -- & 1,697\\
+observations & 1,051,010 & 1,320,211 & 1,193,764 & 1,177,457 & 2,371,221\\
 \bottomrule
@@
 The outcome in all regressions is estimated agricultural yield, defined as tons per hectare.
-The first two columns split the sample based on the *maximum* observed exposure.
+The first two columns split the sample based on the median exposure (by village).
+The second two columns split the sample based on initial -- the first year available -- agricultural productivity.
 * p<0.10 ** p<0.05 *** p<0.01
@@
 
-Column three looks add the possible multiplicative effects of rainfall shocks and pollution shocks on agricultural productivity.
-In previous tables, rainfall always entered the regression positively -- as we would expect -- indicating that higher rainfall in a given season leads to higher agricultural productivity.
-However, an interesting question is whether rainfall and pollution shocks might interact.
-The positive coefficient on the interaction term between days of wind and rainfall during the agricultural season indicates that there is a multiplicative effect of shocks; when wind days are higher (bad) and rainfall is lower (bad), the overall negative effects are even worse than when just one shock happens.
-Considering that climate change is leading to more variable rainfall and more negative rainfall shocks, these results are potentially worrisome, especially for smallholders.
-
-Column four adds a quadratic in wind to see if the negative effects of pollution on agricultural productivity are increase or decreasing in wind.
-In column four, the coefficient on the linear term is 0.0006 while the coefficient on the quadratic term is -0.0002, meaning that the effect of wind is decreasing for any days of wind over three The negative coefficient also indicates that this negative effect is consistently decreasing, meaning the effect gets worse as average wind days gets higher.
-Note that this is somewhat different from the regressions in the first two columns, which split the sample by the \emph{maximum} days of wind observed in the sample, not the mean.
-
-In the final column I add to the regression an interaction between wind and initial agricultural productivity (defined in 2002, the first year observed in the productivity data).
-The linear term drops out of the regression since it is invariant within villages, leaving just the interaction term.
-The negative coefficient on the interaction term of -0.001 indicates that the effect of pollution is more negative in areas where agricultural productivity is higher.
-Note that agricultural productivity is defined in logs, so this can be interpreted as a percentage effect, not a level effect.
-Apparently, wind-driven pollution is particularly detrimental to agricultural productivity in the most productive places.
+The coefficient on the level term for pollution is negative.
+Since rainfall is defined as a z score, this represents the effect of pollution only when rainfall is at its mean value for a given season.
+The interaction term, on the other hand, is positive.
+Since higher values of rainfall are generally good, this positive coefficient represents a compounding effect of pollution and rainfall; when rainfall is low (bad), the effects of particulate matter on agricultural productivity is larger (more negative).
+Another way to think about this, however, is that higher values of rainfall dampen the effects of pollution shocks.
+This could have very profound effects on welfare, since this essentially increases the variance of agricultural productivity across years (which is already increasing due to climate change).
 
@@
 This drops any pre-plant-opening observations -- all of which are mechanically zeros for wind -- to prevent endogeneity due to the opening.
-The coefficients in Table \ref{tab:yieldtablepostplant} show that the overall effects are essentially unchanged, with just a small change in coefficient magnitude (driven mostly by rounding).
-
-The second robustness check includes leads for wind to see if they predict current year productivity.
+The coefficients in Table \ref{tab:yieldtablepostplant} show that the overall effects are essentially unchanged across any specifications.
+Comparing changes in coefficients can also help reduce any effects driven by other polluting industries located near coal plants; since the pre-opening exposure variable is zero, the coefficients in Table \ref{tab:yieldtablepostplant} would likely increase in magnitude if the effects were driven by non-coal pollution.
+
+The second robustness check includes leads and lags for wind to see if they predict current year productivity.
 I present these results in \ref{tab:yieldtableleads}.
-The second column includes district-year-season fixed effects, while the first does not.
-The results make clear the importance of including trends or district-year fixed effects.
-In the first column, leads seem to predict productivity.
-However, this relationship disappears in the second column, indicating that the previous results with district-year fixed effects are likely the most reliable.
+The first column is the reduced form effect and shows that only current wind predicts productivity; leads and lags have no impact.
+We see similar results in column two -- the IV results -- with a negative coefficient for current pollution, only.
+However, the standard errors are much larger than the main results -- around five times as large -- making it hard to come to any conclusions.
 
@@
 
-\caption{\label{tab:labortablerural}Wind direction and labor allocation, by location}
+\caption{\label{tab:labortablerural}Wind direction and labor allocation by rural status}
 \centering
@@
 
+\begin{table}
+
+\caption{\label{tab:labortableruralage}Wind direction and labor allocation by age (rural only)}
+\centering
+\begin{threeparttable}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}}
+\toprule
+  & all & all & self & wage & farm & non-farm\\
+\midrule
+\textbf{younger than 35} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+wind & -0.077* & -0.085** & -0.061* & -0.024 & -0.001 & -0.085**\\
+ & (0.046) & (0.034) & (0.032) & (0.032) & (0.034) & (0.038)\\
+DV mean & 3.106 & 3.106 & 1.794 & 1.312 & 1.061 & 2.044\\
+observations & 263,965 & 263,923 & 263,923 & 263,923 & 263,923 & 263,923\\
+\midrule
+\textbf{45 or older} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+wind & -0.131** & -0.165*** & -0.054 & -0.111*** & -0.070* & -0.095**\\
+ & (0.054) & (0.043) & (0.054) & (0.041) & (0.036) & (0.042)\\
+DV mean & 3.584 & 3.584 & 2.481 & 1.103 & 1.444 & 2.140\\
+observations & 166,184 & 166,155 & 166,155 & 166,155 & 166,155 & 166,155\\
+\midrule
+controls & No & Yes & Yes & Yes & Yes & Yes\\
+\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+district & Yes & Yes & Yes & Yes & Yes & Yes\\
+state-month & Yes & Yes & Yes & Yes & Yes & Yes\\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}[para]
+\item Note: Standard errors are in parentheses and are clustered at the district level.
+Control variables include female, age, age squared, and (years of) education.
+The outcome in each regression is days worked in the previous seven days.
+* p<0.10 ** p<0.05 *** p<0.01
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+
 To put some of these changes in context, an additional day of exposure to pollution from wind increases self-employment for the younger half of the sample by around six percent of the mean and increases farm employment for the younger half by almost 10 percent of the mean.
@@
 
-\hypertarget{nightlights}{%
-\subsection{Nightlights}\label{nightlights}}
+\hypertarget{month-specific-exposure}{%
+\subsection{Month-specific exposure}\label{month-specific-exposure}}
 
@@
 
-\caption{\label{tab:ntltable}Wind direction and nightlight growth}
-\centering
-\begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
-\toprule
-  & (1) & (2) & (3) & (4)\\
-\midrule
-wind (lag) & -0.024*** & -0.024*** &  & \\
- & (0.003) & (0.003) &  & \\
-wind (lead) &  &  & -0.001 & -0.001\\
- &  &  & (0.002) & (0.002)\\
-weather & no & yes & no & yes\\
-\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-village & Yes & Yes & Yes & Yes\\
-year & Yes & Yes & Yes & Yes\\
-\textbf{varying slopes:} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-year (by village) & Yes & Yes & Yes & Yes\\
-\midrule
-observations & 1,694,535 & 1,694,535 & 1,807,504 & 1,807,504\\
-\bottomrule
-\end{tabular}
-\begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the village level.
-The outcome in all regressions is logged nightlights.
-Rainfall is included in columns (2) and (4).
-* p<0.10 ** p<0.05 *** p<0.01
+\caption{\label{tab:yieldtablemonthly}Month-specific exposure}
+\centering
+\begin{threeparttable}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\toprule
+  & monsoon & winter\\
+\midrule
+Month 1 & -0.808*** & 0.788\\
+(June/Nov.) & (0.066) & (0.588)\\
+Month 1 & 0.049 & -0.351\\
+(July/Dec.) & (0.060) & (0.967)\\
+Month 3 & -0.229*** & 4.758\\
+(Aug./Jan.) & (0.034) & (3.756)\\
+Month 4 & -0.215*** & -2.086***\\
+(Sept./Feb.) & (0.035) & (0.503)\\
+Month 5 & 0.149* & -3.788\\
+(Oct./March) & (0.082) & (3.210)\\
+\textbf{fixed effects:} & \textbf{} & \textbf{}\\
+village & Yes & Yes\\
+year & Yes & Yes\\
+\midrule
+F (month 1) & 1,430 & 566\\
+F (month 2) & 1,029 & 897\\
+F (month 3) & 2,713 & 358\\
+F (month 4) & 2,131 & 389\\
+F (month 5) & 592 & 565\\
+observations & 1,266,922 & 1,120,495\\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}[para]
+\item Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions is logged nightlights. * p<0.10 ** p<0.05 *** p<0.01
 \end{tablenotes}
@@
 \begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2cm}>{\centering\arraybackslash}p{4.5cm}>{\centering\arraybackslash}p{3.5cm}>{\centering\arraybackslash}p{3.5cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2cm}>{\centering\arraybackslash}p{5cm}>{\centering\arraybackslash}p{3cm}>{\centering\arraybackslash}p{3cm}}
 \toprule
@@
 pollution & Hammer et al. (2020) & global & monthly\\
-agriculture & Angopadhyay et al. (2022) & global & two seasons/year\\
+agriculture & Angopadhyay et al. (2022) & India & two seasons/year\\
 nightlights & Asher et al. (2021) & India & yearly\\
 weather & TerraClimate & global & monthly\\
+labor & National Sample Survey (NSS) & India & daily interviews\\
 \bottomrule
@@
 \item Global Energy Monitor: globalenergymonitor.org/projects/global-coal-plant-tracker. NCAR: climatedataguide.ucar.edu/. TerraClimate: www.climatologylab.org/terraclimate.html
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+
+\FloatBarrier
+\newpage
+
+\begin{table}
+
+\caption{\label{tab:yieldtablehet1st}Heterogeneity in the effects of pollution on productivity, first-stage results}
+\centering
+\begin{threeparttable}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2.8cm}>{\centering\arraybackslash}p{1.7cm}>{\centering\arraybackslash}p{1.7cm}>{\centering\arraybackslash}p{1.7cm}>{\centering\arraybackslash}p{1.7cm}>{\centering\arraybackslash}p{1.7cm}>{\centering\arraybackslash}p{1.7cm}}
+\toprule
+\multicolumn{1}{c}{ } & \multicolumn{1}{c}{(1)} & \multicolumn{1}{c}{(2)} & \multicolumn{1}{c}{(3)} & \multicolumn{1}{c}{(4)} & \multicolumn{2}{c}{(5)} \\
+\cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-7}
+  & PM & PM & PM & PM & PM & PM*rain\\
+\midrule
+wind & 0.003*** & 0.000*** & 0.003*** & 0.002*** & 0.001*** & -0.032***\\
+ & (0.000) & (0.000) & (0.000) & (0.000) & (0.000) & (0.001)\\
+wind times rain &  &  &  &  & 0.001*** & 0.009***\\
+ &  &  &  &  & (0.000) & (0.000)\\
+weather & Yes & Yes & Yes & Yes & Yes & Yes\\
+\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+village-season & Yes & Yes & Yes & Yes & Yes & Yes\\
+year & Yes & Yes & Yes & Yes & Yes & Yes\\
+\midrule
+observations & 1,051,010 & 1,320,211 & 1,193,764 & 1,177,457 & 2,371,221 & 2,371,221\\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}[para]
+\item Note: Standard errors are in parentheses and are clustered at the village level.
+The outcome in all regressions is estimated agricultural yield, defined as tons per hectare.
+The first two columns split the sample based on the median exposure (by village).
+The second two columns split the sample based on initial -- the first year available -- agricultural productivity.
+* p<0.10 ** p<0.05 *** p<0.01
 \end{tablenotes}
@@
 \midrule
-wind & -0.003*** & -0.002*** & -0.0006*** & -0.002*** & -0.003***\\
- & (0.0002) & (0.0002) & (9.56e-5) & (0.0002) & (0.0003)\\
-rain (z) &  & 0.018*** & 0.008*** & 0.061*** & 0.003***\\
- &  & (0.0005) & (0.001) & (0.002) & (0.0005)\\
-mean temp (10s) &  & 0.0008*** & -0.005*** & 0.005*** & -0.008***\\
- &  & (5.28e-5) & (0.0004) & (0.0001) & (0.0001)\\
+particulate matter & -1.153*** & -1.028*** & -2.270*** & -0.538*** & -1.688***\\
+(log PM 2.5) & (0.079) & (0.083) & (0.386) & (0.044) & (0.166)\\
+rain (z) &  & 0.017*** & 0.018*** & 0.073*** & -0.002***\\
+ &  & (0.001) & (0.002) & (0.002) & (0.001)\\
+mean temp (10s) &  & -0.000*** & -0.008*** & 0.005*** & -0.007***\\
+ &  & (0.000) & (0.001) & (0.000) & (0.000)\\
 \textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
@@
 observations & 1,886,956 & 1,872,177 & 1,872,177 & 996,509 & 875,668\\
+\midrule
+\textbf{first stage:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+wind & 0.002*** & 0.002*** & 0.000*** & 0.004*** & 0.002***\\
+ & (0.000) & (0.000) & (0.000) & (0.000) & (0.000)\\
+rain (z) &  & -0.002*** & 0.004*** & 0.023*** & -0.003***\\
+ &  & (0.000) & (0.000) & (0.000) & \vphantom{1} (0.000)\\
+mean temp (10s) &  & -0.001*** & -0.001*** & -0.000*** & 0.000***\\
+ &  & (0.000) & (0.000) & (0.000) & (0.000)\\
 \bottomrule
@@
 \midrule
-wind (lag) & 0.004*** & 0.0005\\
- & (0.0007) & (0.0003)\\
-wind & -0.002*** & -0.0006***\\
- & (0.0004) & (0.0002)\\
-wind (lead) & -0.004*** & 0.0000\\
- & (0.0001) & (0.0000)\\
-rain (z) & 0.022*** & 0.006***\\
- & (0.0004) & (0.001)\\
-mean temp (10s) & 0.0005*** & -0.007***\\
- & (6.04e-5) & (0.0004)\\
+wind (lag) & 0.0005 & \\
+ & (0.0003) & \\
+wind & -0.0006*** & \\
+ & (0.0002) & \\
+wind (lead) & -0.0001 & \\
+ & (0.0001) & \\
+PM (lag) &  & 0.4340\\
+ &  & (0.8666)\\
+PM &  & -0.6942***\\
+ &  & (0.1497)\\
+PM (lead) &  & -0.0958\\
+ &  & (0.5137)\\
+weather & Yes & Yes\\
 \textbf{fixed effects:} & \textbf{} & \textbf{}\\
 village-season & Yes & Yes\\
-year & Yes & No\\
-district-year-season & No & Yes\\
-\midrule
+district-year-season & Yes & Yes\\
+\midrule
+F (lag) & -- & 162\\
+F & -- & 410\\
+F (lead) & -- & 289\\
 observations & 1,976,205 & 1,976,205\\
@@
 
+\FloatBarrier
+\newpage
+
+\begin{figure}
+\includegraphics{draft_files/figure-latex/monthplot-1} \caption[Average farm days by season, rural only]{Average farm days by season, rural only}\label{fig:monthplot}\floatfoot*{Note: The line graph shows the mean number of farm days worked per week based on the month of the relevant agricultural season.
+The monsoon season is defined as June (month 1) to October (month 5) while the rainy season is defined as November (month 1) to March (month 5) of the following year.
+The data is from the NSS.}
+\end{figure}
+
 \end{document}

### 2023-02-21 presentation/slides/slides.Rmd
lots of updates!

@@
 ---
-title: "Pollution, agricultural productivity, and development: Evidence from coal in plants in India"
+title: "Pollution, agricultural productivity, and development: Evidence from coal plants in India"
 subtitle: 
@@
 labortable <- readRDS("pollution_development/draft/tables/labortable.rds")
+labortablerural <- readRDS("pollution_development/draft/tables/labortablerural.rds")
+labortableurban <- readRDS("pollution_development/draft/tables/labortableurban.rds")
+labortableruralyoung <- readRDS("pollution_development/draft/tables/labortableruralyoung.rds")
+labortableruralold <- readRDS("pollution_development/draft/tables/labortableruralold.rds")
+wagestable <- readRDS("pollution_development/draft/tables/wagestable.rds")
 labortablemonsoon <- readRDS("pollution_development/draft/tables/labortablemonsoon.rds")
@@
 laborplot <- readRDS("pollution_development/draft/tables/laborplot.rds")
+yieldtablemonthly <- readRDS("pollution_development/draft/tables/yieldtablemonthly.rds")
 
@@
 
-## Validating wind direction - particulate matter and exposure
+## Validating wind direction - particulate matter and exposure, monthly level
 
@@
 
+## Agricultural productivity and pollution leads/lags
+
+```{r yieldtableleads, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+# Table
+colnames(yieldtableleads) <- c("wind", "IV - PM (log)")
+yieldtableleads[1:6,2] <- yieldtableleads[7:12,2]
+yieldtableleads <- yieldtableleads[-c(7:12),]
+yieldtableleads <- yieldtableleads[-c(7:10),]
+rownames(yieldtableleads)[c(1,3,5)] <- c("lag", "current", "lead")
+kable(
+      yieldtableleads,
+      align = "cc", booktabs = TRUE, linesep = ""
+      ) %>%
+  footnote(
+           "Note: Standard errors are in parentheses and are clustered at the village level.
+The outcome in all regressions estimated agricultural yield (tons per hectare).
+* p<0.10 ** p<0.05 *** p<0.01",
+           general_title = "",
+           threeparttable = TRUE,
+           footnote_as_chunk = TRUE
+           ) %>%
+  column_spec(1, width = "3cm") %>%
+  column_spec(c(2:3),width = "2cm") %>%
+  kable_classic_2()
+
+```
+
+---
+
 ## Taking stock
@@
 
-## Wind direction and labor allocation, by age
+## Wind direction and labor allocation, by rural status
 
-```{r labortableold, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+```{r labortablerural, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
 
 # Table
-labortableold <- labortableold[,-1]
-colnames(labortableold) <- c("all", "self", "wage", "farm", "non-farm")
+colnames(labortablerural) <- c("all", "all", "self", "wage", "farm", "non-farm")
+colnames(labortableurban) <- c("all", "all", "self", "wage", "farm", "non-farm")
+
+mat1 <- matrix(c("", "", "", "", "", ""), nrow = 1)
+colnames(mat1) <- c("all", "all", "self", "wage", "farm", "non-farm")
+rownames(mat1) <- c("rural")
+mat2 <- matrix(c("", "", "", "", "", ""), nrow = 1)
+colnames(mat2) <- c("all", "all", "self", "wage", "farm", "non-farm")
+rownames(mat2) <- c("urban")
+
+labortablerural2 <- rbind(mat1, labortablerural[c(1:2, 7:8),], mat2, labortableurban[-c(3:6),], labortableurban[c(3:6),])
+
 kable(
-      labortableold,
-      align = "ccccc", booktabs = TRUE, linesep = ""
+      labortablerural2[-c(nrow(labortablerural2)-2, nrow(labortablerural2)-1, nrow(labortablerural2)),],
+      align = "cccccc", booktabs = TRUE, linesep = ""
       ) %>%
@@
            ) %>%
-  column_spec(1, width = "4cm") %>%
-  column_spec(c(2:6),width = "1.5cm") %>%
-  row_spec(c(10), hline_after = TRUE) %>%
-  row_spec(c(6,9), bold = TRUE) %>%
+  column_spec(1, width = "3cm") %>%
+  column_spec(c(2:7),width = "1.5cm") %>%
   kable_classic_2()
@@
 
-## Large negative effects on older half
+## Wind direction and labor allocation, by age (rural only)
 
-- Older half of sample decreases overall labor allocation
-  - Driven by self-employment and nonfarm employment
-- Younger half actually increases allocation!
-- Consistent with health effects only on older adults, though not conclusive
+```{r labortableruralage, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+# Table
+colnames(labortableruralyoung) <- c("all", "all", "self", "wage", "farm", "non-farm")
+colnames(labortableruralold) <- c("all", "all", "self", "wage", "farm", "non-farm")
+
+mat1 <- matrix(c("", "", "", "", "", ""), nrow = 1)
+colnames(mat1) <- c("all", "all", "self", "wage", "farm", "non-farm")
+rownames(mat1) <- c("younger than 35")
+mat2 <- matrix(c("", "", "", "", "", ""), nrow = 1)
+colnames(mat2) <- c("all", "all", "self", "wage", "farm", "non-farm")
+rownames(mat2) <- c("45 or older")
+
+labortableruralyoung2 <- rbind(mat1, labortableruralyoung[c(1:2, 7:8),], mat2, labortableruralold[-c(3:6),], labortableruralold[c(3:6),])
+
+kable(
+      labortableruralyoung2[-c(nrow(labortableruralyoung2)-2, nrow(labortableruralyoung2)-1, nrow(labortableruralyoung2)),],
+      align = "cccccc", booktabs = TRUE, linesep = ""
+      ) %>%
+  footnote(
+           "Note: Standard errors are in parentheses and are clustered at the district level.
+Control variables include female, age, age squared, and (years of) education.
+The outcome in each regression is days worked in the previous seven days.
+* p<0.10 ** p<0.05 *** p<0.01",
+           general_title = "",
+           threeparttable = TRUE,
+           footnote_as_chunk = TRUE
+           ) %>%
+  column_spec(1, width = "3cm") %>%
+  column_spec(c(2:7),width = "1.5cm") %>%
+  kable_classic_2()
+
+```
+
+---
+
+## One last check
+
+- Farm labor most common in July 
+- Have monthly values of wind and pollution...
+  - Heterogeneity by month of the season?
+  - Monsoon season only
+
+---
+
+## Heterogeneity by month of the monsoon season
+
+```{r yieldtablemonthly, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+# Table
+colnames(yieldtablemonthly) <- c("(1)", "(2)")
+
+kable(
+      yieldtablemonthly[c(1:10, nrow(yieldtablemonthly)),],
+      align = "cc", booktabs = TRUE, linesep = ""
+      ) %>%
+  footnote(
+           "Note: Standard errors are in parentheses and are clustered at the village level.
+The outcome in all regressions estimated agricultural yield (tons per hectare).
+* p<0.10 ** p<0.05 *** p<0.01",
+           general_title = "",
+           threeparttable = TRUE,
+           footnote_as_chunk = TRUE
+           ) %>%
+  column_spec(1, width = "3cm") %>%
+  column_spec(c(2:3),width = "1.5cm") %>%
+  kable_classic_2()
+
+```
 
@@
 # Table
+ntltable <- ntltable[,c(1,3)]
 colnames(ntltable) <- c("(1)", "(2)")

### 2023-02-21 analyses/regressions.R

@@ -130,11 +130,11 @@ ggsave("pollution_development/draft/tables/devplot.png")
 
 
 
-
-pol1 <- feols(mean ~ wind | village + month, data = pollution, cluster = c("village"))
-pol2 <- feols(mean ~ wind | village + month^distfe, data = pollution, cluster = c("village"))
-pol3 <- feols(mean ~ wind | village + month, data = pollution %>% filter(year>=2002 & year<=2013), cluster = c("village"))
-pol4 <- feols(mean ~ wind | village + month^distfe, data = pollution %>% filter(year>=2002 & year<=2013), cluster = c("village"))
+pollution <- pollution %>% mutate(wind = wind*0.1)
+pol1 <- feols(log(mean) ~ wind | village + month, data = pollution, cluster = c("village"))
+pol2 <- feols(log(mean) ~ wind | village + month^distfe, data = pollution, cluster = c("village"))
+pol3 <- feols(log(mean) ~ wind | village + month, data = pollution %>% filter(year>=2002 & year<=2013), cluster = c("village"))
+pol4 <- feols(log(mean) ~ wind | village + month^distfe, data = pollution %>% filter(year>=2002 & year<=2013), cluster = c("village"))
 
 pollutiontable <- etable(
                           pol1, pol2, pol3, pol4, 
@@ -148,7 +148,7 @@ pollutiontable <- etable(
                           )
 pollutiontable <- pollutiontable[-c(7:8),]
 pollutiontable <- as.matrix(pollutiontable)
-rownames(pollutiontable) <- c("wind", "", "fixed effects:",
+rownames(pollutiontable) <- c("wind (10s)", "", "fixed effects:",
                                "village", "month", "district-month", "observations")
 pollutiontable[c(3),] <- " "
 saveRDS(pollutiontable, "pollution_development/draft/tables/pollutiontable.rds") 
@@ -212,6 +212,7 @@ summary(df$wind*10)
 
 # Mean, divided by ten (for presentation)
 df <- df %>% mutate(temp_mean = (tmax_mean + tmin_mean)/20)
+df$pm25 <- log(df$pm25)
 
 
 yield1 <- feols(log(yield) ~ wind | village^season + year, data = df, cluster = c("village"))
@@ -286,7 +287,7 @@ yieldtabletwo_merge <- rbind(yieldtabletwo[-c(12:13),], new_row, yieldtabletwo1[
 # Turn to matrix so that duplicate row names are allowed
[…]

### 2023-02-21 cleaning/4.cleaning_plants.R

@@ -60,6 +60,9 @@ buffer10km <- st_as_sf(buffer10km)
 # Create buffer of 30 km around the plants
 buffer30km <- gBuffer( as_Spatial(plants), width = 30000, byid = TRUE )
 buffer30km <- st_as_sf(buffer30km)
+# Create buffer of 100 km around the plants
+buffer100km <- gBuffer( as_Spatial(plants), width = 100000, byid = TRUE )
+buffer100km <- st_as_sf(buffer100km)
 
 # Create centroids of villages
 centroids_villages <- gCentroid( as_Spatial(villages), byid = TRUE )
@@ -78,9 +81,21 @@ village_within_30km <- (lengths(centroids_villages_30km) > 0)==T
 # Just keep those that fall within 30km (replacing the centroids_villages_30km object)
 centroids_villages_30km <- centroids_villages[village_within_30km,]
 
+# Now let's figure out which are within the 100km buffer
+centroids_villages_100km <- st_intersects(centroids_villages, buffer100km)
+village_within_100km <- (lengths(centroids_villages_100km) > 0)==T 
+
+# Just keep those that fall within 100km (replacing the centroids_villages_100km object)
+centroids_villages_100km <- centroids_villages[village_within_100km,]
+
+
 # Save villages
 st_write(centroids_villages_30km, "data/spatial/centroids_villages_30km", 
            driver = "ESRI Shapefile", append = F)
+st_write(centroids_villages_100km, "data/spatial/centroids_villages_100km", 
+           driver = "ESRI Shapefile", append = F)
+
+
 
 ####### Loaded villages and plants into QGIS to do distances (much faster than doing it here)
 # Load distance matrix (takes a while; this csv is >4GB)
@@ -149,17 +164,16 @@ dist_matrix$angle <- dist_matrix$angle_original
 dist_matrix$angle[dist_matrix$angle<0] <- dist_matrix$angle[dist_matrix$angle<0] + 360
 
 dist_matrix <- dist_matrix %>% dplyr::select(-c("x_vil", "y_vil", "x_plants", "y_plants"))
+write.csv(dist_matrix, paste0("data/clean/wind_ntl/dist_matrix_angles.csv"))
 
 # Takes a while to get here. Save so we can start here if we want
-write.csv(dist_matrix, paste0("data/clean/wind_ntl/dist_matrix_angles.csv"))
[…]

### 2023-02-21 cleaning/5.cleaning_precip_temp.R

@@ -74,7 +74,20 @@ for (year in 1990:2020){
   file.remove(file)
   print(year)
 }
-  
 
+ntl_temp <- c()
+for (year in 1990:2020){
+   tmax <- read_csv(paste0("data/clean/terra/tmax", year, ".csv"))
+   tmin <- read_csv(paste0("data/clean/terra/tmin", year, ".csv"))
+   tmax[,2:13] <- (tmax[,2:13] + tmin[,2:13])/2
+   tmax <- tmax %>%
+            mutate(temp = rowMeans(tmax[,2:13]),
+                   year = year) %>%
+            dplyr::select(shrid, temp, year)
+   
+   
+   ntl_temp <- rbind(ntl_temp, tmax)
+}
+write_csv(ntl_temp, paste0("data/clean/village_temp.csv"))
   
   

### 2023-02-21 cleaning/6.cleaning_pm25.R

@@ -39,7 +39,7 @@ months <- c()
 for (month in 1:9){
   months <- c(months, paste0("0", month))
 }
-months <- c(months, "11", "12")
+months <- c(months, "10", "11", "12")
 pollution <- c()
 for (year in 1998:2015){
   for (month in months){
@@ -77,32 +77,40 @@ write_csv(pollution, paste0("data/clean/pm25/all_combined.csv"))
 
 # Now clean to merge into ag productivity (try IV?)
 pollution <- read_csv(paste0("data/clean/pm25/all_combined.csv"))
+# Don't need these months
+pollution <- pollution %>% filter(!(month %in% c(4, 5)), is.na(mean)==F)
+
 pollution$season <- ""
 pollution$season[pollution$month %in% c(6:10)] <- "monsoon"
 pollution$season[pollution$month %in% c(11:12) | pollution$month %in% c(1:3)] <- "winter"
-pollution$year[pollution$season=="winter"] <- pollution$year[pollution$season=="winter"] - 1
+pollution$year[pollution$month %in% c(1:3)] <- pollution$year[pollution$month %in% c(1:3)] - 1
 
 pollution$m1 <- NA
-pollution$m1[pollution$month==6 | pollution$month==11] <- pollution$mean[pollution$month==6 | pollution$month==11]
+pollution$m1[pollution$month==6] <- pollution$mean[pollution$month==6]
+pollution$m1[pollution$month==11] <- pollution$mean[pollution$month==11]
 pollution$m2 <- NA
-pollution$m2[pollution$month==7 | pollution$month==12] <- pollution$mean[pollution$month==7 | pollution$month==12]
+pollution$m2[pollution$month==7] <- pollution$mean[pollution$month==7]
+pollution$m2[pollution$month==12] <- pollution$mean[pollution$month==12]
 pollution$m3 <- NA
-pollution$m3[pollution$month==8 | pollution$month==1] <- pollution$mean[pollution$month==8 | pollution$month==1]
+pollution$m3[pollution$month==8] <- pollution$mean[pollution$month==8]
+pollution$m3[pollution$month==1] <- pollution$mean[pollution$month==1]
 pollution$m4 <- NA
-pollution$m4[pollution$month==9 | pollution$month==2] <- pollution$mean[pollution$month==9 | pollution$month==2]
+pollution$m4[pollution$month==9] <- pollution$mean[pollution$month==9]
+pollution$m4[pollution$month==2] <- pollution$mean[pollution$month==2]
 pollution$m5 <- NA
[…]

### 2023-02-21 cleaning/7.cleaning_ag_productivity_village.R

@@ -36,6 +36,7 @@ rm(villages_centroids)
 
 
 
+
 for (year in 2002:2013){
   
   # MONSOON --------------------------------------------------------------------
@@ -66,11 +67,15 @@ for (year in 2002:2013){
   
 }
 
+# date of coal plant opening
+dist_matrix <- read.csv(paste0("data/clean/wind_ntl/dist_matrix_angles.csv"))
 
 # The months for monsoon are Jun to Oct
 # The months for Winter are Nov to Mar
 # 2002 to 2013
 for (year in 2002:2013){
+  # Is there a single coal plant within 30 km that year?
+  
   # Monsoon is this year only
   rainfall <- read_csv(paste0("data/clean/terra/precip", year, ".csv"))
   rainfall <- rainfall[, c(1, 7:11)]
@@ -229,21 +234,41 @@ for (year in 2002:2013){
   # Monsoon is this year only
   village_wind <- read_csv(paste0("data/clean/wind_ntl/months/y", year, "m1.csv")) %>% as_tibble()
   villages <- village_wind %>% dplyr::select(shrid)
-  month_temp <- 1
-  for (month in 6:10){
+  month_temp <- 0
+  for (month in 5:10){
     village_wind <- read_csv(paste0("data/clean/wind_ntl/months/y", year, "m", month, ".csv")) %>% as_tibble()
     colnames(village_wind) <- c("shrid", paste0("m", month_temp))
     villages <- villages %>% left_join(village_wind, by = "shrid")
     month_temp <- month_temp + 1
   }
   villages$season <- "monsoon"
-  villages <- villages %>% mutate(days_sums = rowSums(villages[,2:6]))
+  villages <- villages %>% mutate(days_sums = rowSums(villages[,3:7]))
   # Just make sure (should already be named this)
-  colnames(villages) <- c("shrid", "m1", "m2", "m3", "m4", "m5", "season", "days_sums")
+  colnames(villages) <- c("shrid", "m0", "m1", "m2", "m3", "m4", "m5", "season", "days_sums")
   
+  # Now values for plants NOT OPEN
+  village_wind <- read_csv(paste0("data/clean/wind_ntl_NOT/months/y", year, "m1.csv")) %>% as_tibble()
+  villages_TEMP <- village_wind %>% dplyr::select(shrid)
+  month_temp <- 0
+  for (month in 5:10){
[…]

### 2023-03-02 draft/draft.Rmd
scripts and draft

@@
 abstract: |
-  | \noindent I show that emissions from a common energy source lead to decreased productivity and growth.
-The rapid increase in population in India from the 1990s through the 2010s led to the rapid construction of new coal power plants throughout the country.
-Using this roll-out combined with random changes in wind direction, I document negative effects of pollution from coal power plants on agricultural productivity throughout the country.
-Results suggest this decrease may be driven by a decline in labor allocation of the oldest half of my sample, who greatly decrease labor allocated to both farm and non-farm employment when exposed to pollution.
-These impacts spill over into aggregate economic growth, as well; nightlight growth is slower following years of higher exposure to pollution.
+  | \noindent Pollution is a growing problem in many developing countries.
+Using the rollout of coal plants to identify high-pollution locations in India and random variation induced by changes in wind direction, I show that pollution has large, negative impacts on agricultural productivity.
+Results indicate that some of this impact is driven by changes in labor allocation patterns in response to increased exposure to pollution.
+I also show that pollution from coal plants, specifically, has a larger effect on agricultural productivity than other types of pollution.
   |
@@
 pollutiontable <- readRDS("pollution_development/draft/tables/pollutiontable.rds")
-yieldtable <- readRDS("pollution_development/draft/tables/yieldtable.rds")
-yieldtablepm <- readRDS("pollution_development/draft/tables/yieldtablepm.rds")
-yieldtabletwo <- readRDS("pollution_development/draft/tables/yieldtabletwo.rds")
-yieldtablehet <- readRDS("pollution_development/draft/tables/yieldtablehet.rds")
-yieldtablehet1st <- readRDS("pollution_development/draft/tables/yieldtablehet1st.rds")
-yieldtablemonthly <- readRDS("pollution_development/draft/tables/yieldtablemonthly.rds")
-pmeffects <- readRDS("pollution_development/draft/tables/yieldtablePM.rds")
-labortable <- readRDS("pollution_development/draft/tables/labortable.rds")
-labortablerural <- readRDS("pollution_development/draft/tables/labortablerural.rds")
-labortableurban <- readRDS("pollution_development/draft/tables/labortableurban.rds")
-labortableruralyoung <- readRDS("pollution_development/draft/tables/labortableruralyoung.rds")
-labortableruralold <- readRDS("pollution_development/draft/tables/labortableruralold.rds")
-wagestable <- readRDS("pollution_development/draft/tables/wagestable.rds")
-labortablemonsoon <- readRDS("pollution_development/draft/tables/labortablemonsoon.rds")
-labortablewinter <- readRDS("pollution_development/draft/tables/labortablewinter.rds")
-labortableold <- readRDS("pollution_development/draft/tables/labortableold.rds")
-labortableoldstate <- readRDS("pollution_development/draft/tables/labortableoldstate.rds")
-ntltable <- readRDS("pollution_development/draft/tables/ntltable.rds")
-yieldtableleads <- readRDS("pollution_development/draft/tables/yieldtableleads.rds")
-yieldtablepostplant <- readRDS("pollution_development/draft/tables/yieldtablepostplant.rds")
+yield1reducedform <- readRDS("pollution_development/draft/tables/yield1reducedform.rds")
+yield2naive <- readRDS("pollution_development/draft/tables/yield2naive.rds")
+yield3ivmain <- readRDS("pollution_development/draft/tables/yield3ivmain.rds")
+yield4monthly <- readRDS("pollution_development/draft/tables/yield4monthly.rds")
+yield5heterogeneity <- readRDS("pollution_development/draft/tables/yield5heterogeneity.rds")
+yield6leads <- readRDS("pollution_development/draft/tables/yield6leads.rds")
+yield7ivdiffindiff <- readRDS("pollution_development/draft/tables/yield7ivdiffindiff.rds")
+labor1all <- readRDS("pollution_development/draft/tables/labor1all.rds")
+labor2rural <- readRDS("pollution_development/draft/tables/labor2rural.rds")
+labor3month <- readRDS("pollution_development/draft/tables/labor3month.rds")
 laborplot <- readRDS("pollution_development/draft/tables/laborplot.rds")
 monthplot <- readRDS("pollution_development/draft/tables/monthplot.rds")
+opplot <- readRDS("pollution_development/draft/tables/opplot.rds")
 
@@
 
-I study the effects of emissions from coal power plants on agricultural productivity, labor allocation, and economic growth in India.
+I study the effects of emissions from coal power plants on agricultural productivity in India.
 Previous work has suggested that there may be harmful direct effects of air pollution [@heck1982assessment; @marshall1997hidden] and water pollution [@reddy2006impact] on agricultural productivity, including work showing that gold mines -- which release large amounts of pollutants into the surrounding environment -- lead to lower agricultural productivity in Ghana [@aragon2016polluting].
@@
 Using a village-level shapefile, I identify all villages that are located within 30km of a coal plant at any time during the sample period, which comprises the 1990s to the 2010s, with differences depending on data availability.
-To measure exposure to emissions from coal plants, I calculate the direction from each coal plant to all villages within 30km.
+
+I use the location of coal plants to identify high-pollution areas and calculate the direction from each coal plant to all villages within 30km.
+I include all areas, regardless of whether the coal plant has opened yet, to allow me to disaggregate the effects of pollution from coal plants with other kinds of pollution.
 Using daily wind data and these spatial variables, I create an identifier separately for each of the one-hundred-thousand-plus villages for whether a given village is located downwind from a coal plant on any given day.
-I then aggregate this exposure variable based on the desired analysis; for agricultural productivity, for example, I create an exposure variable measuring the total number of exposure days in the five-month agricultural season, separately for the main monsoon season and the smaller winter season.
+I then aggregate this exposure variable based on the desired analysis; for agricultural productivity, for example, I create an exposure variable measuring the total number of exposure days in the five-month agricultural season, focusing only on the monsoon (Kharif) season.
 Consistent with expectations, wind direction is a strong predictor of particulate matter at the monthly level.
@@
 Using newly available estimates of India-wide agricultural (land) productivity, results show that the total number of days of being downwind from a coal power plant has a negative effect on overall agricultural productivity.
-Just a single additional day of being downwind from a coal power plant leads to a decrease in predicted yields of 0.3 percent.
-The within-village absolute deviation of wind exposure is around eight days, meaning that median changes in wind direction are responsible for fluctuations in agricultural productivity of around 2.4 percent, with changes of more than five percent from season to season being common.
-Using wind direction as an instrument for particulate matter, an increase of PM 2.5 by one microgram per cubic meter leads to an even larger decrease in agricultural productivity, at around 20 percent.
+Just a single additional day of being downwind from a coal power plant leads to a decrease in predicted yields of 0.6 percent.
+Combined with the within-village absolute deviation of wind exposure, results indicate that median changes in wind direction are responsible for fluctuations in agricultural productivity of around 1.5 percent.
+
+Using wind direction as an instrument for particulate matter, an increase of PM 2.5 by one percent leads to an even larger decrease in agricultural productivity, at around 0.73 percent.
+Given within-village changes across years, pollution could be responsible for swings of 7-to-14 percent in agricultural productivity across years.
 This IV strategy is similar to that in @deryugina2019mortality, though they use general "high-pollution" wind directions while I concentrate only on wind directions from the location of a coal power plant.
@@
 
-These effects are apparently concentrated in villages that do not experience the most extreme exposure to pollution; villages with higher maximum exposure experience effects only around 10 percent as large as villages with lower maximum exposure.
-Additional heterogeneity analysis shows that, despite the results with respect to the maximum, the effects are increasing in *mean* exposure.
-I also show that the effects are largest in villages with the highest initial agricultural productivity and that pollution and rainfall shocks are compounding; higher exposure along with less rainfall increases the negative effect of either shock individually.
+These effects are apparently larger in villages that experience higher median exposure to pollution and in villages with higher initial yields.
+I also show that rainfall shocks are compounding; higher exposure along with less rainfall increases the negative effect of either shock individually.
 As weather shocks become more common across the globe, the rise of pollution in India may be particularly problematic for farmers.
 
-Robustness checks show that this effect is concentrated in the years after the construction of the power plant, which helps decrease concerns around endogeneity since, conditional on the village fixed effect and weather (rainfall and temperature), wind direction across years and seasons is exogenous.
-Additionally, leads for wind direction do not predict current agricultural productivity, again supporting the identifying assumptions.
-
-While I am not able to explicitly test effects on land productivity, I look at how adults allocate labor over the last seven days using the National Sample Survey and creating a district-level exposure measure based on the timing of the survey for each household.
-An additional day of wind over the previous week decreases overall labor allocation by `r paste0(substr(labortable[1,2],2,6))` days.
-However, the overall results show stark heterogeneity based on age; the older half of the sample greatly decreases labor allocation in response to pollution exposure while the younger half increases overall labor allocation.
-Older adults decrease labor allocation to both farm and non-farm labor, while younger adults increase labor allocated to farm labor.
-This is suggestive evidence of productivity effects being driven by the health and productivity effects on individuals, since older adults are more susceptible to pollution and they also greatly decrease non-farm employment, which presumably will no suffer any possible direct effects of pollution on productivity.
-The preponderance of results seems to indicate that effects on labor are the most likely mechanism.
-
-The final set of results explores whether there are impacts of pollution on aggregate economic development.
-Using nightlights, I show that lagged wind -- i.e.
-pollution exposure in the prior year -- has a negative impact on economic development as proxied by nightlights.
-Going from the 25th percentile to the 75th percentile of wind exposure is synonymous with a decrease in the growth of nightlights within a village by around 0.7 percent.
-While this is a relatively small magnitude, it is worth noting that there are more than 100,000 villages in this sample.
-While the effect for a single village might be small, the overall, aggregate effect on growth is quite substantial.
-To put this number in perspective, @castello2018higher argue that a standard-deviation increase in the share of adults with a university education leads to 12 percent higher nightlights in India, while @storeygard2016farther shows that a four-fold increase in oil prices over six years leads to seven percent lower nightlights at a distance of around 500km from the nearest large port in sub-Saharan Africa.
-In other words, the 0.7 percent change in nightlights is quite large given this is annual variation.
+To explore the mechanisms driving this effect, I also use data on labor allocation from the National Sample Survey (NSS).
+While there do not appear to be large changes in labor allocation patterns when looking at the season as a whole, there is a clear pattern when we break down exposure to pollution by month: higher levels of pollution at the beginning of the monsoon season (June and July) lead to lower levels of farm labor allocation, but no changes in non-farm labor allocation.
+This matches up well with a similar exercise with agricultural yields: pollution in June is the only month in which there is a negative effect of pollution on agricultural yield.
+In tandem, these two results point to effects on labor as one mechanism driving the total effects, though I cannot rule out direct effects of pollution on land productivity.
+
+The final set of results tries to disentangle the effects of pollution from the effects of coal pollution, specifically.
+Combining instrumental variables with a differences-in-differences strategy, I show that agricultural productivity has an elasticity of around -0.15 with respect to non-coal pollution.
+The opening of a coal plant, however, triples the elasticity, to more than -0.45, indicating that coal plant emissions are particularly harmful to agricultural productivity.
+This is consistent with coal emissions being quite different from other forms of pollution; for example, coal emissions contain much more mercury than other forms of pollution.
 
@@
 
-```{r plants, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Coal plants in India from 1990 to 2010", fig.footer = list("Note: The top figure shows the location of coal plants in 1990.
-The bottom figure shows the location of coal plants in 2010.")}
-
-plant1 <- ggplot() +
-                theme_map() +
-                geom_sf(data = india_shape2, fill = NA, lwd = 0.01) +
-                geom_sf(data = plants1990, size = 0.75, color = "black") +
-                labs(x = "lon",
-                     y = "lat",
-                     subtitle = "Figure A: 1990")
-
-plant2 <- ggplot() +
-                theme_map() +
-                geom_sf(data = india_shape2, fill = NA, lwd = 0.01) +
-                geom_sf(data = plants2010, size = 0.75, color = "black") +
-                labs(x = "lon",
-                     y = "lat",
-                     subtitle = "Figure B: 2010")
-
-ggarrange(plant1, plant2)
-
-```
-
-Figure \@ref(fig:plants) displays the location of coal plants in India for two specific years: 1990 and 2010.
+Figure \@ref(fig:plants) in the appendix displays the location of coal plants in India for two specific years: 1990 and 2010.
 There is a clear increase in the prevalence of coal plants across the country over the two decades.
@@
 
-[^10]: I define "towards" as within five degrees to help take into account that I am using village centroids, using the x and y components of wind speed.
-
-```{r windexample, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 8, fig.cap = "Wind direction and aggregation examples (2010-01-01)", fig.footer = list("Note: The top figure shows the average wind direction on January 1st, 2010, with zero degrees indicating north.
-The points are the location of coal plants on that date.
-The bottom figure shows the distribution of pollution exposure in a specific district -- Guna district in Madhya Pradesh -- on the same date.")}
-
-p1 <- ggplot() + 
-        theme_map() +
-        geom_sf(data = india_shape, fill = NA) +
-        geom_tile(data = windexample, aes(x = x, y = y, fill = wind_example), alpha = 0.3) +
-        scale_fill_viridis_c("wind direction\n(degrees)", option = "viridis") +
-        geom_sf(data = plants2010, size = 0.5) +
-        labs(x = "lon",
-             y = "lat",
-             subtitle = "Figure A: Wind direction") +
-        theme(legend.position = c(0.6, 0.05),
-              legend.direction = "vertical",
-              legend.key.height = unit(0.4, 'cm'),
-              legend.text = element_text(size = 8),
-              legend.background=element_blank() )
-
-p2 <- ggplot() + 
-        theme_map() +
-        geom_sf(data = districts, fill = NA) +
-        geom_sf(data = villages %>% filter(is.na(days_sum)==F), aes(fill = days_sum), lwd = NA) +
-        scale_fill_viridis_c("exposure (days)", option = "viridis") +
-        labs(x = "lon",
-             y = "lat",
-             subtitle = "Figure B: Aggregation to district") +
-        theme(legend.position = c(0.5, 0.05),
-              legend.direction = "horizontal",
-              legend.key.width = unit(0.3, 'cm'),
-              legend.text = element_text(size = 8),
-              legend.background=element_blank() )
-
-ggarrange(p1, p2,
-          nrow = 2)
-
-```
-
-The top panel in Figure \@ref(fig:windexample) is an example wind direction raster.
+[^10]: I define "towards" as within five degrees to help take into account that I am using village centroids, using the x and y components of wind speed to calculate angles.
+
+The top panel in Figure \@ref(fig:windexample) in the appendix is an example wind direction raster.
 The raster shows wind direction for the entirety of India on January 1st, 2010, as well as the location of coal power plants on that date.
@@
 
-Finally, I use nightlights as a measure of economic growth [@henderson2012measuring].
-I also take this data directly from SHRUG and @almn2021.
-This covers 1992 through 2013.
-Some of the older data may not be ideal [@gibson2021night], but this is the only way to document changes in economic growth at a more disaggregated level and at a regular basis.
-Although satellite imagery is an attractive alternative [@burke2021using], high-resolution satellite imagery does not go back to the early 1990s.
-
 ## Identification
@@
 
-To calculate the effects of exposure to pollution on different outcomes, I rely on fixed effects and the differential exposure driven by plausibly exogenous wind direction, conditional on total rainfall and average temperature.
-The regression is of the form \begin{gather} y_{it} = \alpha_{i} + \gamma_{t} + \beta wind_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it}, \end{gather} where $y_{it}$ is outcome $y$ for unit $i$ at time $t$, $\alpha$ is a geographic fixed effects, $\gamma_t$ is time fixed effects, $wind_{it}$ is exposure to pollution -- defined using wind direction -- $rain_{it}$ is total rainfall during the season, and $temp_{it}$ is average temperature during the season.
-Although I use $i$ and $t$ as subscripts for the fixed effects, note that these change depending on the unit of analysis and the outcome.
-I explicitly state the level of fixed effects in the tables and results below.
-
-Identification relies on changes in wind direction across time for the same geographic units.
-In other words, conditional on the fixed effects and weather, I assume changes in wind direction are as good as random and uncorrelated with the outcome -- e.g.
-agricultural productivity -- except through exposure to pollution.
-One major threat to identification is that the building of a coal plant could lead to a spurious correlation between the exposure variable and the outcome.
-Specifically, since exposure is defined as zero prior to the building of a coal plant, the exposure variable itself might be endogenous even though wind direction is not.
-To help allay these concerns, I estimate additional regressions restricting the sample to only time periods after which a given location has a coal plant.
-I also present results using leads to show that future wind direction does not seem to predict current outcomes.
-
-The inclusion of rainfall and temperature is meant to reduce concerns that wind direction may not affect agricultural productivity but, instead, be correlated with it through changes in weather.
-For example, prevailing winds from one direction may be associated with dry weather, while winds from another direction may be more likely to carry moisture and rain.
-It is not a priori clear how this would bias the estimates with respect to pollution and coal plants, however.
-The differences in weather patterns is very context specific and can differ even for villages located near one another.
-Nonetheless, adding rainfall and temperature should alleviate some of these concerns.
-
-A large question looms with respect to identification: what are the characteristics of areas near the construction of new coal plants?
-While I examine the identifying assumptions in a panel context below, I begin with this static question here.
+I am primarily interested in the effects of pollution on agricultural productivity. We might consider a regression of the following form:
+
+\begin{gather} y_{it} = \alpha_{i} + \gamma_{t} + \beta PM_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it}, \end{gather}
+
+where $y_{it}$ is outcome $y$ for unit $i$ at time $t$, $\alpha$ is a geographic fixed effects, $\gamma_t$ is time fixed effects, $PM_{it}$ is the concentration of particulate matter 2.5, $rain_{it}$ is total rainfall during the season, and $temp_{it}$ is average temperature during the season.
+
+However, even with the fixed effects, this regression likely suffers from endogeneity; places with more pollution may be very different from places with less pollution.
+It is difficult to sign the direction of the bias a priori, because it depends on how the coal plants are placed.
+For example, if coal plants are placed in places in need of power -- for example, due to rapid growth -- then the estimates might be upwards biased.
+On the other hand, if official purposefully place coal plants in poorer areas for political reasons, then the estimates might be downwards biased.
+
+So what are the characteristics of areas near the construction of new coal plants?
 Table \@ref(tab:plantresultstable) presents four separate regressions.
@@
 
+Given the likely endogeneity, I instead turn to a plausibly exogenous source of variation in pollution: wind direction.
+While the location of a coal plant might be endogenous, the amount of pollution received by a specific village in a given year can vary randomly based on wind direction.
+By counting the number of days in a given year that a village is downwind from a polluted area and conditioning on village fixed effects and weather (rainfall and temperature), variation across years is plausibly exogenous.
+
+Consider the reduced form effect of wind exposure on agricultural productivity:
+\begin{gather} y_{it} = \alpha_{i} + \gamma_{t} + \beta wind_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it}, \end{gather}
+
+where variables are defined identically to above but with wind as the primary variable of interest instead of pollution.
+Combining these into a two-stage procedure, we can instrument for pollution using wind direction; small changes in wind direction can lead to large changes in pollution exposure for villages located near point sources of pollution, allowing me to identify the causal effects of pollution by isolating the random component driven by wind direction.
+This is similar in spirit to @deryugina2019mortality.
+
+Identification relies on changes in wind direction across time for the same geographic units.
+In other words, conditional on the fixed effects and weather, I assume changes in wind direction are as good as random and uncorrelated with the outcome -- e.g.
+agricultural productivity -- except through exposure to pollution.
+One major threat to identification is that the building of a coal plant could lead to a spurious correlation between the exposure variable and the outcome.
+Specifically, since exposure is defined as zero prior to the building of a coal plant, the exposure variable itself might be endogenous even though wind direction is not.
+To help allay these concerns, I estimate additional regressions restricting the sample to only time periods after which a given location has a coal plant.
+I also present results using leads to show that future wind direction does not seem to predict current outcomes.
+
+The inclusion of rainfall and temperature is meant to reduce concerns that wind direction may not affect agricultural productivity but, instead, be correlated with it through changes in weather.
+For example, prevailing winds from one direction may be associated with dry weather, while winds from another direction may be more likely to carry moisture and rain.
+It is not a priori clear how this would bias the estimates with respect to pollution and coal plants, however.
+The differences in weather patterns is very context specific and can differ even for villages located near one another.
+Nonetheless, adding rainfall and temperature should alleviate some of these concerns.
+
 It is worth taking the time to validate the use of wind direction as a measure of exposure to pollution.
-Consider the data used for the regressions presented in Table \@ref(tab:pollutiontable).
+Consider the data used for the regressions presented in Table \@ref(tab:pollutiontable) of the appendix.
 The outcome variable is particulate matter -- specifically, $\mathrm{PM_{2.5}}$, which is particulate matter no larger than 2.5 micrometers in diameter -- which comes from @hammer2020global.
@@
 [^12]: www.epa.gov/airmarkets/power-plants-and-neighboring-communities
-
-```{r pollutiontable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-# Table
-colnames(pollutiontable) <- c("(1)", "(2)", "(3)", "(4)")
-kable(
-      pollutiontable, caption = "Wind direction and particulate matter",
-      align = "cccc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
-           ) %>%
-  column_spec(1, width = "4cm") %>%
-  column_spec(c(2:5),width = "2cm") %>%
-  row_spec(c(3), bold = TRUE) %>%
-  row_spec(6, hline_after = TRUE) %>%
-  add_header_above(c(" ", "1998-2015" = 2, "2002-2013" = 2)) %>%
-  kable_classic_2()
-
-```
 
@@
 
+### Identifying effects of pollution from coal plants
+
+I use the location of coal plants to identify locations of high pollution, but there can be other pollution sources located near coal plants, as well.
+I nonetheless want to identify the effects of pollution coming specifically from coal plants.
+I do this by combining the two-stage procedure above with a differences-in-differences framework.
+Specifically, I create a new variable identifying the exact year the coal plant opens in a given area and calculate the change in the effects of pollution on agricultural productivity.
+
+Consider the following regression of interest:
+\begin{gather} 
+y_{it} = \alpha_{i} + \gamma_{t} + \omega pm_{it} + \phi open_{it} + \beta pm_{it}\times open_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it},
+\end{gather}
+
+where the interaction term $pm_{it}\times open_{it}$ indicates the change in the effect of pollution following the opening of the coal plant.
+Note that this is not the same as saying that pollution increased; instead, it is focusing on the change in the effect of a change in pollution of one unit on the outcome of interest, agricultural yield.
+I instrument for both $pm_{it}$ and $pm_{it}\times open_{it}$ using wind and the interaction between wind and the open indicator, or $wind_{it}\times open_{it}$.
+
 # Results
@@
 
-This section presents the main results of this paper.
-I present three sets of results: one set looking at agricultural productivity, another looking at labor allocation, and a final set looking at nightlights, which are a proxy for economic development.
-I go through each of these in turn.
+This section presents the main results of this paper. I present two main sets of results: one set looking at agricultural productivity and another looking at labor allocation.
 
@@
 
-I first present results for agricultural productivity.
+I first present results for agricultural productivity, with the reduced form.
 The outcome of all the regressions in this section is the log of agricultural land productivity, which is defined as tons per hectare.
 The first set of results are in Table \@ref(tab:yieldtable).
-The unit of analysis is the vilalge-year-season -- from 2002 to 2013 -- the geographic fixed effect is the village-season, and the temporal fixed effect is the year.
-The exposure variable is defined for the entire season.
-Since @gangopadhyay2022new define both the monsoon (kharif) season and the winter (rabi) season as five months long, I do as well, using the same five months as the authors.
-This implies that the exposure variable ranges from zero to slightly more than 150.
+The unit of analysis is the village-year -- from 2002 to 2013 -- focusing only on the monsoon (kharif) season.
+The data used to construct the yield measure is calculated using five months of data, so the wind variable is defined for the same five months and is the sum of days in which the wind is pointed towards the village centroid from a coal plant.
 
 The first column presents the most simple results, with only the wind exposure variable and the fixed effects.
-The second column adds a rainfall -- defined as the total rainfall in the season -- and average temperature variable, the addition of which is motivated by the fact that changes in wind direction could be correlated with changes in weather.
-However, the coefficient on the wind exposure variable is completely unchanged by the addition of the two weather variables, indicating that weather may not be as worrisome of a confounder as thought.
-The direction of the weather coefficients is also consistent with expectations; higher rainfall is good for agricultural productivity while higher temperature is not.
-
-Column three adds district-by-year-by-season fixed effects as an additional robustness checks.
-This addition decreases the magnitude of the exposure variable but it remains negative and significant.
-However, part of this may be due to the fact that the new fixed effects soak up much of the variation in the exposure variable, as shown in Figure \@ref(fig:windexample).
-The last two columns separate the effects into the monsoon season and the winter season, but the overall effect is roughly similar, though rainfall has a larger effect in the monsoon season than the winter season, which is unsurprising given the winter season is driven by irrigation in many areas.
+The second column adds a rainfall and average temperature variable, while the third column includes higher-order weather variables, as well.
+While the coefficient decreases slightly in magnitude as we move from the first to the second column and the second to the third column, the decrease is relatively small.
+There is a consistent, negative effect of wind exposure on agricultural productivity.
 
@@
 # Table
-colnames(yieldtable) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
+colnames(yield1reducedform) <- c("(1)", "(2)", "(3)")
 kable(
-      yieldtable, caption = "Wind direction and agricultural productivity",
-      align = "ccccc", booktabs = TRUE, linesep = ""
+      yield1reducedform, caption = "Wind direction and agricultural productivity",
+      align = "ccc", booktabs = TRUE, linesep = ""
       ) %>%
@@
   column_spec(1, width = "3cm") %>%
-  column_spec(c(2:6),width = "2cm") %>%
-  row_spec(7, bold = TRUE) %>%
-  row_spec(11, hline_after = TRUE) %>%
-  add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
+  column_spec(c(2:4),width = "2.5cm") %>%
+  row_spec(13, bold = TRUE) %>%
+  row_spec(15, hline_after = TRUE) %>%
+  #add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
   kable_classic_2()
@@
 To put the size of the coefficient in context, it is worth digging a little deeper into the exposure variable.
-The mean within-village absolute deviation in wind exposure is approximately 8.06 days, meaning that the average change in agricultural productivity from year-to-year due to nothing but changes in wind patterns carrying particulate matter is around 2.4 percent relative to the mean.
-This implies that swings in agricultural productivity -- due to wind direction -- of more than five percent are probably quite common, since the absolute deviation includes deviations below and above the mean.
+The mean within-village absolute deviation in wind exposure is approximately 2.42 days, meaning that the average change in agricultural productivity from year-to-year due to nothing but changes in wind patterns carrying particulate matter is around 1.5 percent relative to the mean.
+This implies that swings in agricultural productivity -- due to wind direction -- of several percent are probably quite common, since the absolute deviation includes deviations below and above the mean.
 
@@
 # Table
-colnames(yieldtablepm) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
+colnames(yield2naive) <- c("(1)", "(2)", "(3)")
 kable(
-      yieldtablepm, caption = "Particulate matter (PM 2.5) and agricultural productivity",
-      align = "ccccc", booktabs = TRUE, linesep = ""
+      yield2naive, caption = "Particulate matter (PM 2.5) and agricultural productivity",
+      align = "ccc", booktabs = TRUE, linesep = ""
       ) %>%
@@
   column_spec(1, width = "3cm") %>%
-  column_spec(c(2:6),width = "2cm") %>%
-  row_spec(7, bold = TRUE) %>%
-  row_spec(11, hline_after = TRUE) %>%
-  add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
+  column_spec(c(2:4),width = "2.5cm") %>%
+  row_spec(5, bold = TRUE) %>%
+  row_spec(7, hline_after = TRUE) %>%
+  #add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
   kable_classic_2()
@@
 Instead, we are actually interested in better understanding the effects of exposure to pollution on agricultural productivity.
-Table \@ref(tab:yieldtablepm) presents the same regressions as Table \@ref(fig:yieldtable), with a key difference: instead of wind exposure, it uses average concentration of $\mathrm{PM_{2.5}}$ -- from @hammer2020global -- during the season as the key explanatory variable.
-
-Several interesting patterns emerge.
-We only observe a negative relationship between particulate matter and agricultural productivity in column three; the coefficient in column two is actually positive.
-Breaking out the effects into the monsoon and winter seasons, we also see a positive relationship between pollution and agricultural productivity.
+Table \@ref(tab:yieldtablepm) presents the same regressions as Table \@ref(fig:yieldtable), with a key difference: instead of wind exposure, it uses logged average concentration of $\mathrm{PM_{2.5}}$ -- from @hammer2020global -- during the season as the key explanatory variable.
+There is again a consistent, negative relationship between particulate matter and agricultural productivity.
+Since both variables are logged, the coefficients are similar to elasticities, indicating that a one-percent increase in pollution leads to a 0.03-to-0.04 decrease in agricultural productivity.
 
@@
 To recover the actual effect of pollution, we need an instrument.
-If, conditional on the fixed effects and weather variables, wind direction is exogenous and only affects agricultural productivity through pollution -- which seems reasonable since it is hard to imagine what other variables it might affect -- then it is a valid instrument for pollution.
-
-I present the IV results in Table \@ref(tab:yieldtabletwo), with the second-stage results in the top panel and the first-stage results in the bottom panel.
-The first-stage results make clear how strong of an instrument wind is; the t-statistic in column one, for example, is almost 48 while it is still well above 10 in column three, which translates to a first-stage F-statistic of 195, meeting even the higher bar for the first-stage F-test argued for by @lee2022valid.
+If, conditional on the fixed effects and weather variables, wind direction is exogenous and only affects agricultural productivity through pollution, then it is a valid instrument for pollution.
+
+I present the IV results in Table \@ref(tab:yieldtabletwo), with the first-stage F-statistic in the bottom.
+The first-stage results make clear how strong of an instrument wind is; all three F-statistics are above 1,000, meeting even the higher bar for the first-stage F-test argued for by @lee2022valid.
 Also note that the coefficients are quite different from those in Table \@ref(tab:pollutiontable) because the unit of analysis in Table \@ref(tab:yieldtabletwo) is the season, while in Table \@ref(tab:pollutiontable) the unit of analysis is the month.
@@
 The second-stage results first indicate that $\mathrm{PM_{2.5}}$ is a very strong predictor of agricultural productivity.
-The smallest coefficient -- that for the monsoon season in column (4) -- and the mean absolute deviation for $\mathrm{PM_{2.5}}$ point to a change in agricultural productivity of around 23 percent, which is much higher than the wind results documented above.
-This makes sense, of course, since the IV is essentially scaling up the effects of wind.
+Using the results in column three, a one percent increase in particulate matter leads to a 0.73 percent decrease in agricultural productivity.
+To put this in further perspective, the mean particulate matter concentration is just short of 40 (39.68) in the monsoon sample, while the within-village absolute deviation in particulate matter is 3.98, meaning year-to-year swings of 7-to-14 percent in agricultural productivity due to pollution exposure are common.
 
@@
 # Table
-colnames(yieldtabletwo) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
+colnames(yield3ivmain) <- c("(1)", "(2)", "(3)")
+rownames(yield3ivmain)[8] <- "F (first stage)"
 kable(
-      yieldtabletwo, caption = "Pollution and agricultural productivity, IV estimates",
-      align = "cccc", booktabs = TRUE, linesep = ""
+      yield3ivmain, caption = "Pollution and agricultural productivity, IV estimates",
+      align = "ccc", booktabs = TRUE, linesep = ""
       ) %>%
@@
   column_spec(1, width = "3cm") %>%
-  column_spec(c(2:6),width = "2cm") %>%
-  row_spec(c(7,13), bold = TRUE) %>%
-  row_spec(c(11:12), hline_after = TRUE) %>%
-  add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
+  column_spec(c(2:4),width = "2cm") %>%
+  row_spec(5, bold = TRUE) %>%
+  row_spec(7, hline_after = TRUE) %>%
+  #add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
   kable_classic_2()
@@
 In other words, for each village, I calculate the median exposure (based on wind) and then use this single value for each village to split the sample into villages above and below median exposure.
-The coefficient in the first column is 2.5 times larger than the coefficient in the second column, indicating much larger effects of pollution in places where it is more common.
+The coefficient in the first column is 2.3 times larger than the coefficient in the second column, indicating much larger effects of pollution in places where it is more common.
 One possible explanation is that exposure is cumulative, leading to larger and larger effects.
@@
 
-Columns two and three split the sample based on median yield at the very beginning of the sample (the first year available, 2002-2003).
+Columns three and four split the sample based on median yield at the very beginning of the sample (the first year available, 2002-2003).
 Column three presents results for villages above the median while column four presents results for villages below the median.
-The largest effects are apparently concentrated in the lower half of the distribution.
-Perhaps more productive areas are more likely to have ways to combat the effects of shocks -- e.g.
-by adjusting the use of other inputs -- or maybe less productive areas are inherently more susceptible to the effects of pollution, perhaps due to direct effects on land productivity.
+The largest effects are apparently concentrated in the upper half of the distribution, meaning the most productive areas are also most susceptible to the effects of pollution.
 
 Given ongoing changes in the climate, a key question is how welfare will change as shocks become more and more common.
-In the last column looks at possible multiplicative effects of shocks by interacting pollution with rainfall and seeing whether the effects compound one another.
+The last column looks at possible multiplicative effects of shocks by interacting pollution with rainfall and seeing whether the effects compound one another.
 Since there are now two endogenous regressors, I instrument for both using wind exposure and an interaction between wind exposure and rainfall.
@@
 # Table
-colnames(yieldtablehet) <- c(">p(50)", "<=p(50)", ">p(50)", "<=p(50)", "")
-rownames(yieldtablehet)[c(1, 3)] <- c("PM 2.5", "PM time rain")
+colnames(yield5heterogeneity) <- c(">p(50)", "<=p(50)", ">p(50)", "<=p(50)", "")
+rownames(yield5heterogeneity)[c(1, 3)] <- c("PM 2.5", "PM times rain")
 kable(
-      yieldtablehet, caption = "Heterogeneity in the effects of pollution on productivity",
+      yield5heterogeneity, caption = "Heterogeneity in the effects of pollution on productivity",
       align = "ccccc", booktabs = TRUE, linesep = ""
@@
   row_spec(c(6), bold = TRUE) %>%
-  add_header_above(c(" ", "exposure" = 2, "yield" = 2, " " = 1)) %>%
+  add_header_above(c(" " = 1, "wind" = 2, "yield" = 2, " " = 1)) %>%
   kable_classic_2()
@@
 
-## Robustness checks
-
-I present two separate robustness checks in the appendix.
-The first set of analyses attempts to separate the possible endogenous location of coal plants from outcomes.
-Specifically, prior to the opening of a new coal plant in an area, all villages receive a zero for wind.
-This means that, even if wind direction is exogenous, the opening of the coal plant could lead to correlations between the error term and the wind variable.
-
-Although the results with district-year-season help mitigate this concern, Table \@ref(tab:yieldtablepostplant) in the appendix checks this possibility in a different way.
-Specifically, I restrict the sample to only village-years *after* the introduction of a coal plant.
-This drops any pre-plant-opening observations -- all of which are mechanically zeros for wind -- to prevent endogeneity due to the opening.
-The coefficients in Table \@ref(tab:yieldtablepostplant) show that the overall effects are essentially unchanged across any specifications.
-Comparing changes in coefficients can also help reduce any effects driven by other polluting industries located near coal plants; since the pre-opening exposure variable is zero, the coefficients in Table \@ref(tab:yieldtablepostplant) would likely increase in magnitude if the effects were driven by non-coal pollution.
-
-The second robustness check includes leads and lags for wind to see if they predict current year productivity.
-I present these results in \@ref(tab:yieldtableleads).
-The first column is the reduced form effect and shows that only current wind predicts productivity; leads and lags have no impact.
-We see similar results in column two -- the IV results -- with a negative coefficient for current pollution, only.
-However, the standard errors are much larger than the main results -- around five times as large -- making it hard to come to any conclusions.
+In the appendix, I present in Table \@ref(tab:yieldtableleads) as a robustness check similar regressions but also including wind/pollution leads.
+The negative effect is only seen with current pollution, not future pollution.
 
@@
 Figure \@ref(fig:laborplot) in the appendix shows this temporal distribution across the five waves of the survey.
-Since I can no longer include district-by-month fixed effects, I instead opt for district-level trends to help take into account aggregate temporal changes in outcomes.
-
-```{r labortablerural, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+Since pollution is not disaggregated at a daily level, I instead stick to the reduced form results with wind.
+
+```{r labortable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
 
 # Table
-colnames(labortablerural) <- c("all", "all", "self", "wage", "farm", "non-farm")
-colnames(labortableurban) <- c("all", "all", "self", "wage", "farm", "non-farm")
+colnames(labor1all) <- c("all", "all", "self", "wage", "farm", "non-farm")
+colnames(labor2rural) <- c("all", "all", "self", "wage", "farm", "non-farm")
 
@@
 colnames(mat1) <- c("all", "all", "self", "wage", "farm", "non-farm")
-rownames(mat1) <- c("rural")
+rownames(mat1) <- c("panel a: all")
 mat2 <- matrix(c("", "", "", "", "", ""), nrow = 1)
 colnames(mat2) <- c("all", "all", "self", "wage", "farm", "non-farm")
-rownames(mat2) <- c("urban")
-
-labortablerural <- rbind(mat1, labortablerural[c(1:2, 7:8),], mat2, labortableurban[-c(3:6),], labortableurban[c(3:6),])
+rownames(mat2) <- c("panel b: rural")
+
+labortablerural <- rbind(mat1, labor1all[c(1:2, 7:8),], mat2, labor2rural[-c(3:6),], labor2rural[c(3:6),])
 
@@
 
-There are some interesting results in Table \@ref(tab:labortable).
-The first two columns show a marginally significant decrease in overall labor allocation (where "labor" is defined as non-domestic self and wage employment).
-The third and fourth columns show that this overall decrease is predominantly from wage employment, not self employment.
-An additional day of wind over the last seven decreases wage employment by 0.034 days in the district.
-Mean wage employment is 1.458 days, meaning one day of wind decreases wage employment by 2.3 percent of the mean.
-Since much of the district is not covered by "treated" villages, this is actually a rather substantive effect on the villages involved.
-
-The last two columns split the effects by sector, into farm and non-farm employment.
-Although agricultural (land) productivity decreases when wind blows in the direction of villages, overall labor allocation to farm labor actually goes *up*, by around 4.5 percent of the mean of 0.784 days.
-Non-farm labor, on the other hand, goes down, by around 2.4 percent of the mean of 2.527 days.
-
-There are two possible explanations for this pattern of results.
-First, it could be that land productivity goes down, leading farmers to increase their farm labor allocation in an attempt to compensate for the decreased productivity.
-A second possibility is that pollution does not affect land productivity at all, at least not directly.
-In this scenario, the worsening air quality could lead to decreased labor productivity, meaning that "effective labor" essentially decreases when pollution is higher.
-
-Table \@ref(tab:labortablemonsoon) and Table \@ref(tab:labortablewinter) in the appendix disaggregate these effects into the monsoon and winter seasons.
-Apparently, the entirety of the effect is driven by changes in labor allocation during the monsoon season.
-The dynamics across seasons apparently lead to large differences in the effects of pollution on labor allocation.
-
-One way to provide suggestive evidence of mechanisms is to explore heterogeneity in the effects.
-Previous research has shown larger effects of pollution on the health of the oldest [@deryugina2019mortality].
-The labor sample includes all adults between 15 and 60 years of age.
-To look at heterogeneity by age, I present an additional set of results interacting the wind variable with an indicator for whether an individual is over 37 years of age.[^13] These results are in Table \@ref(tab:labortableold), which otherwise presents the same specifications in the last five columns as \@ref(tab:labortable).
-
-[^13]: This splits the sample into two equal bins of age: those who are 15-37 and those who are 38-60.
-
-Table \@ref(tab:labortableold) presents a very clear set of results: the effects are always larger (more negative) for older adults.
-Particularly interestingly, we now see that the mean-zero effect on self employment in the previous table is driven by two stark contrasts: a large increase for the younger group and a large decrease for the older group.
-For the youngest group, this increase appears in farm labor, indicating likely increases in farm self-employment.
-The oldest, on the other hand, have large decreases in both farm and non-farm labor, with larger decreases in the latter.
-
-```{r labortableruralage, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+The top panel presents results for all respondents during the five-month monsoon season while the bottom panel presents results for respondents in rural areas.
+Interestingly, we do not see any signs of large, negative effects of wind exposure on labor allocation across any sectors, with the possible exception of wage employment in rural areas.
+However, the coefficient for self-employment in rural areas is of the same magnitude and *positive*, but imprecisely estimated, making it hard to come to any definitive conclusions, at least in the aggregate.
+
+## Month-specific exposure
+
+While overall results for labor allocation are inconclusive, perhaps we can learn more by looking at temporally disaggregated results, focusing on May through December.
+Although the yield data was created by looking at five specific months of the year -- June through October -- this is partially because this is the time of year when remote-sensing data is most able to pick up overall changes in crop cover.
+Preparation, however, starts before that, with operations such as ploughing often taking place prior to June.
+
+Since the NSS data has survey dates and the both the pollution and wind data can be aggregated to the month, we can look at how exposure during different months of the agricultural season (differentially) affect labor allocation and overall agricultural productivity.
+I begin with labor allocation in Table \@ref(tab:labor3month).
+As before, the exposure measure is just the wind variable (over the last seven days), since I can directly match daily wind direction with the timing of the survey, but I cannot do that for pollution, which is monthly.
+The first column presents results for farm labor and the second column presents results for non-farm labor.
+
+Higher exposure leads to significantly less farm labor labor in June and July, but there are no discernible effects in other months of the year.
+There are likewise no discernible effects on non-farm labor, but some of the coefficients are rather imprecisely estimate.
+To put the effects for June and July in context and taking the coefficients at face value, an additional day of wind in the last seven days decreases farm labor in June by around `r paste0(reportPercent(0.136/1.209))` percent of the mean and in July by around `r paste0(reportPercent(0.285/1.671))` percent of the mean.
+
+```{r labor3month, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
 
 # Table
-colnames(labortableruralyoung) <- c("all", "all", "self", "wage", "farm", "non-farm")
-colnames(labortableruralold) <- c("all", "all", "self", "wage", "farm", "non-farm")
-
-mat1 <- matrix(c("", "", "", "", "", ""), nrow = 1)
-colnames(mat1) <- c("all", "all", "self", "wage", "farm", "non-farm")
-rownames(mat1) <- c("younger than 35")
-mat2 <- matrix(c("", "", "", "", "", ""), nrow = 1)
-colnames(mat2) <- c("all", "all", "self", "wage", "farm", "non-farm")
-rownames(mat2) <- c("45 or older")
-
-labortableruralyoung <- rbind(mat1, labortableruralyoung[c(1:2, 7:8),], mat2, labortableruralold[-c(3:6),], labortableruralold[c(3:6),])
-
+colnames(labor3month) <- c("farm", "non-farm")
 kable(
-      labortableruralyoung, caption = "Wind direction and labor allocation by age (rural only)",
-      align = "cccccc", booktabs = TRUE, linesep = ""
+      labor3month, caption = "Wind and labor allocation, by month",
+      align = "cc", booktabs = TRUE, linesep = ""
       ) %>%
   footnote(
-           "Note: Standard errors are in parentheses and are clustered at the district level.
-Control variables include female, age, age squared, and (years of) education.
-The outcome in each regression is days worked in the previous seven days.
+           "Note: Standard errors are in parentheses and are clustered at the village level. The outcome in column one is .  * p<0.10 ** p<0.05 *** p<0.01", 
+           general_title = "",
+           threeparttable = TRUE,
+           footnote_as_chunk = TRUE
+           ) %>%
+  column_spec(1, width = "3cm") %>%
+  column_spec(c(2:3),width = "2cm") %>%
+  row_spec(c(18), hline_after = TRUE) %>%
+  row_spec(c(16), bold = TRUE) %>%
+  kable_classic_2()
+
+```
+
+I perform a similar exercise with yield, the results of which are in Table \@ref(tab:yield4monthly).
+I include only the five months of the year that are used in the calculation of the yield estimates, or June through October.
+The results are again from two-stage least squares, with the F-statistics for each month presented at the bottom; monthly wind remains a strong predictor of pollution.
+Column one presents results without controls for weather while column two presents results with controls for weather; the differences in coefficients make clear how important the inclusion of weather is here.
+
+Focusing on column two, there is only one month where we see a clear effect of pollution on yield: June.
+Given that we also see negative effects of pollution on farm labor allocation in June, this is suggestive -- but not conclusive -- evidence that pollution works through labor.
+Note that this does not rule out pollution also having direct effects on agricultural productivity, however.
+
+```{r yield4monthly, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+# Table
+colnames(yield4monthly) <- c("(1)", "(2)")
+kable(
+      yield4monthly, caption = "Pollution and productivity, by month",
+      align = "cc", booktabs = TRUE, linesep = ""
+      ) %>%
+  footnote(
+           "Note: Standard errors are in parentheses and are clustered at the village level.
+The outcome in all regressions is estimated agricultural yield, defined as tons per hectare.
 * p<0.10 ** p<0.05 *** p<0.01",
@@
   column_spec(1, width = "3cm") %>%
-  column_spec(c(2:7),width = "1.5cm") %>%
-  row_spec(c(5,10), hline_after = TRUE) %>%
-  row_spec(c(1,6,12), bold = TRUE) %>%
+  column_spec(c(2:3),width = "2cm") %>%
+  row_spec(12, bold = TRUE) %>%
+  row_spec(14, hline_after = TRUE) %>%
+  #add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
   kable_classic_2()
@@
 
-To put some of these changes in context, an additional day of exposure to pollution from wind increases self-employment for the younger half of the sample by around six percent of the mean and increases farm employment for the younger half by almost 10 percent of the mean.
-For the older half, on the other hand, an additional day of exposure decreases self-employment and non-farm employment by around five percent of their means.
-
-While speculative, one possible explanation for this pattern is a decrease in labor productivity.
-While the youngest may be able to readjust and increase their overall self-employment labor allocation, perhaps the oldest cannot due to, for example, health effects of pollution, which are well documented [@bentayeb2012adverse; @pope2000epidemiology].[^14] We see decreases in overall agricultural productivity, despite an increase in overall farm labor.
-We also see decreases in non-farm employment, indicating that individuals are reallocating labor from the non-farm sector to the farm sector.
-
-[^14]: Much of the literature also suggests the youngest are at significant risk of mortality and morbidities. However, given my focus on labor allocation, I only discuss older adults here.
-
-It is worth noting that the larger negative effects for older adults is more consistent with a labor story than a land story; it is hard to imagine these patterns emerging if the only effects were through land productivity changes.
-The overall decrease in non-farm employment is more compatible with a labor health/productivity story, though with general equilibrium effects in agriculture spillover over into the non-farm sector, this need not necessarily be the case.
-
-In addition, Table \@ref(tab:labortableoldstate) in the appendix presents one additional set of results to support a labor story.
-Changes in land productivity should be relatively fixed in a short period of time.
-For example, if we focus on just a single month in a single district, the overall effects driven by changes in land productivity should be constant even though pollution may have differential effects throughout the month on labor productivity, which responds more immediately to changes in ar quality.
-Table \@ref(tab:labortableoldstate) changes the specification to include district-by-year-by-month fixed effects.
-In other words, the results compare people in the same district in the same month, but who may have differential exposure to pollution simply because they were interviewed at a slightly different time.
-
-The overall pattern of results stays the same when we shift the focus.
-Overall labor allocation increases for the youngest, but decreases for the oldest, and this is driven completely by self employment.
-There are decreases in both farm and non-farm employment for the oldest, while the youngest show increases in non-farm employment, but not farm employment -- which is a change from the previous results.
-The overall story remains the same, however, which is further evidence of effects being driven primarily through changes in labor allocation and productivity.
-
-Finally, Table \@ref(tab:wagestable) in the appendix shows that there are no significant changes in daily wages, though the overall direction of wage changes would be ambiguous, anyway.
-While reduced productivity might lead to lower wages, decreased labor supply could also put upward pressure on wages.
-
-## Month-specific exposure
-
-Pollution leads to lower agricultural productivity as well as changes in labor allocation.
-This leads to an obvious question: might higher levels of pollution also directly affect economic growth?
-To explore this question, I look at changes in nightlights relative to changes in pollution.
-It seems reasonable to think that pollution might affect nightlights with a lag, so I include the lag of pollution rather than contemporaneous pollution.
-
-```{r yieldtablemonthly, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+A final caveat about these month-specific results is that they do not take into account possible interactions between different months.
+For example, yield might respond strongly to average pollution in June and July, which would not be captured in Table \@ref(tab:yield4monthly).
+
+## Coal plants or overall pollution?
+
+Finally, while I identify high-pollution areas using the location of coal plants, an important question is how much of the overall effect is driven by coal plants and how much is driven by other polluting industries located in areas near coal plants.
+I try to answer this question in Table \@ref(tab:yield7ivdiffindiff), where I implement an instrumental variables differences-in-differences, essentially instrumenting for pollution prior to the opening of a coal plant with wind and instrumenting for the interaction term (pollution times plant opening) with wind times plant opening.
+
+```{r yield7ivdiffindiff, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
 
 # Table
-
+colnames(yield7ivdiffindiff) <- c("(1)", "(2)", "(3)")
 kable(
-      yieldtablemonthly, caption = "Month-specific exposure",
-      align = "cc", booktabs = TRUE, linesep = ""
+      yield7ivdiffindiff, caption = "Agricultural productivity and pollution source",
+      align = "ccc", booktabs = TRUE, linesep = ""
       ) %>%
   footnote(
-           "Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions is logged nightlights. * p<0.10 ** p<0.05 *** p<0.01", 
+           "Note: Standard errors are in parentheses and are clustered at the village level.
+The outcome in all regressions is estimated agricultural yield, defined as tons per hectare.
+* p<0.10 ** p<0.05 *** p<0.01",
            general_title = "",
@@
   column_spec(1, width = "3cm") %>%
-  column_spec(c(2:3),width = "2cm") %>%
-  row_spec(c(13), hline_after = TRUE) %>%
-  row_spec(c(11), bold = TRUE) %>%
+  column_spec(c(2:4),width = "2cm") %>%
+  row_spec(c(7), bold = TRUE) %>%
+  row_spec(9, hline_after = TRUE) %>%
   kable_classic_2()
@@
 
-I present these results in Table \@ref(tab:ntltable).
-To help simplify the coefficients, I recode wind to be the proportion of days in the year that a given village experiences wind in their direction, from a coal plant.
-Mean wind in this new specification is 0.066, indicating that a village in the sample experiences wind in its direction on around 6.6 percent of days, or 24 days of the year.
-
-We see a strong, negative relationship between pollution (lagged wind) and nightlights.
-Going from the 25th percentile to the 75th percentile of wind exposure is synonymous with a decrease in the growth of nightlights within a village by around 0.7 percent.
-While this is a relatively small magnitude, it is worth noting that there are more than 100,000 villages in this sample.
-While the effect for a single village might be small, the overall, aggregate effect on growth is quite substantial.
-To put this number in perspective, @castello2018higher argue that a standard-deviation increase in the share of adults with a university education leads to 12 percent higher nightlights in India, while @storeygard2016farther shows that a four-fold increase in oil prices over six years leads to seven percent lower nightlights at a distance of around 500km from the nearest large port in sub-Saharan Africa.
-In other words, the 0.7 percent change in nightlights is quite large given this is annual variation.
+Columns one through three of Table \@ref(tab:yield7ivdiffindiff) present specifications with progressively increasing weather controls.
+Column three presents the most extensive set of controls and I focus on that column here.
+First, there are now two first stages but the F-test indicates that the instruments are strong for both, including the interaction term (F = 232).
+Second, the first row shows the effect of pollution on agricultural productivity prior to the opening of a coal plant; the coefficient is negative but relatively smaller.
+Third, the interaction term is strongly negative and more than twice as large in magnitude as the level term, indicating a very clear effect specific to coal plants.
+This could be due to differences in emission patterns for different kinds of plants and factors; coal plants, for example, are responsible for much more mercury than other kinds of polluting industries.^[https://www.epa.gov/sites/default/files/2016-12/documents/nei2014v1_tsd.pdf]
 
@@
 
+Pollution is a growing problem throughout the developing world.
+India, for example, has some of the worst air quality on earth.
 The rapid growth of coal power plants has also led to increases in emissions from the burning of coal.
-In this paper, I show some of the negative effects associated with this burning: decreased agricultural productivity.
-I find clear indications that pollution from coal plants leads to lower agricultural productivity.
-Importantly, I use exogenous changes in wind direction to measure exposure to pollution and show that these effects are robust to a range of alternative specifications.
-
-Results indicate that these effects may be driven by changes in labor allocation patterns.
-In addition, the negative labor effects are concentrated in older adults, which is consistent with previous research on the effects of pollution on health in adults.
-Despite this, I am unable to rule out direct effects on land productivity with the data available.
-
-Finally, these negative effects are also seen in aggregate economic development, with pollution leading to decreases in the growth of nightlights.
-These effects, though small for an individual village, are quite large in overall terms, indicating that pollution from coal plants can have effects on overall economic development.
+In this paper, I show some of the negative effects associated with pollution, in general, and coal pollution, in particular.
+Using exogenous variation in pollution driven by changes in wind direction, results here show that pollution has a large, negative effect on agricultural productivity in India.
+
+These negative effects are at least partly due to changes in labor allocation decisions; the largest negative effects on agricultural productivity and on labor allocation happen in the same month of the season: June.
+Nonetheless, I am unable to rule out direct effects on land productivity with the data available.
+
+Using instrumental variables in a differences-in-differences set up, I also show that the effects of pollution from coal, in particular, are particularly harmful to agricultural productivity.
+Changes of an equivalent increase in pollution lead to three times larger negative effects on agricultural productivity following the opening of a coal plant, compared to just before.
 
@@
 data[4,1] <- "Hammer et al. (2020)"
-data[5,1] <- "Angopadhyay et al. (2022)"
+data[5,1] <- "Gangopadhya et al. (2022)"
 data[6,1] <- "Asher et al. (2021)"
@@
 ```
-```{r yieldtablehet1st, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+```{r plants, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Coal plants in India from 1990 to 2010", fig.footer = list("Note: The top figure shows the location of coal plants in 1990.
+The bottom figure shows the location of coal plants in 2010.")}
+
+plant1 <- ggplot() +
+                theme_map() +
+                geom_sf(data = india_shape2, fill = NA, lwd = 0.01) +
+                geom_sf(data = plants1990, size = 0.75, color = "black") +
+                labs(x = "lon",
+                     y = "lat",
+                     subtitle = "Figure A: 1990")
+
+plant2 <- ggplot() +
+                theme_map() +
+                geom_sf(data = india_shape2, fill = NA, lwd = 0.01) +
+                geom_sf(data = plants2010, size = 0.75, color = "black") +
+                labs(x = "lon",
+                     y = "lat",
+                     subtitle = "Figure B: 2010")
+
+ggarrange(plant1, plant2)
+
+```
+
+```{=tex}
+\FloatBarrier
+\newpage
+```
+
+```{r windexample, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 8, fig.cap = "Wind direction and aggregation examples (2010-01-01)", fig.footer = list("Note: The top figure shows the average wind direction on January 1st, 2010, with zero degrees indicating north.
+The points are the location of coal plants on that date.
+The bottom figure shows the distribution of pollution exposure in a specific district -- Guna district in Madhya Pradesh -- on the same date.")}
+
+p1 <- ggplot() + 
+        theme_map() +
+        geom_sf(data = india_shape, fill = NA) +
+        geom_tile(data = windexample, aes(x = x, y = y, fill = wind_example), alpha = 0.3) +
+        scale_fill_viridis_c("wind direction\n(degrees)", option = "viridis") +
+        geom_sf(data = plants2010, size = 0.5) +
+        labs(x = "lon",
+             y = "lat",
+             subtitle = "Figure A: Wind direction") +
+        theme(legend.position = c(0.6, 0.05),
+              legend.direction = "vertical",
+              legend.key.height = unit(0.4, 'cm'),
+              legend.text = element_text(size = 8),
+              legend.background=element_blank() )
+
+p2 <- ggplot() + 
+        theme_map() +
+        geom_sf(data = districts, fill = NA) +
+        geom_sf(data = villages %>% filter(is.na(days_sum)==F), aes(fill = days_sum), lwd = NA) +
+        scale_fill_viridis_c("exposure (days)", option = "viridis") +
+        labs(x = "lon",
+             y = "lat",
+             subtitle = "Figure B: Aggregation to district") +
+        theme(legend.position = c(0.5, 0.05),
+              legend.direction = "horizontal",
+              legend.key.width = unit(0.3, 'cm'),
+              legend.text = element_text(size = 8),
+              legend.background=element_blank() )
+
+ggarrange(p1, p2,
+          nrow = 2)
+
+```
+
+```{=tex}
+\FloatBarrier
+\newpage
+```
+
+```{r pollutiontable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
 
 # Table
-colnames(yieldtablehet1st) <- c("PM", "PM", "PM", "PM", "PM", "PM*rain")
+colnames(pollutiontable) <- c("(1)", "(2)", "(3)", "(4)")
 kable(
-      yieldtablehet1st, caption = "Heterogeneity in the effects of pollution on productivity, first-stage results",
-      align = "cccccc", booktabs = TRUE, linesep = ""
+      pollutiontable, caption = "Wind direction and particulate matter",
+      align = "cccc", booktabs = TRUE, linesep = ""
       ) %>%
   footnote(
-           "Note: Standard errors are in parentheses and are clustered at the village level.
-The outcome in all regressions is estimated agricultural yield, defined as tons per hectare.
-The first two columns split the sample based on the median exposure (by village).
-The second two columns split the sample based on initial -- the first year available -- agricultural productivity.
-* p<0.10 ** p<0.05 *** p<0.01",
+           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
            general_title = "",
@@
            ) %>%
-  column_spec(1, width = "2.8cm") %>%
-  column_spec(c(2:7),width = "1.7cm") %>%
-  row_spec(c(8), hline_after = TRUE) %>%
-  row_spec(c(6), bold = TRUE) %>%
-  add_header_above(c(" ", "(1)" = 1, "(2)" = 1, "(3)" = 1, "(4)" = 1, "(5)" = 2)) %>%
+  column_spec(1, width = "4cm") %>%
+  column_spec(c(2:5),width = "2cm") %>%
+  row_spec(c(3), bold = TRUE) %>%
+  row_spec(6, hline_after = TRUE) %>%
+  add_header_above(c(" ", "1998-2015" = 2, "2002-2013" = 2)) %>%
   kable_classic_2()
@@
 ```
-```{r yieldtablepostplant, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+```{r yieldtableleads, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
 
 # Table
-colnames(yieldtablepostplant) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
+colnames(yield6leads) <- c("(1)", "(2)", "(3)")
 kable(
-      yieldtablepostplant, caption = "Wind direction and agricultural productivity, post-coal plant only",
-      align = "ccccc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the village level.
-The sample is restricted to village-years after a coal plant was built near the village.
-The outcome in all regressions estimated agricultural yield (tons per hectare).
-* p<0.10 ** p<0.05 *** p<0.01",
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
-           ) %>%
-  column_spec(1, width = "3cm") %>%
-  column_spec(c(2:6),width = "2cm") %>%
-  row_spec(c(7, 13), bold = TRUE) %>%
-  row_spec(c(11:12), hline_after = TRUE) %>%
-  add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  kable_classic_2()
-
-```
-
-```{=tex}
-\FloatBarrier
-\newpage
-```
-
-```{r yieldtableleads, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-# Table
-colnames(yieldtableleads) <- c("(1)", "(2)")
-kable(
-      yieldtableleads, caption = "Agricultural productivity and pollution leads",
-      align = "cc", booktabs = TRUE, linesep = ""
+      yield6leads, caption = "Agricultural productivity and pollution leads",
+      align = "ccc", booktabs = TRUE, linesep = ""
       ) %>%
@@
   column_spec(1, width = "3cm") %>%
-  column_spec(c(2:3),width = "2cm") %>%
-  row_spec(c(16), hline_after = TRUE) %>%
-  row_spec(c(14), bold = TRUE) %>%
+  column_spec(c(2:4),width = "2cm") %>%
+  row_spec(c(9), hline_after = TRUE) %>%
+  row_spec(c(7), bold = TRUE) %>%
   kable_classic_2()
@@
 ```
-```{r labortablemonsoon, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-# Table
-colnames(labortablemonsoon) <- c("all", "all", "self", "wage", "farm", "non-farm")
-kable(
-      labortablemonsoon, caption = "Wind direction and labor allocation, monsoon season only",
-      align = "cccccc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the district level.
-Control variables include female, age, age squared, and (years of) education.
-* p<0.10 ** p<0.05 *** p<0.01",
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
-           ) %>%
-  column_spec(1, width = "3cm") %>%
-  column_spec(c(2:7),width = "1.5cm") %>%
-  row_spec(c(8), hline_after = TRUE) %>%
-  row_spec(c(4,7), bold = TRUE) %>%
-  kable_classic_2()
-
-```
-
-```{=tex}
-\FloatBarrier
-\newpage
-```
-```{r labortablewinter, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-# Table
-colnames(labortablewinter) <- c("all", "all", "self", "wage", "farm", "non-farm")
-kable(
-      labortablewinter, caption = "Wind direction and labor allocation, winter season only",
-      align = "cccccc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the district level.
-Control variables include female, age, age squared, and (years of) education.
-The outcome in each regression is days worked in the previous seven days.
-* p<0.10 ** p<0.05 *** p<0.01",
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
-           ) %>%
-  column_spec(1, width = "3cm") %>%
-  column_spec(c(2:7),width = "1.5cm") %>%
-  row_spec(c(8), hline_after = TRUE) %>%
-  row_spec(c(4,7), bold = TRUE) %>%
-  kable_classic_2()
-
-```
-
-```{=tex}
-\FloatBarrier
-\newpage
-```
-```{r labortableoldstate, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-# Table
-colnames(labortableoldstate) <- c("all", "self", "wage", "farm", "non-farm")
-kable(
-      labortableoldstate, caption = "Wind direction and labor allocation, heterogeneity by age",
-      align = "ccccc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the district level.
-Control variables include female, age, age squared, and (years of) education.
-The outcome in each regression is days worked in the previous seven days.
-* p<0.10 ** p<0.05 *** p<0.01",
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
-           ) %>%
-  column_spec(1, width = "4cm") %>%
-  column_spec(c(2:6),width = "1.5cm") %>%
-  row_spec(c(7), hline_after = TRUE) %>%
-  row_spec(c(6), bold = TRUE) %>%
-  kable_classic_2()
-
-```
-
-```{=tex}
-\FloatBarrier
-\newpage
-```
-```{r wagestable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-# Table
-colnames(wagestable) <- c("farm", "non-farm", "all")
-kable(
-      wagestable, caption = "Wind direction and wages",
-      align = "ccc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the district level.
-Control variables include female, age, age squared, and (years of) education.
-The outcome in each regression is the daily wage (log Rs per day).
-* p<0.10 ** p<0.05 *** p<0.01",
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
-           ) %>%
-  column_spec(1, width = "3cm") %>%
-  column_spec(c(2:4),width = "1.5cm") %>%
-  row_spec(c(8), hline_after = TRUE) %>%
-  row_spec(c(4,7), bold = TRUE) %>%
-  kable_classic_2()
-
-```
-
-```{=tex}
-\FloatBarrier
-\newpage
-```
 ```{r monthplot, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Average farm days by season, rural only", fig.footer = list("Note: The line graph shows the mean number of farm days worked per week based on the month of the relevant agricultural season.

### 2023-03-02 draft/draft.tex
scripts and draft

@@
 \author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
-\date{2023-02-20}
+\date{2023-03-02}
 
@@
 \begin{abstract}
-\noindent I show that emissions from a common energy source lead to decreased productivity and growth.
-The rapid increase in population in India from the 1990s through the 2010s led to the rapid construction of new coal power plants throughout the country.
-Using this roll-out combined with random changes in wind direction, I document negative effects of pollution from coal power plants on agricultural productivity throughout the country.
-Results suggest this decrease may be driven by a decline in labor allocation of the oldest half of my sample, who greatly decrease labor allocated to both farm and non-farm employment when exposed to pollution.
-These impacts spill over into aggregate economic growth, as well; nightlight growth is slower following years of higher exposure to pollution.\\
+\noindent Pollution is a growing problem in many developing countries.
+Using the rollout of coal plants to identify high-pollution locations in India and random variation induced by changes in wind direction, I show that pollution has large, negative impacts on agricultural productivity.
+Results indicate that some of this impact is driven by changes in labor allocation patterns in response to increased exposure to pollution.
+I also show that pollution from coal plants, specifically, has a larger effect on agricultural productivity than other types of pollution.\\
 \strut \\
@@
 
-I study the effects of emissions from coal power plants on agricultural productivity, labor allocation, and economic growth in India.
+I study the effects of emissions from coal power plants on agricultural productivity in India.
 Previous work has suggested that there may be harmful direct effects of air pollution (Heck et al.
@@
 Using newly available estimates of India-wide agricultural (land) productivity, results show that the total number of days of being downwind from a coal power plant has a negative effect on overall agricultural productivity.
-Just a single additional day of being downwind from a coal power plant leads to a decrease in predicted yields of 0.3 percent.
-The within-village absolute deviation of wind exposure is around eight days, meaning that median changes in wind direction are responsible for fluctuations in agricultural productivity of around 2.4 percent, with changes of more than five percent from season to season being common.
-Using wind direction as an instrument for particulate matter, an increase of PM 2.5 by one microgram per cubic meter leads to an even larger decrease in agricultural productivity, at around 20 percent.
+Just a single additional day of being downwind from a coal power plant leads to a decrease in predicted yields of 0.6 percent.
+Combined with the within-village absolute deviation of wind exposure, results indicate that median changes in wind direction are responsible for fluctuations in agricultural productivity of around 1.5 percent.
+
+Using wind direction as an instrument for particulate matter, an increase of PM 2.5 by one percent leads to an even larger decrease in agricultural productivity, at around 0.73 percent.
+Given within-village changes across years, pollution could be responsible for swings of 7-to-14 percent in agricultural productivity across years.
 This IV strategy is similar to that in Deryugina et al.
@@
 
-These effects are apparently concentrated in villages that do not experience the most extreme exposure to pollution; villages with higher maximum exposure experience effects only around 10 percent as large as villages with lower maximum exposure.
-Additional heterogeneity analysis shows that, despite the results with respect to the maximum, the effects are increasing in \emph{mean} exposure.
-I also show that the effects are largest in villages with the highest initial agricultural productivity and that pollution and rainfall shocks are compounding; higher exposure along with less rainfall increases the negative effect of either shock individually.
+These effects are apparently larger in villages that experience higher median exposure to pollution and in villages with higher initial yields.
+I also show that rainfall shocks are compounding; higher exposure along with less rainfall increases the negative effect of either shock individually.
 As weather shocks become more common across the globe, the rise of pollution in India may be particularly problematic for farmers.
 
-Robustness checks show that this effect is concentrated in the years after the construction of the power plant, which helps decrease concerns around endogeneity since, conditional on the village fixed effect and weather (rainfall and temperature), wind direction across years and seasons is exogenous.
-Additionally, leads for wind direction do not predict current agricultural productivity, again supporting the identifying assumptions.
-
-While I am not able to explicitly test effects on land productivity, I look at how adults allocate labor over the last seven days using the National Sample Survey and creating a district-level exposure measure based on the timing of the survey for each household.
-An additional day of wind over the previous week decreases overall labor allocation by 0.055 days.
-However, the overall results show stark heterogeneity based on age; the older half of the sample greatly decreases labor allocation in response to pollution exposure while the younger half increases overall labor allocation.
-Older adults decrease labor allocation to both farm and non-farm labor, while younger adults increase labor allocated to farm labor.
-This is suggestive evidence of productivity effects being driven by the health and productivity effects on individuals, since older adults are more susceptible to pollution and they also greatly decrease non-farm employment, which presumably will no suffer any possible direct effects of pollution on productivity.
-The preponderance of results seems to indicate that effects on labor are the most likely mechanism.
-
-The final set of results explores whether there are impacts of pollution on aggregate economic development.
-Using nightlights, I show that lagged wind -- i.e.~pollution exposure in the prior year -- has a negative impact on economic development as proxied by nightlights.
-Going from the 25th percentile to the 75th percentile of wind exposure is synonymous with a decrease in the growth of nightlights within a village by around 0.7 percent.
-While this is a relatively small magnitude, it is worth noting that there are more than 100,000 villages in this sample.
-While the effect for a single village might be small, the overall, aggregate effect on growth is quite substantial.
-To put this number in perspective, Castelló-Climent, Chaudhary, and Mukhopadhyay (2018) argue that a standard-deviation increase in the share of adults with a university education leads to 12 percent higher nightlights in India, while Storeygard (2016) shows that a four-fold increase in oil prices over six years leads to seven percent lower nightlights at a distance of around 500km from the nearest large port in sub-Saharan Africa.
-In other words, the 0.7 percent change in nightlights is quite large given this is annual variation.
+To explore the mechanisms driving this effect, I also use data on labor allocation from the National Sample Survey (NSS).
+While there do not appear to be large changes in labor allocation patterns when looking at the season as a whole, there is a clear pattern when we break down exposure to pollution by month: higher levels of pollution at the beginning of the monsoon season (June and July) lead to lower levels of farm labor allocation, but no changes in non-farm labor allocation.
+This matches up well with a similar exercise with agricultural yields: pollution in June is the only month in which there is a negative effect of pollution on agricultural yield.
+In tandem, these two results point to effects on labor as one mechanism driving the total effects, though I cannot rule out direct effects of pollution on land productivity.
+
+The final set of results tries to disentangle the effects of pollution from the effects of coal pollution, specifically.
+Combining instrumental variables with a differences-in-differences strategy, I show that agricultural productivity has an elasticity of around -0.15 with respect to non-coal pollution.
+The opening of a coal plant, however, triples the elasticity, to more than -0.45, indicating that coal plant emissions are particularly harmful to agricultural productivity.
+This is consistent with coal emissions being quite different from other forms of pollution; for example, coal emissions contain much more mercury than other forms of pollution.
 
@@
 
-\begin{figure}
-\includegraphics{draft_files/figure-latex/plants-1} \caption[Coal plants in India from 1990 to 2010]{Coal plants in India from 1990 to 2010}\label{fig:plants}\floatfoot*{Note: The top figure shows the location of coal plants in 1990.
-The bottom figure shows the location of coal plants in 2010.}
-\end{figure}
-
-Figure \ref{fig:plants} displays the location of coal plants in India for two specific years: 1990 and 2010.
+Figure \ref{fig:plants} in the appendix displays the location of coal plants in India for two specific years: 1990 and 2010.
 There is a clear increase in the prevalence of coal plants across the country over the two decades.
@@
 I then calculate the direction from coal plants to all village centroids within that 30km radius.
-To define exposure, I then pull daily wind direction data from the National Center for Atmospheric Research.\footnote{climatedataguide.ucar.edu/} For each day, I document whether the wind is blowing towards each village centroid.\footnote{I define ``towards'' as within five degrees to help take into account that I am using village centroids, using the x and y components of wind speed.} I then temporally aggregate this daily data depending on the temporal definition of the corresponding outcome.
+To define exposure, I then pull daily wind direction data from the National Center for Atmospheric Research.\footnote{climatedataguide.ucar.edu/} For each day, I document whether the wind is blowing towards each village centroid.\footnote{I define ``towards'' as within five degrees to help take into account that I am using village centroids, using the x and y components of wind speed to calculate angles.} I then temporally aggregate this daily data depending on the temporal definition of the corresponding outcome.
 For example, agricultural productivity is defined across five months (e.g.~the monsoon season is from June to October) so I count the total days a given village is exposed to wind blowing from \emph{any} coal plant within 30km during those five months.
 
-\begin{figure}
-\includegraphics{draft_files/figure-latex/windexample-1} \caption[Wind direction and aggregation examples (2010-01-01)]{Wind direction and aggregation examples (2010-01-01)}\label{fig:windexample}\floatfoot*{Note: The top figure shows the average wind direction on January 1st, 2010, with zero degrees indicating north.
-The points are the location of coal plants on that date.
-The bottom figure shows the distribution of pollution exposure in a specific district -- Guna district in Madhya Pradesh -- on the same date.}
-\end{figure}
-
-The top panel in Figure \ref{fig:windexample} is an example wind direction raster.
+The top panel in Figure \ref{fig:windexample} in the appendix is an example wind direction raster.
 The raster shows wind direction for the entirety of India on January 1st, 2010, as well as the location of coal power plants on that date.
@@
 
-Finally, I use nightlights as a measure of economic growth (Henderson, Storeygard, and Weil 2012).
-I also take this data directly from SHRUG and Asher et al.
-(2021).
-This covers 1992 through 2013.
-Some of the older data may not be ideal (Gibson et al.
-2021), but this is the only way to document changes in economic growth at a more disaggregated level and at a regular basis.
-Although satellite imagery is an attractive alternative (Burke et al.
-2021), high-resolution satellite imagery does not go back to the early 1990s.
-
 \hypertarget{identification}{%
@@
 
-To calculate the effects of exposure to pollution on different outcomes, I rely on fixed effects and the differential exposure driven by plausibly exogenous wind direction, conditional on total rainfall and average temperature.
-The regression is of the form \begin{gather} y_{it} = \alpha_{i} + \gamma_{t} + \beta wind_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it}, \end{gather} where \(y_{it}\) is outcome \(y\) for unit \(i\) at time \(t\), \(\alpha\) is a geographic fixed effects, \(\gamma_t\) is time fixed effects, \(wind_{it}\) is exposure to pollution -- defined using wind direction -- \(rain_{it}\) is total rainfall during the season, and \(temp_{it}\) is average temperature during the season.
-Although I use \(i\) and \(t\) as subscripts for the fixed effects, note that these change depending on the unit of analysis and the outcome.
-I explicitly state the level of fixed effects in the tables and results below.
-
-Identification relies on changes in wind direction across time for the same geographic units.
-In other words, conditional on the fixed effects and weather, I assume changes in wind direction are as good as random and uncorrelated with the outcome -- e.g.~agricultural productivity -- except through exposure to pollution.
-One major threat to identification is that the building of a coal plant could lead to a spurious correlation between the exposure variable and the outcome.
-Specifically, since exposure is defined as zero prior to the building of a coal plant, the exposure variable itself might be endogenous even though wind direction is not.
-To help allay these concerns, I estimate additional regressions restricting the sample to only time periods after which a given location has a coal plant.
-I also present results using leads to show that future wind direction does not seem to predict current outcomes.
-
-The inclusion of rainfall and temperature is meant to reduce concerns that wind direction may not affect agricultural productivity but, instead, be correlated with it through changes in weather.
-For example, prevailing winds from one direction may be associated with dry weather, while winds from another direction may be more likely to carry moisture and rain.
-It is not a priori clear how this would bias the estimates with respect to pollution and coal plants, however.
-The differences in weather patterns is very context specific and can differ even for villages located near one another.
-Nonetheless, adding rainfall and temperature should alleviate some of these concerns.
-
-A large question looms with respect to identification: what are the characteristics of areas near the construction of new coal plants?
-While I examine the identifying assumptions in a panel context below, I begin with this static question here.
+I am primarily interested in the effects of pollution on agricultural productivity. We might consider a regression of the following form:
+
+\begin{gather} y_{it} = \alpha_{i} + \gamma_{t} + \beta PM_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it}, \end{gather}
+
+where \(y_{it}\) is outcome \(y\) for unit \(i\) at time \(t\), \(\alpha\) is a geographic fixed effects, \(\gamma_t\) is time fixed effects, \(PM_{it}\) is the concentration of particulate matter 2.5, \(rain_{it}\) is total rainfall during the season, and \(temp_{it}\) is average temperature during the season.
+
+However, even with the fixed effects, this regression likely suffers from endogeneity; places with more pollution may be very different from places with less pollution.
+It is difficult to sign the direction of the bias a priori, because it depends on how the coal plants are placed.
+For example, if coal plants are placed in places in need of power -- for example, due to rapid growth -- then the estimates might be upwards biased.
+On the other hand, if official purposefully place coal plants in poorer areas for political reasons, then the estimates might be downwards biased.
+
+So what are the characteristics of areas near the construction of new coal plants?
 Table \ref{tab:plantresultstable} presents four separate regressions.
@@
 \midrule
-pop (log) & 0.068*** & 0.000 & 0.009*** & 0.002***\\
+pop (log) & 0.010*** & 0.001*** & 0.017*** & 0.004***\\
  & (0.000) & (0.000) & (0.000) & (0.000)\\
-literacy (prop) & 0.127*** & -0.005*** & 0.044*** & 0.030***\\
- & (0.002) & (0.001) & (0.002) & (0.002)\\
-has paved road & -0.037*** & 0.003*** & -0.008*** & -0.000\\
- & (0.001) & (0.000) & (0.001) & (0.001)\\
-ag productivity &  &  & 0.033*** & 0.005***\\
+area (log) & -0.006*** & 0.001** & -0.015*** & -0.001***\\
+ & (0.001) & (0.000) & (0.000) & (0.000)\\
+literacy (prop) & 0.024*** & 0.017*** & 0.032*** & 0.029***\\
+ & (0.003) & (0.002) & (0.002) & (0.002)\\
+ag productivity &  &  & 0.026*** & 0.004***\\
  &  &  & (0.001) & (0.000)\\
@@
 \midrule
-observations & 490,882 & 447,288 & 516,798 & 476,517\\
+observations & 283,021 & 264,265 & 508,784 & 469,004\\
 \bottomrule
@@
 
+Given the likely endogeneity, I instead turn to a plausibly exogenous source of variation in pollution: wind direction.
+While the location of a coal plant might be endogenous, the amount of pollution received by a specific village in a given year can vary randomly based on wind direction.
+By counting the number of days in a given year that a village is downwind from a polluted area and conditioning on village fixed effects and weather (rainfall and temperature), variation across years is plausibly exogenous.
+
+Consider the reduced form effect of wind exposure on agricultural productivity:
+\begin{gather} y_{it} = \alpha_{i} + \gamma_{t} + \beta wind_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it}, \end{gather}
+
+where variables are defined identically to above but with wind as the primary variable of interest instead of pollution.
+Combining these into a two-stage procedure, we can instrument for pollution using wind direction; small changes in wind direction can lead to large changes in pollution exposure for villages located near point sources of pollution, allowing me to identify the causal effects of pollution by isolating the random component driven by wind direction.
+This is similar in spirit to Deryugina et al.
+(2019).
+
+Identification relies on changes in wind direction across time for the same geographic units.
+In other words, conditional on the fixed effects and weather, I assume changes in wind direction are as good as random and uncorrelated with the outcome -- e.g.~agricultural productivity -- except through exposure to pollution.
+One major threat to identification is that the building of a coal plant could lead to a spurious correlation between the exposure variable and the outcome.
+Specifically, since exposure is defined as zero prior to the building of a coal plant, the exposure variable itself might be endogenous even though wind direction is not.
+To help allay these concerns, I estimate additional regressions restricting the sample to only time periods after which a given location has a coal plant.
+I also present results using leads to show that future wind direction does not seem to predict current outcomes.
+
+The inclusion of rainfall and temperature is meant to reduce concerns that wind direction may not affect agricultural productivity but, instead, be correlated with it through changes in weather.
+For example, prevailing winds from one direction may be associated with dry weather, while winds from another direction may be more likely to carry moisture and rain.
+It is not a priori clear how this would bias the estimates with respect to pollution and coal plants, however.
+The differences in weather patterns is very context specific and can differ even for villages located near one another.
+Nonetheless, adding rainfall and temperature should alleviate some of these concerns.
+
 It is worth taking the time to validate the use of wind direction as a measure of exposure to pollution.
-Consider the data used for the regressions presented in Table \ref{tab:pollutiontable}.
+Consider the data used for the regressions presented in Table \ref{tab:pollutiontable} of the appendix.
 The outcome variable is particulate matter -- specifically, \(\mathrm{PM_{2.5}}\), which is particulate matter no larger than 2.5 micrometers in diameter -- which comes from Hammer et al.
@@
 \(\mathrm{PM_{2.5}}\) is one of the harmful byproducts from coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.\footnote{www.epa.gov/airmarkets/power-plants-and-neighboring-communities} This data is available at the monthly level, so in Table \ref{tab:pollutiontable}, I aggregate wind exposure to the monthly level, as well.
-
-\begin{table}
-
-\caption{\label{tab:pollutiontable}Wind direction and particulate matter}
-\centering
-\begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{1998-2015} & \multicolumn{2}{c}{2002-2013} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
-  & (1) & (2) & (3) & (4)\\
-\midrule
-wind (10s) & 0.002*** & 0.002*** & 0.006*** & 0.001***\\
- & (0.000) & (0.000) & (0.001) & (0.000)\\
-\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-village & Yes & Yes & Yes & Yes\\
-month & Yes & No & Yes & No\\
-district-month & No & Yes & No & Yes\\
-\midrule
-observations & 24,376,458 & 24,376,458 & 16,250,973 & 16,250,973\\
-\bottomrule
-\end{tabular}
-\begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
 
@@
 
+\hypertarget{identifying-effects-of-pollution-from-coal-plants}{%
+\subsubsection{Identifying effects of pollution from coal plants}\label{identifying-effects-of-pollution-from-coal-plants}}
+
+I use the location of coal plants to identify locations of high pollution, but there can be other pollution sources located near coal plants, as well.
+I nonetheless want to identify the effects of pollution coming specifically from coal plants.
+I do this by combining the two-stage procedure above with a differences-in-differences framework.
+Specifically, I create a new variable identifying the exact year the coal plant opens in a given area and calculate the change in the effects of pollution on agricultural productivity.
+
+Consider the following regression of interest:
+\begin{gather} 
+y_{it} = \alpha_{i} + \gamma_{t} + \omega pm_{it} + \phi open_{it} + \beta pm_{it}\times open_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it},
+\end{gather}
+
+where the interaction term \(pm_{it}\times open_{it}\) indicates the change in the effect of pollution following the opening of the coal plant.
+Note that this is not the same as saying that pollution increased; instead, it is focusing on the change in the effect of a change in pollution of one unit on the outcome of interest, agricultural yield.
+I instrument for both \(pm_{it}\) and \(pm_{it}\times open_{it}\) using wind and the interaction between wind and the open indicator, or \(wind_{it}\times open_{it}\).
+
 \hypertarget{results}{%
@@
 
-I first present results for agricultural productivity.
+I first present results for agricultural productivity, with the reduced form.
 The outcome of all the regressions in this section is the log of agricultural land productivity, which is defined as tons per hectare.
 The first set of results are in Table \ref{tab:yieldtable}.
-The unit of analysis is the vilalge-year-season -- from 2002 to 2013 -- the geographic fixed effect is the village-season, and the temporal fixed effect is the year.
-The exposure variable is defined for the entire season.
-Since Gangopadhyay et al.
-(2022) define both the monsoon (kharif) season and the winter (rabi) season as five months long, I do as well, using the same five months as the authors.
-This implies that the exposure variable ranges from zero to slightly more than 150.
+The unit of analysis is the village-year -- from 2002 to 2013 -- focusing only on the monsoon (kharif) season.
+The data used to construct the yield measure is calculated using five months of data, so the wind variable is defined for the same five months and is the sum of days in which the wind is pointed towards the village centroid from a coal plant.
 
 The first column presents the most simple results, with only the wind exposure variable and the fixed effects.
-The second column adds a rainfall -- defined as the total rainfall in the season -- and average temperature variable, the addition of which is motivated by the fact that changes in wind direction could be correlated with changes in weather.
-However, the coefficient on the wind exposure variable is completely unchanged by the addition of the two weather variables, indicating that weather may not be as worrisome of a confounder as thought.
-The direction of the weather coefficients is also consistent with expectations; higher rainfall is good for agricultural productivity while higher temperature is not.
-
-Column three adds district-by-year-by-season fixed effects as an additional robustness checks.
-This addition decreases the magnitude of the exposure variable but it remains negative and significant.
-However, part of this may be due to the fact that the new fixed effects soak up much of the variation in the exposure variable, as shown in Figure \ref{fig:windexample}.
-The last two columns separate the effects into the monsoon season and the winter season, but the overall effect is roughly similar, though rainfall has a larger effect in the monsoon season than the winter season, which is unsurprising given the winter season is driven by irrigation in many areas.
+The second column adds a rainfall and average temperature variable, while the third column includes higher-order weather variables, as well.
+While the coefficient decreases slightly in magnitude as we move from the first to the second column and the second to the third column, the decrease is relatively small.
+There is a consistent, negative effect of wind exposure on agricultural productivity.
 
@@
 \begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}}
 \toprule
-\multicolumn{1}{c}{ } & \multicolumn{3}{c}{all} & \multicolumn{1}{c}{monsoon} & \multicolumn{1}{c}{winter} \\
-\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6}
-  & (1) & (2) & (3) & (4) & (5)\\
-\midrule
-wind & -0.003*** & -0.003*** & -0.001*** & -0.002*** & -0.004***\\
- & (0.000) & (0.000) & (0.000) & (0.000) & (0.000)\\
-rain (z) &  & 0.029*** & 0.008*** & 0.075*** & 0.006***\\
- &  & (0.000) & (0.001) & (0.002) & (0.000)\\
-mean temp (10s) &  & -0.000*** & -0.007*** & -0.002*** & -0.008***\\
- &  & (0.000) & (0.000) & (0.000) & (0.000)\\
-\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-village-season & Yes & Yes & Yes & No & No\\
-year & Yes & Yes & No & Yes & Yes\\
-district-year-season & No & No & Yes & No & No\\
-village & No & No & No & Yes & Yes\\
-\midrule
-observations & 2,387,417 & 2,371,221 & 2,371,221 & 1,256,951 & 1,114,270\\
+  & (1) & (2) & (3)\\
+\midrule
+wind days & -0.0078*** & -0.0069*** & -0.0062***\\
+ & (0.0007) & (0.0007) & (0.0006)\\
+rain (z) &  & 0.0854*** & -0.0424***\\
+ &  & (0.0018) & (0.0046)\\
+mean temp (10s) &  & -0.0021*** & 0.1321***\\
+ &  & (0.0001) & (0.0034)\\
+rain square &  &  & -0.0010***\\
+ &  &  & (0.0001)\\
+temp square &  &  & -0.0002***\\
+ &  &  & \vphantom{1} (0.0000)\\
+rain times temp &  &  & 0.0003***\\
+ &  &  & (0.0000)\\
+\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{}\\
+village & Yes & Yes & Yes\\
+year & Yes & Yes & Yes\\
+\midrule
+observations & 1,266,922 & 1,266,922 & 1,266,922\\
 \bottomrule
@@
 \begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}}
 \toprule
-\multicolumn{1}{c}{ } & \multicolumn{3}{c}{all} & \multicolumn{1}{c}{monsoon} & \multicolumn{1}{c}{winter} \\
-\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6}
-  & (1) & (2) & (3) & (4) & (5)\\
-\midrule
-particulate matter & 0.0000 & 0.0329*** & -0.0787*** & -0.0509*** & 0.0216***\\
-(log PM 2.5) & (0.0000) & (0.0028) & (0.0040) & (0.0033) & (0.0052)\\
-rain (z) &  & 0.0290*** & 0.0079*** & 0.0760*** & 0.0064***\\
- &  & (0.0004) & (0.0013) & (0.0017) & (0.0005)\\
-mean temp (10s) &  & -0.0003*** & -0.0071*** & -0.0020*** & -0.0080***\\
- &  & (0.0001) & (0.0004) & (0.0001) & (0.0001)\\
-\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-village-season & Yes & Yes & Yes & No & No\\
-year & Yes & Yes & No & Yes & Yes\\
-district-year-season & No & No & Yes & No & No\\
-village & No & No & No & Yes & Yes\\
-\midrule
-observations & 2,387,417 & 2,371,221 & 2,371,221 & 1,256,951 & 1,114,270\\
+  & (1) & (2) & (3)\\
+\midrule
+particulate matter & -0.042*** & -0.044*** & -0.032***\\
+(log PM 2.5) & (0.003) & (0.003) & (0.003)\\
+weather & No & Yes & Yes\\
+weather (expanded) & No & No & Yes\\
+\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{}\\
+village & Yes & Yes & Yes\\
+year & Yes & Yes & Yes\\
+\midrule
+observations & 1,266,922 & 1,266,922 & 1,266,922\\
 \bottomrule
@@
 Instead, we are actually interested in better understanding the effects of exposure to pollution on agricultural productivity.
-Table \ref{tab:yieldtablepm} presents the same regressions as Table \ref{fig:yieldtable}, with a key difference: instead of wind exposure, it uses average concentration of \(\mathrm{PM_{2.5}}\) -- from Hammer et al.
+Table \ref{tab:yieldtablepm} presents the same regressions as Table \ref{fig:yieldtable}, with a key difference: instead of wind exposure, it uses logged average concentration of \(\mathrm{PM_{2.5}}\) -- from Hammer et al.
 (2020) -- during the season as the key explanatory variable.
-
-Several interesting patterns emerge.
-We only observe a negative relationship between particulate matter and agricultural productivity in column three; the coefficient in column two is actually positive.
-Breaking out the effects into the monsoon and winter seasons, we also see a positive relationship between pollution and agricultural productivity.
+There is again a consistent, negative relationship between particulate matter and agricultural productivity.
+Since both variables are logged, the coefficients are similar to elasticities, indicating that a one-percent increase in pollution leads to a 0.03-to-0.04 decrease in agricultural productivity.
 
@@
 To recover the actual effect of pollution, we need an instrument.
-If, conditional on the fixed effects and weather variables, wind direction is exogenous and only affects agricultural productivity through pollution -- which seems reasonable since it is hard to imagine what other variables it might affect -- then it is a valid instrument for pollution.
-
-I present the IV results in Table \ref{tab:yieldtabletwo}, with the second-stage results in the top panel and the first-stage results in the bottom panel.
-The first-stage results make clear how strong of an instrument wind is; the t-statistic in column one, for example, is almost 48 while it is still well above 10 in column three, which translates to a first-stage F-statistic of 195, meeting even the higher bar for the first-stage F-test argued for by D.
+If, conditional on the fixed effects and weather variables, wind direction is exogenous and only affects agricultural productivity through pollution, then it is a valid instrument for pollution.
+
+I present the IV results in Table \ref{tab:yieldtabletwo}, with the first-stage F-statistic in the bottom.
+The first-stage results make clear how strong of an instrument wind is; all three F-statistics are above 1,000, meeting even the higher bar for the first-stage F-test argued for by D.
 S.
@@
 The second-stage results first indicate that \(\mathrm{PM_{2.5}}\) is a very strong predictor of agricultural productivity.
-The smallest coefficient -- that for the monsoon season in column (4) -- and the mean absolute deviation for \(\mathrm{PM_{2.5}}\) point to a change in agricultural productivity of around 23 percent, which is much higher than the wind results documented above.
-This makes sense, of course, since the IV is essentially scaling up the effects of wind.
+Using the results in column three, a one percent increase in particulate matter leads to a 0.73 percent decrease in agricultural productivity.
+To put this in further perspective, the mean particulate matter concentration is just short of 40 (39.68) in the monsoon sample, while the within-village absolute deviation in particulate matter is 3.98, meaning year-to-year swings of 7-to-14 percent in agricultural productivity due to pollution exposure are common.
 
@@
 \begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
 \toprule
-\multicolumn{1}{c}{ } & \multicolumn{3}{c}{all} & \multicolumn{1}{c}{monsoon} & \multicolumn{1}{c}{winter} \\
-\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6}
-  & (1) & (2) & (3) & (4) & (5)\\
-\midrule
-particulate matter & -1.190*** & -1.027*** & -1.671*** & -0.525*** & -1.741***\\
-(log PM 2.5) & (0.070) & (0.071) & (0.235) & (0.043) & (0.139)\\
-rain (z) &  & 0.028*** & 0.012*** & 0.080*** & 0.001**\\
- &  & (0.000) & (0.001) & (0.002) & (0.001)\\
-mean temp (10s) &  & -0.001*** & -0.009*** & -0.002*** & -0.007***\\
- &  & (0.000) & (0.000) & (0.000) & \vphantom{2} (0.000)\\
-\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-village-season & Yes & Yes & Yes & No & No\\
-year & Yes & Yes & No & Yes & Yes\\
-district-year-season & No & No & Yes & No & No\\
-village & No & No & No & Yes & Yes\\
-\midrule
-observations & 2,387,417 & 2,371,221 & 2,371,221 & 1,256,951 & 1,114,270\\
-\midrule
-\textbf{first stage:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-wind & 0.003*** & 0.003*** & 0.000*** & 0.004*** & 0.002***\\
- & (0.000) & (0.000) & (0.000) & (0.000) & (0.000)\\
-rain (z) &  & -0.001*** & 0.002*** & 0.009*** & -0.003***\\
- &  & (0.000) & (0.000) & (0.000) & \vphantom{1} (0.000)\\
-mean temp (10s) &  & -0.000*** & -0.001*** & 0.001*** & 0.001***\\
- &  & (0.000) & (0.000) & (0.000) & (0.000)\\
+  & (1) & (2) & (3)\\
+\midrule
+particulate matter & -0.941*** & -0.813*** & -0.728***\\
+(log PM 2.5) & (0.078) & (0.076) & (0.075)\\
+weather & No & Yes & Yes\\
+weather (expanded) & No & No & Yes\\
+\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{}\\
+village & Yes & Yes & Yes\\
+year & Yes & Yes & Yes\\
+\midrule
+F (first stage) & 1,235 & 1,262 & 1,263\\
+observations & 1,266,922 & 1,266,922 & 1,266,922\\
 \bottomrule
@@
 
-Columns two and three split the sample based on median yield at the very beginning of the sample (the first year available, 2002-2003).
+Columns three and four split the sample based on median yield at the very beginning of the sample (the first year available, 2002-2003).
 Column three presents results for villages above the median while column four presents results for villages below the median.
-The largest effects are apparently concentrated in the lower half of the distribution.
-Perhaps more productive areas are more likely to have ways to combat the effects of shocks -- e.g.~by adjusting the use of other inputs -- or maybe less productive areas are inherently more susceptible to the effects of pollution, perhaps due to direct effects on land productivity.
+The largest effects are apparently concentrated in the upper half of the distribution, meaning the most productive areas are also most susceptible to the effects of pollution.
 
 Given ongoing changes in the climate, a key question is how welfare will change as shocks become more and more common.
-In the last column looks at possible multiplicative effects of shocks by interacting pollution with rainfall and seeing whether the effects compound one another.
+The last column looks at possible multiplicative effects of shocks by interacting pollution with rainfall and seeing whether the effects compound one another.
 Since there are now two endogenous regressors, I instrument for both using wind exposure and an interaction between wind exposure and rainfall.
@@
 \toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{exposure} & \multicolumn{2}{c}{yield} & \multicolumn{1}{c}{ } \\
+\multicolumn{1}{c}{ } & \multicolumn{2}{c}{wind} & \multicolumn{2}{c}{yield} & \multicolumn{1}{c}{ } \\
 \cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
@@
 \midrule
-PM 2.5 & -1.082*** & -2.523*** & -0.439*** & -1.711*** & -0.809***\\
- & (0.073) & (0.624) & (0.046) & (0.146) & (0.060)\\
-PM time rain &  &  &  &  & 0.180***\\
- &  &  &  &  & (0.008)\\
+PM 2.5 & -0.767*** & -0.328*** & -1.130*** & -0.690*** & -0.409**\\
+ & (0.104) & (0.095) & (0.060) & (0.146) & (0.160)\\
+PM times rain &  &  &  &  & 0.061***\\
+ &  &  &  &  & (0.014)\\
 weather & Yes & Yes & Yes & Yes & Yes\\
 \textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-village-season & Yes & Yes & Yes & Yes & Yes\\
+village & Yes & Yes & Yes & Yes & Yes\\
 year & Yes & Yes & Yes & Yes & Yes\\
 \midrule
-F (1st stage, PM) & 6,488 & 60 & 2,867 & 2,150 & 2,958\\
-F (1st stage, PM times rain) & -- & -- & -- & -- & 1,697\\
-observations & 1,051,010 & 1,320,211 & 1,193,764 & 1,177,457 & 2,371,221\\
+F (1st stage, PM) & 684 & 491 & 688 & 484 & 865\\
+F (1st stage, PM times rain) &  &  &  &  & 1,437\\
+observations & 617,804 & 649,118 & 634,342 & 632,580 & 1,266,922\\
 \bottomrule
@@
 
-\hypertarget{robustness-checks}{%
-\subsection{Robustness checks}\label{robustness-checks}}
-
-I present two separate robustness checks in the appendix.
-The first set of analyses attempts to separate the possible endogenous location of coal plants from outcomes.
-Specifically, prior to the opening of a new coal plant in an area, all villages receive a zero for wind.
-This means that, even if wind direction is exogenous, the opening of the coal plant could lead to correlations between the error term and the wind variable.
-
-Although the results with district-year-season help mitigate this concern, Table \ref{tab:yieldtablepostplant} in the appendix checks this possibility in a different way.
-Specifically, I restrict the sample to only village-years \emph{after} the introduction of a coal plant.
-This drops any pre-plant-opening observations -- all of which are mechanically zeros for wind -- to prevent endogeneity due to the opening.
-The coefficients in Table \ref{tab:yieldtablepostplant} show that the overall effects are essentially unchanged across any specifications.
-Comparing changes in coefficients can also help reduce any effects driven by other polluting industries located near coal plants; since the pre-opening exposure variable is zero, the coefficients in Table \ref{tab:yieldtablepostplant} would likely increase in magnitude if the effects were driven by non-coal pollution.
-
-The second robustness check includes leads and lags for wind to see if they predict current year productivity.
-I present these results in \ref{tab:yieldtableleads}.
-The first column is the reduced form effect and shows that only current wind predicts productivity; leads and lags have no impact.
-We see similar results in column two -- the IV results -- with a negative coefficient for current pollution, only.
-However, the standard errors are much larger than the main results -- around five times as large -- making it hard to come to any conclusions.
+In the appendix, I present in Table \ref{tab:yieldtableleads} as a robustness check similar regressions but also including wind/pollution leads.
+The negative effect is only seen with current pollution, not future pollution.
 
@@
 Figure \ref{fig:laborplot} in the appendix shows this temporal distribution across the five waves of the survey.
-Since I can no longer include district-by-month fixed effects, I instead opt for district-level trends to help take into account aggregate temporal changes in outcomes.
+Since pollution is not disaggregated at a daily level, I instead stick to the reduced form results with wind.
 
@@
 
-\caption{\label{tab:labortablerural}Wind direction and labor allocation by rural status}
+\caption{\label{tab:labortable}Wind direction and labor allocation by rural status}
 \centering
@@
 \midrule
-\textbf{rural} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-wind & -0.112*** & -0.118*** & -0.050* & -0.068** & -0.002 & -0.116***\\
- & (0.036) & (0.029) & (0.030) & (0.031) & (0.028) & (0.034)\\
-DV mean & 3.499 & 3.499 & 2.172 & 1.327 & 1.260 & 2.240\\
-observations & 529,562 & 529,466 & 529,466 & 529,466 & 529,466 & 529,466\\
-\midrule
-\textbf{urban} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-wind & -0.020 & -0.030 & -0.011 & -0.019 & 0.023* & -0.053\\
- & (0.033) & (0.041) & (0.039) & (0.031) & (0.014) & (0.048)\\
-DV mean & 3.040 & 3.040 & 1.393 & 1.647 & 0.101 & 2.939\\
-observations & 369,483 & 369,390 & 369,390 & 369,390 & 369,390 & 369,390\\
+\textbf{panel a: all} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+wind & -0.011 & -0.009 & 0.030 & -0.049 & -0.009 & -0.010\\
+ & (0.031) & (0.034) & (0.039) & (0.034) & (0.034) & (0.033)\\
+DV mean & 3.303 & 3.303 & 1.862 & 1.442 & 0.789 & 2.514\\
+observations & 523,075 & 522,957 & 522,957 & 522,957 & 522,957 & 522,957\\
+\midrule
+\textbf{panel b: rural} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+wind & -0.006 & -0.034 & 0.068 & -0.102*** & -0.024 & -0.010\\
+ & (0.044) & (0.040) & (0.043) & (0.035) & (0.043) & (0.036)\\
+DV mean & 3.494 & 3.494 & 2.184 & 1.311 & 1.271 & 2.223\\
+observations & 307,381 & 307,325 & 307,325 & 307,325 & 307,325 & 307,325\\
 \midrule
@@
 
-There are some interesting results in Table \ref{tab:labortable}.
-The first two columns show a marginally significant decrease in overall labor allocation (where ``labor'' is defined as non-domestic self and wage employment).
-The third and fourth columns show that this overall decrease is predominantly from wage employment, not self employment.
-An additional day of wind over the last seven decreases wage employment by 0.034 days in the district.
-Mean wage employment is 1.458 days, meaning one day of wind decreases wage employment by 2.3 percent of the mean.
-Since much of the district is not covered by ``treated'' villages, this is actually a rather substantive effect on the villages involved.
-
-The last two columns split the effects by sector, into farm and non-farm employment.
-Although agricultural (land) productivity decreases when wind blows in the direction of villages, overall labor allocation to farm labor actually goes \emph{up}, by around 4.5 percent of the mean of 0.784 days.
-Non-farm labor, on the other hand, goes down, by around 2.4 percent of the mean of 2.527 days.
-
-There are two possible explanations for this pattern of results.
-First, it could be that land productivity goes down, leading farmers to increase their farm labor allocation in an attempt to compensate for the decreased productivity.
-A second possibility is that pollution does not affect land productivity at all, at least not directly.
-In this scenario, the worsening air quality could lead to decreased labor productivity, meaning that ``effective labor'' essentially decreases when pollution is higher.
-
-Table \ref{tab:labortablemonsoon} and Table \ref{tab:labortablewinter} in the appendix disaggregate these effects into the monsoon and winter seasons.
-Apparently, the entirety of the effect is driven by changes in labor allocation during the monsoon season.
-The dynamics across seasons apparently lead to large differences in the effects of pollution on labor allocation.
-
-One way to provide suggestive evidence of mechanisms is to explore heterogeneity in the effects.
-Previous research has shown larger effects of pollution on the health of the oldest (Deryugina et al.
-2019).
-The labor sample includes all adults between 15 and 60 years of age.
-To look at heterogeneity by age, I present an additional set of results interacting the wind variable with an indicator for whether an individual is over 37 years of age.\footnote{This splits the sample into two equal bins of age: those who are 15-37 and those who are 38-60.} These results are in Table \ref{tab:labortableold}, which otherwise presents the same specifications in the last five columns as \ref{tab:labortable}.
-
-Table \ref{tab:labortableold} presents a very clear set of results: the effects are always larger (more negative) for older adults.
-Particularly interestingly, we now see that the mean-zero effect on self employment in the previous table is driven by two stark contrasts: a large increase for the younger group and a large decrease for the older group.
-For the youngest group, this increase appears in farm labor, indicating likely increases in farm self-employment.
-The oldest, on the other hand, have large decreases in both farm and non-farm labor, with larger decreases in the latter.
+The top panel presents results for all respondents during the five-month monsoon season while the bottom panel presents results for respondents in rural areas.
+Interestingly, we do not see any signs of large, negative effects of wind exposure on labor allocation across any sectors, with the possible exception of wage employment in rural areas.
+However, the coefficient for self-employment in rural areas is of the same magnitude and \emph{positive}, but imprecisely estimated, making it hard to come to any definitive conclusions, at least in the aggregate.
+
+\hypertarget{month-specific-exposure}{%
+\subsection{Month-specific exposure}\label{month-specific-exposure}}
+
+While overall results for labor allocation are inconclusive, perhaps we can learn more by looking at temporally disaggregated results, focusing on May through December.
+Although the yield data was created by looking at five specific months of the year -- June through October -- this is partially because this is the time of year when remote-sensing data is most able to pick up overall changes in crop cover.
+Preparation, however, starts before that, with operations such as ploughing often taking place prior to June.
+
+Since the NSS data has survey dates and the both the pollution and wind data can be aggregated to the month, we can look at how exposure during different months of the agricultural season (differentially) affect labor allocation and overall agricultural productivity.
+I begin with labor allocation in Table \ref{tab:labor3month}.
+As before, the exposure measure is just the wind variable (over the last seven days), since I can directly match daily wind direction with the timing of the survey, but I cannot do that for pollution, which is monthly.
+The first column presents results for farm labor and the second column presents results for non-farm labor.
+
+Higher exposure leads to significantly less farm labor labor in June and July, but there are no discernible effects in other months of the year.
+There are likewise no discernible effects on non-farm labor, but some of the coefficients are rather imprecisely estimate.
+To put the effects for June and July in context and taking the coefficients at face value, an additional day of wind in the last seven days decreases farm labor in June by around 11.2 percent of the mean and in July by around 17.1 percent of the mean.
 
@@
 
-\caption{\label{tab:labortableruralage}Wind direction and labor allocation by age (rural only)}
+\caption{\label{tab:labor3month}Wind and labor allocation, by month}
 \centering
 \begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
 \toprule
-  & all & all & self & wage & farm & non-farm\\
-\midrule
-\textbf{younger than 35} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-wind & -0.077* & -0.085** & -0.061* & -0.024 & -0.001 & -0.085**\\
- & (0.046) & (0.034) & (0.032) & (0.032) & (0.034) & (0.038)\\
-DV mean & 3.106 & 3.106 & 1.794 & 1.312 & 1.061 & 2.044\\
-observations & 263,965 & 263,923 & 263,923 & 263,923 & 263,923 & 263,923\\
-\midrule
-\textbf{45 or older} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-wind & -0.131** & -0.165*** & -0.054 & -0.111*** & -0.070* & -0.095**\\
- & (0.054) & (0.043) & (0.054) & (0.041) & (0.036) & (0.042)\\
-DV mean & 3.584 & 3.584 & 2.481 & 1.103 & 1.444 & 2.140\\
-observations & 166,184 & 166,155 & 166,155 & 166,155 & 166,155 & 166,155\\
-\midrule
-controls & No & Yes & Yes & Yes & Yes & Yes\\
-\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-district & Yes & Yes & Yes & Yes & Yes & Yes\\
-state-month & Yes & Yes & Yes & Yes & Yes & Yes\\
+  & farm & non-farm\\
+\midrule
+May & -0.006 & -0.039\\
+ & (0.068) & (0.070)\\
+June & -0.136* & 0.114\\
+ & (0.080) & (0.099)\\
+July & -0.285** & 0.176\\
+ & (0.142) & (0.134)\\
+Aug. & 0.048 & -0.010\\
+ & (0.086) & (0.069)\\
+Sept. & 0.076 & -0.082\\
+ & (0.085) & (0.091)\\
+Oct. & 0.013 & -0.083\\
+ & (0.076) & (0.059)\\
+Nov. & -0.063 & -0.013\\
+ & (0.062) & (0.075)\\
+controls & Yes & Yes\\
+\textbf{fixed effects:} & \textbf{} & \textbf{}\\
+district & Yes & Yes\\
+year-month & Yes & Yes\\
+\midrule
+observations & 307,325 & 307,325\\
 \bottomrule
@@
 \begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the district level.
-Control variables include female, age, age squared, and (years of) education.
-The outcome in each regression is days worked in the previous seven days.
+\item Note: Standard errors are in parentheses and are clustered at the village level. The outcome in column one is .  * p<0.10 ** p<0.05 *** p<0.01
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+
+I perform a similar exercise with yield, the results of which are in Table \ref{tab:yield4monthly}.
+I include only the five months of the year that are used in the calculation of the yield estimates, or June through October.
+The results are again from two-stage least squares, with the F-statistics for each month presented at the bottom; monthly wind remains a strong predictor of pollution.
+Column one presents results without controls for weather while column two presents results with controls for weather; the differences in coefficients make clear how important the inclusion of weather is here.
+
+Focusing on column two, there is only one month where we see a clear effect of pollution on yield: June.
+Given that we also see negative effects of pollution on farm labor allocation in June, this is suggestive -- but not conclusive -- evidence that pollution works through labor.
+Note that this does not rule out pollution also having direct effects on agricultural productivity, however.
+
+\begin{table}
+
+\caption{\label{tab:yield4monthly}Pollution and productivity, by month}
+\centering
+\begin{threeparttable}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\toprule
+  & (1) & (2)\\
+\midrule
+June & -0.325** & -0.736***\\
+ & (0.135) & (0.062)\\
+July & -0.778*** & 0.051\\
+ & (0.151) & (0.095)\\
+Aug. & -1.201*** & 0.087\\
+ & (0.455) & (0.070)\\
+Sept. & 1.270*** & -0.140\\
+ & (0.427) & (0.180)\\
+Oct. & -3.013*** & 0.334\\
+ & (0.963) & (0.291)\\
+weather & No & Yes\\
+\textbf{fixed effects:} & \textbf{} & \textbf{}\\
+village & Yes & Yes\\
+year & Yes & Yes\\
+\midrule
+F (June) & 1,063 & 972\\
+F (July) & 2,090 & 2,200\\
+F (Aug.) & 1,646 & 2,340\\
+F (Sept.) & 2,025 & 2,024\\
+F (Oct.) & 473 & 402\\
+observations & 1,266,922 & 1,266,922\\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}[para]
+\item Note: Standard errors are in parentheses and are clustered at the village level.
+The outcome in all regressions is estimated agricultural yield, defined as tons per hectare.
 * p<0.10 ** p<0.05 *** p<0.01
@@
 
-To put some of these changes in context, an additional day of exposure to pollution from wind increases self-employment for the younger half of the sample by around six percent of the mean and increases farm employment for the younger half by almost 10 percent of the mean.
-For the older half, on the other hand, an additional day of exposure decreases self-employment and non-farm employment by around five percent of their means.
-
-While speculative, one possible explanation for this pattern is a decrease in labor productivity.
-While the youngest may be able to readjust and increase their overall self-employment labor allocation, perhaps the oldest cannot due to, for example, health effects of pollution, which are well documented (Bentayeb et al.
-2012; Pope III 2000).\footnote{Much of the literature also suggests the youngest are at significant risk of mortality and morbidities.
-However, given my focus on labor allocation, I only discuss older adults here.} We see decreases in overall agricultural productivity, despite an increase in overall farm labor.
-We also see decreases in non-farm employment, indicating that individuals are reallocating labor from the non-farm sector to the farm sector.
-
-It is worth noting that the larger negative effects for older adults is more consistent with a labor story than a land story; it is hard to imagine these patterns emerging if the only effects were through land productivity changes.
-The overall decrease in non-farm employment is more compatible with a labor health/productivity story, though with general equilibrium effects in agriculture spillover over into the non-farm sector, this need not necessarily be the case.
-
-In addition, Table \ref{tab:labortableoldstate} in the appendix presents one additional set of results to support a labor story.
-Changes in land productivity should be relatively fixed in a short period of time.
-For example, if we focus on just a single month in a single district, the overall effects driven by changes in land productivity should be constant even though pollution may have differential effects throughout the month on labor productivity, which responds more immediately to changes in ar quality.
-Table \ref{tab:labortableoldstate} changes the specification to include district-by-year-by-month fixed effects.
-In other words, the results compare people in the same district in the same month, but who may have differential exposure to pollution simply because they were interviewed at a slightly different time.
-
-The overall pattern of results stays the same when we shift the focus.
-Overall labor allocation increases for the youngest, but decreases for the oldest, and this is driven completely by self employment.
-There are decreases in both farm and non-farm employment for the oldest, while the youngest show increases in non-farm employment, but not farm employment -- which is a change from the previous results.
-The overall story remains the same, however, which is further evidence of effects being driven primarily through changes in labor allocation and productivity.
-
-Finally, Table \ref{tab:wagestable} in the appendix shows that there are no significant changes in daily wages, though the overall direction of wage changes would be ambiguous, anyway.
-While reduced productivity might lead to lower wages, decreased labor supply could also put upward pressure on wages.
-
-\hypertarget{month-specific-exposure}{%
-\subsection{Month-specific exposure}\label{month-specific-exposure}}
-
-Pollution leads to lower agricultural productivity as well as changes in labor allocation.
-This leads to an obvious question: might higher levels of pollution also directly affect economic growth?
-To explore this question, I look at changes in nightlights relative to changes in pollution.
-It seems reasonable to think that pollution might affect nightlights with a lag, so I include the lag of pollution rather than contemporaneous pollution.
+A final caveat about these month-specific results is that they do not take into account possible interactions between different months.
+For example, yield might respond strongly to average pollution in June and July, which would not be captured in Table \ref{tab:yield4monthly}.
+
+\hypertarget{coal-plants-or-overall-pollution}{%
+\subsection{Coal plants or overall pollution?}\label{coal-plants-or-overall-pollution}}
+
+Finally, while I identify high-pollution areas using the location of coal plants, an important question is how much of the overall effect is driven by coal plants and how much is driven by other polluting industries located in areas near coal plants.
+I try to answer this question in Table \ref{tab:yield7ivdiffindiff}, where I implement an instrumental variables differences-in-differences, essentially instrumenting for pollution prior to the opening of a coal plant with wind and instrumenting for the interaction term (pollution times plant opening) with wind times plant opening.
 
@@
 
-\caption{\label{tab:yieldtablemonthly}Month-specific exposure}
+\caption{\label{tab:yield7ivdiffindiff}Agricultural productivity and pollution source}
 \centering
 \begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
 \toprule
-  & monsoon & winter\\
-\midrule
-Month 1 & -0.808*** & 0.788\\
-(June/Nov.) & (0.066) & (0.588)\\
-Month 1 & 0.049 & -0.351\\
-(July/Dec.) & (0.060) & (0.967)\\
-Month 3 & -0.229*** & 4.758\\
-(Aug./Jan.) & (0.034) & (3.756)\\
-Month 4 & -0.215*** & -2.086***\\
-(Sept./Feb.) & (0.035) & (0.503)\\
-Month 5 & 0.149* & -3.788\\
-(Oct./March) & (0.082) & (3.210)\\
-\textbf{fixed effects:} & \textbf{} & \textbf{}\\
-village & Yes & Yes\\
-year & Yes & Yes\\
-\midrule
-F (month 1) & 1,430 & 566\\
-F (month 2) & 1,029 & 897\\
-F (month 3) & 2,713 & 358\\
-F (month 4) & 2,131 & 389\\
-F (month 5) & 592 & 565\\
-observations & 1,266,922 & 1,120,495\\
+  & (1) & (2) & (3)\\
+\midrule
+PM 2.5 & -0.299*** & -0.107* & -0.147**\\
+ & (0.066) & (0.063) & (0.063)\\
+PM 2.5 times coal & -0.328*** & -0.321*** & -0.304***\\
+ & (0.026) & (0.024) & (0.023)\\
+weather & No & Yes & Yes\\
+weather (expanded) & No & No & Yes\\
+\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{}\\
+year & Yes & Yes & Yes\\
+village & Yes & Yes & Yes\\
+\midrule
+F (pm) & 766 & 776 & 788\\
+F (pm times open) & 199 & 222 & 232\\
+observations & 1,266,922 & 1,266,922 & 1,266,922\\
 \bottomrule
@@
 \begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions is logged nightlights. * p<0.10 ** p<0.05 *** p<0.01
+\item Note: Standard errors are in parentheses and are clustered at the village level.
+The outcome in all regressions is estimated agricultural yield, defined as tons per hectare.
+* p<0.10 ** p<0.05 *** p<0.01
 \end{tablenotes}
@@
 
-I present these results in Table \ref{tab:ntltable}.
-To help simplify the coefficients, I recode wind to be the proportion of days in the year that a given village experiences wind in their direction, from a coal plant.
-Mean wind in this new specification is 0.066, indicating that a village in the sample experiences wind in its direction on around 6.6 percent of days, or 24 days of the year.
-
-We see a strong, negative relationship between pollution (lagged wind) and nightlights.
-Going from the 25th percentile to the 75th percentile of wind exposure is synonymous with a decrease in the growth of nightlights within a village by around 0.7 percent.
-While this is a relatively small magnitude, it is worth noting that there are more than 100,000 villages in this sample.
-While the effect for a single village might be small, the overall, aggregate effect on growth is quite substantial.
-To put this number in perspective, Castelló-Climent, Chaudhary, and Mukhopadhyay (2018) argue that a standard-deviation increase in the share of adults with a university education leads to 12 percent higher nightlights in India, while Storeygard (2016) shows that a four-fold increase in oil prices over six years leads to seven percent lower nightlights at a distance of around 500km from the nearest large port in sub-Saharan Africa.
-In other words, the 0.7 percent change in nightlights is quite large given this is annual variation.
+Columns one through three of Table \ref{tab:yield7ivdiffindiff} present specifications with progressively increasing weather controls.
+Column three presents the most extensive set of controls and I focus on that column here.
+First, there are now two first stages but the F-test indicates that the instruments are strong for both, including the interaction term (F = 232).
+Second, the first row shows the effect of pollution on agricultural productivity prior to the opening of a coal plant; the coefficient is negative but relatively smaller.
+Third, the interaction term is strongly negative and more than twice as large in magnitude as the level term, indicating a very clear effect specific to coal plants.
+This could be due to differences in emission patterns for different kinds of plants and factors; coal plants, for example, are responsible for much more mercury than other kinds of polluting industries.\footnote{\url{https://www.epa.gov/sites/default/files/2016-12/documents/nei2014v1_tsd.pdf}}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-bentayeb2012adverse}{}}%
-Bentayeb, Malek, Marzia Simoni, Nour Baiz, Dan Norback, Sandra Baldacci, Sara Maio, Giovanni Viegi, Isabella Annesi-Maesano, Geriatric Study in Europe on Health Effects of Air Quality in Nursing Homes (GERIE) Group, et al.
-2012.
-{``Adverse Respiratory Effects of Outdoor Air Pollution in the Elderly.''} \emph{{The International Journal of Tuberculosis and Lung Disease}} 16 (9): 1149--61.
-
 \leavevmode\vadjust pre{\hypertarget{ref-brunekreef2002air}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-burke2021using}{}}%
-Burke, Marshall, Anne Driscoll, David B Lobell, and Stefano Ermon. 2021. {``{Using satellite imagery to understand and promote sustainable development}.''} \emph{Science} 371 (6535): eabe8628.
-
 \leavevmode\vadjust pre{\hypertarget{ref-burney2014recent}{}}%
 Burney, Jennifer, and V Ramanathan. 2014. {``{Recent climate and air pollution impacts on Indian agriculture}.''} \emph{{Proceedings of the National Academy of Sciences}} 111 (46): 16319--24.
-
-\leavevmode\vadjust pre{\hypertarget{ref-castello2018higher}{}}%
-Castelló-Climent, Amparo, Latika Chaudhary, and Abhiroop Mukhopadhyay.
-2018.
-{``{Higher education and prosperity: From Catholic missionaries to luminosity in India}.''} \emph{{The Economic Journal}} 128 (616): 3039--75.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-gibson2021night}{}}%
-Gibson, John, Susan Olivia, Geua Boe-Gibson, and Chao Li. 2021. {``{Which night lights data should we use in economics, and where?}''} \emph{{Journal of Development Economics}} 149: 102602.
-
 \leavevmode\vadjust pre{\hypertarget{ref-graff2012impact}{}}%
@@
 Heft-Neal, Sam, Jennifer Burney, Eran Bendavid, Kara Voss, and Marshall Burke. 2019. {``{Air pollution and infant mortality: Evidence from Saharan dust}.''} {National Bureau of Economic Research}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-henderson2012measuring}{}}%
-Henderson, J Vernon, Adam Storeygard, and David N Weil. 2012. {``Measuring Economic Growth from Outer Space.''} \emph{{American Economic Review}} 102 (2): 994--1028.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-pope2000epidemiology}{}}%
-Pope III, C Arden.
-2000.
-{``Epidemiology of Fine Particulate Air Pollution and Human Health: Biologic Mechanisms and Who's at Risk?''} \emph{{Environmental Health Perspectives}} 108 (suppl 4): 713--23.
-
 \leavevmode\vadjust pre{\hypertarget{ref-pope2006health}{}}%
@@
 Rud, Juan Pablo. 2012. {``{Electricity provision and industrial development: Evidence from India}.''} \emph{{Journal of Development Economics}} 97 (2): 352--67.
-
-\leavevmode\vadjust pre{\hypertarget{ref-storeygard2016farther}{}}%
-Storeygard, Adam. 2016. {``{Farther on down the road: transport costs, trade and urban growth in sub-Saharan Africa}.''} \emph{{The Review of Economic Studies}} 83 (3): 1263--95.
 
@@
 pollution & Hammer et al. (2020) & global & monthly\\
-agriculture & Angopadhyay et al. (2022) & India & two seasons/year\\
+agriculture & Gangopadhya et al. (2022) & India & two seasons/year\\
 nightlights & Asher et al. (2021) & India & yearly\\
@@
 
+\begin{figure}
+\includegraphics{draft_files/figure-latex/plants-1} \caption[Coal plants in India from 1990 to 2010]{Coal plants in India from 1990 to 2010}\label{fig:plants}\floatfoot*{Note: The top figure shows the location of coal plants in 1990.
+The bottom figure shows the location of coal plants in 2010.}
+\end{figure}
+
+\FloatBarrier
+\newpage
+
+\begin{figure}
+\includegraphics{draft_files/figure-latex/windexample-1} \caption[Wind direction and aggregation examples (2010-01-01)]{Wind direction and aggregation examples (2010-01-01)}\label{fig:windexample}\floatfoot*{Note: The top figure shows the average wind direction on January 1st, 2010, with zero degrees indicating north.
+The points are the location of coal plants on that date.
+The bottom figure shows the distribution of pollution exposure in a specific district -- Guna district in Madhya Pradesh -- on the same date.}
+\end{figure}
+
+\FloatBarrier
+\newpage
+
 \begin{table}
 
-\caption{\label{tab:yieldtablehet1st}Heterogeneity in the effects of pollution on productivity, first-stage results}
+\caption{\label{tab:pollutiontable}Wind direction and particulate matter}
 \centering
 \begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2.8cm}>{\centering\arraybackslash}p{1.7cm}>{\centering\arraybackslash}p{1.7cm}>{\centering\arraybackslash}p{1.7cm}>{\centering\arraybackslash}p{1.7cm}>{\centering\arraybackslash}p{1.7cm}>{\centering\arraybackslash}p{1.7cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
 \toprule
-\multicolumn{1}{c}{ } & \multicolumn{1}{c}{(1)} & \multicolumn{1}{c}{(2)} & \multicolumn{1}{c}{(3)} & \multicolumn{1}{c}{(4)} & \multicolumn{2}{c}{(5)} \\
-\cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-7}
-  & PM & PM & PM & PM & PM & PM*rain\\
-\midrule
-wind & 0.003*** & 0.000*** & 0.003*** & 0.002*** & 0.001*** & -0.032***\\
- & (0.000) & (0.000) & (0.000) & (0.000) & (0.000) & (0.001)\\
-wind times rain &  &  &  &  & 0.001*** & 0.009***\\
- &  &  &  &  & (0.000) & (0.000)\\
-weather & Yes & Yes & Yes & Yes & Yes & Yes\\
-\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-village-season & Yes & Yes & Yes & Yes & Yes & Yes\\
-year & Yes & Yes & Yes & Yes & Yes & Yes\\
-\midrule
-observations & 1,051,010 & 1,320,211 & 1,193,764 & 1,177,457 & 2,371,221 & 2,371,221\\
+\multicolumn{1}{c}{ } & \multicolumn{2}{c}{1998-2015} & \multicolumn{2}{c}{2002-2013} \\
+\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
+  & (1) & (2) & (3) & (4)\\
+\midrule
+wind (10s) & 0.0023*** & 0.0016*** & 0.0057*** & 0.0015***\\
+ & (0.0005) & (0.0002) & (0.0005) & (0.0002)\\
+\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+village & Yes & Yes & Yes & Yes\\
+month & Yes & No & Yes & No\\
+district-month & No & Yes & No & Yes\\
+\midrule
+observations & 24,376,458 & 24,376,458 & 16,250,973 & 16,250,973\\
 \bottomrule
@@
 \begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the village level.
-The outcome in all regressions is estimated agricultural yield, defined as tons per hectare.
-The first two columns split the sample based on the median exposure (by village).
-The second two columns split the sample based on initial -- the first year available -- agricultural productivity.
-* p<0.10 ** p<0.05 *** p<0.01
+\item Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01
 \end{tablenotes}
@@
 
-\begin{table}[H]
-
-\caption{\label{tab:yieldtablepostplant}Wind direction and agricultural productivity, post-coal plant only}
-\centering
-\begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{3}{c}{all} & \multicolumn{1}{c}{monsoon} & \multicolumn{1}{c}{winter} \\
-\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6}
-  & (1) & (2) & (3) & (4) & (5)\\
-\midrule
-particulate matter & -1.153*** & -1.028*** & -2.270*** & -0.538*** & -1.688***\\
-(log PM 2.5) & (0.079) & (0.083) & (0.386) & (0.044) & (0.166)\\
-rain (z) &  & 0.017*** & 0.018*** & 0.073*** & -0.002***\\
- &  & (0.001) & (0.002) & (0.002) & (0.001)\\
-mean temp (10s) &  & -0.000*** & -0.008*** & 0.005*** & -0.007***\\
- &  & (0.000) & (0.001) & (0.000) & (0.000)\\
-\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-village-season & Yes & Yes & Yes & No & No\\
-year & Yes & Yes & No & Yes & Yes\\
-district-year-season & No & No & Yes & No & No\\
-village & No & No & No & Yes & Yes\\
-\midrule
-observations & 1,886,956 & 1,872,177 & 1,872,177 & 996,509 & 875,668\\
-\midrule
-\textbf{first stage:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-wind & 0.002*** & 0.002*** & 0.000*** & 0.004*** & 0.002***\\
- & (0.000) & (0.000) & (0.000) & (0.000) & (0.000)\\
-rain (z) &  & -0.002*** & 0.004*** & 0.023*** & -0.003***\\
- &  & (0.000) & (0.000) & (0.000) & \vphantom{1} (0.000)\\
-mean temp (10s) &  & -0.001*** & -0.001*** & -0.000*** & 0.000***\\
- &  & (0.000) & (0.000) & (0.000) & (0.000)\\
-\bottomrule
-\end{tabular}
-\begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the village level.
-The sample is restricted to village-years after a coal plant was built near the village.
-The outcome in all regressions estimated agricultural yield (tons per hectare).
-* p<0.10 ** p<0.05 *** p<0.01
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
-\FloatBarrier
-\newpage
-
 \begin{table}
@@
 \begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
 \toprule
-  & (1) & (2)\\
-\midrule
-wind (lag) & 0.0005 & \\
- & (0.0003) & \\
-wind & -0.0006*** & \\
- & (0.0002) & \\
-wind (lead) & -0.0001 & \\
- & (0.0001) & \\
-PM (lag) &  & 0.4340\\
- &  & (0.8666)\\
-PM &  & -0.6942***\\
- &  & (0.1497)\\
-PM (lead) &  & -0.0958\\
- &  & (0.5137)\\
-weather & Yes & Yes\\
-\textbf{fixed effects:} & \textbf{} & \textbf{}\\
-village-season & Yes & Yes\\
-district-year-season & Yes & Yes\\
-\midrule
-F (lag) & -- & 162\\
-F & -- & 410\\
-F (lead) & -- & 289\\
-observations & 1,976,205 & 1,976,205\\
+  & (1) & (2) & (3)\\
+\midrule
+PM & -1.1524*** & -1.0915*** & -1.0387***\\
+ & (0.0755) & (0.0736) & (0.0731)\\
+PM (lead) & 0.0681 & 0.2890*** & 0.3091***\\
+ & (0.0716) & (0.0755) & (0.0727)\\
+weather & No & Yes & Yes\\
+weather (expanded) & No & No & Yes\\
+\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{}\\
+village & Yes & Yes & Yes\\
+year & Yes & Yes & Yes\\
+\midrule
+F & 974 & 993 & 1,014\\
+F (lead) & 612 & 585 & 571\\
+observations & 1,161,265 & 1,161,265 & 1,161,265\\
 \bottomrule
@@
 
-\begin{table}
-
-\caption{\label{tab:labortablemonsoon}Wind direction and labor allocation, monsoon season only}
-\centering
-\begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}}
-\toprule
-  & all & all & self & wage & farm & non-farm\\
-\midrule
-wind & -0.120*** & -0.119** & -0.067 & -0.052 & 0.025 & -0.144**\\
- & (0.046) & (0.046) & (0.042) & (0.045) & (0.036) & (0.062)\\
-controls & No & Yes & Yes & Yes & Yes & Yes\\
-\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-district & Yes & Yes & Yes & Yes & Yes & Yes\\
-state-month & Yes & Yes & Yes & Yes & Yes & Yes\\
-\textbf{DV mean} & \textbf{3.324} & \textbf{3.324} & \textbf{1.883} & \textbf{1.441} & \textbf{0.810} & \textbf{2.513}\\
-observations & 359,652 & 359,576 & 359,576 & 359,576 & 359,576 & 359,576\\
-\midrule
-\bottomrule
-\end{tabular}
-\begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the district level. Control variables include female, age, age squared, and (years of) education. * p<0.10 ** p<0.05 *** p<0.01
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
-\FloatBarrier
-\newpage
-
-\begin{table}
-
-\caption{\label{tab:labortablewinter}Wind direction and labor allocation, winter season only}
-\centering
-\begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}}
-\toprule
-  & all & all & self & wage & farm & non-farm\\
-\midrule
-wind & -0.012 & -0.032 & -0.026 & -0.006 & -0.020 & -0.012\\
- & (0.040) & (0.037) & (0.040) & (0.028) & (0.030) & (0.036)\\
-controls & No & Yes & Yes & Yes & Yes & Yes\\
-\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-district & Yes & Yes & Yes & Yes & Yes & Yes\\
-state-month & Yes & Yes & Yes & Yes & Yes & Yes\\
-\textbf{DV mean} & \textbf{3.339} & \textbf{3.339} & \textbf{1.865} & \textbf{1.474} & \textbf{0.790} & \textbf{2.550}\\
-observations & 375,956 & 375,883 & 375,883 & 375,883 & 375,883 & 375,883\\
-\midrule
-\bottomrule
-\end{tabular}
-\begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the district level.
-Control variables include female, age, age squared, and (years of) education.
-The outcome in each regression is days worked in the previous seven days.
-* p<0.10 ** p<0.05 *** p<0.01
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
-\FloatBarrier
-\newpage
-
-\begin{table}
-
-\caption{\label{tab:labortableoldstate}Wind direction and labor allocation, heterogeneity by age}
-\centering
-\begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}}
-\toprule
-  & all & self & wage & farm & non-farm\\
-\midrule
-wind & 0.100*** & 0.117*** & -0.018 & 0.026 & 0.074***\\
- & (0.027) & (0.029) & (0.026) & (0.024) & (0.028)\\
-wind times age>=38 & -0.203*** & -0.201*** & -0.002 & -0.070*** & -0.133***\\
- & (0.037) & (0.032) & (0.039) & (0.018) & (0.031)\\
-controls & Yes & Yes & Yes & Yes & Yes\\
-\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-district-year-month & Yes & Yes & Yes & Yes & Yes\\
-\midrule
-observations & 898,856 & 898,856 & 898,856 & 898,856 & 898,856\\
-\bottomrule
-\end{tabular}
-\begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the district level.
-Control variables include female, age, age squared, and (years of) education.
-The outcome in each regression is days worked in the previous seven days.
-* p<0.10 ** p<0.05 *** p<0.01
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
-\FloatBarrier
-\newpage
-
-\begin{table}
-
-\caption{\label{tab:wagestable}Wind direction and wages}
-\centering
-\begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}}
-\toprule
-  & farm & non-farm & all\\
-\midrule
-wind & 0.037 & 0.010 & 0.009\\
- & (0.031) & (0.010) & (0.011)\\
-controls & Yes & Yes & Yes\\
-\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{}\\
-district & Yes & Yes & Yes\\
-year & Yes & Yes & Yes\\
-\textbf{varying slopes:} & \textbf{} & \textbf{} & \textbf{}\\
-year (by district) & Yes & Yes & Yes\\
-\midrule
-observations & 16,534 & 149,973 & 133,526\\
-\bottomrule
-\end{tabular}
-\begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the district level.
-Control variables include female, age, age squared, and (years of) education.
-The outcome in each regression is the daily wage (log Rs per day).
-* p<0.10 ** p<0.05 *** p<0.01
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
-\FloatBarrier
-\newpage
-
 \begin{figure}

### 2023-03-02 analyses/1.openings.R

@@ -1,0 +1,79 @@
+# Purpose: This script estimates correlates of where plants open. save the results to load into the markdown script.
+# Author: Josh Merfeld
+# Date: February 25th, 2023
+
+rm(list = ls())
+
+library(tidyverse)
+library(lubridate)
+library(viridis)
+library(fixest)
+library(sf)
+
+
+# Sets wd to folder this script is in so we can create relative paths instead of absolute paths
+setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
+# We want to get the wd up to the main folder
+# Need to go up two levels
+setwd("../../")
+# Double check
+getwd()    # check
+
+
+
+# Where do plants open?
+villages90 <- read_csv("data/clean/census_plants/villages90.csv")
+villages90 <- villages90 %>% mutate(
+                                    literacy = pc91_pca_p_lit/pc91_pca_tot_p,
+                                    pop = pc91_pca_tot_p,
+                                    scst_prop = (pc91_pca_p_sc + pc91_pca_p_st)/pop,
+                                    tar_road = pc91_vd_tar_road
+                                    )
+villages00 <- read_csv("data/clean/census_plants/villages00.csv")
+villages00 <- villages00 %>% group_by(shrid) %>%
+                             mutate(
+                                    literacy = sum(pc01_pca_p_lit)/sum(pc01_pca_tot_p),
+                                    pop = sum(pc01_pca_tot_p),
+                                    scst_prop = (sum(pc01_pca_p_sc) + sum(pc01_pca_p_st))/pop,
+                                    tar_road = max(pc01_vd_tar_road)
+                                    ) %>%
+                              filter(row_number()==1) %>%
+                              ungroup()
+# monsoon at beginning
+monsoon <- read_csv("data/clean/census_plants/monsoon2002.csv")
+monsoon <- monsoon %>% group_by(shrid) %>% 
+                        mutate(ag_prod = log(mean(mean))) %>%
+                        filter(row_number()==1) %>%
+                        ungroup()
+villages00 <- villages00 %>% left_join(monsoon, by = "shrid")
+# rename to get same row in table
+villages90 <- villages90 %>% rename(area = pc91_vd_area)
+villages00 <- villages00 %>% rename(area = pc01_vd_area)
+
+
+reg1 <- feols(plants90 ~ log(pop) + log(area) + literacy, data = villages90, vcov = "hetero")
[…]

### 2023-03-02 analyses/2.pollution_wind_validation.R

@@ -1,0 +1,102 @@
+# Purpose: This script shows that wind does indeed predict pollution. I save the results to load into the markdown script
+# Author: Josh Merfeld
+# Date: February 25th, 2023
+
+rm(list = ls())
+
+library(tidyverse)
+library(lubridate)
+library(viridis)
+library(fixest)
+library(sf)
+
+
+# Sets wd to folder this script is in so we can create relative paths instead of absolute paths
+setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
+# We want to get the wd up to the main folder
+# Need to go up two levels
+setwd("../../")
+# Double check
+getwd()    # check
+
+
+
+pollution <- read_csv(paste0("data/clean/pm25/all_combined.csv"))
+pollution <- pollution %>% mutate(
+                                  date = as_date(paste0(year, "-", month, "-01")),
+                                  state = substr(shrid, 1, 2),
+                                  village = shrid,
+                                  month = as.numeric(date),
+                                  wind = sums
+                                  )
+pollution <- pollution %>% mutate()
+pollution <- panel(pollution, ~ village + date)
+
+# And district identifiers
+villages_overlap <- read_sf("data/spatial/villages_overlap/villages_overlap.shp")
+villages_overlap <- villages_overlap %>% mutate(shrid = paste0(pc11_s_id, "-", pc11_tv_id),
+                                                distfe = paste0(pc11_s_id, "-", pc11_d_id)) %>%
+                                          as_tibble() %>%
+                                          dplyr::select(shrid, distfe) %>%
+                                          group_by(shrid) %>%
+                                          filter(row_number()==1) %>%
+                                          ungroup()
+pollution <- pollution %>% left_join(villages_overlap, by = "shrid")
+
+pollution <- pollution %>% group_by(shrid) %>%
+                              mutate(pm_abs_dev = abs(mean - mean(mean)),
+                                     pm_dev = (mean - mean(mean)),
+                                     wind_dev = (wind - mean(wind))) %>%
+                              ungroup()
+pollution$month_int <- month(pollution$date)
+pollution <- pollution %>% group_by(shrid, month_int) %>%
+                              mutate(m_pm_dev = (mean - mean(mean)),
+                                     m_wind_dev = (wind - mean(wind))) %>%
[…]

### 2023-03-02 analyses/3.yield.R

@@ -1,0 +1,497 @@
+# Purpose: This script estimates effects for yields. I save the results to load into the markdown script
+# Author: Josh Merfeld
+# Date: February 25th, 2023
+
+rm(list = ls())
+
+library(tidyverse)
+library(lubridate)
+library(viridis)
+library(fixest)
+library(sf)
+library(openxlsx)
+
+
+# Sets wd to folder this script is in so we can create relative paths instead of absolute paths
+setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
+# We want to get the wd up to the main folder
+# Need to go up two levels
+setwd("../../")
+# Double check
+getwd()    # check
+
+
+
+# Cleaning ---------------------------------
+
+df <- read_csv("data/clean/ag_productivity/all_combined.csv")
+pollution <- read_csv(paste0("data/clean/pm25/ag_merge.csv"))
+df <- df %>% 
+        left_join(pollution, by = c("shrid", "year", "season"))
+rm(pollution)
+df <- df %>% group_by(shrid, season) %>% 
+              mutate(
+                     yield = mean,
+                     rain_z = (log(rain_total) - mean(log(rain_total))/sd(log(rain_total))),
+                     wind_abs_dev = abs(days_sums - mean(days_sums)),
+                     pm25_abs_dev = abs(pm25 - mean(pm25)),
+                     wind = days_sums/10,
+                     pm1 = log(pm1),
+                     pm2 = log(pm2),
+                     pm3 = log(pm3),
+                     pm4 = log(pm4),
+                     pm5 = log(pm5),
+                     pmavg1 = log(pm1),
+                     pmavg2 = log((pm1 + pm2)/2),
+                     pmavg3 = log((pm1 + pm2 + pm3)/3),
+                     pmavg4 = log((pm1 + pm2 + pm3 + pm4)/4),
+                     pmavg5 = log((pm1 + pm2 + pm3 + pm4 + pm5)/5),
+                     temp_m1 = (tmax_m1 + tmin_m1)/20,
+                     temp_m2 = (tmax_m2 + tmin_m2)/20,
+                     temp_m3 = (tmax_m3 + tmin_m3)/20,
+                     temp_m4 = (tmax_m4 + tmin_m4)/20,
+                     temp_m5 = (tmax_m5 + tmin_m5)/20,
+                     rain1_z = (log(rain_m1) - mean(log(rain_m1))/sd(log(rain_m1))),
[…]

### 2023-03-02 analyses/4.nss.R

@@ -1,0 +1,289 @@
+# Purpose: This script estimates effects with the NSS data. I save the results to load into the markdown script
+# Author: Josh Merfeld
+# Date: February 25th, 2023
+
+rm(list = ls())
+
+library(tidyverse)
+library(lubridate)
+library(viridis)
+library(fixest)
+library(sf)
+
+
+# Sets wd to folder this script is in so we can create relative paths instead of absolute paths
+setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
+# We want to get the wd up to the main folder
+# Need to go up two levels
+setwd("../../")
+# Double check
+getwd()    # check
+
+
+
+
+
+df <- read_csv("data/clean/nss/merged_week.csv")
+df <- df %>% filter(age>=15)
+df <- df %>% mutate(
+                    month_int = month(date),
+                    year = year(date),
+                    month = as.numeric(date),
+                    educ = as.factor(educ),
+                    wind = days_sum,
+                    district = distfe
+                    )
+pollution <- read_csv(paste0("data/clean/pm25/nss_merge.csv"))
+pollution <- pollution %>% mutate(
+                                  temp_mean = (month_tmin + month_tmax)/200,
+                                  season_temp_mean = (season_tmin + season_tmax)/200,
+                                  month_precip = log(month_precip + 1),
+                                  season_precip = log(season_precip + 1),
+                                  pm25 = log(pm25)
+                                  )
+df <- df %>% left_join(pollution, by = c("state_merge", "district_merge", "year", "month_int"))
+
+# MONSOON ONLY
+df <- df %>% filter(month_int %in% c(5,6,7,8,9,10,11))
+
+# control variables
+# NOTE: "season' variables are defined as the season UP TO THAT POINT, so interact with month_int factor
+setFixest_fml(..ctrl = ~ poly(female, 1) + poly(age, 2) + educ + temp_mean*month_precip + temp_mean^2 + month_precip^2 + 
+                as_factor(month_int)*season_temp_mean*season_precip + as_factor(month_int)*season_temp_mean^2 + as_factor(month_int)*season_precip^2)
+
+
+
+
+# means
+summary(df$days_f[df$month_int==6 & df$rural==1])
+summary(df$days_f[df$month_int==7 & df$rural==1])
+summary(df$days_f[df$month_int==8 & df$rural==1])
+summary(df$days_f[df$month_int==9 & df$rural==1])
[…]

### 2023-03-02 analyses/5.icrisat.R

@@ -1,0 +1,84 @@
+# Purpose: This script estimates effects with the NSS data. I save the results to load into the markdown script
+# Author: Josh Merfeld
+# Date: February 25th, 2023
+
+rm(list = ls())
+
+library(tidyverse)
+library(lubridate)
+library(readxl)
+
+
+
+# Sets wd to folder this script is in so we can create relative paths instead of absolute paths
+setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
+# We want to get the wd up to the main folder
+# Need to go up two levels
+setwd("../../")
+# Double check
+getwd()    # check
+
+
+
+df <- read_xlsx("/Users/Josh/Dropbox/Datasets/VDSA/EastIndia/2011/Cultivation/Cult_ip.xlsx")
+df <- rbind(df, read_xlsx("/Users/Josh/Dropbox/Datasets/VDSA/EastIndia/2012/Cultivation/Cult_ip.xlsx"))
+df$year <- paste0("20", substr(df$Sur_mon_yr, 4, 5))
+df$month <- substr(df$Sur_mon_yr, 1, 2)
+df$day <- "01"
+df$date <- as_date(paste0(df$year, "-", df$month, "-", df$day))
+
+
+
+
+# create new variables based on operations
+# first to lower
+df$Operation <- str_to_lower(df$Operation)
+df$op_plough <- str_detect(df$Operation, "plough")
+df$op_transplant <- str_detect(df$Operation, "transpl")
+df$op_sowing <- str_detect(df$Operation, "sow")
+df$op_fertilizer <- str_detect(df$Operation, "fertiliz")
+df$op_weeding <- str_detect(df$Operation, "weed")
+df$op_threshing <- str_detect(df$Operation, "threshing")
+df$op_harvest <- str_detect(df$Operation, "harvest")
+df <- df %>% filter(Season=="Kharif")
+
+df <- df %>%
+      group_by(date) %>%
+      mutate(
+             op_plough = mean(op_plough),
+             op_transplant = mean(op_transplant),
+             op_sowing = mean(op_sowing),
+             op_fertilizer = mean(op_fertilizer),
+             op_weeding = mean(op_weeding),
+             op_threshing = mean(op_threshing),
+             op_harvest = mean(op_harvest)
+             ) %>%
+      filter(row_number()==1) %>%
+      ungroup() %>%
+      dplyr::select(date, op_plough, op_transplant, op_sowing, op_fertilizer, op_weeding, op_threshing, op_harvest)
+
+
+
+
+
[…]

### 2023-03-02 analyses/5.nightlights.R

@@ -1,0 +1,138 @@
+# Purpose: This script estimates effects. I save the results to load into the markdown script
+# Author: Josh Merfeld
+# Date: December 12th, 2022
+
+rm(list = ls())
+
+library(tidyverse)
+library(lubridate)
+library(viridis)
+library(fixest)
+library(sf)
+
+
+# Sets wd to folder this script is in so we can create relative paths instead of absolute paths
+setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
+# We want to get the wd up to the main folder
+# Need to go up two levels
+setwd("../../")
+# Double check
+getwd()    # check
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+# Nightlights
+ntl <- read_csv("data/clean/village_ntl.csv")
+ntl <- ntl %>% mutate(state = substr(shrid, 1, 2), 
+                      wind = sums/365,
+                      rain_z = (log(rain) - mean(log(rain))/sd(log(rain))),
+                      village = shrid
+                      )
+
+pollution <- as_tibble(read_csv(paste0("data/clean/pm25/ntl_merge.csv")))
+ntl <- pollution %>% left_join(ntl, by = c("shrid", "year"))
+# ntltemp <- read_csv(paste0("data/clean/village_temp.csv"))
+# ntl <- ntl %>% left_join(ntltemp, by = c("shrid", "year"))
+
+# And district identifiers
+villages_overlap <- read_sf("data/spatial/villages_overlap/villages_overlap.shp")
+villages_overlap <- villages_overlap %>% mutate(shrid = paste0(pc11_s_id, "-", pc11_tv_id),
+                                                distfe = paste0(pc11_s_id, "-", pc11_d_id)) %>%
+                                          as_tibble() %>%
+                                          dplyr::select(shrid, distfe) %>%
+                                          group_by(shrid) %>%
+                                          filter(row_number()==1) %>%
+                                          ungroup()
+ntl <- ntl %>% left_join(villages_overlap, by = "shrid")
+ntl$pm25 <- log(ntl$pm25)
+
+ntl <- panel(ntl, ~ shrid + year, duplicate.method = "first")
+
+
+
+
+
+ntl1 <- feols(log(total_light + 1) ~ 1 | village + year | l(pm25, 1) ~ l(wind, 1),
+              data = ntl,
[…]

### 2023-03-02 analyses/regressions.R

@@ -1,2029 +1,0 @@
-# Purpose: This script estimates effects. I save the results to load into the markdown script
-# Author: Josh Merfeld
-# Date: December 12th, 2022
-
-rm(list = ls())
-
-library(tidyverse)
-library(lubridate)
-library(viridis)
-library(fixest)
-library(sf)
-
-
-# Sets wd to folder this script is in so we can create relative paths instead of absolute paths
-setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
-# We want to get the wd up to the main folder
-# Need to go up two levels
-setwd("../../")
-# Double check
-getwd()    # check
-
-
-
-
-# Where do plants open?
-villages90 <- read_csv("data/clean/census_plants/villages90.csv")
-villages90 <- villages90 %>% mutate(
-                                    literacy = pc91_pca_p_lit/pc91_pca_tot_p,
-                                    pop = pc91_pca_tot_p,
-                                    scst_prop = (pc91_pca_p_sc + pc91_pca_p_st)/pop,
-                                    tar_road = pc91_vd_tar_road
-                                    )
-villages00 <- read_csv("data/clean/census_plants/villages00.csv")
-villages00 <- villages00 %>% group_by(shrid) %>%
-                             mutate(
-                                    literacy = sum(pc01_pca_p_lit)/sum(pc01_pca_tot_p),
-                                    pop = sum(pc01_pca_tot_p),
-                                    scst_prop = (sum(pc01_pca_p_sc) + sum(pc01_pca_p_st))/pop,
-                                    tar_road = max(pc01_vd_tar_road)
-                                    ) %>%
-                              filter(row_number()==1) %>%
-                              ungroup()
-monsoon <- read_csv("data/clean/census_plants/monsoon2002.csv")
-monsoon <- monsoon %>% group_by(shrid) %>% 
-                        mutate(ag_prod = log(mean(mean))) %>%
-                        filter(row_number()==1) %>%
-                        ungroup()
-villages00 <- villages00 %>% left_join(monsoon, by = "shrid")
-
-
-reg1 <- feols(plants90 ~ log(pop) + literacy + tar_road, data = villages90, vcov = "hetero")
-reg2 <- feols(plants00 ~ log(pop) + literacy + tar_road, data = villages90 %>% filter(plants90==0), vcov = "hetero")
-reg3 <- feols(plants00 ~ ag_prod + log(pop) + literacy + tar_road, data = villages00, vcov = "hetero")
[…]

### 2023-03-02 cleaning/4.cleaning_plants.R

@@ -184,8 +184,8 @@ rad_deg <- function(rad){
 
 # GET PREVIOUS SEVEN DAYS (before start of loop)
 # Date
-date <- as.Date("2000-08-22", "%Y-%m-%d")
-final_date <- as.Date("2016-12-31", "%Y-%m-%d")
+date <- as.Date("2014-08-21", "%Y-%m-%d")
+final_date <- as.Date("2015-03-10", "%Y-%m-%d")
 
 wind_last_week <- dist_matrix %>% dplyr::select("shrid", "plant_id", "angle", "angle_original", "year_built", "year_retired")
 wind_last_week <- wind_last_week %>% group_by(shrid) %>% filter(row_number()==1) %>% ungroup()
@@ -554,7 +554,7 @@ while (date<=final_date){
 ## Aggregating to month/year ------------------------------------------------------------------------------------------
 ### PHEW! IT'S DONE! Now we need to pull all those daily csvs and aggregate them into months and years.
 # From 1990 through 2015
-today <- as.Date("1999-01-01", "%Y-%m-%d")
+today <- as.Date("2002-01-01", "%Y-%m-%d")
 final_date <- as.Date("2015-12-31", "%Y-%m-%d")
 
 while (today<=final_date){
@@ -572,152 +572,154 @@ while (today<=final_date){
     }
   }
   
-
-  # Going to set this up to redownload if there is an issue reading the csv
-  today_value <- NULL
-  try(
-    today_value <- read_csv(paste0("data/clean/wind_ntl/days/date_", year(today + 1), "-", month(today + 1), "-", day(today + 1), ".csv"))
-  )
-  if (is.null(today_value)==T){
-    wind_last_week <- dist_matrix %>% dplyr::select("shrid", "plant_id", "angle", "angle_original", "year_built", "year_retired")
-    wind_last_week <- wind_last_week %>% group_by(shrid) %>% filter(row_number()==1) %>% ungroup()
-    for (lag in 1:7){
-      wind_last_week_temp <- wind_last_week
-      wind_last_week_temp$sum <- 0
-      
-      # do it for TOMORROW
-      dl_year <- year(today + 1 - lag)
-      dl_month <- month(today + 1 - lag)
-      dl_day <- day(today + 1 - lag)
-      if (str_length(dl_month)==1){
-        dl_month <- paste0("0", dl_month)
-      }
-      if (str_length(dl_day)==1){
[…]

### 2023-03-02 cleaning/5.cleaning_precip_temp.R

@@ -23,7 +23,7 @@ setwd("../../")
 getwd()    # check
 
 
-
+# villages
 villages_centroids <- read_sf("data/spatial/centroids_villages_30km")
 villages_centroids <- as_tibble(villages_centroids) %>% dplyr::select(shrid)
 villages_centroids$within30 <- 1
@@ -31,6 +31,9 @@ villages <- read_sf("data/spatial/shapefiles/village.shp")
 villages$shrid <- paste0(villages$pc11_s_id, "-", villages$pc11_tv_id)
 villages <- villages %>% left_join(villages_centroids, by = "shrid")
 villages <- villages %>% filter(within30==1) %>% dplyr::select(shrid)
+
+# districts
+districts <- read_sf("data/spatial/districts/districts.shp")
 
 
 # Always using this
@@ -41,10 +44,12 @@ for (year in 1990:2020){
   file <- paste0("TerraClimate_ppt_", year, ".nc")
   curl::curl_download(file.path(url, file), file)
   r <- stack(file, varname = "ppt")
-  
+  # villages
   extracted_values <- exact_extract(r, villages, fun = "mean", append_cols = "shrid")
-  
   write_csv(extracted_values, paste0("data/clean/terra/precip", year, ".csv"))
+  # districts
+  extracted_values <- exact_extract(r, districts, fun = "mean", append_cols = c("ST_CEN_CD", "DT_CEN_CD"))
+  write_csv(extracted_values, paste0("data/clean/terra_district/precip", year, ".csv"))
   file.remove(file)
   print(year)
 }
@@ -54,10 +59,12 @@ for (year in 1990:2020){
   file <- paste0("TerraClimate_tmax_", year, ".nc")
   curl::curl_download(file.path(url, file), file)
   r <- stack(file, varname = "tmax")
-  
+  # villages
   extracted_values <- exact_extract(r, villages, fun = "mean", append_cols = "shrid")
-  
   write_csv(extracted_values, paste0("data/clean/terra/tmax", year, ".csv"))
+  # districts
+  extracted_values <- exact_extract(r, districts, fun = "mean", append_cols = c("ST_CEN_CD", "DT_CEN_CD"))
+  write_csv(extracted_values, paste0("data/clean/terra_district/tmax", year, ".csv"))
   file.remove(file)
   print(year)
 }
[…]

### 2023-03-02 cleaning/8.cleaning_overlap_merging_nss.R

@@ -19,8 +19,6 @@ setwd("../../")
 getwd()    # check
 
 
-
-
 # Village overlap data
 villages <- read_sf(paste0("data/spatial/villages_overlap/villages_overlap.shp"))
 villages <- villages %>% mutate(shrid = paste0(pc11_s_id, "-", pc11_tv_id),
@@ -59,7 +57,7 @@ date_vec <- unique(df_labor$date[is.na(df_labor$date)==F])
 # First, make the df smaller by dropping districts that do not match to the villages
 village_merge <- villages %>% mutate(state_merge = as.numeric(state),
                                      district_merge = as.numeric(district)) %>%
-                              select("state_merge", "district_merge")
+                              dplyr::select("state_merge", "district_merge")
 village_merge$merge <- 1
 village_merge <- village_merge %>% group_by(state_merge, district_merge) %>%
                                     filter(row_number()==1) %>%
@@ -73,7 +71,6 @@ df_labor <- df_labor %>% filter(is.na(merge)==F)
 
 pollution <- read_csv(paste0("data/clean/pm25/all_combined.csv"))
 pollution <- pollution %>% rename(pm25 = mean)
-
 
 # Going to go through the dates and process date by date
 df_labor_merged <- c()
@@ -92,7 +89,7 @@ for (day in 1:length(date_vec)){
   pollution_merge <- pollution %>% filter(year==year(date_vec[day]) & month==month(date_vec[day]))
   # there are a couple mistakes in terms of duplicate ids. fix those
   pollution_merge <- pollution_merge %>% group_by(shrid) %>% filter(row_number()==1) %>% ungroup()
-  pollution_merge <- pollution_merge %>% select(shrid, pm25)
+  pollution_merge <- pollution_merge %>% dplyr::select(shrid, pm25)
   village_wind <- village_wind %>% left_join(pollution_merge, by = "shrid")
   
   # monthly wind
@@ -101,7 +98,7 @@ for (day in 1:length(date_vec)){
   village_month_wind <- village_month_wind %>% group_by(shrid) %>% filter(row_number()==1) %>% ungroup()
   # rename
   colnames(village_month_wind) <- c("shrid", "month_sum")
-  village_month_wind <- village_month_wind %>% select(shrid, month_sum)
+  village_month_wind <- village_month_wind %>% dplyr::select(shrid, month_sum)
   village_wind <- village_wind %>% left_join(village_month_wind, by = "shrid")
   
   # Weighted mean by AREA. Any parts of the district without villages are ZEROS
[…]

### 2023-03-02 draft/draft.Rmd
draft update

@@
 I first plot the location of coal plants in India by year.
-Using a village-level shapefile, I identify all villages that are located within 30km of a coal plant at any time during the sample period, which comprises the 1990s to the 2010s, with differences depending on data availability.
+Using a village-level shapefile, I identify all villages that are located within 30km of a coal plant at any time during the sample period, which comprises the 2000s to the 2010s, with differences across outcomes depending on data availability.
 
@@
 Also in China, @he2019severe find small but significant negative effects of pollution on manufacturing productivity, while @chen2022effect show pollution leads to labor out migration.
-Perhaps relevant to my labor allocation results, @hanna2015effect show that decreases in pollution lead to increases in labor allocation, while I show the same relationship among the oldest cohort.
+Perhaps relevant to my labor allocation results, @hanna2015effect show that decreases in pollution lead to increases in labor allocation, while I show the same relationship in at least a subset of months of the agricultural season.
 The papers most related to this one are @aragon2016polluting, who show that gold mines in Ghana have substantial negative impacts on agricultural productivity in the surrounding areas, and @burney2014recent, who model the effects of climate change and pollutants on agricultural productivity in India.
@@
 
-The main goal of this paper is to examine whether exposure to emissions from coal plants affects agricultural productivity, labor allocation, and economic growth.
-I use several sources of data, which I describe here and in Table \@ref(tab:data) in the appendix.
+The main goal of this paper is to examine whether exposure to pollution and emissions from coal plants, in particular, affect agricultural productivity.
+I use several sources of data, which I describe here and list in Table \@ref(tab:data) of the appendix.
 
@@
 This means that any measurement error will be constant across time and subsumed by the village fixed effects.
-While the measure of output is also noisy, the use of satellite-derived data and a static cropland mask reduces concerns about measurement error being responsible for the results below.
+While the measure of output is also noisy, the use of satellite-derived data and a static cropland mask reduces concerns about measurement error being responsible for the results below, especially since common sources of measurement error are unlikely to be correlated with wind direction.
 The authors define the monsoon season as June to October and the winter season as November to March.
-I keep these definitions when matching across data below.
+I focus on only the monsoon season and keep the authors' definition when matching across data below.
 This data is also publicly available from Nature's data-sharing website.[^7] Since the main regressions below use village fixed effects, any differences in cropping patterns across areas should not bias estimates on changes, as long as these patterns do not change markedly in response to pollution.
@@
 To aggregate this data up to a useful administrative unit, I use village-level shapefiles provided by @almn2021 and publicly available on the SHRUG platform.[^8] I aggregate the agricultural productivity data to the village level by extracting mean productivity for each feature in the shapefile.
-I do this separately for each season -- monsoon and winter -- and each year, from 2002 to 2013.
+I do this separately for the monsoon season each year, from 2002 to 2013.
 
@@
 I then calculate the direction from coal plants to all village centroids within that 30km radius.
-To define exposure, I then pull daily wind direction data from the National Center for Atmospheric Research.[^9] For each day, I document whether the wind is blowing towards each village centroid.[^10] I then temporally aggregate this daily data depending on the temporal definition of the corresponding outcome.
+To define exposure, I then pull daily wind direction data from the National Center for Atmospheric Research.[^9] For each day, I document whether the wind is blowing towards each village centroid[^10] and then temporally aggregate this daily data depending on the temporal definition of the corresponding outcome.
 For example, agricultural productivity is defined across five months (e.g.
@@
 I am primarily interested in the effects of pollution on agricultural productivity. We might consider a regression of the following form:
-
 \begin{gather} y_{it} = \alpha_{i} + \gamma_{t} + \beta PM_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it}, \end{gather}
 
-where $y_{it}$ is outcome $y$ for unit $i$ at time $t$, $\alpha$ is a geographic fixed effects, $\gamma_t$ is time fixed effects, $PM_{it}$ is the concentration of particulate matter 2.5, $rain_{it}$ is total rainfall during the season, and $temp_{it}$ is average temperature during the season.
+where $y_{it}$ is outcome $y$ for unit $i$ at time $t$, $\alpha$ is a geographic fixed effects, $\gamma_t$ is time fixed effects, $PM_{it}$ is the (log) concentration of particulate matter 2.5, $rain_{it}$ is total rainfall during the season, and $temp_{it}$ is average temperature during the season.
 
@@
 For example, if coal plants are placed in places in need of power -- for example, due to rapid growth -- then the estimates might be upwards biased.
-On the other hand, if official purposefully place coal plants in poorer areas for political reasons, then the estimates might be downwards biased.
+On the other hand, if officials purposefully place coal plants in poorer areas for political reasons, then the estimates might be downwards biased.
 
@@
 Villages with higher populations and higher literacy in 1991 are more likely to be near a coal plant at the time.
-Interestingly, however, they are less likely to have a paved road.
 
@@
 This column restricts the sample to only villages that do not already have a coal plant within 30km in 1991.
-Population is no longer a significant predictor while the coefficients on literacy and the presence of a paved road have flipped.
+Population and literacy both remain strong predictors of the building of a new coal plant.
 
@@
 Identification relies on changes in wind direction across time for the same geographic units.
-In other words, conditional on the fixed effects and weather, I assume changes in wind direction are as good as random and uncorrelated with the outcome -- e.g.
-agricultural productivity -- except through exposure to pollution.
-One major threat to identification is that the building of a coal plant could lead to a spurious correlation between the exposure variable and the outcome.
-Specifically, since exposure is defined as zero prior to the building of a coal plant, the exposure variable itself might be endogenous even though wind direction is not.
-To help allay these concerns, I estimate additional regressions restricting the sample to only time periods after which a given location has a coal plant.
-I also present results using leads to show that future wind direction does not seem to predict current outcomes.
+In other words, conditional on the fixed effects and weather, I assume changes in wind direction are as good as random and uncorrelated with the outcome -- primarily agricultural productivity in this paper -- except through exposure to pollution.
 
@@
 Across all four columns, the story is the same: when the wind is blowing in the direction of a village from a coal plant within 30 kilometers, estimated $\mathrm{PM_{2.5}}$ is substantially higher.
-The regressions indicate that one additional day of wind in the direction of the village increases mean $\mathrm{PM_{2.5}}$ for the month by between `r paste0(substr(pollutiontable[1,2],1,5))` and `r paste0(substr(pollutiontable[1,3],1,5))` $\mathrm{\mu g/m^3}$.
 
@@
 \begin{gather} 
-y_{it} = \alpha_{i} + \gamma_{t} + \omega pm_{it} + \phi open_{it} + \beta pm_{it}\times open_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it},
+y_{it} = \alpha_{i} + \gamma_{t} + \omega PM_{it} + \phi open_{it} + \beta PM_{it}\times open_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it},
 \end{gather}
 
-where the interaction term $pm_{it}\times open_{it}$ indicates the change in the effect of pollution following the opening of the coal plant.
+where the interaction term $PM_{it}\times open_{it}$ indicates the change in the effect of pollution following the opening of the coal plant.
 Note that this is not the same as saying that pollution increased; instead, it is focusing on the change in the effect of a change in pollution of one unit on the outcome of interest, agricultural yield.
-I instrument for both $pm_{it}$ and $pm_{it}\times open_{it}$ using wind and the interaction between wind and the open indicator, or $wind_{it}\times open_{it}$.
+I instrument for both $PM_{it}$ and $PM_{it}\times open_{it}$ using wind and the interaction between wind and the open indicator, or $wind_{it}\times open_{it}$.
 
@@
 
-I first present results for agricultural productivity, with the reduced form.
+I first present the reduced form results for agricultural productivity.
 The outcome of all the regressions in this section is the log of agricultural land productivity, which is defined as tons per hectare.
@@
 Instead, we are actually interested in better understanding the effects of exposure to pollution on agricultural productivity.
-Table \@ref(tab:yieldtablepm) presents the same regressions as Table \@ref(fig:yieldtable), with a key difference: instead of wind exposure, it uses logged average concentration of $\mathrm{PM_{2.5}}$ -- from @hammer2020global -- during the season as the key explanatory variable.
+Table \@ref(tab:yieldtablepm) presents the same regressions as Table \@ref(tab:yieldtable), with a key difference: instead of wind exposure, it uses log average concentration of $\mathrm{PM_{2.5}}$ -- from @hammer2020global -- during the season as the key explanatory variable.
 There is again a consistent, negative relationship between particulate matter and agricultural productivity.
-Since both variables are logged, the coefficients are similar to elasticities, indicating that a one-percent increase in pollution leads to a 0.03-to-0.04 decrease in agricultural productivity.
+Since both variables are logged, the coefficients are similar to elasticities, indicating that a one-percent increase in pollution leads to a 0.03-to-0.04 percent decrease in agricultural productivity.
 
 Recall, however, the results from above: coal plants were placed in places with higher average agricultural productivity.
-This means that the simple regressions in \@ref(tab:yieldtablepm) might actually be biased upwards: although pollution might negatively affect productivity, pollution is higher in more productive areas due to the endogenous placement of coal plants in the study areas.
+This means that the simple regressions in Table \@ref(tab:yieldtablepm) might actually be biased upwards: although pollution might negatively affect productivity, pollution is higher in more productive areas due to the endogenous placement of coal plants in the study areas.
 To recover the actual effect of pollution, we need an instrument.
@@
 
-I present the IV results in Table \@ref(tab:yieldtabletwo), with the first-stage F-statistic in the bottom.
+I present the IV results in Table \@ref(tab:yieldtabletwo), with the first-stage F-statistic at the bottom.
 The first-stage results make clear how strong of an instrument wind is; all three F-statistics are above 1,000, meeting even the higher bar for the first-stage F-test argued for by @lee2022valid.
@@
 All regressions use the IV strategy detailed above and I include the F statistic from the first stage regressions (of which there are two in the last column) at the bottom of the table.
-Table \@ref(tab:yieldtablehet1st) in the appendix presents the first-stage results for all columns.
 

### 2023-03-02 draft/draft.tex
draft update

@@
 Also in China, He, Liu, and Salvo (2019) find small but significant negative effects of pollution on manufacturing productivity, while Chen, Oliva, and Zhang (2022) show pollution leads to labor out migration.
-Perhaps relevant to my labor allocation results, Hanna and Oliva (2015) show that decreases in pollution lead to increases in labor allocation, while I show the same relationship among the oldest cohort.
+Perhaps relevant to my labor allocation results, Hanna and Oliva (2015) show that decreases in pollution lead to increases in labor allocation, while I show the same relationship in at least a subset of months of the agricultural season.
 The papers most related to this one are Aragón and Rud (2016), who show that gold mines in Ghana have substantial negative impacts on agricultural productivity in the surrounding areas, and Burney and Ramanathan (2014), who model the effects of climate change and pollutants on agricultural productivity in India.
@@
 
-The main goal of this paper is to examine whether exposure to emissions from coal plants affects agricultural productivity, labor allocation, and economic growth.
-I use several sources of data, which I describe here and in Table \ref{tab:data} in the appendix.
+The main goal of this paper is to examine whether exposure to pollution and emissions from coal plants, in particular, affect agricultural productivity.
+I use several sources of data, which I describe here and list in Table \ref{tab:data} of the appendix.
 
@@
 This means that any measurement error will be constant across time and subsumed by the village fixed effects.
-While the measure of output is also noisy, the use of satellite-derived data and a static cropland mask reduces concerns about measurement error being responsible for the results below.
+While the measure of output is also noisy, the use of satellite-derived data and a static cropland mask reduces concerns about measurement error being responsible for the results below, especially since common sources of measurement error are unlikely to be correlated with wind direction.
 The authors define the monsoon season as June to October and the winter season as November to March.
-I keep these definitions when matching across data below.
+I focus on only the monsoon season and keep the authors' definition when matching across data below.
 This data is also publicly available from Nature's data-sharing website.\footnote{springernature.figshare.com} Since the main regressions below use village fixed effects, any differences in cropping patterns across areas should not bias estimates on changes, as long as these patterns do not change markedly in response to pollution.
@@
 (2021) and publicly available on the SHRUG platform.\footnote{www.devdatalab.org/shrug\_download/} I aggregate the agricultural productivity data to the village level by extracting mean productivity for each feature in the shapefile.
-I do this separately for each season -- monsoon and winter -- and each year, from 2002 to 2013.
+I do this separately for the monsoon season each year, from 2002 to 2013.
 
@@
 I then calculate the direction from coal plants to all village centroids within that 30km radius.
-To define exposure, I then pull daily wind direction data from the National Center for Atmospheric Research.\footnote{climatedataguide.ucar.edu/} For each day, I document whether the wind is blowing towards each village centroid.\footnote{I define ``towards'' as within five degrees to help take into account that I am using village centroids, using the x and y components of wind speed to calculate angles.} I then temporally aggregate this daily data depending on the temporal definition of the corresponding outcome.
+To define exposure, I then pull daily wind direction data from the National Center for Atmospheric Research.\footnote{climatedataguide.ucar.edu/} For each day, I document whether the wind is blowing towards each village centroid\footnote{I define ``towards'' as within five degrees to help take into account that I am using village centroids, using the x and y components of wind speed to calculate angles.} and then temporally aggregate this daily data depending on the temporal definition of the corresponding outcome.
 For example, agricultural productivity is defined across five months (e.g.~the monsoon season is from June to October) so I count the total days a given village is exposed to wind blowing from \emph{any} coal plant within 30km during those five months.
@@
 I am primarily interested in the effects of pollution on agricultural productivity. We might consider a regression of the following form:
-
 \begin{gather} y_{it} = \alpha_{i} + \gamma_{t} + \beta PM_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it}, \end{gather}
 
-where \(y_{it}\) is outcome \(y\) for unit \(i\) at time \(t\), \(\alpha\) is a geographic fixed effects, \(\gamma_t\) is time fixed effects, \(PM_{it}\) is the concentration of particulate matter 2.5, \(rain_{it}\) is total rainfall during the season, and \(temp_{it}\) is average temperature during the season.
+where \(y_{it}\) is outcome \(y\) for unit \(i\) at time \(t\), \(\alpha\) is a geographic fixed effects, \(\gamma_t\) is time fixed effects, \(PM_{it}\) is the (log) concentration of particulate matter 2.5, \(rain_{it}\) is total rainfall during the season, and \(temp_{it}\) is average temperature during the season.
 
@@
 Identification relies on changes in wind direction across time for the same geographic units.
-In other words, conditional on the fixed effects and weather, I assume changes in wind direction are as good as random and uncorrelated with the outcome -- e.g.~agricultural productivity -- except through exposure to pollution.
-One major threat to identification is that the building of a coal plant could lead to a spurious correlation between the exposure variable and the outcome.
-Specifically, since exposure is defined as zero prior to the building of a coal plant, the exposure variable itself might be endogenous even though wind direction is not.
-To help allay these concerns, I estimate additional regressions restricting the sample to only time periods after which a given location has a coal plant.
-I also present results using leads to show that future wind direction does not seem to predict current outcomes.
+In other words, conditional on the fixed effects and weather, I assume changes in wind direction are as good as random and uncorrelated with the outcome -- primarily agricultural productivity in this paper -- except through exposure to pollution.
 
@@
 Across all four columns, the story is the same: when the wind is blowing in the direction of a village from a coal plant within 30 kilometers, estimated \(\mathrm{PM_{2.5}}\) is substantially higher.
-The regressions indicate that one additional day of wind in the direction of the village increases mean \(\mathrm{PM_{2.5}}\) for the month by between 0.00 and 0.00 \(\mathrm{\mu g/m^3}\).
 
@@
 \begin{gather} 
-y_{it} = \alpha_{i} + \gamma_{t} + \omega pm_{it} + \phi open_{it} + \beta pm_{it}\times open_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it},
+y_{it} = \alpha_{i} + \gamma_{t} + \omega PM_{it} + \phi open_{it} + \beta PM_{it}\times open_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it},
 \end{gather}
 
-where the interaction term \(pm_{it}\times open_{it}\) indicates the change in the effect of pollution following the opening of the coal plant.
+where the interaction term \(PM_{it}\times open_{it}\) indicates the change in the effect of pollution following the opening of the coal plant.
 Note that this is not the same as saying that pollution increased; instead, it is focusing on the change in the effect of a change in pollution of one unit on the outcome of interest, agricultural yield.
-I instrument for both \(pm_{it}\) and \(pm_{it}\times open_{it}\) using wind and the interaction between wind and the open indicator, or \(wind_{it}\times open_{it}\).
+I instrument for both \(PM_{it}\) and \(PM_{it}\times open_{it}\) using wind and the interaction between wind and the open indicator, or \(wind_{it}\times open_{it}\).
 
@@
 Instead, we are actually interested in better understanding the effects of exposure to pollution on agricultural productivity.
-Table \ref{tab:yieldtablepm} presents the same regressions as Table \ref{fig:yieldtable}, with a key difference: instead of wind exposure, it uses logged average concentration of \(\mathrm{PM_{2.5}}\) -- from Hammer et al.
+Table \ref{tab:yieldtablepm} presents the same regressions as Table \ref{tab:yieldtable}, with a key difference: instead of wind exposure, it uses log average concentration of \(\mathrm{PM_{2.5}}\) -- from Hammer et al.
 (2020) -- during the season as the key explanatory variable.
 There is again a consistent, negative relationship between particulate matter and agricultural productivity.
-Since both variables are logged, the coefficients are similar to elasticities, indicating that a one-percent increase in pollution leads to a 0.03-to-0.04 decrease in agricultural productivity.
+Since both variables are logged, the coefficients are similar to elasticities, indicating that a one-percent increase in pollution leads to a 0.03-to-0.04 percent decrease in agricultural productivity.
 
 Recall, however, the results from above: coal plants were placed in places with higher average agricultural productivity.
-This means that the simple regressions in \ref{tab:yieldtablepm} might actually be biased upwards: although pollution might negatively affect productivity, pollution is higher in more productive areas due to the endogenous placement of coal plants in the study areas.
+This means that the simple regressions in Table \ref{tab:yieldtablepm} might actually be biased upwards: although pollution might negatively affect productivity, pollution is higher in more productive areas due to the endogenous placement of coal plants in the study areas.
 To recover the actual effect of pollution, we need an instrument.
@@
 
-I present the IV results in Table \ref{tab:yieldtabletwo}, with the first-stage F-statistic in the bottom.
+I present the IV results in Table \ref{tab:yieldtabletwo}, with the first-stage F-statistic at the bottom.
 The first-stage results make clear how strong of an instrument wind is; all three F-statistics are above 1,000, meeting even the higher bar for the first-stage F-test argued for by D.
-S.
 Lee et al.
@@
 All regressions use the IV strategy detailed above and I include the F statistic from the first stage regressions (of which there are two in the last column) at the bottom of the table.
-Table \ref{tab:yieldtablehet1st} in the appendix presents the first-stage results for all columns.
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-lee2022valid}{}}%
-Lee, David S, Justin McCrary, Marcelo J Moreira, and Jack Porter. 2022. {``{Valid t-ratio Inference for IV}.''} \emph{{American Economic Review}} 112 (10): 3260--90.
+Lee, David, Justin McCrary, Marcelo J Moreira, and Jack Porter. 2022. {``{Valid t-ratio Inference for IV}.''} \emph{{American Economic Review}} 112 (10): 3260--90.
 

### 2023-03-02 draft/draft.Rmd
draft

@@
 ---
-title: "Pollution, agricultural productivity, and development: Evidence from coal plants in India^[I thank Marc Bellemare and David Sungho Park for comments and suggestions that greatly improved the quality of the paper.
+title: "Blowing in the wind: Pollution and agricultural productivity in India^[I thank Marc Bellemare and David Sungho Park for comments and suggestions that greatly improved the quality of the paper.
 The usual disclaimer applies.

### 2023-03-02 draft/draft.tex
draft

@@
 \hypersetup{
-  pdftitle={Pollution, agricultural productivity, and development: Evidence from coal plants in India},
+  pdftitle={Blowing in the wind: Pollution and agricultural productivity in India},
   pdfauthor={Joshua D. Merfeld},
@@
 
-\title{Pollution, agricultural productivity, and development: Evidence from coal plants in India\footnote{I thank Marc Bellemare and David Sungho Park for comments and suggestions that greatly improved the quality of the paper.
+\title{Blowing in the wind: Pollution and agricultural productivity in India\footnote{I thank Marc Bellemare and David Sungho Park for comments and suggestions that greatly improved the quality of the paper.
 The usual disclaimer applies.

### 2023-04-07 draft/draft.Rmd
changes to draft, analyses, and cleaning

@@
 ---
-title: "Blowing in the wind: Pollution and agricultural productivity in India^[I thank Marc Bellemare and David Sungho Park for comments and suggestions that greatly improved the quality of the paper.
+title: "Blowing in the wind: Pollution and agricultural productivity in India^[I thank Ariel Ortiz-Bobea, Marc Bellemare, and David Sungho Park for comments and suggestions that greatly improved the quality of the paper.
 The usual disclaimer applies.
@@
 library(raster)
+library(viridis)
 loadfonts()  # Must be run once in each session (markdown is a new "session" each time you render/knit)
@@
 yield3ivmain <- readRDS("pollution_development/draft/tables/yield3ivmain.rds")
+yield3ivmain_lead <- readRDS("pollution_development/draft/tables/yield3ivmain_lead.rds")
 yield4monthly <- readRDS("pollution_development/draft/tables/yield4monthly.rds")
@@
 opplot <- readRDS("pollution_development/draft/tables/opplot.rds")
+coef_vec <- readRDS("pollution_development/draft/tables/randomization_coef_vec.rds")
+coefs_distfe <- readRDS("pollution_development/draft/tables/randomization_coefs_distfe.rds")
+true_value <- readRDS("pollution_development/draft/tables/randomization_true_value.rds")
 
@@
 Second, I use a different identification strategy from either paper, relying on wind blowing pollution from coal plants.
+
+@sanders2022adaptation
 
@@
 colnames(plantresultstable) <- c("1991", "2001", "2001", "2011")
+rownames(plantresultstable) <- c("Population (log)", "", "Area (log)", "", "Literacy (prop)", "", "Ag. productivity", "", 
+                                "Sub-sample", "Observations")
 kable(
@@
 colnames(yield1reducedform) <- c("(1)", "(2)", "(3)")
+
+rownames(yield1reducedform) <- c("Wind days", " ", "Rain (z)", "", "Mean temperature (10s)", "",
+                                  "Rain squared", " ", "Temperature squared", " ", "Rain times Temperature", " ",
+                                  "Fixed effects:", "Village", "Year",
+                                  "Observations")
 kable(
@@
            ) %>%
-  column_spec(1, width = "3cm") %>%
+  column_spec(1, width = "4cm") %>%
   column_spec(c(2:4),width = "2.5cm") %>%
@@
 colnames(yield2naive) <- c("(1)", "(2)", "(3)")
+rownames(yield2naive) <- c("Particulate matter", "(log PM 2.5)", 
+                           "Weather", "Weather (expanded)",
+                           "Fixed effects:", "Village", "Year",
+                           "Observations")
 kable(
@@
 # Table
+yield3ivmain <- yield3ivmain[,-1]
 colnames(yield3ivmain) <- c("(1)", "(2)", "(3)")
-rownames(yield3ivmain)[8] <- "F (first stage)"
+rownames(yield3ivmain) <- c("Particulate matter", "(log PM 2.5)",
+                            "Weather", "Weather (expanded)",
+                            "Fixed effects:", "Village", "Year",
+                            "F (first stage)", 
+                            "Observations")
 kable(
@@
            ) %>%
-  column_spec(1, width = "3cm") %>%
+  column_spec(1, width = "4cm") %>%
   column_spec(c(2:4),width = "2cm") %>%
@@
 
+## Robustness checks
+
+```{r randomplot, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 8, fig.cap = "Randomization inference", fig.footer = list("Note: The figure presents the distribution of coefficients by randomly assigning wind direction across villages within the same year.
+The \"all\" distribution is created by randomly assigning a single wind value to each village, sampled with replacement across all villages in a given year.
+The \"within district\" randomly samples with replacement only from within the same district in the same year.")}
+
+pal <- viridis(4)
+colors <- c(paste0(pal[1]), paste0(pal[2]))
+labels <- c("All", "Within district")
+
+# plot
+gg1 <- ggplot() + 
+        geom_density(aes(x = coef_vec, fill = "All"), alpha = 0.5) +
+        geom_density(aes(x = coefs_distfe, fill = "Within district"), alpha = 0.5) +
+        scale_fill_manual(
+          name = "Randomization:",
+          values = colors,
+          labels = labels
+        ) +
+        theme_minimal() +
+        xlab("Distribution of coefficients") +
+        ylab("Density")
+
+colors <- c(paste0(pal[3]))
+labels <- c("actual")
+
+gg1 <- gg1 + 
+        geom_vline(aes(xintercept = true_value, color = "actual")) +
+        scale_color_manual(
+          name = "",
+          values = colors,
+          labels = labels
+        ) +
+        theme(legend.position = c(0.4, 0.6))
+
+gg1
+
+```
+
+```{r yieldtableleads, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+# Table
+colnames(yield3ivmain_lead) <- c("(1)", "(2)")
+rownames(yield3ivmain_lead) <- c("Particulate matter", "(log PM 2.5)",
+                                 "Particulate matter", "(one-year lead)",
+                                 "Weather",
+                                 "Fixed effects:", "Village", "Year",
+                                 "F (current)", "F (lead)", 
+                                 "Observations")
+kable(
+      yield3ivmain_lead, caption = "Agricultural productivity and pollution leads",
+      align = "cc", booktabs = TRUE, linesep = ""
+      ) %>%
+  footnote(
+           "Note: Standard errors are in parentheses and are clustered at the village level.
+The outcome in all regressions estimated agricultural yield (tons per hectare).
+* p<0.10 ** p<0.05 *** p<0.01",
+           general_title = "",
+           threeparttable = TRUE,
+           footnote_as_chunk = TRUE
+           ) %>%
+  column_spec(1, width = "3.5cm") %>%
+  column_spec(c(2:3),width = "2cm") %>%
+  row_spec(c(8), hline_after = TRUE) %>%
+  row_spec(c(6), bold = TRUE) %>%
+  kable_classic_2()
+
+```
+
 ## Heterogeneity
@@
 rownames(yield5heterogeneity)[c(1, 3)] <- c("PM 2.5", "PM times rain")
+rownames(yield5heterogeneity) <- c("Particulate matter", "(log PM 2.5)", "PM times Rain", "", "Weather",
+                                   "Fixed effects:", "Village", "Year", 
+                                   "F (1st stage, PM)", "F (1st stage, PM times rain)", 
+                                   "Observations")
 kable(
@@
   row_spec(c(6), bold = TRUE) %>%
-  add_header_above(c(" " = 1, "wind" = 2, "yield" = 2, " " = 1)) %>%
+  add_header_above(c(" " = 1, "Wind" = 2, "Yield" = 2, " " = 1)) %>%
   kable_classic_2()
@@
 # Table
-colnames(labor1all) <- c("all", "all", "self", "wage", "farm", "non-farm")
-colnames(labor2rural) <- c("all", "all", "self", "wage", "farm", "non-farm")
+colnames(labor1all) <- c("All", "All", "Self", "Wage", "Farm", "Non-farm")
+colnames(labor2rural) <- c("All", "All", "Self", "Wage", "Farm", "Non-farm")
 
 mat1 <- matrix(c("", "", "", "", "", ""), nrow = 1)
-colnames(mat1) <- c("all", "all", "self", "wage", "farm", "non-farm")
-rownames(mat1) <- c("panel a: all")
+colnames(mat1) <- c("All", "All", "Self", "Wage", "Farm", "Non-farm")
+rownames(mat1) <- c("Panel A: all")
 mat2 <- matrix(c("", "", "", "", "", ""), nrow = 1)
-colnames(mat2) <- c("all", "all", "self", "wage", "farm", "non-farm")
-rownames(mat2) <- c("panel b: rural")
+colnames(mat2) <- c("All", "All", "Self", "Wage", "Farm", "Non-farm")
+rownames(mat2) <- c("Panel B: Rural")
 
 labortablerural <- rbind(mat1, labor1all[c(1:2, 7:8),], mat2, labor2rural[-c(3:6),], labor2rural[c(3:6),])
+rownames(labortablerural)[c(2,7)] <- "Wind"
+rownames(labortablerural)[c(5,10)] <- "Observations"
+rownames(labortablerural)[11] <- "Controls"
+rownames(labortablerural)[12] <- "Fixed effects:"
+rownames(labortablerural)[13] <- "District"
+rownames(labortablerural)[14] <- "State-month"
 
@@
 
+labor3month <- cbind(labor3month, yield4monthly[-c(15:19),2])
+
 # Table
-colnames(labor3month) <- c("farm", "non-farm")
+colnames(labor3month) <- c("Farm", "Non-farm", "Yield")
 kable(
-      labor3month, caption = "Wind and labor allocation, by month",
-      align = "cc", booktabs = TRUE, linesep = ""
+      labor3month, caption = "Heterogeneity by month",
+      align = "ccc", booktabs = TRUE, linesep = ""
       ) %>%
   footnote(
-           "Note: Standard errors are in parentheses and are clustered at the village level. The outcome in column one is .  * p<0.10 ** p<0.05 *** p<0.01", 
+           "Note: Standard errors are in parentheses and are clustered at the village level.
+The outcome in column one is farm labor, the outcome is column two is non-farm labor, and the outcome in column three is village-level productivity.
+The independent variable for labor is wind, while the IV approach above is still used for yield.
+* p<0.10 ** p<0.05 *** p<0.01",
            general_title = "",
@@
            ) %>%
-  column_spec(1, width = "3cm") %>%
-  column_spec(c(2:3),width = "2cm") %>%
-  row_spec(c(18), hline_after = TRUE) %>%
-  row_spec(c(16), bold = TRUE) %>%
+  column_spec(1, width = "2.5cm") %>%
+  column_spec(c(2:4),width = "2.5cm") %>%
+  row_spec(c(14), hline_after = TRUE) %>%
+  row_spec(c(12), bold = TRUE) %>%
+  add_header_above(c(" " = 1, "Labor" = 2, " " = 1)) %>%
   kable_classic_2()
@@
 
-I perform a similar exercise with yield, the results of which are in Table \@ref(tab:yield4monthly).
+I perform a similar exercise with yield, the results of which are in Table \@ref(tab:labor3month).
 I include only the five months of the year that are used in the calculation of the yield estimates, or June through October.
@@
 
-```{r yield4monthly, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+A final caveat about these month-specific results is that they do not take into account possible interactions between different months.
+For example, yield might respond strongly to average pollution in June and July, which would not be captured in Table \@ref(tab:labor3month).
+
+## Coal plants or overall pollution?
+
+Finally, while I identify high-pollution areas using the location of coal plants, an important question is how much of the overall effect is driven by coal plants and how much is driven by other polluting industries located in areas near coal plants.
+I try to answer this question in Table \@ref(tab:yield7ivdiffindiff), where I implement an instrumental variables differences-in-differences, essentially instrumenting for pollution prior to the opening of a coal plant with wind and instrumenting for the interaction term (pollution times plant opening) with wind times plant opening.
+
+```{r yield7ivdiffindiff, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
 
 # Table
-colnames(yield4monthly) <- c("(1)", "(2)")
+colnames(yield7ivdiffindiff) <- c("(1)", "(2)", "(3)")
+rownames(yield7ivdiffindiff) <- c("Particulate matter", "(log PM 2.5)", 
+                                  "PM 2.5 times Coal", "",
+                                  "Weather", "Weather (expanded)",
+                                  "Fixed effects:", "Year", "Village",
+                                  "F (PM)", "F (PM times Coal)", "Observations")
 kable(
-      yield4monthly, caption = "Pollution and productivity, by month",
-      align = "cc", booktabs = TRUE, linesep = ""
+      yield7ivdiffindiff, caption = "Agricultural productivity and pollution source",
+      align = "ccc", booktabs = TRUE, linesep = ""
       ) %>%
@@
            ) %>%
-  column_spec(1, width = "3cm") %>%
-  column_spec(c(2:3),width = "2cm") %>%
-  row_spec(12, bold = TRUE) %>%
-  row_spec(14, hline_after = TRUE) %>%
-  #add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
-  kable_classic_2()
-
-```
-
-A final caveat about these month-specific results is that they do not take into account possible interactions between different months.
-For example, yield might respond strongly to average pollution in June and July, which would not be captured in Table \@ref(tab:yield4monthly).
-
-## Coal plants or overall pollution?
-
-Finally, while I identify high-pollution areas using the location of coal plants, an important question is how much of the overall effect is driven by coal plants and how much is driven by other polluting industries located in areas near coal plants.
-I try to answer this question in Table \@ref(tab:yield7ivdiffindiff), where I implement an instrumental variables differences-in-differences, essentially instrumenting for pollution prior to the opening of a coal plant with wind and instrumenting for the interaction term (pollution times plant opening) with wind times plant opening.
-
-```{r yield7ivdiffindiff, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-# Table
-colnames(yield7ivdiffindiff) <- c("(1)", "(2)", "(3)")
-kable(
-      yield7ivdiffindiff, caption = "Agricultural productivity and pollution source",
-      align = "ccc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the village level.
-The outcome in all regressions is estimated agricultural yield, defined as tons per hectare.
-* p<0.10 ** p<0.05 *** p<0.01",
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
-           ) %>%
-  column_spec(1, width = "3cm") %>%
+  column_spec(1, width = "3.5cm") %>%
   column_spec(c(2:4),width = "2cm") %>%
@@
 
-data <- matrix(NA, nrow = 8, ncol = 3)
-rownames(data) <- c("shapefile", "coal plants", "wind", "pollution", "agriculture", "nightlights", "weather", "labor")
+data <- matrix(NA, nrow = 7, ncol = 3)
+rownames(data) <- c("shapefile", "coal plants", "wind", "pollution", "agriculture", "weather", "labor")
 colnames(data) <- c("source", "geographic coverage", "temporal coverage")
@@
 data[5,1] <- "Gangopadhya et al. (2022)"
-data[6,1] <- "Asher et al. (2021)"
-data[7,1] <- "TerraClimate"
-data[8,1] <- "National Sample Survey (NSS)"
+data[6,1] <- "TerraClimate"
+data[7,1] <- "National Sample Survey (NSS)"
 data[1,2] <- "India"
@@
 data[5,2] <- "India"
-data[6,2] <- "India"
-data[7,2] <- "global"
-data[8,2] <- "India"
+data[6,2] <- "global"
+data[7,2] <- "India"
 data[1,3] <- " "
@@
 data[5,3] <- "two seasons/year"
-data[6,3] <- "yearly"
-data[7,3] <- "monthly"
-data[8,3] <- "daily interviews"
+data[6,3] <- "monthly"
+data[7,3] <- "daily interviews"
 
@@
                      y = "lat",
-                     subtitle = "Figure A: 1990")
+                     subtitle = "A. 1990")
 
@@
                      y = "lat",
-                     subtitle = "Figure B: 2010")
+                     subtitle = "B. 2010")
 
@@
              y = "lat",
-             subtitle = "Figure A: Wind direction") +
+             subtitle = "A. Wind direction") +
         theme(legend.position = c(0.6, 0.05),
@@
              y = "lat",
-             subtitle = "Figure B: Aggregation to district") +
+             subtitle = "B. Aggregation to district") +
         theme(legend.position = c(0.5, 0.05),
@@
 
-```{r yieldtableleads, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-# Table
-colnames(yield6leads) <- c("(1)", "(2)", "(3)")
-kable(
-      yield6leads, caption = "Agricultural productivity and pollution leads",
-      align = "ccc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the village level.
-The outcome in all regressions estimated agricultural yield (tons per hectare).
-* p<0.10 ** p<0.05 *** p<0.01",
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
-           ) %>%
-  column_spec(1, width = "3cm") %>%
-  column_spec(c(2:4),width = "2cm") %>%
-  row_spec(c(9), hline_after = TRUE) %>%
-  row_spec(c(7), bold = TRUE) %>%
-  kable_classic_2()
-
-```
-
 ```{=tex}

### 2023-04-07 draft/draft.tex
changes to draft, analyses, and cleaning

@@
 
-\title{Blowing in the wind: Pollution and agricultural productivity in India\footnote{I thank Marc Bellemare and David Sungho Park for comments and suggestions that greatly improved the quality of the paper.
+\title{Blowing in the wind: Pollution and agricultural productivity in India\footnote{I thank Ariel Ortiz-Bobea, Marc Bellemare, and David Sungho Park for comments and suggestions that greatly improved the quality of the paper.
 The usual disclaimer applies.
@@
 \author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
-\date{2023-03-02}
+\date{2023-04-07}
 
@@
 
+Sanders and Barreca (2022)
+
 I also contribute to the more general literature on the effects of pollution on health.
@@
 \midrule
-pop (log) & 0.010*** & 0.001*** & 0.017*** & 0.004***\\
+Population (log) & 0.010*** & 0.001*** & 0.017*** & 0.004***\\
  & (0.000) & (0.000) & (0.000) & (0.000)\\
-area (log) & -0.006*** & 0.001** & -0.015*** & -0.001***\\
+Area (log) & -0.006*** & 0.001** & -0.015*** & -0.001***\\
  & (0.001) & (0.000) & (0.000) & (0.000)\\
-literacy (prop) & 0.024*** & 0.017*** & 0.032*** & 0.029***\\
+Literacy (prop) & 0.024*** & 0.017*** & 0.032*** & 0.029***\\
  & (0.003) & (0.002) & (0.002) & (0.002)\\
-ag productivity &  &  & 0.026*** & 0.004***\\
+Ag. productivity &  &  & 0.026*** & 0.004***\\
  &  &  & (0.001) & (0.000)\\
-sub-sample & all & no plant & all & no plant\\
-\midrule
-observations & 283,021 & 264,265 & 508,784 & 469,004\\
+Sub-sample & all & no plant & all & no plant\\
+\midrule
+Observations & 283,021 & 264,265 & 508,784 & 469,004\\
 \bottomrule
@@
 \begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}}
 \toprule
@@
 \midrule
-wind days & -0.0078*** & -0.0069*** & -0.0062***\\
+Wind days & -0.0078*** & -0.0069*** & -0.0062***\\
  & (0.0007) & (0.0007) & (0.0006)\\
-rain (z) &  & 0.0854*** & -0.0424***\\
+Rain (z) &  & 0.0854*** & -0.0424***\\
  &  & (0.0018) & (0.0046)\\
-mean temp (10s) &  & -0.0021*** & 0.1321***\\
+Mean temperature (10s) &  & -0.0021*** & 0.1321***\\
  &  & (0.0001) & (0.0034)\\
-rain square &  &  & -0.0010***\\
+Rain squared &  &  & -0.0010***\\
  &  &  & (0.0001)\\
-temp square &  &  & -0.0002***\\
+Temperature squared &  &  & -0.0002***\\
  &  &  & \vphantom{1} (0.0000)\\
-rain times temp &  &  & 0.0003***\\
+Rain times Temperature &  &  & 0.0003***\\
  &  &  & (0.0000)\\
-\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{}\\
-village & Yes & Yes & Yes\\
-year & Yes & Yes & Yes\\
-\midrule
-observations & 1,266,922 & 1,266,922 & 1,266,922\\
+\textbf{Fixed effects:} & \textbf{} & \textbf{} & \textbf{}\\
+Village & Yes & Yes & Yes\\
+Year & Yes & Yes & Yes\\
+\midrule
+Observations & 1,266,922 & 1,266,922 & 1,266,922\\
 \bottomrule
@@
 \midrule
-particulate matter & -0.042*** & -0.044*** & -0.032***\\
+Particulate matter & -0.042*** & -0.044*** & -0.032***\\
 (log PM 2.5) & (0.003) & (0.003) & (0.003)\\
-weather & No & Yes & Yes\\
-weather (expanded) & No & No & Yes\\
-\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{}\\
-village & Yes & Yes & Yes\\
-year & Yes & Yes & Yes\\
-\midrule
-observations & 1,266,922 & 1,266,922 & 1,266,922\\
+Weather & No & Yes & Yes\\
+Weather (expanded) & No & No & Yes\\
+\textbf{Fixed effects:} & \textbf{} & \textbf{} & \textbf{}\\
+Village & Yes & Yes & Yes\\
+Year & Yes & Yes & Yes\\
+\midrule
+Observations & 1,266,922 & 1,266,922 & 1,266,922\\
 \bottomrule
@@
 \begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
 \toprule
@@
 \midrule
-particulate matter & -0.941*** & -0.813*** & -0.728***\\
+Particulate matter & -0.941*** & -0.813*** & -0.728***\\
 (log PM 2.5) & (0.078) & (0.076) & (0.075)\\
-weather & No & Yes & Yes\\
-weather (expanded) & No & No & Yes\\
-\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{}\\
-village & Yes & Yes & Yes\\
-year & Yes & Yes & Yes\\
-\midrule
-F (first stage) & 1,235 & 1,262 & 1,263\\
-observations & 1,266,922 & 1,266,922 & 1,266,922\\
+Weather & No & Yes & Yes\\
+Weather (expanded) & No & No & Yes\\
+\textbf{Fixed effects:} & \textbf{} & \textbf{} & \textbf{}\\
+Village & Yes & Yes & Yes\\
+Year & Yes & Yes & Yes\\
+\midrule
+F (first stage) & 1,236 & 1,263 & 1,264\\
+Observations & 1,266,922 & 1,266,922 & 1,266,922\\
 \bottomrule
@@
 
+\hypertarget{robustness-checks}{%
+\subsection{Robustness checks}\label{robustness-checks}}
+
+\begin{figure}
+\includegraphics{draft_files/figure-latex/randomplot-1} \caption[Randomization inference]{Randomization inference}\label{fig:randomplot}\floatfoot*{Note: The figure presents the distribution of coefficients by randomly assigning wind direction across villages within the same year.
+The "all" distribution is created by randomly assigning a single wind value to each village, sampled with replacement across all villages in a given year.
+The "within district" randomly samples with replacement only from within the same district in the same year.}
+\end{figure}
+
+\begin{table}
+
+\caption{\label{tab:yieldtableleads}Agricultural productivity and pollution leads}
+\centering
+\begin{threeparttable}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3.5cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\toprule
+  & (1) & (2)\\
+\midrule
+Particulate matter & -1.152*** & -1.067***\\
+(log PM 2.5) & (0.075) & (0.071)\\
+Particulate matter & 0.068 & 0.268***\\
+(one-year lead) & (0.072) & (0.074)\\
+Weather & No & Yes\\
+\textbf{Fixed effects:} & \textbf{} & \textbf{}\\
+Village & Yes & Yes\\
+Year & Yes & Yes\\
+\midrule
+F (current) & 974 & 1,036\\
+F (lead) & 612 & 597\\
+Observations & 1,161,265 & 1,161,265\\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}[para]
+\item Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions estimated agricultural yield (tons per hectare). * p<0.10 ** p<0.05 *** p<0.01
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+
 \hypertarget{heterogeneity}{%
@@
 \toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{wind} & \multicolumn{2}{c}{yield} & \multicolumn{1}{c}{ } \\
+\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Wind} & \multicolumn{2}{c}{Yield} & \multicolumn{1}{c}{ } \\
 \cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
@@
 \midrule
-PM 2.5 & -0.767*** & -0.328*** & -1.130*** & -0.690*** & -0.409**\\
- & (0.104) & (0.095) & (0.060) & (0.146) & (0.160)\\
-PM times rain &  &  &  &  & 0.061***\\
+Particulate matter & -0.767*** & -0.328*** & -1.130*** & -0.690*** & -0.409**\\
+(log PM 2.5) & (0.104) & (0.095) & (0.060) & (0.146) & (0.160)\\
+PM times Rain &  &  &  &  & 0.061***\\
  &  &  &  &  & (0.014)\\
-weather & Yes & Yes & Yes & Yes & Yes\\
-\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-village & Yes & Yes & Yes & Yes & Yes\\
-year & Yes & Yes & Yes & Yes & Yes\\
+Weather & Yes & Yes & Yes & Yes & Yes\\
+\textbf{Fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+Village & Yes & Yes & Yes & Yes & Yes\\
+Year & Yes & Yes & Yes & Yes & Yes\\
 \midrule
@@
 F (1st stage, PM times rain) &  &  &  &  & 1,437\\
-observations & 617,804 & 649,118 & 634,342 & 632,580 & 1,266,922\\
+Observations & 617,804 & 649,118 & 634,342 & 632,580 & 1,266,922\\
 \bottomrule
@@
 \toprule
-  & all & all & self & wage & farm & non-farm\\
-\midrule
-\textbf{panel a: all} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-wind & -0.011 & -0.009 & 0.030 & -0.049 & -0.009 & -0.010\\
+  & All & All & Self & Wage & Farm & Non-farm\\
+\midrule
+\textbf{Panel A: all} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+Wind & -0.011 & -0.009 & 0.030 & -0.049 & -0.009 & -0.010\\
  & (0.031) & (0.034) & (0.039) & (0.034) & (0.034) & (0.033)\\
 DV mean & 3.303 & 3.303 & 1.862 & 1.442 & 0.789 & 2.514\\
-observations & 523,075 & 522,957 & 522,957 & 522,957 & 522,957 & 522,957\\
-\midrule
-\textbf{panel b: rural} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-wind & -0.006 & -0.034 & 0.068 & -0.102*** & -0.024 & -0.010\\
+Observations & 523,075 & 522,957 & 522,957 & 522,957 & 522,957 & 522,957\\
+\midrule
+\textbf{Panel B: Rural} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+Wind & -0.006 & -0.034 & 0.068 & -0.102*** & -0.024 & -0.010\\
  & (0.044) & (0.040) & (0.043) & (0.035) & (0.043) & (0.036)\\
 DV mean & 3.494 & 3.494 & 2.184 & 1.311 & 1.271 & 2.223\\
-observations & 307,381 & 307,325 & 307,325 & 307,325 & 307,325 & 307,325\\
-\midrule
-controls & No & Yes & Yes & Yes & Yes & Yes\\
-\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-district & Yes & Yes & Yes & Yes & Yes & Yes\\
-state-month & Yes & Yes & Yes & Yes & Yes & Yes\\
+Observations & 307,381 & 307,325 & 307,325 & 307,325 & 307,325 & 307,325\\
+\midrule
+Controls & No & Yes & Yes & Yes & Yes & Yes\\
+\textbf{Fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+District & Yes & Yes & Yes & Yes & Yes & Yes\\
+State-month & Yes & Yes & Yes & Yes & Yes & Yes\\
 \bottomrule
@@
 
-\caption{\label{tab:labor3month}Wind and labor allocation, by month}
+\caption{\label{tab:labor3month}Heterogeneity by month}
 \centering
 \begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}}
 \toprule
-  & farm & non-farm\\
-\midrule
-May & -0.006 & -0.039\\
- & (0.068) & (0.070)\\
-June & -0.136* & 0.114\\
- & (0.080) & (0.099)\\
-July & -0.285** & 0.176\\
- & (0.142) & (0.134)\\
-Aug. & 0.048 & -0.010\\
- & (0.086) & (0.069)\\
-Sept. & 0.076 & -0.082\\
- & (0.085) & (0.091)\\
-Oct. & 0.013 & -0.083\\
- & (0.076) & (0.059)\\
-Nov. & -0.063 & -0.013\\
- & (0.062) & (0.075)\\
-controls & Yes & Yes\\
-\textbf{fixed effects:} & \textbf{} & \textbf{}\\
-district & Yes & Yes\\
-year-month & Yes & Yes\\
-\midrule
-observations & 307,325 & 307,325\\
+\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Labor} & \multicolumn{1}{c}{ } \\
+\cmidrule(l{3pt}r{3pt}){2-3}
+  & Farm & Non-farm & Yield\\
+\midrule
+June & -0.128* & 0.123 & -0.736***\\
+ & (0.076) & (0.095) & (0.062)\\
+July & -0.278* & 0.185 & 0.051\\
+ & (0.142) & (0.133) & (0.095)\\
+Aug. & 0.055 & -0.002 & 0.087\\
+ & (0.083) & (0.068) & (0.070)\\
+Sept. & 0.086 & -0.072 & -0.140\\
+ & (0.081) & (0.093) & (0.180)\\
+Oct. & 0.022 & -0.074 & 0.334\\
+ & (0.071) & (0.057) & (0.291)\\
+controls & Yes & Yes & Yes\\
+\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{}\\
+district & Yes & Yes & Yes\\
+year-month & Yes & Yes & Yes\\
+\midrule
+observations & 307,325 & 307,325 & 1,266,922\\
 \bottomrule
@@
 \begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the village level. The outcome in column one is .  * p<0.10 ** p<0.05 *** p<0.01
+\item Note: Standard errors are in parentheses and are clustered at the village level.
+The outcome in column one is farm labor, the outcome is column two is non-farm labor, and the outcome in column three is village-level productivity.
+The independent variable for labor is wind, while the IV approach above is still used for yield.
+* p<0.10 ** p<0.05 *** p<0.01
 \end{tablenotes}
@@
 
-I perform a similar exercise with yield, the results of which are in Table \ref{tab:yield4monthly}.
+I perform a similar exercise with yield, the results of which are in Table \ref{tab:labor3month}.
 I include only the five months of the year that are used in the calculation of the yield estimates, or June through October.
@@
 
+A final caveat about these month-specific results is that they do not take into account possible interactions between different months.
+For example, yield might respond strongly to average pollution in June and July, which would not be captured in Table \ref{tab:labor3month}.
+
+\hypertarget{coal-plants-or-overall-pollution}{%
+\subsection{Coal plants or overall pollution?}\label{coal-plants-or-overall-pollution}}
+
+Finally, while I identify high-pollution areas using the location of coal plants, an important question is how much of the overall effect is driven by coal plants and how much is driven by other polluting industries located in areas near coal plants.
+I try to answer this question in Table \ref{tab:yield7ivdiffindiff}, where I implement an instrumental variables differences-in-differences, essentially instrumenting for pollution prior to the opening of a coal plant with wind and instrumenting for the interaction term (pollution times plant opening) with wind times plant opening.
+
 \begin{table}
 
-\caption{\label{tab:yield4monthly}Pollution and productivity, by month}
+\caption{\label{tab:yield7ivdiffindiff}Agricultural productivity and pollution source}
 \centering
 \begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3.5cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
 \toprule
-  & (1) & (2)\\
-\midrule
-June & -0.325** & -0.736***\\
- & (0.135) & (0.062)\\
-July & -0.778*** & 0.051\\
- & (0.151) & (0.095)\\
-Aug. & -1.201*** & 0.087\\
- & (0.455) & (0.070)\\
-Sept. & 1.270*** & -0.140\\
- & (0.427) & (0.180)\\
-Oct. & -3.013*** & 0.334\\
- & (0.963) & (0.291)\\
-weather & No & Yes\\
-\textbf{fixed effects:} & \textbf{} & \textbf{}\\
-village & Yes & Yes\\
-year & Yes & Yes\\
-\midrule
-F (June) & 1,063 & 972\\
-F (July) & 2,090 & 2,200\\
-F (Aug.) & 1,646 & 2,340\\
-F (Sept.) & 2,025 & 2,024\\
-F (Oct.) & 473 & 402\\
-observations & 1,266,922 & 1,266,922\\
+  & (1) & (2) & (3)\\
+\midrule
+Particulate matter & -0.299*** & -0.107* & -0.147**\\
+(log PM 2.5) & (0.066) & (0.063) & (0.063)\\
+PM 2.5 times Coal & -0.328*** & -0.321*** & -0.304***\\
+ & (0.026) & (0.024) & (0.023)\\
+Weather & No & Yes & Yes\\
+Weather (expanded) & No & No & Yes\\
+\textbf{Fixed effects:} & \textbf{} & \textbf{} & \textbf{}\\
+Year & Yes & Yes & Yes\\
+Village & Yes & Yes & Yes\\
+\midrule
+F (PM) & 766 & 776 & 788\\
+F (PM times Coal) & 199 & 222 & 232\\
+Observations & 1,266,922 & 1,266,922 & 1,266,922\\
 \bottomrule
@@
 
-A final caveat about these month-specific results is that they do not take into account possible interactions between different months.
-For example, yield might respond strongly to average pollution in June and July, which would not be captured in Table \ref{tab:yield4monthly}.
-
-\hypertarget{coal-plants-or-overall-pollution}{%
-\subsection{Coal plants or overall pollution?}\label{coal-plants-or-overall-pollution}}
-
-Finally, while I identify high-pollution areas using the location of coal plants, an important question is how much of the overall effect is driven by coal plants and how much is driven by other polluting industries located in areas near coal plants.
-I try to answer this question in Table \ref{tab:yield7ivdiffindiff}, where I implement an instrumental variables differences-in-differences, essentially instrumenting for pollution prior to the opening of a coal plant with wind and instrumenting for the interaction term (pollution times plant opening) with wind times plant opening.
-
-\begin{table}
-
-\caption{\label{tab:yield7ivdiffindiff}Agricultural productivity and pollution source}
-\centering
-\begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
-\toprule
-  & (1) & (2) & (3)\\
-\midrule
-PM 2.5 & -0.299*** & -0.107* & -0.147**\\
- & (0.066) & (0.063) & (0.063)\\
-PM 2.5 times coal & -0.328*** & -0.321*** & -0.304***\\
- & (0.026) & (0.024) & (0.023)\\
-weather & No & Yes & Yes\\
-weather (expanded) & No & No & Yes\\
-\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{}\\
-year & Yes & Yes & Yes\\
-village & Yes & Yes & Yes\\
-\midrule
-F (pm) & 766 & 776 & 788\\
-F (pm times open) & 199 & 222 & 232\\
-observations & 1,266,922 & 1,266,922 & 1,266,922\\
-\bottomrule
-\end{tabular}
-\begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the village level.
-The outcome in all regressions is estimated agricultural yield, defined as tons per hectare.
-* p<0.10 ** p<0.05 *** p<0.01
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
 Columns one through three of Table \ref{tab:yield7ivdiffindiff} present specifications with progressively increasing weather controls.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-sanders2022adaptation}{}}%
+Sanders, Nicholas J, and Alan I Barreca.
+2022.
+{``{Adaptation to Environmental Change: Agriculture and the Unexpected Incidence of the Acid Rain Program}.''} \emph{{American Economic Journal: Economic Policy}} 14 (1): 373--401.
+
 \leavevmode\vadjust pre{\hypertarget{ref-van2017long}{}}%
@@
 agriculture & Gangopadhya et al. (2022) & India & two seasons/year\\
-nightlights & Asher et al. (2021) & India & yearly\\
 weather & TerraClimate & global & monthly\\
@@
 
-\begin{table}
-
-\caption{\label{tab:yieldtableleads}Agricultural productivity and pollution leads}
-\centering
-\begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
-\toprule
-  & (1) & (2) & (3)\\
-\midrule
-PM & -1.1524*** & -1.0915*** & -1.0387***\\
- & (0.0755) & (0.0736) & (0.0731)\\
-PM (lead) & 0.0681 & 0.2890*** & 0.3091***\\
- & (0.0716) & (0.0755) & (0.0727)\\
-weather & No & Yes & Yes\\
-weather (expanded) & No & No & Yes\\
-\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{}\\
-village & Yes & Yes & Yes\\
-year & Yes & Yes & Yes\\
-\midrule
-F & 974 & 993 & 1,014\\
-F (lead) & 612 & 585 & 571\\
-observations & 1,161,265 & 1,161,265 & 1,161,265\\
-\bottomrule
-\end{tabular}
-\begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions estimated agricultural yield (tons per hectare). * p<0.10 ** p<0.05 *** p<0.01
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
 \FloatBarrier

### 2023-04-07 presentation/slides/slides.Rmd
changes to draft, analyses, and cleaning

--- 2023-02-11 presentation/slides/slides.Rmd
@@
 knitr::opts_chunk$set(
-  fig.width=9, fig.height=3.5, fig.retina=3,
+                      fig.width = 9, fig.height = 3.5, fig.retina = 3,
                       out.width = "100%",
@@
 library(lubridate)
-library(icons)
 library(png)
@@
 pollutiontable <- readRDS("pollution_development/draft/tables/pollutiontable.rds")
-yieldtable <- readRDS("pollution_development/draft/tables/yieldtable.rds")
-yieldtablepm <- readRDS("pollution_development/draft/tables/yieldtablepm.rds")
-yieldtabletwo <- readRDS("pollution_development/draft/tables/yieldtabletwo.rds")
-yieldtablehet <- readRDS("pollution_development/draft/tables/yieldtablehet.rds")
-pmeffects <- readRDS("pollution_development/draft/tables/yieldtablePM.rds")
-labortable <- readRDS("pollution_development/draft/tables/labortable.rds")
-labortablemonsoon <- readRDS("pollution_development/draft/tables/labortablemonsoon.rds")
-labortablewinter <- readRDS("pollution_development/draft/tables/labortablewinter.rds")
-labortableold <- readRDS("pollution_development/draft/tables/labortableold.rds")
-ntltable <- readRDS("pollution_development/draft/tables/ntltable.rds")
-yieldtableleads <- readRDS("pollution_development/draft/tables/yieldtableleads.rds")
-yieldtablepostplant <- readRDS("pollution_development/draft/tables/yieldtablepostplant.rds")
+yield1reducedform <- readRDS("pollution_development/draft/tables/yield1reducedform.rds")
+yield2naive <- readRDS("pollution_development/draft/tables/yield2naive.rds")
+yield3ivmain <- readRDS("pollution_development/draft/tables/yield3ivmain.rds")
+yield4monthly <- readRDS("pollution_development/draft/tables/yield4monthly.rds")
+yield5heterogeneity <- readRDS("pollution_development/draft/tables/yield5heterogeneity.rds")
+yield6leads <- readRDS("pollution_development/draft/tables/yield6leads.rds")
+yield7ivdiffindiff <- readRDS("pollution_development/draft/tables/yield7ivdiffindiff.rds")
+labor1all <- readRDS("pollution_development/draft/tables/labor1all.rds")
+labor2rural <- readRDS("pollution_development/draft/tables/labor2rural.rds")
+labor3month <- readRDS("pollution_development/draft/tables/labor3month.rds")
 laborplot <- readRDS("pollution_development/draft/tables/laborplot.rds")
+monthplot <- readRDS("pollution_development/draft/tables/monthplot.rds")
+opplot <- readRDS("pollution_development/draft/tables/opplot.rds")
 
@@
 
-![]("../../../pollution_development/draft/tables/energy.png")
+- india villages complain coal "power plant" (Adaniwatch)
 
@@
 
-## Coal and energy in India
+```{r energypng, echo = FALSE, message = FALSE, warning = FALSE}
 
-![]("../../../pollution_development/draft/tables/energy.png")
+knitr::include_graphics(
+                        path = "energy.png",
+                        auto_pdf = TRUE
+                        )
+
+```
 
@@
   - Exposure defined as wind direction from coal plants towards villages
-  - One additional day of exposure per season: decrease of 0.3 percent (yields)
+  - Also instrument for particulate matter 2.5 (PM 2.5)
 
@@
 
-- Adults decrease labor allocation
-  - Driven by large decrease in older half of sample; younger half increases allocation to farm labor
-- More pollution in previous year also leads to lower nightlight growth
-  - 25th to 75th percentile of exposure leads to decrease in nightlights of 0.7 percent
+- Effects on labor allocation decisions, as well
+- Largest effects on labor and yield in same month
+  - Around very beginning of the monsoon season (June)
+
+---
+
+## General idea
+- Coal plants open in a given year
+- Plot wind direction each day
+  - Exposure means wind blowing from coal plant to village (within 30km)
+  - Aggregate up to desired level (e.g. season, month, week)
 
@@
 ```
-
----
-
-## General idea
-- Coal plants open in a given year
-- Plot wind direction each day
-  - Exposure means wind blowing from coal plant to village (within 30km)
-  - Aggregate up to desired level (e.g. season, month, week)
 
@@
 
+## Empirical strategy
+
+First stage
+
+$$pm_{it} = \alpha^1_{i} + \gamma^1_{t} + \phi^1 wind_{it} + \psi^1 weather_{it} + \varepsilon^1_{it}$$
+
+Second stage
+
+$$yield_{it} = \alpha^2_{i} + \gamma^2_{t} + \phi^2 \hat{pm}_{it} + \psi^2 weather_{it} + \varepsilon^2_{it}$$
+
+- i indexes villages (districts for labor)
+- t indexes time (where time differs based on the analysis)
+- standard errors clustered at village (or district) throughout
+- common use of stars (one indicates 0.10) throughout
+
+---
+
 ## Where do coal plants open?
@@
       align = "cccc", booktabs = TRUE, linesep = "", escape = FALSE
-      ) %>%
-  footnote(
-           "Note: Robust standard errors are in parentheses.
-The outcome in the first column is whether the village is within 30km of a coal plant in 1991.
-The second column is whether a village in 1991 will have a coal plant in 2001, conditional on not having one in 1991.
-The last two columns are similarly defined, except using 2001 and 2011 as the years.
-* p<0.10 ** p<0.05 *** p<0.01",
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE,
-           escape = FALSE
       ) %>%
@@
 
-## Validating wind direction - particulate matter and exposure
+## Reduced form - wind and yield
+
+```{r yieldtable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+# Table
+colnames(yield1reducedform) <- c("(1)", "(2)", "(3)")
+kable(
+      yield1reducedform,
+      align = "ccc", booktabs = TRUE, linesep = ""
+      ) %>%
+  column_spec(1, width = "3cm") %>%
+  column_spec(c(2:4),width = "2.5cm") %>%
+  row_spec(13, bold = TRUE) %>%
+  row_spec(15, hline_after = TRUE) %>%
+  #add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
+  kable_classic_2()
+
+```
+
+---
+
+## First stage - particulate matter and exposure, monthly level
 
@@
       align = "cccc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
       ) %>%
@@
 
-## Effects of wind direction on agricultural productivity
+## PM and agricultural productivity
 
-```{r yieldtable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+```{r yieldtablepm, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
 
 # Table
-colnames(yieldtable) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
+colnames(yield2naive) <- c("(1)", "(2)", "(3)")
 kable(
-      yieldtable,
-      align = "ccccc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
+      yield2naive,
+      align = "ccc", booktabs = TRUE, linesep = ""
       ) %>%
   column_spec(1, width = "3cm") %>%
-  column_spec(c(2:6),width = "2cm") %>%
-  row_spec(7, bold = TRUE) %>%
-  row_spec(11, hline_after = TRUE) %>%
-  add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
+  column_spec(c(2:4),width = "2.5cm") %>%
+  row_spec(5, bold = TRUE) %>%
+  row_spec(7, hline_after = TRUE) %>%
+  #add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
   kable_classic_2()
@@
 
----
-
-## Effects of PM on agricultural productivity
-
 - Likely endogenous
   - Upward biased?
-
-```{r yieldtablepm, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-# Table
-colnames(yieldtablepm) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
-kable(
-      yieldtablepm[-c(3:6),],
-      align = "ccccc", booktabs = TRUE, linesep = ""
-      ) %>%
-  # footnote(
-  #          "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
-  #          general_title = "",
-  #          threeparttable = TRUE,
-  #          footnote_as_chunk = TRUE
-  #          ) %>%
-  column_spec(1, width = "3cm") %>%
-  column_spec(c(2:6),width = "2cm") %>%
-  row_spec(3, bold = TRUE) %>%
-  row_spec(7, hline_after = TRUE) %>%
-  add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
-  kable_classic_2()
-
-```
 
@@
 
-## Effects of PM on agricultural productivity, IV estimates
+## Effects of PM on agricultural productivity (IV)
 
@@
 # Table
-colnames(yieldtabletwo) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
+colnames(yield3ivmain) <- c("(1)", "(2)", "(3)")
+rownames(yield3ivmain)[8] <- "F (first stage)"
 kable(
-      yieldtabletwo[-c(3:6, 12, 16:19),],
-      align = "cccc", booktabs = TRUE, linesep = ""
+      yield3ivmain,
+      align = "ccc", booktabs = TRUE, linesep = ""
       ) %>%
-  # footnote(
-  #          "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
-  #          general_title = "",
-  #          threeparttable = TRUE,
-  #          footnote_as_chunk = TRUE
-  #          ) %>%
   column_spec(1, width = "3cm") %>%
-  column_spec(c(2:6),width = "2cm") %>%
-  row_spec(c(3,8), bold = TRUE) %>%
-  row_spec(c(2), hline_after = TRUE) %>%
-  add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
+  column_spec(c(2:4),width = "2cm") %>%
+  row_spec(5, bold = TRUE) %>%
+  row_spec(7, hline_after = TRUE) %>%
+  #add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
   kable_classic_2()
@@
 - Only post coal plant opening
+
+---
+
+## Heterogeneity
+
+```{r heterogeneity, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+# Table
+colnames(yield5heterogeneity) <- c(">p(50)", "<=p(50)", ">p(50)", "<=p(50)", "")
+rownames(yield5heterogeneity)[c(1, 3)] <- c("PM 2.5", "PM times rain")
+kable(
+      yield5heterogeneity,
+      align = "ccccc", booktabs = TRUE, linesep = ""
+      ) %>%
+  column_spec(1, width = "4.5cm") %>%
+  column_spec(c(2:6),width = "1.8cm") %>%
+  row_spec(c(8), hline_after = TRUE) %>%
+  row_spec(c(6), bold = TRUE) %>%
+  add_header_above(c(" " = 1, "wind" = 2, "yield" = 2, " " = 1)) %>%
+  kable_classic_2()
+
+```
 
@@
 
-## Wind direction and labor allocation, by age
+## Wind direction and labor allocation, by rural status
 
-```{r labortableold, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+```{r labortablerural, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
 
 # Table
-labortableold <- labortableold[,-1]
-colnames(labortableold) <- c("all", "self", "wage", "farm", "non-farm")
+colnames(labor1all) <- c("all", "all", "self", "wage", "farm", "non-farm")
+colnames(labor2rural) <- c("all", "all", "self", "wage", "farm", "non-farm")
+
+mat1 <- matrix(c("", "", "", "", "", ""), nrow = 1)
+colnames(mat1) <- c("all", "all", "self", "wage", "farm", "non-farm")
+rownames(mat1) <- c("panel a: all")
+mat2 <- matrix(c("", "", "", "", "", ""), nrow = 1)
+colnames(mat2) <- c("all", "all", "self", "wage", "farm", "non-farm")
+rownames(mat2) <- c("panel b: rural")
+
+labortablerural <- rbind(mat1, labor1all[c(1:2, 7:8),], mat2, labor2rural[-c(3:6),], labor2rural[c(3:6),])
+
 kable(
-      labortableold,
-      align = "ccccc", booktabs = TRUE, linesep = ""
+      labortablerural,
+      align = "cccccc", booktabs = TRUE, linesep = ""
       ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the district level.
-Control variables include female, age, age squared, and (years of) education.
-* p<0.10 ** p<0.05 *** p<0.01",
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
-           ) %>%
-  column_spec(1, width = "4cm") %>%
-  column_spec(c(2:6),width = "1.5cm") %>%
-  row_spec(c(10), hline_after = TRUE) %>%
-  row_spec(c(6,9), bold = TRUE) %>%
+  column_spec(1, width = "3cm") %>%
+  column_spec(c(2:7),width = "1.5cm") %>%
+  row_spec(c(5,10), hline_after = TRUE) %>%
+  row_spec(c(1,6,12), bold = TRUE) %>%
   kable_classic_2()
@@
 
-## Large negative effects on older half
+## One last check
 
-- Older half of sample decreases overall labor allocation
-  - Driven by self-employment and nonfarm employment
-- Younger half actually increases allocation!
-- Consistent with health effects only on older adults, though not conclusive
+- Farm labor most common in July 
+- Have monthly values of wind and pollution...
+  - Heterogeneity by month of the season?
+  - Monsoon season only
 
@@
 
-## Effects on aggregate growth?
+## Heterogeneity by month, labor
 
-- We know pollution affects productivity and labor allocation
-- Might it also affect growth?
-- Proxy for growth with nightlights
-  - Look at previous year's exposure and current year's nightlights
-  - "Growth" of nightlights
-  - Village FE
+```{r labortablemonthly, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+# Table
+colnames(labor3month) <- c("farm", "non-farm")
+labor3month <- labor3month[-c(1:2, 13:14),]
+labor3month <- labor3month[c(1:14),]
+kable(
+      labor3month,
+      align = "cc", booktabs = TRUE, linesep = ""
+      ) %>%
+  column_spec(1, width = "3cm") %>%
+  column_spec(c(2:3),width = "2cm") %>%
+  row_spec(c(14), hline_after = TRUE) %>%
+  row_spec(c(12), bold = TRUE) %>%
+  kable_classic_2()
+
+```
 
@@
 
-## Wind direction and growth of nightlights
+## Heterogeneity by month, yield
+
+```{r yieldtablemonthly, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+# Table
+colnames(yield4monthly) <- c("(1)", "(2)")
+yield4monthly <- yield4monthly[c(1:14),]
+kable(
+      yield4monthly,
+      align = "cc", booktabs = TRUE, linesep = ""
+      ) %>%
+  column_spec(1, width = "3cm") %>%
+  column_spec(c(2:3),width = "2cm") %>%
+  row_spec(12, bold = TRUE) %>%
+  row_spec(14, hline_after = TRUE) %>%
+  kable_classic_2()
+
+```
+
+---
+
+## Coal plants or overall pollution?
+
+- I use the location of coal plants to identify high-pollution areas
+- However, some coal plants open part-way through sample
+- Differences-in-differences combined with IV
+  - Differences in coal pollution, specifically, and other forms of pollution
+
+---
+
+## Coal plants or overall pollution?
 
@@
 # Table
-colnames(ntltable) <- c("(1)", "(2)")
+colnames(yield7ivdiffindiff) <- c("(1)", "(2)", "(3)")
 kable(
-      ntltable,
-      align = "cc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
+      yield7ivdiffindiff, caption = "Agricultural productivity and pollution source",
+      align = "ccc", booktabs = TRUE, linesep = ""
       ) %>%
   column_spec(1, width = "3cm") %>%
-  column_spec(c(2:3),width = "2cm") %>%
-  row_spec(c(9), hline_after = TRUE) %>%
-  row_spec(c(5,8), bold = TRUE) %>%
+  column_spec(c(2:4),width = "2cm") %>%
+  row_spec(c(7), bold = TRUE) %>%
+  row_spec(9, hline_after = TRUE) %>%
   kable_classic_2()
@@
   - Instrumenting for particulate matter, effects are even larger (more than 20 percent)
-- Clear changes in labor allocation
-  - Oldest half of sample
-- Slower growth of nightlights
+- Clear changes in labor allocation at beginning of the season
+  - Yield only changes w.r.t. early-season pollution
 
@@
 - 30km radius indicates relatively long-range effects
+
+---
+
+class: center, middle
+
+<font size = "40"> Thank you! </font>
+
+[https://joshmerfeld.github.io](https://joshmerfeld.github.io) <br>
+Twitter: [@Josh\_Merfeld](twitter.com/Josh_Merfeld)

### 2023-04-07 analyses/3.yield.R

@@ -10,6 +10,7 @@ library(viridis)
 library(fixest)
 library(sf)
 library(openxlsx)
+library(parallel)
 
 
 # Sets wd to folder this script is in so we can create relative paths instead of absolute paths
@@ -134,14 +135,15 @@ summary(df$pm25_abs_dev)
 
 
 
+setFixest_fml(..ctrl1 = ~ rain_z + temp_mean)
+setFixest_fml(..ctrl2 = ~ rain_z + temp_mean + rain_z^2 + temp_mean^2 + rain_z*temp_mean)
+
+
+
 
 # Regressions ---------------------------------
 
 ## reduced form ---------------------------------
-
-setFixest_fml(..ctrl1 = ~ rain_z + temp_mean)
-setFixest_fml(..ctrl2 = ~ rain_z + temp_mean + rain_z^2 + temp_mean^2 + rain_z*temp_mean)
-
 
 yield1 <- feols(log(yield) ~ wind | shrid + year, data = df, cluster = c("shrid"))
 yield2 <- feols(log(yield) ~ wind + ..ctrl1 | shrid + year, data = df, cluster = c("shrid"))
@@ -165,6 +167,161 @@ rownames(yieldtable) <- c("wind days", " ", "rain (z)", "", "mean temp (10s)", "
                           "fixed effects:", "village", "year",
                           "observations")
 saveRDS(yieldtable, "pollution_development/draft/tables/yield1reducedform.rds")
+
+
+
+
+
+## placebo test
+wind_year <- df %>%
+              dplyr::select(wind, year) %>%
+              arrange(year) %>%
+              mutate(row = row_number()) %>% # row number in entire thing
+              group_by(year) %>%
+              mutate(
+                     row_min = min(row), # min row number (in entire column)
+                     row_max = max(row) # max
+                     ) %>%
+              ungroup() %>%
+              arrange(year)
+df_placebo <- df %>%
+              dplyr::select(-wind) %>%
+              arrange(year)
+
+samplefrom <- cbind(wind_year$row_min, wind_year$row_max)
+
+
+# create function to sample
+sample_pre_fun <- function(a) {
+  sample(x = min(a):max(a), size = 1, replace = T)
+}
+
+# now wrap that function in apply
+sample_fun <- function(a) {
+  sample_vec <- apply(a, 1, sample_pre_fun)
+}
+
+f <- function(i) {
+  # create vector with random sampling
[…]

### 2023-04-07 analyses/4.nss.R

@@ -33,23 +33,20 @@ df <- df %>% mutate(
                     wind = days_sum,
                     district = distfe
                     )
-pollution <- read_csv(paste0("data/clean/pm25/nss_merge.csv"))
-pollution <- pollution %>% mutate(
-                                  temp_mean = (month_tmin + month_tmax)/200,
-                                  season_temp_mean = (season_tmin + season_tmax)/200,
-                                  month_precip = log(month_precip + 1),
-                                  season_precip = log(season_precip + 1),
-                                  pm25 = log(pm25)
-                                  )
-df <- df %>% left_join(pollution, by = c("state_merge", "district_merge", "year", "month_int"))
 
 # MONSOON ONLY
 df <- df %>% filter(month_int %in% c(5,6,7,8,9,10,11))
+# create seasons means
+df <- df %>% 
+      group_by(year, district) %>%
+      mutate(temp_mean = mean((month_tmin + month_tmax)/2),
+             precip_sum = sum(month_precip)) %>%
+      ungroup()
+
 
 # control variables
 # NOTE: "season' variables are defined as the season UP TO THAT POINT, so interact with month_int factor
-setFixest_fml(..ctrl = ~ poly(female, 1) + poly(age, 2) + educ + temp_mean*month_precip + temp_mean^2 + month_precip^2 + 
-                as_factor(month_int)*season_temp_mean*season_precip + as_factor(month_int)*season_temp_mean^2 + as_factor(month_int)*season_precip^2)
+setFixest_fml(..ctrl = ~ poly(female, 1) + poly(age, 2) + educ + temp_mean*precip_sum + month_precip + temp_mean^2 + precip_sum^2)
 
 
 
@@ -201,10 +198,10 @@ df <- df %>% mutate(
   wind6 = wind*(month_int==11)
 )
 
-labor1 <- feols(days_f ~ wind0 + wind1 + wind2 + wind3 + wind4 + wind5 + wind6 + ..ctrl | district + month,
+labor1 <- feols(days_f ~ wind1 + wind2 + wind3 + wind4 + wind5 + ..ctrl | district + month,
                 data = df %>% filter(rural==1),
                 cluster = "district")
-labor2 <- feols(days_nf ~ wind0 + wind1 + wind2 + wind3 + wind4 + wind5 + wind6 + ..ctrl | district + month,
+labor2 <- feols(days_nf ~ wind1 + wind2 + wind3 + wind4 + wind5 + ..ctrl | district + month,
[…]

### 2023-04-07 cleaning/5.cleaning_precip_temp.R

@@ -23,6 +23,18 @@ setwd("../../")
 getwd()    # check
 
 
+# districts
+villages <- read_sf("data/spatial/shapefiles/village.shp")
+# create district shapefile
+villages <- villages %>% 
+  mutate(district = paste0(pc11_s_id, "-", pc11_d_id))
+villages_spatial <- as_Spatial(villages)
+
+districts <- gUnaryUnion(villages_spatial, id = villages_spatial@data$district)
+districts <- st_as_sf(districts)
+districts <- cbind(districts, unique(villages$district))
+colnames(districts)[1] <- c("district_id")
+
 # villages
 villages_centroids <- read_sf("data/spatial/centroids_villages_30km")
 villages_centroids <- as_tibble(villages_centroids) %>% dplyr::select(shrid)
@@ -32,14 +44,12 @@ villages$shrid <- paste0(villages$pc11_s_id, "-", villages$pc11_tv_id)
 villages <- villages %>% left_join(villages_centroids, by = "shrid")
 villages <- villages %>% filter(within30==1) %>% dplyr::select(shrid)
 
-# districts
-districts <- read_sf("data/spatial/districts/districts.shp")
 
 
 # Always using this
 url <- "thredds.northwestknowledge.net:8080/thredds/fileServer/TERRACLIMATE_ALL/data/"
-# PRECIP --------------------------------------------------------------------
 for (year in 1990:2020){
+  # PRECIP --------------------------------------------------------------------
   # This changes by year
   file <- paste0("TerraClimate_ppt_", year, ".nc")
   curl::curl_download(file.path(url, file), file)
@@ -48,13 +58,11 @@ for (year in 1990:2020){
   extracted_values <- exact_extract(r, villages, fun = "mean", append_cols = "shrid")
   write_csv(extracted_values, paste0("data/clean/terra/precip", year, ".csv"))
   # districts
-  extracted_values <- exact_extract(r, districts, fun = "mean", append_cols = c("ST_CEN_CD", "DT_CEN_CD"))
+  extracted_values <- exact_extract(r, districts, fun = "mean", append_cols = c("district_id"))
   write_csv(extracted_values, paste0("data/clean/terra_district/precip", year, ".csv"))
   file.remove(file)
-  print(year)
-}
-# MAX TEMP --------------------------------------------------------------------
-for (year in 1990:2020){
+
[…]

### 2023-04-07 cleaning/6.cleaning_pm25.R

@@ -12,6 +12,7 @@ library(raster)
 library(ncdf4)
 library(lubridate)
 library(exactextractr)  
+library(rgeos)
 
 
 # Sets wd to folder this script is in so we can create relative paths instead of absolute paths
@@ -53,10 +54,10 @@ for (year in 1998:2015){
     # extract
     extracted_values <- exact_extract(r, villages, fun = "mean", append_cols = "shrid")
     extracted_values <- as_tibble(extracted_values) %>% 
-                        group_by(shrid) %>% 
-                        mutate(mean = mean(mean)) %>% 
-                        filter(row_number()==1) %>%
-                        ungroup()
+      group_by(shrid) %>% 
+      mutate(mean = mean(mean)) %>% 
+      filter(row_number()==1) %>%
+      ungroup()
     
     # And wind direction
     village_wind <- read_csv(paste0("data/clean/wind_ntl/months/y", year, "m", as.numeric(month), ".csv")) %>% as_tibble()
@@ -103,19 +104,19 @@ pollution$m5[pollution$month==3] <- pollution$mean[pollution$month==3]
 
 # Now shrid, year, season
 pollution <- pollution %>% 
-                group_by(shrid, year, season) %>% 
-                mutate(
-                       pm25 = mean(mean, na.rm = T),
-                       pm1 = mean(m1, na.rm = T),
-                       pm2 = mean(m2, na.rm = T),
-                       pm3 = mean(m3, na.rm = T),
-                       pm4 = mean(m4, na.rm = T),
-                       pm5 = mean(m5, na.rm = T)
-                       ) %>%
-                filter(row_number()==1) %>%
-                ungroup() %>%
-                dplyr::select(shrid, year, season, pm25,
-                              pm1, pm2, pm3, pm4, pm5)
+  group_by(shrid, year, season) %>% 
+  mutate(
+    pm25 = mean(mean, na.rm = T),
+    pm1 = mean(m1, na.rm = T),
+    pm2 = mean(m2, na.rm = T),
+    pm3 = mean(m3, na.rm = T),
+    pm4 = mean(m4, na.rm = T),
+    pm5 = mean(m5, na.rm = T)
+  ) %>%
+  filter(row_number()==1) %>%
+  ungroup() %>%
+  dplyr::select(shrid, year, season, pm25,
+                pm1, pm2, pm3, pm4, pm5)
 
 write_csv(pollution, paste0("data/clean/pm25/ag_merge.csv"))
 
[…]

### 2023-04-07 cleaning/8.cleaning_overlap_merging_nss.R

@@ -7,6 +7,7 @@ rm(list = ls())
 library(tidyverse)
 library(lubridate)
 library(readxl)
+library(rgeos)
 library(sf)
 
 
@@ -22,9 +23,12 @@ getwd()    # check
 # Village overlap data
 villages <- read_sf(paste0("data/spatial/villages_overlap/villages_overlap.shp"))
 villages <- villages %>% mutate(shrid = paste0(pc11_s_id, "-", pc11_tv_id),
+                                district_id = paste0(pc11_s_id, "-", pc11_d_id),
                                 area_overlap = area,
                                 area_district = exp(area_2)) %>%
-                          dplyr::select(shrid, state = ST_CEN_CD, district = DT_CEN_CD, state_name = ST_NM, district_name = DISTRICT, area_overlap, area_district)
+                          dplyr::select(shrid, state = ST_CEN_CD, district = DT_CEN_CD, state_name = ST_NM, district_name = DISTRICT, area_overlap, area_district, district_id)
+
+# now clean villages more 
 villages <- as_tibble(villages) %>% dplyr::select(-geometry)
 villages <- villages %>% 
               group_by(state, district) %>% 
@@ -36,6 +40,7 @@ villages <- villages %>%
 villages$area_weight[is.na(villages$area_weight)==T] <- villages$area_weight_alt[is.na(villages$area_weight)==T]
 villages$area_weight[villages$area_weight>villages$area_weight_alt] <- villages$area_weight_alt[villages$area_weight>villages$area_weight_alt]
 
+              
 
 
 # WEEKLY --------------------------------------------------------------------------------
@@ -68,9 +73,9 @@ df_labor <- df_labor %>% mutate(state_merge = as.numeric(state_merge),
 df_labor <- df_labor %>% filter(is.na(merge)==F)
 
 
+# district pollution
+pollution <- as_tibble(read_csv("data/clean/pm25_district/all_combined_district.csv"))
 
-pollution <- read_csv(paste0("data/clean/pm25/all_combined.csv"))
-pollution <- pollution %>% rename(pm25 = mean)
 
 # Going to go through the dates and process date by date
 df_labor_merged <- c()
@@ -88,9 +93,9 @@ for (day in 1:length(date_vec)){
   # And also pm
   pollution_merge <- pollution %>% filter(year==year(date_vec[day]) & month==month(date_vec[day]))
   # there are a couple mistakes in terms of duplicate ids. fix those
[…]

### 2023-04-11 draft/draft.Rmd
vastly updated draft

@@
 ---
-title: "Blowing in the wind: Pollution and agricultural productivity in India^[I thank Ariel Ortiz-Bobea, Marc Bellemare, and David Sungho Park for comments and suggestions that greatly improved the quality of the paper.
+title: "Air pollution and agricultural productivity in a developing country^[I thank Ariel Ortiz-Bobea, Marc Bellemare, and David Sungho Park for comments and suggestions that greatly improved the quality of the paper.
 The usual disclaimer applies.
@@
 abstract: |
-  | \noindent Pollution is a growing problem in many developing countries.
-Using the rollout of coal plants to identify high-pollution locations in India and random variation induced by changes in wind direction, I show that pollution has large, negative impacts on agricultural productivity.
-Results indicate that some of this impact is driven by changes in labor allocation patterns in response to increased exposure to pollution.
-I also show that pollution from coal plants, specifically, has a larger effect on agricultural productivity than other types of pollution.
+  | \noindent I document large negative externalities in the agricultural sector of India due to air pollution.
+Using random variation in pollution exposure induced by changes in wind direction, I show that higher levels of particulate matter lead to decreased agricultural productivity.
+Based on within-village average deviations in pollution exposure, year-to-year changes in productivity can be above 10 percent.
+Combining wind direction with the rollout of coal plants, I show that pollution from coal plants, specifically, has a larger effect on agricultural productivity than other types of pollution.
+Given that the agricultural sector is a refuge for the poor in many developing countries, these results suggest that the effects of pollution may hit the poorest particularly hard.
   |
@@
 yield7ivdiffindiff <- readRDS("pollution_development/draft/tables/yield7ivdiffindiff.rds")
+yield8cumulative <- readRDS("pollution_development/draft/tables/yield8cumulative.rds")
 labor1all <- readRDS("pollution_development/draft/tables/labor1all.rds")
@@
 true_value <- readRDS("pollution_development/draft/tables/randomization_true_value.rds")
+coef_current <- readRDS("pollution_development/draft/tables/coef_current.rds")
+coef_lead <- readRDS("pollution_development/draft/tables/coef_lead.rds")
 
@@
 
-Coal is one of the most common energy sources throughout the world.
-In 2021, energy generation from coal reached an all-time high,[^1] driven in part by rapid population growth and rising energy prices.
-India is no exception to these trends.
-From 1990 to 2010, India's population grew by more than 40 percent.[^2] This population growth led to a large increase in demand for power.
-While natural gas has become a popular alternative in much of the world, the Indian government met much of the increased demand through the construction of coal power plants.
-During those same two decades, power generation from all coal units of at least 30 megawatts in the country increased by almost 140 percent, from 42.4 gigawatts to more than 100 gigawatts.[^3] There are no doubt positive effects from this increase in power generation; for example, electricity may lead to economic and productivity growth [@dinkelman2011effects; @kline2014local; @rud2012electricity; @van2017long], though electrification alone may not be a sufficient condition [@lee2020does].
+While reliance on coal is declining in much of the world, it continues to be an important source of energy in many countries.
+In 2021, energy generation from coal reached an all-time high,[^1] driven in part by rapid population growth in developing countries, including India.
+From 1990 to 2010, India's population grew by more than 40 percent, leading to a large increase in demand for power.[^2] While natural gas has become a popular alternative in much of the world, the Indian government met much of the increased demand through the construction of coal power plants.
+During those same two decades, power generation from all coal units of at least 30 megawatts in the country more than doubled, from 42.4 gigawatts to more than 100 gigawatts.[^3]
 
@@
 
+There are no doubt positive effects from this increase in power generation; for example, electricity may lead to economic and productivity growth [@dinkelman2011effects; @kline2014local; @rud2012electricity; @van2017long], though electrification alone may not be a sufficient condition [@lee2020does].
 These improvements do not come without downsides, however, especially when it comes to power derived from coal.
@@
 
-I study the effects of emissions from coal power plants on agricultural productivity in India.
+In addition to coal, pollution more generally is increasing in much of the world.
+Carbon dioxide equivalent per capita has almost double over the last 50 years in India, more than doubled in Indonesia, and almost tripled in China.^[https://edgar.jrc.ec.europa.eu] Of course, much of this increase is due to the development process itself; as countries develop, their economies transform, with more people moving out of agriculture and into manufacturing and services [@herrendorf2014growth].
+Despite this development, agriculture remains a key sector in many developing countries, employing a large proportion of the poor [@dercon2009rural].
+This also means agriculture is a key sector when it comes to poverty reduction [@christiaensen2011evolving].
+Given the large effects of pollution on health and previously documented impacts on agriculture, this raises a important question: do rising pollution levels in developing countries threaten agricultural productivity?
+
+In this paper, I study the effects of pollution on agricultural productivity in India.
 Previous work has suggested that there may be harmful direct effects of air pollution [@heck1982assessment; @marshall1997hidden] and water pollution [@reddy2006impact] on agricultural productivity, including work showing that gold mines -- which release large amounts of pollutants into the surrounding environment -- lead to lower agricultural productivity in Ghana [@aragon2016polluting].
-These effects can work through the well known effects on human health and productivity, but also directly through land productivity.
+These effects can work through the well-known effects on human health and productivity, but also directly through land productivity.
 Black carbon in soot, for example, can directly affect solar radiation reaching plants on the surface [@ramanathan2008global; @burney2014recent].
-Relative to previous work in economics, this paper tackles the broader case of coal plants, which are a leading source of power throughout the world.
-I show that air pollution from these plants can lead to lower agricultural productivity and even decreased economic growth.
+Relative to previous work in economics, this paper tackles pollution more broadly, before turning to a specific source of pollution: coal plants.
+I show that air pollution leads to lower agricultural productivity in India and that pollution from coal plants is particularly harmful to agriculture.
 
@@
 On the one hand, governments may locate new plants in fast-growing areas, while on the other hand, they may decide the exact location of plants based on the political power of local citizens [@kopas2020environmental].
-In the case of the data used here, coal plants are constructed in areas with higher population and higher literacy (Table \@ref(tab:plantresultstable)).
+In the case of the data I use here, coal plants are constructed in areas with higher population and higher literacy.
 In terms of agriculture, higher productivity predicts more coal plants in a given year as well as the construction of a coal plant over the next decade, indicating possible bias when estimating OLS regressions of agricultural productivity on pollution levels.
 
-To identify the effects of air pollution on agricultural productivity, I instead combine the roll out of coal plants in India with plausibly exogenous changes in wind direction -- in a similar spirit to @deryugina2019mortality -- to measure the effects of pollution from coal plants on key development outcomes.
+To identify the effects of air pollution on agricultural productivity, I identify areas of high pollution emission using the location of coal plants.
+While coal plants may add emissions, I show that they are located in areas with relatively higher levels of pollution, in general.
+Using these locations as point sources of pollution, I use plausibly exogenous changes in wind direction -- in a similar spirit to @deryugina2019mortality -- to measure the effects of pollution on key agricultural outcomes.
 I first plot the location of coal plants in India by year.
 Using a village-level shapefile, I identify all villages that are located within 30km of a coal plant at any time during the sample period, which comprises the 2000s to the 2010s, with differences across outcomes depending on data availability.
-
-I use the location of coal plants to identify high-pollution areas and calculate the direction from each coal plant to all villages within 30km.
-I include all areas, regardless of whether the coal plant has opened yet, to allow me to disaggregate the effects of pollution from coal plants with other kinds of pollution.
-Using daily wind data and these spatial variables, I create an identifier separately for each of the one-hundred-thousand-plus villages for whether a given village is located downwind from a coal plant on any given day.
+I then calculate the direction from each coal plant to all villages within 30km.
+I include all areas, regardless of whether the coal plant has opened yet, to allow me to disaggregate the effects of pollution from coal plants with other kinds of pollution later in the paper.
+Using daily wind data and these spatial variables, I create an identifier separately for every village for whether a given village is located downwind from a coal plant on any given day.
 I then aggregate this exposure variable based on the desired analysis; for agricultural productivity, for example, I create an exposure variable measuring the total number of exposure days in the five-month agricultural season, focusing only on the monsoon (Kharif) season.
-Consistent with expectations, wind direction is a strong predictor of particulate matter at the monthly level.
-More days being located downwind from a coal power plant leads to higher average concentrations of particulate matter 2.5.
-
-Using newly available estimates of India-wide agricultural (land) productivity, results show that the total number of days of being downwind from a coal power plant has a negative effect on overall agricultural productivity.
-Just a single additional day of being downwind from a coal power plant leads to a decrease in predicted yields of 0.6 percent.
-Combined with the within-village absolute deviation of wind exposure, results indicate that median changes in wind direction are responsible for fluctuations in agricultural productivity of around 1.5 percent.
-
-Using wind direction as an instrument for particulate matter, an increase of PM 2.5 by one percent leads to an even larger decrease in agricultural productivity, at around 0.73 percent.
-Given within-village changes across years, pollution could be responsible for swings of 7-to-14 percent in agricultural productivity across years.
-This IV strategy is similar to that in @deryugina2019mortality, though they use general "high-pollution" wind directions while I concentrate only on wind directions from the location of a coal power plant.
-The IV estimates indicate that the effect of pollution is more negative than naïve OLS estimates indicate, which is consistent with the evidence on the construction of coal plants; higher levels of pollution are correlated with higher levels of agricultural productivity and economic growth, a relationship that the IV strategy adjusts downwards.
-
-These effects are apparently larger in villages that experience higher median exposure to pollution and in villages with higher initial yields.
-I also show that rainfall shocks are compounding; higher exposure along with less rainfall increases the negative effect of either shock individually.
-As weather shocks become more common across the globe, the rise of pollution in India may be particularly problematic for farmers.
-
-To explore the mechanisms driving this effect, I also use data on labor allocation from the National Sample Survey (NSS).
+More days being located downwind from a high-pollution location leads to substantially higher average concentrations of particulate matter 2.5 (PM 2.5).^[PM 2.5 is defined as particulate matter with a diameter of less than 2.5 micrometers.]
+
+Using wind direction as an instrument for particulate matter, an increase of PM 2.5 by one percent leads to a large decrease in agricultural productivity, at around 0.73 percent.
+Given within-village changes in exposure, pollution could be responsible for swings of 7-to-14 percent in agricultural productivity across years.
+The IV estimates indicate that the effect of pollution is more negative than naïve OLS estimates, which is consistent with the evidence on the construction of coal plants; higher levels of pollution are correlated with higher levels of agricultural productivity, a relationship that the IV strategy adjusts downwards.
+Robustness checks confirm these main results.
+Randomization inference shows that the estimates are robust to alternative measures of uncertainty, while leads of pollution do not affect current agricultural productivity.
+
+Heterogeneity analysis shows that the negative effects of pollution are larger in villages that experience higher median exposure to pollution and in villages with higher initial yields.
+Finally, given how many poor farmers rely on rainfall and how weather shocks increase across the globe, I also look at possible multiplicative effects of shocks by interacting the pollution variables with a rainfall variable.
+Higher exposure along with less rainfall increases the negative effect of either shock individually, meaning these shocks are compounding.
+Given concerns regarding agricultural productivity and climate change, the rise of pollution in India may be particularly problematic for poor farmers.
+
+I find that these effects are mostly transmitted by reductions in farm labor during narrow windows of the growing season, using labor allocation data from the National Sample Survey (NSS).
 While there do not appear to be large changes in labor allocation patterns when looking at the season as a whole, there is a clear pattern when we break down exposure to pollution by month: higher levels of pollution at the beginning of the monsoon season (June and July) lead to lower levels of farm labor allocation, but no changes in non-farm labor allocation.
 This matches up well with a similar exercise with agricultural yields: pollution in June is the only month in which there is a negative effect of pollution on agricultural yield.
-In tandem, these two results point to effects on labor as one mechanism driving the total effects, though I cannot rule out direct effects of pollution on land productivity.
-
-The final set of results tries to disentangle the effects of pollution from the effects of coal pollution, specifically.
+In tandem, these two results point to effects on labor as one mechanism driving the total effects, though I cannot rule out direct effects of pollution on land productivity with my data.
+
+The final set of analyses tries to disentangle the effects of pollution from the effects of coal pollution, specifically, by using wind direction and the rollout of coal plants across the country.
 Combining instrumental variables with a differences-in-differences strategy, I show that agricultural productivity has an elasticity of around -0.15 with respect to non-coal pollution.
-The opening of a coal plant, however, triples the elasticity, to more than -0.45, indicating that coal plant emissions are particularly harmful to agricultural productivity.
+The opening of a coal plant, however, triples the elasticity, to more than -0.45, indicating that coal plant emissions are particularly harmful to agricultural productivity, at least in India.
 This is consistent with coal emissions being quite different from other forms of pollution; for example, coal emissions contain much more mercury than other forms of pollution.
@@
 Perhaps relevant to my labor allocation results, @hanna2015effect show that decreases in pollution lead to increases in labor allocation, while I show the same relationship in at least a subset of months of the agricultural season.
-The papers most related to this one are @aragon2016polluting, who show that gold mines in Ghana have substantial negative impacts on agricultural productivity in the surrounding areas, and @burney2014recent, who model the effects of climate change and pollutants on agricultural productivity in India.
-This paper differs in two important ways.
-First, I explore effects of pollution from coal plants, which are much more ubiquitous than gold mines across the world.
+Several papers are closely related to this one.
+@aragon2016polluting show that gold mines in Ghana have substantial negative impacts on agricultural productivity in the surrounding areas.
+@burney2014recent model the effects of climate change and pollutants on agricultural productivity in India.
+Finally, @sanders2022adaptation actually find that coal emissions can be good for agriculture in the United States, due to the presence of sulfur, which can increase agricultural productivity.
+This paper differs in three important ways.
+First, I explore effects of pollution more generally as well as from coal plants, which are much more ubiquitous than gold mines across the world.
 Second, I use a different identification strategy from either paper, relying on wind blowing pollution from coal plants.
-
-@sanders2022adaptation
+Finally, agriculture in India and the United States can be quite different, with the latter being much more mechanized.
+This means effects of pollution could differ substantially.
 
@@
 
-# Data and methods
+# Data
 
 \label{data}
-
-In this section, I describe the different data sources and the identification strategy.
-I also spend time validating the use of wind direction as a proxy for particulate matter as well as looking at the correlates of construction of coal power plants.
-
-## Data
-
-\label{data2}
 
@@
 
-## Identification
+# Methods
 
@@
 So what are the characteristics of areas near the construction of new coal plants?
-Table \@ref(tab:plantresultstable) presents four separate regressions.
+Table \@ref(tab:plantresultstable) presents four separate regressions, using state-level fixed effects due to the spatial variation in coal plants.
 The first two columns use 1991 census data -- downloaded from SHRUG -- and the second two columns use 2001 census data, all at the village level.
@@
 colnames(plantresultstable) <- c("1991", "2001", "2001", "2011")
-rownames(plantresultstable) <- c("Population (log)", "", "Area (log)", "", "Literacy (prop)", "", "Ag. productivity", "", 
+rownames(plantresultstable) <- c("Population (log)", "", "Area (log)", "", "Literacy (prop)", "", "Ag. productivity", "", "Pollution", "(log PM 2.5)",
                                 "Sub-sample", "Observations")
@@
   footnote(
-           "Note: Robust standard errors are in parentheses.
+           "Note: State fixed effects are included in regressions.
+Standard errors are in parentheses and clustered at the state level.
 The outcome in the first column is whether the village is within 30km of a coal plant in 1991.
@@
   column_spec(c(2:5),width = "2cm") %>%
-  row_spec(9, hline_after = TRUE) %>%
+  row_spec(11, hline_after = TRUE) %>%
   add_header_above(c(" ", "1991 census" = 2, "2001 census" = 2)) %>%
@@
 They present similar results to those in the first two columns, but use the 2001 census data and a dummy for the presence of a coal plant in 2001 and 2011 in columns three and four, respectively.
-Importantly, I include agricultural productivity in the monsoon season of 2002 -- the earliest year available -- as an additional predictor.
+Importantly, I include agricultural productivity in the monsoon season of 2002 -- the earliest year available -- as well as average pollution during the monsoon season for 1998 through 2000 as additional predictors.
 In 2001, agricultural productivity is a strong predictor for the presence of a coal plant.
@@
 However, note that this is not the same as looking at wind direction, though comparing the OLS and IV results below leads to the same conclusion regarding selection.
+Pollution, on the other hand, is not significant but is positive in both columns (and is also logged, just like productivity).
 
@@
 Consider the data used for the regressions presented in Table \@ref(tab:pollutiontable) of the appendix.
-The outcome variable is particulate matter -- specifically, $\mathrm{PM_{2.5}}$, which is particulate matter no larger than 2.5 micrometers in diameter -- which comes from @hammer2020global.
-$\mathrm{PM_{2.5}}$ is one of the harmful byproducts from coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.[^12] This data is available at the monthly level, so in Table \@ref(tab:pollutiontable), I aggregate wind exposure to the monthly level, as well.
+The outcome variable is particulate matter -- specifically, $\mathrm{PM 2.5}$, which is particulate matter no larger than 2.5 micrometers in diameter -- which comes from @hammer2020global.
+$\mathrm{PM 2.5}$ is one of the harmful byproducts from coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.[^12] This data is available at the monthly level, so in Table \@ref(tab:pollutiontable), I aggregate wind exposure to the monthly level, as well.
 
@@
 The first two columns include all available years of data while the last two columns restrict estimation only to the years for which the agricultural productivity regressions are estimated.
-Across all four columns, the story is the same: when the wind is blowing in the direction of a village from a coal plant within 30 kilometers, estimated $\mathrm{PM_{2.5}}$ is substantially higher.
-
-To put these numbers into perspective, the World Health Organization's updated guidelines are that average annual exposure should not exceed 15 $\mathrm{\mu g/m^3}$.
+Across all four columns, the story is the same: when the wind is blowing in the direction of a village from a coal plant within 30 kilometers, estimated $\mathrm{PM 2.5}$ is substantially higher.
+
+To put context behind pollution numbers, the World Health Organization's updated guidelines are that average annual exposure should not exceed 15 $\mathrm{\mu g/m^3}$.
 For the month, an increase at the midpoint of the estimated range is equal to an increase of 0.2 percent of the maximum recommended mean concentration from the WHO.
@@
 
-### Identifying effects of pollution from coal plants
+## Identifying effects of pollution from coal plants
 
@@
 \label{results}
-
-This section presents the main results of this paper. I present two main sets of results: one set looking at agricultural productivity and another looking at labor allocation.
-
-## Agricultural productivity
 
@@
 Instead, we are actually interested in better understanding the effects of exposure to pollution on agricultural productivity.
-Table \@ref(tab:yieldtablepm) presents the same regressions as Table \@ref(tab:yieldtable), with a key difference: instead of wind exposure, it uses log average concentration of $\mathrm{PM_{2.5}}$ -- from @hammer2020global -- during the season as the key explanatory variable.
+Table \@ref(tab:yieldtablepm) presents the same regressions as Table \@ref(tab:yieldtable), with a key difference: instead of wind exposure, it uses log average concentration of $\mathrm{PM 2.5}$ -- from @hammer2020global -- during the season as the key explanatory variable.
 There is again a consistent, negative relationship between particulate matter and agricultural productivity.
@@
 
-The second-stage results first indicate that $\mathrm{PM_{2.5}}$ is a very strong predictor of agricultural productivity.
+The second-stage results first indicate that $\mathrm{PM 2.5}$ is a very strong predictor of agricultural productivity.
 Using the results in column three, a one percent increase in particulate matter leads to a 0.73 percent decrease in agricultural productivity.
@@
 ```
-
-## Robustness checks
 
@@
 
+## Robustness checks
+
+In this section, I report the results of two separate robustness checks for the main results presented above.
+In the previous results, standard errors are clustered at the level of the village, which is also the level at which treatment (exposure to pollution) is assigned as well as the level at which agricultural productivity is extracted.
+Past work has shown that standard errors might be underestimated in such contexts [@bertrand2004much].
+This is a particular concern in the present case, since agricultural productivity is extracted at the village level but estimated using less coarse data that has high spatial correlation [@gangopadhyay2022new].
+
+To help remedy concerns, I perform randomization inference (e.g.
+@barrios2012clustering) by randomly assigning observed treatment across villages and repeating this process 1,000 times.
+I do this in two separate ways.
+First, I take all villages in a given year and randomly assign, with replacement, treatment intensity within the year.
+Second, I do this separately by selecting from within district-years only, to help take into account district-year correlation in outcomes and as a way to mitigate possible concerns that traditional clustered standard errors may actually be much too conservative [@abadie2023should].
+Since this is an instrumental variables procedure, I randomly select the wind and pollution variable _together_.
+
+I present these results in Figure \@ref(fig:randomplot).
+The purple density is the resulting distribution of estimated treatment effects when selecting from all villages within the same year.
+The density of treatment effects is relatively tight around 0, with the left tail only getting slightly beyond -0.05, compared to the most conservative actual estimated treatment effect of -0.73.
+
+The within-district correlation in outcomes and treatment is apparent when instead randomly selecting from only villages within the same district in the same year.
+The distribution is still approximately normal but has a mode of almost -0.4.
+Nonetheless, the actual treatment effect is larger in magnitude than every single one of the 1,000 bootstrapped replications.
+If we were to use the standard deviation of the resulting distribution as a standard error, we would get a confidence interval around the main effect of approximately `r paste0(reportDec(-0.73-1.96*sd(coefs_distfe)))` to `r paste0(reportDec(-0.73+1.96*sd(coefs_distfe)))`.
+The results are apparently robust to alternative specifications of uncertainty
+
+The second robustness check includes leads of pollution and wind direction.
+I present these results in Table \@ref(tab:yieldtableleads).
+Interestingly, there is actually a positive coefficient for the lead in the second column.
+However, if we perform randomization inference, it appears that the confidence intervals might be slightly underestimated; the positive effect size is in the 11th percentile of the distribution in Figure \@ref(fig:randomplotleads), which for a two-tailed test would not even have a p-value of less than 0.2.
+Overall, there is no indication that leads predict any changes in current agricultural productivity.
+
 ```{r yieldtableleads, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
@@
 One possible explanation is that exposure is cumulative, leading to larger and larger effects.
-I return to this in column five.
+I return to the effects of cumulative shocks in column five.
 
@@
 The largest effects are apparently concentrated in the upper half of the distribution, meaning the most productive areas are also most susceptible to the effects of pollution.
+This could have important implications for the overall effects of the placement of any point sources of pollution.
 
@@
 Another way to think about this, however, is that higher values of rainfall dampen the effects of pollution shocks.
-This could have very profound effects on welfare, since this essentially increases the variance of agricultural productivity across years (which is already increasing due to climate change).
-
-In the appendix, I present in Table \@ref(tab:yieldtableleads) as a robustness check similar regressions but also including wind/pollution leads.
-The negative effect is only seen with current pollution, not future pollution.
+This could have very profound effects on welfare, since this essentially increases the variance of agricultural productivity across years (which is already increasing due to climate change), at least as it relates to pollution exposure..
 
@@
 Table \@ref(tab:labortable) presents results for labor allocation over the last seven days using the National Sample Survey and the methodology described above.
-Since the labor recall is for seven days, labor allocation ranges from zero to seven, as does the wind variable.
-Recall that the data is at the district level, not the village level.
+Since the labor recall is for seven days, labor allocation ranges from zero to seven, as does the wind variable.^[I also tried an instrumental variables strategy with the seven-day wind and monthly pollution, but the first stage lacked power due to the difference in temporal aggregation across the two key variables (pollution and wind).] Recall that the data is at the district level, not the village level.
 However, it is also more temporally disaggregated, with rolling interviews throughout the year-long survey.
@@
 
-While overall results for labor allocation are inconclusive, perhaps we can learn more by looking at temporally disaggregated results, focusing on May through December.
-Although the yield data was created by looking at five specific months of the year -- June through October -- this is partially because this is the time of year when remote-sensing data is most able to pick up overall changes in crop cover.
-Preparation, however, starts before that, with operations such as ploughing often taking place prior to June.
+While overall results for labor allocation are inconclusive, perhaps we can learn more by looking at temporally disaggregated results, focusing on June through October.
+I can create separate variables for exposure by each month of the season, keeping the unit of analysis the individual (for labor) or the village-season (for yields).
 
 Since the NSS data has survey dates and the both the pollution and wind data can be aggregated to the month, we can look at how exposure during different months of the agricultural season (differentially) affect labor allocation and overall agricultural productivity.
-I begin with labor allocation in Table \@ref(tab:labor3month).
+I begin with labor allocation in columns one and two of Table \@ref(tab:labor3month).
 As before, the exposure measure is just the wind variable (over the last seven days), since I can directly match daily wind direction with the timing of the survey, but I cannot do that for pollution, which is monthly.
@@
 
-I perform a similar exercise with yield, the results of which are in Table \@ref(tab:labor3month).
+I perform a similar exercise with yield, the results of which are in the third column Table \@ref(tab:labor3month).
+I return to the IV strategy, however, to be consistent with the previous yield results.
 I include only the five months of the year that are used in the calculation of the yield estimates, or June through October.
-The results are again from two-stage least squares, with the F-statistics for each month presented at the bottom; monthly wind remains a strong predictor of pollution.
-Column one presents results without controls for weather while column two presents results with controls for weather; the differences in coefficients make clear how important the inclusion of weather is here.
-
-Focusing on column two, there is only one month where we see a clear effect of pollution on yield: June.
-Given that we also see negative effects of pollution on farm labor allocation in June, this is suggestive -- but not conclusive -- evidence that pollution works through labor.
-Note that this does not rule out pollution also having direct effects on agricultural productivity, however.
+
+There is only one month where we see a clear effect of pollution on yield: June.
+Given that we also see negative effects of pollution on farm labor allocation in June, this is suggestive -- but not conclusive -- evidence that pollution at least partly works through labor.
+Note that this does not rule out pollution also having direct effects on agricultural land productivity, however.
 
@@
 For example, yield might respond strongly to average pollution in June and July, which would not be captured in Table \@ref(tab:labor3month).
+As one additional check, Table \@ref(tab:yield8cumulative) in the appendix presents five separate instrumental variable regressions, which look at the effects of cumulative exposure to pollution separately by month.
+Across all five months, cumulative exposure negatively affects agricultural productivity, with the largest coefficient actually showing up for October, the last month of the season.
 
@@
 Column three presents the most extensive set of controls and I focus on that column here.
-First, there are now two first stages but the F-test indicates that the instruments are strong for both, including the interaction term (F = 232).
-Second, the first row shows the effect of pollution on agricultural productivity prior to the opening of a coal plant; the coefficient is negative but relatively smaller.
-Third, the interaction term is strongly negative and more than twice as large in magnitude as the level term, indicating a very clear effect specific to coal plants.
-This could be due to differences in emission patterns for different kinds of plants and factors; coal plants, for example, are responsible for much more mercury than other kinds of polluting industries.^[https://www.epa.gov/sites/default/files/2016-12/documents/nei2014v1_tsd.pdf]
+First, there are now two first stages but the F-tests indicate that the instruments are strong for both, including the interaction term (F = 232).
+Second, the first row shows the effect of pollution on agricultural productivity prior to the opening of a coal plant; the coefficient is negative, indicating negative effects of non-coal pollution.
+Third, the interaction term is strongly negative and more than twice as large in magnitude as the level term, indicating a very clear effect specific to coal plants, in addition to the overall negative effect of pollution.
+
+Importantly, since these results come from an IV, both coefficients denote the change in agricultural yield due to an increase in pollution by the same units.
+In other words, this is not identifying differences in overall pollution, per se, but rather differences in the effects of an increase of pollution by the same percentage.
+These different effects could be due to, for example, differences in emission patterns for different kinds of plants and factories; coal plants, for example, are responsible for much more mercury than other kinds of polluting industries.^[https://www.epa.gov/sites/default/files/2016-12/documents/nei2014v1_tsd.pdf]
 
@@
 
-Pollution is a growing problem throughout the developing world.
-India, for example, has some of the worst air quality on earth.
-The rapid growth of coal power plants has also led to increases in emissions from the burning of coal.
 In this paper, I show some of the negative effects associated with pollution, in general, and coal pollution, in particular.
@@
 
-These results, of course, do not address possible positive effects of the building of additional power plants, which can lead to increased electricity access.
-As such, the main conclusion is that the location of new power generation facilities may be particularly important for multiple aspects of economic activity, not just health and mortality.
+These results document relatively large, local externalities in the agricultural sector in a developing country.
+Since the sector is a refuge for some of the poorest citizens in the world, the results raise important questions about the current speed of the energy transition as it relates to sustainable development and poverty reduction, more generally.
+While we often frame the debate around greenhouse gasses in terms of climate change, there are important local reasons to reduce pollution emissions, as well.
+The technology for greener energy exists; at this point, it is simply a question of policies [@markard2018next; blazquez2020some].
+
+These results, of course, do not address possible positive effects of the building of additional power plants, which can lead to increased electricity access, or shifting more of the population into non-agricultural sectors, which can raise standards of living.
+As such, the main conclusion is that the location of pollution emissions -- especially from coal plants -- may be particularly important for multiple aspects of economic activity, not just health and mortality.
 
@@
 
+```{r randomplotleads, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 8, fig.cap = "Randomization inference for leads", fig.footer = list("Note: The figure presents the distribution of coefficients by randomly assigning a single wind value to each village, sampled with replacement across all villages in a given district in the same year.
+\"Current\" refers to the coefficient on current pollution, while \"lead\" refers to the coefficient on next year's pollution in the same table.")}
+
+pal <- viridis(4)
+colors <- c(paste0(pal[1]), paste0(pal[3]))
+labels <- c("current", "lead")
+
+# plot
+ggplot() + 
+    geom_density(aes(x = coef_current, fill = "current"), alpha = 0.5) +
+    geom_density(aes(x = coef_lead, fill = "lead"), alpha = 0.5) +
+    scale_fill_manual(
+      name = "coefficient:",
+      values = colors,
+      labels = labels
+    ) +
+    theme_minimal() +
+    xlab("distribution of coefficients")
+
+```
+
 ```{=tex}
@@
 ```
+
+```{=tex}
+\FloatBarrier
+\newpage
+```
+```{r yield8cumulative, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+# Table
+colnames(yield8cumulative) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
+kable(
+      yield8cumulative, caption = "Cumulative exposure to pollution during the season",
+      align = "ccccc", booktabs = TRUE, linesep = ""
+      ) %>%
+  footnote(
+           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
+           general_title = "",
+           threeparttable = TRUE,
+           footnote_as_chunk = TRUE
+           ) %>%
+  column_spec(1, width = "2.25cm") %>%
+  column_spec(c(2:6),width = "2cm") %>%
+  row_spec(c(12), bold = TRUE) %>%
+  row_spec(14, hline_after = TRUE) %>%
+  kable_classic_2()
+
+```

### 2023-04-11 draft/draft.tex
vastly updated draft

@@
 \hypersetup{
-  pdftitle={Blowing in the wind: Pollution and agricultural productivity in India},
+  pdftitle={Air pollution and agricultural productivity in a developing country},
   pdfauthor={Joshua D. Merfeld},
@@
 
-\title{Blowing in the wind: Pollution and agricultural productivity in India\footnote{I thank Ariel Ortiz-Bobea, Marc Bellemare, and David Sungho Park for comments and suggestions that greatly improved the quality of the paper.
+\title{Air pollution and agricultural productivity in a developing country\footnote{I thank Ariel Ortiz-Bobea, Marc Bellemare, and David Sungho Park for comments and suggestions that greatly improved the quality of the paper.
 The usual disclaimer applies.
@@
 \author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
-\date{2023-04-07}
+\date{2023-04-11}
 
@@
 \begin{abstract}
-\noindent Pollution is a growing problem in many developing countries.
-Using the rollout of coal plants to identify high-pollution locations in India and random variation induced by changes in wind direction, I show that pollution has large, negative impacts on agricultural productivity.
-Results indicate that some of this impact is driven by changes in labor allocation patterns in response to increased exposure to pollution.
-I also show that pollution from coal plants, specifically, has a larger effect on agricultural productivity than other types of pollution.\\
+\noindent I document large negative externalities in the agricultural sector of India due to air pollution.
+Using random variation in pollution exposure induced by changes in wind direction, I show that higher levels of particulate matter lead to decreased agricultural productivity.
+Based on within-village average deviations in pollution exposure, year-to-year changes in productivity can be above 10 percent.
+Combining wind direction with the rollout of coal plants, I show that pollution from coal plants, specifically, has a larger effect on agricultural productivity than other types of pollution.
+Given that the agricultural sector is a refuge for the poor in many developing countries, these results suggest that the effects of pollution may hit the poorest particularly hard.\\
 \strut \\
@@
 
-Coal is one of the most common energy sources throughout the world.
-In 2021, energy generation from coal reached an all-time high,\footnote{www.iea.org/reports/coal-fired-electricity} driven in part by rapid population growth and rising energy prices.
-India is no exception to these trends.
-From 1990 to 2010, India's population grew by more than 40 percent.\footnote{data.worldbank.org} This population growth led to a large increase in demand for power.
-While natural gas has become a popular alternative in much of the world, the Indian government met much of the increased demand through the construction of coal power plants.
-During those same two decades, power generation from all coal units of at least 30 megawatts in the country increased by almost 140 percent, from 42.4 gigawatts to more than 100 gigawatts.\footnote{globalenergymonitor.org/} There are no doubt positive effects from this increase in power generation; for example, electricity may lead to economic and productivity growth (Dinkelman 2011; Kline and Moretti 2014; Rud 2012; Van de Walle et al.
+While reliance on coal is declining in much of the world, it continues to be an important source of energy in many countries.
+In 2021, energy generation from coal reached an all-time high,\footnote{www.iea.org/reports/coal-fired-electricity} driven in part by rapid population growth in developing countries, including India.
+From 1990 to 2010, India's population grew by more than 40 percent, leading to a large increase in demand for power.\footnote{data.worldbank.org} While natural gas has become a popular alternative in much of the world, the Indian government met much of the increased demand through the construction of coal power plants.
+During those same two decades, power generation from all coal units of at least 30 megawatts in the country more than doubled, from 42.4 gigawatts to more than 100 gigawatts.\footnote{globalenergymonitor.org/}
+
+There are no doubt positive effects from this increase in power generation; for example, electricity may lead to economic and productivity growth (Dinkelman 2011; Kline and Moretti 2014; Rud 2012; Van de Walle et al.
 2017), though electrification alone may not be a sufficient condition (K.
 Lee, Miguel, and Wolfram 2020).
-
 These improvements do not come without downsides, however, especially when it comes to power derived from coal.
@@
 
-I study the effects of emissions from coal power plants on agricultural productivity in India.
+In addition to coal, pollution more generally is increasing in much of the world.
+Carbon dioxide equivalent per capita has almost double over the last 50 years in India, more than doubled in Indonesia, and almost tripled in China.\footnote{\url{https://edgar.jrc.ec.europa.eu}} Of course, much of this increase is due to the development process itself; as countries develop, their economies transform, with more people moving out of agriculture and into manufacturing and services (Herrendorf, Rogerson, and Valentinyi 2014).
+Despite this development, agriculture remains a key sector in many developing countries, employing a large proportion of the poor (Dercon 2009).
+This also means agriculture is a key sector when it comes to poverty reduction (Christiaensen, Demery, and Kuhl 2011).
+Given the large effects of pollution on health and previously documented impacts on agriculture, this raises a important question: do rising pollution levels in developing countries threaten agricultural productivity?
+
+In this paper, I study the effects of pollution on agricultural productivity in India.
 Previous work has suggested that there may be harmful direct effects of air pollution (Heck et al.
@@
 1997) and water pollution (Reddy and Behera 2006) on agricultural productivity, including work showing that gold mines -- which release large amounts of pollutants into the surrounding environment -- lead to lower agricultural productivity in Ghana (Aragón and Rud 2016).
-These effects can work through the well known effects on human health and productivity, but also directly through land productivity.
+These effects can work through the well-known effects on human health and productivity, but also directly through land productivity.
 Black carbon in soot, for example, can directly affect solar radiation reaching plants on the surface (Ramanathan and Carmichael 2008; Burney and Ramanathan 2014).
-Relative to previous work in economics, this paper tackles the broader case of coal plants, which are a leading source of power throughout the world.
-I show that air pollution from these plants can lead to lower agricultural productivity and even decreased economic growth.
+Relative to previous work in economics, this paper tackles pollution more broadly, before turning to a specific source of pollution: coal plants.
+I show that air pollution leads to lower agricultural productivity in India and that pollution from coal plants is particularly harmful to agriculture.
 
@@
 2020).
-In the case of the data used here, coal plants are constructed in areas with higher population and higher literacy (Table \ref{tab:plantresultstable}).
+In the case of the data I use here, coal plants are constructed in areas with higher population and higher literacy.
 In terms of agriculture, higher productivity predicts more coal plants in a given year as well as the construction of a coal plant over the next decade, indicating possible bias when estimating OLS regressions of agricultural productivity on pollution levels.
 
-To identify the effects of air pollution on agricultural productivity, I instead combine the roll out of coal plants in India with plausibly exogenous changes in wind direction -- in a similar spirit to Deryugina et al.
-(2019) -- to measure the effects of pollution from coal plants on key development outcomes.
+To identify the effects of air pollution on agricultural productivity, I identify areas of high pollution emission using the location of coal plants.
+While coal plants may add emissions, I show that they are located in areas with relatively higher levels of pollution, in general.
+Using these locations as point sources of pollution, I use plausibly exogenous changes in wind direction -- in a similar spirit to Deryugina et al.
+(2019) -- to measure the effects of pollution on key agricultural outcomes.
 I first plot the location of coal plants in India by year.
 Using a village-level shapefile, I identify all villages that are located within 30km of a coal plant at any time during the sample period, which comprises the 2000s to the 2010s, with differences across outcomes depending on data availability.
-
-I use the location of coal plants to identify high-pollution areas and calculate the direction from each coal plant to all villages within 30km.
-I include all areas, regardless of whether the coal plant has opened yet, to allow me to disaggregate the effects of pollution from coal plants with other kinds of pollution.
-Using daily wind data and these spatial variables, I create an identifier separately for each of the one-hundred-thousand-plus villages for whether a given village is located downwind from a coal plant on any given day.
+I then calculate the direction from each coal plant to all villages within 30km.
+I include all areas, regardless of whether the coal plant has opened yet, to allow me to disaggregate the effects of pollution from coal plants with other kinds of pollution later in the paper.
+Using daily wind data and these spatial variables, I create an identifier separately for every village for whether a given village is located downwind from a coal plant on any given day.
 I then aggregate this exposure variable based on the desired analysis; for agricultural productivity, for example, I create an exposure variable measuring the total number of exposure days in the five-month agricultural season, focusing only on the monsoon (Kharif) season.
-Consistent with expectations, wind direction is a strong predictor of particulate matter at the monthly level.
-More days being located downwind from a coal power plant leads to higher average concentrations of particulate matter 2.5.
-
-Using newly available estimates of India-wide agricultural (land) productivity, results show that the total number of days of being downwind from a coal power plant has a negative effect on overall agricultural productivity.
-Just a single additional day of being downwind from a coal power plant leads to a decrease in predicted yields of 0.6 percent.
-Combined with the within-village absolute deviation of wind exposure, results indicate that median changes in wind direction are responsible for fluctuations in agricultural productivity of around 1.5 percent.
-
-Using wind direction as an instrument for particulate matter, an increase of PM 2.5 by one percent leads to an even larger decrease in agricultural productivity, at around 0.73 percent.
-Given within-village changes across years, pollution could be responsible for swings of 7-to-14 percent in agricultural productivity across years.
-This IV strategy is similar to that in Deryugina et al.
-(2019), though they use general ``high-pollution'' wind directions while I concentrate only on wind directions from the location of a coal power plant.
-The IV estimates indicate that the effect of pollution is more negative than naïve OLS estimates indicate, which is consistent with the evidence on the construction of coal plants; higher levels of pollution are correlated with higher levels of agricultural productivity and economic growth, a relationship that the IV strategy adjusts downwards.
-
-These effects are apparently larger in villages that experience higher median exposure to pollution and in villages with higher initial yields.
-I also show that rainfall shocks are compounding; higher exposure along with less rainfall increases the negative effect of either shock individually.
-As weather shocks become more common across the globe, the rise of pollution in India may be particularly problematic for farmers.
-
-To explore the mechanisms driving this effect, I also use data on labor allocation from the National Sample Survey (NSS).
+More days being located downwind from a high-pollution location leads to substantially higher average concentrations of particulate matter 2.5 (PM 2.5).\footnote{PM 2.5 is defined as particulate matter with a diameter of less than 2.5 micrometers.}
+
+Using wind direction as an instrument for particulate matter, an increase of PM 2.5 by one percent leads to a large decrease in agricultural productivity, at around 0.73 percent.
+Given within-village changes in exposure, pollution could be responsible for swings of 7-to-14 percent in agricultural productivity across years.
+The IV estimates indicate that the effect of pollution is more negative than naïve OLS estimates, which is consistent with the evidence on the construction of coal plants; higher levels of pollution are correlated with higher levels of agricultural productivity, a relationship that the IV strategy adjusts downwards.
+Robustness checks confirm these main results.
+Randomization inference shows that the estimates are robust to alternative measures of uncertainty, while leads of pollution do not affect current agricultural productivity.
+
+Heterogeneity analysis shows that the negative effects of pollution are larger in villages that experience higher median exposure to pollution and in villages with higher initial yields.
+Finally, given how many poor farmers rely on rainfall and how weather shocks increase across the globe, I also look at possible multiplicative effects of shocks by interacting the pollution variables with a rainfall variable.
+Higher exposure along with less rainfall increases the negative effect of either shock individually, meaning these shocks are compounding.
+Given concerns regarding agricultural productivity and climate change, the rise of pollution in India may be particularly problematic for poor farmers.
+
+I find that these effects are mostly transmitted by reductions in farm labor during narrow windows of the growing season, using labor allocation data from the National Sample Survey (NSS).
 While there do not appear to be large changes in labor allocation patterns when looking at the season as a whole, there is a clear pattern when we break down exposure to pollution by month: higher levels of pollution at the beginning of the monsoon season (June and July) lead to lower levels of farm labor allocation, but no changes in non-farm labor allocation.
 This matches up well with a similar exercise with agricultural yields: pollution in June is the only month in which there is a negative effect of pollution on agricultural yield.
-In tandem, these two results point to effects on labor as one mechanism driving the total effects, though I cannot rule out direct effects of pollution on land productivity.
-
-The final set of results tries to disentangle the effects of pollution from the effects of coal pollution, specifically.
+In tandem, these two results point to effects on labor as one mechanism driving the total effects, though I cannot rule out direct effects of pollution on land productivity with my data.
+
+The final set of analyses tries to disentangle the effects of pollution from the effects of coal pollution, specifically, by using wind direction and the rollout of coal plants across the country.
 Combining instrumental variables with a differences-in-differences strategy, I show that agricultural productivity has an elasticity of around -0.15 with respect to non-coal pollution.
-The opening of a coal plant, however, triples the elasticity, to more than -0.45, indicating that coal plant emissions are particularly harmful to agricultural productivity.
+The opening of a coal plant, however, triples the elasticity, to more than -0.45, indicating that coal plant emissions are particularly harmful to agricultural productivity, at least in India.
 This is consistent with coal emissions being quite different from other forms of pollution; for example, coal emissions contain much more mercury than other forms of pollution.
@@
 Perhaps relevant to my labor allocation results, Hanna and Oliva (2015) show that decreases in pollution lead to increases in labor allocation, while I show the same relationship in at least a subset of months of the agricultural season.
-The papers most related to this one are Aragón and Rud (2016), who show that gold mines in Ghana have substantial negative impacts on agricultural productivity in the surrounding areas, and Burney and Ramanathan (2014), who model the effects of climate change and pollutants on agricultural productivity in India.
-This paper differs in two important ways.
-First, I explore effects of pollution from coal plants, which are much more ubiquitous than gold mines across the world.
+Several papers are closely related to this one.
+Aragón and Rud (2016) show that gold mines in Ghana have substantial negative impacts on agricultural productivity in the surrounding areas.
+Burney and Ramanathan (2014) model the effects of climate change and pollutants on agricultural productivity in India.
+Finally, Sanders and Barreca (2022) actually find that coal emissions can be good for agriculture in the United States, due to the presence of sulfur, which can increase agricultural productivity.
+This paper differs in three important ways.
+First, I explore effects of pollution more generally as well as from coal plants, which are much more ubiquitous than gold mines across the world.
 Second, I use a different identification strategy from either paper, relying on wind blowing pollution from coal plants.
-
-Sanders and Barreca (2022)
+Finally, agriculture in India and the United States can be quite different, with the latter being much more mechanized.
+This means effects of pollution could differ substantially.
 
@@
 
-\hypertarget{data-and-methods}{%
-\section{Data and methods}\label{data-and-methods}}
+\hypertarget{data}{%
+\section{Data}\label{data}}
 
 \label{data}
-
-In this section, I describe the different data sources and the identification strategy.
-I also spend time validating the use of wind direction as a proxy for particulate matter as well as looking at the correlates of construction of coal power plants.
-
-\hypertarget{data}{%
-\subsection{Data}\label{data}}
-
-\label{data2}
 
@@
 
-\hypertarget{identification}{%
-\subsection{Identification}\label{identification}}
+\hypertarget{methods}{%
+\section{Methods}\label{methods}}
 
@@
 So what are the characteristics of areas near the construction of new coal plants?
-Table \ref{tab:plantresultstable} presents four separate regressions.
+Table \ref{tab:plantresultstable} presents four separate regressions, using state-level fixed effects due to the spatial variation in coal plants.
 The first two columns use 1991 census data -- downloaded from SHRUG -- and the second two columns use 2001 census data, all at the village level.
@@
 \midrule
-Population (log) & 0.010*** & 0.001*** & 0.017*** & 0.004***\\
- & (0.000) & (0.000) & (0.000) & (0.000)\\
-Area (log) & -0.006*** & 0.001** & -0.015*** & -0.001***\\
- & (0.001) & (0.000) & (0.000) & (0.000)\\
-Literacy (prop) & 0.024*** & 0.017*** & 0.032*** & 0.029***\\
- & (0.003) & (0.002) & (0.002) & (0.002)\\
-Ag. productivity &  &  & 0.026*** & 0.004***\\
- &  &  & (0.001) & (0.000)\\
+Population (log) & 0.009 & 0.007* & 0.013* & 0.005\\
+ & (0.006) & (0.004) & (0.006) & (0.004)\\
+Area (log) & -0.007* & -0.004 & -0.007* & -0.004\\
+ & (0.004) & (0.005) & (0.004) & (0.003)\\
+Literacy (prop) & 0.080** & 0.037 & 0.073* & 0.036\\
+ & (0.029) & (0.028) & (0.039) & (0.026)\\
+Ag. productivity &  &  & 0.014** & 0.028***\\
+ &  &  & (0.005) & (0.009)\\
+Pollution &  &  & 0.047 & 0.014\\
+(log PM 2.5) &  &  & (0.045) & (0.018)\\
 Sub-sample & all & no plant & all & no plant\\
@@
 \begin{tablenotes}[para]
-\item Note: Robust standard errors are in parentheses.
+\item Note: State fixed effects are included in regressions.
+Standard errors are in parentheses and clustered at the state level.
 The outcome in the first column is whether the village is within 30km of a coal plant in 1991.
@@
 Consider the data used for the regressions presented in Table \ref{tab:pollutiontable} of the appendix.
-The outcome variable is particulate matter -- specifically, \(\mathrm{PM_{2.5}}\), which is particulate matter no larger than 2.5 micrometers in diameter -- which comes from Hammer et al.
+The outcome variable is particulate matter -- specifically, \(\mathrm{PM 2.5}\), which is particulate matter no larger than 2.5 micrometers in diameter -- which comes from Hammer et al.
 (2020).
-\(\mathrm{PM_{2.5}}\) is one of the harmful byproducts from coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.\footnote{www.epa.gov/airmarkets/power-plants-and-neighboring-communities} This data is available at the monthly level, so in Table \ref{tab:pollutiontable}, I aggregate wind exposure to the monthly level, as well.
+\(\mathrm{PM 2.5}\) is one of the harmful byproducts from coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.\footnote{www.epa.gov/airmarkets/power-plants-and-neighboring-communities} This data is available at the monthly level, so in Table \ref{tab:pollutiontable}, I aggregate wind exposure to the monthly level, as well.
 
@@
 The first two columns include all available years of data while the last two columns restrict estimation only to the years for which the agricultural productivity regressions are estimated.
-Across all four columns, the story is the same: when the wind is blowing in the direction of a village from a coal plant within 30 kilometers, estimated \(\mathrm{PM_{2.5}}\) is substantially higher.
-
-To put these numbers into perspective, the World Health Organization's updated guidelines are that average annual exposure should not exceed 15 \(\mathrm{\mu g/m^3}\).
+Across all four columns, the story is the same: when the wind is blowing in the direction of a village from a coal plant within 30 kilometers, estimated \(\mathrm{PM 2.5}\) is substantially higher.
+
+To put context behind pollution numbers, the World Health Organization's updated guidelines are that average annual exposure should not exceed 15 \(\mathrm{\mu g/m^3}\).
 For the month, an increase at the midpoint of the estimated range is equal to an increase of 0.2 percent of the maximum recommended mean concentration from the WHO.
@@
 \hypertarget{identifying-effects-of-pollution-from-coal-plants}{%
-\subsubsection{Identifying effects of pollution from coal plants}\label{identifying-effects-of-pollution-from-coal-plants}}
+\subsection{Identifying effects of pollution from coal plants}\label{identifying-effects-of-pollution-from-coal-plants}}
 
@@
 \label{results}
-
-This section presents the main results of this paper. I present two main sets of results: one set looking at agricultural productivity and another looking at labor allocation.
-
-\hypertarget{agricultural-productivity}{%
-\subsection{Agricultural productivity}\label{agricultural-productivity}}
 
@@
 Instead, we are actually interested in better understanding the effects of exposure to pollution on agricultural productivity.
-Table \ref{tab:yieldtablepm} presents the same regressions as Table \ref{tab:yieldtable}, with a key difference: instead of wind exposure, it uses log average concentration of \(\mathrm{PM_{2.5}}\) -- from Hammer et al.
+Table \ref{tab:yieldtablepm} presents the same regressions as Table \ref{tab:yieldtable}, with a key difference: instead of wind exposure, it uses log average concentration of \(\mathrm{PM 2.5}\) -- from Hammer et al.
 (2020) -- during the season as the key explanatory variable.
@@
 
-The second-stage results first indicate that \(\mathrm{PM_{2.5}}\) is a very strong predictor of agricultural productivity.
+The second-stage results first indicate that \(\mathrm{PM 2.5}\) is a very strong predictor of agricultural productivity.
 Using the results in column three, a one percent increase in particulate matter leads to a 0.73 percent decrease in agricultural productivity.
@@
 
-\hypertarget{robustness-checks}{%
-\subsection{Robustness checks}\label{robustness-checks}}
-
 \begin{figure}
@@
 \end{figure}
+
+\hypertarget{robustness-checks}{%
+\subsection{Robustness checks}\label{robustness-checks}}
+
+In this section, I report the results of two separate robustness checks for the main results presented above.
+In the previous results, standard errors are clustered at the level of the village, which is also the level at which treatment (exposure to pollution) is assigned as well as the level at which agricultural productivity is extracted.
+Past work has shown that standard errors might be underestimated in such contexts (Bertrand, Duflo, and Mullainathan 2004).
+This is a particular concern in the present case, since agricultural productivity is extracted at the village level but estimated using less coarse data that has high spatial correlation (Gangopadhyay et al.
+2022).
+
+To help remedy concerns, I perform randomization inference (e.g.
+Barrios et al.
+(2012)) by randomly assigning observed treatment across villages and repeating this process 1,000 times.
+I do this in two separate ways.
+First, I take all villages in a given year and randomly assign, with replacement, treatment intensity within the year.
+Second, I do this separately by selecting from within district-years only, to help take into account district-year correlation in outcomes and as a way to mitigate possible concerns that traditional clustered standard errors may actually be much too conservative (Abadie et al.
+2023).
+Since this is an instrumental variables procedure, I randomly select the wind and pollution variable \emph{together}.
+
+I present these results in Figure \ref{fig:randomplot}.
+The purple density is the resulting distribution of estimated treatment effects when selecting from all villages within the same year.
+The density of treatment effects is relatively tight around 0, with the left tail only getting slightly beyond -0.05, compared to the most conservative actual estimated treatment effect of -0.73.
+
+The within-district correlation in outcomes and treatment is apparent when instead randomly selecting from only villages within the same district in the same year.
+The distribution is still approximately normal but has a mode of almost -0.4.
+Nonetheless, the actual treatment effect is larger in magnitude than every single one of the 1,000 bootstrapped replications.
+If we were to use the standard deviation of the resulting distribution as a standard error, we would get a confidence interval around the main effect of approximately -0.899 to -0.561.
+The results are apparently robust to alternative specifications of uncertainty
+
+The second robustness check includes leads of pollution and wind direction.
+I present these results in Table \ref{tab:yieldtableleads}.
+Interestingly, there is actually a positive coefficient for the lead in the second column.
+However, if we perform randomization inference, it appears that the confidence intervals might be slightly underestimated; the positive effect size is in the 11th percentile of the distribution in Figure \ref{fig:randomplotleads}, which for a two-tailed test would not even have a p-value of less than 0.2.
+Overall, there is no indication that leads predict any changes in current agricultural productivity.
 
@@
 Another way to think about this, however, is that higher values of rainfall dampen the effects of pollution shocks.
-This could have very profound effects on welfare, since this essentially increases the variance of agricultural productivity across years (which is already increasing due to climate change).
-
-In the appendix, I present in Table \ref{tab:yieldtableleads} as a robustness check similar regressions but also including wind/pollution leads.
-The negative effect is only seen with current pollution, not future pollution.
+This could have very profound effects on welfare, since this essentially increases the variance of agricultural productivity across years (which is already increasing due to climate change), at least as it relates to pollution exposure..
 
@@
 Table \ref{tab:labortable} presents results for labor allocation over the last seven days using the National Sample Survey and the methodology described above.
-Since the labor recall is for seven days, labor allocation ranges from zero to seven, as does the wind variable.
-Recall that the data is at the district level, not the village level.
+Since the labor recall is for seven days, labor allocation ranges from zero to seven, as does the wind variable.\footnote{I also tried an instrumental variables strategy with the seven-day wind and monthly pollution, but the first stage lacked power due to the difference in temporal aggregation across the two key variables (pollution and wind).} Recall that the data is at the district level, not the village level.
 However, it is also more temporally disaggregated, with rolling interviews throughout the year-long survey.
@@
 
-While overall results for labor allocation are inconclusive, perhaps we can learn more by looking at temporally disaggregated results, focusing on May through December.
-Although the yield data was created by looking at five specific months of the year -- June through October -- this is partially because this is the time of year when remote-sensing data is most able to pick up overall changes in crop cover.
-Preparation, however, starts before that, with operations such as ploughing often taking place prior to June.
+While overall results for labor allocation are inconclusive, perhaps we can learn more by looking at temporally disaggregated results, focusing on June through October.
+I can create separate variables for exposure by each month of the season, keeping the unit of analysis the individual (for labor) or the village-season (for yields).
 
 Since the NSS data has survey dates and the both the pollution and wind data can be aggregated to the month, we can look at how exposure during different months of the agricultural season (differentially) affect labor allocation and overall agricultural productivity.
-I begin with labor allocation in Table \ref{tab:labor3month}.
+I begin with labor allocation in columns one and two of Table \ref{tab:labor3month}.
 As before, the exposure measure is just the wind variable (over the last seven days), since I can directly match daily wind direction with the timing of the survey, but I cannot do that for pollution, which is monthly.
@@
 
-I perform a similar exercise with yield, the results of which are in Table \ref{tab:labor3month}.
+I perform a similar exercise with yield, the results of which are in the third column Table \ref{tab:labor3month}.
+I return to the IV strategy, however, to be consistent with the previous yield results.
 I include only the five months of the year that are used in the calculation of the yield estimates, or June through October.
-The results are again from two-stage least squares, with the F-statistics for each month presented at the bottom; monthly wind remains a strong predictor of pollution.
-Column one presents results without controls for weather while column two presents results with controls for weather; the differences in coefficients make clear how important the inclusion of weather is here.
-
-Focusing on column two, there is only one month where we see a clear effect of pollution on yield: June.
-Given that we also see negative effects of pollution on farm labor allocation in June, this is suggestive -- but not conclusive -- evidence that pollution works through labor.
-Note that this does not rule out pollution also having direct effects on agricultural productivity, however.
+
+There is only one month where we see a clear effect of pollution on yield: June.
+Given that we also see negative effects of pollution on farm labor allocation in June, this is suggestive -- but not conclusive -- evidence that pollution at least partly works through labor.
+Note that this does not rule out pollution also having direct effects on agricultural land productivity, however.
 
@@
 For example, yield might respond strongly to average pollution in June and July, which would not be captured in Table \ref{tab:labor3month}.
+As one additional check, Table \ref{tab:yield8cumulative} in the appendix presents five separate instrumental variable regressions, which look at the effects of cumulative exposure to pollution separately by month.
+Across all five months, cumulative exposure negatively affects agricultural productivity, with the largest coefficient actually showing up for October, the last month of the season.
 
@@
 Column three presents the most extensive set of controls and I focus on that column here.
-First, there are now two first stages but the F-test indicates that the instruments are strong for both, including the interaction term (F = 232).
-Second, the first row shows the effect of pollution on agricultural productivity prior to the opening of a coal plant; the coefficient is negative but relatively smaller.
-Third, the interaction term is strongly negative and more than twice as large in magnitude as the level term, indicating a very clear effect specific to coal plants.
-This could be due to differences in emission patterns for different kinds of plants and factors; coal plants, for example, are responsible for much more mercury than other kinds of polluting industries.\footnote{\url{https://www.epa.gov/sites/default/files/2016-12/documents/nei2014v1_tsd.pdf}}
+First, there are now two first stages but the F-tests indicate that the instruments are strong for both, including the interaction term (F = 232).
+Second, the first row shows the effect of pollution on agricultural productivity prior to the opening of a coal plant; the coefficient is negative, indicating negative effects of non-coal pollution.
+Third, the interaction term is strongly negative and more than twice as large in magnitude as the level term, indicating a very clear effect specific to coal plants, in addition to the overall negative effect of pollution.
+
+Importantly, since these results come from an IV, both coefficients denote the change in agricultural yield due to an increase in pollution by the same units.
+In other words, this is not identifying differences in overall pollution, per se, but rather differences in the effects of an increase of pollution by the same percentage.
+These different effects could be due to, for example, differences in emission patterns for different kinds of plants and factories; coal plants, for example, are responsible for much more mercury than other kinds of polluting industries.\footnote{\url{https://www.epa.gov/sites/default/files/2016-12/documents/nei2014v1_tsd.pdf}}
 
@@
 
-These results, of course, do not address possible positive effects of the building of additional power plants, which can lead to increased electricity access.
-As such, the main conclusion is that the location of new power generation facilities may be particularly important for multiple aspects of economic activity, not just health and mortality.
+These results document relatively large, local externalities in the agricultural sector in a developing country.
+Since the sector is a refuge for some of the poorest citizens in the world, the results raise important questions about the current speed of the energy transition as it relates to sustainable development and poverty reduction, more generally.
+While we often frame the debate around greenhouse gasses in terms of climate change, there are important local reasons to reduce pollution emissions, as well.
+The technology for greener energy exists; at this point, it is simply a question of policies {[}Markard (2018); blazquez2020some{]}.
+
+These results, of course, do not address possible positive effects of the building of additional power plants, which can lead to increased electricity access, or shifting more of the population into non-agricultural sectors, which can raise standards of living.
+As such, the main conclusion is that the location of pollution emissions -- especially from coal plants -- may be particularly important for multiple aspects of economic activity, not just health and mortality.
 
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-abadie2023should}{}}%
+Abadie, Alberto, Susan Athey, Guido W Imbens, and Jeffrey M Wooldridge. 2023. {``When Should You Adjust Standard Errors for Clustering?''} \emph{{The Quarterly Journal of Economics}} 138 (1): 1--35.
+
 \leavevmode\vadjust pre{\hypertarget{ref-abay2022nonclassical}{}}%
@@
 2022.
-{``{Nonclassical Measurement Error and Farmers?
-Response to Information Reveal Behavioral Anomalies}.''}
+{``{Nonclassical Measurement Error and Farmers' Response to Information Reveal Behavioral Anomalies}.''}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-barrios2012clustering}{}}%
+Barrios, Thomas, Rebecca Diamond, Guido W Imbens, and Michal Kolesár.
+2012.
+{``Clustering, Spatial Correlations, and Randomization Inference.''} \emph{{Journal of the American Statistical Association}} 107 (498): 578--91.
+
 \leavevmode\vadjust pre{\hypertarget{ref-bencko1977health}{}}%
 Bencko, Vladimir, and Karel Symon. 1977. {``Health Aspects of Burning Coal with a High Arsenic Content.''} \emph{{Environmental Research}} 13 (3): 378--85.
+
+\leavevmode\vadjust pre{\hypertarget{ref-bertrand2004much}{}}%
+Bertrand, Marianne, Esther Duflo, and Sendhil Mullainathan. 2004. {``How Much Should We Trust Differences-in-Differences Estimates?''} \emph{{The Quarterly Journal of Economics}} 119 (1): 249--75.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-christiaensen2011evolving}{}}%
+Christiaensen, Luc, Lionel Demery, and Jesper Kuhl.
+2011.
+{``{The (evolving) role of agriculture in poverty reduction --- An empirical perspective}.''} \emph{{Journal of Development Economics}} 96 (2): 239--54.
+
 \leavevmode\vadjust pre{\hypertarget{ref-currie2014we}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-dercon2009rural}{}}%
+Dercon, Stefan. 2009. {``{Rural poverty: Old challenges in new contexts}.''} \emph{{The World Bank Research Observer}} 24 (1): 1--28.
+
 \leavevmode\vadjust pre{\hypertarget{ref-deryugina2019mortality}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-herrendorf2014growth}{}}%
+Herrendorf, Berthold, Richard Rogerson, and Akos Valentinyi. 2014. {``Growth and Structural Transformation.''} \emph{{Handbook of Economic Growth}} 2: 855--941.
+
 \leavevmode\vadjust pre{\hypertarget{ref-iea2022}{}}%
@@
 {``{Health and air quality benefits of policies to reduce coal-fired power plant emissions: a case study in North Carolina}.''} \emph{{Environmental Science \& Technology}} 48 (17): 10019--27.
+
+\leavevmode\vadjust pre{\hypertarget{ref-markard2018next}{}}%
+Markard, Jochen. 2018. {``The Next Phase of the Energy Transition and Its Implications for Research and Policy.''} \emph{{Nature Energy}} 3 (8): 628--33.
 
@@
 
+\begin{figure}
+\includegraphics{draft_files/figure-latex/randomplotleads-1} \caption[Randomization inference for leads]{Randomization inference for leads}\label{fig:randomplotleads}\floatfoot*{Note: The figure presents the distribution of coefficients by randomly assigning a single wind value to each village, sampled with replacement across all villages in a given district in the same year.
+"Current" refers to the coefficient on current pollution, while "lead" refers to the coefficient on next year's pollution in the same table.}
+\end{figure}
+
 \FloatBarrier
@@
 
+\FloatBarrier
+\newpage
+
+\begin{table}
+
+\caption{\label{tab:yield8cumulative}Cumulative exposure to pollution during the season}
+\centering
+\begin{threeparttable}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2.25cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\toprule
+  & (1) & (2) & (3) & (4) & (5)\\
+\midrule
+June & -0.643*** &  &  &  & \\
+ & (0.064) &  &  &  & \\
+July &  & -0.658*** &  &  & \\
+ &  & (0.045) &  &  & \\
+Aug. &  &  & -0.531*** &  & \\
+ &  &  & (0.049) &  & \\
+Sept. &  &  &  & -0.672*** & \\
+ &  &  &  & (0.047) & \\
+Oct. &  &  &  &  & -0.943***\\
+ &  &  &  &  & (0.069)\\
+weather & Yes & Yes & Yes & Yes & Yes\\
+\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+village & Yes & Yes & Yes & Yes & Yes\\
+year & Yes & Yes & Yes & Yes & Yes\\
+\midrule
+F & 1,327 & 2,830 & 2,933 & 2,413 & 1,515\\
+observations & 1,266,922 & 1,266,922 & 1,266,922 & 1,266,922 & 1,266,922\\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}[para]
+\item Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+
 \end{document}

### 2023-04-11 analyses/1.openings.R

@@ -49,12 +49,34 @@ villages00 <- villages00 %>% left_join(monsoon, by = "shrid")
 # rename to get same row in table
 villages90 <- villages90 %>% rename(area = pc91_vd_area)
 villages00 <- villages00 %>% rename(area = pc01_vd_area)
+# pollution
+pollution <- read_csv(paste0("data/clean/pm25/plant_openings199801.csv"))
+pollution <- pollution %>%
+              dplyr::select(shrid)
+for (year in c("1998", "1999", "2000")){
+  for (month in c("01", "02", "03", "04", "05", "06",
+                  "07", "08", "09", "10", "11", "12")){
+    extracted_values <- read_csv(paste0("data/clean/pm25/plant_openings", year, month, ".csv"))
+    colnames(extracted_values) <- c("shrid", paste0(year, month))
+    pollution <- pollution %>%
+                  left_join(extracted_values, by = "shrid")
+  }
+}
+pollution <- pollution[, c(1, 7:11, 19:23, 31:35)]
+pollution$pm25 <- apply(pollution[,-1], 1, FUN = mean)
+pollution <- pollution %>%
+              dplyr::select(shrid, pm25)
+villages00 <- villages00 %>% left_join(pollution, by = "shrid")
+
+villages90$state <- substr(villages90$shrid, 1, 2)
+villages00$state <- substr(villages00$shrid, 1, 2)
 
 
-reg1 <- feols(plants90 ~ log(pop) + log(area) + literacy, data = villages90, vcov = "hetero")
-reg2 <- feols(plants00 ~ log(pop) + log(area) + literacy, data = villages90 %>% filter(plants90==0), vcov = "hetero")
-reg3 <- feols(plants00 ~ ag_prod + log(pop) + log(area) + literacy, data = villages00, vcov = "hetero")
-reg4 <- feols(plants10 ~ ag_prod + log(pop) + log(area) + literacy, data = villages00 %>% filter(plants00==0), vcov = "hetero")
+
+reg1 <- feols(plants90 ~ log(pop) + log(area) + literacy | state, data = villages90, vcov = "cluster")
+reg2 <- feols(plants00 ~ log(pop) + log(area) + literacy | state, data = villages90 %>% filter(plants90==0), vcov = "cluster")
+reg3 <- feols(plants00 ~ ag_prod + log(pm25) + log(pop) + log(area) + literacy | state, data = villages00, vcov = "cluster")
[…]

### 2023-04-11 analyses/3.yield.R

@@ -37,16 +37,21 @@ df <- df %>% group_by(shrid, season) %>%
                      wind_abs_dev = abs(days_sums - mean(days_sums)),
                      pm25_abs_dev = abs(pm25 - mean(pm25)),
                      wind = days_sums/10,
-                     pm1 = log(pm1),
-                     pm2 = log(pm2),
-                     pm3 = log(pm3),
-                     pm4 = log(pm4),
-                     pm5 = log(pm5),
                      pmavg1 = log(pm1),
                      pmavg2 = log((pm1 + pm2)/2),
                      pmavg3 = log((pm1 + pm2 + pm3)/3),
                      pmavg4 = log((pm1 + pm2 + pm3 + pm4)/4),
                      pmavg5 = log((pm1 + pm2 + pm3 + pm4 + pm5)/5),
+                     pm1 = log(pm1),
+                     pm2 = log(pm2),
+                     pm3 = log(pm3),
+                     pm4 = log(pm4),
+                     pm5 = log(pm5),
+                     windavg1 = m1,
+                     windavg2 = (m1 + m2)/2,
+                     windavg3 = (m1 + m2 + m3)/3,
+                     windavg4 = (m1 + m2 + m3 + m4)/4,
+                     windavg5 = (m1 + m2 + m3 + m4 + m5)/5,
                      temp_m1 = (tmax_m1 + tmin_m1)/20,
                      temp_m2 = (tmax_m2 + tmin_m2)/20,
                      temp_m3 = (tmax_m3 + tmin_m3)/20,
@@ -455,9 +460,6 @@ saveRDS(yieldtable, "pollution_development/draft/tables/yield3ivmain_lead.rds")
 
 
 
-
-
-
 ## placebo test
 wind_year <- df %>%
               dplyr::select(wind, pm25, year) %>%
@@ -617,6 +619,118 @@ row # larger than every single randomized value!
 
 
 
+
+
+
+
+
+
+
+
+# randomization for current/lead
+# create function to sample
+sample_pre_fun <- function(a) {
+  sample(x = min(a):max(a), size = 1, replace = T)
+}
+
+# now wrap that function in apply
+sample_fun <- function(a) {
+  sample_vec <- apply(a, 1, sample_pre_fun)
+}
+
+# Also get distribution if we resample wind from WITHIN DISTRICTS ONLY
+## placebo test
+df <- as_tibble(df)
+wind_year <- df %>%
[…]

### 2023-04-11 analyses/4.nss.R

@@ -9,6 +9,7 @@ library(lubridate)
 library(viridis)
 library(fixest)
 library(sf)
+library(tree)
 
 
 # Sets wd to folder this script is in so we can create relative paths instead of absolute paths
@@ -278,6 +279,47 @@ laborplot <- ggplot(data = df) +
        y = "households interviewed") +
   theme_minimal()
 saveRDS(laborplot, "pollution_development/draft/tables/laborplot.rds")
+
+
+
+
+
+
+
+
+
+# treatment effect heterogeneity ------------------------------------
+# first get centered values
+df_centered <- df %>%
+                group_by(district) %>%
+                mutate(
+                       month1 = as.numeric(month_int==6),
+                       month2 = as.numeric(month_int==7),
+                       month3 = as.numeric(month_int==8),
+                       month4 = as.numeric(month_int==9),
+                       month5 = as.numeric(month_int==10),
+                       wind = wind - mean(wind),
+                       wind1 = wind*(month_int==6) - mean(wind*(month_int==6)),
+                       wind2 = wind*(month_int==7) - mean(wind*(month_int==7)),
+                       wind3 = wind*(month_int==8) - mean(wind*(month_int==8)),
+                       wind4 = wind*(month_int==9) - mean(wind*(month_int==9)),
+                       wind5 = wind*(month_int==10) - mean(wind*(month_int==10)),
+                       educ = as.numeric(educ) - mean(as.numeric(educ)),
+                       distfe = min(row_number())
+                       ) %>%
+                ungroup() %>%
+                dplyr::select("month_int", "female", "age", "educ", "temp_mean", "precip_sum",
+                              "wind1", "wind2", "wind3", "wind4", "wind5",
+                              "days_f", "days_nf", "wind", "district", "distfe") %>%
+                filter(month_int %in% c(6, 7, 8, 9, 10))
+df_centered <- df_centered[complete.cases(df_centered),]
+xgboost <- xgboost(data = as.matrix(df_centered[,c("wind1", "wind2", "wind3", "wind4", "wind5", 
+                                             "female", "age", "educ", "temp_mean", "precip_sum")]),
+                   label = df_centered$days_f,
+                   nrounds = 250,
+                   subsample = 0.6,
+                   colsample_bytree = 0.6
+                   )
 
 
 

### 2023-04-11 cleaning/6.cleaning_pm25.R

@@ -25,6 +25,30 @@ getwd()    # check
 
 
 
+# first extract pollution in villages for the 2001 census
+villages <- read_sf("data/spatial/shapefiles/village.shp")
+villages$shrid <- paste0(villages$pc11_s_id, "-", villages$pc11_tv_id)
+for (year in c("1998", "1999", "2000")){
+  for (month in c("01", "02", "03", "04", "05", "06",
+                  "07", "08", "09", "10", "11", "12")){
+    r <- raster(paste0("data/raw/pm25/V5GL03.HybridPM25.Asia.", year, month, "-", year, month, ".nc"))
+    
+    # crop first (makes extraction MUCH faster)
+    r <- crop(r, villages)
+    
+    # extract
+    extracted_values <- exact_extract(r, villages, fun = "mean", append_cols = "shrid")
+    extracted_values <- as_tibble(extracted_values) %>% 
+                          group_by(shrid) %>% 
+                          mutate(mean = mean(mean)) %>% 
+                          filter(row_number()==1) %>%
+                          ungroup()
+    colnames(extracted_values) <- c("shrid", "pm25")
+    write_csv(extracted_values, paste0("data/clean/pm25/plant_openings", year, month, ".csv"))
+  }
+}
+
+
 villages_centroids <- read_sf("data/spatial/centroids_villages_30km")
 villages_centroids <- as_tibble(villages_centroids) %>% dplyr::select(shrid)
 villages_centroids$within30 <- 1

### 2023-04-12 draft/draft.Rmd
updating some figures

@@
 abstract: |
-  | \noindent I document large negative externalities in the agricultural sector of India due to air pollution.
-Using random variation in pollution exposure induced by changes in wind direction, I show that higher levels of particulate matter lead to decreased agricultural productivity.
-Based on within-village average deviations in pollution exposure, year-to-year changes in productivity can be above 10 percent.
-Combining wind direction with the rollout of coal plants, I show that pollution from coal plants, specifically, has a larger effect on agricultural productivity than other types of pollution.
+  | \noindent I document externalities in the agricultural sector of India due to air pollution.
+Using variation in pollution induced by changes in wind, I show that higher levels of particulate matter lead to decreased agricultural productivity, with changes in productivity of 10 percent being common.
+Combining wind direction with the rollout of coal plants, results indicate pollution from coal plants has a larger effect on agricultural productivity than other types of pollution.
 Given that the agricultural sector is a refuge for the poor in many developing countries, these results suggest that the effects of pollution may hit the poorest particularly hard.
@@
 library(viridis)
+library(stars)
 loadfonts()  # Must be run once in each session (markdown is a new "session" each time you render/knit)
@@
 The rest of the paper is organized as follows.
-I first discuss data and methodology.
-I break this section up into \autoref{data2}, which covers the multiple sources of data used in this paper; \autoref{identification}, which discusses the empirical strategy and validates the wind direction variable using data on particulate matter.
+I first discuss data and then methodology, in two separate sections.
+The methodology section discusses the empirical strategy and validates the wind direction variable using data on particulate matter.
 I present the main results in \autoref{results} before concluding in \autoref{conclusion}.
@@
 
+windexample2 <- rasterFromXYZ(windexample)
+windexample2 <- mask(windexample2, india_shape)
+windexample2 <- rasterToPoints(windexample2)
+windexample2 <- as.data.frame(windexample2)
 p1 <- ggplot() + 
@@
         geom_sf(data = india_shape, fill = NA) +
-        geom_tile(data = windexample, aes(x = x, y = y, fill = wind_example), alpha = 0.3) +
+        geom_tile(data = windexample2, aes(x = x, y = y, fill = wind_example), alpha = 0.3) +
         scale_fill_viridis_c("wind direction\n(degrees)", option = "viridis") +
@@
              subtitle = "B. Aggregation to district") +
-        theme(legend.position = c(0.5, 0.05),
+        theme(legend.position = c(0.45, 0.05),
               legend.direction = "horizontal",
-              legend.key.width = unit(0.3, 'cm'),
+              legend.key.width = unit(0.8, 'cm'),
               legend.text = element_text(size = 8),
-              legend.background=element_blank() )
+              legend.background=element_blank() ) +
+        guides(fill = guide_colourbar(title.position="top", title.hjust = 0.5))
 

### 2023-04-12 draft/draft.tex
updating some figures

@@
 \author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
-\date{2023-04-11}
+\date{2023-04-12}
 
@@
 \begin{abstract}
-\noindent I document large negative externalities in the agricultural sector of India due to air pollution.
-Using random variation in pollution exposure induced by changes in wind direction, I show that higher levels of particulate matter lead to decreased agricultural productivity.
-Based on within-village average deviations in pollution exposure, year-to-year changes in productivity can be above 10 percent.
-Combining wind direction with the rollout of coal plants, I show that pollution from coal plants, specifically, has a larger effect on agricultural productivity than other types of pollution.
+\noindent I document externalities in the agricultural sector of India due to air pollution.
+Using variation in pollution induced by changes in wind, I show that higher levels of particulate matter lead to decreased agricultural productivity, with changes in productivity of 10 percent being common.
+Combining wind direction with the rollout of coal plants, results indicate pollution from coal plants has a larger effect on agricultural productivity than other types of pollution.
 Given that the agricultural sector is a refuge for the poor in many developing countries, these results suggest that the effects of pollution may hit the poorest particularly hard.\\

### 2023-04-13 analyses/5.icrisat.R
Updating readme

@@ -1,84 +1,0 @@
-# Purpose: This script estimates effects with the NSS data. I save the results to load into the markdown script
-# Author: Josh Merfeld
-# Date: February 25th, 2023
-
-rm(list = ls())
-
-library(tidyverse)
-library(lubridate)
-library(readxl)
-
-
-
-# Sets wd to folder this script is in so we can create relative paths instead of absolute paths
-setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
-# We want to get the wd up to the main folder
-# Need to go up two levels
-setwd("../../")
-# Double check
-getwd()    # check
-
-
-
-df <- read_xlsx("/Users/Josh/Dropbox/Datasets/VDSA/EastIndia/2011/Cultivation/Cult_ip.xlsx")
-df <- rbind(df, read_xlsx("/Users/Josh/Dropbox/Datasets/VDSA/EastIndia/2012/Cultivation/Cult_ip.xlsx"))
-df$year <- paste0("20", substr(df$Sur_mon_yr, 4, 5))
-df$month <- substr(df$Sur_mon_yr, 1, 2)
-df$day <- "01"
-df$date <- as_date(paste0(df$year, "-", df$month, "-", df$day))
-
-
-
-
-# create new variables based on operations
-# first to lower
-df$Operation <- str_to_lower(df$Operation)
-df$op_plough <- str_detect(df$Operation, "plough")
-df$op_transplant <- str_detect(df$Operation, "transpl")
-df$op_sowing <- str_detect(df$Operation, "sow")
-df$op_fertilizer <- str_detect(df$Operation, "fertiliz")
-df$op_weeding <- str_detect(df$Operation, "weed")
-df$op_threshing <- str_detect(df$Operation, "threshing")
-df$op_harvest <- str_detect(df$Operation, "harvest")
-df <- df %>% filter(Season=="Kharif")
-
-df <- df %>%
-      group_by(date) %>%
-      mutate(
-             op_plough = mean(op_plough),
-             op_transplant = mean(op_transplant),
-             op_sowing = mean(op_sowing),
-             op_fertilizer = mean(op_fertilizer),
-             op_weeding = mean(op_weeding),
-             op_threshing = mean(op_threshing),
-             op_harvest = mean(op_harvest)
-             ) %>%
-      filter(row_number()==1) %>%
-      ungroup() %>%
-      dplyr::select(date, op_plough, op_transplant, op_sowing, op_fertilizer, op_weeding, op_threshing, op_harvest)
-
-
-
-
-
[…]

### 2023-04-13 analyses/5.nightlights.R

@@ -1,138 +1,0 @@
-# Purpose: This script estimates effects. I save the results to load into the markdown script
-# Author: Josh Merfeld
-# Date: December 12th, 2022
-
-rm(list = ls())
-
-library(tidyverse)
-library(lubridate)
-library(viridis)
-library(fixest)
-library(sf)
-
-
-# Sets wd to folder this script is in so we can create relative paths instead of absolute paths
-setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
-# We want to get the wd up to the main folder
-# Need to go up two levels
-setwd("../../")
-# Double check
-getwd()    # check
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-# Nightlights
-ntl <- read_csv("data/clean/village_ntl.csv")
-ntl <- ntl %>% mutate(state = substr(shrid, 1, 2), 
-                      wind = sums/365,
-                      rain_z = (log(rain) - mean(log(rain))/sd(log(rain))),
-                      village = shrid
-                      )
-
-pollution <- as_tibble(read_csv(paste0("data/clean/pm25/ntl_merge.csv")))
-ntl <- pollution %>% left_join(ntl, by = c("shrid", "year"))
-# ntltemp <- read_csv(paste0("data/clean/village_temp.csv"))
-# ntl <- ntl %>% left_join(ntltemp, by = c("shrid", "year"))
-
-# And district identifiers
-villages_overlap <- read_sf("data/spatial/villages_overlap/villages_overlap.shp")
-villages_overlap <- villages_overlap %>% mutate(shrid = paste0(pc11_s_id, "-", pc11_tv_id),
-                                                distfe = paste0(pc11_s_id, "-", pc11_d_id)) %>%
-                                          as_tibble() %>%
-                                          dplyr::select(shrid, distfe) %>%
-                                          group_by(shrid) %>%
-                                          filter(row_number()==1) %>%
-                                          ungroup()
-ntl <- ntl %>% left_join(villages_overlap, by = "shrid")
-ntl$pm25 <- log(ntl$pm25)
-
-ntl <- panel(ntl, ~ shrid + year, duplicate.method = "first")
-
-
-
-
-
-ntl1 <- feols(log(total_light + 1) ~ 1 | village + year | l(pm25, 1) ~ l(wind, 1),
-              data = ntl,
[…]

### 2023-04-13 cleaning/4.cleaning_plants.R

@@ -184,7 +184,7 @@ rad_deg <- function(rad){
 
 # GET PREVIOUS SEVEN DAYS (before start of loop)
 # Date
-date <- as.Date("2014-08-21", "%Y-%m-%d")
+date <- as.Date("1990-01-01", "%Y-%m-%d")
 final_date <- as.Date("2015-03-10", "%Y-%m-%d")
 
 wind_last_week <- dist_matrix %>% dplyr::select("shrid", "plant_id", "angle", "angle_original", "year_built", "year_retired")
@@ -338,6 +338,7 @@ for (lag in 1:7){
   wind_last_week_merge <- wind_last_week_merge %>% left_join(wind_last_week_temp2, by = "shrid")
   wind_last_week_merge$sum[is.na(wind_last_week_merge$sum)==T] <- 0
   wind_last_week_ALL[, 6 + lag] <- wind_last_week_merge$sum
+  
   # now those during years coal plant actually in operation
   wind_last_week_temp <- wind_last_week_temp[dl_year>=wind_last_week_temp$year_built & dl_year<=wind_last_week_temp$year_retired,]
   wind_last_week_temp <- wind_last_week_temp %>% group_by(shrid) %>% mutate(sum = max(sum)) %>% filter(row_number()==1) %>% ungroup() %>% dplyr::select("shrid", "sum")

### 2023-04-13 presentation/slides/pollution-slides.Rmd
slides and readme

--- 2023-04-07 presentation/slides/slides.Rmd
@@
 # Table
+yield3ivmain <- yield3ivmain[,-1]
 colnames(yield3ivmain) <- c("(1)", "(2)", "(3)")
@@
 colnames(labor3month) <- c("farm", "non-farm")
-labor3month <- labor3month[-c(1:2, 13:14),]
-labor3month <- labor3month[c(1:14),]
 kable(
@@
 colnames(yield4monthly) <- c("(1)", "(2)")
-yield4monthly <- yield4monthly[c(1:14),]
+yield4monthly <- yield4monthly[c(1:14, nrow(yield4monthly)),]
 kable(

### 2023-07-10 draft-short/draft.Rmd
updated draft; labor intensity

--- 2023-04-12 draft/draft.Rmd
@@
 ---
-title: "Air pollution and agricultural productivity in a developing country^[I thank Ariel Ortiz-Bobea, Marc Bellemare, and David Sungho Park for comments and suggestions that greatly improved the quality of the paper.
-The usual disclaimer applies.
-All scripts used in this paper are publicly available on my GitHub page: github.com/JoshMerfeld/pollution_development]"
+title: Air pollution and agricultural productivity in a developing country^[I thank Ariel Ortiz-Bobea, Marc Bellemare, and David Sungho Park for comments and suggestions that greatly improved the quality of the paper.
+The usual disclaimer applies.] # "All scripts used in this paper are publicly available on my GitHub page: github.com/JoshMerfeld/pollution_development"
 author:
@@
   | \noindent I document externalities in the agricultural sector of India due to air pollution.
-Using variation in pollution induced by changes in wind, I show that higher levels of particulate matter lead to decreased agricultural productivity, with changes in productivity of 10 percent being common.
+Using variation in pollution induced by changes in wind, I show that higher levels of pollution lead to decreased agricultural productivity, with large changes in productivity being common.
 Combining wind direction with the rollout of coal plants, results indicate pollution from coal plants has a larger effect on agricultural productivity than other types of pollution.
@@
 # Sets wd to folder this script is in so we can create relative paths instead of absolute paths
-setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
+#setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
 # We want to get the wd up to the main folder
@@
 yield3ivmain <- readRDS("pollution_development/draft/tables/yield3ivmain.rds")
+yield3ivmain_firststage <- readRDS("pollution_development/draft/tables/yield3ivmain_firststage.rds")
+yield3ivmain_firststage <- yield3ivmain_firststage[,-1]
 yield3ivmain_lead <- readRDS("pollution_development/draft/tables/yield3ivmain_lead.rds")
@@
 
-There are no doubt positive effects from this increase in power generation; for example, electricity may lead to economic and productivity growth [@dinkelman2011effects; @kline2014local; @rud2012electricity; @van2017long], though electrification alone may not be a sufficient condition [@lee2020does].
-These improvements do not come without downsides, however, especially when it comes to power derived from coal.
+While there are clear positive effects from this increase in power generation -- for example, electricity may lead to economic and productivity growth [@dinkelman2011effects; @kline2014local; @rud2012electricity; @van2017long], though electrification alone may not be a sufficient condition [@lee2020does] -- these improvements do not come without downsides, especially when it comes to power derived from coal.
 Pollutants from coal contribute to climate change and the phase out of coal energy is therefore a major climate goal [@iea2022].
@@
 
-In addition to coal, pollution more generally is increasing in much of the world.
+In addition to pollution from coal, pollution more generally is also increasing in much of the world.
 Carbon dioxide equivalent per capita has almost double over the last 50 years in India, more than doubled in Indonesia, and almost tripled in China.^[https://edgar.jrc.ec.europa.eu] Of course, much of this increase is due to the development process itself; as countries develop, their economies transform, with more people moving out of agriculture and into manufacturing and services [@herrendorf2014growth].
@@
 
-Using wind direction as an instrument for particulate matter, an increase of PM 2.5 by one percent leads to a large decrease in agricultural productivity, at around 0.73 percent.
-Given within-village changes in exposure, pollution could be responsible for swings of 7-to-14 percent in agricultural productivity across years.
+Using wind direction as an instrument for particulate matter, an increase of PM 2.5 by one percent leads to a large decrease in agricultural productivity, at around 0.44 percent in the most conservative specification.
 The IV estimates indicate that the effect of pollution is more negative than naïve OLS estimates, which is consistent with the evidence on the construction of coal plants; higher levels of pollution are correlated with higher levels of agricultural productivity, a relationship that the IV strategy adjusts downwards.
@@
 
-I find that these effects are mostly transmitted by reductions in farm labor during narrow windows of the growing season, using labor allocation data from the National Sample Survey (NSS).
-While there do not appear to be large changes in labor allocation patterns when looking at the season as a whole, there is a clear pattern when we break down exposure to pollution by month: higher levels of pollution at the beginning of the monsoon season (June and July) lead to lower levels of farm labor allocation, but no changes in non-farm labor allocation.
-This matches up well with a similar exercise with agricultural yields: pollution in June is the only month in which there is a negative effect of pollution on agricultural yield.
-In tandem, these two results point to effects on labor as one mechanism driving the total effects, though I cannot rule out direct effects of pollution on land productivity with my data.
-
 The final set of analyses tries to disentangle the effects of pollution from the effects of coal pollution, specifically, by using wind direction and the rollout of coal plants across the country.
-Combining instrumental variables with a differences-in-differences strategy, I show that agricultural productivity has an elasticity of around -0.15 with respect to non-coal pollution.
-The opening of a coal plant, however, triples the elasticity, to more than -0.45, indicating that coal plant emissions are particularly harmful to agricultural productivity, at least in India.
+Combining instrumental variables with a differences-in-differences strategy, I show that agricultural productivity has an elasticity of around -0.22 with respect to non-coal pollution.
+The opening of a coal plant, however, doubles the elasticity, to more than -0.44, indicating that coal plant emissions are particularly harmful to agricultural productivity, at least in India.
 This is consistent with coal emissions being quite different from other forms of pollution; for example, coal emissions contain much more mercury than other forms of pollution.
+
+An important caveat to these results is that I cannot disentangle exactly which pollutants drive results.
+While I use PM 2.5 as a measure of overall exposure to pollution, effects could be driven by any pollutants carried by the wind.
+This leaves open an important avenue of future research regarding the effects of different types of pollutants on agricultural productivity.
 
@@
 Finally, @sanders2022adaptation actually find that coal emissions can be good for agriculture in the United States, due to the presence of sulfur, which can increase agricultural productivity.
-This paper differs in three important ways.
+This paper differs from the previous literature in three important ways.
 First, I explore effects of pollution more generally as well as from coal plants, which are much more ubiquitous than gold mines across the world.
-Second, I use a different identification strategy from either paper, relying on wind blowing pollution from coal plants.
+Second, I use a different identification strategy from most other papers, relying on wind blowing pollution from coal plants.
 Finally, agriculture in India and the United States can be quite different, with the latter being much more mechanized.
 This means effects of pollution could differ substantially.
-
-I also contribute to the more general literature on the effects of pollution on health.
-Pollution has substantial negative effects on health in both developing and developed countries [@arceo2016does] and we have known of these negative effects for many years [@brunekreef2002air; @kampa2008human; @pope2006health].
-@currie2014we review the literature on the health effects of pollution during adolescence; we have robust evidence of pollution exposure leading to increases in infant mortality, including in developing countries [@heft2018robust; @heft2019air].
-Since these effects are often driven by heart and lung disease and asthma [@kampa2008human], this my explain the labor allocation effects I identify in this paper.
 
@@
 
-To help understand some of the mechanisms driving the relationship between exposure to emissions and agricultural productivity, I also include household survey data.
-Unfortunately, I am not aware of any publicly available survey data that covers multiple years and has identifiers below the district level.
-For example, the ARIS-REDS data[^11] includes GPS identifiers for one year, but that only allows GPS matching for a total of two years: 1999 and 2007/2008.
-The DHS, on the other hand, only has state-level identifiers (except for the newest two waves) and is also not a labor survey.
-Instead, I use the National Sample Survey (NSS) from the Ministry of Statistics and Programme Implementation.
-
-[^11]: www.ncaer.org/project/additional-rural-incomes-survey-aris-rural-economic-demographic-survey-reds
-
-The NSS is a nationally representative survey that was conducted every couple of years until relatively recently.
-Many rounds of the NSS also included a module on employment that asked about labor allocation over the previous seven days.
-I use the 61st, 62nd, 64th, 66th, and 68th rounds of the NSS, all of which include the employment module.
-The publicly available data also includes the date of survey, which allows me to match pollution exposure to the exact day of the survey.
-
-The biggest issue with using the NSS is that it includes only district-level identifiers.
-The main exposure variables, on the other hand, are defined at a lower level of aggregation, at the village.
-I choose to aggregate this village-level exposure to the district level as shown in the second panel of Figure \@ref(fig:windexample) -- the district is a randomly selected district in the data, Guna district in Madhya Pradesh.
-I use the area overlap of the villages with the district to define a weighted average of exposure for the district at a given time.
-I calculate the area of overlap for each village and the district as well as the overall area of the district and use these variables to define weights for individual villages.
-Any part of the district that does not overlap with one of the villages -- the white area in Figure \@ref(fig:windexample) -- does not have any exposure and, as such, receives a value of zero for exposure.
-
 # Methods
@@
 I am primarily interested in the effects of pollution on agricultural productivity. We might consider a regression of the following form:
-\begin{gather} y_{it} = \alpha_{i} + \gamma_{t} + \beta PM_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it}, \end{gather}
+\begin{gather} y_{it} = \alpha_{i} + \gamma_{t} + \beta PM_{it} + f(rain_{it}, temp_{it}) + \varepsilon_{it}, \end{gather}
 
 where $y_{it}$ is outcome $y$ for unit $i$ at time $t$, $\alpha$ is a geographic fixed effects, $\gamma_t$ is time fixed effects, $PM_{it}$ is the (log) concentration of particulate matter 2.5, $rain_{it}$ is total rainfall during the season, and $temp_{it}$ is average temperature during the season.
+In different specifications, I use different forms for $f()$; the simplest specification uses just rainfall and temperature, but in more flexible specifications I turn rainfall into a binned variable -- based on a z score -- for each month of the monsoon season and include all of the bins along with interactions with temperature and squared temperature in each month.
 
@@
 
-So what are the characteristics of areas near the construction of new coal plants?
-Table \@ref(tab:plantresultstable) presents four separate regressions, using state-level fixed effects due to the spatial variation in coal plants.
+What are the characteristics of areas near the construction of new coal plants?
+Table \@ref(tab:plantresultstable) in the appendix presents four separate regressions, using state-level fixed effects due to the spatial variation in coal plants.
 The first two columns use 1991 census data -- downloaded from SHRUG -- and the second two columns use 2001 census data, all at the village level.
@@
 Villages with higher populations and higher literacy in 1991 are more likely to be near a coal plant at the time.
-
-```{r plantresultstable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-# Table
-colnames(plantresultstable) <- c("1991", "2001", "2001", "2011")
-rownames(plantresultstable) <- c("Population (log)", "", "Area (log)", "", "Literacy (prop)", "", "Ag. productivity", "", "Pollution", "(log PM 2.5)",
-                                "Sub-sample", "Observations")
-kable(
-      plantresultstable, caption = "Local characteristics and the construction of coal plants",
-      align = "cccc", booktabs = TRUE, linesep = "", escape = FALSE
-      ) %>%
-  footnote(
-           "Note: State fixed effects are included in regressions.
-Standard errors are in parentheses and clustered at the state level.
-The outcome in the first column is whether the village is within 30km of a coal plant in 1991.
-The second column is whether a village in 1991 will have a coal plant in 2001, conditional on not having one in 1991.
-The last two columns are similarly defined, except using 2001 and 2011 as the years.
-* p<0.10 ** p<0.05 *** p<0.01",
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE,
-           escape = FALSE
-           ) %>%
-  column_spec(1, width = "3cm") %>%
-  column_spec(c(2:5),width = "2cm") %>%
-  row_spec(11, hline_after = TRUE) %>%
-  add_header_above(c(" ", "1991 census" = 2, "2001 census" = 2)) %>%
-  kable_classic_2()
-
-```
 
@@
 Consider the reduced form effect of wind exposure on agricultural productivity:
-\begin{gather} y_{it} = \alpha_{i} + \gamma_{t} + \beta wind_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it}, \end{gather}
+\begin{gather} y_{it} = \alpha_{i} + \gamma_{t} + \beta wind_{it} + f(rain_{it}, temp_{it}) + \varepsilon_{it}, \end{gather}
 
@@
 The differences in weather patterns is very context specific and can differ even for villages located near one another.
-Nonetheless, adding rainfall and temperature should alleviate some of these concerns.
+Nonetheless, adding rainfall and temperature should alleviate some of these concerns, especially the most flexible specifications.
 
@@
 The outcome variable is particulate matter -- specifically, $\mathrm{PM 2.5}$, which is particulate matter no larger than 2.5 micrometers in diameter -- which comes from @hammer2020global.
-$\mathrm{PM 2.5}$ is one of the harmful byproducts from coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.[^12] This data is available at the monthly level, so in Table \@ref(tab:pollutiontable), I aggregate wind exposure to the monthly level, as well.
+$\mathrm{PM 2.5}$ is one of the harmful byproducts from coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.[^12] The main results look at productivity during the main monsoon season, so in Table \@ref(tab:pollutiontable), I aggregate wind exposure to the season.
 
@@
 
-Both the particulate matter data and the wind data is at the village level, which means the regressions in Table \@ref(tab:pollutiontable) are at the village level.
-As such, the geographic fixed effects are village fixed effects and the temporal fixed effects are month fixed effects.
+Both the particulate matter data and the wind data is at the village level, which means the regressions in Table \@ref(tab:pollutiontable) are at the village/season level.
+As such, the geographic fixed effects are village fixed effects and the temporal fixed effects are season/year fixed effects.
 Since the exposure variable is at the village level and we follow villages over time, the standard errors are also clustered at the village level.
-In columns two and four, I switch the month fixed effects for district-month fixed effects, which considers only villages in the same district in the same month.
-
-In addition to serving as an example, Table \@ref(tab:pollutiontable) also helps validate the use of wind direction as a proxy for exposure to pollution.
-The first two columns include all available years of data while the last two columns restrict estimation only to the years for which the agricultural productivity regressions are estimated.
-Across all four columns, the story is the same: when the wind is blowing in the direction of a village from a coal plant within 30 kilometers, estimated $\mathrm{PM 2.5}$ is substantially higher.
+Columns one through four include more and more flexible specifications for the weather variable.
+The last column includes the most flexible specification, which turns rainfall into a binned variable for each month of the season and interacts these bins with temperature and temperature squared, also in each month of the season.
+
+```{r pollutiontable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+# Table
+colnames(yield3ivmain_firststage) <- c("(1)", "(2)", "(3)", "(4)")
+
+kable(
+      yield3ivmain_firststage, caption = "Wind direction and pollution during the monsoon season",
+      align = "cccc", booktabs = TRUE, linesep = ""
+      ) %>%
+  footnote(
+           "Note: Standard errors are in parentheses and are clustered at the village level.
+The outcome in all regressions is average pollution during the monsoon season.
+* p<0.10 ** p<0.05 *** p<0.01",
+           general_title = "",
+           threeparttable = TRUE,
+           footnote_as_chunk = TRUE
+           ) %>%
+  column_spec(1, width = "4cm") %>%
+  column_spec(c(2:5),width = "2.5cm") %>%
+  row_spec(6, bold = TRUE) %>%
+  row_spec(8, hline_after = TRUE) %>%
+  #add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
+  kable_classic_2()
+
+```
+
+Across all columns, the coefficient on wind direction is quite consistent, ranging from 0.0083 to 0.0086.
+These coefficients are also very precisely estimated, leading to t-statistics of more than 30 in all columns.
+Wind direction is clearly a strong predictor of pollution exposure across the season, even conditioning on flexible specifications for weather.
 
@@
 \begin{gather} 
-y_{it} = \alpha_{i} + \gamma_{t} + \omega PM_{it} + \phi open_{it} + \beta PM_{it}\times open_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it},
+y_{it} = \alpha_{i} + \gamma_{t} + \omega PM_{it} + \phi open_{it} + \beta PM_{it}\times open_{it} + f(rain_{it}, temp_{it}) + \varepsilon_{it},
 \end{gather}
@@
 
-I first present the reduced form results for agricultural productivity.
+I first present the naïve results for agricultural productivity, regression productivity on pollution.
 The outcome of all the regressions in this section is the log of agricultural land productivity, which is defined as tons per hectare.
-The first set of results are in Table \@ref(tab:yieldtable).
+These results are in Table \@ref(tab:yieldtablepm).
 The unit of analysis is the village-year -- from 2002 to 2013 -- focusing only on the monsoon (kharif) season.
-The data used to construct the yield measure is calculated using five months of data, so the wind variable is defined for the same five months and is the sum of days in which the wind is pointed towards the village centroid from a coal plant.
-
-The first column presents the most simple results, with only the wind exposure variable and the fixed effects.
-The second column adds a rainfall and average temperature variable, while the third column includes higher-order weather variables, as well.
-While the coefficient decreases slightly in magnitude as we move from the first to the second column and the second to the third column, the decrease is relatively small.
-There is a consistent, negative effect of wind exposure on agricultural productivity.
-
-```{r yieldtable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+The data used to construct the yield measure is calculated using five months of data, and the pollution variable is defined as average pollution during the entire season at the village centroid.
+
+```{r yieldtablepm, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
 
 # Table
-colnames(yield1reducedform) <- c("(1)", "(2)", "(3)")
-
-rownames(yield1reducedform) <- c("Wind days", " ", "Rain (z)", "", "Mean temperature (10s)", "",
-                                  "Rain squared", " ", "Temperature squared", " ", "Rain times Temperature", " ",
-                                  "Fixed effects:", "Village", "Year",
-                                  "Observations")
+yield2naive <- yield2naive[,-1]
+colnames(yield2naive) <- c("(1)", "(2)", "(3)")
+rownames(yield2naive)[4] <- "weather (expanded, bins)"
+yield2naive[4,3] <- "Yes"
 kable(
-      yield1reducedform, caption = "Wind direction and agricultural productivity",
+      yield2naive, caption = "Particulate matter (PM 2.5) and agricultural productivity",
       align = "ccc", booktabs = TRUE, linesep = ""
@@
   column_spec(c(2:4),width = "2.5cm") %>%
-  row_spec(13, bold = TRUE) %>%
-  row_spec(15, hline_after = TRUE) %>%
+  row_spec(5, bold = TRUE) %>%
+  row_spec(7, hline_after = TRUE) %>%
   #add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
@@
 
-To put the size of the coefficient in context, it is worth digging a little deeper into the exposure variable.
-The mean within-village absolute deviation in wind exposure is approximately 2.42 days, meaning that the average change in agricultural productivity from year-to-year due to nothing but changes in wind patterns carrying particulate matter is around 1.5 percent relative to the mean.
-This implies that swings in agricultural productivity -- due to wind direction -- of several percent are probably quite common, since the absolute deviation includes deviations below and above the mean.
-
-```{r yieldtablepm, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+The first column presents the most simple results, with only the wind exposure variable and the fixed effects.
+The second column adds a rainfall and average temperature variable, while the third column includes the most flexible weather specification.
+There is a consistent, negative effect of wind exposure on agricultural productivity.
+Since both variables are logged, the coefficients are similar to elasticities, indicating that a one-percent increase in pollution leads to a 0.045 percent decrease in agricultural productivity, on average, across the three columns.
+
+Recall, however, the results from above: coal plants were placed in places with higher average agricultural productivity.
+This means that the simple regressions in Table \@ref(tab:yieldtablepm) might actually be biased upwards: although pollution might negatively affect productivity, pollution is higher in more productive areas due to the endogenous placement of coal plants in the study areas.
+To recover the actual effect of pollution, we need an instrument.
+If, conditional on the fixed effects and weather variables, wind direction is exogenous and only affects agricultural productivity through pollution, then it is a valid instrument for pollution.
+
+I present the IV results in Table \@ref(tab:yieldtabletwo), with the first-stage F-statistic at the bottom.
+The first-stage results make clear how strong of an instrument wind is; all three F-statistics are above 1,000, meeting even the higher bar for the first-stage F-test argued for by @lee2022valid.
+Also note that the coefficients are quite different from those in Table \@ref(tab:pollutiontable) because the unit of analysis in Table \@ref(tab:yieldtabletwo) is the season, while in Table \@ref(tab:pollutiontable) the unit of analysis is the month.
+
+```{r yieldtabletwo, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
 
 # Table
-colnames(yield2naive) <- c("(1)", "(2)", "(3)")
-rownames(yield2naive) <- c("Particulate matter", "(log PM 2.5)", 
-                           "Weather", "Weather (expanded)",
-                           "Fixed effects:", "Village", "Year",
-                           "Observations")
+yield3ivmain <- yield3ivmain[,-1]
+colnames(yield3ivmain) <- c("(1)", "(2)", "(3)", "(4)")
+
 kable(
-      yield2naive, caption = "Particulate matter (PM 2.5) and agricultural productivity",
-      align = "ccc", booktabs = TRUE, linesep = ""
+      yield3ivmain, caption = "Pollution and agricultural productivity, IV estimates",
+      align = "cccc", booktabs = TRUE, linesep = ""
       ) %>%
@@
   column_spec(1, width = "4cm") %>%
-  column_spec(c(2:4),width = "2.5cm") %>%
-  row_spec(5, bold = TRUE) %>%
-  row_spec(7, hline_after = TRUE) %>%
+  column_spec(c(2:5),width = "2cm") %>%
+  row_spec(6, bold = TRUE) %>%
+  row_spec(8, hline_after = TRUE) %>%
   #add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
@@
 
-Of course, it is not wind, per se, that we are interested in.
-Instead, we are actually interested in better understanding the effects of exposure to pollution on agricultural productivity.
-Table \@ref(tab:yieldtablepm) presents the same regressions as Table \@ref(tab:yieldtable), with a key difference: instead of wind exposure, it uses log average concentration of $\mathrm{PM 2.5}$ -- from @hammer2020global -- during the season as the key explanatory variable.
-There is again a consistent, negative relationship between particulate matter and agricultural productivity.
-Since both variables are logged, the coefficients are similar to elasticities, indicating that a one-percent increase in pollution leads to a 0.03-to-0.04 percent decrease in agricultural productivity.
-
-Recall, however, the results from above: coal plants were placed in places with higher average agricultural productivity.
-This means that the simple regressions in Table \@ref(tab:yieldtablepm) might actually be biased upwards: although pollution might negatively affect productivity, pollution is higher in more productive areas due to the endogenous placement of coal plants in the study areas.
-To recover the actual effect of pollution, we need an instrument.
-If, conditional on the fixed effects and weather variables, wind direction is exogenous and only affects agricultural productivity through pollution, then it is a valid instrument for pollution.
-
-I present the IV results in Table \@ref(tab:yieldtabletwo), with the first-stage F-statistic at the bottom.
-The first-stage results make clear how strong of an instrument wind is; all three F-statistics are above 1,000, meeting even the higher bar for the first-stage F-test argued for by @lee2022valid.
-Also note that the coefficients are quite different from those in Table \@ref(tab:pollutiontable) because the unit of analysis in Table \@ref(tab:yieldtabletwo) is the season, while in Table \@ref(tab:pollutiontable) the unit of analysis is the month.
-
 The second-stage results first indicate that $\mathrm{PM 2.5}$ is a very strong predictor of agricultural productivity.
-Using the results in column three, a one percent increase in particulate matter leads to a 0.73 percent decrease in agricultural productivity.
-To put this in further perspective, the mean particulate matter concentration is just short of 40 (39.68) in the monsoon sample, while the within-village absolute deviation in particulate matter is 3.98, meaning year-to-year swings of 7-to-14 percent in agricultural productivity due to pollution exposure are common.
+Using the results in column four -- again, the most flexible weather specification -- a one-percent increase in particulate matter leads to a 0.44 percent decrease in agricultural productivity.
+To put this in further perspective, the mean particulate matter concentration is just short of 40 (39.68) in the monsoon sample, while the within-village absolute deviation in particulate matter is 3.98, meaning year-to-year swings of 4-to-7 percent in agricultural productivity due to pollution exposure are common.
 
 The comparison of the results to those in the previous table also confirm the concerns around endogenous placement of coal plants.
-The coefficients in Table \@ref(tab:yieldtabletwo) are always more negative than those in Table \@ref(tab:yieldtablepm), and sometimes substantially so.
+The coefficients in Table \@ref(tab:yieldtabletwo) are always more negative than those in Table \@ref(tab:yieldtablepm), and substantially so.
 In other words, the simple regression of agricultural productivity on pollution results in coefficients on pollution that are upwards biased, since higher levels of pollution occur in more productive areas, on average.
@@
 
-```{r yieldtabletwo, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+## Leads for wind and pollution
+
+In this section, I report the results of a robustness check to help validate the use of wind as an instrument for pollution.
+Specifically, I use leads of wind and particulate matter and show that these do not predict current agricultural productivity.
+I present these results in \@ref(tab:yieldtableleads).
+
+```{r yieldtableleads, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
 
 # Table
-yield3ivmain <- yield3ivmain[,-1]
-colnames(yield3ivmain) <- c("(1)", "(2)", "(3)")
-rownames(yield3ivmain) <- c("Particulate matter", "(log PM 2.5)",
-                            "Weather", "Weather (expanded)",
-                            "Fixed effects:", "Village", "Year",
-                            "F (first stage)", 
-                            "Observations")
+colnames(yield3ivmain_lead) <- c("(1)", "(2)")
 kable(
-      yield3ivmain, caption = "Pollution and agricultural productivity, IV estimates",
-      align = "ccc", booktabs = TRUE, linesep = ""
+      yield3ivmain_lead, caption = "Agricultural productivity and pollution leads",
+      align = "cc", booktabs = TRUE, linesep = ""
+      ) %>%
+  footnote(
+           "Note: Standard errors are in parentheses and are clustered at the village level.
+The outcome in all regressions estimated agricultural yield (tons per hectare).
+* p<0.10 ** p<0.05 *** p<0.01",
+           general_title = "",
+           threeparttable = TRUE,
+           footnote_as_chunk = TRUE
+           ) %>%
+  column_spec(1, width = "3.5cm") %>%
+  column_spec(c(2:3),width = "2cm") %>%
+  row_spec(c(7), hline_after = TRUE) %>%
+  row_spec(c(5), bold = TRUE) %>%
+  kable_classic_2()
+
+```
+
+The first column looks at the effects of a one-year lead, while the second column looks at the effects of a two-year lead.
+In both cases, the coefficient is much smaller and insignificant than the main results in \@ref(tab:yieldtabletwo).
+Although the coefficients are negative, they are less than 20 percent the magnitude of the previous results.
+In addition, the first-stage F-statistic shows that the instrument continues to be strong, despite the loss of observations when using leads.
+
+## Coal plants or overall pollution?
+
+Finally, while I identify high-pollution areas using the location of coal plants, an important question is how much of the overall effect is driven by coal plants and how much is driven by other polluting industries located in areas near coal plants.
+I try to answer this question in Table \@ref(tab:yield7ivdiffindiff), where I implement an instrumental variables differences-in-differences, essentially instrumenting for pollution prior to the opening of a coal plant with wind and instrumenting for the interaction term (pollution times plant opening) with wind times plant opening.
+
+```{r yield7ivdiffindiff, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+# Table
+colnames(yield7ivdiffindiff) <- c("(1)", "(2)")
+
+kable(
+      yield7ivdiffindiff, caption = "Agricultural productivity and pollution source",
+      align = "cc", booktabs = TRUE, linesep = ""
       ) %>%
@@
            ) %>%
-  column_spec(1, width = "4cm") %>%
-  column_spec(c(2:4),width = "2cm") %>%
-  row_spec(5, bold = TRUE) %>%
-  row_spec(7, hline_after = TRUE) %>%
-  #add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
-  kable_classic_2()
-
-```
-
-```{r randomplot, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 8, fig.cap = "Randomization inference", fig.footer = list("Note: The figure presents the distribution of coefficients by randomly assigning wind direction across villages within the same year.
-The \"all\" distribution is created by randomly assigning a single wind value to each village, sampled with replacement across all villages in a given year.
-The \"within district\" randomly samples with replacement only from within the same district in the same year.")}
-
-pal <- viridis(4)
-colors <- c(paste0(pal[1]), paste0(pal[2]))
-labels <- c("All", "Within district")
-
-# plot
-gg1 <- ggplot() + 
-        geom_density(aes(x = coef_vec, fill = "All"), alpha = 0.5) +
-        geom_density(aes(x = coefs_distfe, fill = "Within district"), alpha = 0.5) +
-        scale_fill_manual(
-          name = "Randomization:",
-          values = colors,
-          labels = labels
-        ) +
-        theme_minimal() +
-        xlab("Distribution of coefficients") +
-        ylab("Density")
-
-colors <- c(paste0(pal[3]))
-labels <- c("actual")
-
-gg1 <- gg1 + 
-        geom_vline(aes(xintercept = true_value, color = "actual")) +
-        scale_color_manual(
-          name = "",
-          values = colors,
-          labels = labels
-        ) +
-        theme(legend.position = c(0.4, 0.6))
-
-gg1
-
-```
-
-## Robustness checks
-
-In this section, I report the results of two separate robustness checks for the main results presented above.
-In the previous results, standard errors are clustered at the level of the village, which is also the level at which treatment (exposure to pollution) is assigned as well as the level at which agricultural productivity is extracted.
-Past work has shown that standard errors might be underestimated in such contexts [@bertrand2004much].
-This is a particular concern in the present case, since agricultural productivity is extracted at the village level but estimated using less coarse data that has high spatial correlation [@gangopadhyay2022new].
-
-To help remedy concerns, I perform randomization inference (e.g.
-@barrios2012clustering) by randomly assigning observed treatment across villages and repeating this process 1,000 times.
-I do this in two separate ways.
-First, I take all villages in a given year and randomly assign, with replacement, treatment intensity within the year.
-Second, I do this separately by selecting from within district-years only, to help take into account district-year correlation in outcomes and as a way to mitigate possible concerns that traditional clustered standard errors may actually be much too conservative [@abadie2023should].
-Since this is an instrumental variables procedure, I randomly select the wind and pollution variable _together_.
-
-I present these results in Figure \@ref(fig:randomplot).
-The purple density is the resulting distribution of estimated treatment effects when selecting from all villages within the same year.
-The density of treatment effects is relatively tight around 0, with the left tail only getting slightly beyond -0.05, compared to the most conservative actual estimated treatment effect of -0.73.
-
-The within-district correlation in outcomes and treatment is apparent when instead randomly selecting from only villages within the same district in the same year.
-The distribution is still approximately normal but has a mode of almost -0.4.
-Nonetheless, the actual treatment effect is larger in magnitude than every single one of the 1,000 bootstrapped replications.
-If we were to use the standard deviation of the resulting distribution as a standard error, we would get a confidence interval around the main effect of approximately `r paste0(reportDec(-0.73-1.96*sd(coefs_distfe)))` to `r paste0(reportDec(-0.73+1.96*sd(coefs_distfe)))`.
-The results are apparently robust to alternative specifications of uncertainty
-
-The second robustness check includes leads of pollution and wind direction.
-I present these results in Table \@ref(tab:yieldtableleads).
-Interestingly, there is actually a positive coefficient for the lead in the second column.
-However, if we perform randomization inference, it appears that the confidence intervals might be slightly underestimated; the positive effect size is in the 11th percentile of the distribution in Figure \@ref(fig:randomplotleads), which for a two-tailed test would not even have a p-value of less than 0.2.
-Overall, there is no indication that leads predict any changes in current agricultural productivity.
-
-```{r yieldtableleads, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-# Table
-colnames(yield3ivmain_lead) <- c("(1)", "(2)")
-rownames(yield3ivmain_lead) <- c("Particulate matter", "(log PM 2.5)",
-                                 "Particulate matter", "(one-year lead)",
-                                 "Weather",
-                                 "Fixed effects:", "Village", "Year",
-                                 "F (current)", "F (lead)", 
-                                 "Observations")
-kable(
-      yield3ivmain_lead, caption = "Agricultural productivity and pollution leads",
-      align = "cc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the village level.
-The outcome in all regressions estimated agricultural yield (tons per hectare).
-* p<0.10 ** p<0.05 *** p<0.01",
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
-           ) %>%
   column_spec(1, width = "3.5cm") %>%
   column_spec(c(2:3),width = "2cm") %>%
-  row_spec(c(8), hline_after = TRUE) %>%
-  row_spec(c(6), bold = TRUE) %>%
-  kable_classic_2()
-
-```
-
-## Heterogeneity
-
-Having established that exposure to pollution has a negative effect on agricultural productivity, I now move to heterogeneity in these effects.
-Table \@ref(tab:yieldtablehet) presents three different analyses: heterogeneity by median exposure, heterogeneity by initial yield, and possible multiplicative effects of shocks.
-All regressions use the IV strategy detailed above and I include the F statistic from the first stage regressions (of which there are two in the last column) at the bottom of the table.
-
-The first two columns split the sample based on median exposure.
-In other words, for each village, I calculate the median exposure (based on wind) and then use this single value for each village to split the sample into villages above and below median exposure.
-The coefficient in the first column is 2.3 times larger than the coefficient in the second column, indicating much larger effects of pollution in places where it is more common.
-One possible explanation is that exposure is cumulative, leading to larger and larger effects.
-I return to the effects of cumulative shocks in column five.
-
-Columns three and four split the sample based on median yield at the very beginning of the sample (the first year available, 2002-2003).
-Column three presents results for villages above the median while column four presents results for villages below the median.
-The largest effects are apparently concentrated in the upper half of the distribution, meaning the most productive areas are also most susceptible to the effects of pollution.
-This could have important implications for the overall effects of the placement of any point sources of pollution.
-
-Given ongoing changes in the climate, a key question is how welfare will change as shocks become more and more common.
-The last column looks at possible multiplicative effects of shocks by interacting pollution with rainfall and seeing whether the effects compound one another.
-Since there are now two endogenous regressors, I instrument for both using wind exposure and an interaction between wind exposure and rainfall.
-
-```{r yieldtablehet, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-# Table
-colnames(yield5heterogeneity) <- c(">p(50)", "<=p(50)", ">p(50)", "<=p(50)", "")
-rownames(yield5heterogeneity)[c(1, 3)] <- c("PM 2.5", "PM times rain")
-rownames(yield5heterogeneity) <- c("Particulate matter", "(log PM 2.5)", "PM times Rain", "", "Weather",
-                                   "Fixed effects:", "Village", "Year", 
-                                   "F (1st stage, PM)", "F (1st stage, PM times rain)", 
-                                   "Observations")
-kable(
-      yield5heterogeneity, caption = "Heterogeneity in the effects of pollution on productivity",
-      align = "ccccc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the village level.
-The outcome in all regressions is estimated agricultural yield, defined as tons per hectare.
-The first two columns split the sample based on the median exposure (by village).
-The second two columns split the sample based on initial -- the first year available -- agricultural productivity.
-* p<0.10 ** p<0.05 *** p<0.01",
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
-           ) %>%
-  column_spec(1, width = "4.5cm") %>%
-  column_spec(c(2:6),width = "1.8cm") %>%
-  row_spec(c(8), hline_after = TRUE) %>%
-  row_spec(c(6), bold = TRUE) %>%
-  add_header_above(c(" " = 1, "Wind" = 2, "Yield" = 2, " " = 1)) %>%
-  kable_classic_2()
-
-```
-
-The coefficient on the level term for pollution is negative.
-Since rainfall is defined as a z score, this represents the effect of pollution only when rainfall is at its mean value for a given season.
-The interaction term, on the other hand, is positive.
-Since higher values of rainfall are generally good, this positive coefficient represents a compounding effect of pollution and rainfall; when rainfall is low (bad), the effects of particulate matter on agricultural productivity is larger (more negative).
-Another way to think about this, however, is that higher values of rainfall dampen the effects of pollution shocks.
-This could have very profound effects on welfare, since this essentially increases the variance of agricultural productivity across years (which is already increasing due to climate change), at least as it relates to pollution exposure..
-
-## Labor allocation
-
-The previous set of results show a consistent negative effect of wind-driven pollution on agricultural productivity.
-A key question is the mechanism through which this happens; is this driven by changes in labor allocation/productivity or by changes in land productivity?
-While I cannot definitely prove the mechanism, I can shed some light on possibilities by looking at labor allocation decisions.
-Since these shocks generally happen after initial planting decisions, it seems unlikely that changes in land *allocation* could drive the results.
-
-Table \@ref(tab:labortable) presents results for labor allocation over the last seven days using the National Sample Survey and the methodology described above.
-Since the labor recall is for seven days, labor allocation ranges from zero to seven, as does the wind variable.^[I also tried an instrumental variables strategy with the seven-day wind and monthly pollution, but the first stage lacked power due to the difference in temporal aggregation across the two key variables (pollution and wind).] Recall that the data is at the district level, not the village level.
-However, it is also more temporally disaggregated, with rolling interviews throughout the year-long survey.
-Figure \@ref(fig:laborplot) in the appendix shows this temporal distribution across the five waves of the survey.
-Since pollution is not disaggregated at a daily level, I instead stick to the reduced form results with wind.
-
-```{r labortable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-# Table
-colnames(labor1all) <- c("All", "All", "Self", "Wage", "Farm", "Non-farm")
-colnames(labor2rural) <- c("All", "All", "Self", "Wage", "Farm", "Non-farm")
-
-mat1 <- matrix(c("", "", "", "", "", ""), nrow = 1)
-colnames(mat1) <- c("All", "All", "Self", "Wage", "Farm", "Non-farm")
-rownames(mat1) <- c("Panel A: all")
-mat2 <- matrix(c("", "", "", "", "", ""), nrow = 1)
-colnames(mat2) <- c("All", "All", "Self", "Wage", "Farm", "Non-farm")
-rownames(mat2) <- c("Panel B: Rural")
-
-labortablerural <- rbind(mat1, labor1all[c(1:2, 7:8),], mat2, labor2rural[-c(3:6),], labor2rural[c(3:6),])
-rownames(labortablerural)[c(2,7)] <- "Wind"
-rownames(labortablerural)[c(5,10)] <- "Observations"
-rownames(labortablerural)[11] <- "Controls"
-rownames(labortablerural)[12] <- "Fixed effects:"
-rownames(labortablerural)[13] <- "District"
-rownames(labortablerural)[14] <- "State-month"
-
-kable(
-      labortablerural, caption = "Wind direction and labor allocation by rural status",
-      align = "cccccc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the district level.
-Control variables include female, age, age squared, and (years of) education.
-The outcome in each regression is days worked in the previous seven days.
-* p<0.10 ** p<0.05 *** p<0.01",
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
-           ) %>%
-  column_spec(1, width = "3cm") %>%
-  column_spec(c(2:7),width = "1.5cm") %>%
-  row_spec(c(5,10), hline_after = TRUE) %>%
-  row_spec(c(1,6,12), bold = TRUE) %>%
-  kable_classic_2()
-
-```
-
-The top panel presents results for all respondents during the five-month monsoon season while the bottom panel presents results for respondents in rural areas.
-Interestingly, we do not see any signs of large, negative effects of wind exposure on labor allocation across any sectors, with the possible exception of wage employment in rural areas.
-However, the coefficient for self-employment in rural areas is of the same magnitude and *positive*, but imprecisely estimated, making it hard to come to any definitive conclusions, at least in the aggregate.
-
-## Month-specific exposure
-
-While overall results for labor allocation are inconclusive, perhaps we can learn more by looking at temporally disaggregated results, focusing on June through October.
-I can create separate variables for exposure by each month of the season, keeping the unit of analysis the individual (for labor) or the village-season (for yields).
-
-Since the NSS data has survey dates and the both the pollution and wind data can be aggregated to the month, we can look at how exposure during different months of the agricultural season (differentially) affect labor allocation and overall agricultural productivity.
-I begin with labor allocation in columns one and two of Table \@ref(tab:labor3month).
-As before, the exposure measure is just the wind variable (over the last seven days), since I can directly match daily wind direction with the timing of the survey, but I cannot do that for pollution, which is monthly.
-The first column presents results for farm labor and the second column presents results for non-farm labor.
-
-Higher exposure leads to significantly less farm labor labor in June and July, but there are no discernible effects in other months of the year.
-There are likewise no discernible effects on non-farm labor, but some of the coefficients are rather imprecisely estimate.
-To put the effects for June and July in context and taking the coefficients at face value, an additional day of wind in the last seven days decreases farm labor in June by around `r paste0(reportPercent(0.136/1.209))` percent of the mean and in July by around `r paste0(reportPercent(0.285/1.671))` percent of the mean.
-
-```{r labor3month, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-labor3month <- cbind(labor3month, yield4monthly[-c(15:19),2])
-
-# Table
-colnames(labor3month) <- c("Farm", "Non-farm", "Yield")
-kable(
-      labor3month, caption = "Heterogeneity by month",
-      align = "ccc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the village level.
-The outcome in column one is farm labor, the outcome is column two is non-farm labor, and the outcome in column three is village-level productivity.
-The independent variable for labor is wind, while the IV approach above is still used for yield.
-* p<0.10 ** p<0.05 *** p<0.01",
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
-           ) %>%
-  column_spec(1, width = "2.5cm") %>%
-  column_spec(c(2:4),width = "2.5cm") %>%
-  row_spec(c(14), hline_after = TRUE) %>%
-  row_spec(c(12), bold = TRUE) %>%
-  add_header_above(c(" " = 1, "Labor" = 2, " " = 1)) %>%
-  kable_classic_2()
-
-```
-
-I perform a similar exercise with yield, the results of which are in the third column Table \@ref(tab:labor3month).
-I return to the IV strategy, however, to be consistent with the previous yield results.
-I include only the five months of the year that are used in the calculation of the yield estimates, or June through October.
-
-There is only one month where we see a clear effect of pollution on yield: June.
-Given that we also see negative effects of pollution on farm labor allocation in June, this is suggestive -- but not conclusive -- evidence that pollution at least partly works through labor.
-Note that this does not rule out pollution also having direct effects on agricultural land productivity, however.
-
-A final caveat about these month-specific results is that they do not take into account possible interactions between different months.
-For example, yield might respond strongly to average pollution in June and July, which would not be captured in Table \@ref(tab:labor3month).
-As one additional check, Table \@ref(tab:yield8cumulative) in the appendix presents five separate instrumental variable regressions, which look at the effects of cumulative exposure to pollution separately by month.
-Across all five months, cumulative exposure negatively affects agricultural productivity, with the largest coefficient actually showing up for October, the last month of the season.
-
-## Coal plants or overall pollution?
-
-Finally, while I identify high-pollution areas using the location of coal plants, an important question is how much of the overall effect is driven by coal plants and how much is driven by other polluting industries located in areas near coal plants.
-I try to answer this question in Table \@ref(tab:yield7ivdiffindiff), where I implement an instrumental variables differences-in-differences, essentially instrumenting for pollution prior to the opening of a coal plant with wind and instrumenting for the interaction term (pollution times plant opening) with wind times plant opening.
-
-```{r yield7ivdiffindiff, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-# Table
-colnames(yield7ivdiffindiff) <- c("(1)", "(2)", "(3)")
-rownames(yield7ivdiffindiff) <- c("Particulate matter", "(log PM 2.5)", 
-                                  "PM 2.5 times Coal", "",
-                                  "Weather", "Weather (expanded)",
-                                  "Fixed effects:", "Year", "Village",
-                                  "F (PM)", "F (PM times Coal)", "Observations")
-kable(
-      yield7ivdiffindiff, caption = "Agricultural productivity and pollution source",
-      align = "ccc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the village level.
-The outcome in all regressions is estimated agricultural yield, defined as tons per hectare.
-* p<0.10 ** p<0.05 *** p<0.01",
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
-           ) %>%
-  column_spec(1, width = "3.5cm") %>%
-  column_spec(c(2:4),width = "2cm") %>%
   row_spec(c(7), bold = TRUE) %>%
@@
 Column three presents the most extensive set of controls and I focus on that column here.
-First, there are now two first stages but the F-tests indicate that the instruments are strong for both, including the interaction term (F = 232).
+First, there are now two first stages but the F-tests indicate that the instruments are strong for both, including the interaction term.
 Second, the first row shows the effect of pollution on agricultural productivity prior to the opening of a coal plant; the coefficient is negative, indicating negative effects of non-coal pollution.
@@
 In other words, this is not identifying differences in overall pollution, per se, but rather differences in the effects of an increase of pollution by the same percentage.
-These different effects could be due to, for example, differences in emission patterns for different kinds of plants and factories; coal plants, for example, are responsible for much more mercury than other kinds of polluting industries.^[https://www.epa.gov/sites/default/files/2016-12/documents/nei2014v1_tsd.pdf]
+These different effects could be due to, among other things, differences in emission patterns for different kinds of plants and factories; coal plants, for example, are responsible for much more mercury than other kinds of polluting industries.^[https://www.epa.gov/sites/default/files/2016-12/documents/nei2014v1_tsd.pdf]
 
@@
 Using exogenous variation in pollution driven by changes in wind direction, results here show that pollution has a large, negative effect on agricultural productivity in India.
-
-These negative effects are at least partly due to changes in labor allocation decisions; the largest negative effects on agricultural productivity and on labor allocation happen in the same month of the season: June.
-Nonetheless, I am unable to rule out direct effects on land productivity with the data available.
 
@@
 While we often frame the debate around greenhouse gasses in terms of climate change, there are important local reasons to reduce pollution emissions, as well.
-The technology for greener energy exists; at this point, it is simply a question of policies [@markard2018next; blazquez2020some].
+The technology for greener energy exists; at this point, it is simply a question of policies [@markard2018next; @blazquez2020some].
 
@@
 ```
+
+```{r plantresultstable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+# Table
+colnames(plantresultstable) <- c("1991", "2001", "2001", "2011")
+rownames(plantresultstable) <- c("Population (log)", "", "Area (log)", "", "Literacy (prop)", "", "Ag. productivity", "", "Pollution", "(log PM 2.5)",
+                                "Sub-sample", "Observations")
+kable(
+      plantresultstable, caption = "Local characteristics and the construction of coal plants",
+      align = "cccc", booktabs = TRUE, linesep = "", escape = FALSE
+      ) %>%
+  footnote(
+           "Note: State fixed effects are included in regressions.
+Standard errors are in parentheses and clustered at the state level.
+The outcome in the first column is whether the village is within 30km of a coal plant in 1991.
+The second column is whether a village in 1991 will have a coal plant in 2001, conditional on not having one in 1991.
+The last two columns are similarly defined, except using 2001 and 2011 as the years.
+* p<0.10 ** p<0.05 *** p<0.01",
+           general_title = "",
+           threeparttable = TRUE,
+           footnote_as_chunk = TRUE,
+           escape = FALSE
+           ) %>%
+  column_spec(1, width = "3cm") %>%
+  column_spec(c(2:5),width = "2cm") %>%
+  row_spec(11, hline_after = TRUE) %>%
+  add_header_above(c(" ", "1991 census" = 2, "2001 census" = 2)) %>%
+  kable_classic_2()
+
+```
+
+```{=tex}
+\FloatBarrier
+\newpage
+```
+
 ```{r plants, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Coal plants in India from 1990 to 2010", fig.footer = list("Note: The top figure shows the location of coal plants in 1990.
@@
 
-```{r windexample, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 8, fig.cap = "Wind direction and aggregation examples (2010-01-01)", fig.footer = list("Note: The top figure shows the average wind direction on January 1st, 2010, with zero degrees indicating north.
-The points are the location of coal plants on that date.
-The bottom figure shows the distribution of pollution exposure in a specific district -- Guna district in Madhya Pradesh -- on the same date.")}
+```{r windexample, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 8, fig.cap = "Wind direction example (2010-01-01)", fig.footer = list("Note: The figure shows the average wind direction on January 1st, 2010, with zero degrees indicating north.
+The points are the location of coal plants on that date.")}
 
@@
         labs(x = "lon",
-             y = "lat",
-             subtitle = "A. Wind direction") +
+             y = "lat") +
         theme(legend.position = c(0.6, 0.05),
@@
 
-ggarrange(p1, p2,
-          nrow = 2)
+p1
 
@@
 ```
-
-```{r pollutiontable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-# Table
-colnames(pollutiontable) <- c("(1)", "(2)", "(3)", "(4)")
-kable(
-      pollutiontable, caption = "Wind direction and particulate matter",
-      align = "cccc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
-           ) %>%
-  column_spec(1, width = "4cm") %>%
-  column_spec(c(2:5),width = "2cm") %>%
-  row_spec(c(3), bold = TRUE) %>%
-  row_spec(6, hline_after = TRUE) %>%
-  add_header_above(c(" ", "1998-2015" = 2, "2002-2013" = 2)) %>%
-  kable_classic_2()
-
-```
-
-```{=tex}
-\FloatBarrier
-\newpage
-```
-
-```{r randomplotleads, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 8, fig.cap = "Randomization inference for leads", fig.footer = list("Note: The figure presents the distribution of coefficients by randomly assigning a single wind value to each village, sampled with replacement across all villages in a given district in the same year.
-\"Current\" refers to the coefficient on current pollution, while \"lead\" refers to the coefficient on next year's pollution in the same table.")}
-
-pal <- viridis(4)
-colors <- c(paste0(pal[1]), paste0(pal[3]))
-labels <- c("current", "lead")
-
-# plot
-ggplot() + 
-    geom_density(aes(x = coef_current, fill = "current"), alpha = 0.5) +
-    geom_density(aes(x = coef_lead, fill = "lead"), alpha = 0.5) +
-    scale_fill_manual(
-      name = "coefficient:",
-      values = colors,
-      labels = labels
-    ) +
-    theme_minimal() +
-    xlab("distribution of coefficients")
-
-```
-
-```{=tex}
-\FloatBarrier
-\newpage
-```
-```{r laborplot, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Timing of household surveys for the National Sample Survey", fig.footer = list("Note: The histogram shows the distribution of survey dates for the four waves of the National Sample Survey (NSS) for the five waves used in this paper.")}
-
-laborplot
-
-```
-
-```{=tex}
-\FloatBarrier
-\newpage
-```
-```{r monthplot, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Average farm days by season, rural only", fig.footer = list("Note: The line graph shows the mean number of farm days worked per week based on the month of the relevant agricultural season.
-The monsoon season is defined as June (month 1) to October (month 5) while the rainy season is defined as November (month 1) to March (month 5) of the following year.
-The data is from the NSS.")}
-
-monthplot
-
-```
-
-```{=tex}
-\FloatBarrier
-\newpage
-```
-```{r yield8cumulative, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-# Table
-colnames(yield8cumulative) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
-kable(
-      yield8cumulative, caption = "Cumulative exposure to pollution during the season",
-      align = "ccccc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
-           ) %>%
-  column_spec(1, width = "2.25cm") %>%
-  column_spec(c(2:6),width = "2cm") %>%
-  row_spec(c(12), bold = TRUE) %>%
-  row_spec(14, hline_after = TRUE) %>%
-  kable_classic_2()
-
-```

### 2023-07-10 draft-short/draft.tex
updated draft; labor intensity

--- 2023-04-12 draft/draft.tex
@@
 \usepackage{amsmath,amssymb}
-\usepackage{lmodern}
 \usepackage{iftex}
@@
 \else % if luatex or xetex
-  \usepackage{unicode-math}
+  \usepackage{unicode-math} % this also loads fontspec
   \defaultfontfeatures{Scale=MatchLowercase}
   \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else
+  % xetex/luatex font selection
 \fi
@@
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
+\urlstyle{same}
 \hypersetup{
@@
 \title{Air pollution and agricultural productivity in a developing country\footnote{I thank Ariel Ortiz-Bobea, Marc Bellemare, and David Sungho Park for comments and suggestions that greatly improved the quality of the paper.
-The usual disclaimer applies.
-All scripts used in this paper are publicly available on my GitHub page: github.com/JoshMerfeld/pollution\_development}}
+The usual disclaimer applies.}}
 \author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
-\date{2023-04-12}
+\date{2023-07-03}
 
@@
 \noindent I document externalities in the agricultural sector of India due to air pollution.
-Using variation in pollution induced by changes in wind, I show that higher levels of particulate matter lead to decreased agricultural productivity, with changes in productivity of 10 percent being common.
+Using variation in pollution induced by changes in wind, I show that higher levels of pollution lead to decreased agricultural productivity, with large changes in productivity being common.
 Combining wind direction with the rollout of coal plants, results indicate pollution from coal plants has a larger effect on agricultural productivity than other types of pollution.
@@
 
-There are no doubt positive effects from this increase in power generation; for example, electricity may lead to economic and productivity growth (Dinkelman 2011; Kline and Moretti 2014; Rud 2012; Van de Walle et al.
+While there are clear positive effects from this increase in power generation -- for example, electricity may lead to economic and productivity growth (Dinkelman 2011; Kline and Moretti 2014; Rud 2012; Van de Walle et al.
 2017), though electrification alone may not be a sufficient condition (K.
-Lee, Miguel, and Wolfram 2020).
-These improvements do not come without downsides, however, especially when it comes to power derived from coal.
+Lee, Miguel, and Wolfram 2020) -- these improvements do not come without downsides, especially when it comes to power derived from coal.
 Pollutants from coal contribute to climate change and the phase out of coal energy is therefore a major climate goal (IEA 2022).
@@
 
-In addition to coal, pollution more generally is increasing in much of the world.
+In addition to pollution from coal, pollution more generally is also increasing in much of the world.
 Carbon dioxide equivalent per capita has almost double over the last 50 years in India, more than doubled in Indonesia, and almost tripled in China.\footnote{\url{https://edgar.jrc.ec.europa.eu}} Of course, much of this increase is due to the development process itself; as countries develop, their economies transform, with more people moving out of agriculture and into manufacturing and services (Herrendorf, Rogerson, and Valentinyi 2014).
@@
 Finally, Sanders and Barreca (2022) actually find that coal emissions can be good for agriculture in the United States, due to the presence of sulfur, which can increase agricultural productivity.
-This paper differs in three important ways.
+This paper differs from the previous literature in three important ways.
 First, I explore effects of pollution more generally as well as from coal plants, which are much more ubiquitous than gold mines across the world.
-Second, I use a different identification strategy from either paper, relying on wind blowing pollution from coal plants.
+Second, I use a different identification strategy from most other papers, relying on wind blowing pollution from coal plants.
 Finally, agriculture in India and the United States can be quite different, with the latter being much more mechanized.
 This means effects of pollution could differ substantially.
-
-I also contribute to the more general literature on the effects of pollution on health.
-Pollution has substantial negative effects on health in both developing and developed countries (Arceo, Hanna, and Oliva 2016) and we have known of these negative effects for many years (Brunekreef and Holgate 2002; Kampa and Castanas 2008; Pope III and Dockery 2006).
-Currie et al.
-(2014) review the literature on the health effects of pollution during adolescence; we have robust evidence of pollution exposure leading to increases in infant mortality, including in developing countries (Heft-Neal et al.
-2018, 2019).
-Since these effects are often driven by heart and lung disease and asthma (Kampa and Castanas 2008), this my explain the labor allocation effects I identify in this paper.
 
@@
 
-To help understand some of the mechanisms driving the relationship between exposure to emissions and agricultural productivity, I also include household survey data.
-Unfortunately, I am not aware of any publicly available survey data that covers multiple years and has identifiers below the district level.
-For example, the ARIS-REDS data\footnote{www.ncaer.org/project/additional-rural-incomes-survey-aris-rural-economic-demographic-survey-reds} includes GPS identifiers for one year, but that only allows GPS matching for a total of two years: 1999 and 2007/2008.
-The DHS, on the other hand, only has state-level identifiers (except for the newest two waves) and is also not a labor survey.
-Instead, I use the National Sample Survey (NSS) from the Ministry of Statistics and Programme Implementation.
-
-The NSS is a nationally representative survey that was conducted every couple of years until relatively recently.
-Many rounds of the NSS also included a module on employment that asked about labor allocation over the previous seven days.
-I use the 61st, 62nd, 64th, 66th, and 68th rounds of the NSS, all of which include the employment module.
-The publicly available data also includes the date of survey, which allows me to match pollution exposure to the exact day of the survey.
-
-The biggest issue with using the NSS is that it includes only district-level identifiers.
-The main exposure variables, on the other hand, are defined at a lower level of aggregation, at the village.
-I choose to aggregate this village-level exposure to the district level as shown in the second panel of Figure \ref{fig:windexample} -- the district is a randomly selected district in the data, Guna district in Madhya Pradesh.
-I use the area overlap of the villages with the district to define a weighted average of exposure for the district at a given time.
-I calculate the area of overlap for each village and the district as well as the overall area of the district and use these variables to define weights for individual villages.
-Any part of the district that does not overlap with one of the villages -- the white area in Figure \ref{fig:windexample} -- does not have any exposure and, as such, receives a value of zero for exposure.
-
 \hypertarget{methods}{%
@@
 I am primarily interested in the effects of pollution on agricultural productivity. We might consider a regression of the following form:
-\begin{gather} y_{it} = \alpha_{i} + \gamma_{t} + \beta PM_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it}, \end{gather}
+\begin{gather} y_{it} = \alpha_{i} + \gamma_{t} + \beta PM_{it} + f(rain_{it}, temp_{it}) + \varepsilon_{it}, \end{gather}
 
 where \(y_{it}\) is outcome \(y\) for unit \(i\) at time \(t\), \(\alpha\) is a geographic fixed effects, \(\gamma_t\) is time fixed effects, \(PM_{it}\) is the (log) concentration of particulate matter 2.5, \(rain_{it}\) is total rainfall during the season, and \(temp_{it}\) is average temperature during the season.
+In different specifications, I use different forms for \(f()\); the simplest specification uses just rainfall and temperature, but in more flexible specifications I turn rainfall into a binned variable -- based on a z score -- for each month of the monsoon season and include all of the bins along with interactions with temperature and squared temperature in each month.
 
@@
 
-So what are the characteristics of areas near the construction of new coal plants?
-Table \ref{tab:plantresultstable} presents four separate regressions, using state-level fixed effects due to the spatial variation in coal plants.
+What are the characteristics of areas near the construction of new coal plants?
+Table \ref{tab:plantresultstable} in the appendix presents four separate regressions, using state-level fixed effects due to the spatial variation in coal plants.
 The first two columns use 1991 census data -- downloaded from SHRUG -- and the second two columns use 2001 census data, all at the village level.
@@
 Villages with higher populations and higher literacy in 1991 are more likely to be near a coal plant at the time.
-
-\begin{table}
-
-\caption{\label{tab:plantresultstable}Local characteristics and the construction of coal plants}
-\centering
-\begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{1991 census} & \multicolumn{2}{c}{2001 census} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
-  & 1991 & 2001 & 2001 & 2011\\
-\midrule
-Population (log) & 0.009 & 0.007* & 0.013* & 0.005\\
- & (0.006) & (0.004) & (0.006) & (0.004)\\
-Area (log) & -0.007* & -0.004 & -0.007* & -0.004\\
- & (0.004) & (0.005) & (0.004) & (0.003)\\
-Literacy (prop) & 0.080** & 0.037 & 0.073* & 0.036\\
- & (0.029) & (0.028) & (0.039) & (0.026)\\
-Ag. productivity &  &  & 0.014** & 0.028***\\
- &  &  & (0.005) & (0.009)\\
-Pollution &  &  & 0.047 & 0.014\\
-(log PM 2.5) &  &  & (0.045) & (0.018)\\
-Sub-sample & all & no plant & all & no plant\\
-\midrule
-Observations & 283,021 & 264,265 & 508,784 & 469,004\\
-\bottomrule
-\end{tabular}
-\begin{tablenotes}[para]
-\item Note: State fixed effects are included in regressions.
-Standard errors are in parentheses and clustered at the state level.
-The outcome in the first column is whether the village is within 30km of a coal plant in 1991.
-The second column is whether a village in 1991 will have a coal plant in 2001, conditional on not having one in 1991.
-The last two columns are similarly defined, except using 2001 and 2011 as the years.
-* p<0.10 ** p<0.05 *** p<0.01
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
 
@@
 (2020).
-\(\mathrm{PM 2.5}\) is one of the harmful byproducts from coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.\footnote{www.epa.gov/airmarkets/power-plants-and-neighboring-communities} This data is available at the monthly level, so in Table \ref{tab:pollutiontable}, I aggregate wind exposure to the monthly level, as well.
-
-Both the particulate matter data and the wind data is at the village level, which means the regressions in Table \ref{tab:pollutiontable} are at the village level.
-As such, the geographic fixed effects are village fixed effects and the temporal fixed effects are month fixed effects.
+\(\mathrm{PM 2.5}\) is one of the harmful byproducts from coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.\footnote{www.epa.gov/airmarkets/power-plants-and-neighboring-communities} The main results look at productivity during the main monsoon season, so in Table \ref{tab:pollutiontable}, I aggregate wind exposure to the season.
+
+Both the particulate matter data and the wind data is at the village level, which means the regressions in Table \ref{tab:pollutiontable} are at the village/season level.
+As such, the geographic fixed effects are village fixed effects and the temporal fixed effects are season/year fixed effects.
 Since the exposure variable is at the village level and we follow villages over time, the standard errors are also clustered at the village level.
-In columns two and four, I switch the month fixed effects for district-month fixed effects, which considers only villages in the same district in the same month.
-
-In addition to serving as an example, Table \ref{tab:pollutiontable} also helps validate the use of wind direction as a proxy for exposure to pollution.
-The first two columns include all available years of data while the last two columns restrict estimation only to the years for which the agricultural productivity regressions are estimated.
-Across all four columns, the story is the same: when the wind is blowing in the direction of a village from a coal plant within 30 kilometers, estimated \(\mathrm{PM 2.5}\) is substantially higher.
+Columns one through four include more and more flexible specifications for the weather variable.
+The last column includes the most flexible specification, which turns rainfall into a binned variable for each month of the season and interacts these bins with temperature and temperature squared, also in each month of the season.
+
+\begin{table}
+
+\caption{\label{tab:pollutiontable}Wind direction and pollution during the monsoon season}
+\centering
+\begin{threeparttable}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}}
+\toprule
+  & (1) & (2) & (3) & (4)\\
+\midrule
+wind & 0.0083*** & 0.0085*** & 0.0083*** & 0.0086***\\
+ & (0.0002) & (0.0002) & (0.0002) & (0.0002)\\
+weather & No & Yes & Yes & Yes\\
+weather (expanded) & No & No & Yes & No\\
+weather (expanded, bins) & No & No & No & Yes\\
+\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+village & Yes & Yes & Yes & Yes\\
+year & Yes & Yes & Yes & Yes\\
+\midrule
+F & 1,236 & 1,266 & 1,205 & 1,264\\
+observations & 1,266,922 & 1,266,922 & 1,266,922 & 1,266,922\\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}[para]
+\item Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions is average pollution during the monsoon season. * p<0.10 ** p<0.05 *** p<0.01
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+
+Across all columns, the coefficient on wind direction is quite consistent, ranging from 0.0083 to 0.0086.
+These coefficients are also very precisely estimated, leading to t-statistics of more than 30 in all columns.
+Wind direction is clearly a strong predictor of pollution exposure across the season, even conditioning on flexible specifications for weather.
 
@@
 
-I first present the reduced form results for agricultural productivity.
+I first present the naïve results for agricultural productivity, regression productivity on pollution.
 The outcome of all the regressions in this section is the log of agricultural land productivity, which is defined as tons per hectare.
-The first set of results are in Table \ref{tab:yieldtable}.
+These results are in Table \ref{tab:yieldtablepm}.
 The unit of analysis is the village-year -- from 2002 to 2013 -- focusing only on the monsoon (kharif) season.
-The data used to construct the yield measure is calculated using five months of data, so the wind variable is defined for the same five months and is the sum of days in which the wind is pointed towards the village centroid from a coal plant.
-
-The first column presents the most simple results, with only the wind exposure variable and the fixed effects.
-The second column adds a rainfall and average temperature variable, while the third column includes higher-order weather variables, as well.
-While the coefficient decreases slightly in magnitude as we move from the first to the second column and the second to the third column, the decrease is relatively small.
-There is a consistent, negative effect of wind exposure on agricultural productivity.
+The data used to construct the yield measure is calculated using five months of data, and the pollution variable is defined as average pollution during the entire season at the village centroid.
 
@@
 
-\caption{\label{tab:yieldtable}Wind direction and agricultural productivity}
+\caption{\label{tab:yieldtablepm}Particulate matter (PM 2.5) and agricultural productivity}
 \centering
@@
 \midrule
-Wind days & -0.0078*** & -0.0069*** & -0.0062***\\
- & (0.0007) & (0.0007) & (0.0006)\\
-Rain (z) &  & 0.0854*** & -0.0424***\\
- &  & (0.0018) & (0.0046)\\
-Mean temperature (10s) &  & -0.0021*** & 0.1321***\\
- &  & (0.0001) & (0.0034)\\
-Rain squared &  &  & -0.0010***\\
- &  &  & (0.0001)\\
-Temperature squared &  &  & -0.0002***\\
- &  &  & \vphantom{1} (0.0000)\\
-Rain times Temperature &  &  & 0.0003***\\
- &  &  & (0.0000)\\
-\textbf{Fixed effects:} & \textbf{} & \textbf{} & \textbf{}\\
-Village & Yes & Yes & Yes\\
-Year & Yes & Yes & Yes\\
-\midrule
-Observations & 1,266,922 & 1,266,922 & 1,266,922\\
+particulate matter & -0.042*** & -0.045*** & -0.046***\\
+(log PM 2.5) & (0.003) & (0.003) & (0.003)\\
+weather & No & Yes & Yes\\
+weather (expanded, bins) & No & No & Yes\\
+\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{}\\
+village & Yes & Yes & Yes\\
+year & Yes & Yes & Yes\\
+\midrule
+observations & 1,266,922 & 1,266,922 & 1,266,922\\
 \bottomrule
@@
 
-To put the size of the coefficient in context, it is worth digging a little deeper into the exposure variable.
-The mean within-village absolute deviation in wind exposure is approximately 2.42 days, meaning that the average change in agricultural productivity from year-to-year due to nothing but changes in wind patterns carrying particulate matter is around 1.5 percent relative to the mean.
-This implies that swings in agricultural productivity -- due to wind direction -- of several percent are probably quite common, since the absolute deviation includes deviations below and above the mean.
+The first column presents the most simple results, with only the wind exposure variable and the fixed effects.
+The second column adds a rainfall and average temperature variable, while the third column includes the most flexible weather specification.
+There is a consistent, negative effect of wind exposure on agricultural productivity.
+Since both variables are logged, the coefficients are similar to elasticities, indicating that a one-percent increase in pollution leads to a 0.045 percent decrease in agricultural productivity, on average, across the three columns.
+
+Recall, however, the results from above: coal plants were placed in places with higher average agricultural productivity.
+This means that the simple regressions in Table \ref{tab:yieldtablepm} might actually be biased upwards: although pollution might negatively affect productivity, pollution is higher in more productive areas due to the endogenous placement of coal plants in the study areas.
+To recover the actual effect of pollution, we need an instrument.
+If, conditional on the fixed effects and weather variables, wind direction is exogenous and only affects agricultural productivity through pollution, then it is a valid instrument for pollution.
+
+I present the IV results in Table \ref{tab:yieldtabletwo}, with the first-stage F-statistic at the bottom.
+The first-stage results make clear how strong of an instrument wind is; all three F-statistics are above 1,000, meeting even the higher bar for the first-stage F-test argued for by D.
+Lee et al.
+(2022).
+Also note that the coefficients are quite different from those in Table \ref{tab:pollutiontable} because the unit of analysis in Table \ref{tab:yieldtabletwo} is the season, while in Table \ref{tab:pollutiontable} the unit of analysis is the month.
 
@@
 
-\caption{\label{tab:yieldtablepm}Particulate matter (PM 2.5) and agricultural productivity}
+\caption{\label{tab:yieldtabletwo}Pollution and agricultural productivity, IV estimates}
 \centering
 \begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
 \toprule
-  & (1) & (2) & (3)\\
-\midrule
-Particulate matter & -0.042*** & -0.044*** & -0.032***\\
-(log PM 2.5) & (0.003) & (0.003) & (0.003)\\
-Weather & No & Yes & Yes\\
-Weather (expanded) & No & No & Yes\\
-\textbf{Fixed effects:} & \textbf{} & \textbf{} & \textbf{}\\
-Village & Yes & Yes & Yes\\
-Year & Yes & Yes & Yes\\
-\midrule
-Observations & 1,266,922 & 1,266,922 & 1,266,922\\
+  & (1) & (2) & (3) & (4)\\
+\midrule
+particulate matter & -0.941*** & -0.769*** & -0.621*** & -0.437***\\
+(log PM 2.5) & (0.078) & (0.075) & (0.080) & (0.075)\\
+weather & No & Yes & Yes & Yes\\
+weather (expanded) & No & No & Yes & No\\
+weather (expanded, bins) & No & No & No & Yes\\
+\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+village & Yes & Yes & Yes & Yes\\
+year & Yes & Yes & Yes & Yes\\
+\midrule
+F & 1,236 & 1,266 & 1,205 & 1,264\\
+observations & 1,266,922 & 1,266,922 & 1,266,922 & 1,266,922\\
 \bottomrule
@@
 
-Of course, it is not wind, per se, that we are interested in.
-Instead, we are actually interested in better understanding the effects of exposure to pollution on agricultural productivity.
-Table \ref{tab:yieldtablepm} presents the same regressions as Table \ref{tab:yieldtable}, with a key difference: instead of wind exposure, it uses log average concentration of \(\mathrm{PM 2.5}\) -- from Hammer et al.
-(2020) -- during the season as the key explanatory variable.
-There is again a consistent, negative relationship between particulate matter and agricultural productivity.
-Since both variables are logged, the coefficients are similar to elasticities, indicating that a one-percent increase in pollution leads to a 0.03-to-0.04 percent decrease in agricultural productivity.
-
-Recall, however, the results from above: coal plants were placed in places with higher average agricultural productivity.
-This means that the simple regressions in Table \ref{tab:yieldtablepm} might actually be biased upwards: although pollution might negatively affect productivity, pollution is higher in more productive areas due to the endogenous placement of coal plants in the study areas.
-To recover the actual effect of pollution, we need an instrument.
-If, conditional on the fixed effects and weather variables, wind direction is exogenous and only affects agricultural productivity through pollution, then it is a valid instrument for pollution.
-
-I present the IV results in Table \ref{tab:yieldtabletwo}, with the first-stage F-statistic at the bottom.
-The first-stage results make clear how strong of an instrument wind is; all three F-statistics are above 1,000, meeting even the higher bar for the first-stage F-test argued for by D.
-Lee et al.
-(2022).
-Also note that the coefficients are quite different from those in Table \ref{tab:pollutiontable} because the unit of analysis in Table \ref{tab:yieldtabletwo} is the season, while in Table \ref{tab:pollutiontable} the unit of analysis is the month.
-
 The second-stage results first indicate that \(\mathrm{PM 2.5}\) is a very strong predictor of agricultural productivity.
-Using the results in column three, a one percent increase in particulate matter leads to a 0.73 percent decrease in agricultural productivity.
-To put this in further perspective, the mean particulate matter concentration is just short of 40 (39.68) in the monsoon sample, while the within-village absolute deviation in particulate matter is 3.98, meaning year-to-year swings of 7-to-14 percent in agricultural productivity due to pollution exposure are common.
+Using the results in column four -- again, the most flexible weather specification -- a one-percent increase in particulate matter leads to a 0.44 percent decrease in agricultural productivity.
+To put this in further perspective, the mean particulate matter concentration is just short of 40 (39.68) in the monsoon sample, while the within-village absolute deviation in particulate matter is 3.98, meaning year-to-year swings of 4-to-7 percent in agricultural productivity due to pollution exposure are common.
 
 The comparison of the results to those in the previous table also confirm the concerns around endogenous placement of coal plants.
-The coefficients in Table \ref{tab:yieldtabletwo} are always more negative than those in Table \ref{tab:yieldtablepm}, and sometimes substantially so.
+The coefficients in Table \ref{tab:yieldtabletwo} are always more negative than those in Table \ref{tab:yieldtablepm}, and substantially so.
 In other words, the simple regression of agricultural productivity on pollution results in coefficients on pollution that are upwards biased, since higher levels of pollution occur in more productive areas, on average.
@@
 
+\hypertarget{leads-for-wind-and-pollution}{%
+\subsection{Leads for wind and pollution}\label{leads-for-wind-and-pollution}}
+
+In this section, I report the results of a robustness check to help validate the use of wind as an instrument for pollution.
+Specifically, I use leads of wind and particulate matter and show that these do not predict current agricultural productivity.
+I present these results in \ref{tab:yieldtableleads}.
+
 \begin{table}
 
-\caption{\label{tab:yieldtabletwo}Pollution and agricultural productivity, IV estimates}
+\caption{\label{tab:yieldtableleads}Agricultural productivity and pollution leads}
 \centering
 \begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3.5cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
 \toprule
-  & (1) & (2) & (3)\\
-\midrule
-Particulate matter & -0.941*** & -0.813*** & -0.728***\\
-(log PM 2.5) & (0.078) & (0.076) & (0.075)\\
-Weather & No & Yes & Yes\\
-Weather (expanded) & No & No & Yes\\
-\textbf{Fixed effects:} & \textbf{} & \textbf{} & \textbf{}\\
-Village & Yes & Yes & Yes\\
-Year & Yes & Yes & Yes\\
-\midrule
-F (first stage) & 1,236 & 1,263 & 1,264\\
-Observations & 1,266,922 & 1,266,922 & 1,266,922\\
+  & (1) & (2)\\
+\midrule
+particulate matter & -0.073 & \\
+(one-year lead) & (0.066) & \\
+particulate matter &  & -0.079\\
+(two-year lead) &  & (0.051)\\
+\textbf{fixed effects:} & \textbf{} & \textbf{}\\
+village & Yes & Yes\\
+year & Yes & Yes\\
+\midrule
+F & 603 & 746\\
+observations & 1,161,265 & 1,055,562\\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}[para]
+\item Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions estimated agricultural yield (tons per hectare). * p<0.10 ** p<0.05 *** p<0.01
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+
+The first column looks at the effects of a one-year lead, while the second column looks at the effects of a two-year lead.
+In both cases, the coefficient is much smaller and insignificant than the main results in \ref{tab:yieldtabletwo}.
+Although the coefficients are negative, they are less than 20 percent the magnitude of the previous results.
+In addition, the first-stage F-statistic shows that the instrument continues to be strong, despite the loss of observations when using leads.
+
+\hypertarget{coal-plants-or-overall-pollution}{%
+\subsection{Coal plants or overall pollution?}\label{coal-plants-or-overall-pollution}}
+
+Finally, while I identify high-pollution areas using the location of coal plants, an important question is how much of the overall effect is driven by coal plants and how much is driven by other polluting industries located in areas near coal plants.
+I try to answer this question in Table \ref{tab:yield7ivdiffindiff}, where I implement an instrumental variables differences-in-differences, essentially instrumenting for pollution prior to the opening of a coal plant with wind and instrumenting for the interaction term (pollution times plant opening) with wind times plant opening.
+
+\begin{table}
+
+\caption{\label{tab:yield7ivdiffindiff}Agricultural productivity and pollution source}
+\centering
+\begin{threeparttable}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3.5cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\toprule
+  & (1) & (2)\\
+\midrule
+PM 2.5 & -0.228*** & -0.223***\\
+ & (0.066) & (0.065)\\
+PM 2.5 times coal & -0.250*** & -0.220***\\
+ & (0.024) & (0.025)\\
+weather & Yes & No\\
+weather (expanded, bins) & No & Yes\\
+\textbf{fixed effects:} & \textbf{} & \textbf{}\\
+year & Yes & Yes\\
+village & Yes & Yes\\
+\midrule
+F (pm) & 769 & 772\\
+F (pm times open) & 231 & 205\\
+observations & 1,266,922 & 1,266,922\\
 \bottomrule
@@
 
-\begin{figure}
-\includegraphics{draft_files/figure-latex/randomplot-1} \caption[Randomization inference]{Randomization inference}\label{fig:randomplot}\floatfoot*{Note: The figure presents the distribution of coefficients by randomly assigning wind direction across villages within the same year.
-The "all" distribution is created by randomly assigning a single wind value to each village, sampled with replacement across all villages in a given year.
-The "within district" randomly samples with replacement only from within the same district in the same year.}
-\end{figure}
-
-\hypertarget{robustness-checks}{%
-\subsection{Robustness checks}\label{robustness-checks}}
-
-In this section, I report the results of two separate robustness checks for the main results presented above.
-In the previous results, standard errors are clustered at the level of the village, which is also the level at which treatment (exposure to pollution) is assigned as well as the level at which agricultural productivity is extracted.
-Past work has shown that standard errors might be underestimated in such contexts (Bertrand, Duflo, and Mullainathan 2004).
-This is a particular concern in the present case, since agricultural productivity is extracted at the village level but estimated using less coarse data that has high spatial correlation (Gangopadhyay et al.
-2022).
-
-To help remedy concerns, I perform randomization inference (e.g.
-Barrios et al.
-(2012)) by randomly assigning observed treatment across villages and repeating this process 1,000 times.
-I do this in two separate ways.
-First, I take all villages in a given year and randomly assign, with replacement, treatment intensity within the year.
-Second, I do this separately by selecting from within district-years only, to help take into account district-year correlation in outcomes and as a way to mitigate possible concerns that traditional clustered standard errors may actually be much too conservative (Abadie et al.
-2023).
-Since this is an instrumental variables procedure, I randomly select the wind and pollution variable \emph{together}.
-
-I present these results in Figure \ref{fig:randomplot}.
-The purple density is the resulting distribution of estimated treatment effects when selecting from all villages within the same year.
-The density of treatment effects is relatively tight around 0, with the left tail only getting slightly beyond -0.05, compared to the most conservative actual estimated treatment effect of -0.73.
-
-The within-district correlation in outcomes and treatment is apparent when instead randomly selecting from only villages within the same district in the same year.
-The distribution is still approximately normal but has a mode of almost -0.4.
-Nonetheless, the actual treatment effect is larger in magnitude than every single one of the 1,000 bootstrapped replications.
-If we were to use the standard deviation of the resulting distribution as a standard error, we would get a confidence interval around the main effect of approximately -0.899 to -0.561.
-The results are apparently robust to alternative specifications of uncertainty
-
-The second robustness check includes leads of pollution and wind direction.
-I present these results in Table \ref{tab:yieldtableleads}.
-Interestingly, there is actually a positive coefficient for the lead in the second column.
-However, if we perform randomization inference, it appears that the confidence intervals might be slightly underestimated; the positive effect size is in the 11th percentile of the distribution in Figure \ref{fig:randomplotleads}, which for a two-tailed test would not even have a p-value of less than 0.2.
-Overall, there is no indication that leads predict any changes in current agricultural productivity.
-
-\begin{table}
-
-\caption{\label{tab:yieldtableleads}Agricultural productivity and pollution leads}
-\centering
-\begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3.5cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
-\toprule
-  & (1) & (2)\\
-\midrule
-Particulate matter & -1.152*** & -1.067***\\
-(log PM 2.5) & (0.075) & (0.071)\\
-Particulate matter & 0.068 & 0.268***\\
-(one-year lead) & (0.072) & (0.074)\\
-Weather & No & Yes\\
-\textbf{Fixed effects:} & \textbf{} & \textbf{}\\
-Village & Yes & Yes\\
-Year & Yes & Yes\\
-\midrule
-F (current) & 974 & 1,036\\
-F (lead) & 612 & 597\\
-Observations & 1,161,265 & 1,161,265\\
-\bottomrule
-\end{tabular}
-\begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions estimated agricultural yield (tons per hectare). * p<0.10 ** p<0.05 *** p<0.01
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
-\hypertarget{heterogeneity}{%
-\subsection{Heterogeneity}\label{heterogeneity}}
-
-Having established that exposure to pollution has a negative effect on agricultural productivity, I now move to heterogeneity in these effects.
-Table \ref{tab:yieldtablehet} presents three different analyses: heterogeneity by median exposure, heterogeneity by initial yield, and possible multiplicative effects of shocks.
-All regressions use the IV strategy detailed above and I include the F statistic from the first stage regressions (of which there are two in the last column) at the bottom of the table.
-
-The first two columns split the sample based on median exposure.
-In other words, for each village, I calculate the median exposure (based on wind) and then use this single value for each village to split the sample into villages above and below median exposure.
-The coefficient in the first column is 2.3 times larger than the coefficient in the second column, indicating much larger effects of pollution in places where it is more common.
-One possible explanation is that exposure is cumulative, leading to larger and larger effects.
-I return to the effects of cumulative shocks in column five.
-
-Columns three and four split the sample based on median yield at the very beginning of the sample (the first year available, 2002-2003).
-Column three presents results for villages above the median while column four presents results for villages below the median.
-The largest effects are apparently concentrated in the upper half of the distribution, meaning the most productive areas are also most susceptible to the effects of pollution.
-This could have important implications for the overall effects of the placement of any point sources of pollution.
-
-Given ongoing changes in the climate, a key question is how welfare will change as shocks become more and more common.
-The last column looks at possible multiplicative effects of shocks by interacting pollution with rainfall and seeing whether the effects compound one another.
-Since there are now two endogenous regressors, I instrument for both using wind exposure and an interaction between wind exposure and rainfall.
-
-\begin{table}
-
-\caption{\label{tab:yieldtablehet}Heterogeneity in the effects of pollution on productivity}
-\centering
-\begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4.5cm}>{\centering\arraybackslash}p{1.8cm}>{\centering\arraybackslash}p{1.8cm}>{\centering\arraybackslash}p{1.8cm}>{\centering\arraybackslash}p{1.8cm}>{\centering\arraybackslash}p{1.8cm}}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Wind} & \multicolumn{2}{c}{Yield} & \multicolumn{1}{c}{ } \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
-  & >p(50) & <=p(50) & >p(50) & <=p(50) & \\
-\midrule
-Particulate matter & -0.767*** & -0.328*** & -1.130*** & -0.690*** & -0.409**\\
-(log PM 2.5) & (0.104) & (0.095) & (0.060) & (0.146) & (0.160)\\
-PM times Rain &  &  &  &  & 0.061***\\
- &  &  &  &  & (0.014)\\
-Weather & Yes & Yes & Yes & Yes & Yes\\
-\textbf{Fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-Village & Yes & Yes & Yes & Yes & Yes\\
-Year & Yes & Yes & Yes & Yes & Yes\\
-\midrule
-F (1st stage, PM) & 684 & 491 & 688 & 484 & 865\\
-F (1st stage, PM times rain) &  &  &  &  & 1,437\\
-Observations & 617,804 & 649,118 & 634,342 & 632,580 & 1,266,922\\
-\bottomrule
-\end{tabular}
-\begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the village level.
-The outcome in all regressions is estimated agricultural yield, defined as tons per hectare.
-The first two columns split the sample based on the median exposure (by village).
-The second two columns split the sample based on initial -- the first year available -- agricultural productivity.
-* p<0.10 ** p<0.05 *** p<0.01
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
-The coefficient on the level term for pollution is negative.
-Since rainfall is defined as a z score, this represents the effect of pollution only when rainfall is at its mean value for a given season.
-The interaction term, on the other hand, is positive.
-Since higher values of rainfall are generally good, this positive coefficient represents a compounding effect of pollution and rainfall; when rainfall is low (bad), the effects of particulate matter on agricultural productivity is larger (more negative).
-Another way to think about this, however, is that higher values of rainfall dampen the effects of pollution shocks.
-This could have very profound effects on welfare, since this essentially increases the variance of agricultural productivity across years (which is already increasing due to climate change), at least as it relates to pollution exposure..
-
-\hypertarget{labor-allocation}{%
-\subsection{Labor allocation}\label{labor-allocation}}
-
-The previous set of results show a consistent negative effect of wind-driven pollution on agricultural productivity.
-A key question is the mechanism through which this happens; is this driven by changes in labor allocation/productivity or by changes in land productivity?
-While I cannot definitely prove the mechanism, I can shed some light on possibilities by looking at labor allocation decisions.
-Since these shocks generally happen after initial planting decisions, it seems unlikely that changes in land \emph{allocation} could drive the results.
-
-Table \ref{tab:labortable} presents results for labor allocation over the last seven days using the National Sample Survey and the methodology described above.
-Since the labor recall is for seven days, labor allocation ranges from zero to seven, as does the wind variable.\footnote{I also tried an instrumental variables strategy with the seven-day wind and monthly pollution, but the first stage lacked power due to the difference in temporal aggregation across the two key variables (pollution and wind).} Recall that the data is at the district level, not the village level.
-However, it is also more temporally disaggregated, with rolling interviews throughout the year-long survey.
-Figure \ref{fig:laborplot} in the appendix shows this temporal distribution across the five waves of the survey.
-Since pollution is not disaggregated at a daily level, I instead stick to the reduced form results with wind.
-
-\begin{table}
-
-\caption{\label{tab:labortable}Wind direction and labor allocation by rural status}
-\centering
-\begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}}
-\toprule
-  & All & All & Self & Wage & Farm & Non-farm\\
-\midrule
-\textbf{Panel A: all} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-Wind & -0.011 & -0.009 & 0.030 & -0.049 & -0.009 & -0.010\\
- & (0.031) & (0.034) & (0.039) & (0.034) & (0.034) & (0.033)\\
-DV mean & 3.303 & 3.303 & 1.862 & 1.442 & 0.789 & 2.514\\
-Observations & 523,075 & 522,957 & 522,957 & 522,957 & 522,957 & 522,957\\
-\midrule
-\textbf{Panel B: Rural} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-Wind & -0.006 & -0.034 & 0.068 & -0.102*** & -0.024 & -0.010\\
- & (0.044) & (0.040) & (0.043) & (0.035) & (0.043) & (0.036)\\
-DV mean & 3.494 & 3.494 & 2.184 & 1.311 & 1.271 & 2.223\\
-Observations & 307,381 & 307,325 & 307,325 & 307,325 & 307,325 & 307,325\\
-\midrule
-Controls & No & Yes & Yes & Yes & Yes & Yes\\
-\textbf{Fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-District & Yes & Yes & Yes & Yes & Yes & Yes\\
-State-month & Yes & Yes & Yes & Yes & Yes & Yes\\
-\bottomrule
-\end{tabular}
-\begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the district level.
-Control variables include female, age, age squared, and (years of) education.
-The outcome in each regression is days worked in the previous seven days.
-* p<0.10 ** p<0.05 *** p<0.01
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
-The top panel presents results for all respondents during the five-month monsoon season while the bottom panel presents results for respondents in rural areas.
-Interestingly, we do not see any signs of large, negative effects of wind exposure on labor allocation across any sectors, with the possible exception of wage employment in rural areas.
-However, the coefficient for self-employment in rural areas is of the same magnitude and \emph{positive}, but imprecisely estimated, making it hard to come to any definitive conclusions, at least in the aggregate.
-
-\hypertarget{month-specific-exposure}{%
-\subsection{Month-specific exposure}\label{month-specific-exposure}}
-
-While overall results for labor allocation are inconclusive, perhaps we can learn more by looking at temporally disaggregated results, focusing on June through October.
-I can create separate variables for exposure by each month of the season, keeping the unit of analysis the individual (for labor) or the village-season (for yields).
-
-Since the NSS data has survey dates and the both the pollution and wind data can be aggregated to the month, we can look at how exposure during different months of the agricultural season (differentially) affect labor allocation and overall agricultural productivity.
-I begin with labor allocation in columns one and two of Table \ref{tab:labor3month}.
-As before, the exposure measure is just the wind variable (over the last seven days), since I can directly match daily wind direction with the timing of the survey, but I cannot do that for pollution, which is monthly.
-The first column presents results for farm labor and the second column presents results for non-farm labor.
-
-Higher exposure leads to significantly less farm labor labor in June and July, but there are no discernible effects in other months of the year.
-There are likewise no discernible effects on non-farm labor, but some of the coefficients are rather imprecisely estimate.
-To put the effects for June and July in context and taking the coefficients at face value, an additional day of wind in the last seven days decreases farm labor in June by around 11.2 percent of the mean and in July by around 17.1 percent of the mean.
-
-\begin{table}
-
-\caption{\label{tab:labor3month}Heterogeneity by month}
-\centering
-\begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Labor} & \multicolumn{1}{c}{ } \\
-\cmidrule(l{3pt}r{3pt}){2-3}
-  & Farm & Non-farm & Yield\\
-\midrule
-June & -0.128* & 0.123 & -0.736***\\
- & (0.076) & (0.095) & (0.062)\\
-July & -0.278* & 0.185 & 0.051\\
- & (0.142) & (0.133) & (0.095)\\
-Aug. & 0.055 & -0.002 & 0.087\\
- & (0.083) & (0.068) & (0.070)\\
-Sept. & 0.086 & -0.072 & -0.140\\
- & (0.081) & (0.093) & (0.180)\\
-Oct. & 0.022 & -0.074 & 0.334\\
- & (0.071) & (0.057) & (0.291)\\
-controls & Yes & Yes & Yes\\
-\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{}\\
-district & Yes & Yes & Yes\\
-year-month & Yes & Yes & Yes\\
-\midrule
-observations & 307,325 & 307,325 & 1,266,922\\
-\bottomrule
-\end{tabular}
-\begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the village level.
-The outcome in column one is farm labor, the outcome is column two is non-farm labor, and the outcome in column three is village-level productivity.
-The independent variable for labor is wind, while the IV approach above is still used for yield.
-* p<0.10 ** p<0.05 *** p<0.01
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
-I perform a similar exercise with yield, the results of which are in the third column Table \ref{tab:labor3month}.
-I return to the IV strategy, however, to be consistent with the previous yield results.
-I include only the five months of the year that are used in the calculation of the yield estimates, or June through October.
-
-There is only one month where we see a clear effect of pollution on yield: June.
-Given that we also see negative effects of pollution on farm labor allocation in June, this is suggestive -- but not conclusive -- evidence that pollution at least partly works through labor.
-Note that this does not rule out pollution also having direct effects on agricultural land productivity, however.
-
-A final caveat about these month-specific results is that they do not take into account possible interactions between different months.
-For example, yield might respond strongly to average pollution in June and July, which would not be captured in Table \ref{tab:labor3month}.
-As one additional check, Table \ref{tab:yield8cumulative} in the appendix presents five separate instrumental variable regressions, which look at the effects of cumulative exposure to pollution separately by month.
-Across all five months, cumulative exposure negatively affects agricultural productivity, with the largest coefficient actually showing up for October, the last month of the season.
-
-\hypertarget{coal-plants-or-overall-pollution}{%
-\subsection{Coal plants or overall pollution?}\label{coal-plants-or-overall-pollution}}
-
-Finally, while I identify high-pollution areas using the location of coal plants, an important question is how much of the overall effect is driven by coal plants and how much is driven by other polluting industries located in areas near coal plants.
-I try to answer this question in Table \ref{tab:yield7ivdiffindiff}, where I implement an instrumental variables differences-in-differences, essentially instrumenting for pollution prior to the opening of a coal plant with wind and instrumenting for the interaction term (pollution times plant opening) with wind times plant opening.
-
-\begin{table}
-
-\caption{\label{tab:yield7ivdiffindiff}Agricultural productivity and pollution source}
-\centering
-\begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3.5cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
-\toprule
-  & (1) & (2) & (3)\\
-\midrule
-Particulate matter & -0.299*** & -0.107* & -0.147**\\
-(log PM 2.5) & (0.066) & (0.063) & (0.063)\\
-PM 2.5 times Coal & -0.328*** & -0.321*** & -0.304***\\
- & (0.026) & (0.024) & (0.023)\\
-Weather & No & Yes & Yes\\
-Weather (expanded) & No & No & Yes\\
-\textbf{Fixed effects:} & \textbf{} & \textbf{} & \textbf{}\\
-Year & Yes & Yes & Yes\\
-Village & Yes & Yes & Yes\\
-\midrule
-F (PM) & 766 & 776 & 788\\
-F (PM times Coal) & 199 & 222 & 232\\
-Observations & 1,266,922 & 1,266,922 & 1,266,922\\
-\bottomrule
-\end{tabular}
-\begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the village level.
-The outcome in all regressions is estimated agricultural yield, defined as tons per hectare.
-* p<0.10 ** p<0.05 *** p<0.01
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
 Columns one through three of Table \ref{tab:yield7ivdiffindiff} present specifications with progressively increasing weather controls.
 Column three presents the most extensive set of controls and I focus on that column here.
-First, there are now two first stages but the F-tests indicate that the instruments are strong for both, including the interaction term (F = 232).
+First, there are now two first stages but the F-tests indicate that the instruments are strong for both, including the interaction term.
 Second, the first row shows the effect of pollution on agricultural productivity prior to the opening of a coal plant; the coefficient is negative, indicating negative effects of non-coal pollution.
@@
 In other words, this is not identifying differences in overall pollution, per se, but rather differences in the effects of an increase of pollution by the same percentage.
-These different effects could be due to, for example, differences in emission patterns for different kinds of plants and factories; coal plants, for example, are responsible for much more mercury than other kinds of polluting industries.\footnote{\url{https://www.epa.gov/sites/default/files/2016-12/documents/nei2014v1_tsd.pdf}}
+These different effects could be due to, among other things, differences in emission patterns for different kinds of plants and factories; coal plants, for example, are responsible for much more mercury than other kinds of polluting industries.\footnote{\url{https://www.epa.gov/sites/default/files/2016-12/documents/nei2014v1_tsd.pdf}}
 
@@
 While we often frame the debate around greenhouse gasses in terms of climate change, there are important local reasons to reduce pollution emissions, as well.
-The technology for greener energy exists; at this point, it is simply a question of policies {[}Markard (2018); blazquez2020some{]}.
+The technology for greener energy exists; at this point, it is simply a question of policies (Markard 2018; Blazquez, Fuentes, and Manzano 2020).
 
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-abadie2023should}{}}%
-Abadie, Alberto, Susan Athey, Guido W Imbens, and Jeffrey M Wooldridge. 2023. {``When Should You Adjust Standard Errors for Clustering?''} \emph{{The Quarterly Journal of Economics}} 138 (1): 1--35.
-
 \leavevmode\vadjust pre{\hypertarget{ref-abay2022nonclassical}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-arceo2016does}{}}%
-Arceo, Eva, Rema Hanna, and Paulina Oliva.
-2016.
-{``{Does the effect of pollution on infant mortality differ between developing and developed countries?
-Evidence from Mexico City}.''} \emph{{The Economic Journal}} 126 (591): 257--80.
-
 \leavevmode\vadjust pre{\hypertarget{ref-almn2021}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-barrios2012clustering}{}}%
-Barrios, Thomas, Rebecca Diamond, Guido W Imbens, and Michal Kolesár.
-2012.
-{``Clustering, Spatial Correlations, and Randomization Inference.''} \emph{{Journal of the American Statistical Association}} 107 (498): 578--91.
-
 \leavevmode\vadjust pre{\hypertarget{ref-bencko1977health}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-bertrand2004much}{}}%
-Bertrand, Marianne, Esther Duflo, and Sendhil Mullainathan. 2004. {``How Much Should We Trust Differences-in-Differences Estimates?''} \emph{{The Quarterly Journal of Economics}} 119 (1): 249--75.
-
-\leavevmode\vadjust pre{\hypertarget{ref-brunekreef2002air}{}}%
-Brunekreef, Bert, and Stephen T Holgate. 2002. {``{Air pollution and health}.''} \emph{{The Lancet}} 360 (9341): 1233--42.
+\leavevmode\vadjust pre{\hypertarget{ref-blazquez2020some}{}}%
+Blazquez, Jorge, Rolando Fuentes, and Baltasar Manzano. 2020. {``On Some Economic Principles of the Energy Transition.''} \emph{{Energy Policy}} 147: 111807.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-currie2014we}{}}%
-Currie, Janet, Joshua Graff Zivin, Jamie Mullins, and Matthew Neidell.
-2014.
-{``{What do we know about short-and long-term effects of early-life exposure to pollution?}''} \emph{{Annual Review Resource Economics}} 6 (1): 217--47.
-
 \leavevmode\vadjust pre{\hypertarget{ref-dercon2009rural}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-heft2018robust}{}}%
-Heft-Neal, Sam, Jennifer Burney, Eran Bendavid, and Marshall Burke. 2018. {``{Robust relationship between air quality and infant mortality in Africa}.''} \emph{Nature} 559 (7713): 254--58.
-
-\leavevmode\vadjust pre{\hypertarget{ref-heft2019air}{}}%
-Heft-Neal, Sam, Jennifer Burney, Eran Bendavid, Kara Voss, and Marshall Burke. 2019. {``{Air pollution and infant mortality: Evidence from Saharan dust}.''} {National Bureau of Economic Research}.
-
 \leavevmode\vadjust pre{\hypertarget{ref-herrendorf2014growth}{}}%
@@
 IEA. 2022. {``{Coal-fired electricity}.''} {International Energy Association}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-kampa2008human}{}}%
-Kampa, Marilena, and Elias Castanas. 2008. {``{Human health effects of air pollution}.''} \emph{{Environmental Pollution}} 151 (2): 362--67.
 
@@
 {International Institute for Environment and Development.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-pope2006health}{}}%
-Pope III, C Arden, and Douglas W Dockery.
-2006.
-{``{Health effects of fine particulate air pollution: lines that connect}.''} \emph{{Journal of the Air \& Waste Management Association}} 56 (6): 709--42.
 
@@
 
+\begin{table}
+
+\caption{\label{tab:plantresultstable}Local characteristics and the construction of coal plants}
+\centering
+\begin{threeparttable}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\toprule
+\multicolumn{1}{c}{ } & \multicolumn{2}{c}{1991 census} & \multicolumn{2}{c}{2001 census} \\
+\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
+  & 1991 & 2001 & 2001 & 2011\\
+\midrule
+Population (log) & 0.009 & 0.007* & 0.013* & 0.005\\
+ & (0.006) & (0.004) & (0.006) & (0.004)\\
+Area (log) & -0.007* & -0.004 & -0.007* & -0.004\\
+ & (0.004) & (0.005) & (0.004) & (0.003)\\
+Literacy (prop) & 0.080** & 0.037 & 0.073* & 0.036\\
+ & (0.029) & (0.028) & (0.039) & (0.026)\\
+Ag. productivity &  &  & 0.014** & 0.028***\\
+ &  &  & (0.005) & (0.009)\\
+Pollution &  &  & 0.047 & 0.014\\
+(log PM 2.5) &  &  & (0.045) & (0.018)\\
+Sub-sample & all & no plant & all & no plant\\
+\midrule
+Observations & 283,021 & 264,265 & 508,784 & 469,004\\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}[para]
+\item Note: State fixed effects are included in regressions.
+Standard errors are in parentheses and clustered at the state level.
+The outcome in the first column is whether the village is within 30km of a coal plant in 1991.
+The second column is whether a village in 1991 will have a coal plant in 2001, conditional on not having one in 1991.
+The last two columns are similarly defined, except using 2001 and 2011 as the years.
+* p<0.10 ** p<0.05 *** p<0.01
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+
+\FloatBarrier
+\newpage
+
 \begin{figure}
@@
 \begin{figure}
-\includegraphics{draft_files/figure-latex/windexample-1} \caption[Wind direction and aggregation examples (2010-01-01)]{Wind direction and aggregation examples (2010-01-01)}\label{fig:windexample}\floatfoot*{Note: The top figure shows the average wind direction on January 1st, 2010, with zero degrees indicating north.
-The points are the location of coal plants on that date.
-The bottom figure shows the distribution of pollution exposure in a specific district -- Guna district in Madhya Pradesh -- on the same date.}
+\includegraphics{draft_files/figure-latex/windexample-1} \caption[Wind direction example (2010-01-01)]{Wind direction example (2010-01-01)}\label{fig:windexample}\floatfoot*{Note: The figure shows the average wind direction on January 1st, 2010, with zero degrees indicating north.
+The points are the location of coal plants on that date.}
 \end{figure}
@@
 
-\begin{table}
-
-\caption{\label{tab:pollutiontable}Wind direction and particulate matter}
-\centering
-\begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{1998-2015} & \multicolumn{2}{c}{2002-2013} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
-  & (1) & (2) & (3) & (4)\\
-\midrule
-wind (10s) & 0.0023*** & 0.0016*** & 0.0057*** & 0.0015***\\
- & (0.0005) & (0.0002) & (0.0005) & (0.0002)\\
-\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-village & Yes & Yes & Yes & Yes\\
-month & Yes & No & Yes & No\\
-district-month & No & Yes & No & Yes\\
-\midrule
-observations & 24,376,458 & 24,376,458 & 16,250,973 & 16,250,973\\
-\bottomrule
-\end{tabular}
-\begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
-\FloatBarrier
-\newpage
-
-\begin{figure}
-\includegraphics{draft_files/figure-latex/randomplotleads-1} \caption[Randomization inference for leads]{Randomization inference for leads}\label{fig:randomplotleads}\floatfoot*{Note: The figure presents the distribution of coefficients by randomly assigning a single wind value to each village, sampled with replacement across all villages in a given district in the same year.
-"Current" refers to the coefficient on current pollution, while "lead" refers to the coefficient on next year's pollution in the same table.}
-\end{figure}
-
-\FloatBarrier
-\newpage
-
-\begin{figure}
-\includegraphics{draft_files/figure-latex/laborplot-1} \caption[Timing of household surveys for the National Sample Survey]{Timing of household surveys for the National Sample Survey}\label{fig:laborplot}\floatfoot*{Note: The histogram shows the distribution of survey dates for the four waves of the National Sample Survey (NSS) for the five waves used in this paper.}
-\end{figure}
-
-\FloatBarrier
-\newpage
-
-\begin{figure}
-\includegraphics{draft_files/figure-latex/monthplot-1} \caption[Average farm days by season, rural only]{Average farm days by season, rural only}\label{fig:monthplot}\floatfoot*{Note: The line graph shows the mean number of farm days worked per week based on the month of the relevant agricultural season.
-The monsoon season is defined as June (month 1) to October (month 5) while the rainy season is defined as November (month 1) to March (month 5) of the following year.
-The data is from the NSS.}
-\end{figure}
-
-\FloatBarrier
-\newpage
-
-\begin{table}
-
-\caption{\label{tab:yield8cumulative}Cumulative exposure to pollution during the season}
-\centering
-\begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2.25cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
-\toprule
-  & (1) & (2) & (3) & (4) & (5)\\
-\midrule
-June & -0.643*** &  &  &  & \\
- & (0.064) &  &  &  & \\
-July &  & -0.658*** &  &  & \\
- &  & (0.045) &  &  & \\
-Aug. &  &  & -0.531*** &  & \\
- &  &  & (0.049) &  & \\
-Sept. &  &  &  & -0.672*** & \\
- &  &  &  & (0.047) & \\
-Oct. &  &  &  &  & -0.943***\\
- &  &  &  &  & (0.069)\\
-weather & Yes & Yes & Yes & Yes & Yes\\
-\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-village & Yes & Yes & Yes & Yes & Yes\\
-year & Yes & Yes & Yes & Yes & Yes\\
-\midrule
-F & 1,327 & 2,830 & 2,933 & 2,413 & 1,515\\
-observations & 1,266,922 & 1,266,922 & 1,266,922 & 1,266,922 & 1,266,922\\
-\bottomrule
-\end{tabular}
-\begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
 \end{document}

### 2023-07-10 draft/draft.Rmd
updated draft; labor intensity

@@
 abstract: |
-  | \noindent I document externalities in the agricultural sector of India due to air pollution.
-Using variation in pollution induced by changes in wind, I show that higher levels of particulate matter lead to decreased agricultural productivity, with changes in productivity of 10 percent being common.
-Combining wind direction with the rollout of coal plants, results indicate pollution from coal plants has a larger effect on agricultural productivity than other types of pollution.
-Given that the agricultural sector is a refuge for the poor in many developing countries, these results suggest that the effects of pollution may hit the poorest particularly hard.
+  | \noindent I document negative externalities of air pollution in the Indian agricultural sector.
+Using variation in pollution induced by changes in wind across years, I show that higher levels of pollution lead to decreased agricultural productivity, with large changes in productivity being common.
+The negative effects of pollution are larger in areas growing more labor-intensive crops, indicating that the pollution works at least partly through direct effects on labor productivity.
+Finally, combining wind direction with the rollout of coal plants, results indicate that pollution from coal plants has a larger effect on agricultural productivity than other types of pollution.
+Given that the agricultural sector is a refuge for the poor in many developing countries, these results suggest that the negative externalities of pollution may hit the poorest particularly hard.
   |
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, dev = "cairo_pdf")
+knitr::opts_chunk$set(echo = FALSE, dev = "png") # NOTE: switched to png instead of pdf to decrease size of the resulting pdf
 
@@
 
-# Sets wd to folder this script is in so we can create relative paths instead of absolute paths
-setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
 # We want to get the wd up to the main folder
@@
                           dplyr::select(shrid, state = ST_CEN_CD, district = DT_CEN_CD, state_name = ST_NM, district_name = DISTRICT)
+village_shp <- villages
 villages <- villages %>% filter(state=="23" & district=="07")
@@
 villages <- villages %>% st_set_crs(st_crs(districts))
+village_shp <- village_shp %>% st_set_crs(st_crs(districts))
 # and crs for plants
@@
 villages <- st_snap(villages, districts, tolerance = 1250)
+crops_all <- read_csv(paste0("data/clean/ag_productivity/crop_area.csv"))
 
@@
 yield2naive <- readRDS("pollution_development/draft/tables/yield2naive.rds")
+yield3ivmainfirststage <- readRDS("pollution_development/draft/tables/yield3ivmain_firststage.rds")
 yield3ivmain <- readRDS("pollution_development/draft/tables/yield3ivmain.rds")
@@
 coef_lead <- readRDS("pollution_development/draft/tables/coef_lead.rds")
+yieldrice <- readRDS("pollution_development/draft/tables/yieldrice.rds")
+yieldricestatefe <- readRDS("pollution_development/draft/tables/yieldricestatefe.rds")
 
@@
 
-There are no doubt positive effects from this increase in power generation; for example, electricity may lead to economic and productivity growth [@dinkelman2011effects; @kline2014local; @rud2012electricity; @van2017long], though electrification alone may not be a sufficient condition [@lee2020does].
-These improvements do not come without downsides, however, especially when it comes to power derived from coal.
+While there are likely positive effects from this increase in power generation -- for example, electricity may lead to economic and productivity growth [@dinkelman2011effects; @kline2014local; @rud2012electricity; @van2017long], though electrification alone may not be a sufficient condition [@lee2020does] -- these improvements do not come without downsides, especially when it comes to power derived from coal.
 Pollutants from coal contribute to climate change and the phase out of coal energy is therefore a major climate goal [@iea2022].
-Emissions from the production and burning of coal are also directly harmful to health and the environment, including pollutants like sulfur dioxide, nitrogen oxides, and mercury, to name but a few.[^4] While the use of coal had been declining, recent events -- including Russia's invasion of Ukraine -- have led to an increase in the demand for coal.[^5]
+Emissions from the production and burning of coal are also directly harmful to human health and the environment; pollutants like sulfur dioxide, nitrogen oxides, methane, and mercury are particularly harmful.[^4] While the use of coal had been declining, recent events -- including Russia's invasion of Ukraine -- have led to an increase in the demand for coal.[^5]
 
@@
 
-In addition to coal, pollution more generally is increasing in much of the world.
-Carbon dioxide equivalent per capita has almost double over the last 50 years in India, more than doubled in Indonesia, and almost tripled in China.^[https://edgar.jrc.ec.europa.eu] Of course, much of this increase is due to the development process itself; as countries develop, their economies transform, with more people moving out of agriculture and into manufacturing and services [@herrendorf2014growth].
+In addition to pollution from coal, pollution more generally is also increasing in much of the world.
+Carbon dioxide equivalent per capita has almost double over the last 50 years in India, more than doubled in Indonesia, and almost tripled in China.^[https://edgar.jrc.ec.europa.eu] Much of this increase is due to the development process itself -- as countries develop, their economies transform, with more people moving out of agriculture and into manufacturing and services [@herrendorf2014growth].
 Despite this development, agriculture remains a key sector in many developing countries, employing a large proportion of the poor [@dercon2009rural].
@@
 On the one hand, governments may locate new plants in fast-growing areas, while on the other hand, they may decide the exact location of plants based on the political power of local citizens [@kopas2020environmental].
-In the case of the data I use here, coal plants are constructed in areas with higher population and higher literacy.
-In terms of agriculture, higher productivity predicts more coal plants in a given year as well as the construction of a coal plant over the next decade, indicating possible bias when estimating OLS regressions of agricultural productivity on pollution levels.
+In the case of the data I use here, coal plants are constructed in areas with higher population and higher levels of pollution.
+In terms of agriculture, higher productivity predicts more coal plants in a given year as well as the construction of a coal plant over the next decade, indicating possible bias when estimating effects of pollution on agricultural productivity.
 
@@
 I include all areas, regardless of whether the coal plant has opened yet, to allow me to disaggregate the effects of pollution from coal plants with other kinds of pollution later in the paper.
+
 Using daily wind data and these spatial variables, I create an identifier separately for every village for whether a given village is located downwind from a coal plant on any given day.
@@
 
-Using wind direction as an instrument for particulate matter, an increase of PM 2.5 by one percent leads to a large decrease in agricultural productivity, at around 0.73 percent.
-Given within-village changes in exposure, pollution could be responsible for swings of 7-to-14 percent in agricultural productivity across years.
+Using wind direction as an instrument for particulate matter, an increase of PM 2.5 by one percent leads to a large decrease in agricultural productivity, with an elasticity of approximately -0.63 in the most conservative specification.
 The IV estimates indicate that the effect of pollution is more negative than naïve OLS estimates, which is consistent with the evidence on the construction of coal plants; higher levels of pollution are correlated with higher levels of agricultural productivity, a relationship that the IV strategy adjusts downwards.
@@
 Heterogeneity analysis shows that the negative effects of pollution are larger in villages that experience higher median exposure to pollution and in villages with higher initial yields.
-Finally, given how many poor farmers rely on rainfall and how weather shocks increase across the globe, I also look at possible multiplicative effects of shocks by interacting the pollution variables with a rainfall variable.
+Finally, given how many poor farmers rely on rainfall and how weather shocks are on the rise across the globe, I also look at possible multiplicative effects of shocks by interacting the pollution variables with a rainfall variable.
 Higher exposure along with less rainfall increases the negative effect of either shock individually, meaning these shocks are compounding.
@@
 
-I find that these effects are mostly transmitted by reductions in farm labor during narrow windows of the growing season, using labor allocation data from the National Sample Survey (NSS).
-While there do not appear to be large changes in labor allocation patterns when looking at the season as a whole, there is a clear pattern when we break down exposure to pollution by month: higher levels of pollution at the beginning of the monsoon season (June and July) lead to lower levels of farm labor allocation, but no changes in non-farm labor allocation.
-This matches up well with a similar exercise with agricultural yields: pollution in June is the only month in which there is a negative effect of pollution on agricultural yield.
-In tandem, these two results point to effects on labor as one mechanism driving the total effects, though I cannot rule out direct effects of pollution on land productivity with my data.
+A key question is the mechanism: does pollution reduce labor productivity, land productivity, or both?
+While previous work from the United States indicates that pollution from coal plants, specifically, can lead to higher _land_ productivity [@sanders2022adaptation], agriculture in India differs greatly from agriculture in the United States.
+To test for possible effects on labor productivity, I look at differential effects based on labor intensity of the predominant crops in different villages.
+Specifically, I separate effects in predominantly rice-growing areas from other areas; since rice is very labor intensive, we would expect to see larger negative effects of pollution in rice-growing regions relative to other regions if one mechanism through which pollution affects productivity is direct effects on labor.
+I show that pollution negatively effects agricultural productivity in non-rice areas but that the effects in rice-growing areas are around 50 percent larger in magnitude.
+These results indicate that pollution likely directly affects labor productivity, though I am not able to rule out direct effects on land, either, due to a lack of data.
 
 The final set of analyses tries to disentangle the effects of pollution from the effects of coal pollution, specifically, by using wind direction and the rollout of coal plants across the country.
-Combining instrumental variables with a differences-in-differences strategy, I show that agricultural productivity has an elasticity of around -0.15 with respect to non-coal pollution.
-The opening of a coal plant, however, triples the elasticity, to more than -0.45, indicating that coal plant emissions are particularly harmful to agricultural productivity, at least in India.
-This is consistent with coal emissions being quite different from other forms of pollution; for example, coal emissions contain much more mercury than other forms of pollution.
+Combining instrumental variables with a differences-in-differences strategy, I show that agricultural productivity has an elasticity of around -0.22 with respect to non-coal pollution.
+The opening of a coal plant, however, doubles the elasticity, to more than -0.44, indicating that coal plant emissions are particularly harmful to agricultural productivity, at least in India.
+This is consistent with coal emissions being quite different from other forms of pollution; coal emissions contain different types and quantities of harmful substances, meaning effects may differ greatly.
+
+An important caveat to these results is that I cannot disentangle exactly which pollutants drive results.
+While I use PM 2.5 as a measure of overall exposure to pollution, effects could be driven by any pollutants carried by the wind.
+This leaves open an important avenue of future research regarding the effects of different types of pollutants on agricultural productivity.
 
 This paper contributes to several strands of literature.
-First, we already have evidence of effects of pollution on different forms of labor activity.
+We already have evidence of effects of pollution on different forms of labor activity.
 For example, general levels of pollution decrease productivity of call center works in China [@chang2019effect], lead people to perform worse on cognitive functions [@ebenstein2016long; @wen2022lower], and also lead to lower levels of farm labor productivity in California, driven by changes in ozone concentration [@graff2012impact].
 Also in China, @he2019severe find small but significant negative effects of pollution on manufacturing productivity, while @chen2022effect show pollution leads to labor out migration.
-Perhaps relevant to my labor allocation results, @hanna2015effect show that decreases in pollution lead to increases in labor allocation, while I show the same relationship in at least a subset of months of the agricultural season.
+Relevant to my results, @hanna2015effect show that decreases in pollution lead to increases in labor allocation in another developing country.
+
 Several papers are closely related to this one.
@@
 Finally, @sanders2022adaptation actually find that coal emissions can be good for agriculture in the United States, due to the presence of sulfur, which can increase agricultural productivity.
-This paper differs in three important ways.
+This paper differs from the previous literature in three important ways.
 First, I explore effects of pollution more generally as well as from coal plants, which are much more ubiquitous than gold mines across the world.
-Second, I use a different identification strategy from either paper, relying on wind blowing pollution from coal plants.
+Second, I use a different identification strategy from most other papers, relying on wind blowing pollution from coal plants.
 Finally, agriculture in India and the United States can be quite different, with the latter being much more mechanized.
-This means effects of pollution could differ substantially.
-
-I also contribute to the more general literature on the effects of pollution on health.
-Pollution has substantial negative effects on health in both developing and developed countries [@arceo2016does] and we have known of these negative effects for many years [@brunekreef2002air; @kampa2008human; @pope2006health].
-@currie2014we review the literature on the health effects of pollution during adolescence; we have robust evidence of pollution exposure leading to increases in infant mortality, including in developing countries [@heft2018robust; @heft2019air].
-Since these effects are often driven by heart and lung disease and asthma [@kampa2008human], this my explain the labor allocation effects I identify in this paper.
-
-The rest of the paper is organized as follows.
-I first discuss data and then methodology, in two separate sections.
-The methodology section discusses the empirical strategy and validates the wind direction variable using data on particulate matter.
-I present the main results in \autoref{results} before concluding in \autoref{conclusion}.
+This means effects of pollution could differ substantially depending on the context, a conclusion that the current results support.
 
@@
 
-To help understand some of the mechanisms driving the relationship between exposure to emissions and agricultural productivity, I also include household survey data.
-Unfortunately, I am not aware of any publicly available survey data that covers multiple years and has identifiers below the district level.
-For example, the ARIS-REDS data[^11] includes GPS identifiers for one year, but that only allows GPS matching for a total of two years: 1999 and 2007/2008.
-The DHS, on the other hand, only has state-level identifiers (except for the newest two waves) and is also not a labor survey.
-Instead, I use the National Sample Survey (NSS) from the Ministry of Statistics and Programme Implementation.
-
-[^11]: www.ncaer.org/project/additional-rural-incomes-survey-aris-rural-economic-demographic-survey-reds
-
-The NSS is a nationally representative survey that was conducted every couple of years until relatively recently.
-Many rounds of the NSS also included a module on employment that asked about labor allocation over the previous seven days.
-I use the 61st, 62nd, 64th, 66th, and 68th rounds of the NSS, all of which include the employment module.
-The publicly available data also includes the date of survey, which allows me to match pollution exposure to the exact day of the survey.
-
-The biggest issue with using the NSS is that it includes only district-level identifiers.
-The main exposure variables, on the other hand, are defined at a lower level of aggregation, at the village.
-I choose to aggregate this village-level exposure to the district level as shown in the second panel of Figure \@ref(fig:windexample) -- the district is a randomly selected district in the data, Guna district in Madhya Pradesh.
-I use the area overlap of the villages with the district to define a weighted average of exposure for the district at a given time.
-I calculate the area of overlap for each village and the district as well as the overall area of the district and use these variables to define weights for individual villages.
-Any part of the district that does not overlap with one of the villages -- the white area in Figure \@ref(fig:windexample) -- does not have any exposure and, as such, receives a value of zero for exposure.
+To help understand some of the mechanisms driving the relationship between exposure to emissions and agricultural productivity, I also look at the differential effects by the labor intensity of the predominant crop in each area.
+For this, I use data from [@monfreda2008farming], who estimate area harvested for different crops across the global around the year 2000.
+I identify areas where rice (paddy) is the most commonly grown crop, under the assumption that rice is more labor intensive than other crops [@vollrath2011agricultural; @michler2020agriculture].
+@michler2020agriculture, for example, shows that rice, on avearge, uses almost 1,800 hours per hectare, while the second-highest crop in his data, cotton, juses just 850.^[He estimates that wheat, another common crop in India, uses just 600 hours per hectare.]
 
@@
 However, even with the fixed effects, this regression likely suffers from endogeneity; places with more pollution may be very different from places with less pollution.
-It is difficult to sign the direction of the bias a priori, because it depends on how the coal plants are placed.
+It is difficult to sign the direction of the bias a priori, because it depends on the local context and how the coal plants are placed.
 For example, if coal plants are placed in places in need of power -- for example, due to rapid growth -- then the estimates might be upwards biased.
@@
 
-So what are the characteristics of areas near the construction of new coal plants?
+What are the characteristics of areas near the construction of new coal plants?
 Table \@ref(tab:plantresultstable) presents four separate regressions, using state-level fixed effects due to the spatial variation in coal plants.
@@
 The first column presents results from regressing a dummy for whether there is a coal plant within 30km in 1991 on several census variables.
-Villages with higher populations and higher literacy in 1991 are more likely to be near a coal plant at the time.
+Villages with higher population densities (higher log population and lower log area) are more likely to be near a coal plant at the time.
+Urban population proportion is not a significant predictor after conditioning on population and area.
 
@@
 colnames(plantresultstable) <- c("1991", "2001", "2001", "2011")
-rownames(plantresultstable) <- c("Population (log)", "", "Area (log)", "", "Literacy (prop)", "", "Ag. productivity", "", "Pollution", "(log PM 2.5)",
+rownames(plantresultstable) <- c("Population (log)", "", "Area (log)", "", "Urban pop (prop)", "", "Ag. productivity", "", "Pollution", "(log PM 2.5)",
                                 "Sub-sample", "Observations")
@@
            "Note: State fixed effects are included in regressions.
-Standard errors are in parentheses and clustered at the state level.
+Robust standard errors are in parentheses.
 The outcome in the first column is whether the village is within 30km of a coal plant in 1991.
@@
 This column restricts the sample to only villages that do not already have a coal plant within 30km in 1991.
-Population and literacy both remain strong predictors of the building of a new coal plant.
+The pattern of significance remains the same in column two.
 
@@
 However, note that this is not the same as looking at wind direction, though comparing the OLS and IV results below leads to the same conclusion regarding selection.
-Pollution, on the other hand, is not significant but is positive in both columns (and is also logged, just like productivity).
-
-Given the likely endogeneity, I instead turn to a plausibly exogenous source of variation in pollution: wind direction.
+
+Pollution is also a significant predictor for the presence of a coal plant in 2001 as well as for whether a coal plant will be constructed before 2011.
+While the coefficient is relatively small, it is actually quite large given that only three percent of villages in column four will get a new plant in the next decade.
+
+Given the likely endogeneity previewed by these summary statistics, I now turn to a plausibly exogenous source of variation in pollution: wind direction.
 While the location of a coal plant might be endogenous, the amount of pollution received by a specific village in a given year can vary randomly based on wind direction.
@@
 It is worth taking the time to validate the use of wind direction as a measure of exposure to pollution.
-Consider the data used for the regressions presented in Table \@ref(tab:pollutiontable) of the appendix.
+Consider the data used for the regressions presented in Table \@ref(tab:yield3ivmainfirststage) of the appendix.
 The outcome variable is particulate matter -- specifically, $\mathrm{PM 2.5}$, which is particulate matter no larger than 2.5 micrometers in diameter -- which comes from @hammer2020global.
-$\mathrm{PM 2.5}$ is one of the harmful byproducts from coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.[^12] This data is available at the monthly level, so in Table \@ref(tab:pollutiontable), I aggregate wind exposure to the monthly level, as well.
+$\mathrm{PM 2.5}$ is one of the harmful byproducts from coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.[^12] This data is available at the monthly level, but I aggregate variables to the season variable since that is the level of analysis in the main results.
 
@@
 
-Both the particulate matter data and the wind data is at the village level, which means the regressions in Table \@ref(tab:pollutiontable) are at the village level.
+Both the particulate matter data and the wind data is at the village level, which means the regressions in Table \@ref(tab:yield3ivmainfirststage) are at the village level.
 As such, the geographic fixed effects are village fixed effects and the temporal fixed effects are month fixed effects.
@@
 
-In addition to serving as an example, Table \@ref(tab:pollutiontable) also helps validate the use of wind direction as a proxy for exposure to pollution.
+In addition to serving as an example, Table \@ref(tab:yield3ivmainfirststage) also helps validate the use of wind direction as a proxy for exposure to pollution.
 The first two columns include all available years of data while the last two columns restrict estimation only to the years for which the agricultural productivity regressions are estimated.
@@
 
-To put context behind pollution numbers, the World Health Organization's updated guidelines are that average annual exposure should not exceed 15 $\mathrm{\mu g/m^3}$.
-For the month, an increase at the midpoint of the estimated range is equal to an increase of 0.2 percent of the maximum recommended mean concentration from the WHO.
-This is just a single day of wind in the direction of a village.
-Since this includes all villages located within 30km of a coal plant, this is evidence that wind direction can have serious repercussions on the health of hundreds of millions of people in India.
+## Identifying mechanisms
+
+To identify possible mechanisms, I identify areas that grow predominantly rice, which is a labor-intensive crop. Specifically, we are interested in the following regression:
+\begin{gather} 
+y_{it} = \alpha_{i} + \gamma_{t} + \omega PM_{it} + \phi rice_{it} + \beta PM_{it}\times rice_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it},
+\end{gather}
+where $PM_{it}$ denotes the effect of pollution in non-rice areas and $PM_{it}\times rice_{it}$ is the differential effect in rice-growing areas.
+Here, I create an indicator variable for rice areas that is time invariant, so $rice_{it}$ drops out of the estimation.
+I instrument for $PM_{it}$ and for $PM_{it}\times rice_{it}$ using wind and wind interacted with the indicator for rice.
 
@@
 I first present the reduced form results for agricultural productivity.
-The outcome of all the regressions in this section is the log of agricultural land productivity, which is defined as tons per hectare.
+The outcome of all regressions in this section is the log of agricultural land productivity, which is defined as tons per hectare.
 The first set of results are in Table \@ref(tab:yieldtable).
-The unit of analysis is the village-year -- from 2002 to 2013 -- focusing only on the monsoon (kharif) season.
+The unit of analysis is the village-season -- from 2002 to 2013 -- focusing only on the monsoon (kharif) season.
 The data used to construct the yield measure is calculated using five months of data, so the wind variable is defined for the same five months and is the sum of days in which the wind is pointed towards the village centroid from a coal plant.
@@
 # Table
+yield1reducedform <- yield1reducedform[,-c(1)]
 colnames(yield1reducedform) <- c("(1)", "(2)", "(3)")
@@
 # Table
+yield2naive <- yield2naive[,-1]
 colnames(yield2naive) <- c("(1)", "(2)", "(3)")
@@
 There is again a consistent, negative relationship between particulate matter and agricultural productivity.
-Since both variables are logged, the coefficients are similar to elasticities, indicating that a one-percent increase in pollution leads to a 0.03-to-0.04 percent decrease in agricultural productivity.
+Since both variables are logged, the coefficients are similar to elasticities, indicating that a one-percent increase in pollution leads to an approximately 0.04 percent decrease in agricultural productivity.
 
@@
 This means that the simple regressions in Table \@ref(tab:yieldtablepm) might actually be biased upwards: although pollution might negatively affect productivity, pollution is higher in more productive areas due to the endogenous placement of coal plants in the study areas.
-To recover the actual effect of pollution, we need an instrument.
+One possible way to recover the actual effect of pollution on agricultural productivity is with an instrument.
 If, conditional on the fixed effects and weather variables, wind direction is exogenous and only affects agricultural productivity through pollution, then it is a valid instrument for pollution.
 
-I present the IV results in Table \@ref(tab:yieldtabletwo), with the first-stage F-statistic at the bottom.
-The first-stage results make clear how strong of an instrument wind is; all three F-statistics are above 1,000, meeting even the higher bar for the first-stage F-test argued for by @lee2022valid.
-Also note that the coefficients are quite different from those in Table \@ref(tab:pollutiontable) because the unit of analysis in Table \@ref(tab:yieldtabletwo) is the season, while in Table \@ref(tab:pollutiontable) the unit of analysis is the month.
+I present these IV results in Table \@ref(tab:yieldtabletwo), with the first-stage F-statistic at the bottom.
+The first-stage results -- the full results are in Table \@ref(tab:yield3ivmainfirststage) of the appendix -- make clear how strong of an instrument wind is; all three F-statistics are above 1,000, meeting even the higher bar for the first-stage F-test argued for by @lee2022valid.
 
 The second-stage results first indicate that $\mathrm{PM 2.5}$ is a very strong predictor of agricultural productivity.
-Using the results in column three, a one percent increase in particulate matter leads to a 0.73 percent decrease in agricultural productivity.
-To put this in further perspective, the mean particulate matter concentration is just short of 40 (39.68) in the monsoon sample, while the within-village absolute deviation in particulate matter is 3.98, meaning year-to-year swings of 7-to-14 percent in agricultural productivity due to pollution exposure are common.
-
-The comparison of the results to those in the previous table also confirm the concerns around endogenous placement of coal plants.
-The coefficients in Table \@ref(tab:yieldtabletwo) are always more negative than those in Table \@ref(tab:yieldtablepm), and sometimes substantially so.
-In other words, the simple regression of agricultural productivity on pollution results in coefficients on pollution that are upwards biased, since higher levels of pollution occur in more productive areas, on average.
-The IV strategy corrects this bias and leads to a robust, negative relationship between pollution and agricultural productivity.
+Using the results in column three, a one percent increase in particulate matter leads to a 0.62 percent decrease in agricultural productivity.
+To put this in further perspective, the mean particulate matter concentration is just short of 40 (39.68) in the monsoon sample, while the within-village absolute deviation in particulate matter is 3.98, meaning year-to-year swings of several percent in agricultural productivity due to pollution exposure are common.
+
+Column four presents one last specification; it includes a much more flexible specification for weather.
+Specifically, I turn the linear z-score weather variable into bins for each month of the season.
+I then interact this rainfall variable with the temperature variable.
+The coefficient in column four shows almost not change from those in column three, relieving some concerns that weather may not be correctly specified, leading to a spurious correlation between particulate matter and agricultural productivity.
 
@@
 yield3ivmain <- yield3ivmain[,-1]
-colnames(yield3ivmain) <- c("(1)", "(2)", "(3)")
+colnames(yield3ivmain) <- c("(1)", "(2)", "(3)", "(4)")
 rownames(yield3ivmain) <- c("Particulate matter", "(log PM 2.5)",
-                            "Weather", "Weather (expanded)",
+                            "Weather", "Weather (expanded)", "Weather (expanded, bins)",
                             "Fixed effects:", "Village", "Year",
@@
       yield3ivmain, caption = "Pollution and agricultural productivity, IV estimates",
-      align = "ccc", booktabs = TRUE, linesep = ""
+      align = "cccc", booktabs = TRUE, linesep = ""
       ) %>%
@@
   column_spec(1, width = "4cm") %>%
-  column_spec(c(2:4),width = "2cm") %>%
-  row_spec(5, bold = TRUE) %>%
-  row_spec(7, hline_after = TRUE) %>%
+  column_spec(c(2:5),width = "2cm") %>%
+  row_spec(6, bold = TRUE) %>%
+  row_spec(8, hline_after = TRUE) %>%
   #add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
@@
 
-```{r randomplot, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 8, fig.cap = "Randomization inference", fig.footer = list("Note: The figure presents the distribution of coefficients by randomly assigning wind direction across villages within the same year.
-The \"all\" distribution is created by randomly assigning a single wind value to each village, sampled with replacement across all villages in a given year.
-The \"within district\" randomly samples with replacement only from within the same district in the same year.")}
-
-pal <- viridis(4)
-colors <- c(paste0(pal[1]), paste0(pal[2]))
-labels <- c("All", "Within district")
-
-# plot
-gg1 <- ggplot() + 
-        geom_density(aes(x = coef_vec, fill = "All"), alpha = 0.5) +
-        geom_density(aes(x = coefs_distfe, fill = "Within district"), alpha = 0.5) +
-        scale_fill_manual(
-          name = "Randomization:",
-          values = colors,
-          labels = labels
-        ) +
-        theme_minimal() +
-        xlab("Distribution of coefficients") +
-        ylab("Density")
-
-colors <- c(paste0(pal[3]))
-labels <- c("actual")
-
-gg1 <- gg1 + 
-        geom_vline(aes(xintercept = true_value, color = "actual")) +
-        scale_color_manual(
-          name = "",
-          values = colors,
-          labels = labels
-        ) +
-        theme(legend.position = c(0.4, 0.6))
-
-gg1
-
-```
+The comparison of the results to those in the previous table also confirm the concerns around endogenous placement of coal plants.
+The coefficients in Table \@ref(tab:yieldtabletwo) are always more negative than those in Table \@ref(tab:yieldtablepm), and sometimes substantially so.
+In other words, the simple regression of agricultural productivity on pollution results in coefficients on pollution that are upwards biased, since higher levels of pollution occur in more productive areas, on average.
+The IV strategy corrects this bias and leads to a robust, negative relationship between pollution and agricultural productivity.
 
@@
 
-I present these results in Figure \@ref(fig:randomplot).
+I present these results in Figure \@ref(fig:randomplot) of the appendix.
 The purple density is the resulting distribution of estimated treatment effects when selecting from all villages within the same year.
-The density of treatment effects is relatively tight around 0, with the left tail only getting slightly beyond -0.05, compared to the most conservative actual estimated treatment effect of -0.73.
+The density of treatment effects is relatively tight around 0, with the left tail only reaching slightly beyond -0.05, compared to the most conservative actual estimated treatment effect of -0.63.
 
@@
 The distribution is still approximately normal but has a mode of almost -0.4.
-Nonetheless, the actual treatment effect is larger in magnitude than every single one of the 1,000 bootstrapped replications.
-If we were to use the standard deviation of the resulting distribution as a standard error, we would get a confidence interval around the main effect of approximately `r paste0(reportDec(-0.73-1.96*sd(coefs_distfe)))` to `r paste0(reportDec(-0.73+1.96*sd(coefs_distfe)))`.
-The results are apparently robust to alternative specifications of uncertainty
+Nonetheless, the actual treatment effect is larger in magnitude than almost all of the bootstrapped replications.
+The results are apparently robust to alternative specifications of uncertainty.
 
@@
 I present these results in Table \@ref(tab:yieldtableleads).
-Interestingly, there is actually a positive coefficient for the lead in the second column.
-However, if we perform randomization inference, it appears that the confidence intervals might be slightly underestimated; the positive effect size is in the 11th percentile of the distribution in Figure \@ref(fig:randomplotleads), which for a two-tailed test would not even have a p-value of less than 0.2.
-Overall, there is no indication that leads predict any changes in current agricultural productivity.
+Neither a one-year nor a two-year lead predicts current agricultural productivity, which supports the identification assumptions.
 
@@
 colnames(yield3ivmain_lead) <- c("(1)", "(2)")
-rownames(yield3ivmain_lead) <- c("Particulate matter", "(log PM 2.5)",
-                                 "Particulate matter", "(one-year lead)",
-                                 "Weather",
+rownames(yield3ivmain_lead) <- c("Particulate matter", "(one-year lead)",
+                                 "Particulate matter", "(two-year lead)",
                                  "Fixed effects:", "Village", "Year",
-                                 "F (current)", "F (lead)", 
+                                 "F", 
                                  "Observations")
@@
   column_spec(c(2:3),width = "2cm") %>%
-  row_spec(c(8), hline_after = TRUE) %>%
-  row_spec(c(6), bold = TRUE) %>%
+  row_spec(c(7), hline_after = TRUE) %>%
+  row_spec(c(5), bold = TRUE) %>%
   kable_classic_2()
@@
 
-Having established that exposure to pollution has a negative effect on agricultural productivity, I now move to heterogeneity in these effects.
+Having established that exposure to pollution has a negative effect on agricultural productivity, I now move to the heterogeneity in these effects.
 Table \@ref(tab:yieldtablehet) presents three different analyses: heterogeneity by median exposure, heterogeneity by initial yield, and possible multiplicative effects of shocks.
@@
 In other words, for each village, I calculate the median exposure (based on wind) and then use this single value for each village to split the sample into villages above and below median exposure.
-The coefficient in the first column is 2.3 times larger than the coefficient in the second column, indicating much larger effects of pollution in places where it is more common.
-One possible explanation is that exposure is cumulative, leading to larger and larger effects.
-I return to the effects of cumulative shocks in column five.
+The coefficient in the first column is slightly larger than the coefficient in the second column, indicating that pollution does not have large differential effects based on exposure.
+One possible explanation is that exposure is not cumulative, but continues to be harmful regardless of the baseline level.
 
@@
 Another way to think about this, however, is that higher values of rainfall dampen the effects of pollution shocks.
-This could have very profound effects on welfare, since this essentially increases the variance of agricultural productivity across years (which is already increasing due to climate change), at least as it relates to pollution exposure..
-
-## Labor allocation
-
-The previous set of results show a consistent negative effect of wind-driven pollution on agricultural productivity.
-A key question is the mechanism through which this happens; is this driven by changes in labor allocation/productivity or by changes in land productivity?
-While I cannot definitely prove the mechanism, I can shed some light on possibilities by looking at labor allocation decisions.
-Since these shocks generally happen after initial planting decisions, it seems unlikely that changes in land *allocation* could drive the results.
-
-Table \@ref(tab:labortable) presents results for labor allocation over the last seven days using the National Sample Survey and the methodology described above.
-Since the labor recall is for seven days, labor allocation ranges from zero to seven, as does the wind variable.^[I also tried an instrumental variables strategy with the seven-day wind and monthly pollution, but the first stage lacked power due to the difference in temporal aggregation across the two key variables (pollution and wind).] Recall that the data is at the district level, not the village level.
-However, it is also more temporally disaggregated, with rolling interviews throughout the year-long survey.
-Figure \@ref(fig:laborplot) in the appendix shows this temporal distribution across the five waves of the survey.
-Since pollution is not disaggregated at a daily level, I instead stick to the reduced form results with wind.
-
-```{r labortable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+This could have very profound effects on welfare, since this essentially increases the variance of agricultural productivity across years (which is already increasing due to climate change), at least as it relates to pollution exposure.
+
+## Crops and labor intensity
+
+A key question about the results is the mechanism.
+While previous results indicate that pollution may actually increase land productivity [@sanders2022adaptation], the results in that paper come from the United States, where agricultural practices vary greatly from those in India.
+Unfortunately, there is no good labor data that would allow me to directly test the effects of pollution on agricultural labor productivity.
+While the National Sample Survey is nationally representative, it does not allow one to back out productivity, only total labor allocation in the preceding week.
+Additional data options -- like the ARIS-REDS or IHDS -- are not collected frequently enough to enable the kind of analysis employed here.
+
+Instead, I look at heterogeneity by the modal crop grown in each area.
+I use data from @monfreda2008farming, who estimate area harvested by crop at a resolution of approximately 10km.
+These estimates come from around the year 2000, at the beginning of my sample period.
+I identify areas where rice is the most commonly grown crop.
+Rice (paddy) is commonly regarded as a labor-intensive crop in India [@vollrath2011agricultural; @michler2020agriculture], meaning that it may be more susceptible to pollution if some of the effects of pollution are through direct effects on labor productivity.
+
+```{r crops, echo = FALSE, message = FALSE, warning = FALSE, dev='png', fig.height = 5, fig.cap = "Most common crop by village", fig.footer = list("Note: The figure shows the most common crop grown in each village, by area harvested.")}
+
+village_shp <- village_shp %>% left_join(crops_all, by = c("shrid"))
+
+# plot
+ggplot() +
+  geom_sf(data = village_shp, aes(fill = most_common), color = NA) +
+  scale_fill_brewer("Most common crop\nby area harvested", palette = "Set3") +
+  geom_sf(data = india_shape2, fill = NA, lwd = 0.01) +
+  theme_minimal()
+
+```
+
+Figure \@ref(fig:crops) shows the distribution of the most common crop across the villages in my sample.
+Though there are large areas where rice is the most commonly grown crop, there is also quite a bit of overlap, especially in the middle parts of the country.
+However, given the geographic concentration of rice production, I will implement a robustness check with state-by-year fixed effects below.
+
+```{r yieldrice, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
 
 # Table
-colnames(labor1all) <- c("All", "All", "Self", "Wage", "Farm", "Non-farm")
-colnames(labor2rural) <- c("All", "All", "Self", "Wage", "Farm", "Non-farm")
-
-mat1 <- matrix(c("", "", "", "", "", ""), nrow = 1)
-colnames(mat1) <- c("All", "All", "Self", "Wage", "Farm", "Non-farm")
-rownames(mat1) <- c("Panel A: all")
-mat2 <- matrix(c("", "", "", "", "", ""), nrow = 1)
-colnames(mat2) <- c("All", "All", "Self", "Wage", "Farm", "Non-farm")
-rownames(mat2) <- c("Panel B: Rural")
-
-labortablerural <- rbind(mat1, labor1all[c(1:2, 7:8),], mat2, labor2rural[-c(3:6),], labor2rural[c(3:6),])
-rownames(labortablerural)[c(2,7)] <- "Wind"
-rownames(labortablerural)[c(5,10)] <- "Observations"
-rownames(labortablerural)[11] <- "Controls"
-rownames(labortablerural)[12] <- "Fixed effects:"
-rownames(labortablerural)[13] <- "District"
-rownames(labortablerural)[14] <- "State-month"
-
+yieldrice <- yieldrice[,-1]
+colnames(yieldrice) <- c("(1)", "(2)")
+rownames(yieldrice) <- c("Particulate matter", "(log PM 2.5)",
+                                "PM 2.5 times rice", " ",
+                                "Weather (expanded)", "Weather (expanded, bins)",
+                                "Fixed effects:", "Village", "Year",
+                                "F (PM 2.5)", "F (PM 2.5 times rice)", 
+                                "Observations")
 kable(
-      labortablerural, caption = "Wind direction and labor allocation by rural status",
-      align = "cccccc", booktabs = TRUE, linesep = ""
+      yieldrice, caption = "Agricultural productivity by most common crop",
+      align = "cc", booktabs = TRUE, linesep = ""
       ) %>%
   footnote(
-           "Note: Standard errors are in parentheses and are clustered at the district level.
-Control variables include female, age, age squared, and (years of) education.
-The outcome in each regression is days worked in the previous seven days.
+           "Note: Standard errors are in parentheses and are clustered at the village level.
+The outcome in all regressions estimated agricultural yield (tons per hectare).
+The rice dummy equals one in villages where rice is the most commonly harvested crop.
 * p<0.10 ** p<0.05 *** p<0.01",
@@
            ) %>%
-  column_spec(1, width = "3cm") %>%
-  column_spec(c(2:7),width = "1.5cm") %>%
-  row_spec(c(5,10), hline_after = TRUE) %>%
-  row_spec(c(1,6,12), bold = TRUE) %>%
+  column_spec(1, width = "3.5cm") %>%
+  column_spec(c(2:3),width = "2cm") %>%
+  row_spec(c(9), hline_after = TRUE) %>%
+  row_spec(c(7), bold = TRUE) %>%
   kable_classic_2()
@@
 
-The top panel presents results for all respondents during the five-month monsoon season while the bottom panel presents results for respondents in rural areas.
-Interestingly, we do not see any signs of large, negative effects of wind exposure on labor allocation across any sectors, with the possible exception of wage employment in rural areas.
-However, the coefficient for self-employment in rural areas is of the same magnitude and *positive*, but imprecisely estimated, making it hard to come to any definitive conclusions, at least in the aggregate.
-
-## Month-specific exposure
-
-While overall results for labor allocation are inconclusive, perhaps we can learn more by looking at temporally disaggregated results, focusing on June through October.
-I can create separate variables for exposure by each month of the season, keeping the unit of analysis the individual (for labor) or the village-season (for yields).
-
-Since the NSS data has survey dates and the both the pollution and wind data can be aggregated to the month, we can look at how exposure during different months of the agricultural season (differentially) affect labor allocation and overall agricultural productivity.
-I begin with labor allocation in columns one and two of Table \@ref(tab:labor3month).
-As before, the exposure measure is just the wind variable (over the last seven days), since I can directly match daily wind direction with the timing of the survey, but I cannot do that for pollution, which is monthly.
-The first column presents results for farm labor and the second column presents results for non-farm labor.
-
-Higher exposure leads to significantly less farm labor labor in June and July, but there are no discernible effects in other months of the year.
-There are likewise no discernible effects on non-farm labor, but some of the coefficients are rather imprecisely estimate.
-To put the effects for June and July in context and taking the coefficients at face value, an additional day of wind in the last seven days decreases farm labor in June by around `r paste0(reportPercent(0.136/1.209))` percent of the mean and in July by around `r paste0(reportPercent(0.285/1.671))` percent of the mean.
-
-```{r labor3month, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-labor3month <- cbind(labor3month, yield4monthly[-c(15:19),2])
-
-# Table
-colnames(labor3month) <- c("Farm", "Non-farm", "Yield")
-kable(
-      labor3month, caption = "Heterogeneity by month",
-      align = "ccc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the village level.
-The outcome in column one is farm labor, the outcome is column two is non-farm labor, and the outcome in column three is village-level productivity.
-The independent variable for labor is wind, while the IV approach above is still used for yield.
-* p<0.10 ** p<0.05 *** p<0.01",
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
-           ) %>%
-  column_spec(1, width = "2.5cm") %>%
-  column_spec(c(2:4),width = "2.5cm") %>%
-  row_spec(c(14), hline_after = TRUE) %>%
-  row_spec(c(12), bold = TRUE) %>%
-  add_header_above(c(" " = 1, "Labor" = 2, " " = 1)) %>%
-  kable_classic_2()
-
-```
-
-I perform a similar exercise with yield, the results of which are in the third column Table \@ref(tab:labor3month).
-I return to the IV strategy, however, to be consistent with the previous yield results.
-I include only the five months of the year that are used in the calculation of the yield estimates, or June through October.
-
-There is only one month where we see a clear effect of pollution on yield: June.
-Given that we also see negative effects of pollution on farm labor allocation in June, this is suggestive -- but not conclusive -- evidence that pollution at least partly works through labor.
-Note that this does not rule out pollution also having direct effects on agricultural land productivity, however.
-
-A final caveat about these month-specific results is that they do not take into account possible interactions between different months.
-For example, yield might respond strongly to average pollution in June and July, which would not be captured in Table \@ref(tab:labor3month).
-As one additional check, Table \@ref(tab:yield8cumulative) in the appendix presents five separate instrumental variable regressions, which look at the effects of cumulative exposure to pollution separately by month.
-Across all five months, cumulative exposure negatively affects agricultural productivity, with the largest coefficient actually showing up for October, the last month of the season.
+Table \@ref(tab:yieldrice) presents the main results in this section.
+I continue to employ an IV strategy, interacting pollution with an indicator for rice and instrumenting this with an interaction between wind and rice.
+The first row shows the effects of overall pollution on agricultural productivity in villages where rice is not the most commonly grown crop.
+Column two, which has the most flexible weather specification, indicates that an increase of particulate matter by one percent decreases productivity in these areas by around 0.4 percent.
+The second row is an interaction term between rice and PM 2.5, meaning the linear combination of the two coefficients gives the overall effect of pollution on productivity in rice-predominant villages.
+The coefficient of -0.2 shows that pollution decreases productivity by around 50 percent more in rice-growing villages than in non-rice growing villages.
+Since rice is more labor intensive, this is evidence that the effects of pollution work at least partly through the effects of pollution on labor productivity.
+This does not, however, rule out any direct effects on land productivity.
+
+Given the geographic concentration of crops, Table \@ref(tab:yieldricestatefe) in the appendix presents the same results but with state-by-year fixed effects, instead of just year fixed effects.
+This will compare the effects of pollution within states in the same year, hopefully decreasing concerns that something other than labor intensity is the main driver of the heterogeneity documented in Table \@ref(tab:yieldrice).
+The results are consistent with those presented here, and show that pollution decreases productivity more in rice-growing villages than in other villages.
 
@@
 # Table
-colnames(yield7ivdiffindiff) <- c("(1)", "(2)", "(3)")
+colnames(yield7ivdiffindiff) <- c("(1)", "(2)")
 rownames(yield7ivdiffindiff) <- c("Particulate matter", "(log PM 2.5)", 
@@
       yield7ivdiffindiff, caption = "Agricultural productivity and pollution source",
-      align = "ccc", booktabs = TRUE, linesep = ""
+      align = "cc", booktabs = TRUE, linesep = ""
       ) %>%
@@
   column_spec(1, width = "3.5cm") %>%
-  column_spec(c(2:4),width = "2cm") %>%
+  column_spec(c(2:3),width = "2cm") %>%
   row_spec(c(7), bold = TRUE) %>%
@@
 
-Columns one through three of Table \@ref(tab:yield7ivdiffindiff) present specifications with progressively increasing weather controls.
-Column three presents the most extensive set of controls and I focus on that column here.
-First, there are now two first stages but the F-tests indicate that the instruments are strong for both, including the interaction term (F = 232).
+Columns one and two of Table \@ref(tab:yield7ivdiffindiff) present specifications with slightly different weather controls.
+Column two presents the most extensive set of controls and I focus on that column here.
+First, there are now two first stages but the F-tests indicate that the instruments are strong for both, including the interaction term (F = 205 in the second column).
 Second, the first row shows the effect of pollution on agricultural productivity prior to the opening of a coal plant; the coefficient is negative, indicating negative effects of non-coal pollution.
-Third, the interaction term is strongly negative and more than twice as large in magnitude as the level term, indicating a very clear effect specific to coal plants, in addition to the overall negative effect of pollution.
+Third, the interaction term is strongly negative and approximately the same in magnitude as the level term, indicating a very clear effect specific to coal plants, in addition to the overall negative effect of pollution.
 
@@
 Using exogenous variation in pollution driven by changes in wind direction, results here show that pollution has a large, negative effect on agricultural productivity in India.
-
-These negative effects are at least partly due to changes in labor allocation decisions; the largest negative effects on agricultural productivity and on labor allocation happen in the same month of the season: June.
+These negative effects are at least partly due to changes in labor productivity; areas growing more labor-intensive crops see larger decreases in agricultural productivity due to pollution.
 Nonetheless, I am unable to rule out direct effects on land productivity with the data available.
@@
 data <- matrix(NA, nrow = 7, ncol = 3)
-rownames(data) <- c("shapefile", "coal plants", "wind", "pollution", "agriculture", "weather", "labor")
+rownames(data) <- c("shapefile", "coal plants", "wind", "pollution", "agriculture", "weather", "crops")
 colnames(data) <- c("source", "geographic coverage", "temporal coverage")
@@
 data[6,1] <- "TerraClimate"
-data[7,1] <- "National Sample Survey (NSS)"
+data[7,1] <- "Monfreda et al. (2008)"
 data[1,2] <- "India"
@@
 data[6,2] <- "global"
-data[7,2] <- "India"
+data[7,2] <- "global"
 data[1,3] <- " "
@@
 data[6,3] <- "monthly"
-data[7,3] <- "daily interviews"
+data[7,3] <- "2000 (year)"
 
@@
 ```
-```{r plants, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Coal plants in India from 1990 to 2010", fig.footer = list("Note: The top figure shows the location of coal plants in 1990.
+```{r plants, echo = FALSE, message = FALSE, warning = FALSE, dev='png', fig.cap = "Coal plants in India from 1990 to 2010", fig.footer = list("Note: The top figure shows the location of coal plants in 1990.
 The bottom figure shows the location of coal plants in 2010.")}
@@
 
-```{r windexample, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 8, fig.cap = "Wind direction and aggregation examples (2010-01-01)", fig.footer = list("Note: The top figure shows the average wind direction on January 1st, 2010, with zero degrees indicating north.
+```{r windexample, echo = FALSE, message = FALSE, warning = FALSE, dev='png', fig.height = 8, fig.cap = "Wind direction and aggregation examples (2010-01-01)", fig.footer = list("Note: The top figure shows the average wind direction on January 1st, 2010, with zero degrees indicating north.
 The points are the location of coal plants on that date.
@@
 
-```{r pollutiontable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+```{r yield3ivmainfirststage, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
 
 # Table
-colnames(pollutiontable) <- c("(1)", "(2)", "(3)", "(4)")
+yield3ivmainfirststage <- yield3ivmainfirststage[,-c(1)]
+colnames(yield3ivmainfirststage) <- c("(1)", "(2)", "(3)", "(4)")
 kable(
-      pollutiontable, caption = "Wind direction and particulate matter",
+      yield3ivmainfirststage, caption = "Wind direction and particulate matter",
       align = "cccc", booktabs = TRUE, linesep = ""
@@
   column_spec(c(2:5),width = "2cm") %>%
-  row_spec(c(3), bold = TRUE) %>%
-  row_spec(6, hline_after = TRUE) %>%
-  add_header_above(c(" ", "1998-2015" = 2, "2002-2013" = 2)) %>%
+  row_spec(6, bold = TRUE) %>%
+  row_spec(8, hline_after = TRUE) %>%
   kable_classic_2()
@@
 
-```{r randomplotleads, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 8, fig.cap = "Randomization inference for leads", fig.footer = list("Note: The figure presents the distribution of coefficients by randomly assigning a single wind value to each village, sampled with replacement across all villages in a given district in the same year.
-\"Current\" refers to the coefficient on current pollution, while \"lead\" refers to the coefficient on next year's pollution in the same table.")}
+```{r randomplot, echo = FALSE, message = FALSE, warning = FALSE, dev='png', fig.height = 4, fig.cap = "Randomization inference", fig.footer = list("Note: The figure presents the distribution of coefficients by randomly assigning wind direction across villages within the same year.
+The \"all\" distribution is created by randomly assigning a single wind value to each village, sampled with replacement across all villages in a given year.
+The \"within district\" randomly samples with replacement only from within the same district in the same year.")}
 
 pal <- viridis(4)
-colors <- c(paste0(pal[1]), paste0(pal[3]))
-labels <- c("current", "lead")
+colors <- c(paste0(pal[1]), paste0(pal[2]))
+labels <- c("All", "Within district")
 
 # plot
-ggplot() + 
-    geom_density(aes(x = coef_current, fill = "current"), alpha = 0.5) +
-    geom_density(aes(x = coef_lead, fill = "lead"), alpha = 0.5) +
+gg1 <- ggplot() + 
+        geom_density(aes(x = coef_vec, fill = "All"), alpha = 0.5) +
+        geom_density(aes(x = coefs_distfe, fill = "Within district"), alpha = 0.5) +
         scale_fill_manual(
-      name = "coefficient:",
+          name = "Randomization:",
           values = colors,
@@
         theme_minimal() +
-    xlab("distribution of coefficients")
+        xlab("Distribution of coefficients") +
+        ylab("Density")
+
+colors <- c(paste0(pal[3]))
+labels <- c("actual")
+
+gg1 <- gg1 + 
+        geom_vline(aes(xintercept = true_value, color = "actual")) +
+        scale_color_manual(
+          name = "",
+          values = colors,
+          labels = labels
+        ) +
+        theme(legend.position = c(0.4, 0.6))
+
+gg1
 
@@
 \newpage
-```
-```{r laborplot, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Timing of household surveys for the National Sample Survey", fig.footer = list("Note: The histogram shows the distribution of survey dates for the four waves of the National Sample Survey (NSS) for the five waves used in this paper.")}
-
-laborplot
-
-```
-
-```{=tex}
-\FloatBarrier
-\newpage
-```
-```{r monthplot, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Average farm days by season, rural only", fig.footer = list("Note: The line graph shows the mean number of farm days worked per week based on the month of the relevant agricultural season.
-The monsoon season is defined as June (month 1) to October (month 5) while the rainy season is defined as November (month 1) to March (month 5) of the following year.
-The data is from the NSS.")}
-
-monthplot
-
 ```
@@
 ```
+
+```{=tex}
+\FloatBarrier
+\newpage
+```
+```{r yieldricestatefe, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+# Table
+yieldricestatefe <- yieldricestatefe[,-1]
+colnames(yieldricestatefe) <- c("(1)", "(2)")
+rownames(yieldricestatefe) <- c("Particulate matter", "(log PM 2.5)",
+                                "PM 2.5 times rice", " ",
+                                "Weather (expanded)", "Weather (expanded, bins)",
+                                "Fixed effects:", "Village", "Year",
+                                "F (PM 2.5)", "F (PM 2.5 times rice)", 
+                                "Observations")
+kable(
+      yieldricestatefe, caption = "Agricultural productivity by most common crop",
+      align = "cc", booktabs = TRUE, linesep = ""
+      ) %>%
+  footnote(
+           "Note: Standard errors are in parentheses and are clustered at the village level.
+The outcome in all regressions estimated agricultural yield (tons per hectare).
+The rice dummy equals one in villages where rice is the most commonly harvested crop.
+* p<0.10 ** p<0.05 *** p<0.01",
+           general_title = "",
+           threeparttable = TRUE,
+           footnote_as_chunk = TRUE
+           ) %>%
+  column_spec(1, width = "3.5cm") %>%
+  column_spec(c(2:3),width = "2cm") %>%
+  row_spec(c(9), hline_after = TRUE) %>%
+  row_spec(c(7), bold = TRUE) %>%
+  kable_classic_2()
+
+```

### 2023-07-10 draft/draft.tex
Updating readme
updated draft; labor intensity

@@
 \author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
-\date{2023-04-12}
+\date{2023-07-10}
 
@@
 \begin{abstract}
-\noindent I document externalities in the agricultural sector of India due to air pollution.
-Using variation in pollution induced by changes in wind, I show that higher levels of particulate matter lead to decreased agricultural productivity, with changes in productivity of 10 percent being common.
-Combining wind direction with the rollout of coal plants, results indicate pollution from coal plants has a larger effect on agricultural productivity than other types of pollution.
-Given that the agricultural sector is a refuge for the poor in many developing countries, these results suggest that the effects of pollution may hit the poorest particularly hard.\\
+\noindent I document negative externalities of air pollution in the Indian agricultural sector.
+Using variation in pollution induced by changes in wind across years, I show that higher levels of pollution lead to decreased agricultural productivity, with large changes in productivity being common.
+The negative effects of pollution are larger in areas growing more labor-intensive crops, indicating that the pollution works at least partly through direct effects on labor productivity.
+Finally, combining wind direction with the rollout of coal plants, results indicate that pollution from coal plants has a larger effect on agricultural productivity than other types of pollution.
+Given that the agricultural sector is a refuge for the poor in many developing countries, these results suggest that the negative externalities of pollution may hit the poorest particularly hard.\\
 \strut \\
@@
 
-There are no doubt positive effects from this increase in power generation; for example, electricity may lead to economic and productivity growth (Dinkelman 2011; Kline and Moretti 2014; Rud 2012; Van de Walle et al.
+While there are likely positive effects from this increase in power generation -- for example, electricity may lead to economic and productivity growth (Dinkelman 2011; Kline and Moretti 2014; Rud 2012; Van de Walle et al.
 2017), though electrification alone may not be a sufficient condition (K.
-Lee, Miguel, and Wolfram 2020).
-These improvements do not come without downsides, however, especially when it comes to power derived from coal.
+Lee, Miguel, and Wolfram 2020) -- these improvements do not come without downsides, especially when it comes to power derived from coal.
 Pollutants from coal contribute to climate change and the phase out of coal energy is therefore a major climate goal (IEA 2022).
-Emissions from the production and burning of coal are also directly harmful to health and the environment, including pollutants like sulfur dioxide, nitrogen oxides, and mercury, to name but a few.\footnote{www.eia.gov/energyexplained/coal/coal-and-the-environment.php} While the use of coal had been declining, recent events -- including Russia's invasion of Ukraine -- have led to an increase in the demand for coal.\footnote{www.npr.org/2022/08/15/1117560560/a-rising-demand-for-coal-amidst-war-in-ukraine}
-
-In addition to coal, pollution more generally is increasing in much of the world.
-Carbon dioxide equivalent per capita has almost double over the last 50 years in India, more than doubled in Indonesia, and almost tripled in China.\footnote{\url{https://edgar.jrc.ec.europa.eu}} Of course, much of this increase is due to the development process itself; as countries develop, their economies transform, with more people moving out of agriculture and into manufacturing and services (Herrendorf, Rogerson, and Valentinyi 2014).
+Emissions from the production and burning of coal are also directly harmful to human health and the environment; pollutants like sulfur dioxide, nitrogen oxides, methane, and mercury are particularly harmful.\footnote{www.eia.gov/energyexplained/coal/coal-and-the-environment.php} While the use of coal had been declining, recent events -- including Russia's invasion of Ukraine -- have led to an increase in the demand for coal.\footnote{www.npr.org/2022/08/15/1117560560/a-rising-demand-for-coal-amidst-war-in-ukraine}
+
+In addition to pollution from coal, pollution more generally is also increasing in much of the world.
+Carbon dioxide equivalent per capita has almost double over the last 50 years in India, more than doubled in Indonesia, and almost tripled in China.\footnote{\url{https://edgar.jrc.ec.europa.eu}} Much of this increase is due to the development process itself -- as countries develop, their economies transform, with more people moving out of agriculture and into manufacturing and services (Herrendorf, Rogerson, and Valentinyi 2014).
 Despite this development, agriculture remains a key sector in many developing countries, employing a large proportion of the poor (Dercon 2009).
@@
 2020).
-In the case of the data I use here, coal plants are constructed in areas with higher population and higher literacy.
-In terms of agriculture, higher productivity predicts more coal plants in a given year as well as the construction of a coal plant over the next decade, indicating possible bias when estimating OLS regressions of agricultural productivity on pollution levels.
+In the case of the data I use here, coal plants are constructed in areas with higher population and higher levels of pollution.
+In terms of agriculture, higher productivity predicts more coal plants in a given year as well as the construction of a coal plant over the next decade, indicating possible bias when estimating effects of pollution on agricultural productivity.
 
@@
 
-I find that these effects are mostly transmitted by reductions in farm labor during narrow windows of the growing season, using labor allocation data from the National Sample Survey (NSS).
-While there do not appear to be large changes in labor allocation patterns when looking at the season as a whole, there is a clear pattern when we break down exposure to pollution by month: higher levels of pollution at the beginning of the monsoon season (June and July) lead to lower levels of farm labor allocation, but no changes in non-farm labor allocation.
-This matches up well with a similar exercise with agricultural yields: pollution in June is the only month in which there is a negative effect of pollution on agricultural yield.
-In tandem, these two results point to effects on labor as one mechanism driving the total effects, though I cannot rule out direct effects of pollution on land productivity with my data.
+A key question is the mechanism: does pollution reduce labor productivity, land productivity, or both?
+While previous work from the United States indicates that pollution from coal plants, specifically, can lead to higher \emph{land} productivity (Sanders and Barreca 2022), agriculture in India differs greatly from agriculture in the United States.
+To test for possible effects on labor productivity, I look at differential effects based on labor intensity of the predominant crops in different villages.
+Specifically, I separate effects in predominantly rice-growing areas from other areas; since rice is very labor intensive, we would expect to see larger negative effects of pollution in rice-growing regions relative to other regions if one mechanism through which pollution affects productivity is direct effects on labor.
+I show that pollution negatively effects agricultural productivity in non-rice areas but that the effects in rice-growing areas are around 50 percent larger in magnitude.
+These results indicate that pollution likely directly affects labor productivity, though I am not able to rule out direct effects on land, either, due to a lack of data.
 
 The final set of analyses tries to disentangle the effects of pollution from the effects of coal pollution, specifically, by using wind direction and the rollout of coal plants across the country.
-Combining instrumental variables with a differences-in-differences strategy, I show that agricultural productivity has an elasticity of around -0.15 with respect to non-coal pollution.
-The opening of a coal plant, however, triples the elasticity, to more than -0.45, indicating that coal plant emissions are particularly harmful to agricultural productivity, at least in India.
-This is consistent with coal emissions being quite different from other forms of pollution; for example, coal emissions contain much more mercury than other forms of pollution.
+Combining instrumental variables with a differences-in-differences strategy, I show that agricultural productivity has an elasticity of around -0.22 with respect to non-coal pollution.
+The opening of a coal plant, however, doubles the elasticity, to more than -0.44, indicating that coal plant emissions are particularly harmful to agricultural productivity, at least in India.
+This is consistent with coal emissions being quite different from other forms of pollution; coal emissions contain different types and quantities of harmful substances, meaning effects may differ greatly.
+
+An important caveat to these results is that I cannot disentangle exactly which pollutants drive results.
+While I use PM 2.5 as a measure of overall exposure to pollution, effects could be driven by any pollutants carried by the wind.
+This leaves open an important avenue of future research regarding the effects of different types of pollutants on agricultural productivity.
 
 This paper contributes to several strands of literature.
-First, we already have evidence of effects of pollution on different forms of labor activity.
+We already have evidence of effects of pollution on different forms of labor activity.
 For example, general levels of pollution decrease productivity of call center works in China (Chang et al.
@@
 Also in China, He, Liu, and Salvo (2019) find small but significant negative effects of pollution on manufacturing productivity, while Chen, Oliva, and Zhang (2022) show pollution leads to labor out migration.
-Perhaps relevant to my labor allocation results, Hanna and Oliva (2015) show that decreases in pollution lead to increases in labor allocation, while I show the same relationship in at least a subset of months of the agricultural season.
+Relevant to my results, Hanna and Oliva (2015) show that decreases in pollution lead to increases in labor allocation in another developing country.
+
 Several papers are closely related to this one.
@@
 Finally, Sanders and Barreca (2022) actually find that coal emissions can be good for agriculture in the United States, due to the presence of sulfur, which can increase agricultural productivity.
-This paper differs in three important ways.
+This paper differs from the previous literature in three important ways.
 First, I explore effects of pollution more generally as well as from coal plants, which are much more ubiquitous than gold mines across the world.
-Second, I use a different identification strategy from either paper, relying on wind blowing pollution from coal plants.
+Second, I use a different identification strategy from most other papers, relying on wind blowing pollution from coal plants.
 Finally, agriculture in India and the United States can be quite different, with the latter being much more mechanized.
-This means effects of pollution could differ substantially.
-
-I also contribute to the more general literature on the effects of pollution on health.
-Pollution has substantial negative effects on health in both developing and developed countries (Arceo, Hanna, and Oliva 2016) and we have known of these negative effects for many years (Brunekreef and Holgate 2002; Kampa and Castanas 2008; Pope III and Dockery 2006).
-Currie et al.
-(2014) review the literature on the health effects of pollution during adolescence; we have robust evidence of pollution exposure leading to increases in infant mortality, including in developing countries (Heft-Neal et al.
-2018, 2019).
-Since these effects are often driven by heart and lung disease and asthma (Kampa and Castanas 2008), this my explain the labor allocation effects I identify in this paper.
-
-The rest of the paper is organized as follows.
-I first discuss data and then methodology, in two separate sections.
-The methodology section discusses the empirical strategy and validates the wind direction variable using data on particulate matter.
-I present the main results in \autoref{results} before concluding in \autoref{conclusion}.
+This means effects of pollution could differ substantially depending on the context, a conclusion that the current results support.
 
@@
 
-To help understand some of the mechanisms driving the relationship between exposure to emissions and agricultural productivity, I also include household survey data.
-Unfortunately, I am not aware of any publicly available survey data that covers multiple years and has identifiers below the district level.
-For example, the ARIS-REDS data\footnote{www.ncaer.org/project/additional-rural-incomes-survey-aris-rural-economic-demographic-survey-reds} includes GPS identifiers for one year, but that only allows GPS matching for a total of two years: 1999 and 2007/2008.
-The DHS, on the other hand, only has state-level identifiers (except for the newest two waves) and is also not a labor survey.
-Instead, I use the National Sample Survey (NSS) from the Ministry of Statistics and Programme Implementation.
-
-The NSS is a nationally representative survey that was conducted every couple of years until relatively recently.
-Many rounds of the NSS also included a module on employment that asked about labor allocation over the previous seven days.
-I use the 61st, 62nd, 64th, 66th, and 68th rounds of the NSS, all of which include the employment module.
-The publicly available data also includes the date of survey, which allows me to match pollution exposure to the exact day of the survey.
-
-The biggest issue with using the NSS is that it includes only district-level identifiers.
-The main exposure variables, on the other hand, are defined at a lower level of aggregation, at the village.
-I choose to aggregate this village-level exposure to the district level as shown in the second panel of Figure \ref{fig:windexample} -- the district is a randomly selected district in the data, Guna district in Madhya Pradesh.
-I use the area overlap of the villages with the district to define a weighted average of exposure for the district at a given time.
-I calculate the area of overlap for each village and the district as well as the overall area of the district and use these variables to define weights for individual villages.
-Any part of the district that does not overlap with one of the villages -- the white area in Figure \ref{fig:windexample} -- does not have any exposure and, as such, receives a value of zero for exposure.
+To help understand some of the mechanisms driving the relationship between exposure to emissions and agricultural productivity, I also look at the differential effects by the labor intensity of the predominant crop in each area.
+For this, I use data from (Monfreda, Ramankutty, and Foley 2008), who estimate area harvested for different crops across the global around the year 2000.
+I identify areas where rice (paddy) is the most commonly grown crop, under the assumption that rice is more labor intensive than other crops (Vollrath 2011; Michler 2020).
+Michler (2020), for example, shows that rice, on avearge, uses almost 1,800 hours per hectare, while the second-highest crop in his data, cotton, juses just 850.\footnote{He estimates that wheat, another common crop in India, uses just 600 hours per hectare.}
 
@@
 
-So what are the characteristics of areas near the construction of new coal plants?
+What are the characteristics of areas near the construction of new coal plants?
 Table \ref{tab:plantresultstable} presents four separate regressions, using state-level fixed effects due to the spatial variation in coal plants.
@@
 \midrule
-Population (log) & 0.009 & 0.007* & 0.013* & 0.005\\
- & (0.006) & (0.004) & (0.006) & (0.004)\\
-Area (log) & -0.007* & -0.004 & -0.007* & -0.004\\
- & (0.004) & (0.005) & (0.004) & (0.003)\\
-Literacy (prop) & 0.080** & 0.037 & 0.073* & 0.036\\
- & (0.029) & (0.028) & (0.039) & (0.026)\\
-Ag. productivity &  &  & 0.014** & 0.028***\\
- &  &  & (0.005) & (0.009)\\
-Pollution &  &  & 0.047 & 0.014\\
-(log PM 2.5) &  &  & (0.045) & (0.018)\\
+Population (log) & 0.012*** & 0.009*** & 0.015*** & 0.006***\\
+ & (0.001) & (0.000) & (0.000) & \vphantom{1} (0.000)\\
+Area (log) & -0.010*** & -0.005*** & -0.010*** & -0.005***\\
+ & (0.001) & (0.000) & (0.000) & (0.000)\\
+Urban pop (prop) & -0.029 & 0.006 & -0.003 & -0.007\\
+ & (0.037) & (0.035) & (0.022) & (0.018)\\
+Ag. productivity &  &  & 0.014*** & 0.028***\\
+ &  &  & (0.001) & (0.001)\\
+Pollution &  &  & 0.051*** & 0.016***\\
+(log PM 2.5) &  &  & (0.002) & (0.001)\\
 Sub-sample & all & no plant & all & no plant\\
 \midrule
-Observations & 283,021 & 264,265 & 508,784 & 469,004\\
+Observations & 283,021 & 264,265 & 508,786 & 469,005\\
 \bottomrule
@@
 \item Note: State fixed effects are included in regressions.
-Standard errors are in parentheses and clustered at the state level.
+Robust standard errors are in parentheses.
 The outcome in the first column is whether the village is within 30km of a coal plant in 1991.
@@
 It is worth taking the time to validate the use of wind direction as a measure of exposure to pollution.
-Consider the data used for the regressions presented in Table \ref{tab:pollutiontable} of the appendix.
+Consider the data used for the regressions presented in Table \ref{tab:yield3ivmainfirststage} of the appendix.
 The outcome variable is particulate matter -- specifically, \(\mathrm{PM 2.5}\), which is particulate matter no larger than 2.5 micrometers in diameter -- which comes from Hammer et al.
 (2020).
-\(\mathrm{PM 2.5}\) is one of the harmful byproducts from coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.\footnote{www.epa.gov/airmarkets/power-plants-and-neighboring-communities} This data is available at the monthly level, so in Table \ref{tab:pollutiontable}, I aggregate wind exposure to the monthly level, as well.
-
-Both the particulate matter data and the wind data is at the village level, which means the regressions in Table \ref{tab:pollutiontable} are at the village level.
+\(\mathrm{PM 2.5}\) is one of the harmful byproducts from coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.\footnote{www.epa.gov/airmarkets/power-plants-and-neighboring-communities} This data is available at the monthly level, but I aggregate variables to the season variable since that is the level of analysis in the main results.
+
+Both the particulate matter data and the wind data is at the village level, which means the regressions in Table \ref{tab:yield3ivmainfirststage} are at the village level.
 As such, the geographic fixed effects are village fixed effects and the temporal fixed effects are month fixed effects.
@@
 
-In addition to serving as an example, Table \ref{tab:pollutiontable} also helps validate the use of wind direction as a proxy for exposure to pollution.
+In addition to serving as an example, Table \ref{tab:yield3ivmainfirststage} also helps validate the use of wind direction as a proxy for exposure to pollution.
 The first two columns include all available years of data while the last two columns restrict estimation only to the years for which the agricultural productivity regressions are estimated.
@@
 
-To put context behind pollution numbers, the World Health Organization's updated guidelines are that average annual exposure should not exceed 15 \(\mathrm{\mu g/m^3}\).
-For the month, an increase at the midpoint of the estimated range is equal to an increase of 0.2 percent of the maximum recommended mean concentration from the WHO.
-This is just a single day of wind in the direction of a village.
-Since this includes all villages located within 30km of a coal plant, this is evidence that wind direction can have serious repercussions on the health of hundreds of millions of people in India.
+\hypertarget{identifying-mechanisms}{%
+\subsection{Identifying mechanisms}\label{identifying-mechanisms}}
+
+To identify possible mechanisms, I identify areas that grow predominantly rice, which is a labor-intensive crop. Specifically, we are interested in the following regression:
+\begin{gather} 
+y_{it} = \alpha_{i} + \gamma_{t} + \omega PM_{it} + \phi rice_{it} + \beta PM_{it}\times rice_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it},
+\end{gather}
+where \(PM_{it}\) denotes the effect of pollution in non-rice areas and \(PM_{it}\times rice_{it}\) is the differential effect in rice-growing areas.
+Here, I create an indicator variable for rice areas that is time invariant, so \(rice_{it}\) drops out of the estimation.
+I instrument for \(PM_{it}\) and for \(PM_{it}\times rice_{it}\) using wind and wind interacted with the indicator for rice.
 
@@
 I first present the reduced form results for agricultural productivity.
-The outcome of all the regressions in this section is the log of agricultural land productivity, which is defined as tons per hectare.
+The outcome of all regressions in this section is the log of agricultural land productivity, which is defined as tons per hectare.
 The first set of results are in Table \ref{tab:yieldtable}.
-The unit of analysis is the village-year -- from 2002 to 2013 -- focusing only on the monsoon (kharif) season.
+The unit of analysis is the village-season -- from 2002 to 2013 -- focusing only on the monsoon (kharif) season.
 The data used to construct the yield measure is calculated using five months of data, so the wind variable is defined for the same five months and is the sum of days in which the wind is pointed towards the village centroid from a coal plant.
@@
 \midrule
-Wind days & -0.0078*** & -0.0069*** & -0.0062***\\
- & (0.0007) & (0.0007) & (0.0006)\\
-Rain (z) &  & 0.0854*** & -0.0424***\\
- &  & (0.0018) & (0.0046)\\
-Mean temperature (10s) &  & -0.0021*** & 0.1321***\\
- &  & (0.0001) & (0.0034)\\
-Rain squared &  &  & -0.0010***\\
- &  &  & (0.0001)\\
-Temperature squared &  &  & -0.0002***\\
+Wind days & -0.0078*** & -0.0065*** & -0.0051***\\
+ & (0.0007) & (0.0007) & (0.0007)\\
+Rain (z) &  & 0.0258*** & 0.2545***\\
+ &  & (0.0004) & (0.0100)\\
+Mean temperature (10s) &  & -0.0023*** & 0.0654***\\
+ &  & (0.0002) & (0.0028)\\
+Rain squared &  &  & -0.0410***\\
+ &  &  & (0.0004)\\
+Temperature squared &  &  & -0.0001***\\
  &  &  & \vphantom{1} (0.0000)\\
-Rain times Temperature &  &  & 0.0003***\\
+Rain times Temperature &  &  & -0.0007***\\
  &  &  & (0.0000)\\
@@
 \midrule
-Particulate matter & -0.042*** & -0.044*** & -0.032***\\
+Particulate matter & -0.042*** & -0.045*** & -0.041***\\
 (log PM 2.5) & (0.003) & (0.003) & (0.003)\\
-Weather & No & Yes & Yes\\
-Weather (expanded) & No & No & Yes\\
+Weather & No & Yes & No\\
+Weather (expanded) & No & No & No\\
 \textbf{Fixed effects:} & \textbf{} & \textbf{} & \textbf{}\\
@@
 This means that the simple regressions in Table \ref{tab:yieldtablepm} might actually be biased upwards: although pollution might negatively affect productivity, pollution is higher in more productive areas due to the endogenous placement of coal plants in the study areas.
-To recover the actual effect of pollution, we need an instrument.
+One possible way to recover the actual effect of pollution on agricultural productivity is with an instrument.
 If, conditional on the fixed effects and weather variables, wind direction is exogenous and only affects agricultural productivity through pollution, then it is a valid instrument for pollution.
 
-I present the IV results in Table \ref{tab:yieldtabletwo}, with the first-stage F-statistic at the bottom.
-The first-stage results make clear how strong of an instrument wind is; all three F-statistics are above 1,000, meeting even the higher bar for the first-stage F-test argued for by D.
+I present these IV results in Table \ref{tab:yieldtabletwo}, with the first-stage F-statistic at the bottom.
+The first-stage results -- the full results are in Table \ref{tab:yield3ivmainfirststage} of the appendix -- make clear how strong of an instrument wind is; all three F-statistics are above 1,000, meeting even the higher bar for the first-stage F-test argued for by D.
 Lee et al.
 (2022).
-Also note that the coefficients are quite different from those in Table \ref{tab:pollutiontable} because the unit of analysis in Table \ref{tab:yieldtabletwo} is the season, while in Table \ref{tab:pollutiontable} the unit of analysis is the month.
 
 The second-stage results first indicate that \(\mathrm{PM 2.5}\) is a very strong predictor of agricultural productivity.
-Using the results in column three, a one percent increase in particulate matter leads to a 0.73 percent decrease in agricultural productivity.
-To put this in further perspective, the mean particulate matter concentration is just short of 40 (39.68) in the monsoon sample, while the within-village absolute deviation in particulate matter is 3.98, meaning year-to-year swings of 7-to-14 percent in agricultural productivity due to pollution exposure are common.
-
-The comparison of the results to those in the previous table also confirm the concerns around endogenous placement of coal plants.
-The coefficients in Table \ref{tab:yieldtabletwo} are always more negative than those in Table \ref{tab:yieldtablepm}, and sometimes substantially so.
-In other words, the simple regression of agricultural productivity on pollution results in coefficients on pollution that are upwards biased, since higher levels of pollution occur in more productive areas, on average.
-The IV strategy corrects this bias and leads to a robust, negative relationship between pollution and agricultural productivity.
+Using the results in column three, a one percent increase in particulate matter leads to a 0.62 percent decrease in agricultural productivity.
+To put this in further perspective, the mean particulate matter concentration is just short of 40 (39.68) in the monsoon sample, while the within-village absolute deviation in particulate matter is 3.98, meaning year-to-year swings of several percent in agricultural productivity due to pollution exposure are common.
+
+Column four presents one last specification; it includes a much more flexible specification for weather.
+Specifically, I turn the linear z-score weather variable into bins for each month of the season.
+I then interact this rainfall variable with the temperature variable.
+The coefficient in column four shows almost not change from those in column three, relieving some concerns that weather may not be correctly specified, leading to a spurious correlation between particulate matter and agricultural productivity.
 
@@
 \begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
 \toprule
-  & (1) & (2) & (3)\\
-\midrule
-Particulate matter & -0.941*** & -0.813*** & -0.728***\\
-(log PM 2.5) & (0.078) & (0.076) & (0.075)\\
-Weather & No & Yes & Yes\\
-Weather (expanded) & No & No & Yes\\
-\textbf{Fixed effects:} & \textbf{} & \textbf{} & \textbf{}\\
-Village & Yes & Yes & Yes\\
-Year & Yes & Yes & Yes\\
-\midrule
-F (first stage) & 1,236 & 1,263 & 1,264\\
-Observations & 1,266,922 & 1,266,922 & 1,266,922\\
+  & (1) & (2) & (3) & (4)\\
+\midrule
+Particulate matter & -0.941*** & -0.769*** & -0.621*** & -0.626***\\
+(log PM 2.5) & (0.078) & (0.075) & (0.080) & (0.077)\\
+Weather & No & Yes & Yes & No\\
+Weather (expanded) & No & No & Yes & No\\
+Weather (expanded, bins) & No & No & No & Yes\\
+\textbf{Fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+Village & Yes & Yes & Yes & Yes\\
+Year & Yes & Yes & Yes & Yes\\
+\midrule
+F (first stage) & 1,236 & 1,266 & 1,205 & 1,246\\
+Observations & 1,266,922 & 1,266,922 & 1,266,922 & 1,266,922\\
 \bottomrule
@@
 
-\begin{figure}
-\includegraphics{draft_files/figure-latex/randomplot-1} \caption[Randomization inference]{Randomization inference}\label{fig:randomplot}\floatfoot*{Note: The figure presents the distribution of coefficients by randomly assigning wind direction across villages within the same year.
-The "all" distribution is created by randomly assigning a single wind value to each village, sampled with replacement across all villages in a given year.
-The "within district" randomly samples with replacement only from within the same district in the same year.}
-\end{figure}
+The comparison of the results to those in the previous table also confirm the concerns around endogenous placement of coal plants.
+The coefficients in Table \ref{tab:yieldtabletwo} are always more negative than those in Table \ref{tab:yieldtablepm}, and sometimes substantially so.
+In other words, the simple regression of agricultural productivity on pollution results in coefficients on pollution that are upwards biased, since higher levels of pollution occur in more productive areas, on average.
+The IV strategy corrects this bias and leads to a robust, negative relationship between pollution and agricultural productivity.
 
@@
 
-I present these results in Figure \ref{fig:randomplot}.
+I present these results in Figure \ref{fig:randomplot} of the appendix.
 The purple density is the resulting distribution of estimated treatment effects when selecting from all villages within the same year.
-The density of treatment effects is relatively tight around 0, with the left tail only getting slightly beyond -0.05, compared to the most conservative actual estimated treatment effect of -0.73.
+The density of treatment effects is relatively tight around 0, with the left tail only reaching slightly beyond -0.05, compared to the most conservative actual estimated treatment effect of -0.63.
 
@@
 The distribution is still approximately normal but has a mode of almost -0.4.
-Nonetheless, the actual treatment effect is larger in magnitude than every single one of the 1,000 bootstrapped replications.
-If we were to use the standard deviation of the resulting distribution as a standard error, we would get a confidence interval around the main effect of approximately -0.899 to -0.561.
-The results are apparently robust to alternative specifications of uncertainty
+Nonetheless, the actual treatment effect is larger in magnitude than almost all of the bootstrapped replications.
+The results are apparently robust to alternative specifications of uncertainty.
 
@@
 I present these results in Table \ref{tab:yieldtableleads}.
-Interestingly, there is actually a positive coefficient for the lead in the second column.
-However, if we perform randomization inference, it appears that the confidence intervals might be slightly underestimated; the positive effect size is in the 11th percentile of the distribution in Figure \ref{fig:randomplotleads}, which for a two-tailed test would not even have a p-value of less than 0.2.
-Overall, there is no indication that leads predict any changes in current agricultural productivity.
+Neither a one-year nor a two-year lead predicts current agricultural productivity, which supports the identification assumptions.
 
@@
 \midrule
-Particulate matter & -1.152*** & -1.067***\\
-(log PM 2.5) & (0.075) & (0.071)\\
-Particulate matter & 0.068 & 0.268***\\
-(one-year lead) & (0.072) & (0.074)\\
-Weather & No & Yes\\
+Particulate matter & -0.073 & \\
+(one-year lead) & (0.066) & \\
+Particulate matter &  & -0.079\\
+(two-year lead) &  & (0.051)\\
 \textbf{Fixed effects:} & \textbf{} & \textbf{}\\
@@
 \midrule
-F (current) & 974 & 1,036\\
-F (lead) & 612 & 597\\
-Observations & 1,161,265 & 1,161,265\\
+F & 603 & 746\\
+Observations & 1,161,265 & 1,055,562\\
 \bottomrule
@@
 
-Having established that exposure to pollution has a negative effect on agricultural productivity, I now move to heterogeneity in these effects.
+Having established that exposure to pollution has a negative effect on agricultural productivity, I now move to the heterogeneity in these effects.
 Table \ref{tab:yieldtablehet} presents three different analyses: heterogeneity by median exposure, heterogeneity by initial yield, and possible multiplicative effects of shocks.
@@
 \midrule
-Particulate matter & -0.767*** & -0.328*** & -1.130*** & -0.690*** & -0.409**\\
-(log PM 2.5) & (0.104) & (0.095) & (0.060) & (0.146) & (0.160)\\
-PM times Rain &  &  &  &  & 0.061***\\
- &  &  &  &  & (0.014)\\
+Particulate matter & -0.415*** & -0.336*** & -1.026*** & -0.568*** & -0.306***\\
+(log PM 2.5) & (0.114) & (0.099) & (0.060) & (0.166) & (0.077)\\
+PM times Rain &  &  &  &  & 0.119***\\
+ &  &  &  &  & (0.002)\\
 Weather & Yes & Yes & Yes & Yes & Yes\\
@@
 \midrule
-F (1st stage, PM) & 684 & 491 & 688 & 484 & 865\\
-F (1st stage, PM times rain) &  &  &  &  & 1,437\\
+F (1st stage, PM) & 597 & 482 & 675 & 428 & 674\\
+F (1st stage, PM times rain) &  &  &  &  & 6,393\\
 Observations & 617,804 & 649,118 & 634,342 & 632,580 & 1,266,922\\
@@
 Another way to think about this, however, is that higher values of rainfall dampen the effects of pollution shocks.
-This could have very profound effects on welfare, since this essentially increases the variance of agricultural productivity across years (which is already increasing due to climate change), at least as it relates to pollution exposure..
-
-\hypertarget{labor-allocation}{%
-\subsection{Labor allocation}\label{labor-allocation}}
-
-The previous set of results show a consistent negative effect of wind-driven pollution on agricultural productivity.
-A key question is the mechanism through which this happens; is this driven by changes in labor allocation/productivity or by changes in land productivity?
-While I cannot definitely prove the mechanism, I can shed some light on possibilities by looking at labor allocation decisions.
-Since these shocks generally happen after initial planting decisions, it seems unlikely that changes in land \emph{allocation} could drive the results.
-
-Table \ref{tab:labortable} presents results for labor allocation over the last seven days using the National Sample Survey and the methodology described above.
-Since the labor recall is for seven days, labor allocation ranges from zero to seven, as does the wind variable.\footnote{I also tried an instrumental variables strategy with the seven-day wind and monthly pollution, but the first stage lacked power due to the difference in temporal aggregation across the two key variables (pollution and wind).} Recall that the data is at the district level, not the village level.
-However, it is also more temporally disaggregated, with rolling interviews throughout the year-long survey.
-Figure \ref{fig:laborplot} in the appendix shows this temporal distribution across the five waves of the survey.
-Since pollution is not disaggregated at a daily level, I instead stick to the reduced form results with wind.
+This could have very profound effects on welfare, since this essentially increases the variance of agricultural productivity across years (which is already increasing due to climate change), at least as it relates to pollution exposure.
+
+\hypertarget{crops-and-labor-intensity}{%
+\subsection{Crops and labor intensity}\label{crops-and-labor-intensity}}
+
+A key question about the results is the mechanism.
+While previous results indicate that pollution may actually increase land productivity (Sanders and Barreca 2022), the results in that paper come from the United States, where agricultural practices vary greatly from those in India.
+Unfortunately, there is no good labor data that would allow me to directly test the effects of pollution on agricultural labor productivity.
+While the National Sample Survey is nationally representative, it does not allow one to back out productivity, only total labor allocation in the preceding week.
+Additional data options -- like the ARIS-REDS or IHDS -- are not collected frequently enough to enable the kind of analysis employed here.
+
+Instead, I look at heterogeneity by the modal crop grown in each area.
+I use data from Monfreda, Ramankutty, and Foley (2008), who estimate area harvested by crop at a resolution of approximately 10km.
+These estimates come from around the year 2000, at the beginning of my sample period.
+I identify areas where rice is the most commonly grown crop.
+Rice (paddy) is commonly regarded as a labor-intensive crop in India (Vollrath 2011; Michler 2020), meaning that it may be more susceptible to pollution if some of the effects of pollution are through direct effects on labor productivity.
+
+\begin{figure}
+\includegraphics{draft_files/figure-latex/crops-1} \caption[Most common crop by village]{Most common crop by village}\label{fig:crops}\floatfoot*{Note: The figure shows the most common crop grown in each village, by area harvested.}
+\end{figure}
+
+Figure \ref{fig:crops} shows the distribution of the most common crop across the villages in my sample.
+Though there are large areas where rice is the most commonly grown crop, there is also quite a bit of overlap, especially in the middle parts of the country.
+However, given the geographic concentration of rice production, I will implement a robustness check with state-by-year fixed effects below.
 
@@
 
-\caption{\label{tab:labortable}Wind direction and labor allocation by rural status}
+\caption{\label{tab:yieldrice}Agricultural productivity by most common crop}
 \centering
 \begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3.5cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
 \toprule
-  & All & All & Self & Wage & Farm & Non-farm\\
-\midrule
-\textbf{Panel A: all} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-Wind & -0.011 & -0.009 & 0.030 & -0.049 & -0.009 & -0.010\\
- & (0.031) & (0.034) & (0.039) & (0.034) & (0.034) & (0.033)\\
-DV mean & 3.303 & 3.303 & 1.862 & 1.442 & 0.789 & 2.514\\
-Observations & 523,075 & 522,957 & 522,957 & 522,957 & 522,957 & 522,957\\
-\midrule
-\textbf{Panel B: Rural} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-Wind & -0.006 & -0.034 & 0.068 & -0.102*** & -0.024 & -0.010\\
- & (0.044) & (0.040) & (0.043) & (0.035) & (0.043) & (0.036)\\
-DV mean & 3.494 & 3.494 & 2.184 & 1.311 & 1.271 & 2.223\\
-Observations & 307,381 & 307,325 & 307,325 & 307,325 & 307,325 & 307,325\\
-\midrule
-Controls & No & Yes & Yes & Yes & Yes & Yes\\
-\textbf{Fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-District & Yes & Yes & Yes & Yes & Yes & Yes\\
-State-month & Yes & Yes & Yes & Yes & Yes & Yes\\
+  & (1) & (2)\\
+\midrule
+Particulate matter & -0.287** & -0.394***\\
+(log PM 2.5) & (0.133) & (0.134)\\
+PM 2.5 times rice & -0.338*** & -0.205***\\
+ & (0.061) & (0.060)\\
+Weather (expanded) & Yes & No\\
+Weather (expanded, bins) & No & Yes\\
+\textbf{Fixed effects:} & \textbf{} & \textbf{}\\
+Village & Yes & Yes\\
+Year & Yes & Yes\\
+\midrule
+F (PM 2.5) & 1,369 & 1,168\\
+F (PM 2.5 times rice) & 2,107 & 2,201\\
+Observations & 1,266,922 & 1,266,922\\
 \bottomrule
@@
 \begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the district level.
-Control variables include female, age, age squared, and (years of) education.
-The outcome in each regression is days worked in the previous seven days.
+\item Note: Standard errors are in parentheses and are clustered at the village level.
+The outcome in all regressions estimated agricultural yield (tons per hectare).
+The rice dummy equals one in villages where rice is the most commonly harvested crop.
 * p<0.10 ** p<0.05 *** p<0.01
@@
 
-The top panel presents results for all respondents during the five-month monsoon season while the bottom panel presents results for respondents in rural areas.
-Interestingly, we do not see any signs of large, negative effects of wind exposure on labor allocation across any sectors, with the possible exception of wage employment in rural areas.
-However, the coefficient for self-employment in rural areas is of the same magnitude and \emph{positive}, but imprecisely estimated, making it hard to come to any definitive conclusions, at least in the aggregate.
-
-\hypertarget{month-specific-exposure}{%
-\subsection{Month-specific exposure}\label{month-specific-exposure}}
-
-While overall results for labor allocation are inconclusive, perhaps we can learn more by looking at temporally disaggregated results, focusing on June through October.
-I can create separate variables for exposure by each month of the season, keeping the unit of analysis the individual (for labor) or the village-season (for yields).
-
-Since the NSS data has survey dates and the both the pollution and wind data can be aggregated to the month, we can look at how exposure during different months of the agricultural season (differentially) affect labor allocation and overall agricultural productivity.
-I begin with labor allocation in columns one and two of Table \ref{tab:labor3month}.
-As before, the exposure measure is just the wind variable (over the last seven days), since I can directly match daily wind direction with the timing of the survey, but I cannot do that for pollution, which is monthly.
-The first column presents results for farm labor and the second column presents results for non-farm labor.
-
-Higher exposure leads to significantly less farm labor labor in June and July, but there are no discernible effects in other months of the year.
-There are likewise no discernible effects on non-farm labor, but some of the coefficients are rather imprecisely estimate.
-To put the effects for June and July in context and taking the coefficients at face value, an additional day of wind in the last seven days decreases farm labor in June by around 11.2 percent of the mean and in July by around 17.1 percent of the mean.
-
-\begin{table}
-
-\caption{\label{tab:labor3month}Heterogeneity by month}
-\centering
-\begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Labor} & \multicolumn{1}{c}{ } \\
-\cmidrule(l{3pt}r{3pt}){2-3}
-  & Farm & Non-farm & Yield\\
-\midrule
-June & -0.128* & 0.123 & -0.736***\\
- & (0.076) & (0.095) & (0.062)\\
-July & -0.278* & 0.185 & 0.051\\
- & (0.142) & (0.133) & (0.095)\\
-Aug. & 0.055 & -0.002 & 0.087\\
- & (0.083) & (0.068) & (0.070)\\
-Sept. & 0.086 & -0.072 & -0.140\\
- & (0.081) & (0.093) & (0.180)\\
-Oct. & 0.022 & -0.074 & 0.334\\
- & (0.071) & (0.057) & (0.291)\\
-controls & Yes & Yes & Yes\\
-\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{}\\
-district & Yes & Yes & Yes\\
-year-month & Yes & Yes & Yes\\
-\midrule
-observations & 307,325 & 307,325 & 1,266,922\\
-\bottomrule
-\end{tabular}
-\begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the village level.
-The outcome in column one is farm labor, the outcome is column two is non-farm labor, and the outcome in column three is village-level productivity.
-The independent variable for labor is wind, while the IV approach above is still used for yield.
-* p<0.10 ** p<0.05 *** p<0.01
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
-I perform a similar exercise with yield, the results of which are in the third column Table \ref{tab:labor3month}.
-I return to the IV strategy, however, to be consistent with the previous yield results.
-I include only the five months of the year that are used in the calculation of the yield estimates, or June through October.
-
-There is only one month where we see a clear effect of pollution on yield: June.
-Given that we also see negative effects of pollution on farm labor allocation in June, this is suggestive -- but not conclusive -- evidence that pollution at least partly works through labor.
-Note that this does not rule out pollution also having direct effects on agricultural land productivity, however.
-
-A final caveat about these month-specific results is that they do not take into account possible interactions between different months.
-For example, yield might respond strongly to average pollution in June and July, which would not be captured in Table \ref{tab:labor3month}.
-As one additional check, Table \ref{tab:yield8cumulative} in the appendix presents five separate instrumental variable regressions, which look at the effects of cumulative exposure to pollution separately by month.
-Across all five months, cumulative exposure negatively affects agricultural productivity, with the largest coefficient actually showing up for October, the last month of the season.
+Table \ref{tab:yieldrice} presents the main results in this section.
+I continue to employ an IV strategy, interacting pollution with an indicator for rice and instrumenting this with an interaction between wind and rice.
+The first row shows the effects of overall pollution on agricultural productivity in villages where rice is not the most commonly grown crop.
+Column two, which has the most flexible weather specification, indicates that an increase of particulate matter by one percent decreases productivity in these areas by around 0.4 percent.
+The second row is an interaction term between rice and PM 2.5, meaning the linear combination of the two coefficients gives the overall effect of pollution on productivity in rice-predominant villages.
+The coefficient of -0.2 shows that pollution decreases productivity by around 50 percent more in rice-growing villages than in non-rice growing villages.
+Since rice is more labor intensive, this is evidence that the effects of pollution work at least partly through the effects of pollution on labor productivity.
+This does not, however, rule out any direct effects on land productivity.
+
+Given the geographic concentration of crops, Table \ref{tab:yieldricestatefe} in the appendix presents the same results but with state-by-year fixed effects, instead of just year fixed effects.
+This will compare the effects of pollution within states in the same year, hopefully decreasing concerns that something other than labor intensity is the main driver of the heterogeneity documented in Table \ref{tab:yieldrice}.
+The results are consistent with those presented here, and show that pollution decreases productivity more in rice-growing villages than in other villages.
 
@@
 \begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3.5cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3.5cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
 \toprule
-  & (1) & (2) & (3)\\
-\midrule
-Particulate matter & -0.299*** & -0.107* & -0.147**\\
-(log PM 2.5) & (0.066) & (0.063) & (0.063)\\
-PM 2.5 times Coal & -0.328*** & -0.321*** & -0.304***\\
- & (0.026) & (0.024) & (0.023)\\
-Weather & No & Yes & Yes\\
-Weather (expanded) & No & No & Yes\\
-\textbf{Fixed effects:} & \textbf{} & \textbf{} & \textbf{}\\
-Year & Yes & Yes & Yes\\
-Village & Yes & Yes & Yes\\
-\midrule
-F (PM) & 766 & 776 & 788\\
-F (PM times Coal) & 199 & 222 & 232\\
-Observations & 1,266,922 & 1,266,922 & 1,266,922\\
+  & (1) & (2)\\
+\midrule
+Particulate matter & -0.228*** & -0.223***\\
+(log PM 2.5) & (0.066) & (0.065)\\
+PM 2.5 times Coal & -0.250*** & -0.220***\\
+ & (0.024) & (0.025)\\
+Weather & Yes & No\\
+Weather (expanded) & No & Yes\\
+\textbf{Fixed effects:} & \textbf{} & \textbf{}\\
+Year & Yes & Yes\\
+Village & Yes & Yes\\
+\midrule
+F (PM) & 769 & 772\\
+F (PM times Coal) & 231 & 205\\
+Observations & 1,266,922 & 1,266,922\\
 \bottomrule
@@
 
-Columns one through three of Table \ref{tab:yield7ivdiffindiff} present specifications with progressively increasing weather controls.
-Column three presents the most extensive set of controls and I focus on that column here.
-First, there are now two first stages but the F-tests indicate that the instruments are strong for both, including the interaction term (F = 232).
+Columns one and two of Table \ref{tab:yield7ivdiffindiff} present specifications with slightly different weather controls.
+Column two presents the most extensive set of controls and I focus on that column here.
+First, there are now two first stages but the F-tests indicate that the instruments are strong for both, including the interaction term (F = 205 in the second column).
 Second, the first row shows the effect of pollution on agricultural productivity prior to the opening of a coal plant; the coefficient is negative, indicating negative effects of non-coal pollution.
-Third, the interaction term is strongly negative and more than twice as large in magnitude as the level term, indicating a very clear effect specific to coal plants, in addition to the overall negative effect of pollution.
+Third, the interaction term is strongly negative and approximately the same in magnitude as the level term, indicating a very clear effect specific to coal plants, in addition to the overall negative effect of pollution.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-brunekreef2002air}{}}%
-Brunekreef, Bert, and Stephen T Holgate. 2002. {``{Air pollution and health}.''} \emph{{The Lancet}} 360 (9341): 1233--42.
+\leavevmode\vadjust pre{\hypertarget{ref-blazquez2020some}{}}%
+Blazquez, Jorge, Rolando Fuentes, and Baltasar Manzano. 2020. {``On Some Economic Principles of the Energy Transition.''} \emph{{Energy Policy}} 147: 111807.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-pope2006health}{}}%
-Pope III, C Arden, and Douglas W Dockery.
-2006.
-{``{Health effects of fine particulate air pollution: lines that connect}.''} \emph{{Journal of the Air \& Waste Management Association}} 56 (6): 709--42.
+\leavevmode\vadjust pre{\hypertarget{ref-michler2020agriculture}{}}%
+Michler, Jeffrey D. 2020. {``{Agriculture in the process of development: A micro-perspective}.''} \emph{{World Development}} 129: 104888.
+
+\leavevmode\vadjust pre{\hypertarget{ref-monfreda2008farming}{}}%
+Monfreda, Chad, Navin Ramankutty, and Jonathan A Foley.
+2008.
+{``{Farming the planet: Geographic distribution of crop areas, yields, physiological types, and net primary production in the year 2000}.''} \emph{{Global Biogeochemical Cycles}} 22 (1).
 
@@
 {``{Long-term gains from electrification in rural India}.''} \emph{{The World Bank Economic Review}} 31 (2): 385--411.
+
+\leavevmode\vadjust pre{\hypertarget{ref-vollrath2011agricultural}{}}%
+Vollrath, Dietrich. 2011. {``The Agricultural Basis of Comparative Development.''} \emph{{Journal of Economic Growth}} 16: 343--70.
 
@@
 weather & TerraClimate & global & monthly\\
-labor & National Sample Survey (NSS) & India & daily interviews\\
+crops & Monfreda et al. (2008) & global & 2000 (year)\\
 \bottomrule
@@
 
-\caption{\label{tab:pollutiontable}Wind direction and particulate matter}
+\caption{\label{tab:yield3ivmainfirststage}Wind direction and particulate matter}
 \centering
@@
 \toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{1998-2015} & \multicolumn{2}{c}{2002-2013} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
   & (1) & (2) & (3) & (4)\\
 \midrule
-wind (10s) & 0.0023*** & 0.0016*** & 0.0057*** & 0.0015***\\
- & (0.0005) & (0.0002) & (0.0005) & (0.0002)\\
+wind & 0.0083*** & 0.0085*** & 0.0083*** & 0.0085***\\
+ & (0.0002) & (0.0002) & (0.0002) & (0.0002)\\
+weather & No & Yes & Yes & No\\
+weather (expanded) & No & No & Yes & No\\
+weather (expanded, bins) & No & No & No & Yes\\
 \textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
 village & Yes & Yes & Yes & Yes\\
-month & Yes & No & Yes & No\\
-district-month & No & Yes & No & Yes\\
-\midrule
-observations & 24,376,458 & 24,376,458 & 16,250,973 & 16,250,973\\
+year & Yes & Yes & Yes & Yes\\
+\midrule
+F & 1,236 & 1,266 & 1,205 & 1,246\\
+observations & 1,266,922 & 1,266,922 & 1,266,922 & 1,266,922\\
 \bottomrule
@@
 \begin{figure}
-\includegraphics{draft_files/figure-latex/randomplotleads-1} \caption[Randomization inference for leads]{Randomization inference for leads}\label{fig:randomplotleads}\floatfoot*{Note: The figure presents the distribution of coefficients by randomly assigning a single wind value to each village, sampled with replacement across all villages in a given district in the same year.
-"Current" refers to the coefficient on current pollution, while "lead" refers to the coefficient on next year's pollution in the same table.}
+\includegraphics{draft_files/figure-latex/randomplot-1} \caption[Randomization inference]{Randomization inference}\label{fig:randomplot}\floatfoot*{Note: The figure presents the distribution of coefficients by randomly assigning wind direction across villages within the same year.
+The "all" distribution is created by randomly assigning a single wind value to each village, sampled with replacement across all villages in a given year.
+The "within district" randomly samples with replacement only from within the same district in the same year.}
 \end{figure}
@@
 \newpage
-
-\begin{figure}
-\includegraphics{draft_files/figure-latex/laborplot-1} \caption[Timing of household surveys for the National Sample Survey]{Timing of household surveys for the National Sample Survey}\label{fig:laborplot}\floatfoot*{Note: The histogram shows the distribution of survey dates for the four waves of the National Sample Survey (NSS) for the five waves used in this paper.}
-\end{figure}
-
-\FloatBarrier
-\newpage
-
-\begin{figure}
-\includegraphics{draft_files/figure-latex/monthplot-1} \caption[Average farm days by season, rural only]{Average farm days by season, rural only}\label{fig:monthplot}\floatfoot*{Note: The line graph shows the mean number of farm days worked per week based on the month of the relevant agricultural season.
-The monsoon season is defined as June (month 1) to October (month 5) while the rainy season is defined as November (month 1) to March (month 5) of the following year.
-The data is from the NSS.}
-\end{figure}
 
@@
 
+\FloatBarrier
+\newpage
+
+\begin{table}
+
+\caption{\label{tab:yieldricestatefe}Agricultural productivity by most common crop}
+\centering
+\begin{threeparttable}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3.5cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\toprule
+  & (1) & (2)\\
+\midrule
+Particulate matter & -0.811*** & -1.034***\\
+(log PM 2.5) & (0.158) & (0.144)\\
+PM 2.5 times rice & -1.204*** & -0.966***\\
+ & (0.095) & (0.092)\\
+Weather (expanded) & Yes & No\\
+Weather (expanded, bins) & No & Yes\\
+\textbf{Fixed effects:} & \textbf{} & \textbf{}\\
+Village & Yes & Yes\\
+Year & Yes & Yes\\
+\midrule
+F (PM 2.5) & 883 & 941\\
+F (PM 2.5 times rice) & 1,355 & 1,395\\
+Observations & 1,266,922 & 1,266,922\\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}[para]
+\item Note: Standard errors are in parentheses and are clustered at the village level.
+The outcome in all regressions estimated agricultural yield (tons per hectare).
+The rice dummy equals one in villages where rice is the most commonly harvested crop.
+* p<0.10 ** p<0.05 *** p<0.01
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+
 \end{document}

### 2023-07-10 analyses/1.openings.R

@@ -12,7 +12,7 @@ library(sf)
 
 
 # Sets wd to folder this script is in so we can create relative paths instead of absolute paths
-setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
+# setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
 # We want to get the wd up to the main folder
 # Need to go up two levels
 setwd("../../")
@@ -21,24 +21,37 @@ getwd()    # check
 
 
 
+villages <- read_sf(paste0("data/spatial/village_new/village_new.shp"))
+villages <- as_tibble(villages) %>% mutate(shrid = paste0(pc11_s_id, "-", pc11_tv_id),
+                                district_id = paste0(pc11_s_id, "-", pc11_d_id),
+                                area_overlap = area) %>%
+                          dplyr::select(shrid, district_id)
+
+
 # Where do plants open?
 villages90 <- read_csv("data/clean/census_plants/villages90.csv")
+villages90$pc91_pca_tot_p_u[is.na(villages90$pc91_pca_tot_p_u)] <- 0
 villages90 <- villages90 %>% mutate(
                                     literacy = pc91_pca_p_lit/pc91_pca_tot_p,
                                     pop = pc91_pca_tot_p,
                                     scst_prop = (pc91_pca_p_sc + pc91_pca_p_st)/pop,
-                                    tar_road = pc91_vd_tar_road
-                                    )
+                                    tar_road = pc91_vd_tar_road,
+                                    urban_pop = pc91_pca_tot_p_u/pc91_pca_tot_p
+                                    ) %>%
+                              left_join(villages, by = c("shrid"))
 villages00 <- read_csv("data/clean/census_plants/villages00.csv")
+villages00$pc01_pca_tot_p_u[is.na(villages00$pc01_pca_tot_p_u)] <- 0
 villages00 <- villages00 %>% group_by(shrid) %>%
                              mutate(
                                     literacy = sum(pc01_pca_p_lit)/sum(pc01_pca_tot_p),
                                     pop = sum(pc01_pca_tot_p),
                                     scst_prop = (sum(pc01_pca_p_sc) + sum(pc01_pca_p_st))/pop,
-                                    tar_road = max(pc01_vd_tar_road)
+                                    tar_road = max(pc01_vd_tar_road),
[…]

### 2023-07-10 analyses/3.yield.R

@@ -14,10 +14,10 @@ library(parallel)
 
 
 # Sets wd to folder this script is in so we can create relative paths instead of absolute paths
-setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
+# setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
 # We want to get the wd up to the main folder
 # Need to go up two levels
-setwd("../../")
+setwd("../")
 # Double check
 getwd()    # check
 
@@ -30,10 +30,11 @@ pollution <- read_csv(paste0("data/clean/pm25/ag_merge.csv"))
 df <- df %>% 
         left_join(pollution, by = c("shrid", "year", "season"))
 rm(pollution)
-df <- df %>% group_by(shrid, season) %>% 
+df <- df %>% group_by(shrid, season) %>%
               mutate(
                      yield = mean,
-                     rain_z = (log(rain_total) - mean(log(rain_total))/sd(log(rain_total))),
+                     rain_z = (log(rain_total + 1) - mean(log(rain_total + 1)))/sd(log(rain_total + 1)),
+                     rain_z2 = ((rain_total) - mean((rain_total)))/sd((rain_total)),
                      wind_abs_dev = abs(days_sums - mean(days_sums)),
                      pm25_abs_dev = abs(pm25 - mean(pm25)),
                      wind = days_sums/10,
@@ -57,11 +58,21 @@ df <- df %>% group_by(shrid, season) %>%
                      temp_m3 = (tmax_m3 + tmin_m3)/20,
                      temp_m4 = (tmax_m4 + tmin_m4)/20,
                      temp_m5 = (tmax_m5 + tmin_m5)/20,
-                     rain1_z = (log(rain_m1) - mean(log(rain_m1))/sd(log(rain_m1))),
-                     rain2_z = (log(rain_m2 + rain_m1) - mean(log(rain_m2 + rain_m1))/sd(log(rain_m2 + rain_m1))),
-                     rain3_z = (log(rain_m3 + rain_m1 + rain_m2) - mean(log(rain_m3 + rain_m1 + rain_m2))/sd(log(rain_m3 + rain_m1 + rain_m2))),
-                     rain4_z = (log(rain_m4 + rain_m1 + rain_m2 + rain_m3) - mean(log(rain_m4 + rain_m1 + rain_m2 + rain_m3))/sd(log(rain_m4 + rain_m1 + rain_m2 + rain_m3))),
[…]

### 2023-07-10 analyses/4.nss.R

@@ -10,13 +10,14 @@ library(viridis)
 library(fixest)
 library(sf)
 library(tree)
+library(marginaleffects)
 
 
 # Sets wd to folder this script is in so we can create relative paths instead of absolute paths
-setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
+#setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
 # We want to get the wd up to the main folder
 # Need to go up two levels
-setwd("../../")
+setwd("../")
 # Double check
 getwd()    # check
 
@@ -36,61 +37,72 @@ df <- df %>% mutate(
                     )
 
 # MONSOON ONLY
-df <- df %>% filter(month_int %in% c(5,6,7,8,9,10,11))
+df <- df %>% filter(month_int %in% c(6,7,8,9,10))
+
+df$precip_sum <- df$month6_precip
+df$precip_sum[month(df$date)==7] <- df$month6_precip[month(df$date)==7] + df$month7_precip[month(df$date)==7]
+df$precip_sum[month(df$date)==8] <- df$month6_precip[month(df$date)==8] + df$month7_precip[month(df$date)==8] + df$month8_precip[month(df$date)==8]
+df$precip_sum[month(df$date)==9] <- df$month6_precip[month(df$date)==9] + df$month7_precip[month(df$date)==9] + df$month8_precip[month(df$date)==9] + df$month9_precip[month(df$date)==9]
+df$precip_sum[month(df$date)==10] <- df$month6_precip[month(df$date)==10] + df$month7_precip[month(df$date)==10] + df$month8_precip[month(df$date)==10] + df$month9_precip[month(df$date)==10] + df$month_precip[month(df$date)==10]
+
 # create seasons means
 df <- df %>% 
       group_by(year, district) %>%
       mutate(temp_mean = mean((month_tmin + month_tmax)/2),
-             precip_sum = sum(month_precip)) %>%
+             precip_sum_season = mean(month_precip)) %>%
       ungroup()
 
+# add mean of rainfall
+means <- read_csv(paste0("data/clean/terra_district/precip_mean_monthALL.csv"))
+means <- means %>%
+          rename(state_merge = ST_CEN_CD, district_merge = DT_CEN_CD) %>%
+          mutate(state_merge = as.numeric(state_merge),
+                 district_merge = as.numeric(district_merge))
+df <- df %>% 
[…]

### 2023-07-10 cleaning/2.cleaning_ntl.R

@@ -26,7 +26,7 @@ getwd()    # check
 
 
 ### Load raw data ----------------------------------------------------------------------------------------------------------------------------------------------
-ntl <- as_tibble(read.csv("daa/raw/shrug-v1.5.samosa-nl-csv/shrug_nl_wide.csv"))
+ntl <- as_tibble(read.csv("data/raw/shrug-v1.5.samosa-nl-csv/shrug_nl_wide.csv"))
 ntl <- ntl %>% select(-num_cells)
 
 # Gather, one variable at a time (we want long by year but wide by variables)

### 2023-07-10 cleaning/4.cleaning_plants.R

@@ -49,6 +49,7 @@ plants$year_retired <- as.numeric(plants$year_retired)
 
 # Add villages
 villages <- read_sf("data/spatial/shapefiles/village.shp")
+villages$shrid <- paste(villages$pc11_s_id, villages$pc11_tv_id, sep = "-")
 
 # Also want to transform both to a better crs for calculating distances
 plants <- st_transform(plants, "EPSG:24378")
@@ -72,7 +73,7 @@ centroids_villages$pc11_d_id <- NA
 centroids_villages$pc11_sd_id <- NA
 centroids_villages$pc11_tv_id <- NA
 centroids_villages$tv_name <- NA
-centroids_villages[,2] <- as_tibble(villages) %>% dplyr::select(shrid)
+centroids_villages$shrid <- villages$shrid
 
 # Now let's figure out which are within the 30km buffer
 centroids_villages_30km <- st_intersects(centroids_villages, buffer30km)
@@ -98,7 +99,6 @@ st_write(centroids_villages_100km, "data/spatial/centroids_villages_100km",
 
 
 ####### Loaded villages and plants into QGIS to do distances (much faster than doing it here)
-# Load distance matrix (takes a while; this csv is >4GB)
 dist_matrix <- read.csv(paste0("data/clean/wind_ntl/distances.csv"))
 
 # Only keep the rows where distance is within 30km
@@ -106,7 +106,7 @@ dist_matrix <- as_tibble(dist_matrix[dist_matrix$Distance<=30000,])
 colnames(dist_matrix) <- c("shrid", "plant_id", "distance")
 
 # Merge in coordinates
-centroids_villages_30km <- centroids_villages_30km %>% select(shrid)
+centroids_villages_30km <- centroids_villages_30km %>% dplyr::select(shrid)
 centroids_villages_30km$x_vil <- NA
 centroids_villages_30km$y_vil <- NA
 for (row in 1:nrow(centroids_villages_30km)){
@@ -124,6 +124,10 @@ for (row in 1:nrow(plants)){
 }
 # And join
 merging <- as_tibble(centroids_villages_30km) %>% dplyr::select(shrid, x_vil, y_vil)
+merging <- merging %>%
+            group_by(shrid) %>%
+            filter(row_number()==1) %>%
+            ungroup()
 dist_matrix <- dist_matrix %>% left_join(
                                          merging,
[…]

### 2023-07-10 cleaning/7.cleaning_ag_productivity_village.R

@@ -1,4 +1,5 @@
 # Purpose: This script cleans data for estimates of agricultural productivity
+# data from here: https://springernature.figshare.com/collections/A_new_two-decade_2001-2019_high-resolution_agricultural_primary_productivity_dataset_for_India/6079104
 # Author: Josh Merfeld
 # Date: December 12th, 2022
 
@@ -27,8 +28,16 @@ getwd()    # check
 villages_centroids <- read_sf("data/spatial/centroids_villages_30km")
 villages_centroids <- as_tibble(villages_centroids) %>% dplyr::select(shrid)
 villages_centroids$within30 <- 1
+villages_centroids <- villages_centroids %>%
+                          group_by(shrid) %>%
+                          filter(row_number()==1) %>%
+                          ungroup()
 villages <- read_sf("data/spatial/shapefiles/village.shp")
 villages$shrid <- paste0(villages$pc11_s_id, "-", villages$pc11_tv_id)
+villages <- villages %>%
+              group_by(shrid) %>%
+              filter(row_number()==1) %>%
+              ungroup()
 villages <- villages %>% left_join(villages_centroids, by = "shrid")
 villages <- villages %>% filter(within30==1) %>% dplyr::select(shrid)
 rm(villages_centroids)
@@ -329,6 +338,43 @@ for (year in 2002:2013){
 write_csv(df, paste0("data/clean/ag_productivity/all_combined.csv"))
 
 
+
+# harvested area
+crops_all <- as_tibble(villages) %>% dplyr::select("shrid")
+crops <- c("rice", "wheat", "maize", "sugarcane", "cotton", "sorghum", "millet", "groundnut", "bean")
+c <- crops[1]
+temp <- raster(paste0("data/raw/croparea/GeoTiff/", c, "/", c, "_HarvestedAreaFraction.tif"))
+temp <- exact_extract(temp, villages, fun = "mean", append_cols = "shrid")
+crops_all$most_common <- paste0(c)
+crops_all$most_common_area <- temp$mean
+crops_all <- crops_all %>%
+              left_join(temp, by = c("shrid"))
+colnames(crops_all)[ncol(crops_all)] <- paste0("area_", c)
+for (c in crops[2:length(crops)]){
+  temp <- raster(paste0("data/raw/croparea/GeoTiff/", c, "/", c, "_HarvestedAreaFraction.tif"))
[…]

### 2023-07-10 cleaning/8.cleaning_overlap_merging_nss.R

@@ -12,10 +12,10 @@ library(sf)
 
 
 # Sets wd to folder this script is in so we can create relative paths instead of absolute paths
-setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
+# setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
 # We want to get the wd up to the main folder
 # Need to go up two levels
-setwd("../../")
+setwd("../")
 # Double check
 getwd()    # check
 
@@ -73,6 +73,29 @@ df_labor <- df_labor %>% mutate(state_merge = as.numeric(state_merge),
 df_labor <- df_labor %>% filter(is.na(merge)==F)
 
 
+# now get averages for each district
+all <- read_csv(paste0("data/clean/terra_district/precip1990.csv"))
+all <- all[,1:2]
+for (month in 6:11){
+       temp <- read_csv(paste0("data/clean/terra_district/precip1990.csv"))
+       temp <- temp[,c(1:2, 2+month)]
+       for (year in 1991:2003){
+              temp2 <- read_csv(paste0("data/clean/terra_district/precip", year, ".csv"))
+              temp2$sum <- temp2[,c(8:(2+month))]
+              temp2 <- temp2[,c(1:2, ncol(temp2))]
+              temp <- temp %>% left_join(temp2, by = c("ST_CEN_CD", "DT_CEN_CD"))
+       }
+       temp[[paste0("precip_mean_month", month)]] <- apply(temp[,3:ncol(temp)], 1, mean)
+       temp[[paste0("precip_sd_month", month)]] <- apply(temp[,3:ncol(temp)], 1, sd)
+
+       temp <- temp %>%
+                dplyr::select(c("ST_CEN_CD", "DT_CEN_CD", starts_with(c("precip_"))))
+
+       all <- all %>% left_join(temp, by = c("ST_CEN_CD", "DT_CEN_CD"))
+}
+write_csv(all, paste0("data/clean/terra_district/precip_mean_monthALL.csv"))
+
+
 # district pollution
 pollution <- as_tibble(read_csv("data/clean/pm25_district/all_combined_district.csv"))
 
@@ -111,13 +134,13 @@ for (day in 1:length(date_vec)){
                   group_by(state, district) %>%
                   mutate(days_sum = weighted.mean(days_sum, area_weight, na.rm = T),
                          month_sum = weighted.mean(month_sum, area_weight, na.rm = T),
-                         #pm25_district = weighted.mean(pm25, area_weight, na.rm = T),
[…]

### 2023-07-13 draft/draft.Rmd
new draft

@@
 ---
-title: "Air pollution and agricultural productivity in a developing country^[I thank Ariel Ortiz-Bobea, Marc Bellemare, and David Sungho Park for comments and suggestions that greatly improved the quality of the paper.
+title: "Air pollution and agricultural productivity in a developing country^[I thank Ariel Ortiz-Bobea, Marc Bellemare, and David Sungho Park for comments and suggestions that greatly improved the quality of the paper, as well as conference attendees at the 2023 KER International Conference.
 The usual disclaimer applies.
@@
 
-While there are likely positive effects from this increase in power generation -- for example, electricity may lead to economic and productivity growth [@dinkelman2011effects; @kline2014local; @rud2012electricity; @van2017long], though electrification alone may not be a sufficient condition [@lee2020does] -- these improvements do not come without downsides, especially when it comes to power derived from coal.
-Pollutants from coal contribute to climate change and the phase out of coal energy is therefore a major climate goal [@iea2022].
+While there may be positive effects from this increase in power generation -- for example, electricity may lead to economic and productivity growth [@dinkelman2011effects; @kline2014local; @rud2012electricity; @van2017long], though electrification alone may not be a sufficient condition [@lee2020does] -- these improvements do not come without downsides, especially when it comes to power derived from coal.
+Pollutants from coal contribute to climate change and the phase out of coal energy is, therefore, a major climate goal [@iea2022].
 Emissions from the production and burning of coal are also directly harmful to human health and the environment; pollutants like sulfur dioxide, nitrogen oxides, methane, and mercury are particularly harmful.[^4] While the use of coal had been declining, recent events -- including Russia's invasion of Ukraine -- have led to an increase in the demand for coal.[^5]
@@
 These effects can work through the well-known effects on human health and productivity, but also directly through land productivity.
-Black carbon in soot, for example, can directly affect solar radiation reaching plants on the surface [@ramanathan2008global; @burney2014recent].
+Black carbon in soot, for example, can directly affect solar radiation reaching plants on the surface [@ramanathan2008global; @burney2014recent], while, on the other hand, sulfur in some pollution sources can actually be productivity enhancing [@sanders2022adaptation].
 Relative to previous work in economics, this paper tackles pollution more broadly, before turning to a specific source of pollution: coal plants.
@@
 
-The main goal of this paper is to examine whether exposure to pollution and emissions from coal plants, in particular, affect agricultural productivity.
+The main goal of this paper is to examine whether exposure to pollution affects agricultural productivity, as well as to provide evidence about the effects of pollution from coal plants, in particular.
 I use several sources of data, which I describe here and list in Table \@ref(tab:data) of the appendix.
@@
 While there is a wealth of literature on measurement error in land [@desiere2018land; @abay2022nonclassical], the authors use a static cropland mask for the productivity estimates.
-This means that any measurement error will be constant across time and subsumed by the village fixed effects.
+This means that any measurement error will hopefully be constant across time and subsumed by the village fixed effects.
 While the measure of output is also noisy, the use of satellite-derived data and a static cropland mask reduces concerns about measurement error being responsible for the results below, especially since common sources of measurement error are unlikely to be correlated with wind direction.
 The authors define the monsoon season as June to October and the winter season as November to March.
-I focus on only the monsoon season and keep the authors' definition when matching across data below.
-This data is also publicly available from Nature's data-sharing website.[^7] Since the main regressions below use village fixed effects, any differences in cropping patterns across areas should not bias estimates on changes, as long as these patterns do not change markedly in response to pollution.
+Due to large differences in the crops grown and the production technologies used across seasons, I focus on only the monsoon season and keep the authors' definition when matching across data below.
+The data on agricultural productivity is publicly available from Nature's data-sharing website.[^7] Since the main regressions below use village fixed effects, any differences in cropping patterns across areas should not bias estimates on changes, as long as these patterns do not change markedly in response to annual pollution levels.
 
@@
 
-To measure exposure to pollution from coal plants, I first locate all village centroids located within 30km from a coal plant in a given year.
+To measure exposure to pollution from point sources around coal plants, I first locate all village centroids located within 30km from a coal plant in a given year.
 I choose 30km due to previous research on the effects of (air) pollution [@aragon2016polluting; @bencko1977health; @li2014health].
-I then calculate the direction from coal plants to all village centroids within that 30km radius.
-To define exposure, I then pull daily wind direction data from the National Center for Atmospheric Research.[^9] For each day, I document whether the wind is blowing towards each village centroid[^10] and then temporally aggregate this daily data depending on the temporal definition of the corresponding outcome.
-For example, agricultural productivity is defined across five months (e.g.
+I then calculate the direction from these point sources to all village centroids within that 30km radius.
+To define exposure, I pull daily wind direction data from the National Center for Atmospheric Research.[^9] For each day, I document whether the wind is blowing towards each village centroid[^10] and then temporally aggregate this daily data depending on the temporal definition of the corresponding outcome.
+For example, agricultural productivity is defined across five months (i.e.
 the monsoon season is from June to October) so I count the total days a given village is exposed to wind blowing from *any* coal plant within 30km during those five months.
@@
 
-[^10]: I define "towards" as within five degrees to help take into account that I am using village centroids, using the x and y components of wind speed to calculate angles.
+[^10]: I define "towards" as within five degrees to help take into account that I am using village centroids.
+The wind data includes both the x and y components of wind speed, which I can use to calculate angles from point sources to village centroids.
 
@@
 This means that the overall exposure for a given area to coal plant emissions changes across time.
-I also pull data on particulate matter to help validate this proxy for exposure.
-This data comes from @hammer2020global and I also aggregate this to the village using the same method detailed above.
+
+For pollution data, I use estimated particulate matter (PM).
+Specifically, I use estimates of PM 2.5, which includes particles with a diameter of 2.5 micrometers or less.
+This data comes from @hammer2020global and is available as a raster with a resolution of 0.01 degrees.
+I aggregate this data to the village level by extracting the mean PM for each feature in the shapefile.
+I do this separately for each month from June 2002 to October 2013.
+The bottom panel in Figure \@ref(fig:windexample) in the appendix is an example of the PM data.
+The figure shows the estimated PM 2.5 from for each village in Guna district of Madhya Pradesh on January 1st, 2010.
+
+While I use PM 2.5 as the estimated level of pollution, it is important to clarify that I cannot explicitly say that PM 2.5, per se, is responsible for any effects.
+Pollution includes a variety of different components, including PM 2.5, PM 10, sulfur dioxide, and nitrogen dioxide, to name but a few.
+While I use PM 2.5 as a proxy for pollution, it is possible that other components are responsible for the effects.
+Additionally, the PM 2.5 data is estimated using satellite data and a machine learning algorithm.
+While this is a common approach in the literature ( see @hammer2020global for more details), it is important to note that the estimates are not direct measurements of pollution.
+However, a valid instrumental variable can help for any measurement error present in the data [@glewwe2012much; @gillen2019experimenting].
 
 To help understand some of the mechanisms driving the relationship between exposure to emissions and agricultural productivity, I also look at the differential effects by the labor intensity of the predominant crop in each area.
-For this, I use data from [@monfreda2008farming], who estimate area harvested for different crops across the global around the year 2000.
+For this, I use data from @monfreda2008farming, who estimate area harvested for different crops across the global around the year 2000.
 I identify areas where rice (paddy) is the most commonly grown crop, under the assumption that rice is more labor intensive than other crops [@vollrath2011agricultural; @michler2020agriculture].
-@michler2020agriculture, for example, shows that rice, on avearge, uses almost 1,800 hours per hectare, while the second-highest crop in his data, cotton, juses just 850.^[He estimates that wheat, another common crop in India, uses just 600 hours per hectare.]
+@michler2020agriculture, for example, shows that rice, on avearge, uses almost 1,800 hours per hectare, while the second-highest crop in his data, cotton, juses just 850.
+He also estimates that wheat, another common crop in India, uses just 600 hours per hectare.
 
@@
 colnames(plantresultstable) <- c("1991", "2001", "2001", "2011")
-rownames(plantresultstable) <- c("Population (log)", "", "Area (log)", "", "Urban pop (prop)", "", "Ag. productivity", "", "Pollution", "(log PM 2.5)",
+rownames(plantresultstable) <- c("Population (log)", "", "Area (log)", "", "Urban pop (prop)", "", "Ag. productivity", "", "Pollution (log PM 2.5)", "",
                                 "Sub-sample", "Observations")
@@
            ) %>%
-  column_spec(1, width = "3cm") %>%
+  column_spec(1, width = "3.5cm") %>%
   column_spec(c(2:5),width = "2cm") %>%
@@
 
-Given the likely endogeneity previewed by these summary statistics, I now turn to a plausibly exogenous source of variation in pollution: wind direction.
-While the location of a coal plant might be endogenous, the amount of pollution received by a specific village in a given year can vary randomly based on wind direction.
+Since pollution is likely to be endogenous, a naïve OLS regression will not return unbiased estimates.
+The experimental ideal would be to randomly assign pollution to villages, but this is obviously not possible.
+Instead, I use an instrumental variables (IV) approach to identify the causal effects of pollution on agricultural productivity.
+The key identifying assumption is that wind direction is exogenous from year to year, conditional on village fixed effects and weather.
+While the location of pollution sources might be endogenous, the amount of pollution received by a specific village in a given year can vary randomly based on wind direction.
 By counting the number of days in a given year that a village is downwind from a polluted area and conditioning on village fixed effects and weather (rainfall and temperature), variation across years is plausibly exogenous.
+Since I include village fixed effects, identification comes from within-village changes in wind direction across years.
 
 Consider the reduced form effect of wind exposure on agricultural productivity:
-\begin{gather} y_{it} = \alpha_{i} + \gamma_{t} + \beta wind_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it}, \end{gather}
+\begin{gather} PM2.5_{it} = \alpha_{i} + \gamma_{t} + \beta wind_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it}, \end{gather}
 
@@
 The differences in weather patterns is very context specific and can differ even for villages located near one another.
-Nonetheless, adding rainfall and temperature should alleviate some of these concerns.
+
+I believe there are two reasons this is unlikely to be the case here.
+First, villages located near one another have very similar weather patterns, but wind direction leads to large differences in pollution exposure even for villages right next to one another.
+Second, in some specifications, I include very flexible specifications for weather.
+These inclusions do not lead to very large changes in the estimated relationship between pollution and agricultural productivity, suggesting that weather is unlikely to be responsible for a spurious correlation between productivity and pollution in the main results.
 
@@
 The outcome variable is particulate matter -- specifically, $\mathrm{PM 2.5}$, which is particulate matter no larger than 2.5 micrometers in diameter -- which comes from @hammer2020global.
-$\mathrm{PM 2.5}$ is one of the harmful byproducts from coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.[^12] This data is available at the monthly level, but I aggregate variables to the season variable since that is the level of analysis in the main results.
+$\mathrm{PM 2.5}$ is one of the harmful byproducts from many industrial processes, including power generation.
+Particulate matter is a major component of the pollution emitted by coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.[^12] This data is available at the monthly level, but I aggregate variables to the season variable since that is the level of analysis in the main results.
 
@@
 Since the exposure variable is at the village level and we follow villages over time, the standard errors are also clustered at the village level.
-In columns two and four, I switch the month fixed effects for district-month fixed effects, which considers only villages in the same district in the same month.
-
-In addition to serving as an example, Table \@ref(tab:yield3ivmainfirststage) also helps validate the use of wind direction as a proxy for exposure to pollution.
-The first two columns include all available years of data while the last two columns restrict estimation only to the years for which the agricultural productivity regressions are estimated.
-Across all four columns, the story is the same: when the wind is blowing in the direction of a village from a coal plant within 30 kilometers, estimated $\mathrm{PM 2.5}$ is substantially higher.
+I add more and more flexible specifications for weather, with the most flexible specification in the fourth column.
+Note that the relationship between wind direction and particulate matter is very similar across all four columns, suggesting that the inclusion of weather does not change the relationship between wind direction and pollution.
+This is important because it suggests that the use of wind direction as a proxy for pollution exposure is not biased by weather patterns.
 
@@
 
-To identify possible mechanisms, I identify areas that grow predominantly rice, which is a labor-intensive crop. Specifically, we are interested in the following regression:
+A key question is how pollution affects agricultural productivity.
+Specifically, does pollution directly affect labor productivity, land productivity, or both?
+Unfortunately, there is a lack of data that would allow me to directly test these mechanisms, since I do not have data on the amount of labor or land used in each village.
+Instead, I look at heterogeneity in the effects of pollution based on the predominant crop grown in each village.
+
+@michler2020agriculture presents estimates of labor intensity for several major crops in India.
+The most labor-intensive crop is rice, which uses almost 1,800 hours per hectare.
+The second-most labor-intensive crop is cotton, which uses just 850 hours per hectare.
+Wheat, another common crop in India, uses just 600 hours per hectare.
+This suggests that pollution should be more harmful for rice than for other crops if there are direct impacts on labor productivity.
+
+To identify this possible mechanism, I identify areas that grow predominantly rice. Specifically, I estimate the following regression:
 \begin{gather} 
@@
 
+While this is a useful first step, it is not a perfect identification strategy.
+First, there could be other differences in the types of areas that grow rice, and these differences could actually be responsible for any heterogeneity.
+Second, this strategy does not allow me to say anything about the effects of pollution on land productivity.
+
 ## Identifying effects of pollution from coal plants
 
-I use the location of coal plants to identify locations of high pollution, but there can be other pollution sources located near coal plants, as well.
-I nonetheless want to identify the effects of pollution coming specifically from coal plants.
+I use the location of coal plants to identify locations of high pollution emission, but there can be other pollution sources located near coal plants, as well.
+I nonetheless also want to identify the effects of pollution coming specifically from coal plants, in addition to the overall effects of pollution in the other parts of the paper.
 I do this by combining the two-stage procedure above with a differences-in-differences framework.
@@
 I first present the reduced form results for agricultural productivity.
-The outcome of all regressions in this section is the log of agricultural land productivity, which is defined as tons per hectare.
+The outcome for all regressions in this section is the log of agricultural land productivity, which is defined as tons per hectare.
 The first set of results are in Table \@ref(tab:yieldtable).
@@
 The first column presents the most simple results, with only the wind exposure variable and the fixed effects.
-The second column adds a rainfall and average temperature variable, while the third column includes higher-order weather variables, as well.
-While the coefficient decreases slightly in magnitude as we move from the first to the second column and the second to the third column, the decrease is relatively small.
-There is a consistent, negative effect of wind exposure on agricultural productivity.
+The second column adds a rainfall and average temperature variable.
+While I leave them out of the table for simplicity, the coefficients are in the expected direction: rainfall is positive and average temperature is negative.
+The third column adds higher-order weather variables for both rainfall and temperature, as well as the interaction between them.
+This attenuates the relationship between wind and productivity somewhat, but the coefficient remains negative and statistically significant.
+
+The final column includes the most flexible form for the weather variables.
+The z-score for rainfall is turned into bins -- with cutoffs at $z=-1$ and $z=1$ -- based on rainfall in each month of the season.
+In other words, over the five months, I create 15 different indicator variables for rainfall based on the z-score in each month.
+One of these indicators drops out for each month of the year, leaving 10.
+I also include interactions between all of these and temperature, as well as between the bins and temperature squared.
+The coefficient on wind remains negative and statistically significant, and is actually slightly larger in magnitude than the results in column three, which is a much less flexible specification for weather.
+This leads me to conclude that the results are not being driven by weather patterns.
 
@@
 # Table
-yield1reducedform <- yield1reducedform[,-c(1)]
-colnames(yield1reducedform) <- c("(1)", "(2)", "(3)")
-
-rownames(yield1reducedform) <- c("Wind days", " ", "Rain (z)", "", "Mean temperature (10s)", "",
-                                  "Rain squared", " ", "Temperature squared", " ", "Rain times Temperature", " ",
-                                  "Fixed effects:", "Village", "Year",
-                                  "Observations")
+colnames(yield1reducedform) <- c("(1)", "(2)", "(3)", "(4)")
+
+rownames(yield1reducedform) <- c("wind days", " ", 
+                                  "weather", "weather (expanded)", "weather (expanded, bins)",
+                                  "fixed effects:", "village", "year",
+                                  "observations")
 kable(
       yield1reducedform, caption = "Wind direction and agricultural productivity",
-      align = "ccc", booktabs = TRUE, linesep = ""
+      align = "cccc", booktabs = TRUE, linesep = ""
       ) %>%
@@
   column_spec(c(2:4),width = "2.5cm") %>%
-  row_spec(13, bold = TRUE) %>%
-  row_spec(15, hline_after = TRUE) %>%
+  row_spec(6, bold = TRUE) %>%
+  row_spec(8, hline_after = TRUE) %>%
   #add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
@@
 The mean within-village absolute deviation in wind exposure is approximately 2.42 days, meaning that the average change in agricultural productivity from year-to-year due to nothing but changes in wind patterns carrying particulate matter is around 1.5 percent relative to the mean.
-This implies that swings in agricultural productivity -- due to wind direction -- of several percent are probably quite common, since the absolute deviation includes deviations below and above the mean.
+This implies that swings in agricultural productivity -- due to wind direction -- of several percent are probably quite common, since the absolute deviation includes deviations above and below the mean.
 
@@
 # Table
-yield2naive <- yield2naive[,-1]
-colnames(yield2naive) <- c("(1)", "(2)", "(3)")
-rownames(yield2naive) <- c("Particulate matter", "(log PM 2.5)", 
-                           "Weather", "Weather (expanded)",
+colnames(yield2naive) <- c("(1)", "(2)", "(3)", "(4)")
+rownames(yield2naive) <- c("Particulate matter (log PM 2.5)", "", 
+                           "Weather", "Weather (expanded)", "weather (expanded, bins)",
                            "Fixed effects:", "Village", "Year",
@@
       yield2naive, caption = "Particulate matter (PM 2.5) and agricultural productivity",
-      align = "ccc", booktabs = TRUE, linesep = ""
+      align = "cccc", booktabs = TRUE, linesep = ""
       ) %>%
@@
            ) %>%
-  column_spec(1, width = "4cm") %>%
-  column_spec(c(2:4),width = "2.5cm") %>%
-  row_spec(5, bold = TRUE) %>%
-  row_spec(7, hline_after = TRUE) %>%
-  #add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
+  column_spec(1, width = "5cm") %>%
+  column_spec(c(2:4),width = "2cm") %>%
+  row_spec(6, bold = TRUE) %>%
+  row_spec(8, hline_after = TRUE) %>%
   kable_classic_2()
@@
 There is again a consistent, negative relationship between particulate matter and agricultural productivity.
-Since both variables are logged, the coefficients are similar to elasticities, indicating that a one-percent increase in pollution leads to an approximately 0.04 percent decrease in agricultural productivity.
+Since both variables are logged, the coefficients are similar to elasticities, indicating that a one-percent increase in pollution leads to an approximately 0.04 percent decrease in agricultural productivity in three of the columns.
 
 Recall, however, the results from above: coal plants were placed in places with higher average agricultural productivity.
-This means that the simple regressions in Table \@ref(tab:yieldtablepm) might actually be biased upwards: although pollution might negatively affect productivity, pollution is higher in more productive areas due to the endogenous placement of coal plants in the study areas.
+This means that the simple regressions in Table \@ref(tab:yieldtablepm) might actually be biased upwards: although pollution might negatively affect productivity, pollution is higher in more productive areas due to the endogenous placement of pollution sources in the study areas.
 One possible way to recover the actual effect of pollution on agricultural productivity is with an instrument.
@@
 Using the results in column three, a one percent increase in particulate matter leads to a 0.62 percent decrease in agricultural productivity.
-To put this in further perspective, the mean particulate matter concentration is just short of 40 (39.68) in the monsoon sample, while the within-village absolute deviation in particulate matter is 3.98, meaning year-to-year swings of several percent in agricultural productivity due to pollution exposure are common.
+To put this in further perspective, the mean particulate matter concentration is just short of 40 (39.68) in the monsoon sample, while the within-village absolute deviation in particulate matter is 3.98, meaning year-to-year swings of several percent in agricultural productivity due to pollution exposure alone are common.
 
 Column four presents one last specification; it includes a much more flexible specification for weather.
-Specifically, I turn the linear z-score weather variable into bins for each month of the season.
-I then interact this rainfall variable with the temperature variable.
-The coefficient in column four shows almost not change from those in column three, relieving some concerns that weather may not be correctly specified, leading to a spurious correlation between particulate matter and agricultural productivity.
+This is the same binned specificaiton as I describe above.
+The coefficient in column four shows almost not change from those in column three, again relieving some concerns that weather may not be correctly specified, leading to a spurious correlation between particulate matter and agricultural productivity.
 
@@
 
-The comparison of the results to those in the previous table also confirm the concerns around endogenous placement of coal plants.
-The coefficients in Table \@ref(tab:yieldtabletwo) are always more negative than those in Table \@ref(tab:yieldtablepm), and sometimes substantially so.
+The comparison of the results to those in the previous table also confirm the concerns around endogenous placement of pollution sources.
+The coefficients in Table \@ref(tab:yieldtabletwo) are substantially more negative than those in Table \@ref(tab:yieldtablepm).
 In other words, the simple regression of agricultural productivity on pollution results in coefficients on pollution that are upwards biased, since higher levels of pollution occur in more productive areas, on average.
@@
 The second robustness check includes leads of pollution and wind direction.
+I instrument for future pollution using future wind direction, which mirrors the strategy employed in the main results.
 I present these results in Table \@ref(tab:yieldtableleads).
@@
 colnames(yield3ivmain_lead) <- c("(1)", "(2)")
-rownames(yield3ivmain_lead) <- c("Particulate matter", "(one-year lead)",
-                                 "Particulate matter", "(two-year lead)",
-                                 "Fixed effects:", "Village", "Year",
+rownames(yield3ivmain_lead) <- c("particulate matter (one-year lead)", "",
+                                  "particulate matter (two-year lead)", "",
+                                  "weather (expanded)",
+                                  "fixed effects:", "village", "year",
                                   "F", 
-                                 "Observations")
+                                  "observations")
 kable(
@@
            ) %>%
-  column_spec(1, width = "3.5cm") %>%
-  column_spec(c(2:3),width = "2cm") %>%
-  row_spec(c(7), hline_after = TRUE) %>%
-  row_spec(c(5), bold = TRUE) %>%
+  column_spec(1, width = "5.5cm") %>%
+  column_spec(c(2:3),width = "1.5cm") %>%
+  row_spec(c(8), hline_after = TRUE) %>%
+  row_spec(c(6), bold = TRUE) %>%
   kable_classic_2()
@@
 
-Having established that exposure to pollution has a negative effect on agricultural productivity, I now move to the heterogeneity in these effects.
+Having established that exposure to pollution has a negative effect on agricultural productivity, I now move to the heterogeneity in these effects to better understand whether certain areas are driving the effects.
 Table \@ref(tab:yieldtablehet) presents three different analyses: heterogeneity by median exposure, heterogeneity by initial yield, and possible multiplicative effects of shocks.
@@
 The first two columns split the sample based on median exposure.
-In other words, for each village, I calculate the median exposure (based on wind) and then use this single value for each village to split the sample into villages above and below median exposure.
-The coefficient in the first column is slightly larger than the coefficient in the second column, indicating that pollution does not have large differential effects based on exposure.
+In other words, for each village, I calculate the median exposure (based on wind) and then use this single value for each village to split the sample into villages above and below the median.
+The coefficient in the first column is only slightly larger than the coefficient in the second column, indicating that pollution does not have large differential effects based on exposure.
 One possible explanation is that exposure is not cumulative, but continues to be harmful regardless of the baseline level.
 
-Columns three and four split the sample based on median yield at the very beginning of the sample (the first year available, 2002-2003).
+Columns three and four split the sample based on median yield at the very beginning of the sample (the first year available, 2002).
 Column three presents results for villages above the median while column four presents results for villages below the median.
@@
 This could have very profound effects on welfare, since this essentially increases the variance of agricultural productivity across years (which is already increasing due to climate change), at least as it relates to pollution exposure.
-
-## Crops and labor intensity
-
-A key question about the results is the mechanism.
-While previous results indicate that pollution may actually increase land productivity [@sanders2022adaptation], the results in that paper come from the United States, where agricultural practices vary greatly from those in India.
-Unfortunately, there is no good labor data that would allow me to directly test the effects of pollution on agricultural labor productivity.
-While the National Sample Survey is nationally representative, it does not allow one to back out productivity, only total labor allocation in the preceding week.
-Additional data options -- like the ARIS-REDS or IHDS -- are not collected frequently enough to enable the kind of analysis employed here.
-
-Instead, I look at heterogeneity by the modal crop grown in each area.
-I use data from @monfreda2008farming, who estimate area harvested by crop at a resolution of approximately 10km.
-These estimates come from around the year 2000, at the beginning of my sample period.
-I identify areas where rice is the most commonly grown crop.
-Rice (paddy) is commonly regarded as a labor-intensive crop in India [@vollrath2011agricultural; @michler2020agriculture], meaning that it may be more susceptible to pollution if some of the effects of pollution are through direct effects on labor productivity.
+Given the importance of variability to welfare [@dercon2002income; @merfeld2023poverty], this could have important implications for the effects of pollution on welfare.
 
@@
 
+## Crops and labor intensity
+
+A key question about the results is the mechanism.
+While previous results indicate that pollution may actually increase land productivity [@sanders2022adaptation], the results in that paper come from the United States, where agricultural practices vary greatly from those in India.
+On the other hand, we also know that pollution can affect labor productivity in other contexts [@chang2019effect; @ebenstein2016long; @graff2012impact; @wen2022lower].
+Unfortunately, there is no available labor data that would allow me to directly test the effects of pollution on agricultural labor productivity.
+While the National Sample Survey is nationally representative, it does not allow one to back out productivity, only total labor allocation in the preceding week.
+Additional data options -- like the ARIS-REDS or IHDS -- are not collected frequently enough to enable the kind of analysis employed here.
+
+Instead, I look at heterogeneity by the modal crop grown in each area.
+I use data from @monfreda2008farming, who estimate area harvested by crop at a resolution of approximately 10km.
+These estimates come from around the year 2000, at the beginning of my sample period.
+I identify areas where rice is the most commonly grown crop.
+Rice (paddy) is commonly regarded as a labor-intensive crop in India [@vollrath2011agricultural; @michler2020agriculture], meaning that it may be more susceptible to pollution if some of the effects of pollution are through direct effects on labor productivity.
+
 Figure \@ref(fig:crops) shows the distribution of the most common crop across the villages in my sample.
@@
 colnames(yieldrice) <- c("(1)", "(2)")
-rownames(yieldrice) <- c("Particulate matter", "(log PM 2.5)",
+rownames(yieldrice) <- c("Particulate matter (log PM 2.5)", "",
                                 "PM 2.5 times rice", " ",
@@
            ) %>%
-  column_spec(1, width = "3.5cm") %>%
+  column_spec(1, width = "5.5cm") %>%
   column_spec(c(2:3),width = "2cm") %>%
@@
 
+Columns one and two of Table \@ref(tab:yield7ivdiffindiff) present specifications with slightly different weather controls.
+Column two presents the most extensive set of controls -- the flexible specification with binned weather variables -- and I focus on that column here.
+First, there are now two first stages but the F-tests indicate that the instruments are strong for both, including the interaction term (F = 205 in the second column).
+Second, the first row shows the effect of pollution on agricultural productivity prior to the opening of a coal plant; the coefficient is negative, indicating negative effects of non-coal pollution.
+Third, the interaction term is strongly negative and approximately the same in magnitude as the level term, indicating a very clear effect specific to coal plants, in addition to the overall negative effect of pollution.
+
 ```{r yield7ivdiffindiff, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
@@
 colnames(yield7ivdiffindiff) <- c("(1)", "(2)")
-rownames(yield7ivdiffindiff) <- c("Particulate matter", "(log PM 2.5)", 
+rownames(yield7ivdiffindiff) <- c("Particulate matter (log PM 2.5)", "", 
                                   "PM 2.5 times Coal", "",
-                                  "Weather", "Weather (expanded)",
+                                  "Weather (expanded)", "Weather (expanded, bins)",
                                   "Fixed effects:", "Year", "Village",
@@
 ```
-
-Columns one and two of Table \@ref(tab:yield7ivdiffindiff) present specifications with slightly different weather controls.
-Column two presents the most extensive set of controls and I focus on that column here.
-First, there are now two first stages but the F-tests indicate that the instruments are strong for both, including the interaction term (F = 205 in the second column).
-Second, the first row shows the effect of pollution on agricultural productivity prior to the opening of a coal plant; the coefficient is negative, indicating negative effects of non-coal pollution.
-Third, the interaction term is strongly negative and approximately the same in magnitude as the level term, indicating a very clear effect specific to coal plants, in addition to the overall negative effect of pollution.
 
@@
 ```
-```{r plants, echo = FALSE, message = FALSE, warning = FALSE, dev='png', fig.cap = "Coal plants in India from 1990 to 2010", fig.footer = list("Note: The top figure shows the location of coal plants in 1990.
-The bottom figure shows the location of coal plants in 2010.")}
+```{r plants, echo = FALSE, message = FALSE, warning = FALSE, dev='png', fig.cap = "Coal plants in India from 1990 to 2010", fig.footer = list("Note: The left figure shows the location of coal plants of at least 30MW in 1990.
+The right figure shows the location of such coal plants in 2010.")}
 

### 2023-07-13 draft/draft.tex
new draft

@@
 
-\title{Air pollution and agricultural productivity in a developing country\footnote{I thank Ariel Ortiz-Bobea, Marc Bellemare, and David Sungho Park for comments and suggestions that greatly improved the quality of the paper.
+\title{Air pollution and agricultural productivity in a developing country\footnote{I thank Ariel Ortiz-Bobea, Marc Bellemare, and David Sungho Park for comments and suggestions that greatly improved the quality of the paper, as well as conference attendees at the 2023 KER International Conference.
 The usual disclaimer applies.
@@
 \author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
-\date{2023-07-10}
+\date{2023-07-13}
 
@@
 
-While there are likely positive effects from this increase in power generation -- for example, electricity may lead to economic and productivity growth (Dinkelman 2011; Kline and Moretti 2014; Rud 2012; Van de Walle et al.
+While there may be positive effects from this increase in power generation -- for example, electricity may lead to economic and productivity growth (Dinkelman 2011; Kline and Moretti 2014; Rud 2012; Van de Walle et al.
 2017), though electrification alone may not be a sufficient condition (K.
 Lee, Miguel, and Wolfram 2020) -- these improvements do not come without downsides, especially when it comes to power derived from coal.
-Pollutants from coal contribute to climate change and the phase out of coal energy is therefore a major climate goal (IEA 2022).
+Pollutants from coal contribute to climate change and the phase out of coal energy is, therefore, a major climate goal (IEA 2022).
 Emissions from the production and burning of coal are also directly harmful to human health and the environment; pollutants like sulfur dioxide, nitrogen oxides, methane, and mercury are particularly harmful.\footnote{www.eia.gov/energyexplained/coal/coal-and-the-environment.php} While the use of coal had been declining, recent events -- including Russia's invasion of Ukraine -- have led to an increase in the demand for coal.\footnote{www.npr.org/2022/08/15/1117560560/a-rising-demand-for-coal-amidst-war-in-ukraine}
@@
 These effects can work through the well-known effects on human health and productivity, but also directly through land productivity.
-Black carbon in soot, for example, can directly affect solar radiation reaching plants on the surface (Ramanathan and Carmichael 2008; Burney and Ramanathan 2014).
+Black carbon in soot, for example, can directly affect solar radiation reaching plants on the surface (Ramanathan and Carmichael 2008; Burney and Ramanathan 2014), while, on the other hand, sulfur in some pollution sources can actually be productivity enhancing (Sanders and Barreca 2022).
 Relative to previous work in economics, this paper tackles pollution more broadly, before turning to a specific source of pollution: coal plants.
@@
 
-The main goal of this paper is to examine whether exposure to pollution and emissions from coal plants, in particular, affect agricultural productivity.
+The main goal of this paper is to examine whether exposure to pollution affects agricultural productivity, as well as to provide evidence about the effects of pollution from coal plants, in particular.
 I use several sources of data, which I describe here and list in Table \ref{tab:data} of the appendix.
@@
 2022), the authors use a static cropland mask for the productivity estimates.
-This means that any measurement error will be constant across time and subsumed by the village fixed effects.
+This means that any measurement error will hopefully be constant across time and subsumed by the village fixed effects.
 While the measure of output is also noisy, the use of satellite-derived data and a static cropland mask reduces concerns about measurement error being responsible for the results below, especially since common sources of measurement error are unlikely to be correlated with wind direction.
 The authors define the monsoon season as June to October and the winter season as November to March.
-I focus on only the monsoon season and keep the authors' definition when matching across data below.
-This data is also publicly available from Nature's data-sharing website.\footnote{springernature.figshare.com} Since the main regressions below use village fixed effects, any differences in cropping patterns across areas should not bias estimates on changes, as long as these patterns do not change markedly in response to pollution.
+Due to large differences in the crops grown and the production technologies used across seasons, I focus on only the monsoon season and keep the authors' definition when matching across data below.
+The data on agricultural productivity is publicly available from Nature's data-sharing website.\footnote{springernature.figshare.com} Since the main regressions below use village fixed effects, any differences in cropping patterns across areas should not bias estimates on changes, as long as these patterns do not change markedly in response to annual pollution levels.
 
@@
 
-To measure exposure to pollution from coal plants, I first locate all village centroids located within 30km from a coal plant in a given year.
+To measure exposure to pollution from point sources around coal plants, I first locate all village centroids located within 30km from a coal plant in a given year.
 I choose 30km due to previous research on the effects of (air) pollution (Aragón and Rud 2016; Bencko and Symon 1977; Li and Gibson 2014).
-I then calculate the direction from coal plants to all village centroids within that 30km radius.
-To define exposure, I then pull daily wind direction data from the National Center for Atmospheric Research.\footnote{climatedataguide.ucar.edu/} For each day, I document whether the wind is blowing towards each village centroid\footnote{I define ``towards'' as within five degrees to help take into account that I am using village centroids, using the x and y components of wind speed to calculate angles.} and then temporally aggregate this daily data depending on the temporal definition of the corresponding outcome.
-For example, agricultural productivity is defined across five months (e.g.~the monsoon season is from June to October) so I count the total days a given village is exposed to wind blowing from \emph{any} coal plant within 30km during those five months.
+I then calculate the direction from these point sources to all village centroids within that 30km radius.
+To define exposure, I pull daily wind direction data from the National Center for Atmospheric Research.\footnote{climatedataguide.ucar.edu/} For each day, I document whether the wind is blowing towards each village centroid\footnote{I define ``towards'' as within five degrees to help take into account that I am using village centroids.
+The wind data includes both the x and y components of wind speed, which I can use to calculate angles from point sources to village centroids.} and then temporally aggregate this daily data depending on the temporal definition of the corresponding outcome.
+For example, agricultural productivity is defined across five months (i.e.~the monsoon season is from June to October) so I count the total days a given village is exposed to wind blowing from \emph{any} coal plant within 30km during those five months.
 
@@
 This means that the overall exposure for a given area to coal plant emissions changes across time.
-I also pull data on particulate matter to help validate this proxy for exposure.
+
+For pollution data, I use estimated particulate matter (PM).
+Specifically, I use estimates of PM 2.5, which includes particles with a diameter of 2.5 micrometers or less.
 This data comes from Hammer et al.
-(2020) and I also aggregate this to the village using the same method detailed above.
+(2020) and is available as a raster with a resolution of 0.01 degrees.
+I aggregate this data to the village level by extracting the mean PM for each feature in the shapefile.
+I do this separately for each month from June 2002 to October 2013.
+The bottom panel in Figure \ref{fig:windexample} in the appendix is an example of the PM data.
+The figure shows the estimated PM 2.5 from for each village in Guna district of Madhya Pradesh on January 1st, 2010.
+
+While I use PM 2.5 as the estimated level of pollution, it is important to clarify that I cannot explicitly say that PM 2.5, per se, is responsible for any effects.
+Pollution includes a variety of different components, including PM 2.5, PM 10, sulfur dioxide, and nitrogen dioxide, to name but a few.
+While I use PM 2.5 as a proxy for pollution, it is possible that other components are responsible for the effects.
+Additionally, the PM 2.5 data is estimated using satellite data and a machine learning algorithm.
+While this is a common approach in the literature ( see Hammer et al.
+(2020) for more details), it is important to note that the estimates are not direct measurements of pollution.
+However, a valid instrumental variable can help for any measurement error present in the data (Glewwe 2012; Gillen, Snowberg, and Yariv 2019).
 
 To help understand some of the mechanisms driving the relationship between exposure to emissions and agricultural productivity, I also look at the differential effects by the labor intensity of the predominant crop in each area.
-For this, I use data from (Monfreda, Ramankutty, and Foley 2008), who estimate area harvested for different crops across the global around the year 2000.
+For this, I use data from Monfreda, Ramankutty, and Foley (2008), who estimate area harvested for different crops across the global around the year 2000.
 I identify areas where rice (paddy) is the most commonly grown crop, under the assumption that rice is more labor intensive than other crops (Vollrath 2011; Michler 2020).
-Michler (2020), for example, shows that rice, on avearge, uses almost 1,800 hours per hectare, while the second-highest crop in his data, cotton, juses just 850.\footnote{He estimates that wheat, another common crop in India, uses just 600 hours per hectare.}
+Michler (2020), for example, shows that rice, on avearge, uses almost 1,800 hours per hectare, while the second-highest crop in his data, cotton, juses just 850.
+He also estimates that wheat, another common crop in India, uses just 600 hours per hectare.
 
@@
 \begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3.5cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
 \toprule
@@
  &  &  & (0.001) & (0.001)\\
-Pollution &  &  & 0.051*** & 0.016***\\
-(log PM 2.5) &  &  & (0.002) & (0.001)\\
+Pollution (log PM 2.5) &  &  & 0.051*** & 0.016***\\
+ &  &  & (0.002) & (0.001)\\
 Sub-sample & all & no plant & all & no plant\\
@@
 (2020).
-\(\mathrm{PM 2.5}\) is one of the harmful byproducts from coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.\footnote{www.epa.gov/airmarkets/power-plants-and-neighboring-communities} This data is available at the monthly level, but I aggregate variables to the season variable since that is the level of analysis in the main results.
+\(\mathrm{PM 2.5}\) is one of the harmful byproducts from many industrial processes, including power generation.
+Particulate matter is a major component of the pollution emitted by coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.\footnote{www.epa.gov/airmarkets/power-plants-and-neighboring-communities} This data is available at the monthly level, but I aggregate variables to the season variable since that is the level of analysis in the main results.
 
@@
 Since the exposure variable is at the village level and we follow villages over time, the standard errors are also clustered at the village level.
-In columns two and four, I switch the month fixed effects for district-month fixed effects, which considers only villages in the same district in the same month.
-
-In addition to serving as an example, Table \ref{tab:yield3ivmainfirststage} also helps validate the use of wind direction as a proxy for exposure to pollution.
-The first two columns include all available years of data while the last two columns restrict estimation only to the years for which the agricultural productivity regressions are estimated.
-Across all four columns, the story is the same: when the wind is blowing in the direction of a village from a coal plant within 30 kilometers, estimated \(\mathrm{PM 2.5}\) is substantially higher.
+I add more and more flexible specifications for weather, with the most flexible specification in the fourth column.
+Note that the relationship between wind direction and particulate matter is very similar across all four columns, suggesting that the inclusion of weather does not change the relationship between wind direction and pollution.
+This is important because it suggests that the use of wind direction as a proxy for pollution exposure is not biased by weather patterns.
 
@@
 
-To identify possible mechanisms, I identify areas that grow predominantly rice, which is a labor-intensive crop. Specifically, we are interested in the following regression:
+A key question is how pollution affects agricultural productivity.
+Specifically, does pollution directly affect labor productivity, land productivity, or both?
+Unfortunately, there is a lack of data that would allow me to directly test these mechanisms, since I do not have data on the amount of labor or land used in each village.
+Instead, I look at heterogeneity in the effects of pollution based on the predominant crop grown in each village.
+
+Michler (2020) presents estimates of labor intensity for several major crops in India.
+The most labor-intensive crop is rice, which uses almost 1,800 hours per hectare.
+The second-most labor-intensive crop is cotton, which uses just 850 hours per hectare.
+Wheat, another common crop in India, uses just 600 hours per hectare.
+This suggests that pollution should be more harmful for rice than for other crops if there are direct impacts on labor productivity.
+
+To identify this possible mechanism, I identify areas that grow predominantly rice. Specifically, I estimate the following regression:
 \begin{gather} 
@@
 
+While this is a useful first step, it is not a perfect identification strategy.
+First, there could be other differences in the types of areas that grow rice, and these differences could actually be responsible for any heterogeneity.
+Second, this strategy does not allow me to say anything about the effects of pollution on land productivity.
+
 \hypertarget{identifying-effects-of-pollution-from-coal-plants}{%
@@
 
-I use the location of coal plants to identify locations of high pollution, but there can be other pollution sources located near coal plants, as well.
-I nonetheless want to identify the effects of pollution coming specifically from coal plants.
+I use the location of coal plants to identify locations of high pollution emission, but there can be other pollution sources located near coal plants, as well.
+I nonetheless also want to identify the effects of pollution coming specifically from coal plants, in addition to the overall effects of pollution in the other parts of the paper.
 I do this by combining the two-stage procedure above with a differences-in-differences framework.
@@
 I first present the reduced form results for agricultural productivity.
-The outcome of all regressions in this section is the log of agricultural land productivity, which is defined as tons per hectare.
+The outcome for all regressions in this section is the log of agricultural land productivity, which is defined as tons per hectare.
 The first set of results are in Table \ref{tab:yieldtable}.
@@
 The first column presents the most simple results, with only the wind exposure variable and the fixed effects.
-The second column adds a rainfall and average temperature variable, while the third column includes higher-order weather variables, as well.
-While the coefficient decreases slightly in magnitude as we move from the first to the second column and the second to the third column, the decrease is relatively small.
-There is a consistent, negative effect of wind exposure on agricultural productivity.
+The second column adds a rainfall and average temperature variable.
+While I leave them out of the table for simplicity, the coefficients are in the expected direction: rainfall is positive and average temperature is negative.
+The third column adds higher-order weather variables for both rainfall and temperature, as well as the interaction between them.
+This attenuates the relationship between wind and productivity somewhat, but the coefficient remains negative and statistically significant.
+
+The final column includes the most flexible form for the weather variables.
+The z-score for rainfall is turned into bins -- with cutoffs at \(z=-1\) and \(z=1\) -- based on rainfall in each month of the season.
+In other words, over the five months, I create 15 different indicator variables for rainfall based on the z-score in each month.
+One of these indicators drops out for each month of the year, leaving 10.
+I also include interactions between all of these and temperature, as well as between the bins and temperature squared.
+The coefficient on wind remains negative and statistically significant, and is actually slightly larger in magnitude than the results in column three, which is a much less flexible specification for weather.
+This leads me to conclude that the results are not being driven by weather patterns.
 
@@
 \begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}c}
 \toprule
-  & (1) & (2) & (3)\\
-\midrule
-Wind days & -0.0078*** & -0.0065*** & -0.0051***\\
- & (0.0007) & (0.0007) & (0.0007)\\
-Rain (z) &  & 0.0258*** & 0.2545***\\
- &  & (0.0004) & (0.0100)\\
-Mean temperature (10s) &  & -0.0023*** & 0.0654***\\
- &  & (0.0002) & (0.0028)\\
-Rain squared &  &  & -0.0410***\\
- &  &  & (0.0004)\\
-Temperature squared &  &  & -0.0001***\\
- &  &  & \vphantom{1} (0.0000)\\
-Rain times Temperature &  &  & -0.0007***\\
- &  &  & (0.0000)\\
-\textbf{Fixed effects:} & \textbf{} & \textbf{} & \textbf{}\\
-Village & Yes & Yes & Yes\\
-Year & Yes & Yes & Yes\\
-\midrule
-Observations & 1,266,922 & 1,266,922 & 1,266,922\\
+  & (1) & (2) & (3) & (4)\\
+\midrule
+wind days & -0.0078*** & -0.0065*** & -0.0051*** & -0.0053***\\
+ & (0.0007) & (0.0007) & (0.0007) & (0.0007)\\
+weather & No & Yes & Yes & No\\
+weather (expanded) & No & No & Yes & No\\
+weather (expanded, bins) & No & No & No & Yes\\
+\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+village & Yes & Yes & Yes & Yes\\
+year & Yes & Yes & Yes & Yes\\
+\midrule
+observations & 1,266,922 & 1,266,922 & 1,266,922 & 1,266,922\\
 \bottomrule
@@
 \begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{5cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}c}
 \toprule
-  & (1) & (2) & (3)\\
-\midrule
-Particulate matter & -0.042*** & -0.045*** & -0.041***\\
-(log PM 2.5) & (0.003) & (0.003) & (0.003)\\
-Weather & No & Yes & No\\
-Weather (expanded) & No & No & No\\
-\textbf{Fixed effects:} & \textbf{} & \textbf{} & \textbf{}\\
-Village & Yes & Yes & Yes\\
-Year & Yes & Yes & Yes\\
-\midrule
-Observations & 1,266,922 & 1,266,922 & 1,266,922\\
+  & (1) & (2) & (3) & (4)\\
+\midrule
+Particulate matter (log PM 2.5) & -0.042*** & -0.045*** & -0.008** & -0.041***\\
+ & (0.003) & (0.003) & (0.003) & (0.003)\\
+Weather & No & Yes & Yes & No\\
+Weather (expanded) & No & No & Yes & No\\
+weather (expanded, bins) & No & No & No & Yes\\
+\textbf{Fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+Village & Yes & Yes & Yes & Yes\\
+Year & Yes & Yes & Yes & Yes\\
+\midrule
+Observations & 1,266,922 & 1,266,922 & 1,266,922 & 1,266,922\\
 \bottomrule
@@
 There is again a consistent, negative relationship between particulate matter and agricultural productivity.
-Since both variables are logged, the coefficients are similar to elasticities, indicating that a one-percent increase in pollution leads to an approximately 0.04 percent decrease in agricultural productivity.
+Since both variables are logged, the coefficients are similar to elasticities, indicating that a one-percent increase in pollution leads to an approximately 0.04 percent decrease in agricultural productivity in three of the columns.
 
 Recall, however, the results from above: coal plants were placed in places with higher average agricultural productivity.
-This means that the simple regressions in Table \ref{tab:yieldtablepm} might actually be biased upwards: although pollution might negatively affect productivity, pollution is higher in more productive areas due to the endogenous placement of coal plants in the study areas.
+This means that the simple regressions in Table \ref{tab:yieldtablepm} might actually be biased upwards: although pollution might negatively affect productivity, pollution is higher in more productive areas due to the endogenous placement of pollution sources in the study areas.
 One possible way to recover the actual effect of pollution on agricultural productivity is with an instrument.
@@
 
-The comparison of the results to those in the previous table also confirm the concerns around endogenous placement of coal plants.
-The coefficients in Table \ref{tab:yieldtabletwo} are always more negative than those in Table \ref{tab:yieldtablepm}, and sometimes substantially so.
+The comparison of the results to those in the previous table also confirm the concerns around endogenous placement of pollution sources.
+The coefficients in Table \ref{tab:yieldtabletwo} are substantially more negative than those in Table \ref{tab:yieldtablepm}.
 In other words, the simple regression of agricultural productivity on pollution results in coefficients on pollution that are upwards biased, since higher levels of pollution occur in more productive areas, on average.
@@
 The second robustness check includes leads of pollution and wind direction.
+I instrument for future pollution using future wind direction, which mirrors the strategy employed in the main results.
 I present these results in Table \ref{tab:yieldtableleads}.
@@
 \begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3.5cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{5.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}}
 \toprule
@@
 \midrule
-Particulate matter & -0.073 & \\
-(one-year lead) & (0.066) & \\
-Particulate matter &  & -0.079\\
-(two-year lead) &  & (0.051)\\
-\textbf{Fixed effects:} & \textbf{} & \textbf{}\\
-Village & Yes & Yes\\
-Year & Yes & Yes\\
-\midrule
-F & 603 & 746\\
-Observations & 1,161,265 & 1,055,562\\
+particulate matter (one-year lead) & -0.033 & \\
+ & (0.067) & \\
+particulate matter (two-year lead) &  & -0.070\\
+ &  & (0.052)\\
+weather (expanded) & No & No\\
+\textbf{fixed effects:} & \textbf{} & \textbf{}\\
+village & Yes & Yes\\
+year & Yes & Yes\\
+\midrule
+F & 592 & 783\\
+observations & 1,161,265 & 1,055,562\\
 \bottomrule
@@
 
-Having established that exposure to pollution has a negative effect on agricultural productivity, I now move to the heterogeneity in these effects.
+Having established that exposure to pollution has a negative effect on agricultural productivity, I now move to the heterogeneity in these effects to better understand whether certain areas are driving the effects.
 Table \ref{tab:yieldtablehet} presents three different analyses: heterogeneity by median exposure, heterogeneity by initial yield, and possible multiplicative effects of shocks.
@@
 This could have very profound effects on welfare, since this essentially increases the variance of agricultural productivity across years (which is already increasing due to climate change), at least as it relates to pollution exposure.
+Given the importance of variability to welfare (Dercon 2002; Merfeld and Morduch 2023), this could have important implications for the effects of pollution on welfare.
+
+\begin{figure}
+\includegraphics{draft_files/figure-latex/crops-1} \caption[Most common crop by village]{Most common crop by village}\label{fig:crops}\floatfoot*{Note: The figure shows the most common crop grown in each village, by area harvested.}
+\end{figure}
 
@@
 While previous results indicate that pollution may actually increase land productivity (Sanders and Barreca 2022), the results in that paper come from the United States, where agricultural practices vary greatly from those in India.
-Unfortunately, there is no good labor data that would allow me to directly test the effects of pollution on agricultural labor productivity.
+On the other hand, we also know that pollution can affect labor productivity in other contexts (Chang et al.
+2019; Ebenstein, Lavy, and Roth 2016; Graff Zivin and Neidell 2012; Wen and Burke 2022).
+Unfortunately, there is no available labor data that would allow me to directly test the effects of pollution on agricultural labor productivity.
 While the National Sample Survey is nationally representative, it does not allow one to back out productivity, only total labor allocation in the preceding week.
@@
 
-\begin{figure}
-\includegraphics{draft_files/figure-latex/crops-1} \caption[Most common crop by village]{Most common crop by village}\label{fig:crops}\floatfoot*{Note: The figure shows the most common crop grown in each village, by area harvested.}
-\end{figure}
-
 Figure \ref{fig:crops} shows the distribution of the most common crop across the villages in my sample.
@@
 \begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3.5cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{5.5cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
 \toprule
@@
 \midrule
-Particulate matter & -0.287** & -0.394***\\
-(log PM 2.5) & (0.133) & (0.134)\\
+Particulate matter (log PM 2.5) & -0.287** & -0.394***\\
+ & (0.133) & (0.134)\\
 PM 2.5 times rice & -0.338*** & -0.205***\\
@@
 
+Columns one and two of Table \ref{tab:yield7ivdiffindiff} present specifications with slightly different weather controls.
+Column two presents the most extensive set of controls -- the flexible specification with binned weather variables -- and I focus on that column here.
+First, there are now two first stages but the F-tests indicate that the instruments are strong for both, including the interaction term (F = 205 in the second column).
+Second, the first row shows the effect of pollution on agricultural productivity prior to the opening of a coal plant; the coefficient is negative, indicating negative effects of non-coal pollution.
+Third, the interaction term is strongly negative and approximately the same in magnitude as the level term, indicating a very clear effect specific to coal plants, in addition to the overall negative effect of pollution.
+
 \begin{table}
@@
 \midrule
-Particulate matter & -0.228*** & -0.223***\\
-(log PM 2.5) & (0.066) & (0.065)\\
+Particulate matter (log PM 2.5) & -0.228*** & -0.223***\\
+ & (0.066) & (0.065)\\
 PM 2.5 times Coal & -0.250*** & -0.220***\\
  & (0.024) & (0.025)\\
-Weather & Yes & No\\
-Weather (expanded) & No & Yes\\
+Weather (expanded) & Yes & No\\
+Weather (expanded, bins) & No & Yes\\
 \textbf{Fixed effects:} & \textbf{} & \textbf{}\\
@@
 
-Columns one and two of Table \ref{tab:yield7ivdiffindiff} present specifications with slightly different weather controls.
-Column two presents the most extensive set of controls and I focus on that column here.
-First, there are now two first stages but the F-tests indicate that the instruments are strong for both, including the interaction term (F = 205 in the second column).
-Second, the first row shows the effect of pollution on agricultural productivity prior to the opening of a coal plant; the coefficient is negative, indicating negative effects of non-coal pollution.
-Third, the interaction term is strongly negative and approximately the same in magnitude as the level term, indicating a very clear effect specific to coal plants, in addition to the overall negative effect of pollution.
-
 Importantly, since these results come from an IV, both coefficients denote the change in agricultural yield due to an increase in pollution by the same units.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-dercon2002income}{}}%
+Dercon, Stefan. 2002. {``Income Risk, Coping Strategies, and Safety Nets.''} \emph{{The World Bank Research Observer}} 17 (2): 141--66.
+
 \leavevmode\vadjust pre{\hypertarget{ref-dercon2009rural}{}}%
-Dercon, Stefan. 2009. {``{Rural poverty: Old challenges in new contexts}.''} \emph{{The World Bank Research Observer}} 24 (1): 1--28.
+---------. 2009. {``{Rural poverty: Old challenges in new contexts}.''} \emph{{The World Bank Research Observer}} 24 (1): 1--28.
 
@@
 {``{A new two-decade (2001--2019) high-resolution agricultural primary productivity dataset for India}.''} \emph{{Scientific Data}} 9 (1): 1--12.
+
+\leavevmode\vadjust pre{\hypertarget{ref-gillen2019experimenting}{}}%
+Gillen, Ben, Erik Snowberg, and Leeat Yariv.
+2019.
+{``{Experimenting with measurement error: Techniques with applications to the Caltech cohort study}.''} \emph{{Journal of Political Economy}} 127 (4): 1826--63.
+
+\leavevmode\vadjust pre{\hypertarget{ref-glewwe2012much}{}}%
+Glewwe, Paul.
+2012.
+{``{How much of observed economic mobility is measurement error?
+IV methods to reduce measurement error bias, with an application to Vietnam}.''} \emph{{The World Bank Economic Review}} 26 (2): 236--64.
 
@@
 {International Institute for Environment and Development.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-merfeld2023poverty}{}}%
+Merfeld, Joshua D, and Jonathan Morduch. 2023. {``Poverty at Higher Frequency.''} \emph{{KDI School of Pub Policy \& Management Paper No. DP23-03}}.
 
@@
 \begin{figure}
-\includegraphics{draft_files/figure-latex/plants-1} \caption[Coal plants in India from 1990 to 2010]{Coal plants in India from 1990 to 2010}\label{fig:plants}\floatfoot*{Note: The top figure shows the location of coal plants in 1990.
-The bottom figure shows the location of coal plants in 2010.}
+\includegraphics{draft_files/figure-latex/plants-1} \caption[Coal plants in India from 1990 to 2010]{Coal plants in India from 1990 to 2010}\label{fig:plants}\floatfoot*{Note: The left figure shows the location of coal plants of at least 30MW in 1990.
+The right figure shows the location of such coal plants in 2010.}
 \end{figure}

### 2023-07-13 analyses/2.pollution_wind_validation.R

@@ -1,102 +1,0 @@
-# Purpose: This script shows that wind does indeed predict pollution. I save the results to load into the markdown script
-# Author: Josh Merfeld
-# Date: February 25th, 2023
-
-rm(list = ls())
-
-library(tidyverse)
-library(lubridate)
-library(viridis)
-library(fixest)
-library(sf)
-
-
-# Sets wd to folder this script is in so we can create relative paths instead of absolute paths
-setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
-# We want to get the wd up to the main folder
-# Need to go up two levels
-setwd("../../")
-# Double check
-getwd()    # check
-
-
-
-pollution <- read_csv(paste0("data/clean/pm25/all_combined.csv"))
-pollution <- pollution %>% mutate(
-                                  date = as_date(paste0(year, "-", month, "-01")),
-                                  state = substr(shrid, 1, 2),
-                                  village = shrid,
-                                  month = as.numeric(date),
-                                  wind = sums
-                                  )
-pollution <- pollution %>% mutate()
-pollution <- panel(pollution, ~ village + date)
-
-# And district identifiers
-villages_overlap <- read_sf("data/spatial/villages_overlap/villages_overlap.shp")
-villages_overlap <- villages_overlap %>% mutate(shrid = paste0(pc11_s_id, "-", pc11_tv_id),
-                                                distfe = paste0(pc11_s_id, "-", pc11_d_id)) %>%
-                                          as_tibble() %>%
-                                          dplyr::select(shrid, distfe) %>%
-                                          group_by(shrid) %>%
-                                          filter(row_number()==1) %>%
-                                          ungroup()
-pollution <- pollution %>% left_join(villages_overlap, by = "shrid")
-
-pollution <- pollution %>% group_by(shrid) %>%
-                              mutate(pm_abs_dev = abs(mean - mean(mean)),
-                                     pm_dev = (mean - mean(mean)),
-                                     wind_dev = (wind - mean(wind))) %>%
-                              ungroup()
-pollution$month_int <- month(pollution$date)
-pollution <- pollution %>% group_by(shrid, month_int) %>%
-                              mutate(m_pm_dev = (mean - mean(mean)),
-                                     m_wind_dev = (wind - mean(wind))) %>%
[…]

### 2023-07-13 analyses/3.yield.R

@@ -187,22 +187,27 @@ yield3 <- feols(log(yield) ~ wind + ..ctrl2 | shrid + year, data = df, cluster =
 yield4 <- feols(log(yield) ~ wind + ..ctrl3 | shrid + year, data = df, cluster = c("shrid"))
 
 yieldtable <- etable(
-                     yield1, yield2, yield3,
+                     yield1, yield2, yield3, yield4,
                      se.below = TRUE,
                      depvar = FALSE,
                      signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.1),
                      digits = "r4",
                      fitstat = c("n"),
-                     coefstat = "se"
+                      group = list(controls = "rain_z",
+                                   conrols_expanded = "rain_z square",
+                                   conrols_expanded2 = "rain1_zbin0"),
+                     coefstat = "se",
+                     keep = c("wind")
                      )
 # rename
-yieldtable <- yieldtable[-c(16:17),]
-yieldtable[13,] <- " "
+yieldtable <- yieldtable[-c(9:10),]
+yieldtable[6,] <- " "
 yieldtable <- as.matrix(yieldtable)
-rownames(yieldtable) <- c("wind days", " ", "rain (z)", "", "mean temp (10s)", "",
-                          "rain square", " ", "temp square", " ", "rain times temp", " ",
+rownames(yieldtable) <- c("wind days", " ", 
+                          "weather", "weather (expanded)", "weather (expanded, bins)",
                           "fixed effects:", "village", "year",
                           "observations")
+yieldtable <- yieldtable[,-1]
 saveRDS(yieldtable, "pollution_development/draft/tables/yield1reducedform.rds")
 
 
@@ -231,7 +236,7 @@ yield3 <- feols(log(yield) ~ pm25 + ..ctrl2 | shrid + year, data = df, cluster =
 yield4 <- feols(log(yield) ~ pm25 + ..ctrl3 | shrid + year, data = df, cluster = c("shrid"))
 
 yieldtable <- etable(
-                      yield1, yield2, yield3,
+                      yield1, yield2, yield3, yield4,
                       se.below = TRUE,
                       depvar = FALSE,
                       signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.1),
[…]

### 2023-07-13 analyses/4.nss.R

@@ -1,372 +1,0 @@
-# Purpose: This script estimates effects with the NSS data. I save the results to load into the markdown script
-# Author: Josh Merfeld
-# Date: February 25th, 2023
-
-rm(list = ls())
-
-library(tidyverse)
-library(lubridate)
-library(viridis)
-library(fixest)
-library(sf)
-library(tree)
-library(marginaleffects)
-
-
-# Sets wd to folder this script is in so we can create relative paths instead of absolute paths
-#setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
-# We want to get the wd up to the main folder
-# Need to go up two levels
-setwd("../")
-# Double check
-getwd()    # check
-
-
-
-
-
-df <- read_csv("data/clean/nss/merged_week.csv")
-df <- df %>% filter(age>=15)
-df <- df %>% mutate(
-                    month_int = month(date),
-                    year = year(date),
-                    month = as.numeric(date),
-                    educ = as.factor(educ),
-                    wind = days_sum,
-                    district = distfe
-                    )
-
-# MONSOON ONLY
-df <- df %>% filter(month_int %in% c(6,7,8,9,10))
-
-df$precip_sum <- df$month6_precip
-df$precip_sum[month(df$date)==7] <- df$month6_precip[month(df$date)==7] + df$month7_precip[month(df$date)==7]
-df$precip_sum[month(df$date)==8] <- df$month6_precip[month(df$date)==8] + df$month7_precip[month(df$date)==8] + df$month8_precip[month(df$date)==8]
-df$precip_sum[month(df$date)==9] <- df$month6_precip[month(df$date)==9] + df$month7_precip[month(df$date)==9] + df$month8_precip[month(df$date)==9] + df$month9_precip[month(df$date)==9]
-df$precip_sum[month(df$date)==10] <- df$month6_precip[month(df$date)==10] + df$month7_precip[month(df$date)==10] + df$month8_precip[month(df$date)==10] + df$month9_precip[month(df$date)==10] + df$month_precip[month(df$date)==10]
-
-# create seasons means
-df <- df %>% 
-      group_by(year, district) %>%
-      mutate(temp_mean = mean((month_tmin + month_tmax)/2),
-             precip_sum_season = mean(month_precip)) %>%
-      ungroup()
-
-# add mean of rainfall
-means <- read_csv(paste0("data/clean/terra_district/precip_mean_monthALL.csv"))
-means <- means %>%
[…]

### 2023-07-13 draft/draft.Rmd
introduction update

@@
 While there may be positive effects from this increase in power generation -- for example, electricity may lead to economic and productivity growth [@dinkelman2011effects; @kline2014local; @rud2012electricity; @van2017long], though electrification alone may not be a sufficient condition [@lee2020does] -- these improvements do not come without downsides, especially when it comes to power derived from coal.
-Pollutants from coal contribute to climate change and the phase out of coal energy is, therefore, a major climate goal [@iea2022].
-Emissions from the production and burning of coal are also directly harmful to human health and the environment; pollutants like sulfur dioxide, nitrogen oxides, methane, and mercury are particularly harmful.[^4] While the use of coal had been declining, recent events -- including Russia's invasion of Ukraine -- have led to an increase in the demand for coal.[^5]
-
-[^4]: www.eia.gov/energyexplained/coal/coal-and-the-environment.php
+Emissions from the production and burning of coal are also directly harmful to human health and the environment; pollutants like sulfur dioxide, nitrogen oxides, methane, and mercury are particularly harmful.
+While the use of coal had been declining, recent events -- including Russia's invasion of Ukraine -- have led to an increase in the demand for coal.[^5]
 
@@
 Given the large effects of pollution on health and previously documented impacts on agriculture, this raises a important question: do rising pollution levels in developing countries threaten agricultural productivity?
+
+Although we have made much progress over the last several decades, poverty and inequality remain important challenges in the developing world [@alvaredo2015recent].
+Poverty is particularly prevalent in rural areas, where the majority of the world's poor live [@dercon2009rural], and in the agricultural sector, where the majority of the world's poor work [@christiaensen2011evolving].
+As such, agricultural productivity is a key determinant of poverty in developing countries [@dercon2009rural], meaning that any threat to agricultural productivity is also a threat to poverty reduction.
 
@@
 We already have evidence of effects of pollution on different forms of labor activity.
-For example, general levels of pollution decrease productivity of call center works in China [@chang2019effect], lead people to perform worse on cognitive functions [@ebenstein2016long; @wen2022lower], and also lead to lower levels of farm labor productivity in California, driven by changes in ozone concentration [@graff2012impact].
+For example, general levels of pollution decrease productivity of call center works in China [@chang2019effect], lead people to perform worse on cognitive functions [@ebenstein2016long; @wen2022lower], lower labor market earnings [@borgschulte2022air], and also lead to lower levels of farm labor productivity in California, driven by changes in ozone concentration [@graff2012impact].
 Also in China, @he2019severe find small but significant negative effects of pollution on manufacturing productivity, while @chen2022effect show pollution leads to labor out migration.

### 2023-07-13 draft/draft.tex
introduction update

@@
 Lee, Miguel, and Wolfram 2020) -- these improvements do not come without downsides, especially when it comes to power derived from coal.
-Pollutants from coal contribute to climate change and the phase out of coal energy is, therefore, a major climate goal (IEA 2022).
-Emissions from the production and burning of coal are also directly harmful to human health and the environment; pollutants like sulfur dioxide, nitrogen oxides, methane, and mercury are particularly harmful.\footnote{www.eia.gov/energyexplained/coal/coal-and-the-environment.php} While the use of coal had been declining, recent events -- including Russia's invasion of Ukraine -- have led to an increase in the demand for coal.\footnote{www.npr.org/2022/08/15/1117560560/a-rising-demand-for-coal-amidst-war-in-ukraine}
+Emissions from the production and burning of coal are also directly harmful to human health and the environment; pollutants like sulfur dioxide, nitrogen oxides, methane, and mercury are particularly harmful.
+While the use of coal had been declining, recent events -- including Russia's invasion of Ukraine -- have led to an increase in the demand for coal.\footnote{www.npr.org/2022/08/15/1117560560/a-rising-demand-for-coal-amidst-war-in-ukraine}
 
@@
 Given the large effects of pollution on health and previously documented impacts on agriculture, this raises a important question: do rising pollution levels in developing countries threaten agricultural productivity?
+
+Although we have made much progress over the last several decades, poverty and inequality remain important challenges in the developing world (Alvaredo and Gasparini 2015).
+Poverty is particularly prevalent in rural areas, where the majority of the world's poor live (Dercon 2009), and in the agricultural sector, where the majority of the world's poor work (Christiaensen, Demery, and Kuhl 2011).
+As such, agricultural productivity is a key determinant of poverty in developing countries (Dercon 2009), meaning that any threat to agricultural productivity is also a threat to poverty reduction.
 
@@
 For example, general levels of pollution decrease productivity of call center works in China (Chang et al.
-2019), lead people to perform worse on cognitive functions (Ebenstein, Lavy, and Roth 2016; Wen and Burke 2022), and also lead to lower levels of farm labor productivity in California, driven by changes in ozone concentration (Graff Zivin and Neidell 2012).
+2019), lead people to perform worse on cognitive functions (Ebenstein, Lavy, and Roth 2016; Wen and Burke 2022), lower labor market earnings (Borgschulte, Molitor, and Zou 2022), and also lead to lower levels of farm labor productivity in California, driven by changes in ozone concentration (Graff Zivin and Neidell 2012).
 Also in China, He, Liu, and Salvo (2019) find small but significant negative effects of pollution on manufacturing productivity, while Chen, Oliva, and Zhang (2022) show pollution leads to labor out migration.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-alvaredo2015recent}{}}%
+Alvaredo, Facundo, and Leonardo Gasparini. 2015. {``Recent Trends in Inequality and Poverty in Developing Countries.''} \emph{{Handbook of Income Distribution}} 2: 697--805.
+
 \leavevmode\vadjust pre{\hypertarget{ref-aragon2016polluting}{}}%
@@
 Blazquez, Jorge, Rolando Fuentes, and Baltasar Manzano. 2020. {``On Some Economic Principles of the Energy Transition.''} \emph{{Energy Policy}} 147: 111807.
+
+\leavevmode\vadjust pre{\hypertarget{ref-borgschulte2022air}{}}%
+Borgschulte, Mark, David Molitor, and Eric Yongchen Zou. 2022. {``{Air pollution and the labor market: Evidence from wildfire smoke}.''} \emph{{Review of Economics and Statistics}}, 1--46.
 
@@
 Herrendorf, Berthold, Richard Rogerson, and Akos Valentinyi. 2014. {``Growth and Structural Transformation.''} \emph{{Handbook of Economic Growth}} 2: 855--941.
-
-\leavevmode\vadjust pre{\hypertarget{ref-iea2022}{}}%
-IEA. 2022. {``{Coal-fired electricity}.''} {International Energy Association}.
 

### 2023-07-20 draft/draft.Rmd
updating draft

@@
 ---
-title: "Air pollution and agricultural productivity in a developing country^[I thank Ariel Ortiz-Bobea, Marc Bellemare, and David Sungho Park for comments and suggestions that greatly improved the quality of the paper, as well as conference attendees at the 2023 KER International Conference.
+title: "Air pollution and agricultural productivity in a developing country^[I thank Ariel Ortiz-Bobea, Marc Bellemare, and David Sungho Park for comments and suggestions that greatly improved the quality of the paper, as well as conference attendees at the 2023 KER International Conference and the University of Reading.
 The usual disclaimer applies.
@@
   | \noindent I document negative externalities of air pollution in the Indian agricultural sector.
-Using variation in pollution induced by changes in wind across years, I show that higher levels of pollution lead to decreased agricultural productivity, with large changes in productivity being common.
-The negative effects of pollution are larger in areas growing more labor-intensive crops, indicating that the pollution works at least partly through direct effects on labor productivity.
-Finally, combining wind direction with the rollout of coal plants, results indicate that pollution from coal plants has a larger effect on agricultural productivity than other types of pollution.
-Given that the agricultural sector is a refuge for the poor in many developing countries, these results suggest that the negative externalities of pollution may hit the poorest particularly hard.
+Using variation in pollution induced by changes in wind across years, I show that higher levels of pollution lead to decreased agricultural productivity.
+The negative effects of pollution are larger in areas growing more labor-intensive crops, indicating that the pollution works partly through direct effects on labor productivity.
+I also show that pollution from coal plants has a larger effect on agricultural productivity than other types of pollution.
+These results suggest that the negative externalities of pollution may hit the poorest particularly hard, given their reliance on agriculture.
   |
@@
 
+<!-- 
+Jonathan's comments:
+- Drop "in a developing country" from the title?
+- Add a sense of magnitudes, specifically the effects of pollution on yields. Like "pollution has increased on avearge by X percent in India, and this has led to a Y percent decrease in yields"
+  - This is important
+- Build up labor productivity bit a little more
+  - Jonathan thinks current writing makes it sound more like a technical point than a big deal (which it is)
+- Look at pollution prior to season? Kind of a "placebo" test
+- Maybe move labor and coal parts up in the introduction
+- Add a citation on the difference between PM 2.5 and PM 10
+  - Jonathan says he didn't understand the paragraph on PM 2.5 until we discussed it. Important to make sure the reader understands it.
+-->
+
 ```{=tex}
@@
 
-While reliance on coal is declining in much of the world, it continues to be an important source of energy in many countries.
 In 2021, energy generation from coal reached an all-time high,[^1] driven in part by rapid population growth in developing countries, including India.
-From 1990 to 2010, India's population grew by more than 40 percent, leading to a large increase in demand for power.[^2] While natural gas has become a popular alternative in much of the world, the Indian government met much of the increased demand through the construction of coal power plants.
-During those same two decades, power generation from all coal units of at least 30 megawatts in the country more than doubled, from 42.4 gigawatts to more than 100 gigawatts.[^3]
+While natural gas has become a popular alternative in much of the world, the Indian government met much of the increased demand through the construction of coal power plants.
+From 1990 to 2010, India's population grew by more than 40 percent, leading to a large increase in demand for power.[^2] During those same two decades, power generation from all coal units of at least 30 megawatts in the country more than doubled, from 42.4 gigawatts to more than 100 gigawatts.[^3]
 
@@
 
-While there may be positive effects from this increase in power generation -- for example, electricity may lead to economic and productivity growth [@dinkelman2011effects; @kline2014local; @rud2012electricity; @van2017long], though electrification alone may not be a sufficient condition [@lee2020does] -- these improvements do not come without downsides, especially when it comes to power derived from coal.
-Emissions from the production and burning of coal are also directly harmful to human health and the environment; pollutants like sulfur dioxide, nitrogen oxides, methane, and mercury are particularly harmful.
-While the use of coal had been declining, recent events -- including Russia's invasion of Ukraine -- have led to an increase in the demand for coal.[^5]
+While electricity may lead to economic and productivity growth [@dinkelman2011effects; @kline2014local; @rud2012electricity; @van2017long], though electrification alone may not be a sufficient condition [@lee2020does], these improvements do not come without downsides, especially when it comes to power derived from coal.
+Emissions from the production and burning of coal -- pollutants like sulfur dioxide, nitrogen oxides, methane, and mercury, for example -- are also directly harmful to human health and the environment.
+Although the use of coal had been declining, recent events -- including Russia's invasion of Ukraine -- have led to an increase in the demand for coal.[^5]
 
@@
 
-In addition to pollution from coal, pollution more generally is also increasing in much of the world.
-Carbon dioxide equivalent per capita has almost double over the last 50 years in India, more than doubled in Indonesia, and almost tripled in China.^[https://edgar.jrc.ec.europa.eu] Much of this increase is due to the development process itself -- as countries develop, their economies transform, with more people moving out of agriculture and into manufacturing and services [@herrendorf2014growth].
-Despite this development, agriculture remains a key sector in many developing countries, employing a large proportion of the poor [@dercon2009rural].
-This also means agriculture is a key sector when it comes to poverty reduction [@christiaensen2011evolving].
-Given the large effects of pollution on health and previously documented impacts on agriculture, this raises a important question: do rising pollution levels in developing countries threaten agricultural productivity?
-
-Although we have made much progress over the last several decades, poverty and inequality remain important challenges in the developing world [@alvaredo2015recent].
-Poverty is particularly prevalent in rural areas, where the majority of the world's poor live [@dercon2009rural], and in the agricultural sector, where the majority of the world's poor work [@christiaensen2011evolving].
-As such, agricultural productivity is a key determinant of poverty in developing countries [@dercon2009rural], meaning that any threat to agricultural productivity is also a threat to poverty reduction.
+Pollution more generally is also increasing in much of the world.
+Carbon dioxide equivalent per capita has almost doubled over the last 50 years in India, more than doubled in Indonesia, and almost tripled in China.^[https://edgar.jrc.ec.europa.eu] Much of this increase is due to the development process itself -- as countries develop, their economies transform, with more people moving out of agriculture and into manufacturing and services [@herrendorf2014growth].
+Despite this development, agriculture remains a key sector in many developing countries, employing a large proportion of the poor [@dercon2009rural], and agriculture is a key sector for poverty reduction [@christiaensen2011evolving].
+This raises an important question: do rising pollution levels in developing countries threaten agricultural productivity?
 
@@
 I then aggregate this exposure variable based on the desired analysis; for agricultural productivity, for example, I create an exposure variable measuring the total number of exposure days in the five-month agricultural season, focusing only on the monsoon (Kharif) season.
-More days being located downwind from a high-pollution location leads to substantially higher average concentrations of particulate matter 2.5 (PM 2.5).^[PM 2.5 is defined as particulate matter with a diameter of less than 2.5 micrometers.]
+More days being located downwind from a high-pollution location leads to substantially higher average concentrations of particulate matter 2.5 (PM 2.5), which is defined as particulate matter with a diameter of less than 2.5 micrometers.
 
@@
 
-The final set of analyses tries to disentangle the effects of pollution from the effects of coal pollution, specifically, by using wind direction and the rollout of coal plants across the country.
+The final set of analyses disentangles the effects of pollution from the effects of coal pollution, specifically, by using wind direction and the rollout of coal plants across the country.
 Combining instrumental variables with a differences-in-differences strategy, I show that agricultural productivity has an elasticity of around -0.22 with respect to non-coal pollution.
@@
 
-This paper contributes to several strands of literature.
-We already have evidence of effects of pollution on different forms of labor activity.
+The results are consistent with evidence of effects of pollution on different forms of labor activity.
 For example, general levels of pollution decrease productivity of call center works in China [@chang2019effect], lead people to perform worse on cognitive functions [@ebenstein2016long; @wen2022lower], lower labor market earnings [@borgschulte2022air], and also lead to lower levels of farm labor productivity in California, driven by changes in ozone concentration [@graff2012impact].
 Also in China, @he2019severe find small but significant negative effects of pollution on manufacturing productivity, while @chen2022effect show pollution leads to labor out migration.
-Relevant to my results, @hanna2015effect show that decreases in pollution lead to increases in labor allocation in another developing country.
+Relevant to my results, @hanna2015effect show that decreases in pollution lead to increases in labor allocation in Mexico.
 

### 2023-07-20 draft/draft.tex
updating draft

@@
 
-\title{Air pollution and agricultural productivity in a developing country\footnote{I thank Ariel Ortiz-Bobea, Marc Bellemare, and David Sungho Park for comments and suggestions that greatly improved the quality of the paper, as well as conference attendees at the 2023 KER International Conference.
+\title{Air pollution and agricultural productivity in a developing country\footnote{I thank Ariel Ortiz-Bobea, Marc Bellemare, and David Sungho Park for comments and suggestions that greatly improved the quality of the paper, as well as conference attendees at the 2023 KER International Conference and the University of Reading.
 The usual disclaimer applies.
@@
 \author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
-\date{2023-07-13}
+\date{2023-07-19}
 
@@
 \noindent I document negative externalities of air pollution in the Indian agricultural sector.
-Using variation in pollution induced by changes in wind across years, I show that higher levels of pollution lead to decreased agricultural productivity, with large changes in productivity being common.
-The negative effects of pollution are larger in areas growing more labor-intensive crops, indicating that the pollution works at least partly through direct effects on labor productivity.
-Finally, combining wind direction with the rollout of coal plants, results indicate that pollution from coal plants has a larger effect on agricultural productivity than other types of pollution.
-Given that the agricultural sector is a refuge for the poor in many developing countries, these results suggest that the negative externalities of pollution may hit the poorest particularly hard.\\
+Using variation in pollution induced by changes in wind across years, I show that higher levels of pollution lead to decreased agricultural productivity.
+The negative effects of pollution are larger in areas growing more labor-intensive crops, indicating that the pollution works partly through direct effects on labor productivity.
+I also show that pollution from coal plants has a larger effect on agricultural productivity than other types of pollution.
+These results suggest that the negative externalities of pollution may hit the poorest particularly hard, given their reliance on agriculture.\\
 \strut \\
@@
 
-While reliance on coal is declining in much of the world, it continues to be an important source of energy in many countries.
 In 2021, energy generation from coal reached an all-time high,\footnote{www.iea.org/reports/coal-fired-electricity} driven in part by rapid population growth in developing countries, including India.
-From 1990 to 2010, India's population grew by more than 40 percent, leading to a large increase in demand for power.\footnote{data.worldbank.org} While natural gas has become a popular alternative in much of the world, the Indian government met much of the increased demand through the construction of coal power plants.
-During those same two decades, power generation from all coal units of at least 30 megawatts in the country more than doubled, from 42.4 gigawatts to more than 100 gigawatts.\footnote{globalenergymonitor.org/}
-
-While there may be positive effects from this increase in power generation -- for example, electricity may lead to economic and productivity growth (Dinkelman 2011; Kline and Moretti 2014; Rud 2012; Van de Walle et al.
+While natural gas has become a popular alternative in much of the world, the Indian government met much of the increased demand through the construction of coal power plants.
+From 1990 to 2010, India's population grew by more than 40 percent, leading to a large increase in demand for power.\footnote{data.worldbank.org} During those same two decades, power generation from all coal units of at least 30 megawatts in the country more than doubled, from 42.4 gigawatts to more than 100 gigawatts.\footnote{globalenergymonitor.org/}
+
+While electricity may lead to economic and productivity growth (Dinkelman 2011; Kline and Moretti 2014; Rud 2012; Van de Walle et al.
 2017), though electrification alone may not be a sufficient condition (K.
-Lee, Miguel, and Wolfram 2020) -- these improvements do not come without downsides, especially when it comes to power derived from coal.
-Emissions from the production and burning of coal are also directly harmful to human health and the environment; pollutants like sulfur dioxide, nitrogen oxides, methane, and mercury are particularly harmful.
-While the use of coal had been declining, recent events -- including Russia's invasion of Ukraine -- have led to an increase in the demand for coal.\footnote{www.npr.org/2022/08/15/1117560560/a-rising-demand-for-coal-amidst-war-in-ukraine}
-
-In addition to pollution from coal, pollution more generally is also increasing in much of the world.
-Carbon dioxide equivalent per capita has almost double over the last 50 years in India, more than doubled in Indonesia, and almost tripled in China.\footnote{\url{https://edgar.jrc.ec.europa.eu}} Much of this increase is due to the development process itself -- as countries develop, their economies transform, with more people moving out of agriculture and into manufacturing and services (Herrendorf, Rogerson, and Valentinyi 2014).
-Despite this development, agriculture remains a key sector in many developing countries, employing a large proportion of the poor (Dercon 2009).
-This also means agriculture is a key sector when it comes to poverty reduction (Christiaensen, Demery, and Kuhl 2011).
-Given the large effects of pollution on health and previously documented impacts on agriculture, this raises a important question: do rising pollution levels in developing countries threaten agricultural productivity?
-
-Although we have made much progress over the last several decades, poverty and inequality remain important challenges in the developing world (Alvaredo and Gasparini 2015).
-Poverty is particularly prevalent in rural areas, where the majority of the world's poor live (Dercon 2009), and in the agricultural sector, where the majority of the world's poor work (Christiaensen, Demery, and Kuhl 2011).
-As such, agricultural productivity is a key determinant of poverty in developing countries (Dercon 2009), meaning that any threat to agricultural productivity is also a threat to poverty reduction.
+Lee, Miguel, and Wolfram 2020), these improvements do not come without downsides, especially when it comes to power derived from coal.
+Emissions from the production and burning of coal -- pollutants like sulfur dioxide, nitrogen oxides, methane, and mercury, for example -- are also directly harmful to human health and the environment.
+Although the use of coal had been declining, recent events -- including Russia's invasion of Ukraine -- have led to an increase in the demand for coal.\footnote{www.npr.org/2022/08/15/1117560560/a-rising-demand-for-coal-amidst-war-in-ukraine}
+
+Pollution more generally is also increasing in much of the world.
+Carbon dioxide equivalent per capita has almost doubled over the last 50 years in India, more than doubled in Indonesia, and almost tripled in China.\footnote{\url{https://edgar.jrc.ec.europa.eu}} Much of this increase is due to the development process itself -- as countries develop, their economies transform, with more people moving out of agriculture and into manufacturing and services (Herrendorf, Rogerson, and Valentinyi 2014).
+Despite this development, agriculture remains a key sector in many developing countries, employing a large proportion of the poor (Dercon 2009), and agriculture is a key sector for poverty reduction (Christiaensen, Demery, and Kuhl 2011).
+This raises an important question: do rising pollution levels in developing countries threaten agricultural productivity?
 
@@
 I then aggregate this exposure variable based on the desired analysis; for agricultural productivity, for example, I create an exposure variable measuring the total number of exposure days in the five-month agricultural season, focusing only on the monsoon (Kharif) season.
-More days being located downwind from a high-pollution location leads to substantially higher average concentrations of particulate matter 2.5 (PM 2.5).\footnote{PM 2.5 is defined as particulate matter with a diameter of less than 2.5 micrometers.}
+More days being located downwind from a high-pollution location leads to substantially higher average concentrations of particulate matter 2.5 (PM 2.5), which is defined as particulate matter with a diameter of less than 2.5 micrometers.
 
@@
 
-This paper contributes to several strands of literature.
-We already have evidence of effects of pollution on different forms of labor activity.
+The results are consistent with evidence of effects of pollution on different forms of labor activity.
 For example, general levels of pollution decrease productivity of call center works in China (Chang et al.
@@
 Also in China, He, Liu, and Salvo (2019) find small but significant negative effects of pollution on manufacturing productivity, while Chen, Oliva, and Zhang (2022) show pollution leads to labor out migration.
-Relevant to my results, Hanna and Oliva (2015) show that decreases in pollution lead to increases in labor allocation in another developing country.
+Relevant to my results, Hanna and Oliva (2015) show that decreases in pollution lead to increases in labor allocation in Mexico.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-alvaredo2015recent}{}}%
-Alvaredo, Facundo, and Leonardo Gasparini. 2015. {``Recent Trends in Inequality and Poverty in Developing Countries.''} \emph{{Handbook of Income Distribution}} 2: 697--805.
-
 \leavevmode\vadjust pre{\hypertarget{ref-aragon2016polluting}{}}%

### 2023-07-20 analyses/3.yield.R

@@ -1174,3 +1174,6 @@ saveRDS(yieldtable, "pollution_development/draft/tables/yield7ivdiffindiff.rds")
 
 
 
+
+
+

### 2023-11-03 draft/draft.Rmd
slides

@@
   - \usepackage{makecell}
+
 ---
@@
   column_spec(c(2:5),width = "2cm") %>%
-  row_spec(11, hline_after = TRUE) %>%
   add_header_above(c(" ", "1991 census" = 2, "2001 census" = 2)) %>%

### 2023-11-03 analyses/3.yield.R

@@ -17,7 +17,6 @@ library(parallel)
 # setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
 # We want to get the wd up to the main folder
 # Need to go up two levels
-setwd("../")
 # Double check
 getwd()    # check
 

### 2023-11-03 analyses/4.nightlights.r

@@ -1,0 +1,215 @@
+# Purpose: This script estimates effects on nightlights
+# Author: Josh Merfeld
+# Date: February 25th, 2023
+
+rm(list = ls())
+
+library(tidyverse)
+library(lubridate)
+library(viridis)
+library(fixest)
+library(sf)
+library(openxlsx)
+library(parallel)
+
+
+# Cleaning ---------------------------------
+
+df <- read_csv("data/clean/ag_productivity/all_combined.csv")
+pollution <- read_csv(paste0("data/clean/pm25/ag_merge.csv"))
+df <- df %>% 
+        left_join(pollution, by = c("shrid", "year", "season"))
+rm(pollution)
+df <- df %>% group_by(shrid, season) %>%
+              mutate(
+                     yield = mean,
+                     rain_z = (log(rain_total + 1) - mean(log(rain_total + 1)))/sd(log(rain_total + 1)),
+                     rain_z2 = ((rain_total) - mean((rain_total)))/sd((rain_total)),
+                     wind_abs_dev = abs(days_sums - mean(days_sums)),
+                     pm25_abs_dev = abs(pm25 - mean(pm25)),
+                     wind = days_sums/10,
+                     pmavg1 = log(pm1),
+                     pmavg2 = log((pm1 + pm2)/2),
+                     pmavg3 = log((pm1 + pm2 + pm3)/3),
+                     pmavg4 = log((pm1 + pm2 + pm3 + pm4)/4),
+                     pmavg5 = log((pm1 + pm2 + pm3 + pm4 + pm5)/5),
+                     pm1 = log(pm1),
+                     pm2 = log(pm2),
+                     pm3 = log(pm3),
+                     pm4 = log(pm4),
+                     pm5 = log(pm5),
+                     windavg1 = m1,
+                     windavg2 = (m1 + m2)/2,
+                     windavg3 = (m1 + m2 + m3)/3,
+                     windavg4 = (m1 + m2 + m3 + m4)/4,
+                     windavg5 = (m1 + m2 + m3 + m4 + m5)/5,
+                     temp_m1 = (tmax_m1 + tmin_m1)/20,
+                     temp_m2 = (tmax_m2 + tmin_m2)/20,
+                     temp_m3 = (tmax_m3 + tmin_m3)/20,
+                     temp_m4 = (tmax_m4 + tmin_m4)/20,
[…]

### 2023-11-03 cleaning/2.cleaning_ntl.R

@@ -14,65 +14,42 @@ library(tidyverse)
 library(ncdf4)
 library(lubridate)
 library(R.utils)
-
+library(exactextractr)
 
 # Sets wd to folder this script is in so we can create relative paths instead of absolute paths
-setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
-# We want to get the wd up to the main folder
-# Need to go up two levels
-setwd("../../")
 # Double check
 getwd()    # check
 
 
 ### Load raw data ----------------------------------------------------------------------------------------------------------------------------------------------
-ntl <- as_tibble(read.csv("data/raw/shrug-v1.5.samosa-nl-csv/shrug_nl_wide.csv"))
-ntl <- ntl %>% select(-num_cells)
-
-# Gather, one variable at a time (we want long by year but wide by variables)
-ntl2 <- ntl %>% select(-c(starts_with("total_light_cal"), starts_with("max_"))) %>%
-                                gather(
-                                       "variable",
-                                       "ntl",
-                                       -shrid
-                                       )
-# Do some cleaning
-ntl2 <- ntl2 %>% mutate(
-                        year = substr(variable, str_length(variable) - 3, str_length(variable))
-                        ) %>%
-                  select(shrid, year, total_light = ntl)
-
-# Same thing for nezt var
-ntl3 <- ntl %>% select(c(starts_with("total_light_cal"), "shrid")) %>%
-                                gather(
-                                       "variable",
-                                       "ntl",
-                                       -shrid
-                                       )
-# Do some cleaning
-ntl3 <- ntl3 %>% mutate(
-                        year = substr(variable, str_length(variable) - 3, str_length(variable))
-                        ) %>%
-                  select(shrid, year, total_light_cal = ntl)
-
-# And the last one
-ntl4 <- ntl %>% select(c(starts_with("max"), "shrid")) %>%
-                                gather(
-                                       "variable",
-                                       "ntl",
-                                       -shrid
-                                       )
-# Do some cleaning
-ntl4 <- ntl4 %>% mutate(
-                        year = substr(variable, str_length(variable) - 3, str_length(variable))
-                        ) %>%
-                  select(shrid, year, max_light = ntl)
[…]

### 2023-11-29 draft/draft.Rmd
presentation

@@
 I also show that pollution from coal plants has a larger effect on agricultural productivity than other types of pollution.
-These results suggest that the negative externalities of pollution may hit the poorest particularly hard, given their reliance on agriculture.
+A back-of-the-envelope calculation indicates that aggregate output losses from pollution are substantial; 2011 total output would have been between 2.3 and 4.5 percent higher if pollution had remained at 2001 levels, depending on the crop.
   |
@@
 
-  result <- formatC(large, big.mark = ",")
+  result <- formatC(large, big.mark = ",", format = "f", drop0trailing = TRUE)
 
@@
 yieldricestatefe <- readRDS("pollution_development/draft/tables/yieldricestatefe.rds")
+monthlycoefs <- readRDS("pollution_development/draft/tables/monthcoefestimates.rds")
 
@@
 
+# load rastlist
+rastlist <- read_csv("data/counterfactual/rastlistclean.csv")
+yield <- read_csv("data/counterfactual/yieldclean.csv")
+dist <- read_sf("data/counterfactual/distclean.shp")
+# merge yield into dist
+dist <- dist %>% left_join(yield, by = c("dist_merge"))
+# counterfactual
+counterfactual <- read_csv("data/counterfactual/sums.csv")
+
 ```
@@
 # This chunk will allow us to add footers below figures
-library(ggplot2)
 
@@
 
+I carry out a simple back-of-the-envelope calculation to determine how much pollution affected agricultural output in the country as a whole in 2011.
+Using the most conservative estimate of the effect of pollution on agricultural productivity, I find that rice output would have been 2.3 percent higher, maize output would have been 4.5 percent higher, and wheat output would have been 3.8 percent higher in 2011 if pollution had remained at 2001 levels.
+
 An important caveat to these results is that I cannot disentangle exactly which pollutants drive results.
@@
 colnames(plantresultstable) <- c("1991", "2001", "2001", "2011")
-rownames(plantresultstable) <- c("Population (log)", "", "Area (log)", "", "Urban pop (prop)", "", "Ag. productivity", "", "Pollution (log PM 2.5)", "",
-                                "Sub-sample", "Observations")
+rownames(plantresultstable) <- c("Population (log)", "", "Area (log)", "", "Ag. productivity", "", "Pollution (log PM 2.5)", "", "Has plant?", "",
+                                "Observations")
 kable(    
@@
 
-rownames(yield1reducedform) <- c("wind days", " ", 
+rownames(yield1reducedform) <- c("wind days (10s)", " ", 
                                   "weather", "weather (expanded)", "weather (expanded, bins)",
@@
   column_spec(1, width = "4cm") %>%
-  column_spec(c(2:4),width = "2.5cm") %>%
+  column_spec(c(2:5),width = "2.5cm") %>%
   row_spec(6, bold = TRUE) %>%
@@
 To put the size of the coefficient in context, it is worth digging a little deeper into the exposure variable.
-The mean within-village absolute deviation in wind exposure is approximately 2.42 days, meaning that the average change in agricultural productivity from year-to-year due to nothing but changes in wind patterns carrying particulate matter is around 1.5 percent relative to the mean.
-This implies that swings in agricultural productivity -- due to wind direction -- of several percent are probably quite common, since the absolute deviation includes deviations above and below the mean.
-
-```{r yieldtablepm, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-# Table
-colnames(yield2naive) <- c("(1)", "(2)", "(3)", "(4)")
-rownames(yield2naive) <- c("Particulate matter (log PM 2.5)", "", 
-                           "Weather", "Weather (expanded)", "weather (expanded, bins)",
-                           "Fixed effects:", "Village", "Year",
-                           "Observations")
-kable(
-      yield2naive, caption = "Particulate matter (PM 2.5) and agricultural productivity",
-      align = "cccc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the village level.
-The outcome in all regressions is estimated agricultural yield, defined as tons per hectare.
-* p<0.10 ** p<0.05 *** p<0.01",
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
-           ) %>%
-  column_spec(1, width = "5cm") %>%
-  column_spec(c(2:4),width = "2cm") %>%
-  row_spec(6, bold = TRUE) %>%
-  row_spec(8, hline_after = TRUE) %>%
-  kable_classic_2()
-
-```
+Wind is scaled by 10 for presentation, meaning that an increase in *exposure* for a village from its mean leads to a 0.5 percent drop in agricultural productivity.
+The standard deviation of the within-village change in exposure to wind from polluted areas is 4.35, meaning a one-standard-deviation increase from the mean leads to a 0.22 percent decrease in agricultural productivity due to nothing but being located downwind from a pollution source.
 
@@
 Instead, we are actually interested in better understanding the effects of exposure to pollution on agricultural productivity.
-Table \@ref(tab:yieldtablepm) presents the same regressions as Table \@ref(tab:yieldtable), with a key difference: instead of wind exposure, it uses log average concentration of $\mathrm{PM 2.5}$ -- from @hammer2020global -- during the season as the key explanatory variable.
+Table \@ref(tab:yieldtablepm) in the appendix presents the same regressions as Table \@ref(tab:yieldtable), with a key difference: instead of wind exposure, it uses log average concentration of $\mathrm{PM 2.5}$ -- from @hammer2020global -- during the season as the key explanatory variable.
 There is again a consistent, negative relationship between particulate matter and agricultural productivity.
@@
 The first-stage results -- the full results are in Table \@ref(tab:yield3ivmainfirststage) of the appendix -- make clear how strong of an instrument wind is; all three F-statistics are above 1,000, meeting even the higher bar for the first-stage F-test argued for by @lee2022valid.
+The coefficient on wind in the first stage is stable at around 0.0085 across specifications.
 
@@
 
-```{r crops, echo = FALSE, message = FALSE, warning = FALSE, dev='png', fig.height = 5, fig.cap = "Most common crop by village", fig.footer = list("Note: The figure shows the most common crop grown in each village, by area harvested.")}
-
-village_shp <- village_shp %>% left_join(crops_all, by = c("shrid"))
-
-# plot
-ggplot() +
-  geom_sf(data = village_shp, aes(fill = most_common), color = NA) +
-  scale_fill_brewer("Most common crop\nby area harvested", palette = "Set3") +
-  geom_sf(data = india_shape2, fill = NA, lwd = 0.01) +
-  theme_minimal()
-
-```
-
 ## Crops and labor intensity
@@
 
-Figure \@ref(fig:crops) shows the distribution of the most common crop across the villages in my sample.
+Figure \@ref(fig:crops) in the appendix shows the distribution of the most common crop across the villages in my sample.
 Though there are large areas where rice is the most commonly grown crop, there is also quite a bit of overlap, especially in the middle parts of the country.
@@
 
+```{r yield8monthly, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Effects of pollution on productivity by month"}
+
+ggplot() +
+  # first add the coefficient estimates
+  geom_point(aes(x = 1:5, y = monthlycoefs$coefs)) +
+  # now CIs
+  geom_linerange(aes(x = 1:5, ymin = monthlycoefs$confint[,1], ymax = monthlycoefs$confint[,2])) +
+  theme_minimal() +
+  labs(x = "Month of the ag. season", y = "Coefficient estimate of log pm 2.5")
+
+```
+
+Another way to tease out differential effects on labor productivity, specifically, is to look at the effects of pollution by month.
+If pollution is working through labor, then we would expect to see larger effects when labor is relatively more important than.
+The harvest season, the fifth month of the agricultural season in the data, stands out as the most likely candidate for two reasons.
+First, labor allocation is generally quite on plots during the harvest season.
+Second, pollution at the very end of the season is unlikely to have large effects on overall land producitivty, as opposed to pollution in prior months.
+
+Figure \@ref(yield8monthly) presents the coefficients of the IV regression separately for each month of the agricultural season, with the very beginning of the season on the left side of the graph and the harvest season on the right side of the graph.
+Interestingly, the coefficient for the fourth month of the season is actually positive.
+It is hard to know exactly why this is the case, but it could be an interaction between the positive effects of coal pollution on agricultural productivity found by @sanders2022adaptation and the fact that very little labor is used just prior to harvest.
+However, the overall explanation is beyond the scope of the current paper.
+
+The other four months all see negative effects of pollution on agricultural productivity.
+However, the effect during the harvest season is the largest, at more than twice the magnitude of any of the other three months.
+In other words, it seems that pollution has the largest effects during harvest, which is consistent with the hypothesis that pollution affects labor productivity.
+This is also consistent with the results in Table \@ref(tab:yieldrice), which show that pollution has larger effects in rice-growing areas, which are more labor intensive.
+
 ## Coal plants or overall pollution?
@@
 Importantly, since these results come from an IV, both coefficients denote the change in agricultural yield due to an increase in pollution by the same units.
-In other words, this is not identifying differences in overall pollution, per se, but rather differences in the effects of an increase of pollution by the same percentage.
+In other words, this is not identifying differences in overall pollution, per se, but rather differences in the effects of an increase of pollution by the same amount.
 These different effects could be due to, for example, differences in emission patterns for different kinds of plants and factories; coal plants, for example, are responsible for much more mercury than other kinds of polluting industries.^[https://www.epa.gov/sites/default/files/2016-12/documents/nei2014v1_tsd.pdf]
+
+# Quantifying the overall effects of pollution on agricultural productivity
+
+Pollution has increased markedly in India over the last few decades.
+We can use the results above to quantify the overall effects of pollution on agricultural productivity.
+However, there are two reasons just using the sample of villages from the main results above might not be indicative of overall changes in India.
+First, the sample of villages have higher levels of pollution than other villages, as shown in Table \@ref(tab:plantresultstable).
+Second, the sample of villages are also located near coal plants, which appear to have larger negative effects on agricultural productivity than other sources of pollution.
+
+For this reason, I instead turn to data from ICRISAT,^[http://data.icrisat.org/dld] which has district-level total production data from different years.
+To track with the timing of the main data used above, I pull total agricultural production from 2001 and 2011 for three of the most common crops in India: rice, wheat, and maize.
+To simplify the calculation, I assume total area devoted to each crop is unchanged across the decade.
+I then merge this with the pollution data from @hammer2020global using a 2011 district-level census shapefile, which has identifiers that match the ICRISAT data.
+Figure \@ref(fig:yielddatamaps) in the appendix shows the distribution of total output for the three crops in 2001.
+The data covers much, but not all, of the country.
+
+In the output data, I observe values for 2001 and 2011.
+I can calculate the overall change in log output across the two years, as well as the overall change in pollution levels.
+I use the most conservative point estimate *without* coal plants in \@ref(tab:yield7ivdiffindiff), which is $-0.223$.
+For each district, I calculate the counterfactual for each crop as follows:
+\begin{gather} \label{eq:counterfactual} \Delta log(\mathrm{counterfactual})_{d}-0.223\times\Delta log(PM2.5)_{d} = \Delta log(\mathrm{actual})_{d}, \end{gather}
+
+where $\mathrm{counterfactual}$ is the value I want to calculate, while $\mathrm{actual}$ is the actual observed change in output.
+I invert the equation to find the predicted counterfactual without the change in pollution.
+In other words, I calculate what the change would have been if pollution had not changed at all for any given district.
+Since this relies on log changes, I restrict the sample only to districts that have output data for a given crop in 2001.
+Due to a few very large changes in pollution in a few areas, I winsorize log counterfactual changes at the 95th percentile for all crops.
+After calculating the counterfactual for each district separately, I sum total output to the country level in order to judge the overall effects of pollution on agricultural productivity.
+
+The overall change depends on the distribution of changes in pollution as well as the distribution of output across the three crops.
+I first present the actual and predicted changes in (log) output in Figures \@ref(fig:yield9counterfactual1), \@ref(fig:yield9counterfactual2), and \@ref(fig:yield9counterfactual3) in the appendix.
+Since pollution increased almost everywhere, the counterfactual change is (almost) always higher than the actual change.
+
+I present the main results of the back-of-the-envelope calculation in Table \@ref(tab:counterfactualsums).
+I align the three crops across the rows, with the 2001 and 2011 total output (summed across all districts, in '000s of tons) in the first two columns.
+Total actual output increased for all three crops, with the largest absolute increase being for wheat -- which increased by around 37 percent in relative terms -- and the largest relative increase being for maize -- which increased by around 60 percent in total output.
+
+The key set of results is the comparison betweens column two and column three.
+Column two gives the actual total output by crop in 2011, while column three presents the counterfactual output, calculated as what would have happened if pollution had not changed at all.
+There is quite a bit of heterogeneity across crops.
+For rice, total output would have been around 2.3 percent higher if pollution had remained at 2001 levels.
+Maize, on the other hand, would have seen a 4.2 percent increase and wheat would have seen a 3.8 percent increase.
+These results underscore the point that the effects of pollution are not uniform across crops, even if we assume equal effects of pollution, due to differences in the distribution of pollution changes across the country.
+
+```{r counterfactualsums, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+options(scipen = 999)
+# sumtables
+sumtable <- matrix(NA, nrow = 3, ncol = 5)
+colnames(sumtable) <- c("", "2001 actual output", "2011 actual output", "2011 counterfactual", "Difference (pct.)")
+sumtable[,1] <- c("Rice", "Wheat", "Maize")
+sumtable[,2] <- round(as_vector(counterfactual[1,1:3]), 0)
+sumtable[,3] <- round(as_vector(counterfactual[1,4:6]), 0)
+sumtable[,4] <- round(as_vector(counterfactual[1,7:9]), 0)
+sumtable[,5] <- round((as_vector(counterfactual[1,10:12]) - 1)*100, 1)
+for (col in 2:4){
+  sumtable[1, col] <- commaNum(as.numeric(sumtable[1, col]))
+  sumtable[2, col] <- commaNum(as.numeric(sumtable[2, col]))
+  sumtable[3, col] <- commaNum(as.numeric(sumtable[3, col]))
+}
+
+kable(
+      sumtable, caption = "Change in output and counterfactual, all districts",
+      align = "lcccc", booktabs = TRUE, linesep = "", row.names = FALSE
+      ) %>%
+  footnote(
+           "Note: Output is the total sum across all districts, in '000s of tons. The counterfactual is calculated as described in Equation 5.", 
+           general_title = "",
+           threeparttable = TRUE,
+           footnote_as_chunk = TRUE
+           ) %>%
+  column_spec(1, width = "2cm") %>%
+  column_spec(c(2:5),width = "2cm") %>%
+  kable_classic_2()
+
+```
 
@@
 ```
+
 ```{r data, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
@@
 
+```{r yieldtablepm, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+
+# Table
+colnames(yield2naive) <- c("(1)", "(2)", "(3)", "(4)")
+rownames(yield2naive) <- c("Particulate matter (log PM 2.5)", "", 
+                           "Weather", "Weather (expanded)", "weather (expanded, bins)",
+                           "Fixed effects:", "Village", "Year",
+                           "Observations")
+kable(
+      yield2naive, caption = "Particulate matter (PM 2.5) and agricultural productivity",
+      align = "cccc", booktabs = TRUE, linesep = ""
+      ) %>%
+  footnote(
+           "Note: Standard errors are in parentheses and are clustered at the village level.
+The outcome in all regressions is estimated agricultural yield, defined as tons per hectare.
+* p<0.10 ** p<0.05 *** p<0.01",
+           general_title = "",
+           threeparttable = TRUE,
+           footnote_as_chunk = TRUE
+           ) %>%
+  column_spec(1, width = "5cm") %>%
+  column_spec(c(2:5),width = "2cm") %>%
+  row_spec(6, bold = TRUE) %>%
+  row_spec(8, hline_after = TRUE) %>%
+  kable_classic_2()
+
+```
+
+```{=tex}
+\FloatBarrier
+\newpage
+```
+
 ```{r yield3ivmainfirststage, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
@@
 
-```{=tex}
-\FloatBarrier
-\newpage
-```
-```{r yield8cumulative, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-
-# Table
-colnames(yield8cumulative) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
-kable(
-      yield8cumulative, caption = "Cumulative exposure to pollution during the season",
-      align = "ccccc", booktabs = TRUE, linesep = ""
-      ) %>%
-  footnote(
-           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
-           general_title = "",
-           threeparttable = TRUE,
-           footnote_as_chunk = TRUE
-           ) %>%
-  column_spec(1, width = "2.25cm") %>%
-  column_spec(c(2:6),width = "2cm") %>%
-  row_spec(c(12), bold = TRUE) %>%
-  row_spec(14, hline_after = TRUE) %>%
-  kable_classic_2()
+```{r crops, echo = FALSE, message = FALSE, warning = FALSE, dev='png', fig.height = 5, fig.cap = "Most common crop by village", fig.footer = list("Note: The figure shows the most common crop grown in each village, by area harvested.")}
+
+village_shp <- village_shp %>% left_join(crops_all, by = c("shrid"))
+
+# plot
+ggplot() +
+  geom_sf(data = village_shp, aes(fill = most_common), color = NA) +
+  scale_fill_brewer("Most common crop\nby area harvested", palette = "Set3") +
+  geom_sf(data = india_shape2, fill = NA, lwd = 0.01) +
+  theme_minimal()
 
@@
 ```
+
+```{=tex}
+\FloatBarrier
+\newpage
+```
+
+```{r yielddatamaps, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "The maps show the district-level yields of each crop in 2001.", fig.footer = "Yield data comes from ICRISAT."}
+
+g1 <- ggplot() +
+        geom_sf(data = dist, color = NA) +
+        geom_sf(data = dist %>% filter(is.na(log(riceproduction))==F), aes(fill = log(riceproduction)), color = NA) +
+        scale_fill_viridis_c("log ('000 tons)", option = "viridis") +
+        theme_minimal() +
+        labs(title = "Rice")
+g2 <- ggplot(dist) +
+        geom_sf(data = dist, color = NA) +
+        geom_sf(data = dist %>% filter(is.na(log(wheatproduction))==F), aes(fill = log(wheatproduction)), color = NA) +
+        scale_fill_viridis_c("log ('000 tons)", option = "viridis") +
+        theme_minimal() +
+        labs(title = "Wheat")
+g3 <- ggplot(dist) +
+        geom_sf(data = dist, color = NA) +
+        geom_sf(data = dist %>% filter(is.na(log(maizeproduction))==F), aes(fill = log(maizeproduction)), color = NA) +
+        scale_fill_viridis_c("log ('000 tons)", option = "viridis") +
+        theme_minimal() +
+        labs(title = "Maize")
+ggarrange(g1, g2, g3, legend = "bottom", common.legend = TRUE)
+
+```
+
+```{=tex}
+\FloatBarrier
+\newpage
+```
+
+\begin{landscape}
+```{r yield9counterfactual1, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Distribution of actual and counterfactual log changes in output, rice", dev = "pdf", fig.width = 8}
+
+ggplot(yield) +
+        geom_density(aes(x = rice_change, fill = "Actual"), color = "transparent", alpha = 0.5) +
+        geom_density(aes(x = ricecounterfactual, fill = "Counterfactual"), color = "transparent", alpha = 0.5) +
+        theme_minimal() +
+        labs(x = "Change in log output", y = "Density", title = "Rice") +
+        scale_fill_manual(name = "Output", values = c("Actual" = "orange", "Counterfactual" = "navy"))
+
+```
+\end{landscape}
+
+```{=tex}
+\FloatBarrier
+\newpage
+```
+
+\begin{landscape}
+```{r yield9counterfactual2, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Distribution of actual and counterfactual log changes in output, wheat", dev = "pdf", fig.width = 8}
+
+ggplot(yield) +
+        geom_density(aes(x = wheat_change, fill = "Actual"), color = "transparent", alpha = 0.5) +
+        geom_density(aes(x = wheatcounterfactual, fill = "Counterfactual"), color = "transparent", alpha = 0.5) +
+        theme_minimal() +
+        labs(x = "Change in log output", y = "Density", title = "Wheat") +
+        scale_fill_manual(name = "Output", values = c("Actual" = "orange", "Counterfactual" = "navy"))
+
+```
+\end{landscape}
+
+```{=tex}
+\FloatBarrier
+\newpage
+```
+
+\begin{landscape}
+```{r yield9counterfactual3, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Distribution of actual and counterfactual log changes in output, maize", dev = "pdf", fig.width = 8}
+
+ggplot(yield) +
+        geom_density(aes(x = maize_change, fill = "Actual"), color = "transparent", alpha = 0.5) +
+        geom_density(aes(x = maizecounterfactual, fill = "Counterfactual"), color = "transparent", alpha = 0.5) +
+        theme_minimal() +
+        labs(x = "Change in log output", y = "Density", title = "Maize") +
+        scale_fill_manual(name = "Output", values = c("Actual" = "orange", "Counterfactual" = "navy"))
+
+```
+\end{landscape}

### 2023-11-29 draft/draft.tex
presentation

@@
 \author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
-\date{2023-07-19}
+\date{2023-11-23}
 
@@
 I also show that pollution from coal plants has a larger effect on agricultural productivity than other types of pollution.
-These results suggest that the negative externalities of pollution may hit the poorest particularly hard, given their reliance on agriculture.\\
+A back-of-the-envelope calculation indicates that aggregate output losses from pollution are substantial; 2011 total output would have been between 2.3 and 4.5 percent higher if pollution had remained at 2001 levels, depending on the crop.\\
 \strut \\
@@
 This is consistent with coal emissions being quite different from other forms of pollution; coal emissions contain different types and quantities of harmful substances, meaning effects may differ greatly.
+
+I carry out a simple back-of-the-envelope calculation to determine how much pollution affected agricultural output in the country as a whole in 2011.
+Using the most conservative estimate of the effect of pollution on agricultural productivity, I find that rice output would have been 2.3 percent higher, maize output would have been 4.5 percent higher, and wheat output would have been 3.8 percent higher in 2011 if pollution had remained at 2001 levels.
 
@@
 \midrule
-Population (log) & 0.012*** & 0.009*** & 0.015*** & 0.006***\\
- & (0.001) & (0.000) & (0.000) & \vphantom{1} (0.000)\\
-Area (log) & -0.010*** & -0.005*** & -0.010*** & -0.005***\\
- & (0.001) & (0.000) & (0.000) & (0.000)\\
-Urban pop (prop) & -0.029 & 0.006 & -0.003 & -0.007\\
- & (0.037) & (0.035) & (0.022) & (0.018)\\
-Ag. productivity &  &  & 0.014*** & 0.028***\\
- &  &  & (0.001) & (0.001)\\
-Pollution (log PM 2.5) &  &  & 0.051*** & 0.016***\\
- &  &  & (0.002) & (0.001)\\
-Sub-sample & all & no plant & all & no plant\\
-\midrule
-Observations & 283,021 & 264,265 & 508,786 & 469,005\\
+Population (log) & -0.107 & 0.291*** & 0.241*** & 0.128\\
+ & (0.097) & (0.083) & (0.079) & (0.080)\\
+Area (log) & 0.240*** & -0.110 & 0.133* & 0.119\\
+ & (0.065) & (0.088) & (0.079) & (0.085)\\
+Ag. productivity &  &  & 0.092 & 0.121\\
+ &  &  & (0.070) & (0.097)\\
+Pollution (log PM 2.5) &  &  & 0.612*** & 0.530**\\
+ &  &  & (0.201) & (0.256)\\
+Has plant? &  & 14.842*** &  & 13.574***\\
+ &  & (0.568) &  & (0.158)\\
+Observations & 209,110 & 209,110 & 408,913 & 415,334\\
 \bottomrule
@@
 \begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}c}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}}
 \toprule
@@
 \midrule
-wind days & -0.0078*** & -0.0065*** & -0.0051*** & -0.0053***\\
+wind days (10s) & -0.0078*** & -0.0065*** & -0.0051*** & -0.0053***\\
  & (0.0007) & (0.0007) & (0.0007) & (0.0007)\\
@@
 To put the size of the coefficient in context, it is worth digging a little deeper into the exposure variable.
-The mean within-village absolute deviation in wind exposure is approximately 2.42 days, meaning that the average change in agricultural productivity from year-to-year due to nothing but changes in wind patterns carrying particulate matter is around 1.5 percent relative to the mean.
-This implies that swings in agricultural productivity -- due to wind direction -- of several percent are probably quite common, since the absolute deviation includes deviations above and below the mean.
-
-\begin{table}
-
-\caption{\label{tab:yieldtablepm}Particulate matter (PM 2.5) and agricultural productivity}
-\centering
-\begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{5cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}c}
-\toprule
-  & (1) & (2) & (3) & (4)\\
-\midrule
-Particulate matter (log PM 2.5) & -0.042*** & -0.045*** & -0.008** & -0.041***\\
- & (0.003) & (0.003) & (0.003) & (0.003)\\
-Weather & No & Yes & Yes & No\\
-Weather (expanded) & No & No & Yes & No\\
-weather (expanded, bins) & No & No & No & Yes\\
-\textbf{Fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-Village & Yes & Yes & Yes & Yes\\
-Year & Yes & Yes & Yes & Yes\\
-\midrule
-Observations & 1,266,922 & 1,266,922 & 1,266,922 & 1,266,922\\
-\bottomrule
-\end{tabular}
-\begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the village level.
-The outcome in all regressions is estimated agricultural yield, defined as tons per hectare.
-* p<0.10 ** p<0.05 *** p<0.01
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
+Wind is scaled by 10 for presentation, meaning that an increase in \emph{exposure} for a village from its mean leads to a 0.5 percent drop in agricultural productivity.
+The standard deviation of the within-village change in exposure to wind from polluted areas is 4.35, meaning a one-standard-deviation increase from the mean leads to a 0.22 percent decrease in agricultural productivity due to nothing but being located downwind from a pollution source.
 
@@
 Instead, we are actually interested in better understanding the effects of exposure to pollution on agricultural productivity.
-Table \ref{tab:yieldtablepm} presents the same regressions as Table \ref{tab:yieldtable}, with a key difference: instead of wind exposure, it uses log average concentration of \(\mathrm{PM 2.5}\) -- from Hammer et al.
+Table \ref{tab:yieldtablepm} in the appendix presents the same regressions as Table \ref{tab:yieldtable}, with a key difference: instead of wind exposure, it uses log average concentration of \(\mathrm{PM 2.5}\) -- from Hammer et al.
 (2020) -- during the season as the key explanatory variable.
@@
 (2022).
+The coefficient on wind in the first stage is stable at around 0.0085 across specifications.
 
@@
 
-\begin{figure}
-\includegraphics{draft_files/figure-latex/crops-1} \caption[Most common crop by village]{Most common crop by village}\label{fig:crops}\floatfoot*{Note: The figure shows the most common crop grown in each village, by area harvested.}
-\end{figure}
-
 \hypertarget{crops-and-labor-intensity}{%
@@
 
-Figure \ref{fig:crops} shows the distribution of the most common crop across the villages in my sample.
+Figure \ref{fig:crops} in the appendix shows the distribution of the most common crop across the villages in my sample.
 Though there are large areas where rice is the most commonly grown crop, there is also quite a bit of overlap, especially in the middle parts of the country.
@@
 
+\begin{figure}
+\includegraphics{draft_files/figure-latex/yield8monthly-1} \caption[Effects of pollution on productivity by month]{Effects of pollution on productivity by month}\label{fig:yield8monthly}
+\end{figure}
+
+Another way to tease out differential effects on labor productivity, specifically, is to look at the effects of pollution by month.
+If pollution is working through labor, then we would expect to see larger effects when labor is relatively more important than.
+The harvest season, the fifth month of the agricultural season in the data, stands out as the most likely candidate for two reasons.
+First, labor allocation is generally quite on plots during the harvest season.
+Second, pollution at the very end of the season is unlikely to have large effects on overall land producitivty, as opposed to pollution in prior months.
+
+Figure \ref{yield8monthly} presents the coefficients of the IV regression separately for each month of the agricultural season, with the very beginning of the season on the left side of the graph and the harvest season on the right side of the graph.
+Interestingly, the coefficient for the fourth month of the season is actually positive.
+It is hard to know exactly why this is the case, but it could be an interaction between the positive effects of coal pollution on agricultural productivity found by Sanders and Barreca (2022) and the fact that very little labor is used just prior to harvest.
+However, the overall explanation is beyond the scope of the current paper.
+
+The other four months all see negative effects of pollution on agricultural productivity.
+However, the effect during the harvest season is the largest, at more than twice the magnitude of any of the other three months.
+In other words, it seems that pollution has the largest effects during harvest, which is consistent with the hypothesis that pollution affects labor productivity.
+This is also consistent with the results in Table \ref{tab:yieldrice}, which show that pollution has larger effects in rice-growing areas, which are more labor intensive.
+
 \hypertarget{coal-plants-or-overall-pollution}{%
@@
 Importantly, since these results come from an IV, both coefficients denote the change in agricultural yield due to an increase in pollution by the same units.
-In other words, this is not identifying differences in overall pollution, per se, but rather differences in the effects of an increase of pollution by the same percentage.
+In other words, this is not identifying differences in overall pollution, per se, but rather differences in the effects of an increase of pollution by the same amount.
 These different effects could be due to, for example, differences in emission patterns for different kinds of plants and factories; coal plants, for example, are responsible for much more mercury than other kinds of polluting industries.\footnote{\url{https://www.epa.gov/sites/default/files/2016-12/documents/nei2014v1_tsd.pdf}}
+
+\hypertarget{quantifying-the-overall-effects-of-pollution-on-agricultural-productivity}{%
+\section{Quantifying the overall effects of pollution on agricultural productivity}\label{quantifying-the-overall-effects-of-pollution-on-agricultural-productivity}}
+
+Pollution has increased markedly in India over the last few decades.
+We can use the results above to quantify the overall effects of pollution on agricultural productivity.
+However, there are two reasons just using the sample of villages from the main results above might not be indicative of overall changes in India.
+First, the sample of villages have higher levels of pollution than other villages, as shown in Table \ref{tab:plantresultstable}.
+Second, the sample of villages are also located near coal plants, which appear to have larger negative effects on agricultural productivity than other sources of pollution.
+
+For this reason, I instead turn to data from ICRISAT,\footnote{\url{http://data.icrisat.org/dld}} which has district-level total production data from different years.
+To track with the timing of the main data used above, I pull total agricultural production from 2001 and 2011 for three of the most common crops in India: rice, wheat, and maize.
+To simplify the calculation, I assume total area devoted to each crop is unchanged across the decade.
+I then merge this with the pollution data from Hammer et al.
+(2020) using a 2011 district-level census shapefile, which has identifiers that match the ICRISAT data.
+Figure \ref{fig:yielddatamaps} in the appendix shows the distribution of total output for the three crops in 2001.
+The data covers much, but not all, of the country.
+
+In the output data, I observe values for 2001 and 2011.
+I can calculate the overall change in log output across the two years, as well as the overall change in pollution levels.
+I use the most conservative point estimate \emph{without} coal plants in \ref{tab:yield7ivdiffindiff}, which is \(-0.223\).
+For each district, I calculate the counterfactual for each crop as follows:
+\begin{gather} \label{eq:counterfactual} \Delta log(\mathrm{counterfactual})_{d}-0.223\times\Delta log(PM2.5)_{d} = \Delta log(\mathrm{actual})_{d}, \end{gather}
+
+where \(\mathrm{counterfactual}\) is the value I want to calculate, while \(\mathrm{actual}\) is the actual observed change in output.
+I invert the equation to find the predicted counterfactual without the change in pollution.
+In other words, I calculate what the change would have been if pollution had not changed at all for any given district.
+Since this relies on log changes, I restrict the sample only to districts that have output data for a given crop in 2001.
+Due to a few very large changes in pollution in a few areas, I winsorize log counterfactual changes at the 95th percentile for all crops.
+After calculating the counterfactual for each district separately, I sum total output to the country level in order to judge the overall effects of pollution on agricultural productivity.
+
+The overall change depends on the distribution of changes in pollution as well as the distribution of output across the three crops.
+I first present the actual and predicted changes in (log) output in Figures \ref{fig:yield9counterfactual1}, \ref{fig:yield9counterfactual2}, and \ref{fig:yield9counterfactual3} in the appendix.
+Since pollution increased almost everywhere, the counterfactual change is (almost) always higher than the actual change.
+
+I present the main results of the back-of-the-envelope calculation in Table \ref{tab:counterfactualsums}.
+I align the three crops across the rows, with the 2001 and 2011 total output (summed across all districts, in '000s of tons) in the first two columns.
+Total actual output increased for all three crops, with the largest absolute increase being for wheat -- which increased by around 37 percent in relative terms -- and the largest relative increase being for maize -- which increased by around 60 percent in total output.
+
+The key set of results is the comparison betweens column two and column three.
+Column two gives the actual total output by crop in 2011, while column three presents the counterfactual output, calculated as what would have happened if pollution had not changed at all.
+There is quite a bit of heterogeneity across crops.
+For rice, total output would have been around 2.3 percent higher if pollution had remained at 2001 levels.
+Maize, on the other hand, would have seen a 4.2 percent increase and wheat would have seen a 3.8 percent increase.
+These results underscore the point that the effects of pollution are not uniform across crops, even if we assume equal effects of pollution, due to differences in the distribution of pollution changes across the country.
+
+\begin{table}
+
+\caption{\label{tab:counterfactualsums}Change in output and counterfactual, all districts}
+\centering
+\begin{threeparttable}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\toprule
+ & 2001 actual output & 2011 actual output & 2011 counterfactual & Difference (pct.)\\
+\midrule
+Rice & 79,962 & 93,238 & 95,405 & 2.3\\
+Wheat & 58,780 & 93,975 & 97,520 & 3.8\\
+Maize & 11,935 & 20,897 & 21,832 & 4.5\\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}[para]
+\item Note: Output is the total sum across all districts, in '000s of tons. The counterfactual is calculated as described in Equation 5.
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
 
@@
 
+\caption{\label{tab:yieldtablepm}Particulate matter (PM 2.5) and agricultural productivity}
+\centering
+\begin{threeparttable}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{5cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
+\toprule
+  & (1) & (2) & (3) & (4)\\
+\midrule
+Particulate matter (log PM 2.5) & -0.042*** & -0.045*** & -0.008** & -0.041***\\
+ & (0.003) & (0.003) & (0.003) & (0.003)\\
+Weather & No & Yes & Yes & No\\
+Weather (expanded) & No & No & Yes & No\\
+weather (expanded, bins) & No & No & No & Yes\\
+\textbf{Fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+Village & Yes & Yes & Yes & Yes\\
+Year & Yes & Yes & Yes & Yes\\
+\midrule
+Observations & 1,266,922 & 1,266,922 & 1,266,922 & 1,266,922\\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}[para]
+\item Note: Standard errors are in parentheses and are clustered at the village level.
+The outcome in all regressions is estimated agricultural yield, defined as tons per hectare.
+* p<0.10 ** p<0.05 *** p<0.01
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+
+\FloatBarrier
+\newpage
+
+\begin{table}
+
 \caption{\label{tab:yield3ivmainfirststage}Wind direction and particulate matter}
@@
 
-\FloatBarrier
-\newpage
-
-\begin{table}
-
-\caption{\label{tab:yield8cumulative}Cumulative exposure to pollution during the season}
-\centering
-\begin{threeparttable}
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2.25cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
-\toprule
-  & (1) & (2) & (3) & (4) & (5)\\
-\midrule
-June & -0.643*** &  &  &  & \\
- & (0.064) &  &  &  & \\
-July &  & -0.658*** &  &  & \\
- &  & (0.045) &  &  & \\
-Aug. &  &  & -0.531*** &  & \\
- &  &  & (0.049) &  & \\
-Sept. &  &  &  & -0.672*** & \\
- &  &  &  & (0.047) & \\
-Oct. &  &  &  &  & -0.943***\\
- &  &  &  &  & (0.069)\\
-weather & Yes & Yes & Yes & Yes & Yes\\
-\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-village & Yes & Yes & Yes & Yes & Yes\\
-year & Yes & Yes & Yes & Yes & Yes\\
-\midrule
-F & 1,327 & 2,830 & 2,933 & 2,413 & 1,515\\
-observations & 1,266,922 & 1,266,922 & 1,266,922 & 1,266,922 & 1,266,922\\
-\bottomrule
-\end{tabular}
-\begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
+\begin{figure}
+\includegraphics{draft_files/figure-latex/crops-1} \caption[Most common crop by village]{Most common crop by village}\label{fig:crops}\floatfoot*{Note: The figure shows the most common crop grown in each village, by area harvested.}
+\end{figure}
 
@@
 
+\FloatBarrier
+\newpage
+
+\begin{figure}
+\includegraphics{draft_files/figure-latex/yielddatamaps-1} \caption[The maps show the district-level yields of each crop in 2001]{The maps show the district-level yields of each crop in 2001.}\label{fig:yielddatamaps}\floatfoot*{Yield data comes from ICRISAT.}
+\end{figure}
+
+\FloatBarrier
+\newpage
+
+\begin{landscape}
+\begin{figure}
+\includegraphics{draft_files/figure-latex/yield9counterfactual1-1} \caption[Distribution of actual and counterfactual log changes in output, rice]{Distribution of actual and counterfactual log changes in output, rice}\label{fig:yield9counterfactual1}
+\end{figure}
+\end{landscape}
+
+\FloatBarrier
+\newpage
+
+\begin{landscape}
+\begin{figure}
+\includegraphics{draft_files/figure-latex/yield9counterfactual2-1} \caption[Distribution of actual and counterfactual log changes in output, wheat]{Distribution of actual and counterfactual log changes in output, wheat}\label{fig:yield9counterfactual2}
+\end{figure}
+\end{landscape}
+
+\FloatBarrier
+\newpage
+
+\begin{landscape}
+\begin{figure}
+\includegraphics{draft_files/figure-latex/yield9counterfactual3-1} \caption[Distribution of actual and counterfactual log changes in output, maize]{Distribution of actual and counterfactual log changes in output, maize}\label{fig:yield9counterfactual3}
+\end{figure}
+\end{landscape}
+
 \end{document}

### 2023-11-29 analyses/3.yield.R

@@ -11,7 +11,7 @@ library(fixest)
 library(sf)
 library(openxlsx)
 library(parallel)
-
+library(pbmcapply)
 
 # Sets wd to folder this script is in so we can create relative paths instead of absolute paths
 # setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
@@ -71,7 +71,8 @@ df <- df %>% group_by(shrid, season) %>%
                      rain4_zbin0 = as.numeric(rain4_z>-1 & rain4_z<1),
                      rain4_zbin1 = as.numeric(rain4_z>=1),
                      rain5_zbin0 = as.numeric(rain5_z>-1 & rain5_z<1),
-                     rain5_zbin1 = as.numeric(rain5_z>=1)
+                     rain5_zbin1 = as.numeric(rain5_z>=1),
+                     state = substr(shrid, 1, 2)
                      )
 
 
@@ -83,7 +84,7 @@ df <- df %>% group_by(shrid, season) %>%
 
 
 # And district identifiers
-villages_overlap <- read_sf("data/spatial/villages_overlap/villages_overlap.shp")
+villages_overlap <- read_sf("data/spatial/villages_overlap/villages_overlap.shp") %>% st_set_crs(st_crs("EPSG:24378"))
 villages_overlap <- villages_overlap %>% 
                       mutate(
                              shrid = paste0(pc11_s_id, "-", pc11_tv_id),
@@ -143,13 +144,20 @@ rm(plant_data)
 # MONSOON ONLY
 df <- df %>%
       filter(season=="monsoon")
+
+
+
+
+
+
 summary(df$wind_abs_dev)
-summary(exp(df$pm25))
 summary(df$pm25_abs_dev)
+summary(exp(df$pm25))
 
 crops <- read_csv(paste0("data/clean/ag_productivity/crop_area.csv"))
 df <- df %>%
       left_join(crops, by = c("shrid"))
+
 
 
 
@@ -173,7 +181,14 @@ setFixest_fml(..ctrl3 = ~ rain1_zbin0 + rain1_zbin1 + rain2_zbin0 + rain2_zbin1 
                 rain1_zbin0*temp_mean^2 + rain1_zbin1*temp_mean^2 + rain2_zbin0*temp_mean^2 + rain2_zbin1*temp_mean^2 + rain3_zbin0*temp_mean^2 + rain3_zbin1*temp_mean^2 +
[…]

### 2023-11-29 analyses/5.counterfactual.R

@@ -1,0 +1,185 @@
+# create counterfactual of yield
+
+library(tidyverse)
+library(sf)
+library(terra)
+library(exactextractr)
+
+
+
+# district shapefile
+dist <- st_read("data/counterfactual/Census_2011/2011_Dist.shp")
+# rename
+dist <- dist %>%
+            dplyr::select(dist_merge = censuscode, geometry)
+
+# now we want pollution levels for each district in 2001 and 2011
+months <- c()
+for (month in 1:9){
+  months <- c(months, paste0("0", month))
+}
+months <- c(months, "10", "11", "12")
+rastlist2001 <- c()
+for (month in months){
+# MONSOON --------------------------------------------------------------------
+# This changes by year
+    r <- rast(paste0("data/raw/pm25/V5GL03.HybridPM25.Asia.2001", month, "-2001", month, ".nc"))
+    # extract
+    r <- exact_extract(r, dist, "mean", append_cols = "dist_merge")
+
+    rastlist2001 <- rbind(rastlist2001, r)
+}
+rastlist2011 <- c()
+for (month in months){
+# MONSOON --------------------------------------------------------------------
+# This changes by year
+    r <- rast(paste0("data/raw/pm25/V5GL03.HybridPM25.Asia.2011", month, "-2011", month, ".nc"))
+    # extract
+    r <- exact_extract(r, dist, "mean", append_cols = "dist_merge")
+
+    rastlist2011 <- rbind(rastlist2011, r)
+}
+# take means at dist_merge by each year
+rastlist2001 <- rastlist2001 %>%
+                    group_by(dist_merge) %>%
+                    summarise(pm25 = mean(mean, na.rm = TRUE))
+rastlist2001$year <- 2001
+rastlist2011 <- rastlist2011 %>%
+                    group_by(dist_merge) %>%
+                    summarise(pm25 = mean(mean, na.rm = TRUE))
+rastlist2011$year <- 2011
+# rbind
+rastlist <- rbind(rastlist2001, rastlist2011)
+
+
+
+
+
+
+
+
+
+
+
+
+
+# load yield data (http://data.icrisat.org/dld/index.html)
+yield <- read_csv("data/counterfactual/yield.csv")
+colnames(yield) <- c("dist", "year", "state", "statename", "distname", "riceproduction", "wheatproduction", "sorghumproduction",
+                        "pearlmilletproduction", "maizeproduction", "fingermilletproduction")
+# add peral and finger together
[…]

### 2023-12-01 draft/draft.Rmd
draft

@@
 opplot <- readRDS("pollution_development/draft/tables/opplot.rds")
-coef_vec <- readRDS("pollution_development/draft/tables/randomization_coef_vec.rds")
-coefs_distfe <- readRDS("pollution_development/draft/tables/randomization_coefs_distfe.rds")
-true_value <- readRDS("pollution_development/draft/tables/randomization_true_value.rds")
+coefsdist <- readRDS("pollution_development/draft/tables/randomization_coefs_distfe.rds")
+truevalue <- readRDS("pollution_development/draft/tables/randomization_true_value.rds")
 coef_current <- readRDS("pollution_development/draft/tables/coef_current.rds")
@@
 Past work has shown that standard errors might be underestimated in such contexts [@bertrand2004much].
-This is a particular concern in the present case, since agricultural productivity is extracted at the village level but estimated using less coarse data that has high spatial correlation [@gangopadhyay2022new].
+This is a particular concern in the present case, since agricultural productivity is extracted at the village level but estimated using data that has high spatial correlation [@gangopadhyay2022new].
 
 To help remedy concerns, I perform randomization inference (e.g.
-@barrios2012clustering) by randomly assigning observed treatment across villages and repeating this process 1,000 times.
-I do this in two separate ways.
-First, I take all villages in a given year and randomly assign, with replacement, treatment intensity within the year.
-Second, I do this separately by selecting from within district-years only, to help take into account district-year correlation in outcomes and as a way to mitigate possible concerns that traditional clustered standard errors may actually be much too conservative [@abadie2023should].
-Since this is an instrumental variables procedure, I randomly select the wind and pollution variable _together_.
+@barrios2012clustering) by randomly assigning treatment (wind and pollution) across villages and repeating this process 1,000 times.
+I do this separately by selecting from within district-years only, to help take into account district-year correlation in outcomes and as a way to mitigate possible concerns that traditional clustered standard errors may actually be much too conservative [@abadie2023should].
+Since this is an instrumental variables procedure, I randomly select the wind and pollution variable _together_, from the same observation.
 
 I present these results in Figure \@ref(fig:randomplot) of the appendix.
-The purple density is the resulting distribution of estimated treatment effects when selecting from all villages within the same year.
-The density of treatment effects is relatively tight around 0, with the left tail only reaching slightly beyond -0.05, compared to the most conservative actual estimated treatment effect of -0.63.
-
-The within-district correlation in outcomes and treatment is apparent when instead randomly selecting from only villages within the same district in the same year.
-The distribution is still approximately normal but has a mode of almost -0.4.
-Nonetheless, the actual treatment effect is larger in magnitude than almost all of the bootstrapped replications.
+The within-district correlation is apparent when selecting villages from within the same district in the same year.
+The distribution is approximately normal but has a mean and median of around -0.4.
+Nonetheless, the actual treatment effect is larger in magnitude than almost all of the bootstrapped replications; only three of the 1,000 replications results in a large coefficient than the true coefficient, leading to a p-value of well less than 0.01.
 The results are apparently robust to alternative specifications of uncertainty.
@@
 
-```{r randomplot, echo = FALSE, message = FALSE, warning = FALSE, dev='png', fig.height = 4, fig.cap = "Randomization inference", fig.footer = list("Note: The figure presents the distribution of coefficients by randomly assigning wind direction across villages within the same year.
-The \"all\" distribution is created by randomly assigning a single wind value to each village, sampled with replacement across all villages in a given year.
-The \"within district\" randomly samples with replacement only from within the same district in the same year.")}
+```{r randomplot, echo = FALSE, message = FALSE, warning = FALSE, dev='png', fig.height = 4, fig.cap = "Randomization inference", fig.footer = list("Note: The figure presents the distribution of coefficients by randomly assigning wind direction and pollution exposure across villages within the same year.
+Sampling is done within year and district.
+The vertical line is the actual coefficient from column (4) of Table 3.")}
 
 pal <- viridis(4)
-colors <- c(paste0(pal[1]), paste0(pal[2]))
-labels <- c("All", "Within district")
+colors <- c(paste0(pal[1]))
 
@@
 gg1 <- ggplot() + 
-        geom_density(aes(x = coef_vec, fill = "All"), alpha = 0.5) +
-        geom_density(aes(x = coefs_distfe, fill = "Within district"), alpha = 0.5) +
+  geom_density(aes(x = coefsdist, fill = "within district"), alpha = 0.5, show.legend = FALSE) +
   scale_fill_manual(
-          name = "Randomization:",
-          values = colors,
-          labels = labels
+    name = "randomization:",
+    values = colors
   ) +
   theme_minimal() +
-        xlab("Distribution of coefficients") +
-        ylab("Density")
+  xlab("distribution of coefficients")
 
@@
 gg1 <- gg1 + 
-        geom_vline(aes(xintercept = true_value, color = "actual")) +
+  geom_vline(aes(xintercept = truevalue, color = "actual")) +
   scale_color_manual(
@@
   ) + 
-        theme(legend.position = c(0.4, 0.6))
+  geom_vline(aes(xintercept = -0.8), color = "transparent")
 

### 2023-12-01 draft/draft.tex
draft

@@
 \author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
-\date{2023-11-23}
+\date{2023-12-01}
 
@@
 Past work has shown that standard errors might be underestimated in such contexts (Bertrand, Duflo, and Mullainathan 2004).
-This is a particular concern in the present case, since agricultural productivity is extracted at the village level but estimated using less coarse data that has high spatial correlation (Gangopadhyay et al.
+This is a particular concern in the present case, since agricultural productivity is extracted at the village level but estimated using data that has high spatial correlation (Gangopadhyay et al.
 2022).
@@
 Barrios et al.
-(2012)) by randomly assigning observed treatment across villages and repeating this process 1,000 times.
-I do this in two separate ways.
-First, I take all villages in a given year and randomly assign, with replacement, treatment intensity within the year.
-Second, I do this separately by selecting from within district-years only, to help take into account district-year correlation in outcomes and as a way to mitigate possible concerns that traditional clustered standard errors may actually be much too conservative (Abadie et al.
+(2012)) by randomly assigning treatment (wind and pollution) across villages and repeating this process 1,000 times.
+I do this separately by selecting from within district-years only, to help take into account district-year correlation in outcomes and as a way to mitigate possible concerns that traditional clustered standard errors may actually be much too conservative (Abadie et al.
 2023).
-Since this is an instrumental variables procedure, I randomly select the wind and pollution variable \emph{together}.
+Since this is an instrumental variables procedure, I randomly select the wind and pollution variable \emph{together}, from the same observation.
 
 I present these results in Figure \ref{fig:randomplot} of the appendix.
-The purple density is the resulting distribution of estimated treatment effects when selecting from all villages within the same year.
-The density of treatment effects is relatively tight around 0, with the left tail only reaching slightly beyond -0.05, compared to the most conservative actual estimated treatment effect of -0.63.
-
-The within-district correlation in outcomes and treatment is apparent when instead randomly selecting from only villages within the same district in the same year.
-The distribution is still approximately normal but has a mode of almost -0.4.
-Nonetheless, the actual treatment effect is larger in magnitude than almost all of the bootstrapped replications.
+The within-district correlation is apparent when selecting villages from within the same district in the same year.
+The distribution is approximately normal but has a mean and median of around -0.4.
+Nonetheless, the actual treatment effect is larger in magnitude than almost all of the bootstrapped replications; only three of the 1,000 replications results in a large coefficient than the true coefficient, leading to a p-value of well less than 0.01.
 The results are apparently robust to alternative specifications of uncertainty.
@@
 \begin{figure}
-\includegraphics{draft_files/figure-latex/randomplot-1} \caption[Randomization inference]{Randomization inference}\label{fig:randomplot}\floatfoot*{Note: The figure presents the distribution of coefficients by randomly assigning wind direction across villages within the same year.
-The "all" distribution is created by randomly assigning a single wind value to each village, sampled with replacement across all villages in a given year.
-The "within district" randomly samples with replacement only from within the same district in the same year.}
+\includegraphics{draft_files/figure-latex/randomplot-1} \caption[Randomization inference]{Randomization inference}\label{fig:randomplot}\floatfoot*{Note: The figure presents the distribution of coefficients by randomly assigning wind direction and pollution exposure across villages within the same year.
+Sampling is done within year and district.
+The vertical line is the actual coefficient from column (4) of Table 3.}
 \end{figure}

### 2023-12-01 analyses/3.yield.R

@@ -149,7 +149,6 @@ df <- df %>%
 
 
 
-
 summary(df$wind_abs_dev)
 summary(df$pm25_abs_dev)
 summary(exp(df$pm25))
@@ -159,6 +158,31 @@ df <- df %>%
       left_join(crops, by = c("shrid"))
 
 
+
+# # get lat/lon
+# villages_overlap <- read_sf("data/spatial/villages_overlap/villages_overlap.shp") %>% st_set_crs(st_crs("EPSG:24378"))
+# villages_overlap <- villages_overlap %>% 
+#                       mutate(
+#                              shrid = paste0(pc11_s_id, "-", pc11_tv_id)
+#                              ) %>%
+#                       dplyr::select(shrid) %>%
+#                       group_by(shrid) %>%
+#                       filter(row_number()==1) %>%
+#                       ungroup()
+# # to centroid
+# villages_overlap <- villages_overlap %>% st_centroid() 
+# villages_overlap <- st_transform(villages_overlap, crs = 4326)
+# shrid <- villages_overlap$shrid
+# # to lat/lon
+# villages_overlap <- st_coordinates(villages_overlap)
+# villages_overlap <- as_tibble(cbind(shrid, villages_overlap))
+# colnames(villages_overlap) <- c("shrid", "lon", "lat")
+
+# villages_overlap$lon <- as.numeric(villages_overlap$lon)
+# villages_overlap$lat <- as.numeric(villages_overlap$lat)
+
+# df <- df %>%
+#       left_join(villages_overlap, by = c("shrid"))
 
 
 
@@ -180,15 +204,6 @@ setFixest_fml(..ctrl3 = ~ rain1_zbin0 + rain1_zbin1 + rain2_zbin0 + rain2_zbin1 
                 temp_mean^2 + 
                 rain1_zbin0*temp_mean^2 + rain1_zbin1*temp_mean^2 + rain2_zbin0*temp_mean^2 + rain2_zbin1*temp_mean^2 + rain3_zbin0*temp_mean^2 + rain3_zbin1*temp_mean^2 +
                 rain4_zbin0*temp_mean^2 + rain4_zbin1*temp_mean^2 + rain5_zbin0*temp_mean^2 + rain5_zbin1*temp_mean^2)
-
-# add centroid
-df <- df %>% left_join(villages_overlap_latlon, by = "shrid")
-vcov_conley(
-  yield4,
-  lat = "centroid_lat",
-  lon = "centroid_lon",
[…]

### 2023-12-02 draft/draft.Rmd
draft

@@
 I first plot the location of coal plants in India by year.
-Using a village-level shapefile, I identify all villages that are located within 30km of a coal plant at any time during the sample period, which comprises the 2000s to the 2010s, with differences across outcomes depending on data availability.
+Using a village-level shapefile, I identify all villages that are located within 30km of a coal plant at any time during the sample period, which comprises the 2000s to the 2010s.
 I then calculate the direction from each coal plant to all villages within 30km.
@@
 
-Heterogeneity analysis shows that the negative effects of pollution are larger in villages that experience higher median exposure to pollution and in villages with higher initial yields.
+Heterogeneity analysis shows that the negative effects of pollution are relatively similar based on median exposure to pollution but larger in villages with higher initial yields.
 Finally, given how many poor farmers rely on rainfall and how weather shocks are on the rise across the globe, I also look at possible multiplicative effects of shocks by interacting the pollution variables with a rainfall variable.
@@
 I then calculate the direction from these point sources to all village centroids within that 30km radius.
-To define exposure, I pull daily wind direction data from the National Center for Atmospheric Research.[^9] For each day, I document whether the wind is blowing towards each village centroid[^10] and then temporally aggregate this daily data depending on the temporal definition of the corresponding outcome.
-For example, agricultural productivity is defined across five months (i.e.
+To define exposure, I pull daily wind direction data from the National Center for Atmospheric Research.[^9] For each day, I document whether the wind is blowing towards each village centroid[^10] and then temporally aggregate this daily data to match the data on productivity.
+Specifically, agricultural productivity is defined across five months (i.e.
 the monsoon season is from June to October) so I count the total days a given village is exposed to wind blowing from *any* coal plant within 30km during those five months.
@@
 While this is a common approach in the literature ( see @hammer2020global for more details), it is important to note that the estimates are not direct measurements of pollution.
-However, a valid instrumental variable can help for any measurement error present in the data [@glewwe2012much; @gillen2019experimenting].
+However, a valid instrumental variable can help correct for any measurement error present in the data [@glewwe2012much; @gillen2019experimenting].
 
@@
 The first two columns use 1991 census data -- downloaded from SHRUG -- and the second two columns use 2001 census data, all at the village level.
-The first column presents results from regressing a dummy for whether there is a coal plant within 30km in 1991 on several census variables.
-Villages with higher population densities (higher log population and lower log area) are more likely to be near a coal plant at the time.
-Urban population proportion is not a significant predictor after conditioning on population and area.
+The first column presents results from regressing a dummy for whether there is a coal plant within the village polygon in 1991 on several census variables.
+It seems that the exact location of a coal plant does not correlate with overall population of the village, but it does correlate with area; in other words, coal plants are more likely to be located in larger villages but, conditional on area, not in places with more or fewer people.
 
@@
 This column restricts the sample to only villages that do not already have a coal plant within 30km in 1991.
-The pattern of significance remains the same in column two.
+We see that the pattern of coefficients changes from column one to column two; higher populations predict the construction of coal plant in the village over the next ten years.
+Note that I do not have information on productivity and pollution from 1991.
 
@@
 Importantly, I include agricultural productivity in the monsoon season of 2002 -- the earliest year available -- as well as average pollution during the monsoon season for 1998 through 2000 as additional predictors.
-In 2001, agricultural productivity is a strong predictor for the presence of a coal plant.
-Moreover, it is also a significant predictor of the construction of a coal plant between 2001 and 2011 (column four).
-Given that the coefficient is positive, this indicates that coal plants are more likely to be built in areas with higher agricultural productivity.
-In other words, any selection in the overall location of coal plants would bias against finding negative effects of pollution on agricultural productivity.
-However, note that this is not the same as looking at wind direction, though comparing the OLS and IV results below leads to the same conclusion regarding selection.
-
-Pollution is also a significant predictor for the presence of a coal plant in 2001 as well as for whether a coal plant will be constructed before 2011.
-While the coefficient is relatively small, it is actually quite large given that only three percent of villages in column four will get a new plant in the next decade.
+Pollution is a strong predictor for the presence of a coal plant in 2001 as well as for whether a coal plant will be constructed before 2011.
+While the coefficient is small, it is actually relatively large given that relatively few villages will get a new plant in the next decade.
+Agricultural producitivty is positively correlated with the presence of a coal plant, but not significantly so.
 
@@
 Consider the reduced form effect of wind exposure on agricultural productivity:
-\begin{gather} PM2.5_{it} = \alpha_{i} + \gamma_{t} + \beta wind_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it}, \end{gather}
+\begin{gather} y_{it} = \alpha_{i} + \gamma_{t} + \beta wind_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it}, \end{gather}
 
@@
 It is worth taking the time to validate the use of wind direction as a measure of exposure to pollution.
-Consider the data used for the regressions presented in Table \@ref(tab:yield3ivmainfirststage) of the appendix.
+Consider the data used for the regressions presented in Table \@ref(tab:yield3ivmainfirststage) of the appendix, which is the first stage of the 2SLS procedure.
 The outcome variable is particulate matter -- specifically, $\mathrm{PM 2.5}$, which is particulate matter no larger than 2.5 micrometers in diameter -- which comes from @hammer2020global.
 $\mathrm{PM 2.5}$ is one of the harmful byproducts from many industrial processes, including power generation.
-Particulate matter is a major component of the pollution emitted by coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.[^12] This data is available at the monthly level, but I aggregate variables to the season variable since that is the level of analysis in the main results.
+Particulate matter is a major component of the pollution emitted by coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.[^12] This data is available at the monthly level, but I aggregate variables to the season level since that is the level of analysis in the main results.
 
@@
 Both the particulate matter data and the wind data is at the village level, which means the regressions in Table \@ref(tab:yield3ivmainfirststage) are at the village level.
-As such, the geographic fixed effects are village fixed effects and the temporal fixed effects are month fixed effects.
+As such, the geographic fixed effects are village fixed effects and the temporal fixed effects are year fixed effects.
 Since the exposure variable is at the village level and we follow villages over time, the standard errors are also clustered at the village level.
@@
 Recall, however, the results from above: coal plants were placed in places with higher average agricultural productivity.
-This means that the simple regressions in Table \@ref(tab:yieldtablepm) might actually be biased upwards: although pollution might negatively affect productivity, pollution is higher in more productive areas due to the endogenous placement of pollution sources in the study areas.
+This means that the simple regressions in Table \@ref(tab:yieldtablepm) might actually be biased upwards; although pollution might negatively affect productivity, pollution is higher in more productive areas due to the endogenous placement of pollution sources in the study areas.
 One possible way to recover the actual effect of pollution on agricultural productivity is with an instrument.
@@
 Column four presents one last specification; it includes a much more flexible specification for weather.
-This is the same binned specificaiton as I describe above.
+This is the same binned specification as I describe above.
 The coefficient in column four shows almost not change from those in column three, again relieving some concerns that weather may not be correctly specified, leading to a spurious correlation between particulate matter and agricultural productivity.
@@
            "Note: Standard errors are in parentheses and are clustered at the village level.
-The outcome in all regressions estimated agricultural yield (tons per hectare).
+The outcome in all regressions is estimated (log) agricultural yield (tons per hectare).
 * p<0.10 ** p<0.05 *** p<0.01",
@@
 
-## Crops and labor intensity
+## Mechanisms
 
@@
            "Note: Standard errors are in parentheses and are clustered at the village level.
-The outcome in all regressions estimated agricultural yield (tons per hectare).
+The outcome in all regressions is estimated (log) agricultural yield (tons per hectare).
 The rice dummy equals one in villages where rice is the most commonly harvested crop.
@@
 
-```{r yield8monthly, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Effects of pollution on productivity by month"}
+```{r yield8monthly, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Effects of pollution on productivity by month", fig.footer = "Coefficients are from a single regression, where the outcome is (log) agricultural yield and pollution in each of the five months of the season is instrumented with wind in each month.
+Standard errors are indicated by the bars and are clustered at the village level."}
 
@@
   geom_linerange(aes(x = 1:5, ymin = monthlycoefs$confint[,1], ymax = monthlycoefs$confint[,2])) +
+  geom_hline(yintercept = 0, linetype = "dashed", color = "gray", alpha = 0.5) +
   theme_minimal() +
@@
 Another way to tease out differential effects on labor productivity, specifically, is to look at the effects of pollution by month.
-If pollution is working through labor, then we would expect to see larger effects when labor is relatively more important than.
+If pollution is working through labor, then we would expect to see larger effects when labor is relatively more important.
 The harvest season, the fifth month of the agricultural season in the data, stands out as the most likely candidate for two reasons.
-First, labor allocation is generally quite on plots during the harvest season.
-Second, pollution at the very end of the season is unlikely to have large effects on overall land producitivty, as opposed to pollution in prior months.
-
-Figure \@ref(yield8monthly) presents the coefficients of the IV regression separately for each month of the agricultural season, with the very beginning of the season on the left side of the graph and the harvest season on the right side of the graph.
+First, labor allocation is generally quite high on plots during the harvest season.
+Second, pollution at the very end of the season is unlikely to have large effects on overall land producitivty, as opposed to pollution in prior months, since growth is generally complete by the end of the season.
+
+Figure \@ref(fig:yield8monthly) presents the coefficients of the IV regression separately for each month of the agricultural season, with the very beginning of the season on the left side of the graph and the harvest season on the right side of the graph.
 Interestingly, the coefficient for the fourth month of the season is actually positive.
@@
 In other words, this is not identifying differences in overall pollution, per se, but rather differences in the effects of an increase of pollution by the same amount.
-These different effects could be due to, for example, differences in emission patterns for different kinds of plants and factories; coal plants, for example, are responsible for much more mercury than other kinds of polluting industries.^[https://www.epa.gov/sites/default/files/2016-12/documents/nei2014v1_tsd.pdf]
-
-# Quantifying the overall effects of pollution on agricultural productivity
+Since I did not find heterogeneity by exposure, it does not seem that this is driven by non-linear effects of pollution on productivity.
+Instead, these different effects could be due to, for example, differences in emission patterns for different kinds of plants and factories; coal plants, for example, are responsible for much more mercury than other kinds of polluting industries.^[https://www.epa.gov/sites/default/files/2016-12/documents/nei2014v1_tsd.pdf]
+
+# Overall effects of pollution on agricultural productivity
 
@@
 For this reason, I instead turn to data from ICRISAT,^[http://data.icrisat.org/dld] which has district-level total production data from different years.
-To track with the timing of the main data used above, I pull total agricultural production from 2001 and 2011 for three of the most common crops in India: rice, wheat, and maize.
-To simplify the calculation, I assume total area devoted to each crop is unchanged across the decade.
+To track with the timing of the main data used above, I pull total agricultural production from 2001 and 2011 for three of the most common crops in India: rice, wheat, and maize.^[Maize might not seem like the first choice for India.
+However, maize is one of the most common crops grown in other developing countries -- especially in Sub-Saharan Africa -- so I use it here as a benchmark.] To simplify the calculation, I assume total area devoted to each crop is unchanged across the decade.
 I then merge this with the pollution data from @hammer2020global using a 2011 district-level census shapefile, which has identifiers that match the ICRISAT data.
@@
 I can calculate the overall change in log output across the two years, as well as the overall change in pollution levels.
-I use the most conservative point estimate *without* coal plants in \@ref(tab:yield7ivdiffindiff), which is $-0.223$.
+I use the most conservative point estimate *without* coal plants in Table \@ref(tab:yield7ivdiffindiff), which is $-0.223$.
 For each district, I calculate the counterfactual for each crop as follows:
@@
 
-```{r yielddatamaps, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "The maps show the district-level yields of each crop in 2001.", fig.footer = "Yield data comes from ICRISAT."}
+```{r yielddatamaps, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "District-level yields in 2001", fig.footer = "Yield data comes from ICRISAT."}
 

### 2023-12-02 draft/draft.tex
draft

@@
 \author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
-\date{2023-12-01}
+\date{2023-12-02}
 
@@
 To define exposure, I pull daily wind direction data from the National Center for Atmospheric Research.\footnote{climatedataguide.ucar.edu/} For each day, I document whether the wind is blowing towards each village centroid\footnote{I define ``towards'' as within five degrees to help take into account that I am using village centroids.
-The wind data includes both the x and y components of wind speed, which I can use to calculate angles from point sources to village centroids.} and then temporally aggregate this daily data depending on the temporal definition of the corresponding outcome.
-For example, agricultural productivity is defined across five months (i.e.~the monsoon season is from June to October) so I count the total days a given village is exposed to wind blowing from \emph{any} coal plant within 30km during those five months.
+The wind data includes both the x and y components of wind speed, which I can use to calculate angles from point sources to village centroids.} and then temporally aggregate this daily data to match the data on productivity.
+Specifically, agricultural productivity is defined across five months (i.e.~the monsoon season is from June to October) so I count the total days a given village is exposed to wind blowing from \emph{any} coal plant within 30km during those five months.
 
@@
 (2020) for more details), it is important to note that the estimates are not direct measurements of pollution.
-However, a valid instrumental variable can help for any measurement error present in the data (Glewwe 2012; Gillen, Snowberg, and Yariv 2019).
+However, a valid instrumental variable can help correct for any measurement error present in the data (Glewwe 2012; Gillen, Snowberg, and Yariv 2019).
 
@@
 It is worth taking the time to validate the use of wind direction as a measure of exposure to pollution.
-Consider the data used for the regressions presented in Table \ref{tab:yield3ivmainfirststage} of the appendix.
+Consider the data used for the regressions presented in Table \ref{tab:yield3ivmainfirststage} of the appendix, which is the first stage of the 2SLS procedure.
 The outcome variable is particulate matter -- specifically, \(\mathrm{PM 2.5}\), which is particulate matter no larger than 2.5 micrometers in diameter -- which comes from Hammer et al.
@@
 \(\mathrm{PM 2.5}\) is one of the harmful byproducts from many industrial processes, including power generation.
-Particulate matter is a major component of the pollution emitted by coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.\footnote{www.epa.gov/airmarkets/power-plants-and-neighboring-communities} This data is available at the monthly level, but I aggregate variables to the season variable since that is the level of analysis in the main results.
+Particulate matter is a major component of the pollution emitted by coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.\footnote{www.epa.gov/airmarkets/power-plants-and-neighboring-communities} This data is available at the monthly level, but I aggregate variables to the season level since that is the level of analysis in the main results.
 
 Both the particulate matter data and the wind data is at the village level, which means the regressions in Table \ref{tab:yield3ivmainfirststage} are at the village level.
-As such, the geographic fixed effects are village fixed effects and the temporal fixed effects are month fixed effects.
+As such, the geographic fixed effects are village fixed effects and the temporal fixed effects are year fixed effects.
 Since the exposure variable is at the village level and we follow villages over time, the standard errors are also clustered at the village level.
@@
 Recall, however, the results from above: coal plants were placed in places with higher average agricultural productivity.
-This means that the simple regressions in Table \ref{tab:yieldtablepm} might actually be biased upwards: although pollution might negatively affect productivity, pollution is higher in more productive areas due to the endogenous placement of pollution sources in the study areas.
+This means that the simple regressions in Table \ref{tab:yieldtablepm} might actually be biased upwards; although pollution might negatively affect productivity, pollution is higher in more productive areas due to the endogenous placement of pollution sources in the study areas.
 One possible way to recover the actual effect of pollution on agricultural productivity is with an instrument.
@@
 \begin{tablenotes}[para]
-\item Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions estimated agricultural yield (tons per hectare). * p<0.10 ** p<0.05 *** p<0.01
+\item Note: Standard errors are in parentheses and are clustered at the village level.
+The outcome in all regressions is estimated (log) agricultural yield (tons per hectare).
+* p<0.10 ** p<0.05 *** p<0.01
 \end{tablenotes}
@@
 
-\hypertarget{crops-and-labor-intensity}{%
-\subsection{Crops and labor intensity}\label{crops-and-labor-intensity}}
+\hypertarget{mechanisms}{%
+\subsection{Mechanisms}\label{mechanisms}}
 
@@
 \item Note: Standard errors are in parentheses and are clustered at the village level.
-The outcome in all regressions estimated agricultural yield (tons per hectare).
+The outcome in all regressions is estimated (log) agricultural yield (tons per hectare).
 The rice dummy equals one in villages where rice is the most commonly harvested crop.
@@
 \begin{figure}
-\includegraphics{draft_files/figure-latex/yield8monthly-1} \caption[Effects of pollution on productivity by month]{Effects of pollution on productivity by month}\label{fig:yield8monthly}
+\includegraphics{draft_files/figure-latex/yield8monthly-1} \caption[Effects of pollution on productivity by month]{Effects of pollution on productivity by month}\label{fig:yield8monthly}\floatfoot*{Coefficients are from a single regression, where the outcome is (log) agricultural yield and pollution in each of the five months of the season is instrumented with wind in each month.
+Standard errors are indicated by the bars and are clustered at the village level.}
 \end{figure}
@@
 Another way to tease out differential effects on labor productivity, specifically, is to look at the effects of pollution by month.
-If pollution is working through labor, then we would expect to see larger effects when labor is relatively more important than.
+If pollution is working through labor, then we would expect to see larger effects when labor is relatively more important.
 The harvest season, the fifth month of the agricultural season in the data, stands out as the most likely candidate for two reasons.
-First, labor allocation is generally quite on plots during the harvest season.
-Second, pollution at the very end of the season is unlikely to have large effects on overall land producitivty, as opposed to pollution in prior months.
-
-Figure \ref{yield8monthly} presents the coefficients of the IV regression separately for each month of the agricultural season, with the very beginning of the season on the left side of the graph and the harvest season on the right side of the graph.
+First, labor allocation is generally quite high on plots during the harvest season.
+Second, pollution at the very end of the season is unlikely to have large effects on overall land producitivty, as opposed to pollution in prior months, since growth is generally complete by the end of the season.
+
+Figure \ref{fig:yield8monthly} presents the coefficients of the IV regression separately for each month of the agricultural season, with the very beginning of the season on the left side of the graph and the harvest season on the right side of the graph.
 Interestingly, the coefficient for the fourth month of the season is actually positive.
@@
 In other words, this is not identifying differences in overall pollution, per se, but rather differences in the effects of an increase of pollution by the same amount.
-These different effects could be due to, for example, differences in emission patterns for different kinds of plants and factories; coal plants, for example, are responsible for much more mercury than other kinds of polluting industries.\footnote{\url{https://www.epa.gov/sites/default/files/2016-12/documents/nei2014v1_tsd.pdf}}
-
-\hypertarget{quantifying-the-overall-effects-of-pollution-on-agricultural-productivity}{%
-\section{Quantifying the overall effects of pollution on agricultural productivity}\label{quantifying-the-overall-effects-of-pollution-on-agricultural-productivity}}
+Since I did not find heterogeneity by exposure, it does not seem that this is driven by non-linear effects of pollution on productivity.
+Instead, these different effects could be due to, for example, differences in emission patterns for different kinds of plants and factories; coal plants, for example, are responsible for much more mercury than other kinds of polluting industries.\footnote{\url{https://www.epa.gov/sites/default/files/2016-12/documents/nei2014v1_tsd.pdf}}
+
+\hypertarget{overall-effects-of-pollution-on-agricultural-productivity}{%
+\section{Overall effects of pollution on agricultural productivity}\label{overall-effects-of-pollution-on-agricultural-productivity}}
 
@@
 For this reason, I instead turn to data from ICRISAT,\footnote{\url{http://data.icrisat.org/dld}} which has district-level total production data from different years.
-To track with the timing of the main data used above, I pull total agricultural production from 2001 and 2011 for three of the most common crops in India: rice, wheat, and maize.
-To simplify the calculation, I assume total area devoted to each crop is unchanged across the decade.
+To track with the timing of the main data used above, I pull total agricultural production from 2001 and 2011 for three of the most common crops in India: rice, wheat, and maize.\footnote{Maize might not seem like the first choice for India.
+However, maize is one of the most common crops grown in other developing countries -- especially in Sub-Saharan Africa -- so I use it here as a benchmark.} To simplify the calculation, I assume total area devoted to each crop is unchanged across the decade.
 I then merge this with the pollution data from Hammer et al.
@@
 I can calculate the overall change in log output across the two years, as well as the overall change in pollution levels.
-I use the most conservative point estimate \emph{without} coal plants in \ref{tab:yield7ivdiffindiff}, which is \(-0.223\).
+I use the most conservative point estimate \emph{without} coal plants in Table \ref{tab:yield7ivdiffindiff}, which is \(-0.223\).
 For each district, I calculate the counterfactual for each crop as follows:
@@
 \begin{figure}
-\includegraphics{draft_files/figure-latex/yielddatamaps-1} \caption[The maps show the district-level yields of each crop in 2001]{The maps show the district-level yields of each crop in 2001.}\label{fig:yielddatamaps}\floatfoot*{Yield data comes from ICRISAT.}
+\includegraphics{draft_files/figure-latex/yielddatamaps-1} \caption[District-level yields in 2001]{District-level yields in 2001}\label{fig:yielddatamaps}\floatfoot*{Yield data comes from ICRISAT.}
 \end{figure}

### 2023-12-02 draft/draft.Rmd
draft

@@
 ---
-title: "Air pollution and agricultural productivity in a developing country^[I thank Ariel Ortiz-Bobea, Marc Bellemare, and David Sungho Park for comments and suggestions that greatly improved the quality of the paper, as well as conference attendees at the 2023 KER International Conference and the University of Reading.
-The usual disclaimer applies.
-All scripts used in this paper are publicly available on my GitHub page: github.com/JoshMerfeld/pollution_development]"
+title: "Air pollution and agricultural productivity in a developing country^[I thank Ariel Ortiz-Bobea, Marc Bellemare, and David Sungho Park for comments and suggestions that greatly improved the quality of the paper, as well as conference/seminar attendees at the 2023 KER International Conference, Seoul National, the University of Reading, and the EEA.
+The usual disclaimer applies.]"
 author:
@@
            "Note: Standard errors are in parentheses and are clustered at the village level.
-The outcome in all regressions is estimated agricultural yield, defined as tons per hectare.
+The outcome in all regressions is estimated (log) agricultural yield, defined as tons per hectare.
 * p<0.10 ** p<0.05 *** p<0.01",
@@
            "Note: Standard errors are in parentheses and are clustered at the village level.
-The outcome in all regressions is estimated agricultural yield, defined as tons per hectare.
+The outcome in all regressions is estimated (log) agricultural yield, defined as tons per hectare.
 The first two columns split the sample based on the median exposure (by village).

### 2023-12-02 draft/draft.tex
draft

@@
 
-\title{Air pollution and agricultural productivity in a developing country\footnote{I thank Ariel Ortiz-Bobea, Marc Bellemare, and David Sungho Park for comments and suggestions that greatly improved the quality of the paper, as well as conference attendees at the 2023 KER International Conference and the University of Reading.
-The usual disclaimer applies.
-All scripts used in this paper are publicly available on my GitHub page: github.com/JoshMerfeld/pollution\_development}}
+\title{Air pollution and agricultural productivity in a developing country\footnote{I thank Ariel Ortiz-Bobea, Marc Bellemare, and David Sungho Park for comments and suggestions that greatly improved the quality of the paper, as well as conference/seminar attendees at the 2023 KER International Conference, Seoul National, the University of Reading, and the EEA.
+The usual disclaimer applies.}}
 \author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
@@
 \item Note: Standard errors are in parentheses and are clustered at the village level.
-The outcome in all regressions is estimated agricultural yield, defined as tons per hectare.
+The outcome in all regressions is estimated (log) agricultural yield, defined as tons per hectare.
 * p<0.10 ** p<0.05 *** p<0.01
@@
 \item Note: Standard errors are in parentheses and are clustered at the village level.
-The outcome in all regressions is estimated agricultural yield, defined as tons per hectare.
+The outcome in all regressions is estimated (log) agricultural yield, defined as tons per hectare.
 The first two columns split the sample based on the median exposure (by village).
